### 2023-12-03 projects/london-clojurians-dec-2023-talk/docs/notebooks.commits.qmd
initial exploration of commit logs

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: notebooks.commits.html}
highlight-style: solarized

---
<style>table {
  border-style: thin;
}
th, td {
  padding: 6px;
}
td {
  text-align: left;
}
th {
  text-align: center;
  background-color: #ddd;
}
tr:nth-child(even) {
  background-color: #f6f6f6;
}
</style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="notebooks.commits_files/md-default0.js" type="text/javascript"></script><script src="notebooks.commits_files/md-default1.js" type="text/javascript"></script><script src="notebooks.commits_files/vega2.js" type="text/javascript"></script><script src="notebooks.commits_files/vega3.js" type="text/javascript"></script><script src="notebooks.commits_files/vega4.js" type="text/javascript"></script>

<div class="sourceClojure">
```clojure
(ns notebooks.commits
  (:require [tablecloth.api :as tc]
            [scicloj.noj.v1.vis :as vis]
            [aerial.hanami.templates :as ht]
            [tech.v3.datatype.datetime :as datetime]
            [tech.v3.datatype.rolling :as rolling]
            [tech.v3.datatype.functional :as fun]
            [data.generate-dataset]))
```
</div>

<div class="sourceClojure">
```clojure
(def date-counts
  (-> data.generate-dataset/commit-dates
      (tc/group-by [:date])
      (tc/aggregate {:n tc/row-count})
      (tc/order-by [:date])))
```
</div>

<div class="sourceClojure">
```clojure
date-counts
```
</div>

_unnamed [5654 2]:

|      :date |  :n |
|------------|----:|
| 2008-03-29 |   7 |
| 2008-04-02 |   4 |
| 2008-04-03 |   9 |
| 2008-04-05 |   2 |
| 2008-04-06 |   1 |
| 2008-04-07 |   5 |
| 2008-04-08 |   9 |
| 2008-04-10 |   1 |
| 2008-04-11 |   1 |
| 2008-04-16 |   1 |
|        ... | ... |
| 2023-11-23 |  86 |
| 2023-11-24 |  95 |
| 2023-11-25 |  32 |
| 2023-11-26 |  54 |
| 2023-11-27 |  66 |
| 2023-11-28 |  95 |
| 2023-11-29 |  87 |
| 2023-11-30 |  91 |
| 2023-12-01 |  43 |
| 2023-12-02 |  32 |
| 2023-12-03 |  19 |

<div class="sourceClojure">
```clojure
(-> date-counts
    (vis/hanami-plot ht/line-chart
                     {:X "date"
                      :XTYPE "temporal"
                      :Y "n"}))
```
</div>

<div><script>vegaEmbed(document.currentScript.parentElement, {"encoding":{"y":{"field":"n","type":"quantitative"},"x":{"field":"date","type":"temporal"}},"mark":{"type":"line","tooltip":true},"width":400,"background":"floralwhite","height":300,"data":{"values":"date,n\n2008-03-29,7\n2008-04-02,4\n2008-04-03,9\n2008-04-05,2\n2008-04-06,1\n2008-04-07,5\n2008-04-08,9\n2008-04-10,1\n2008-04-11,1\n2008-04-16,1\n2008-04-18,1\n2008-04-19,1\n2008-04-21,1\n2008-04-24,3\n2008-04-25,1\n2008-04-29,1\n2008-04-30,4\n2008-05-01,1\n2008-05-02,2\n2008-05-03,2\n2008-05-07,3\n2008-05-08,2\n2008-05-10,6\n2008-05-11,1\n2008-05-12,1\n2008-05-13,2\n2008-05-14,1\n2008-05-15,1\n2008-05-16,2\n2008-05-17,9\n2008-05-18,12\n2008-05-22,1\n2008-05-24,3\n2008-05-25,2\n2008-05-26,1\n2008-05-27,1\n2008-05-28,1\n2008-05-29,4\n2008-05-31,1\n2008-06-01,2\n2008-06-02,1\n2008-06-03,3\n2008-06-04,2\n2008-06-05,1\n2008-06-06,1\n2008-06-07,1\n2008-06-08,5\n2008-06-09,3\n2008-06-11,2\n2008-06-12,2\n2008-06-13,2\n2008-06-14,4\n2008-06-15,5\n2008-06-18,1\n2008-06-19,3\n2008-06-27,1\n2008-06-28,6\n2008-06-29,6\n2008-06-30,2\n2008-07-03,2\n2008-07-04,1\n2008-07-06,5\n2008-07-07,2\n2008-07-08,1\n2008-07-09,8\n2008-07-10,5\n2008-07-11,6\n2008-07-12,4\n2008-07-13,3\n2008-07-14,1\n2008-07-15,5\n2008-07-21,1\n2008-07-22,5\n2008-07-24,2\n2008-07-25,2\n2008-07-26,7\n2008-07-27,3\n2008-07-29,8\n2008-08-05,1\n2008-08-06,4\n2008-08-07,7\n2008-08-08,1\n2008-08-09,1\n2008-08-10,1\n2008-08-11,1\n2008-08-12,4\n2008-08-13,1\n2008-08-14,1\n2008-08-15,3\n2008-08-16,9\n2008-08-17,4\n2008-08-19,7\n2008-08-20,3\n2008-08-21,1\n2008-08-25,2\n2008-08-27,2\n2008-08-28,2\n2008-08-29,5\n2008-08-30,8\n2008-09-02,8\n2008-09-03,7\n2008-09-04,1\n2008-09-05,1\n2008-09-06,2\n2008-09-07,3\n2008-09-08,29\n2008-09-09,4\n2008-09-10,3\n2008-09-11,2\n2008-09-12,3\n2008-09-13,11\n2008-09-14,2\n2008-09-15,8\n2008-09-16,11\n2008-09-17,8\n2008-09-18,1\n2008-09-19,2\n2008-09-20,8\n2008-09-21,7\n2008-09-22,3\n2008-09-23,4\n2008-09-24,4\n2008-09-25,6\n2008-09-26,4\n2008-09-27,2\n2008-09-28,7\n2008-10-01,6\n2008-10-02,10\n2008-10-03,5\n2008-10-04,3\n2008-10-05,5\n2008-10-06,11\n2008-10-07,2\n2008-10-08,6\n2008-10-09,3\n2008-10-10,4\n2008-10-11,2\n2008-10-12,4\n2008-10-13,2\n2008-10-14,1\n2008-10-15,10\n2008-10-16,1\n2008-10-17,4\n2008-10-18,2\n2008-10-19,8\n2008-10-21,1\n2008-10-22,5\n2008-10-23,4\n2008-10-25,2\n2008-10-26,3\n2008-10-27,1\n2008-10-28,2\n2008-10-29,1\n2008-10-30,5\n2008-10-31,5\n2008-11-01,9\n2008-11-02,1\n2008-11-03,7\n2008-11-08,3\n2008-11-09,5\n2008-11-11,1\n2008-11-12,5\n2008-11-13,23\n2008-11-14,2\n2008-11-15,8\n2008-11-16,17\n2008-11-17,3\n2008-11-18,25\n2008-11-19,3\n2008-11-20,2\n2008-11-21,3\n2008-11-22,3\n2008-11-23,7\n2008-11-24,1\n2008-11-25,6\n2008-11-26,3\n2008-11-27,3\n2008-11-28,5\n2008-11-29,5\n2008-11-30,5\n2008-12-01,7\n2008-12-02,3\n2008-12-03,7\n2008-12-04,5\n2008-12-05,6\n2008-12-07,3\n2008-12-08,4\n2008-12-09,1\n2008-12-10,6\n2008-12-11,5\n2008-12-12,1\n2008-12-13,12\n2008-12-14,8\n2008-12-15,9\n2008-12-16,6\n2008-12-17,6\n2008-12-19,18\n2008-12-20,1\n2008-12-21,3\n2008-12-22,3\n2008-12-23,10\n2008-12-24,3\n2008-12-25,1\n2008-12-26,1\n2008-12-27,13\n2008-12-28,13\n2008-12-29,15\n2008-12-30,10\n2008-12-31,2\n2009-01-01,4\n2009-01-02,5\n2009-01-03,2\n2009-01-04,14\n2009-01-05,3\n2009-01-06,7\n2009-01-07,1\n2009-01-08,17\n2009-01-09,4\n2009-01-10,3\n2009-01-11,9\n2009-01-12,19\n2009-01-13,5\n2009-01-14,4\n2009-01-15,11\n2009-01-16,10\n2009-01-17,36\n2009-01-18,36\n2009-01-19,22\n2009-01-20,12\n2009-01-21,16\n2009-01-22,8\n2009-01-23,13\n2009-01-24,7\n2009-01-25,15\n2009-01-26,13\n2009-01-27,6\n2009-01-28,5\n2009-01-29,7\n2009-01-30,5\n2009-01-31,2\n2009-02-01,2\n2009-02-02,4\n2009-02-03,1\n2009-02-04,3\n2009-02-05,1\n2009-02-06,8\n2009-02-07,3\n2009-02-08,3\n2009-02-10,22\n2009-02-11,14\n2009-02-12,18\n2009-02-13,6\n2009-02-14,5\n2009-02-15,9\n2009-02-16,5\n2009-02-17,8\n2009-02-18,7\n2009-02-19,4\n2009-02-20,8\n2009-02-21,10\n2009-02-22,15\n2009-02-23,12\n2009-02-24,27\n2009-02-25,7\n2009-02-26,11\n2009-02-27,19\n2009-02-28,3\n2009-03-01,5\n2009-03-02,3\n2009-03-03,8\n2009-03-04,7\n2009-03-05,2\n2009-03-06,6\n2009-03-07,3\n2009-03-08,2\n2009-03-09,6\n2009-03-10,4\n2009-03-11,7\n2009-03-12,13\n2009-03-13,4\n2009-03-14,8\n2009-03-15,7\n2009-03-16,9\n2009-03-17,9\n2009-03-18,11\n2009-03-19,13\n2009-03-20,8\n2009-03-21,13\n2009-03-22,11\n2009-03-23,8\n2009-03-24,4\n2009-03-25,4\n2009-03-26,4\n2009-03-27,9\n2009-03-28,10\n2009-03-29,17\n2009-03-30,9\n2009-03-31,7\n2009-04-01,8\n2009-04-02,1\n2009-04-03,5\n2009-04-04,8\n2009-04-05,16\n2009-04-06,15\n2009-04-07,5\n2009-04-08,7\n2009-04-09,15\n2009-04-10,5\n2009-04-11,2\n2009-04-12,3\n2009-04-13,1\n2009-04-14,9\n2009-04-15,17\n2009-04-16,9\n2009-04-17,19\n2009-04-18,8\n2009-04-19,21\n2009-04-20,28\n2009-04-21,24\n2009-04-22,13\n2009-04-23,25\n2009-04-24,8\n2009-04-25,8\n2009-04-26,6\n2009-04-27,4\n2009-04-28,2\n2009-04-29,6\n2009-04-30,1\n2009-05-01,8\n2009-05-02,1\n2009-05-03,12\n2009-05-04,17\n2009-05-05,7\n2009-05-06,9\n2009-05-07,13\n2009-05-08,4\n2009-05-09,15\n2009-05-10,7\n2009-05-11,13\n2009-05-12,3\n2009-05-13,2\n2009-05-14,17\n2009-05-15,3\n2009-05-16,5\n2009-05-17,5\n2009-05-18,6\n2009-05-19,6\n2009-05-20,3\n2009-05-21,5\n2009-05-22,2\n2009-05-23,16\n2009-05-24,7\n2009-05-25,7\n2009-05-26,4\n2009-05-27,7\n2009-05-28,3\n2009-05-29,7\n2009-05-30,4\n2009-05-31,19\n2009-06-01,7\n2009-06-02,10\n2009-06-03,13\n2009-06-04,7\n2009-06-05,17\n2009-06-06,8\n2009-06-07,1\n2009-06-08,11\n2009-06-09,8\n2009-06-10,4\n2009-06-11,7\n2009-06-12,4\n2009-06-13,12\n2009-06-14,4\n2009-06-15,3\n2009-06-16,8\n2009-06-17,2\n2009-06-18,2\n2009-06-19,18\n2009-06-20,1\n2009-06-21,1\n2009-06-22,9\n2009-06-23,13\n2009-06-24,14\n2009-06-25,3\n2009-06-26,17\n2009-06-27,7\n2009-06-28,10\n2009-06-29,3\n2009-06-30,6\n2009-07-01,9\n2009-07-02,17\n2009-07-03,6\n2009-07-04,4\n2009-07-05,6\n2009-07-06,9\n2009-07-07,7\n2009-07-08,13\n2009-07-09,7\n2009-07-10,6\n2009-07-11,2\n2009-07-12,8\n2009-07-13,5\n2009-07-14,7\n2009-07-15,5\n2009-07-16,6\n2009-07-17,3\n2009-07-18,5\n2009-07-19,4\n2009-07-20,14\n2009-07-21,6\n2009-07-23,10\n2009-07-24,11\n2009-07-25,3\n2009-07-26,9\n2009-07-27,5\n2009-07-28,21\n2009-07-29,17\n2009-07-30,5\n2009-07-31,6\n2009-08-01,11\n2009-08-02,4\n2009-08-03,9\n2009-08-04,5\n2009-08-05,3\n2009-08-06,1\n2009-08-07,7\n2009-08-08,7\n2009-08-09,1\n2009-08-10,4\n2009-08-11,2\n2009-08-12,5\n2009-08-13,4\n2009-08-14,9\n2009-08-15,6\n2009-08-16,2\n2009-08-17,21\n2009-08-18,14\n2009-08-19,13\n2009-08-20,6\n2009-08-21,5\n2009-08-22,1\n2009-08-23,9\n2009-08-24,4\n2009-08-25,10\n2009-08-26,2\n2009-08-27,5\n2009-08-28,7\n2009-08-29,8\n2009-08-30,3\n2009-08-31,4\n2009-09-01,8\n2009-09-02,8\n2009-09-03,19\n2009-09-04,22\n2009-09-05,4\n2009-09-06,8\n2009-09-07,15\n2009-09-08,16\n2009-09-09,14\n2009-09-10,16\n2009-09-11,17\n2009-09-12,4\n2009-09-13,11\n2009-09-14,12\n2009-09-15,12\n2009-09-16,10\n2009-09-17,9\n2009-09-18,19\n2009-09-19,9\n2009-09-20,4\n2009-09-21,4\n2009-09-22,6\n2009-09-23,21\n2009-09-25,2\n2009-09-26,4\n2009-09-27,2\n2009-09-28,8\n2009-09-29,5\n2009-09-30,4\n2009-10-01,6\n2009-10-02,2\n2009-10-03,12\n2009-10-04,4\n2009-10-05,10\n2009-10-06,12\n2009-10-07,30\n2009-10-08,8\n2009-10-09,9\n2009-10-10,2\n2009-10-11,1\n2009-10-13,2\n2009-10-14,5\n2009-10-15,4\n2009-10-16,8\n2009-10-17,3\n2009-10-18,5\n2009-10-19,5\n2009-10-21,4\n2009-10-22,7\n2009-10-23,15\n2009-10-24,29\n2009-10-25,7\n2009-10-26,20\n2009-10-27,9\n2009-10-28,8\n2009-10-29,5\n2009-10-30,2\n2009-10-31,10\n2009-11-01,17\n2009-11-02,24\n2009-11-03,7\n2009-11-04,9\n2009-11-05,8\n2009-11-06,10\n2009-11-07,2\n2009-11-08,16\n2009-11-09,14\n2009-11-10,13\n2009-11-11,19\n2009-11-12,12\n2009-11-13,11\n2009-11-14,33\n2009-11-15,18\n2009-11-16,13\n2009-11-17,15\n2009-11-18,23\n2009-11-19,23\n2009-11-20,20\n2009-11-21,23\n2009-11-22,18\n2009-11-23,22\n2009-11-24,12\n2009-11-25,14\n2009-11-26,17\n2009-11-27,18\n2009-11-28,26\n2009-11-29,14\n2009-11-30,16\n2009-12-01,27\n2009-12-02,22\n2009-12-03,9\n2009-12-04,8\n2009-12-05,10\n2009-12-06,10\n2009-12-07,8\n2009-12-08,9\n2009-12-09,9\n2009-12-10,22\n2009-12-11,10\n2009-12-12,9\n2009-12-13,6\n2009-12-14,8\n2009-12-15,8\n2009-12-16,14\n2009-12-17,5\n2009-12-18,15\n2009-12-19,19\n2009-12-20,12\n2009-12-21,16\n2009-12-22,12\n2009-12-23,12\n2009-12-24,13\n2009-12-25,11\n2009-12-26,32\n2009-12-27,25\n2009-12-28,16\n2009-12-29,21\n2009-12-30,27\n2009-12-31,13\n2010-01-01,14\n2010-01-02,14\n2010-01-03,12\n2010-01-04,28\n2010-01-05,7\n2010-01-06,4\n2010-01-07,5\n2010-01-08,18\n2010-01-09,10\n2010-01-10,20\n2010-01-11,39\n2010-01-12,31\n2010-01-13,21\n2010-01-14,27\n2010-01-15,6\n2010-01-16,11\n2010-01-17,7\n2010-01-18,10\n2010-01-19,11\n2010-01-20,26\n2010-01-21,16\n2010-01-22,20\n2010-01-23,8\n2010-01-24,18\n2010-01-25,24\n2010-01-26,46\n2010-01-27,32\n2010-01-28,28\n2010-01-29,30\n2010-01-30,43\n2010-01-31,20\n2010-02-01,43\n2010-02-02,14\n2010-02-03,8\n2010-02-04,6\n2010-02-05,10\n2010-02-06,16\n2010-02-07,11\n2010-02-08,13\n2010-02-09,26\n2010-02-10,26\n2010-02-11,49\n2010-02-12,38\n2010-02-13,20\n2010-02-14,9\n2010-02-15,19\n2010-02-16,24\n2010-02-17,20\n2010-02-18,25\n2010-02-19,29\n2010-02-20,24\n2010-02-21,37\n2010-02-22,44\n2010-02-23,24\n2010-02-24,14\n2010-02-25,10\n2010-02-26,12\n2010-02-27,14\n2010-02-28,45\n2010-03-01,35\n2010-03-02,50\n2010-03-03,31\n2010-03-04,20\n2010-03-05,15\n2010-03-06,20\n2010-03-07,25\n2010-03-08,24\n2010-03-09,18\n2010-03-10,24\n2010-03-11,13\n2010-03-12,17\n2010-03-13,8\n2010-03-14,35\n2010-03-15,33\n2010-03-16,23\n2010-03-17,40\n2010-03-18,25\n2010-03-19,13\n2010-03-20,6\n2010-03-21,8\n2010-03-22,9\n2010-03-23,15\n2010-03-24,28\n2010-03-25,15\n2010-03-26,40\n2010-03-27,10\n2010-03-28,20\n2010-03-29,34\n2010-03-30,39\n2010-03-31,26\n2010-04-01,39\n2010-04-02,36\n2010-04-03,20\n2010-04-04,19\n2010-04-05,31\n2010-04-06,36\n2010-04-07,31\n2010-04-08,30\n2010-04-09,22\n2010-04-10,26\n2010-04-11,26\n2010-04-12,62\n2010-04-13,41\n2010-04-14,30\n2010-04-15,36\n2010-04-16,40\n2010-04-17,34\n2010-04-18,30\n2010-04-19,21\n2010-04-20,25\n2010-04-21,23\n2010-04-22,30\n2010-04-23,32\n2010-04-24,23\n2010-04-25,44\n2010-04-26,29\n2010-04-27,22\n2010-04-28,42\n2010-04-29,55\n2010-04-30,19\n2010-05-01,21\n2010-05-02,39\n2010-05-03,42\n2010-05-04,32\n2010-05-05,42\n2010-05-06,41\n2010-05-07,16\n2010-05-08,14\n2010-05-09,13\n2010-05-10,16\n2010-05-11,21\n2010-05-12,13\n2010-05-13,17\n2010-05-14,37\n2010-05-15,25\n2010-05-16,7\n2010-05-17,18\n2010-05-18,19\n2010-05-19,15\n2010-05-20,24\n2010-05-21,28\n2010-05-22,26\n2010-05-23,6\n2010-05-24,7\n2010-05-25,19\n2010-05-26,26\n2010-05-27,22\n2010-05-28,20\n2010-05-29,24\n2010-05-30,29\n2010-05-31,4\n2010-06-01,14\n2010-06-02,20\n2010-06-03,22\n2010-06-04,14\n2010-06-05,14\n2010-06-06,20\n2010-06-07,31\n2010-06-08,30\n2010-06-09,36\n2010-06-10,20\n2010-06-11,37\n2010-06-12,34\n2010-06-13,17\n2010-06-14,27\n2010-06-15,17\n2010-06-16,34\n2010-06-17,44\n2010-06-18,30\n2010-06-19,23\n2010-06-20,24\n2010-06-21,24\n2010-06-22,25\n2010-06-23,24\n2010-06-24,29\n2010-06-25,33\n2010-06-26,28\n2010-06-27,21\n2010-06-28,13\n2010-06-29,43\n2010-06-30,36\n2010-07-01,17\n2010-07-02,14\n2010-07-03,16\n2010-07-04,21\n2010-07-05,14\n2010-07-06,24\n2010-07-07,23\n2010-07-08,42\n2010-07-09,19\n2010-07-10,19\n2010-07-11,11\n2010-07-12,22\n2010-07-13,25\n2010-07-14,28\n2010-07-15,13\n2010-07-16,38\n2010-07-17,12\n2010-07-18,12\n2010-07-19,19\n2010-07-20,41\n2010-07-21,28\n2010-07-22,24\n2010-07-23,36\n2010-07-24,21\n2010-07-25,7\n2010-07-26,46\n2010-07-27,25\n2010-07-28,32\n2010-07-29,16\n2010-07-30,25\n2010-07-31,21\n2010-08-01,40\n2010-08-02,20\n2010-08-03,20\n2010-08-04,19\n2010-08-05,28\n2010-08-06,24\n2010-08-07,34\n2010-08-08,18\n2010-08-09,28\n2010-08-10,19\n2010-08-11,23\n2010-08-12,24\n2010-08-13,32\n2010-08-14,16\n2010-08-15,30\n2010-08-16,48\n2010-08-17,38\n2010-08-18,23\n2010-08-19,36\n2010-08-20,33\n2010-08-21,60\n2010-08-22,17\n2010-08-23,50\n2010-08-24,38\n2010-08-25,47\n2010-08-26,50\n2010-08-27,51\n2010-08-28,33\n2010-08-29,43\n2010-08-30,38\n2010-08-31,29\n2010-09-01,22\n2010-09-02,20\n2010-09-03,20\n2010-09-04,16\n2010-09-05,9\n2010-09-06,29\n2010-09-07,20\n2010-09-08,15\n2010-09-09,40\n2010-09-10,34\n2010-09-11,32\n2010-09-12,18\n2010-09-13,42\n2010-09-14,39\n2010-09-15,24\n2010-09-16,33\n2010-09-17,59\n2010-09-18,35\n2010-09-19,37\n2010-09-20,61\n2010-09-21,43\n2010-09-22,49\n2010-09-23,44\n2010-09-24,53\n2010-09-25,33\n2010-09-26,21\n2010-09-27,42\n2010-09-28,22\n2010-09-29,49\n2010-09-30,28\n2010-10-01,33\n2010-10-02,32\n2010-10-03,30\n2010-10-04,41\n2010-10-05,40\n2010-10-06,30\n2010-10-07,26\n2010-10-08,24\n2010-10-09,10\n2010-10-10,18\n2010-10-11,31\n2010-10-12,47\n2010-10-13,24\n2010-10-14,33\n2010-10-15,34\n2010-10-16,31\n2010-10-17,18\n2010-10-18,25\n2010-10-19,47\n2010-10-20,42\n2010-10-21,13\n2010-10-22,13\n2010-10-23,17\n2010-10-24,31\n2010-10-25,38\n2010-10-26,41\n2010-10-27,34\n2010-10-28,37\n2010-10-29,36\n2010-10-30,15\n2010-10-31,26\n2010-11-01,32\n2010-11-02,42\n2010-11-03,33\n2010-11-04,32\n2010-11-05,18\n2010-11-06,39\n2010-11-07,37\n2010-11-08,45\n2010-11-09,60\n2010-11-10,62\n2010-11-11,47\n2010-11-12,45\n2010-11-13,24\n2010-11-14,43\n2010-11-15,38\n2010-11-16,46\n2010-11-17,33\n2010-11-18,51\n2010-11-19,48\n2010-11-20,50\n2010-11-21,13\n2010-11-22,43\n2010-11-23,60\n2010-11-24,50\n2010-11-25,35\n2010-11-26,60\n2010-11-27,76\n2010-11-28,61\n2010-11-29,28\n2010-11-30,47\n2010-12-01,40\n2010-12-02,36\n2010-12-03,32\n2010-12-04,34\n2010-12-05,38\n2010-12-06,37\n2010-12-07,28\n2010-12-08,24\n2010-12-09,26\n2010-12-10,5\n2010-12-11,17\n2010-12-12,21\n2010-12-13,32\n2010-12-14,23\n2010-12-15,29\n2010-12-16,33\n2010-12-17,30\n2010-12-18,38\n2010-12-19,19\n2010-12-20,18\n2010-12-21,35\n2010-12-22,42\n2010-12-23,34\n2010-12-24,17\n2010-12-25,13\n2010-12-26,15\n2010-12-27,31\n2010-12-28,40\n2010-12-29,67\n2010-12-30,18\n2010-12-31,19\n2011-01-01,20\n2011-01-02,39\n2011-01-03,50\n2011-01-04,38\n2011-01-05,45\n2011-01-06,37\n2011-01-07,27\n2011-01-08,38\n2011-01-09,33\n2011-01-10,43\n2011-01-11,49\n2011-01-12,45\n2011-01-13,53\n2011-01-14,45\n2011-01-15,29\n2011-01-16,7\n2011-01-17,35\n2011-01-18,26\n2011-01-19,31\n2011-01-20,26\n2011-01-21,57\n2011-01-22,28\n2011-01-23,44\n2011-01-24,26\n2011-01-25,26\n2011-01-26,38\n2011-01-27,39\n2011-01-28,23\n2011-01-29,21\n2011-01-30,52\n2011-01-31,19\n2011-02-01,28\n2011-02-02,31\n2011-02-03,24\n2011-02-04,16\n2011-02-05,34\n2011-02-06,16\n2011-02-07,36\n2011-02-08,24\n2011-02-09,22\n2011-02-10,38\n2011-02-11,19\n2011-02-12,17\n2011-02-13,28\n2011-02-14,20\n2011-02-15,20\n2011-02-16,30\n2011-02-17,39\n2011-02-18,31\n2011-02-19,12\n2011-02-20,11\n2011-02-21,29\n2011-02-22,29\n2011-02-23,24\n2011-02-24,49\n2011-02-25,12\n2011-02-26,10\n2011-02-27,19\n2011-02-28,14\n2011-03-01,22\n2011-03-02,48\n2011-03-03,28\n2011-03-04,26\n2011-03-05,13\n2011-03-06,19\n2011-03-07,15\n2011-03-08,30\n2011-03-09,28\n2011-03-10,22\n2011-03-11,31\n2011-03-12,24\n2011-03-13,26\n2011-03-14,71\n2011-03-15,54\n2011-03-16,22\n2011-03-17,42\n2011-03-18,22\n2011-03-19,17\n2011-03-20,31\n2011-03-21,46\n2011-03-22,44\n2011-03-23,55\n2011-03-24,44\n2011-03-25,40\n2011-03-26,22\n2011-03-27,28\n2011-03-28,18\n2011-03-29,44\n2011-03-30,48\n2011-03-31,51\n2011-04-01,49\n2011-04-02,30\n2011-04-03,10\n2011-04-04,29\n2011-04-05,27\n2011-04-06,8\n2011-04-07,20\n2011-04-08,27\n2011-04-09,36\n2011-04-10,14\n2011-04-11,25\n2011-04-12,40\n2011-04-13,16\n2011-04-14,15\n2011-04-15,50\n2011-04-16,28\n2011-04-17,47\n2011-04-18,65\n2011-04-19,56\n2011-04-20,47\n2011-04-21,80\n2011-04-22,59\n2011-04-23,65\n2011-04-24,50\n2011-04-25,62\n2011-04-26,51\n2011-04-27,37\n2011-04-28,59\n2011-04-29,54\n2011-04-30,23\n2011-05-01,50\n2011-05-02,49\n2011-05-03,27\n2011-05-04,52\n2011-05-05,33\n2011-05-06,26\n2011-05-07,12\n2011-05-08,18\n2011-05-09,22\n2011-05-10,26\n2011-05-11,34\n2011-05-12,27\n2011-05-13,29\n2011-05-14,12\n2011-05-15,22\n2011-05-16,25\n2011-05-17,20\n2011-05-18,36\n2011-05-19,27\n2011-05-20,14\n2011-05-21,14\n2011-05-22,45\n2011-05-23,37\n2011-05-24,43\n2011-05-25,23\n2011-05-26,18\n2011-05-27,42\n2011-05-28,79\n2011-05-29,57\n2011-05-30,39\n2011-05-31,73\n2011-06-01,37\n2011-06-02,58\n2011-06-03,30\n2011-06-04,58\n2011-06-05,10\n2011-06-06,30\n2011-06-07,44\n2011-06-08,59\n2011-06-09,38\n2011-06-10,101\n2011-06-11,36\n2011-06-12,46\n2011-06-13,50\n2011-06-14,83\n2011-06-15,37\n2011-06-16,67\n2011-06-17,116\n2011-06-18,98\n2011-06-19,49\n2011-06-20,71\n2011-06-21,68\n2011-06-22,82\n2011-06-23,58\n2011-06-24,131\n2011-06-25,51\n2011-06-26,34\n2011-06-27,88\n2011-06-28,81\n2011-06-29,64\n2011-06-30,46\n2011-07-01,118\n2011-07-02,49\n2011-07-03,37\n2011-07-04,29\n2011-07-05,50\n2011-07-06,42\n2011-07-07,44\n2011-07-08,57\n2011-07-09,27\n2011-07-10,26\n2011-07-11,40\n2011-07-12,56\n2011-07-13,55\n2011-07-14,59\n2011-07-15,160\n2011-07-16,83\n2011-07-17,147\n2011-07-18,139\n2011-07-19,105\n2011-07-20,94\n2011-07-21,28\n2011-07-22,80\n2011-07-23,35\n2011-07-24,28\n2011-07-25,42\n2011-07-26,73\n2011-07-27,76\n2011-07-28,48\n2011-07-29,66\n2011-07-30,66\n2011-07-31,54\n2011-08-01,45\n2011-08-02,79\n2011-08-03,70\n2011-08-04,81\n2011-08-05,85\n2011-08-06,36\n2011-08-07,73\n2011-08-08,25\n2011-08-09,60\n2011-08-10,58\n2011-08-11,73\n2011-08-12,70\n2011-08-13,34\n2011-08-14,38\n2011-08-15,37\n2011-08-16,66\n2011-08-17,35\n2011-08-18,51\n2011-08-19,54\n2011-08-20,36\n2011-08-21,65\n2011-08-22,40\n2011-08-23,33\n2011-08-24,65\n2011-08-25,59\n2011-08-26,66\n2011-08-27,31\n2011-08-28,45\n2011-08-29,32\n2011-08-30,38\n2011-08-31,60\n2011-09-01,58\n2011-09-02,63\n2011-09-03,38\n2011-09-04,25\n2011-09-05,40\n2011-09-06,62\n2011-09-07,91\n2011-09-08,64\n2011-09-09,56\n2011-09-10,37\n2011-09-11,51\n2011-09-12,41\n2011-09-13,40\n2011-09-14,69\n2011-09-15,48\n2011-09-16,44\n2011-09-17,14\n2011-09-18,28\n2011-09-19,29\n2011-09-20,59\n2011-09-21,48\n2011-09-22,34\n2011-09-23,104\n2011-09-24,84\n2011-09-25,48\n2011-09-26,81\n2011-09-27,84\n2011-09-28,88\n2011-09-29,99\n2011-09-30,110\n2011-10-01,120\n2011-10-02,90\n2011-10-03,57\n2011-10-04,67\n2011-10-05,71\n2011-10-06,48\n2011-10-07,74\n2011-10-08,59\n2011-10-09,69\n2011-10-10,75\n2011-10-11,78\n2011-10-12,72\n2011-10-13,82\n2011-10-14,61\n2011-10-15,61\n2011-10-16,41\n2011-10-17,104\n2011-10-18,62\n2011-10-19,88\n2011-10-20,75\n2011-10-21,97\n2011-10-22,45\n2011-10-23,58\n2011-10-24,85\n2011-10-25,109\n2011-10-26,66\n2011-10-27,45\n2011-10-28,57\n2011-10-29,53\n2011-10-30,50\n2011-10-31,89\n2011-11-01,61\n2011-11-02,78\n2011-11-03,74\n2011-11-04,85\n2011-11-05,84\n2011-11-06,43\n2011-11-07,53\n2011-11-08,48\n2011-11-09,71\n2011-11-10,35\n2011-11-11,72\n2011-11-12,115\n2011-11-13,90\n2011-11-14,64\n2011-11-15,49\n2011-11-16,45\n2011-11-17,79\n2011-11-18,61\n2011-11-19,81\n2011-11-20,78\n2011-11-21,63\n2011-11-22,89\n2011-11-23,66\n2011-11-24,60\n2011-11-25,90\n2011-11-26,108\n2011-11-27,88\n2011-11-28,64\n2011-11-29,72\n2011-11-30,76\n2011-12-01,107\n2011-12-02,73\n2011-12-03,65\n2011-12-04,74\n2011-12-05,80\n2011-12-06,116\n2011-12-07,96\n2011-12-08,113\n2011-12-09,101\n2011-12-10,88\n2011-12-11,142\n2011-12-12,121\n2011-12-13,118\n2011-12-14,86\n2011-12-15,73\n2011-12-16,84\n2011-12-17,56\n2011-12-18,86\n2011-12-19,85\n2011-12-20,89\n2011-12-21,58\n2011-12-22,46\n2011-12-23,65\n2011-12-24,39\n2011-12-25,18\n2011-12-26,39\n2011-12-27,55\n2011-12-28,47\n2011-12-29,71\n2011-12-30,85\n2011-12-31,73\n2012-01-01,56\n2012-01-02,76\n2012-01-03,60\n2012-01-04,121\n2012-01-05,77\n2012-01-06,65\n2012-01-07,60\n2012-01-08,82\n2012-01-09,72\n2012-01-10,106\n2012-01-11,96\n2012-01-12,82\n2012-01-13,145\n2012-01-14,65\n2012-01-15,64\n2012-01-16,71\n2012-01-17,79\n2012-01-18,88\n2012-01-19,82\n2012-01-20,71\n2012-01-21,57\n2012-01-22,61\n2012-01-23,69\n2012-01-24,107\n2012-01-25,99\n2012-01-26,92\n2012-01-27,70\n2012-01-28,44\n2012-01-29,84\n2012-01-30,54\n2012-01-31,71\n2012-02-01,95\n2012-02-02,86\n2012-02-03,87\n2012-02-04,82\n2012-02-05,115\n2012-02-06,189\n2012-02-07,116\n2012-02-08,122\n2012-02-09,106\n2012-02-10,97\n2012-02-11,94\n2012-02-12,90\n2012-02-13,91\n2012-02-14,106\n2012-02-15,93\n2012-02-16,119\n2012-02-17,122\n2012-02-18,109\n2012-02-19,161\n2012-02-20,120\n2012-02-21,144\n2012-02-22,117\n2012-02-23,78\n2012-02-24,68\n2012-02-25,104\n2012-02-26,81\n2012-02-27,87\n2012-02-28,110\n2012-02-29,71\n2012-03-01,83\n2012-03-02,65\n2012-03-03,82\n2012-03-04,133\n2012-03-05,139\n2012-03-06,107\n2012-03-07,77\n2012-03-08,102\n2012-03-09,111\n2012-03-10,100\n2012-03-11,166\n2012-03-12,78\n2012-03-13,100\n2012-03-14,57\n2012-03-15,53\n2012-03-16,99\n2012-03-17,88\n2012-03-18,67\n2012-03-19,160\n2012-03-20,100\n2012-03-21,101\n2012-03-22,99\n2012-03-23,123\n2012-03-24,93\n2012-03-25,142\n2012-03-26,111\n2012-03-27,84\n2012-03-28,67\n2012-03-29,76\n2012-03-30,99\n2012-03-31,109\n2012-04-01,118\n2012-04-02,124\n2012-04-03,109\n2012-04-04,175\n2012-04-05,118\n2012-04-06,92\n2012-04-07,147\n2012-04-08,80\n2012-04-09,135\n2012-04-10,120\n2012-04-11,166\n2012-04-12,77\n2012-04-13,67\n2012-04-14,94\n2012-04-15,102\n2012-04-16,123\n2012-04-17,88\n2012-04-18,130\n2012-04-19,86\n2012-04-20,108\n2012-04-21,117\n2012-04-22,95\n2012-04-23,112\n2012-04-24,106\n2012-04-25,95\n2012-04-26,85\n2012-04-27,107\n2012-04-28,120\n2012-04-29,165\n2012-04-30,127\n2012-05-01,86\n2012-05-02,107\n2012-05-03,92\n2012-05-04,107\n2012-05-05,80\n2012-05-06,103\n2012-05-07,131\n2012-05-08,73\n2012-05-09,105\n2012-05-10,81\n2012-05-11,140\n2012-05-12,83\n2012-05-13,109\n2012-05-14,67\n2012-05-15,71\n2012-05-16,86\n2012-05-17,118\n2012-05-18,101\n2012-05-19,77\n2012-05-20,110\n2012-05-21,77\n2012-05-22,81\n2012-05-23,70\n2012-05-24,79\n2012-05-25,60\n2012-05-26,65\n2012-05-27,46\n2012-05-28,74\n2012-05-29,48\n2012-05-30,90\n2012-05-31,95\n2012-06-01,84\n2012-06-02,31\n2012-06-03,73\n2012-06-04,98\n2012-06-05,63\n2012-06-06,66\n2012-06-07,57\n2012-06-08,78\n2012-06-09,54\n2012-06-10,122\n2012-06-11,95\n2012-06-12,87\n2012-06-13,113\n2012-06-14,79\n2012-06-15,107\n2012-06-16,106\n2012-06-17,96\n2012-06-18,55\n2012-06-19,79\n2012-06-20,94\n2012-06-21,54\n2012-06-22,44\n2012-06-23,67\n2012-06-24,64\n2012-06-25,96\n2012-06-26,73\n2012-06-27,108\n2012-06-28,40\n2012-06-29,87\n2012-06-30,44\n2012-07-01,61\n2012-07-02,91\n2012-07-03,118\n2012-07-04,96\n2012-07-05,123\n2012-07-06,170\n2012-07-07,117\n2012-07-08,136\n2012-07-09,81\n2012-07-10,74\n2012-07-11,96\n2012-07-12,97\n2012-07-13,74\n2012-07-14,83\n2012-07-15,64\n2012-07-16,95\n2012-07-17,119\n2012-07-18,102\n2012-07-19,70\n2012-07-20,66\n2012-07-21,102\n2012-07-22,96\n2012-07-23,107\n2012-07-24,100\n2012-07-25,76\n2012-07-26,96\n2012-07-27,81\n2012-07-28,93\n2012-07-29,74\n2012-07-30,61\n2012-07-31,73\n2012-08-01,100\n2012-08-02,128\n2012-08-03,79\n2012-08-04,106\n2012-08-05,72\n2012-08-06,109\n2012-08-07,107\n2012-08-08,119\n2012-08-09,82\n2012-08-10,100\n2012-08-11,60\n2012-08-12,77\n2012-08-13,68\n2012-08-14,88\n2012-08-15,116\n2012-08-16,94\n2012-08-17,99\n2012-08-18,141\n2012-08-19,46\n2012-08-20,61\n2012-08-21,75\n2012-08-22,84\n2012-08-23,76\n2012-08-24,102\n2012-08-25,71\n2012-08-26,83\n2012-08-27,91\n2012-08-28,80\n2012-08-29,60\n2012-08-30,78\n2012-08-31,61\n2012-09-01,57\n2012-09-02,41\n2012-09-03,86\n2012-09-04,88\n2012-09-05,73\n2012-09-06,100\n2012-09-07,106\n2012-09-08,57\n2012-09-09,40\n2012-09-10,65\n2012-09-11,59\n2012-09-12,99\n2012-09-13,55\n2012-09-14,64\n2012-09-15,40\n2012-09-16,52\n2012-09-17,57\n2012-09-18,51\n2012-09-19,38\n2012-09-20,47\n2012-09-21,53\n2012-09-22,69\n2012-09-23,40\n2012-09-24,32\n2012-09-25,45\n2012-09-26,54\n2012-09-27,59\n2012-09-28,61\n2012-09-29,58\n2012-09-30,62\n2012-10-01,96\n2012-10-02,101\n2012-10-03,98\n2012-10-04,78\n2012-10-05,107\n2012-10-06,59\n2012-10-07,48\n2012-10-08,60\n2012-10-09,76\n2012-10-10,126\n2012-10-11,58\n2012-10-12,70\n2012-10-13,37\n2012-10-14,70\n2012-10-15,79\n2012-10-16,77\n2012-10-17,33\n2012-10-18,86\n2012-10-19,62\n2012-10-20,36\n2012-10-21,18\n2012-10-22,55\n2012-10-23,78\n2012-10-24,58\n2012-10-25,87\n2012-10-26,112\n2012-10-27,48\n2012-10-28,53\n2012-10-29,38\n2012-10-30,92\n2012-10-31,54\n2012-11-01,109\n2012-11-02,119\n2012-11-03,71\n2012-11-04,59\n2012-11-05,91\n2012-11-06,39\n2012-11-07,56\n2012-11-08,66\n2012-11-09,56\n2012-11-10,39\n2012-11-11,20\n2012-11-12,68\n2012-11-13,66\n2012-11-14,58\n2012-11-15,47\n2012-11-16,70\n2012-11-17,67\n2012-11-18,38\n2012-11-19,95\n2012-11-20,114\n2012-11-21,87\n2012-11-22,49\n2012-11-23,80\n2012-11-24,70\n2012-11-25,35\n2012-11-26,71\n2012-11-27,79\n2012-11-28,64\n2012-11-29,58\n2012-11-30,36\n2012-12-01,56\n2012-12-02,58\n2012-12-03,84\n2012-12-04,103\n2012-12-05,118\n2012-12-06,46\n2012-12-07,58\n2012-12-08,98\n2012-12-09,43\n2012-12-10,70\n2012-12-11,91\n2012-12-12,74\n2012-12-13,88\n2012-12-14,74\n2012-12-15,54\n2012-12-16,81\n2012-12-17,121\n2012-12-18,111\n2012-12-19,112\n2012-12-20,103\n2012-12-21,63\n2012-12-22,59\n2012-12-23,91\n2012-12-24,99\n2012-12-25,60\n2012-12-26,61\n2012-12-27,90\n2012-12-28,59\n2012-12-29,69\n2012-12-30,66\n2012-12-31,110\n2013-01-01,65\n2013-01-02,73\n2013-01-03,92\n2013-01-04,93\n2013-01-05,67\n2013-01-06,84\n2013-01-07,176\n2013-01-08,150\n2013-01-09,143\n2013-01-10,195\n2013-01-11,158\n2013-01-12,99\n2013-01-13,156\n2013-01-14,114\n2013-01-15,109\n2013-01-16,182\n2013-01-17,139\n2013-01-18,96\n2013-01-19,82\n2013-01-20,116\n2013-01-21,88\n2013-01-22,124\n2013-01-23,120\n2013-01-24,108\n2013-01-25,143\n2013-01-26,142\n2013-01-27,69\n2013-01-28,128\n2013-01-29,114\n2013-01-30,120\n2013-01-31,96\n2013-02-01,69\n2013-02-02,104\n2013-02-03,156\n2013-02-04,90\n2013-02-05,116\n2013-02-06,81\n2013-02-07,117\n2013-02-08,123\n2013-02-09,80\n2013-02-10,83\n2013-02-11,100\n2013-02-12,115\n2013-02-13,153\n2013-02-14,95\n2013-02-15,146\n2013-02-16,111\n2013-02-17,118\n2013-02-18,119\n2013-02-19,120\n2013-02-20,142\n2013-02-21,127\n2013-02-22,120\n2013-02-23,44\n2013-02-24,131\n2013-02-25,166\n2013-02-26,176\n2013-02-27,113\n2013-02-28,116\n2013-03-01,112\n2013-03-02,72\n2013-03-03,73\n2013-03-04,123\n2013-03-05,111\n2013-03-06,102\n2013-03-07,99\n2013-03-08,109\n2013-03-09,108\n2013-03-10,130\n2013-03-11,121\n2013-03-12,85\n2013-03-13,118\n2013-03-14,129\n2013-03-15,139\n2013-03-16,98\n2013-03-17,137\n2013-03-18,139\n2013-03-19,114\n2013-03-20,153\n2013-03-21,88\n2013-03-22,151\n2013-03-23,105\n2013-03-24,107\n2013-03-25,137\n2013-03-26,114\n2013-03-27,107\n2013-03-28,113\n2013-03-29,150\n2013-03-30,83\n2013-03-31,95\n2013-04-01,124\n2013-04-02,132\n2013-04-03,116\n2013-04-04,135\n2013-04-05,80\n2013-04-06,90\n2013-04-07,149\n2013-04-08,178\n2013-04-09,164\n2013-04-10,120\n2013-04-11,111\n2013-04-12,138\n2013-04-13,123\n2013-04-14,132\n2013-04-15,111\n2013-04-16,90\n2013-04-17,90\n2013-04-18,143\n2013-04-19,110\n2013-04-20,79\n2013-04-21,128\n2013-04-22,93\n2013-04-23,70\n2013-04-24,85\n2013-04-25,83\n2013-04-26,95\n2013-04-27,82\n2013-04-28,96\n2013-04-29,91\n2013-04-30,130\n2013-05-01,117\n2013-05-02,134\n2013-05-03,113\n2013-05-04,97\n2013-05-05,124\n2013-05-06,120\n2013-05-07,98\n2013-05-08,112\n2013-05-09,155\n2013-05-10,135\n2013-05-11,88\n2013-05-12,64\n2013-05-13,123\n2013-05-14,88\n2013-05-15,162\n2013-05-16,136\n2013-05-17,143\n2013-05-18,71\n2013-05-19,89\n2013-05-20,106\n2013-05-21,136\n2013-05-22,115\n2013-05-23,91\n2013-05-24,148\n2013-05-25,70\n2013-05-26,47\n2013-05-27,135\n2013-05-28,150\n2013-05-29,88\n2013-05-30,102\n2013-05-31,82\n2013-06-01,79\n2013-06-02,86\n2013-06-03,87\n2013-06-04,100\n2013-06-05,121\n2013-06-06,112\n2013-06-07,167\n2013-06-08,103\n2013-06-09,83\n2013-06-10,132\n2013-06-11,109\n2013-06-12,133\n2013-06-13,160\n2013-06-14,178\n2013-06-15,73\n2013-06-16,138\n2013-06-17,111\n2013-06-18,112\n2013-06-19,126\n2013-06-20,146\n2013-06-21,122\n2013-06-22,122\n2013-06-23,98\n2013-06-24,86\n2013-06-25,128\n2013-06-26,160\n2013-06-27,115\n2013-06-28,103\n2013-06-29,57\n2013-06-30,76\n2013-07-01,132\n2013-07-02,134\n2013-07-03,95\n2013-07-04,62\n2013-07-05,131\n2013-07-06,90\n2013-07-07,92\n2013-07-08,142\n2013-07-09,111\n2013-07-10,109\n2013-07-11,93\n2013-07-12,104\n2013-07-13,39\n2013-07-14,67\n2013-07-15,125\n2013-07-16,114\n2013-07-17,89\n2013-07-18,107\n2013-07-19,142\n2013-07-20,50\n2013-07-21,58\n2013-07-22,166\n2013-07-23,149\n2013-07-24,169\n2013-07-25,71\n2013-07-26,160\n2013-07-27,50\n2013-07-28,100\n2013-07-29,108\n2013-07-30,206\n2013-07-31,181\n2013-08-01,188\n2013-08-02,143\n2013-08-03,121\n2013-08-04,82\n2013-08-05,117\n2013-08-06,138\n2013-08-07,143\n2013-08-08,117\n2013-08-09,149\n2013-08-10,69\n2013-08-11,68\n2013-08-12,98\n2013-08-13,89\n2013-08-14,108\n2013-08-15,115\n2013-08-16,99\n2013-08-17,62\n2013-08-18,66\n2013-08-19,157\n2013-08-20,126\n2013-08-21,129\n2013-08-22,122\n2013-08-23,124\n2013-08-24,101\n2013-08-25,109\n2013-08-26,175\n2013-08-27,155\n2013-08-28,159\n2013-08-29,116\n2013-08-30,133\n2013-08-31,112\n2013-09-01,103\n2013-09-02,83\n2013-09-03,115\n2013-09-04,141\n2013-09-05,115\n2013-09-06,77\n2013-09-07,99\n2013-09-08,93\n2013-09-09,163\n2013-09-10,143\n2013-09-11,86\n2013-09-12,137\n2013-09-13,135\n2013-09-14,83\n2013-09-15,130\n2013-09-16,123\n2013-09-17,104\n2013-09-18,97\n2013-09-19,104\n2013-09-20,118\n2013-09-21,88\n2013-09-22,88\n2013-09-23,105\n2013-09-24,219\n2013-09-25,188\n2013-09-26,108\n2013-09-27,123\n2013-09-28,81\n2013-09-29,83\n2013-09-30,99\n2013-10-01,80\n2013-10-02,99\n2013-10-03,110\n2013-10-04,121\n2013-10-05,103\n2013-10-06,141\n2013-10-07,155\n2013-10-08,177\n2013-10-09,117\n2013-10-10,117\n2013-10-11,202\n2013-10-12,103\n2013-10-13,140\n2013-10-14,94\n2013-10-15,128\n2013-10-16,167\n2013-10-17,96\n2013-10-18,148\n2013-10-19,89\n2013-10-20,122\n2013-10-21,113\n2013-10-22,160\n2013-10-23,115\n2013-10-24,134\n2013-10-25,160\n2013-10-26,58\n2013-10-27,110\n2013-10-28,191\n2013-10-29,113\n2013-10-30,169\n2013-10-31,143\n2013-11-01,119\n2013-11-02,161\n2013-11-03,104\n2013-11-04,103\n2013-11-05,165\n2013-11-06,121\n2013-11-07,156\n2013-11-08,177\n2013-11-09,85\n2013-11-10,81\n2013-11-11,194\n2013-11-12,180\n2013-11-13,163\n2013-11-14,106\n2013-11-15,97\n2013-11-16,39\n2013-11-17,93\n2013-11-18,127\n2013-11-19,129\n2013-11-20,126\n2013-11-21,109\n2013-11-22,124\n2013-11-23,76\n2013-11-24,107\n2013-11-25,117\n2013-11-26,136\n2013-11-27,233\n2013-11-28,37\n2013-11-29,94\n2013-11-30,138\n2013-12-01,94\n2013-12-02,119\n2013-12-03,152\n2013-12-04,123\n2013-12-05,141\n2013-12-06,144\n2013-12-07,131\n2013-12-08,116\n2013-12-09,149\n2013-12-10,173\n2013-12-11,118\n2013-12-12,172\n2013-12-13,165\n2013-12-14,115\n2013-12-15,84\n2013-12-16,122\n2013-12-17,150\n2013-12-18,124\n2013-12-19,140\n2013-12-20,131\n2013-12-21,106\n2013-12-22,97\n2013-12-23,82\n2013-12-24,50\n2013-12-25,29\n2013-12-26,60\n2013-12-27,98\n2013-12-28,66\n2013-12-29,106\n2013-12-30,173\n2013-12-31,167\n2014-01-01,74\n2014-01-02,128\n2014-01-03,201\n2014-01-04,123\n2014-01-05,93\n2014-01-06,155\n2014-01-07,237\n2014-01-08,207\n2014-01-09,202\n2014-01-10,203\n2014-01-11,124\n2014-01-12,112\n2014-01-13,135\n2014-01-14,119\n2014-01-15,186\n2014-01-16,197\n2014-01-17,107\n2014-01-18,105\n2014-01-19,150\n2014-01-20,116\n2014-01-21,143\n2014-01-22,164\n2014-01-23,234\n2014-01-24,136\n2014-01-25,123\n2014-01-26,102\n2014-01-27,164\n2014-01-28,166\n2014-01-29,151\n2014-01-30,136\n2014-01-31,102\n2014-02-01,73\n2014-02-02,122\n2014-02-03,167\n2014-02-04,209\n2014-02-05,171\n2014-02-06,161\n2014-02-07,137\n2014-02-08,131\n2014-02-09,208\n2014-02-10,227\n2014-02-11,212\n2014-02-12,135\n2014-02-13,191\n2014-02-14,161\n2014-02-15,152\n2014-02-16,101\n2014-02-17,184\n2014-02-18,307\n2014-02-19,165\n2014-02-20,122\n2014-02-21,183\n2014-02-22,139\n2014-02-23,132\n2014-02-24,181\n2014-02-25,173\n2014-02-26,161\n2014-02-27,113\n2014-02-28,103\n2014-03-01,91\n2014-03-02,76\n2014-03-03,126\n2014-03-04,109\n2014-03-05,119\n2014-03-06,162\n2014-03-07,115\n2014-03-08,118\n2014-03-09,73\n2014-03-10,140\n2014-03-11,158\n2014-03-12,119\n2014-03-13,139\n2014-03-14,179\n2014-03-15,110\n2014-03-16,80\n2014-03-17,144\n2014-03-18,114\n2014-03-19,132\n2014-03-20,129\n2014-03-21,167\n2014-03-22,132\n2014-03-23,102\n2014-03-24,187\n2014-03-25,148\n2014-03-26,223\n2014-03-27,145\n2014-03-28,162\n2014-03-29,91\n2014-03-30,100\n2014-03-31,147\n2014-04-01,116\n2014-04-02,154\n2014-04-03,107\n2014-04-04,103\n2014-04-05,94\n2014-04-06,86\n2014-04-07,140\n2014-04-08,142\n2014-04-09,118\n2014-04-10,149\n2014-04-11,90\n2014-04-12,82\n2014-04-13,94\n2014-04-14,76\n2014-04-15,128\n2014-04-16,157\n2014-04-17,130\n2014-04-18,191\n2014-04-19,127\n2014-04-20,102\n2014-04-21,179\n2014-04-22,153\n2014-04-23,167\n2014-04-24,232\n2014-04-25,182\n2014-04-26,119\n2014-04-27,187\n2014-04-28,161\n2014-04-29,190\n2014-04-30,170\n2014-05-01,137\n2014-05-02,151\n2014-05-03,107\n2014-05-04,108\n2014-05-05,193\n2014-05-06,114\n2014-05-07,247\n2014-05-08,230\n2014-05-09,264\n2014-05-10,175\n2014-05-11,127\n2014-05-12,144\n2014-05-13,212\n2014-05-14,236\n2014-05-15,154\n2014-05-16,155\n2014-05-17,140\n2014-05-18,105\n2014-05-19,145\n2014-05-20,156\n2014-05-21,149\n2014-05-22,190\n2014-05-23,196\n2014-05-24,102\n2014-05-25,150\n2014-05-26,124\n2014-05-27,134\n2014-05-28,148\n2014-05-29,185\n2014-05-30,174\n2014-05-31,119\n2014-06-01,182\n2014-06-02,133\n2014-06-03,195\n2014-06-04,183\n2014-06-05,186\n2014-06-06,254\n2014-06-07,141\n2014-06-08,130\n2014-06-09,230\n2014-06-10,203\n2014-06-11,164\n2014-06-12,212\n2014-06-13,169\n2014-06-14,129\n2014-06-15,100\n2014-06-16,160\n2014-06-17,175\n2014-06-18,154\n2014-06-19,197\n2014-06-20,146\n2014-06-21,44\n2014-06-22,81\n2014-06-23,153\n2014-06-24,174\n2014-06-25,174\n2014-06-26,168\n2014-06-27,173\n2014-06-28,96\n2014-06-29,63\n2014-06-30,139\n2014-07-01,160\n2014-07-02,179\n2014-07-03,151\n2014-07-04,134\n2014-07-05,121\n2014-07-06,144\n2014-07-07,151\n2014-07-08,120\n2014-07-09,160\n2014-07-10,160\n2014-07-11,198\n2014-07-12,116\n2014-07-13,125\n2014-07-14,115\n2014-07-15,127\n2014-07-16,242\n2014-07-17,147\n2014-07-18,199\n2014-07-19,103\n2014-07-20,159\n2014-07-21,241\n2014-07-22,189\n2014-07-23,149\n2014-07-24,215\n2014-07-25,136\n2014-07-26,105\n2014-07-27,80\n2014-07-28,126\n2014-07-29,176\n2014-07-30,145\n2014-07-31,108\n2014-08-01,205\n2014-08-02,125\n2014-08-03,93\n2014-08-04,162\n2014-08-05,229\n2014-08-06,253\n2014-08-07,226\n2014-08-08,177\n2014-08-09,133\n2014-08-10,164\n2014-08-11,199\n2014-08-12,162\n2014-08-13,121\n2014-08-14,157\n2014-08-15,123\n2014-08-16,96\n2014-08-17,139\n2014-08-18,134\n2014-08-19,186\n2014-08-20,182\n2014-08-21,205\n2014-08-22,169\n2014-08-23,93\n2014-08-24,110\n2014-08-25,167\n2014-08-26,218\n2014-08-27,188\n2014-08-28,237\n2014-08-29,163\n2014-08-30,126\n2014-08-31,127\n2014-09-01,181\n2014-09-02,240\n2014-09-03,195\n2014-09-04,181\n2014-09-05,177\n2014-09-06,119\n2014-09-07,226\n2014-09-08,133\n2014-09-09,176\n2014-09-10,207\n2014-09-11,146\n2014-09-12,145\n2014-09-13,89\n2014-09-14,125\n2014-09-15,205\n2014-09-16,189\n2014-09-17,165\n2014-09-18,232\n2014-09-19,192\n2014-09-20,124\n2014-09-21,143\n2014-09-22,171\n2014-09-23,192\n2014-09-24,207\n2014-09-25,183\n2014-09-26,201\n2014-09-27,268\n2014-09-28,335\n2014-09-29,244\n2014-09-30,256\n2014-10-01,321\n2014-10-02,241\n2014-10-03,199\n2014-10-04,118\n2014-10-05,93\n2014-10-06,149\n2014-10-07,173\n2014-10-08,169\n2014-10-09,337\n2014-10-10,155\n2014-10-11,89\n2014-10-12,128\n2014-10-13,135\n2014-10-14,190\n2014-10-15,181\n2014-10-16,187\n2014-10-17,181\n2014-10-18,139\n2014-10-19,157\n2014-10-20,155\n2014-10-21,165\n2014-10-22,124\n2014-10-23,211\n2014-10-24,158\n2014-10-25,86\n2014-10-26,65\n2014-10-27,132\n2014-10-28,135\n2014-10-29,227\n2014-10-30,187\n2014-10-31,156\n2014-11-01,86\n2014-11-02,81\n2014-11-03,152\n2014-11-04,167\n2014-11-05,216\n2014-11-06,229\n2014-11-07,208\n2014-11-08,110\n2014-11-09,205\n2014-11-10,161\n2014-11-11,165\n2014-11-12,167\n2014-11-13,162\n2014-11-14,174\n2014-11-15,153\n2014-11-16,204\n2014-11-17,219\n2014-11-18,203\n2014-11-19,180\n2014-11-20,129\n2014-11-21,168\n2014-11-22,164\n2014-11-23,141\n2014-11-24,109\n2014-11-25,176\n2014-11-26,154\n2014-11-27,90\n2014-11-28,93\n2014-11-29,116\n2014-11-30,174\n2014-12-01,184\n2014-12-02,134\n2014-12-03,129\n2014-12-04,126\n2014-12-05,132\n2014-12-06,115\n2014-12-07,160\n2014-12-08,218\n2014-12-09,158\n2014-12-10,127\n2014-12-11,140\n2014-12-12,135\n2014-12-13,136\n2014-12-14,160\n2014-12-15,196\n2014-12-16,191\n2014-12-17,138\n2014-12-18,104\n2014-12-19,205\n2014-12-20,147\n2014-12-21,196\n2014-12-22,194\n2014-12-23,142\n2014-12-24,124\n2014-12-25,70\n2014-12-26,112\n2014-12-27,193\n2014-12-28,118\n2014-12-29,103\n2014-12-30,214\n2014-12-31,134\n2015-01-01,103\n2015-01-02,223\n2015-01-03,188\n2015-01-04,143\n2015-01-05,165\n2015-01-06,185\n2015-01-07,145\n2015-01-08,252\n2015-01-09,185\n2015-01-10,160\n2015-01-11,146\n2015-01-12,186\n2015-01-13,211\n2015-01-14,131\n2015-01-15,187\n2015-01-16,174\n2015-01-17,164\n2015-01-18,148\n2015-01-19,233\n2015-01-20,139\n2015-01-21,160\n2015-01-22,156\n2015-01-23,161\n2015-01-24,134\n2015-01-25,120\n2015-01-26,174\n2015-01-27,166\n2015-01-28,217\n2015-01-29,146\n2015-01-30,205\n2015-01-31,185\n2015-02-01,126\n2015-02-02,182\n2015-02-03,148\n2015-02-04,182\n2015-02-05,256\n2015-02-06,218\n2015-02-07,156\n2015-02-08,163\n2015-02-09,195\n2015-02-10,232\n2015-02-11,189\n2015-02-12,216\n2015-02-13,152\n2015-02-14,120\n2015-02-15,203\n2015-02-16,177\n2015-02-17,215\n2015-02-18,241\n2015-02-19,190\n2015-02-20,266\n2015-02-21,285\n2015-02-22,205\n2015-02-23,278\n2015-02-24,197\n2015-02-25,285\n2015-02-26,175\n2015-02-27,185\n2015-02-28,158\n2015-03-01,140\n2015-03-02,247\n2015-03-03,230\n2015-03-04,200\n2015-03-05,283\n2015-03-06,271\n2015-03-07,187\n2015-03-08,117\n2015-03-09,267\n2015-03-10,204\n2015-03-11,276\n2015-03-12,207\n2015-03-13,289\n2015-03-14,155\n2015-03-15,107\n2015-03-16,314\n2015-03-17,266\n2015-03-18,183\n2015-03-19,228\n2015-03-20,217\n2015-03-21,101\n2015-03-22,187\n2015-03-23,230\n2015-03-24,254\n2015-03-25,222\n2015-03-26,228\n2015-03-27,242\n2015-03-28,157\n2015-03-29,108\n2015-03-30,252\n2015-03-31,201\n2015-04-01,237\n2015-04-02,175\n2015-04-03,222\n2015-04-04,107\n2015-04-05,91\n2015-04-06,198\n2015-04-07,181\n2015-04-08,215\n2015-04-09,230\n2015-04-10,141\n2015-04-11,94\n2015-04-12,72\n2015-04-13,180\n2015-04-14,258\n2015-04-15,178\n2015-04-16,210\n2015-04-17,246\n2015-04-18,111\n2015-04-19,139\n2015-04-20,134\n2015-04-21,160\n2015-04-22,186\n2015-04-23,193\n2015-04-24,201\n2015-04-25,140\n2015-04-26,164\n2015-04-27,185\n2015-04-28,176\n2015-04-29,262\n2015-04-30,216\n2015-05-01,188\n2015-05-02,142\n2015-05-03,172\n2015-05-04,173\n2015-05-05,174\n2015-05-06,202\n2015-05-07,157\n2015-05-08,222\n2015-05-09,97\n2015-05-10,97\n2015-05-11,193\n2015-05-12,176\n2015-05-13,193\n2015-05-14,235\n2015-05-15,190\n2015-05-16,116\n2015-05-17,159\n2015-05-18,312\n2015-05-19,185\n2015-05-20,242\n2015-05-21,164\n2015-05-22,175\n2015-05-23,75\n2015-05-24,115\n2015-05-25,141\n2015-05-26,248\n2015-05-27,220\n2015-05-28,196\n2015-05-29,196\n2015-05-30,152\n2015-05-31,109\n2015-06-01,193\n2015-06-02,184\n2015-06-03,155\n2015-06-04,134\n2015-06-05,128\n2015-06-06,98\n2015-06-07,105\n2015-06-08,161\n2015-06-09,176\n2015-06-10,267\n2015-06-11,184\n2015-06-12,195\n2015-06-13,108\n2015-06-14,182\n2015-06-15,195\n2015-06-16,240\n2015-06-17,279\n2015-06-18,209\n2015-06-19,251\n2015-06-20,157\n2015-06-21,120\n2015-06-22,282\n2015-06-23,223\n2015-06-24,233\n2015-06-25,199\n2015-06-26,156\n2015-06-27,79\n2015-06-28,159\n2015-06-29,192\n2015-06-30,234\n2015-07-01,252\n2015-07-02,191\n2015-07-03,151\n2015-07-04,107\n2015-07-05,95\n2015-07-06,155\n2015-07-07,224\n2015-07-08,201\n2015-07-09,189\n2015-07-10,184\n2015-07-11,142\n2015-07-12,130\n2015-07-13,149\n2015-07-14,235\n2015-07-15,166\n2015-07-16,258\n2015-07-17,236\n2015-07-18,182\n2015-07-19,129\n2015-07-20,237\n2015-07-21,275\n2015-07-22,151\n2015-07-23,211\n2015-07-24,181\n2015-07-25,125\n2015-07-26,125\n2015-07-27,206\n2015-07-28,211\n2015-07-29,176\n2015-07-30,194\n2015-07-31,210\n2015-08-01,134\n2015-08-02,153\n2015-08-03,224\n2015-08-04,221\n2015-08-05,221\n2015-08-06,217\n2015-08-07,163\n2015-08-08,99\n2015-08-09,95\n2015-08-10,197\n2015-08-11,186\n2015-08-12,225\n2015-08-13,203\n2015-08-14,170\n2015-08-15,93\n2015-08-16,108\n2015-08-17,112\n2015-08-18,170\n2015-08-19,198\n2015-08-20,200\n2015-08-21,203\n2015-08-22,105\n2015-08-23,119\n2015-08-24,186\n2015-08-25,175\n2015-08-26,205\n2015-08-27,197\n2015-08-28,174\n2015-08-29,52\n2015-08-30,127\n2015-08-31,199\n2015-09-01,219\n2015-09-02,168\n2015-09-03,251\n2015-09-04,153\n2015-09-05,102\n2015-09-06,82\n2015-09-07,114\n2015-09-08,196\n2015-09-09,272\n2015-09-10,228\n2015-09-11,241\n2015-09-12,136\n2015-09-13,108\n2015-09-14,216\n2015-09-15,197\n2015-09-16,199\n2015-09-17,209\n2015-09-18,197\n2015-09-19,148\n2015-09-20,168\n2015-09-21,194\n2015-09-22,193\n2015-09-23,152\n2015-09-24,166\n2015-09-25,215\n2015-09-26,161\n2015-09-27,139\n2015-09-28,253\n2015-09-29,211\n2015-09-30,232\n2015-10-01,206\n2015-10-02,213\n2015-10-03,129\n2015-10-04,114\n2015-10-05,227\n2015-10-06,240\n2015-10-07,202\n2015-10-08,219\n2015-10-09,190\n2015-10-10,146\n2015-10-11,145\n2015-10-12,125\n2015-10-13,218\n2015-10-14,247\n2015-10-15,201\n2015-10-16,184\n2015-10-17,112\n2015-10-18,128\n2015-10-19,276\n2015-10-20,190\n2015-10-21,251\n2015-10-22,145\n2015-10-23,159\n2015-10-24,104\n2015-10-25,153\n2015-10-26,212\n2015-10-27,179\n2015-10-28,166\n2015-10-29,180\n2015-10-30,188\n2015-10-31,65\n2015-11-01,120\n2015-11-02,225\n2015-11-03,196\n2015-11-04,229\n2015-11-05,211\n2015-11-06,181\n2015-11-07,111\n2015-11-08,167\n2015-11-09,213\n2015-11-10,256\n2015-11-11,297\n2015-11-12,223\n2015-11-13,290\n2015-11-14,139\n2015-11-15,131\n2015-11-16,221\n2015-11-17,189\n2015-11-18,208\n2015-11-19,235\n2015-11-20,245\n2015-11-21,160\n2015-11-22,150\n2015-11-23,212\n2015-11-24,316\n2015-11-25,256\n2015-11-26,96\n2015-11-27,114\n2015-11-28,152\n2015-11-29,160\n2015-11-30,165\n2015-12-01,230\n2015-12-02,192\n2015-12-03,171\n2015-12-04,149\n2015-12-05,138\n2015-12-06,268\n2015-12-07,200\n2015-12-08,248\n2015-12-09,213\n2015-12-10,241\n2015-12-11,232\n2015-12-12,95\n2015-12-13,120\n2015-12-14,220\n2015-12-15,258\n2015-12-16,230\n2015-12-17,262\n2015-12-18,226\n2015-12-19,163\n2015-12-20,131\n2015-12-21,220\n2015-12-22,153\n2015-12-23,192\n2015-12-24,96\n2015-12-25,53\n2015-12-26,85\n2015-12-27,96\n2015-12-28,139\n2015-12-29,166\n2015-12-30,156\n2015-12-31,128\n2016-01-01,131\n2016-01-02,151\n2016-01-03,154\n2016-01-04,185\n2016-01-05,226\n2016-01-06,276\n2016-01-07,310\n2016-01-08,242\n2016-01-09,197\n2016-01-10,159\n2016-01-11,330\n2016-01-12,212\n2016-01-13,261\n2016-01-14,294\n2016-01-15,263\n2016-01-16,228\n2016-01-17,200\n2016-01-18,203\n2016-01-19,214\n2016-01-20,195\n2016-01-21,268\n2016-01-22,209\n2016-01-23,169\n2016-01-24,214\n2016-01-25,251\n2016-01-26,224\n2016-01-27,211\n2016-01-28,238\n2016-01-29,316\n2016-01-30,187\n2016-01-31,253\n2016-02-01,271\n2016-02-02,307\n2016-02-03,221\n2016-02-04,209\n2016-02-05,201\n2016-02-06,165\n2016-02-07,163\n2016-02-08,191\n2016-02-09,217\n2016-02-10,192\n2016-02-11,218\n2016-02-12,291\n2016-02-13,137\n2016-02-14,145\n2016-02-15,182\n2016-02-16,249\n2016-02-17,238\n2016-02-18,224\n2016-02-19,258\n2016-02-20,136\n2016-02-21,168\n2016-02-22,211\n2016-02-23,256\n2016-02-24,232\n2016-02-25,242\n2016-02-26,232\n2016-02-27,179\n2016-02-28,246\n2016-02-29,257\n2016-03-01,208\n2016-03-02,178\n2016-03-03,208\n2016-03-04,208\n2016-03-05,127\n2016-03-06,119\n2016-03-07,208\n2016-03-08,236\n2016-03-09,191\n2016-03-10,136\n2016-03-11,180\n2016-03-12,98\n2016-03-13,97\n2016-03-14,142\n2016-03-15,180\n2016-03-16,245\n2016-03-17,232\n2016-03-18,188\n2016-03-19,119\n2016-03-20,159\n2016-03-21,216\n2016-03-22,200\n2016-03-23,228\n2016-03-24,184\n2016-03-25,272\n2016-03-26,145\n2016-03-27,101\n2016-03-28,241\n2016-03-29,204\n2016-03-30,153\n2016-03-31,177\n2016-04-01,245\n2016-04-02,175\n2016-04-03,175\n2016-04-04,206\n2016-04-05,190\n2016-04-06,189\n2016-04-07,174\n2016-04-08,182\n2016-04-09,182\n2016-04-10,124\n2016-04-11,205\n2016-04-12,218\n2016-04-13,233\n2016-04-14,231\n2016-04-15,155\n2016-04-16,174\n2016-04-17,61\n2016-04-18,197\n2016-04-19,197\n2016-04-20,270\n2016-04-21,174\n2016-04-22,166\n2016-04-23,131\n2016-04-24,89\n2016-04-25,185\n2016-04-26,155\n2016-04-27,147\n2016-04-28,129\n2016-04-29,182\n2016-04-30,114\n2016-05-01,114\n2016-05-02,268\n2016-05-03,157\n2016-05-04,184\n2016-05-05,176\n2016-05-06,162\n2016-05-07,147\n2016-05-08,80\n2016-05-09,190\n2016-05-10,205\n2016-05-11,190\n2016-05-12,140\n2016-05-13,192\n2016-05-14,77\n2016-05-15,98\n2016-05-16,187\n2016-05-17,144\n2016-05-18,160\n2016-05-19,197\n2016-05-20,183\n2016-05-21,135\n2016-05-22,150\n2016-05-23,211\n2016-05-24,236\n2016-05-25,167\n2016-05-26,160\n2016-05-27,131\n2016-05-28,97\n2016-05-29,118\n2016-05-30,162\n2016-05-31,206\n2016-06-01,280\n2016-06-02,224\n2016-06-03,255\n2016-06-04,130\n2016-06-05,119\n2016-06-06,176\n2016-06-07,212\n2016-06-08,211\n2016-06-09,209\n2016-06-10,195\n2016-06-11,178\n2016-06-12,129\n2016-06-13,292\n2016-06-14,211\n2016-06-15,196\n2016-06-16,192\n2016-06-17,199\n2016-06-18,116\n2016-06-19,103\n2016-06-20,271\n2016-06-21,233\n2016-06-22,220\n2016-06-23,175\n2016-06-24,165\n2016-06-25,67\n2016-06-26,116\n2016-06-27,271\n2016-06-28,214\n2016-06-29,193\n2016-06-30,243\n2016-07-01,218\n2016-07-02,170\n2016-07-03,102\n2016-07-04,123\n2016-07-05,148\n2016-07-06,222\n2016-07-07,204\n2016-07-08,172\n2016-07-09,115\n2016-07-10,115\n2016-07-11,175\n2016-07-12,189\n2016-07-13,168\n2016-07-14,164\n2016-07-15,159\n2016-07-16,91\n2016-07-17,135\n2016-07-18,210\n2016-07-19,208\n2016-07-20,199\n2016-07-21,155\n2016-07-22,184\n2016-07-23,98\n2016-07-24,117\n2016-07-25,200\n2016-07-26,233\n2016-07-27,221\n2016-07-28,267\n2016-07-29,219\n2016-07-30,122\n2016-07-31,133\n2016-08-01,178\n2016-08-02,170\n2016-08-03,131\n2016-08-04,230\n2016-08-05,155\n2016-08-06,71\n2016-08-07,105\n2016-08-08,158\n2016-08-09,205\n2016-08-10,227\n2016-08-11,206\n2016-08-12,194\n2016-08-13,66\n2016-08-14,97\n2016-08-15,165\n2016-08-16,170\n2016-08-17,191\n2016-08-18,180\n2016-08-19,181\n2016-08-20,126\n2016-08-21,143\n2016-08-22,207\n2016-08-23,227\n2016-08-24,206\n2016-08-25,242\n2016-08-26,140\n2016-08-27,104\n2016-08-28,111\n2016-08-29,192\n2016-08-30,182\n2016-08-31,254\n2016-09-01,265\n2016-09-02,143\n2016-09-03,74\n2016-09-04,122\n2016-09-05,130\n2016-09-06,213\n2016-09-07,177\n2016-09-08,169\n2016-09-09,174\n2016-09-10,65\n2016-09-11,89\n2016-09-12,153\n2016-09-13,119\n2016-09-14,141\n2016-09-15,179\n2016-09-16,190\n2016-09-17,85\n2016-09-18,150\n2016-09-19,200\n2016-09-20,169\n2016-09-21,149\n2016-09-22,170\n2016-09-23,131\n2016-09-24,76\n2016-09-25,78\n2016-09-26,163\n2016-09-27,187\n2016-09-28,259\n2016-09-29,154\n2016-09-30,150\n2016-10-01,98\n2016-10-02,103\n2016-10-03,199\n2016-10-04,213\n2016-10-05,263\n2016-10-06,212\n2016-10-07,226\n2016-10-08,135\n2016-10-09,93\n2016-10-10,181\n2016-10-11,166\n2016-10-12,224\n2016-10-13,221\n2016-10-14,194\n2016-10-15,97\n2016-10-16,126\n2016-10-17,141\n2016-10-18,146\n2016-10-19,114\n2016-10-20,142\n2016-10-21,182\n2016-10-22,95\n2016-10-23,124\n2016-10-24,132\n2016-10-25,140\n2016-10-26,148\n2016-10-27,124\n2016-10-28,123\n2016-10-29,156\n2016-10-30,81\n2016-10-31,82\n2016-11-01,127\n2016-11-02,146\n2016-11-03,156\n2016-11-04,148\n2016-11-05,78\n2016-11-06,103\n2016-11-07,156\n2016-11-08,120\n2016-11-09,147\n2016-11-10,148\n2016-11-11,114\n2016-11-12,76\n2016-11-13,158\n2016-11-14,179\n2016-11-15,214\n2016-11-16,142\n2016-11-17,184\n2016-11-18,162\n2016-11-19,97\n2016-11-20,103\n2016-11-21,162\n2016-11-22,112\n2016-11-23,86\n2016-11-24,75\n2016-11-25,56\n2016-11-26,47\n2016-11-27,78\n2016-11-28,141\n2016-11-29,169\n2016-11-30,138\n2016-12-01,137\n2016-12-02,147\n2016-12-03,123\n2016-12-04,107\n2016-12-05,214\n2016-12-06,213\n2016-12-07,190\n2016-12-08,168\n2016-12-09,156\n2016-12-10,87\n2016-12-11,106\n2016-12-12,210\n2016-12-13,227\n2016-12-14,192\n2016-12-15,163\n2016-12-16,89\n2016-12-17,95\n2016-12-18,90\n2016-12-19,144\n2016-12-20,138\n2016-12-21,152\n2016-12-22,159\n2016-12-23,122\n2016-12-24,67\n2016-12-25,57\n2016-12-26,75\n2016-12-27,111\n2016-12-28,117\n2016-12-29,129\n2016-12-30,75\n2016-12-31,23\n2017-01-01,59\n2017-01-02,113\n2017-01-03,135\n2017-01-04,186\n2017-01-05,137\n2017-01-06,141\n2017-01-07,80\n2017-01-08,85\n2017-01-09,111\n2017-01-10,183\n2017-01-11,188\n2017-01-12,157\n2017-01-13,159\n2017-01-14,58\n2017-01-15,124\n2017-01-16,116\n2017-01-17,156\n2017-01-18,146\n2017-01-19,98\n2017-01-20,103\n2017-01-21,83\n2017-01-22,133\n2017-01-23,166\n2017-01-24,133\n2017-01-25,183\n2017-01-26,135\n2017-01-27,153\n2017-01-28,103\n2017-01-29,65\n2017-01-30,154\n2017-01-31,137\n2017-02-01,125\n2017-02-02,143\n2017-02-03,102\n2017-02-04,73\n2017-02-05,100\n2017-02-06,131\n2017-02-07,134\n2017-02-08,128\n2017-02-09,122\n2017-02-10,114\n2017-02-11,65\n2017-02-12,116\n2017-02-13,110\n2017-02-14,151\n2017-02-15,109\n2017-02-16,84\n2017-02-17,96\n2017-02-18,86\n2017-02-19,72\n2017-02-20,109\n2017-02-21,153\n2017-02-22,166\n2017-02-23,109\n2017-02-24,117\n2017-02-25,58\n2017-02-26,77\n2017-02-27,102\n2017-02-28,125\n2017-03-01,135\n2017-03-02,144\n2017-03-03,139\n2017-03-04,85\n2017-03-05,71\n2017-03-06,149\n2017-03-07,132\n2017-03-08,149\n2017-03-09,152\n2017-03-10,194\n2017-03-11,118\n2017-03-12,63\n2017-03-13,102\n2017-03-14,204\n2017-03-15,166\n2017-03-16,131\n2017-03-17,147\n2017-03-18,83\n2017-03-19,96\n2017-03-20,111\n2017-03-21,201\n2017-03-22,140\n2017-03-23,163\n2017-03-24,157\n2017-03-25,102\n2017-03-26,79\n2017-03-27,146\n2017-03-28,157\n2017-03-29,130\n2017-03-30,121\n2017-03-31,105\n2017-04-01,35\n2017-04-02,109\n2017-04-03,149\n2017-04-04,129\n2017-04-05,201\n2017-04-06,169\n2017-04-07,195\n2017-04-08,63\n2017-04-09,66\n2017-04-10,152\n2017-04-11,168\n2017-04-12,235\n2017-04-13,189\n2017-04-14,148\n2017-04-15,93\n2017-04-16,69\n2017-04-17,128\n2017-04-18,196\n2017-04-19,198\n2017-04-20,133\n2017-04-21,131\n2017-04-22,81\n2017-04-23,66\n2017-04-24,148\n2017-04-25,171\n2017-04-26,176\n2017-04-27,169\n2017-04-28,181\n2017-04-29,84\n2017-04-30,81\n2017-05-01,161\n2017-05-02,182\n2017-05-03,193\n2017-05-04,170\n2017-05-05,122\n2017-05-06,80\n2017-05-07,127\n2017-05-08,180\n2017-05-09,152\n2017-05-10,225\n2017-05-11,111\n2017-05-12,173\n2017-05-13,76\n2017-05-14,52\n2017-05-15,134\n2017-05-16,140\n2017-05-17,161\n2017-05-18,175\n2017-05-19,199\n2017-05-20,102\n2017-05-21,145\n2017-05-22,170\n2017-05-23,158\n2017-05-24,141\n2017-05-25,132\n2017-05-26,181\n2017-05-27,83\n2017-05-28,78\n2017-05-29,122\n2017-05-30,188\n2017-05-31,152\n2017-06-01,189\n2017-06-02,201\n2017-06-03,91\n2017-06-04,113\n2017-06-05,164\n2017-06-06,125\n2017-06-07,175\n2017-06-08,184\n2017-06-09,154\n2017-06-10,86\n2017-06-11,37\n2017-06-12,179\n2017-06-13,196\n2017-06-14,169\n2017-06-15,178\n2017-06-16,149\n2017-06-17,89\n2017-06-18,85\n2017-06-19,142\n2017-06-20,157\n2017-06-21,143\n2017-06-22,152\n2017-06-23,185\n2017-06-24,77\n2017-06-25,77\n2017-06-26,140\n2017-06-27,239\n2017-06-28,181\n2017-06-29,154\n2017-06-30,162\n2017-07-01,99\n2017-07-02,108\n2017-07-03,129\n2017-07-04,162\n2017-07-05,180\n2017-07-06,174\n2017-07-07,153\n2017-07-08,67\n2017-07-09,76\n2017-07-10,184\n2017-07-11,200\n2017-07-12,204\n2017-07-13,181\n2017-07-14,157\n2017-07-15,115\n2017-07-16,152\n2017-07-17,117\n2017-07-18,168\n2017-07-19,162\n2017-07-20,150\n2017-07-21,142\n2017-07-22,86\n2017-07-23,94\n2017-07-24,113\n2017-07-25,152\n2017-07-26,97\n2017-07-27,139\n2017-07-28,108\n2017-07-29,55\n2017-07-30,60\n2017-07-31,132\n2017-08-01,175\n2017-08-02,107\n2017-08-03,118\n2017-08-04,123\n2017-08-05,88\n2017-08-06,36\n2017-08-07,131\n2017-08-08,119\n2017-08-09,116\n2017-08-10,71\n2017-08-11,116\n2017-08-12,40\n2017-08-13,94\n2017-08-14,149\n2017-08-15,146\n2017-08-16,167\n2017-08-17,160\n2017-08-18,120\n2017-08-19,31\n2017-08-20,43\n2017-08-21,84\n2017-08-22,114\n2017-08-23,144\n2017-08-24,114\n2017-08-25,118\n2017-08-26,48\n2017-08-27,85\n2017-08-28,94\n2017-08-29,119\n2017-08-30,135\n2017-08-31,127\n2017-09-01,116\n2017-09-02,84\n2017-09-03,70\n2017-09-04,113\n2017-09-05,123\n2017-09-06,99\n2017-09-07,95\n2017-09-08,125\n2017-09-09,59\n2017-09-10,56\n2017-09-11,148\n2017-09-12,127\n2017-09-13,123\n2017-09-14,120\n2017-09-15,129\n2017-09-16,40\n2017-09-17,59\n2017-09-18,115\n2017-09-19,137\n2017-09-20,98\n2017-09-21,96\n2017-09-22,82\n2017-09-23,88\n2017-09-24,86\n2017-09-25,144\n2017-09-26,169\n2017-09-27,150\n2017-09-28,151\n2017-09-29,136\n2017-09-30,95\n2017-10-01,94\n2017-10-02,138\n2017-10-03,165\n2017-10-04,114\n2017-10-05,95\n2017-10-06,113\n2017-10-07,80\n2017-10-08,95\n2017-10-09,122\n2017-10-10,175\n2017-10-11,127\n2017-10-12,179\n2017-10-13,197\n2017-10-14,84\n2017-10-15,60\n2017-10-16,128\n2017-10-17,138\n2017-10-18,154\n2017-10-19,163\n2017-10-20,118\n2017-10-21,82\n2017-10-22,49\n2017-10-23,96\n2017-10-24,142\n2017-10-25,160\n2017-10-26,140\n2017-10-27,126\n2017-10-28,81\n2017-10-29,99\n2017-10-30,107\n2017-10-31,157\n2017-11-01,157\n2017-11-02,151\n2017-11-03,136\n2017-11-04,76\n2017-11-05,63\n2017-11-06,155\n2017-11-07,124\n2017-11-08,153\n2017-11-09,119\n2017-11-10,90\n2017-11-11,57\n2017-11-12,86\n2017-11-13,129\n2017-11-14,142\n2017-11-15,137\n2017-11-16,147\n2017-11-17,175\n2017-11-18,101\n2017-11-19,95\n2017-11-20,140\n2017-11-21,127\n2017-11-22,152\n2017-11-23,98\n2017-11-24,90\n2017-11-25,59\n2017-11-26,86\n2017-11-27,95\n2017-11-28,114\n2017-11-29,140\n2017-11-30,131\n2017-12-01,129\n2017-12-02,98\n2017-12-03,85\n2017-12-04,150\n2017-12-05,174\n2017-12-06,168\n2017-12-07,96\n2017-12-08,71\n2017-12-09,96\n2017-12-10,106\n2017-12-11,133\n2017-12-12,144\n2017-12-13,96\n2017-12-14,102\n2017-12-15,133\n2017-12-16,69\n2017-12-17,80\n2017-12-18,83\n2017-12-19,107\n2017-12-20,137\n2017-12-21,132\n2017-12-22,108\n2017-12-23,34\n2017-12-24,37\n2017-12-25,35\n2017-12-26,65\n2017-12-27,129\n2017-12-28,126\n2017-12-29,86\n2017-12-30,99\n2017-12-31,60\n2018-01-01,44\n2018-01-02,79\n2018-01-03,93\n2018-01-04,131\n2018-01-05,125\n2018-01-06,63\n2018-01-07,58\n2018-01-08,180\n2018-01-09,139\n2018-01-10,155\n2018-01-11,128\n2018-01-12,92\n2018-01-13,116\n2018-01-14,92\n2018-01-15,116\n2018-01-16,106\n2018-01-17,145\n2018-01-18,152\n2018-01-19,110\n2018-01-20,100\n2018-01-21,67\n2018-01-22,154\n2018-01-23,122\n2018-01-24,215\n2018-01-25,153\n2018-01-26,127\n2018-01-27,81\n2018-01-28,79\n2018-01-29,146\n2018-01-30,159\n2018-01-31,128\n2018-02-01,97\n2018-02-02,124\n2018-02-03,51\n2018-02-04,82\n2018-02-05,163\n2018-02-06,183\n2018-02-07,107\n2018-02-08,151\n2018-02-09,174\n2018-02-10,147\n2018-02-11,131\n2018-02-12,135\n2018-02-13,170\n2018-02-14,151\n2018-02-15,180\n2018-02-16,128\n2018-02-17,125\n2018-02-18,119\n2018-02-19,128\n2018-02-20,148\n2018-02-21,115\n2018-02-22,136\n2018-02-23,173\n2018-02-24,103\n2018-02-25,115\n2018-02-26,168\n2018-02-27,174\n2018-02-28,180\n2018-03-01,126\n2018-03-02,168\n2018-03-03,112\n2018-03-04,76\n2018-03-05,164\n2018-03-06,123\n2018-03-07,143\n2018-03-08,138\n2018-03-09,161\n2018-03-10,111\n2018-03-11,47\n2018-03-12,175\n2018-03-13,146\n2018-03-14,175\n2018-03-15,141\n2018-03-16,135\n2018-03-17,116\n2018-03-18,113\n2018-03-19,199\n2018-03-20,199\n2018-03-21,207\n2018-03-22,147\n2018-03-23,145\n2018-03-24,161\n2018-03-25,131\n2018-03-26,170\n2018-03-27,160\n2018-03-28,192\n2018-03-29,142\n2018-03-30,152\n2018-03-31,151\n2018-04-01,79\n2018-04-02,150\n2018-04-03,229\n2018-04-04,209\n2018-04-05,157\n2018-04-06,221\n2018-04-07,119\n2018-04-08,162\n2018-04-09,133\n2018-04-10,196\n2018-04-11,172\n2018-04-12,206\n2018-04-13,203\n2018-04-14,89\n2018-04-15,98\n2018-04-16,176\n2018-04-17,166\n2018-04-18,155\n2018-04-19,140\n2018-04-20,166\n2018-04-21,69\n2018-04-22,91\n2018-04-23,156\n2018-04-24,199\n2018-04-25,144\n2018-04-26,148\n2018-04-27,147\n2018-04-28,145\n2018-04-29,161\n2018-04-30,176\n2018-05-01,197\n2018-05-02,175\n2018-05-03,162\n2018-05-04,159\n2018-05-05,68\n2018-05-06,100\n2018-05-07,184\n2018-05-08,227\n2018-05-09,129\n2018-05-10,150\n2018-05-11,195\n2018-05-12,115\n2018-05-13,77\n2018-05-14,178\n2018-05-15,194\n2018-05-16,178\n2018-05-17,198\n2018-05-18,150\n2018-05-19,70\n2018-05-20,68\n2018-05-21,140\n2018-05-22,175\n2018-05-23,186\n2018-05-24,152\n2018-05-25,116\n2018-05-26,38\n2018-05-27,136\n2018-05-28,169\n2018-05-29,149\n2018-05-30,170\n2018-05-31,229\n2018-06-01,228\n2018-06-02,50\n2018-06-03,58\n2018-06-04,167\n2018-06-05,150\n2018-06-06,200\n2018-06-07,232\n2018-06-08,186\n2018-06-09,113\n2018-06-10,88\n2018-06-11,160\n2018-06-12,216\n2018-06-13,180\n2018-06-14,176\n2018-06-15,178\n2018-06-16,93\n2018-06-17,51\n2018-06-18,112\n2018-06-19,117\n2018-06-20,135\n2018-06-21,126\n2018-06-22,94\n2018-06-23,73\n2018-06-24,81\n2018-06-25,137\n2018-06-26,208\n2018-06-27,127\n2018-06-28,152\n2018-06-29,134\n2018-06-30,98\n2018-07-01,128\n2018-07-02,144\n2018-07-03,127\n2018-07-04,70\n2018-07-05,117\n2018-07-06,115\n2018-07-07,55\n2018-07-08,57\n2018-07-09,104\n2018-07-10,133\n2018-07-11,238\n2018-07-12,135\n2018-07-13,244\n2018-07-14,74\n2018-07-15,86\n2018-07-16,178\n2018-07-17,195\n2018-07-18,201\n2018-07-19,149\n2018-07-20,122\n2018-07-21,89\n2018-07-22,81\n2018-07-23,169\n2018-07-24,170\n2018-07-25,134\n2018-07-26,136\n2018-07-27,135\n2018-07-28,79\n2018-07-29,79\n2018-07-30,144\n2018-07-31,120\n2018-08-01,150\n2018-08-02,108\n2018-08-03,147\n2018-08-04,65\n2018-08-05,66\n2018-08-06,139\n2018-08-07,142\n2018-08-08,117\n2018-08-09,79\n2018-08-10,84\n2018-08-11,103\n2018-08-12,79\n2018-08-13,99\n2018-08-14,126\n2018-08-15,106\n2018-08-16,143\n2018-08-17,114\n2018-08-18,68\n2018-08-19,34\n2018-08-20,83\n2018-08-21,75\n2018-08-22,108\n2018-08-23,84\n2018-08-24,85\n2018-08-25,66\n2018-08-26,50\n2018-08-27,77\n2018-08-28,114\n2018-08-29,116\n2018-08-30,112\n2018-08-31,114\n2018-09-01,99\n2018-09-02,92\n2018-09-03,87\n2018-09-04,122\n2018-09-05,125\n2018-09-06,104\n2018-09-07,139\n2018-09-08,83\n2018-09-09,57\n2018-09-10,110\n2018-09-11,119\n2018-09-12,119\n2018-09-13,141\n2018-09-14,147\n2018-09-15,67\n2018-09-16,96\n2018-09-17,104\n2018-09-18,140\n2018-09-19,159\n2018-09-20,125\n2018-09-21,141\n2018-09-22,61\n2018-09-23,84\n2018-09-24,139\n2018-09-25,107\n2018-09-26,159\n2018-09-27,140\n2018-09-28,144\n2018-09-29,92\n2018-09-30,95\n2018-10-01,157\n2018-10-02,173\n2018-10-03,186\n2018-10-04,155\n2018-10-05,137\n2018-10-06,65\n2018-10-07,35\n2018-10-08,135\n2018-10-09,172\n2018-10-10,130\n2018-10-11,119\n2018-10-12,140\n2018-10-13,98\n2018-10-14,66\n2018-10-15,134\n2018-10-16,183\n2018-10-17,152\n2018-10-18,150\n2018-10-19,146\n2018-10-20,125\n2018-10-21,146\n2018-10-22,134\n2018-10-23,152\n2018-10-24,143\n2018-10-25,116\n2018-10-26,133\n2018-10-27,136\n2018-10-28,119\n2018-10-29,149\n2018-10-30,250\n2018-10-31,174\n2018-11-01,251\n2018-11-02,168\n2018-11-03,86\n2018-11-04,287\n2018-11-05,157\n2018-11-06,161\n2018-11-07,157\n2018-11-08,158\n2018-11-09,205\n2018-11-10,168\n2018-11-11,235\n2018-11-12,95\n2018-11-13,179\n2018-11-14,205\n2018-11-15,171\n2018-11-16,171\n2018-11-17,86\n2018-11-18,67\n2018-11-19,160\n2018-11-20,210\n2018-11-21,177\n2018-11-22,80\n2018-11-23,121\n2018-11-24,92\n2018-11-25,117\n2018-11-26,131\n2018-11-27,143\n2018-11-28,169\n2018-11-29,177\n2018-11-30,188\n2018-12-01,153\n2018-12-02,128\n2018-12-03,225\n2018-12-04,137\n2018-12-05,219\n2018-12-06,163\n2018-12-07,169\n2018-12-08,59\n2018-12-09,67\n2018-12-10,142\n2018-12-11,130\n2018-12-12,214\n2018-12-13,115\n2018-12-14,110\n2018-12-15,95\n2018-12-16,91\n2018-12-17,166\n2018-12-18,164\n2018-12-19,129\n2018-12-20,132\n2018-12-21,82\n2018-12-22,75\n2018-12-23,49\n2018-12-24,47\n2018-12-25,25\n2018-12-26,50\n2018-12-27,103\n2018-12-28,123\n2018-12-29,60\n2018-12-30,92\n2018-12-31,91\n2019-01-01,73\n2019-01-02,133\n2019-01-03,107\n2019-01-04,141\n2019-01-05,74\n2019-01-06,93\n2019-01-07,113\n2019-01-08,137\n2019-01-09,178\n2019-01-10,185\n2019-01-11,133\n2019-01-12,102\n2019-01-13,83\n2019-01-14,155\n2019-01-15,149\n2019-01-16,139\n2019-01-17,175\n2019-01-18,165\n2019-01-19,158\n2019-01-20,101\n2019-01-21,181\n2019-01-22,186\n2019-01-23,164\n2019-01-24,173\n2019-01-25,112\n2019-01-26,61\n2019-01-27,83\n2019-01-28,103\n2019-01-29,106\n2019-01-30,117\n2019-01-31,93\n2019-02-01,156\n2019-02-02,101\n2019-02-03,101\n2019-02-04,138\n2019-02-05,129\n2019-02-06,158\n2019-02-07,123\n2019-02-08,175\n2019-02-09,115\n2019-02-10,122\n2019-02-11,150\n2019-02-12,151\n2019-02-13,157\n2019-02-14,150\n2019-02-15,164\n2019-02-16,178\n2019-02-17,100\n2019-02-18,161\n2019-02-19,226\n2019-02-20,197\n2019-02-21,217\n2019-02-22,154\n2019-02-23,91\n2019-02-24,102\n2019-02-25,241\n2019-02-26,164\n2019-02-27,191\n2019-02-28,164\n2019-03-01,117\n2019-03-02,43\n2019-03-03,150\n2019-03-04,129\n2019-03-05,128\n2019-03-06,113\n2019-03-07,109\n2019-03-08,101\n2019-03-09,109\n2019-03-10,85\n2019-03-11,133\n2019-03-12,110\n2019-03-13,104\n2019-03-14,104\n2019-03-15,116\n2019-03-16,66\n2019-03-17,66\n2019-03-18,105\n2019-03-19,119\n2019-03-20,137\n2019-03-21,172\n2019-03-22,118\n2019-03-23,73\n2019-03-24,50\n2019-03-25,141\n2019-03-26,159\n2019-03-27,107\n2019-03-28,100\n2019-03-29,134\n2019-03-30,92\n2019-03-31,70\n2019-04-01,139\n2019-04-02,114\n2019-04-03,113\n2019-04-04,97\n2019-04-05,125\n2019-04-06,37\n2019-04-07,51\n2019-04-08,154\n2019-04-09,118\n2019-04-10,169\n2019-04-11,109\n2019-04-12,107\n2019-04-13,42\n2019-04-14,58\n2019-04-15,108\n2019-04-16,135\n2019-04-17,73\n2019-04-18,114\n2019-04-19,112\n2019-04-20,92\n2019-04-21,70\n2019-04-22,104\n2019-04-23,84\n2019-04-24,100\n2019-04-25,115\n2019-04-26,92\n2019-04-27,44\n2019-04-28,106\n2019-04-29,126\n2019-04-30,78\n2019-05-01,116\n2019-05-02,108\n2019-05-03,78\n2019-05-04,50\n2019-05-05,85\n2019-05-06,115\n2019-05-07,120\n2019-05-08,124\n2019-05-09,108\n2019-05-10,80\n2019-05-11,86\n2019-05-12,69\n2019-05-13,102\n2019-05-14,100\n2019-05-15,80\n2019-05-16,99\n2019-05-17,89\n2019-05-18,51\n2019-05-19,64\n2019-05-20,111\n2019-05-21,97\n2019-05-22,103\n2019-05-23,141\n2019-05-24,68\n2019-05-25,75\n2019-05-26,75\n2019-05-27,58\n2019-05-28,79\n2019-05-29,87\n2019-05-30,125\n2019-05-31,120\n2019-06-01,69\n2019-06-02,74\n2019-06-03,112\n2019-06-04,88\n2019-06-05,138\n2019-06-06,95\n2019-06-07,100\n2019-06-08,95\n2019-06-09,88\n2019-06-10,97\n2019-06-11,105\n2019-06-12,104\n2019-06-13,88\n2019-06-14,89\n2019-06-15,47\n2019-06-16,41\n2019-06-17,78\n2019-06-18,68\n2019-06-19,115\n2019-06-20,86\n2019-06-21,105\n2019-06-22,43\n2019-06-23,79\n2019-06-24,92\n2019-06-25,77\n2019-06-26,97\n2019-06-27,96\n2019-06-28,92\n2019-06-29,92\n2019-06-30,57\n2019-07-01,94\n2019-07-02,164\n2019-07-03,101\n2019-07-04,89\n2019-07-05,67\n2019-07-06,59\n2019-07-07,37\n2019-07-08,146\n2019-07-09,89\n2019-07-10,136\n2019-07-11,109\n2019-07-12,102\n2019-07-13,78\n2019-07-14,65\n2019-07-15,122\n2019-07-16,142\n2019-07-17,113\n2019-07-18,103\n2019-07-19,102\n2019-07-20,92\n2019-07-21,41\n2019-07-22,97\n2019-07-23,82\n2019-07-24,124\n2019-07-25,129\n2019-07-26,115\n2019-07-27,70\n2019-07-28,67\n2019-07-29,98\n2019-07-30,106\n2019-07-31,76\n2019-08-01,140\n2019-08-02,77\n2019-08-03,86\n2019-08-04,52\n2019-08-05,159\n2019-08-06,132\n2019-08-07,142\n2019-08-08,154\n2019-08-09,146\n2019-08-10,87\n2019-08-11,124\n2019-08-12,175\n2019-08-13,204\n2019-08-14,204\n2019-08-15,150\n2019-08-16,149\n2019-08-17,114\n2019-08-18,116\n2019-08-19,131\n2019-08-20,135\n2019-08-21,162\n2019-08-22,159\n2019-08-23,193\n2019-08-24,115\n2019-08-25,59\n2019-08-26,106\n2019-08-27,123\n2019-08-28,172\n2019-08-29,120\n2019-08-30,86\n2019-08-31,75\n2019-09-01,87\n2019-09-02,126\n2019-09-03,125\n2019-09-04,144\n2019-09-05,145\n2019-09-06,109\n2019-09-07,105\n2019-09-08,111\n2019-09-09,131\n2019-09-10,144\n2019-09-11,148\n2019-09-12,104\n2019-09-13,161\n2019-09-14,83\n2019-09-15,75\n2019-09-16,109\n2019-09-17,131\n2019-09-18,112\n2019-09-19,115\n2019-09-20,167\n2019-09-21,156\n2019-09-22,94\n2019-09-23,107\n2019-09-24,147\n2019-09-25,138\n2019-09-26,144\n2019-09-27,119\n2019-09-28,29\n2019-09-29,50\n2019-09-30,132\n2019-10-01,142\n2019-10-02,138\n2019-10-03,129\n2019-10-04,110\n2019-10-05,74\n2019-10-06,70\n2019-10-07,118\n2019-10-08,114\n2019-10-09,123\n2019-10-10,137\n2019-10-11,151\n2019-10-12,70\n2019-10-13,103\n2019-10-14,110\n2019-10-15,170\n2019-10-16,130\n2019-10-17,122\n2019-10-18,122\n2019-10-19,85\n2019-10-20,51\n2019-10-21,126\n2019-10-22,94\n2019-10-23,99\n2019-10-24,154\n2019-10-25,119\n2019-10-26,105\n2019-10-27,92\n2019-10-28,130\n2019-10-29,146\n2019-10-30,164\n2019-10-31,184\n2019-11-01,187\n2019-11-02,102\n2019-11-03,110\n2019-11-04,121\n2019-11-05,161\n2019-11-06,123\n2019-11-07,119\n2019-11-08,131\n2019-11-09,71\n2019-11-10,85\n2019-11-11,113\n2019-11-12,94\n2019-11-13,124\n2019-11-14,100\n2019-11-15,135\n2019-11-16,97\n2019-11-17,110\n2019-11-18,111\n2019-11-19,123\n2019-11-20,127\n2019-11-21,115\n2019-11-22,92\n2019-11-23,61\n2019-11-24,90\n2019-11-25,114\n2019-11-26,106\n2019-11-27,144\n2019-11-28,114\n2019-11-29,115\n2019-11-30,116\n2019-12-01,103\n2019-12-02,141\n2019-12-03,116\n2019-12-04,122\n2019-12-05,157\n2019-12-06,158\n2019-12-07,95\n2019-12-08,110\n2019-12-09,113\n2019-12-10,149\n2019-12-11,142\n2019-12-12,187\n2019-12-13,145\n2019-12-14,113\n2019-12-15,89\n2019-12-16,125\n2019-12-17,201\n2019-12-18,176\n2019-12-19,206\n2019-12-20,180\n2019-12-21,90\n2019-12-22,79\n2019-12-23,49\n2019-12-24,110\n2019-12-25,76\n2019-12-26,78\n2019-12-27,124\n2019-12-28,117\n2019-12-29,60\n2019-12-30,156\n2019-12-31,126\n2020-01-01,99\n2020-01-02,142\n2020-01-03,95\n2020-01-04,120\n2020-01-05,81\n2020-01-06,125\n2020-01-07,161\n2020-01-08,169\n2020-01-09,192\n2020-01-10,251\n2020-01-11,164\n2020-01-12,219\n2020-01-13,224\n2020-01-14,193\n2020-01-15,248\n2020-01-16,255\n2020-01-17,247\n2020-01-18,117\n2020-01-19,187\n2020-01-20,138\n2020-01-21,139\n2020-01-22,199\n2020-01-23,185\n2020-01-24,143\n2020-01-25,64\n2020-01-26,89\n2020-01-27,135\n2020-01-28,101\n2020-01-29,114\n2020-01-30,155\n2020-01-31,165\n2020-02-01,89\n2020-02-02,111\n2020-02-03,141\n2020-02-04,120\n2020-02-05,145\n2020-02-06,133\n2020-02-07,149\n2020-02-08,137\n2020-02-09,119\n2020-02-10,152\n2020-02-11,162\n2020-02-12,134\n2020-02-13,170\n2020-02-14,149\n2020-02-15,95\n2020-02-16,66\n2020-02-17,129\n2020-02-18,175\n2020-02-19,133\n2020-02-20,128\n2020-02-21,133\n2020-02-22,99\n2020-02-23,165\n2020-02-24,153\n2020-02-25,146\n2020-02-26,155\n2020-02-27,144\n2020-02-28,115\n2020-02-29,54\n2020-03-01,56\n2020-03-02,166\n2020-03-03,142\n2020-03-04,120\n2020-03-05,148\n2020-03-06,119\n2020-03-07,92\n2020-03-08,143\n2020-03-09,112\n2020-03-10,111\n2020-03-11,110\n2020-03-12,82\n2020-03-13,76\n2020-03-14,74\n2020-03-15,80\n2020-03-16,91\n2020-03-17,106\n2020-03-18,132\n2020-03-19,92\n2020-03-20,115\n2020-03-21,89\n2020-03-22,105\n2020-03-23,138\n2020-03-24,144\n2020-03-25,116\n2020-03-26,99\n2020-03-27,112\n2020-03-28,123\n2020-03-29,76\n2020-03-30,107\n2020-03-31,177\n2020-04-01,158\n2020-04-02,160\n2020-04-03,145\n2020-04-04,130\n2020-04-05,95\n2020-04-06,152\n2020-04-07,157\n2020-04-08,217\n2020-04-09,160\n2020-04-10,151\n2020-04-11,127\n2020-04-12,106\n2020-04-13,208\n2020-04-14,175\n2020-04-15,146\n2020-04-16,149\n2020-04-17,110\n2020-04-18,148\n2020-04-19,151\n2020-04-20,155\n2020-04-21,141\n2020-04-22,147\n2020-04-23,146\n2020-04-24,136\n2020-04-25,98\n2020-04-26,139\n2020-04-27,180\n2020-04-28,118\n2020-04-29,166\n2020-04-30,220\n2020-05-01,237\n2020-05-02,133\n2020-05-03,142\n2020-05-04,132\n2020-05-05,218\n2020-05-06,229\n2020-05-07,184\n2020-05-08,160\n2020-05-09,152\n2020-05-10,160\n2020-05-11,187\n2020-05-12,158\n2020-05-13,200\n2020-05-14,212\n2020-05-15,199\n2020-05-16,113\n2020-05-17,142\n2020-05-18,161\n2020-05-19,162\n2020-05-20,186\n2020-05-21,122\n2020-05-22,142\n2020-05-23,128\n2020-05-24,134\n2020-05-25,143\n2020-05-26,220\n2020-05-27,144\n2020-05-28,100\n2020-05-29,108\n2020-05-30,213\n2020-05-31,86\n2020-06-01,115\n2020-06-02,107\n2020-06-03,169\n2020-06-04,147\n2020-06-05,115\n2020-06-06,107\n2020-06-07,93\n2020-06-08,116\n2020-06-09,143\n2020-06-10,84\n2020-06-11,177\n2020-06-12,125\n2020-06-13,61\n2020-06-14,98\n2020-06-15,191\n2020-06-16,157\n2020-06-17,199\n2020-06-18,192\n2020-06-19,146\n2020-06-20,60\n2020-06-21,116\n2020-06-22,128\n2020-06-23,130\n2020-06-24,160\n2020-06-25,120\n2020-06-26,167\n2020-06-27,164\n2020-06-28,113\n2020-06-29,167\n2020-06-30,155\n2020-07-01,126\n2020-07-02,134\n2020-07-03,146\n2020-07-04,59\n2020-07-05,102\n2020-07-06,120\n2020-07-07,124\n2020-07-08,154\n2020-07-09,164\n2020-07-10,171\n2020-07-11,79\n2020-07-12,63\n2020-07-13,85\n2020-07-14,90\n2020-07-15,129\n2020-07-16,138\n2020-07-17,95\n2020-07-18,71\n2020-07-19,99\n2020-07-20,140\n2020-07-21,189\n2020-07-22,212\n2020-07-23,176\n2020-07-24,183\n2020-07-25,55\n2020-07-26,100\n2020-07-27,137\n2020-07-28,190\n2020-07-29,118\n2020-07-30,147\n2020-07-31,139\n2020-08-01,91\n2020-08-02,99\n2020-08-03,168\n2020-08-04,168\n2020-08-05,151\n2020-08-06,167\n2020-08-07,155\n2020-08-08,78\n2020-08-09,111\n2020-08-10,126\n2020-08-11,111\n2020-08-12,121\n2020-08-13,97\n2020-08-14,137\n2020-08-15,54\n2020-08-16,56\n2020-08-17,162\n2020-08-18,98\n2020-08-19,139\n2020-08-20,128\n2020-08-21,133\n2020-08-22,93\n2020-08-23,68\n2020-08-24,149\n2020-08-25,165\n2020-08-26,201\n2020-08-27,135\n2020-08-28,108\n2020-08-29,84\n2020-08-30,42\n2020-08-31,100\n2020-09-01,134\n2020-09-02,120\n2020-09-03,100\n2020-09-04,110\n2020-09-05,53\n2020-09-06,62\n2020-09-07,129\n2020-09-08,153\n2020-09-09,148\n2020-09-10,141\n2020-09-11,162\n2020-09-12,78\n2020-09-13,58\n2020-09-14,133\n2020-09-15,126\n2020-09-16,134\n2020-09-17,94\n2020-09-18,102\n2020-09-19,65\n2020-09-20,86\n2020-09-21,136\n2020-09-22,144\n2020-09-23,137\n2020-09-24,146\n2020-09-25,148\n2020-09-26,78\n2020-09-27,82\n2020-09-28,158\n2020-09-29,122\n2020-09-30,108\n2020-10-01,146\n2020-10-02,125\n2020-10-03,55\n2020-10-04,62\n2020-10-05,99\n2020-10-06,95\n2020-10-07,137\n2020-10-08,125\n2020-10-09,117\n2020-10-10,67\n2020-10-11,62\n2020-10-12,133\n2020-10-13,80\n2020-10-14,115\n2020-10-15,104\n2020-10-16,122\n2020-10-17,106\n2020-10-18,86\n2020-10-19,156\n2020-10-20,156\n2020-10-21,162\n2020-10-22,180\n2020-10-23,93\n2020-10-24,81\n2020-10-25,75\n2020-10-26,125\n2020-10-27,130\n2020-10-28,147\n2020-10-29,118\n2020-10-30,99\n2020-10-31,93\n2020-11-01,103\n2020-11-02,103\n2020-11-03,121\n2020-11-04,86\n2020-11-05,153\n2020-11-06,129\n2020-11-07,116\n2020-11-08,51\n2020-11-09,129\n2020-11-10,150\n2020-11-11,112\n2020-11-12,117\n2020-11-13,161\n2020-11-14,69\n2020-11-15,95\n2020-11-16,144\n2020-11-17,155\n2020-11-18,142\n2020-11-19,119\n2020-11-20,113\n2020-11-21,77\n2020-11-22,69\n2020-11-23,97\n2020-11-24,107\n2020-11-25,122\n2020-11-26,131\n2020-11-27,120\n2020-11-28,93\n2020-11-29,116\n2020-11-30,139\n2020-12-01,89\n2020-12-02,152\n2020-12-03,130\n2020-12-04,128\n2020-12-05,112\n2020-12-06,69\n2020-12-07,153\n2020-12-08,110\n2020-12-09,122\n2020-12-10,124\n2020-12-11,107\n2020-12-12,78\n2020-12-13,90\n2020-12-14,87\n2020-12-15,125\n2020-12-16,143\n2020-12-17,114\n2020-12-18,96\n2020-12-19,55\n2020-12-20,66\n2020-12-21,150\n2020-12-22,129\n2020-12-23,134\n2020-12-24,70\n2020-12-25,38\n2020-12-26,68\n2020-12-27,93\n2020-12-28,116\n2020-12-29,151\n2020-12-30,114\n2020-12-31,53\n2021-01-01,67\n2021-01-02,70\n2021-01-03,60\n2021-01-04,113\n2021-01-05,108\n2021-01-06,106\n2021-01-07,133\n2021-01-08,133\n2021-01-09,99\n2021-01-10,114\n2021-01-11,115\n2021-01-12,119\n2021-01-13,167\n2021-01-14,178\n2021-01-15,184\n2021-01-16,116\n2021-01-17,130\n2021-01-18,115\n2021-01-19,155\n2021-01-20,204\n2021-01-21,137\n2021-01-22,141\n2021-01-23,130\n2021-01-24,67\n2021-01-25,132\n2021-01-26,180\n2021-01-27,161\n2021-01-28,162\n2021-01-29,207\n2021-01-30,95\n2021-01-31,149\n2021-02-01,176\n2021-02-02,157\n2021-02-03,191\n2021-02-04,156\n2021-02-05,214\n2021-02-06,159\n2021-02-07,94\n2021-02-08,121\n2021-02-09,134\n2021-02-10,126\n2021-02-11,138\n2021-02-12,115\n2021-02-13,121\n2021-02-14,115\n2021-02-15,148\n2021-02-16,127\n2021-02-17,149\n2021-02-18,153\n2021-02-19,130\n2021-02-20,120\n2021-02-21,98\n2021-02-22,160\n2021-02-23,150\n2021-02-24,185\n2021-02-25,171\n2021-02-26,162\n2021-02-27,130\n2021-02-28,189\n2021-03-01,218\n2021-03-02,138\n2021-03-03,172\n2021-03-04,202\n2021-03-05,152\n2021-03-06,128\n2021-03-07,115\n2021-03-08,195\n2021-03-09,161\n2021-03-10,190\n2021-03-11,172\n2021-03-12,149\n2021-03-13,141\n2021-03-14,77\n2021-03-15,167\n2021-03-16,161\n2021-03-17,115\n2021-03-18,144\n2021-03-19,145\n2021-03-20,88\n2021-03-21,150\n2021-03-22,110\n2021-03-23,157\n2021-03-24,145\n2021-03-25,116\n2021-03-26,126\n2021-03-27,153\n2021-03-28,111\n2021-03-29,178\n2021-03-30,178\n2021-03-31,170\n2021-04-01,170\n2021-04-02,152\n2021-04-03,100\n2021-04-04,119\n2021-04-05,129\n2021-04-06,200\n2021-04-07,164\n2021-04-08,120\n2021-04-09,178\n2021-04-10,136\n2021-04-11,106\n2021-04-12,155\n2021-04-13,114\n2021-04-14,138\n2021-04-15,175\n2021-04-16,130\n2021-04-17,100\n2021-04-18,80\n2021-04-19,137\n2021-04-20,107\n2021-04-21,174\n2021-04-22,136\n2021-04-23,141\n2021-04-24,94\n2021-04-25,124\n2021-04-26,190\n2021-04-27,96\n2021-04-28,131\n2021-04-29,135\n2021-04-30,166\n2021-05-01,80\n2021-05-02,94\n2021-05-03,71\n2021-05-04,96\n2021-05-05,117\n2021-05-06,162\n2021-05-07,133\n2021-05-08,131\n2021-05-09,94\n2021-05-10,151\n2021-05-11,159\n2021-05-12,156\n2021-05-13,172\n2021-05-14,181\n2021-05-15,75\n2021-05-16,94\n2021-05-17,144\n2021-05-18,117\n2021-05-19,157\n2021-05-20,195\n2021-05-21,160\n2021-05-22,104\n2021-05-23,75\n2021-05-24,141\n2021-05-25,177\n2021-05-26,186\n2021-05-27,141\n2021-05-28,146\n2021-05-29,66\n2021-05-30,83\n2021-05-31,120\n2021-06-01,151\n2021-06-02,132\n2021-06-03,171\n2021-06-04,98\n2021-06-05,63\n2021-06-06,64\n2021-06-07,133\n2021-06-08,117\n2021-06-09,181\n2021-06-10,166\n2021-06-11,100\n2021-06-12,56\n2021-06-13,54\n2021-06-14,89\n2021-06-15,95\n2021-06-16,134\n2021-06-17,149\n2021-06-18,116\n2021-06-19,48\n2021-06-20,60\n2021-06-21,115\n2021-06-22,122\n2021-06-23,136\n2021-06-24,116\n2021-06-25,110\n2021-06-26,68\n2021-06-27,104\n2021-06-28,118\n2021-06-29,110\n2021-06-30,129\n2021-07-01,154\n2021-07-02,91\n2021-07-03,49\n2021-07-04,81\n2021-07-05,139\n2021-07-06,141\n2021-07-07,113\n2021-07-08,96\n2021-07-09,118\n2021-07-10,111\n2021-07-11,104\n2021-07-12,131\n2021-07-13,126\n2021-07-14,174\n2021-07-15,140\n2021-07-16,102\n2021-07-17,68\n2021-07-18,55\n2021-07-19,87\n2021-07-20,145\n2021-07-21,125\n2021-07-22,118\n2021-07-23,107\n2021-07-24,37\n2021-07-25,45\n2021-07-26,80\n2021-07-27,152\n2021-07-28,145\n2021-07-29,148\n2021-07-30,118\n2021-07-31,98\n2021-08-01,81\n2021-08-02,137\n2021-08-03,129\n2021-08-04,129\n2021-08-05,103\n2021-08-06,143\n2021-08-07,112\n2021-08-08,94\n2021-08-09,102\n2021-08-10,106\n2021-08-11,140\n2021-08-12,113\n2021-08-13,113\n2021-08-14,109\n2021-08-15,71\n2021-08-16,124\n2021-08-17,95\n2021-08-18,148\n2021-08-19,124\n2021-08-20,93\n2021-08-21,94\n2021-08-22,80\n2021-08-23,156\n2021-08-24,135\n2021-08-25,170\n2021-08-26,125\n2021-08-27,122\n2021-08-28,89\n2021-08-29,80\n2021-08-30,125\n2021-08-31,160\n2021-09-01,158\n2021-09-02,106\n2021-09-03,125\n2021-09-04,96\n2021-09-05,89\n2021-09-06,128\n2021-09-07,124\n2021-09-08,99\n2021-09-09,145\n2021-09-10,132\n2021-09-11,80\n2021-09-12,82\n2021-09-13,148\n2021-09-14,129\n2021-09-15,146\n2021-09-16,183\n2021-09-17,157\n2021-09-18,103\n2021-09-19,133\n2021-09-20,133\n2021-09-21,146\n2021-09-22,146\n2021-09-23,145\n2021-09-24,84\n2021-09-25,108\n2021-09-26,64\n2021-09-27,122\n2021-09-28,170\n2021-09-29,143\n2021-09-30,142\n2021-10-01,104\n2021-10-02,73\n2021-10-03,123\n2021-10-04,96\n2021-10-05,141\n2021-10-06,190\n2021-10-07,152\n2021-10-08,118\n2021-10-09,158\n2021-10-10,118\n2021-10-11,137\n2021-10-12,139\n2021-10-13,110\n2021-10-14,100\n2021-10-15,114\n2021-10-16,87\n2021-10-17,51\n2021-10-18,152\n2021-10-19,135\n2021-10-20,126\n2021-10-21,156\n2021-10-22,137\n2021-10-23,68\n2021-10-24,63\n2021-10-25,126\n2021-10-26,92\n2021-10-27,130\n2021-10-28,112\n2021-10-29,135\n2021-10-30,70\n2021-10-31,86\n2021-11-01,116\n2021-11-02,129\n2021-11-03,163\n2021-11-04,188\n2021-11-05,145\n2021-11-06,116\n2021-11-07,57\n2021-11-08,143\n2021-11-09,128\n2021-11-10,178\n2021-11-11,140\n2021-11-12,98\n2021-11-13,54\n2021-11-14,63\n2021-11-15,136\n2021-11-16,146\n2021-11-17,175\n2021-11-18,133\n2021-11-19,148\n2021-11-20,38\n2021-11-21,72\n2021-11-22,114\n2021-11-23,135\n2021-11-24,120\n2021-11-25,81\n2021-11-26,96\n2021-11-27,67\n2021-11-28,72\n2021-11-29,128\n2021-11-30,101\n2021-12-01,124\n2021-12-02,100\n2021-12-03,79\n2021-12-04,49\n2021-12-05,64\n2021-12-06,122\n2021-12-07,113\n2021-12-08,104\n2021-12-09,125\n2021-12-10,122\n2021-12-11,38\n2021-12-12,94\n2021-12-13,144\n2021-12-14,107\n2021-12-15,143\n2021-12-16,145\n2021-12-17,102\n2021-12-18,62\n2021-12-19,88\n2021-12-20,120\n2021-12-21,126\n2021-12-22,163\n2021-12-23,107\n2021-12-24,48\n2021-12-25,16\n2021-12-26,29\n2021-12-27,111\n2021-12-28,95\n2021-12-29,133\n2021-12-30,88\n2021-12-31,44\n2022-01-01,33\n2022-01-02,44\n2022-01-03,85\n2022-01-04,147\n2022-01-05,120\n2022-01-06,110\n2022-01-07,127\n2022-01-08,95\n2022-01-09,85\n2022-01-10,150\n2022-01-11,118\n2022-01-12,180\n2022-01-13,153\n2022-01-14,157\n2022-01-15,74\n2022-01-16,86\n2022-01-17,151\n2022-01-18,150\n2022-01-19,182\n2022-01-20,156\n2022-01-21,156\n2022-01-22,92\n2022-01-23,115\n2022-01-24,141\n2022-01-25,125\n2022-01-26,158\n2022-01-27,109\n2022-01-28,130\n2022-01-29,103\n2022-01-30,67\n2022-01-31,152\n2022-02-01,153\n2022-02-02,160\n2022-02-03,104\n2022-02-04,124\n2022-02-05,38\n2022-02-06,70\n2022-02-07,129\n2022-02-08,133\n2022-02-09,152\n2022-02-10,144\n2022-02-11,110\n2022-02-12,48\n2022-02-13,43\n2022-02-14,108\n2022-02-15,138\n2022-02-16,140\n2022-02-17,171\n2022-02-18,108\n2022-02-19,63\n2022-02-20,82\n2022-02-21,165\n2022-02-22,120\n2022-02-23,163\n2022-02-24,168\n2022-02-25,122\n2022-02-26,37\n2022-02-27,42\n2022-02-28,97\n2022-03-01,90\n2022-03-02,122\n2022-03-03,103\n2022-03-04,112\n2022-03-05,92\n2022-03-06,62\n2022-03-07,118\n2022-03-08,96\n2022-03-09,110\n2022-03-10,131\n2022-03-11,101\n2022-03-12,54\n2022-03-13,53\n2022-03-14,133\n2022-03-15,121\n2022-03-16,129\n2022-03-17,123\n2022-03-18,121\n2022-03-19,62\n2022-03-20,49\n2022-03-21,94\n2022-03-22,126\n2022-03-23,152\n2022-03-24,152\n2022-03-25,145\n2022-03-26,133\n2022-03-27,103\n2022-03-28,170\n2022-03-29,125\n2022-03-30,146\n2022-03-31,152\n2022-04-01,123\n2022-04-02,92\n2022-04-03,77\n2022-04-04,174\n2022-04-05,182\n2022-04-06,118\n2022-04-07,114\n2022-04-08,156\n2022-04-09,105\n2022-04-10,93\n2022-04-11,135\n2022-04-12,153\n2022-04-13,118\n2022-04-14,149\n2022-04-15,169\n2022-04-16,70\n2022-04-17,55\n2022-04-18,125\n2022-04-19,126\n2022-04-20,108\n2022-04-21,147\n2022-04-22,173\n2022-04-23,90\n2022-04-24,100\n2022-04-25,101\n2022-04-26,150\n2022-04-27,158\n2022-04-28,134\n2022-04-29,107\n2022-04-30,50\n2022-05-01,62\n2022-05-02,106\n2022-05-03,104\n2022-05-04,133\n2022-05-05,167\n2022-05-06,140\n2022-05-07,44\n2022-05-08,71\n2022-05-09,147\n2022-05-10,119\n2022-05-11,151\n2022-05-12,132\n2022-05-13,112\n2022-05-14,96\n2022-05-15,47\n2022-05-16,149\n2022-05-17,141\n2022-05-18,117\n2022-05-19,158\n2022-05-20,112\n2022-05-21,85\n2022-05-22,83\n2022-05-23,154\n2022-05-24,149\n2022-05-25,113\n2022-05-26,113\n2022-05-27,102\n2022-05-28,99\n2022-05-29,56\n2022-05-30,107\n2022-05-31,106\n2022-06-01,130\n2022-06-02,173\n2022-06-03,134\n2022-06-04,70\n2022-06-05,88\n2022-06-06,110\n2022-06-07,113\n2022-06-08,139\n2022-06-09,142\n2022-06-10,109\n2022-06-11,56\n2022-06-12,76\n2022-06-13,95\n2022-06-14,94\n2022-06-15,127\n2022-06-16,200\n2022-06-17,202\n2022-06-18,49\n2022-06-19,62\n2022-06-20,116\n2022-06-21,112\n2022-06-22,131\n2022-06-23,114\n2022-06-24,110\n2022-06-25,52\n2022-06-26,56\n2022-06-27,76\n2022-06-28,72\n2022-06-29,120\n2022-06-30,118\n2022-07-01,87\n2022-07-02,32\n2022-07-03,39\n2022-07-04,89\n2022-07-05,120\n2022-07-06,106\n2022-07-07,138\n2022-07-08,109\n2022-07-09,42\n2022-07-10,44\n2022-07-11,92\n2022-07-12,135\n2022-07-13,157\n2022-07-14,124\n2022-07-15,177\n2022-07-16,56\n2022-07-17,41\n2022-07-18,143\n2022-07-19,119\n2022-07-20,211\n2022-07-21,128\n2022-07-22,162\n2022-07-23,44\n2022-07-24,99\n2022-07-25,152\n2022-07-26,158\n2022-07-27,133\n2022-07-28,214\n2022-07-29,133\n2022-07-30,79\n2022-07-31,70\n2022-08-01,130\n2022-08-02,123\n2022-08-03,105\n2022-08-04,125\n2022-08-05,127\n2022-08-06,71\n2022-08-07,55\n2022-08-08,108\n2022-08-09,99\n2022-08-10,145\n2022-08-11,115\n2022-08-12,110\n2022-08-13,100\n2022-08-14,95\n2022-08-15,133\n2022-08-16,130\n2022-08-17,131\n2022-08-18,116\n2022-08-19,143\n2022-08-20,82\n2022-08-21,78\n2022-08-22,110\n2022-08-23,130\n2022-08-24,181\n2022-08-25,132\n2022-08-26,105\n2022-08-27,79\n2022-08-28,66\n2022-08-29,173\n2022-08-30,174\n2022-08-31,174\n2022-09-01,137\n2022-09-02,135\n2022-09-03,63\n2022-09-04,81\n2022-09-05,168\n2022-09-06,207\n2022-09-07,194\n2022-09-08,162\n2022-09-09,135\n2022-09-10,54\n2022-09-11,53\n2022-09-12,123\n2022-09-13,120\n2022-09-14,137\n2022-09-15,158\n2022-09-16,110\n2022-09-17,77\n2022-09-18,86\n2022-09-19,165\n2022-09-20,177\n2022-09-21,156\n2022-09-22,113\n2022-09-23,137\n2022-09-24,94\n2022-09-25,44\n2022-09-26,122\n2022-09-27,121\n2022-09-28,159\n2022-09-29,139\n2022-09-30,122\n2022-10-01,81\n2022-10-02,57\n2022-10-03,155\n2022-10-04,127\n2022-10-05,114\n2022-10-06,139\n2022-10-07,88\n2022-10-08,61\n2022-10-09,66\n2022-10-10,130\n2022-10-11,138\n2022-10-12,126\n2022-10-13,128\n2022-10-14,147\n2022-10-15,70\n2022-10-16,65\n2022-10-17,125\n2022-10-18,192\n2022-10-19,107\n2022-10-20,121\n2022-10-21,90\n2022-10-22,66\n2022-10-23,60\n2022-10-24,99\n2022-10-25,84\n2022-10-26,79\n2022-10-27,95\n2022-10-28,79\n2022-10-29,43\n2022-10-30,34\n2022-10-31,133\n2022-11-01,101\n2022-11-02,124\n2022-11-03,105\n2022-11-04,86\n2022-11-05,62\n2022-11-06,67\n2022-11-07,101\n2022-11-08,117\n2022-11-09,109\n2022-11-10,152\n2022-11-11,89\n2022-11-12,43\n2022-11-13,67\n2022-11-14,122\n2022-11-15,115\n2022-11-16,171\n2022-11-17,149\n2022-11-18,123\n2022-11-19,77\n2022-11-20,100\n2022-11-21,133\n2022-11-22,142\n2022-11-23,154\n2022-11-24,114\n2022-11-25,116\n2022-11-26,32\n2022-11-27,42\n2022-11-28,155\n2022-11-29,130\n2022-11-30,149\n2022-12-01,124\n2022-12-02,157\n2022-12-03,51\n2022-12-04,56\n2022-12-05,123\n2022-12-06,133\n2022-12-07,125\n2022-12-08,111\n2022-12-09,139\n2022-12-10,66\n2022-12-11,81\n2022-12-12,102\n2022-12-13,143\n2022-12-14,96\n2022-12-15,163\n2022-12-16,159\n2022-12-17,70\n2022-12-18,68\n2022-12-19,93\n2022-12-20,132\n2022-12-21,119\n2022-12-22,119\n2022-12-23,84\n2022-12-24,54\n2022-12-25,22\n2022-12-26,44\n2022-12-27,64\n2022-12-28,57\n2022-12-29,85\n2022-12-30,55\n2022-12-31,39\n2023-01-01,40\n2023-01-02,93\n2023-01-03,99\n2023-01-04,109\n2023-01-05,114\n2023-01-06,110\n2023-01-07,60\n2023-01-08,56\n2023-01-09,144\n2023-01-10,104\n2023-01-11,144\n2023-01-12,117\n2023-01-13,154\n2023-01-14,60\n2023-01-15,40\n2023-01-16,118\n2023-01-17,111\n2023-01-18,128\n2023-01-19,175\n2023-01-20,169\n2023-01-21,68\n2023-01-22,88\n2023-01-23,154\n2023-01-24,103\n2023-01-25,117\n2023-01-26,128\n2023-01-27,135\n2023-01-28,47\n2023-01-29,74\n2023-01-30,149\n2023-01-31,175\n2023-02-01,136\n2023-02-02,159\n2023-02-03,79\n2023-02-04,80\n2023-02-05,54\n2023-02-06,125\n2023-02-07,142\n2023-02-08,132\n2023-02-09,119\n2023-02-10,102\n2023-02-11,47\n2023-02-12,50\n2023-02-13,129\n2023-02-14,110\n2023-02-15,101\n2023-02-16,116\n2023-02-17,75\n2023-02-18,59\n2023-02-19,73\n2023-02-20,103\n2023-02-21,100\n2023-02-22,115\n2023-02-23,99\n2023-02-24,155\n2023-02-25,61\n2023-02-26,43\n2023-02-27,128\n2023-02-28,138\n2023-03-01,106\n2023-03-02,101\n2023-03-03,127\n2023-03-04,44\n2023-03-05,46\n2023-03-06,129\n2023-03-07,113\n2023-03-08,97\n2023-03-09,104\n2023-03-10,129\n2023-03-11,59\n2023-03-12,54\n2023-03-13,104\n2023-03-14,109\n2023-03-15,109\n2023-03-16,132\n2023-03-17,96\n2023-03-18,79\n2023-03-19,65\n2023-03-20,71\n2023-03-21,75\n2023-03-22,118\n2023-03-23,105\n2023-03-24,87\n2023-03-25,45\n2023-03-26,41\n2023-03-27,119\n2023-03-28,138\n2023-03-29,123\n2023-03-30,148\n2023-03-31,137\n2023-04-01,67\n2023-04-02,52\n2023-04-03,116\n2023-04-04,101\n2023-04-05,124\n2023-04-06,94\n2023-04-07,79\n2023-04-08,53\n2023-04-09,47\n2023-04-10,80\n2023-04-11,103\n2023-04-12,134\n2023-04-13,149\n2023-04-14,100\n2023-04-15,52\n2023-04-16,60\n2023-04-17,86\n2023-04-18,117\n2023-04-19,86\n2023-04-20,71\n2023-04-21,79\n2023-04-22,45\n2023-04-23,38\n2023-04-24,119\n2023-04-25,111\n2023-04-26,120\n2023-04-27,82\n2023-04-28,84\n2023-04-29,66\n2023-04-30,22\n2023-05-01,71\n2023-05-02,85\n2023-05-03,72\n2023-05-04,92\n2023-05-05,94\n2023-05-06,27\n2023-05-07,34\n2023-05-08,88\n2023-05-09,88\n2023-05-10,115\n2023-05-11,116\n2023-05-12,113\n2023-05-13,28\n2023-05-14,14\n2023-05-15,96\n2023-05-16,98\n2023-05-17,95\n2023-05-18,124\n2023-05-19,97\n2023-05-20,44\n2023-05-21,70\n2023-05-22,118\n2023-05-23,106\n2023-05-24,108\n2023-05-25,73\n2023-05-26,106\n2023-05-27,42\n2023-05-28,37\n2023-05-29,54\n2023-05-30,98\n2023-05-31,95\n2023-06-01,103\n2023-06-02,92\n2023-06-03,42\n2023-06-04,35\n2023-06-05,68\n2023-06-06,71\n2023-06-07,61\n2023-06-08,84\n2023-06-09,83\n2023-06-10,37\n2023-06-11,29\n2023-06-12,78\n2023-06-13,78\n2023-06-14,90\n2023-06-15,80\n2023-06-16,98\n2023-06-17,54\n2023-06-18,30\n2023-06-19,80\n2023-06-20,115\n2023-06-21,121\n2023-06-22,104\n2023-06-23,73\n2023-06-24,38\n2023-06-25,41\n2023-06-26,126\n2023-06-27,94\n2023-06-28,74\n2023-06-29,102\n2023-06-30,118\n2023-07-01,46\n2023-07-02,35\n2023-07-03,107\n2023-07-04,67\n2023-07-05,90\n2023-07-06,132\n2023-07-07,110\n2023-07-08,25\n2023-07-09,26\n2023-07-10,85\n2023-07-11,90\n2023-07-12,83\n2023-07-13,113\n2023-07-14,121\n2023-07-15,27\n2023-07-16,23\n2023-07-17,48\n2023-07-18,63\n2023-07-19,52\n2023-07-20,67\n2023-07-21,78\n2023-07-22,38\n2023-07-23,41\n2023-07-24,113\n2023-07-25,99\n2023-07-26,103\n2023-07-27,88\n2023-07-28,88\n2023-07-29,52\n2023-07-30,36\n2023-07-31,86\n2023-08-01,84\n2023-08-02,104\n2023-08-03,94\n2023-08-04,89\n2023-08-05,46\n2023-08-06,41\n2023-08-07,53\n2023-08-08,100\n2023-08-09,111\n2023-08-10,180\n2023-08-11,82\n2023-08-12,84\n2023-08-13,33\n2023-08-14,100\n2023-08-15,93\n2023-08-16,177\n2023-08-17,118\n2023-08-18,93\n2023-08-19,57\n2023-08-20,61\n2023-08-21,90\n2023-08-22,75\n2023-08-23,115\n2023-08-24,109\n2023-08-25,78\n2023-08-26,48\n2023-08-27,45\n2023-08-28,104\n2023-08-29,79\n2023-08-30,129\n2023-08-31,104\n2023-09-01,82\n2023-09-02,39\n2023-09-03,25\n2023-09-04,43\n2023-09-05,114\n2023-09-06,90\n2023-09-07,120\n2023-09-08,72\n2023-09-09,74\n2023-09-10,43\n2023-09-11,74\n2023-09-12,76\n2023-09-13,59\n2023-09-14,79\n2023-09-15,76\n2023-09-16,55\n2023-09-17,24\n2023-09-18,99\n2023-09-19,77\n2023-09-20,65\n2023-09-21,71\n2023-09-22,54\n2023-09-23,26\n2023-09-24,49\n2023-09-25,79\n2023-09-26,80\n2023-09-27,113\n2023-09-28,91\n2023-09-29,76\n2023-09-30,33\n2023-10-01,28\n2023-10-02,80\n2023-10-03,80\n2023-10-04,88\n2023-10-05,77\n2023-10-06,96\n2023-10-07,51\n2023-10-08,48\n2023-10-09,88\n2023-10-10,76\n2023-10-11,56\n2023-10-12,90\n2023-10-13,73\n2023-10-14,30\n2023-10-15,34\n2023-10-16,97\n2023-10-17,118\n2023-10-18,125\n2023-10-19,93\n2023-10-20,81\n2023-10-21,55\n2023-10-22,44\n2023-10-23,53\n2023-10-24,84\n2023-10-25,102\n2023-10-26,97\n2023-10-27,89\n2023-10-28,44\n2023-10-29,69\n2023-10-30,127\n2023-10-31,134\n2023-11-01,112\n2023-11-02,85\n2023-11-03,84\n2023-11-04,52\n2023-11-05,63\n2023-11-06,114\n2023-11-07,140\n2023-11-08,99\n2023-11-09,99\n2023-11-10,98\n2023-11-11,67\n2023-11-12,52\n2023-11-13,100\n2023-11-14,106\n2023-11-15,122\n2023-11-16,113\n2023-11-17,107\n2023-11-18,47\n2023-11-19,35\n2023-11-20,84\n2023-11-21,108\n2023-11-22,91\n2023-11-23,86\n2023-11-24,95\n2023-11-25,32\n2023-11-26,54\n2023-11-27,66\n2023-11-28,95\n2023-11-29,87\n2023-11-30,91\n2023-12-01,43\n2023-12-02,32\n2023-12-03,19\n","format":{"type":"csv"}}});</script></div>

<div class="sourceClojure">
```clojure
(defn add-temporal-field [ds tf]
  (-> ds
      (tc/add-column tf (fn [ds]
                          (->> ds
                               :date
                               (datetime/long-temporal-field tf))))))
```
</div>

<div class="sourceClojure">
```clojure
(def processed-date-counts
  (-> date-counts
      (add-temporal-field :day-of-week)
      (add-temporal-field :day-of-year)
      (add-temporal-field :years)))
```
</div>

<div class="sourceClojure">
```clojure
(-> processed-date-counts
    :day-of-year
    fun/reduce-max)
```
</div>

<div class="printedClojure">
```clojure
366

```
</div>

<div class="sourceClojure">
```clojure
(-> processed-date-counts
    (tc/select-rows (fn [row]
                      (-> row :years (#{2008 2012 2016 2020}))))
    (vis/hanami-plot ht/line-chart
                     {:X "day-of-year"
                      :Y "n"
                      :COLOR "years"
                      :OPACITY 0.5}))
```
</div>

<div><script>vegaEmbed(document.currentScript.parentElement, {"encoding":{"y":{"field":"n","type":"quantitative"},"color":{"field":"years","type":"nominal"},"opacity":{"value":0.5},"x":{"field":"day-of-year","type":"quantitative"}},"mark":{"type":"line","tooltip":true},"width":400,"background":"floralwhite","height":300,"data":{"values":"date,n,day-of-week,day-of-year,years\n2008-03-29,7,6,89,2008\n2008-04-02,4,3,93,2008\n2008-04-03,9,4,94,2008\n2008-04-05,2,6,96,2008\n2008-04-06,1,7,97,2008\n2008-04-07,5,1,98,2008\n2008-04-08,9,2,99,2008\n2008-04-10,1,4,101,2008\n2008-04-11,1,5,102,2008\n2008-04-16,1,3,107,2008\n2008-04-18,1,5,109,2008\n2008-04-19,1,6,110,2008\n2008-04-21,1,1,112,2008\n2008-04-24,3,4,115,2008\n2008-04-25,1,5,116,2008\n2008-04-29,1,2,120,2008\n2008-04-30,4,3,121,2008\n2008-05-01,1,4,122,2008\n2008-05-02,2,5,123,2008\n2008-05-03,2,6,124,2008\n2008-05-07,3,3,128,2008\n2008-05-08,2,4,129,2008\n2008-05-10,6,6,131,2008\n2008-05-11,1,7,132,2008\n2008-05-12,1,1,133,2008\n2008-05-13,2,2,134,2008\n2008-05-14,1,3,135,2008\n2008-05-15,1,4,136,2008\n2008-05-16,2,5,137,2008\n2008-05-17,9,6,138,2008\n2008-05-18,12,7,139,2008\n2008-05-22,1,4,143,2008\n2008-05-24,3,6,145,2008\n2008-05-25,2,7,146,2008\n2008-05-26,1,1,147,2008\n2008-05-27,1,2,148,2008\n2008-05-28,1,3,149,2008\n2008-05-29,4,4,150,2008\n2008-05-31,1,6,152,2008\n2008-06-01,2,7,153,2008\n2008-06-02,1,1,154,2008\n2008-06-03,3,2,155,2008\n2008-06-04,2,3,156,2008\n2008-06-05,1,4,157,2008\n2008-06-06,1,5,158,2008\n2008-06-07,1,6,159,2008\n2008-06-08,5,7,160,2008\n2008-06-09,3,1,161,2008\n2008-06-11,2,3,163,2008\n2008-06-12,2,4,164,2008\n2008-06-13,2,5,165,2008\n2008-06-14,4,6,166,2008\n2008-06-15,5,7,167,2008\n2008-06-18,1,3,170,2008\n2008-06-19,3,4,171,2008\n2008-06-27,1,5,179,2008\n2008-06-28,6,6,180,2008\n2008-06-29,6,7,181,2008\n2008-06-30,2,1,182,2008\n2008-07-03,2,4,185,2008\n2008-07-04,1,5,186,2008\n2008-07-06,5,7,188,2008\n2008-07-07,2,1,189,2008\n2008-07-08,1,2,190,2008\n2008-07-09,8,3,191,2008\n2008-07-10,5,4,192,2008\n2008-07-11,6,5,193,2008\n2008-07-12,4,6,194,2008\n2008-07-13,3,7,195,2008\n2008-07-14,1,1,196,2008\n2008-07-15,5,2,197,2008\n2008-07-21,1,1,203,2008\n2008-07-22,5,2,204,2008\n2008-07-24,2,4,206,2008\n2008-07-25,2,5,207,2008\n2008-07-26,7,6,208,2008\n2008-07-27,3,7,209,2008\n2008-07-29,8,2,211,2008\n2008-08-05,1,2,218,2008\n2008-08-06,4,3,219,2008\n2008-08-07,7,4,220,2008\n2008-08-08,1,5,221,2008\n2008-08-09,1,6,222,2008\n2008-08-10,1,7,223,2008\n2008-08-11,1,1,224,2008\n2008-08-12,4,2,225,2008\n2008-08-13,1,3,226,2008\n2008-08-14,1,4,227,2008\n2008-08-15,3,5,228,2008\n2008-08-16,9,6,229,2008\n2008-08-17,4,7,230,2008\n2008-08-19,7,2,232,2008\n2008-08-20,3,3,233,2008\n2008-08-21,1,4,234,2008\n2008-08-25,2,1,238,2008\n2008-08-27,2,3,240,2008\n2008-08-28,2,4,241,2008\n2008-08-29,5,5,242,2008\n2008-08-30,8,6,243,2008\n2008-09-02,8,2,246,2008\n2008-09-03,7,3,247,2008\n2008-09-04,1,4,248,2008\n2008-09-05,1,5,249,2008\n2008-09-06,2,6,250,2008\n2008-09-07,3,7,251,2008\n2008-09-08,29,1,252,2008\n2008-09-09,4,2,253,2008\n2008-09-10,3,3,254,2008\n2008-09-11,2,4,255,2008\n2008-09-12,3,5,256,2008\n2008-09-13,11,6,257,2008\n2008-09-14,2,7,258,2008\n2008-09-15,8,1,259,2008\n2008-09-16,11,2,260,2008\n2008-09-17,8,3,261,2008\n2008-09-18,1,4,262,2008\n2008-09-19,2,5,263,2008\n2008-09-20,8,6,264,2008\n2008-09-21,7,7,265,2008\n2008-09-22,3,1,266,2008\n2008-09-23,4,2,267,2008\n2008-09-24,4,3,268,2008\n2008-09-25,6,4,269,2008\n2008-09-26,4,5,270,2008\n2008-09-27,2,6,271,2008\n2008-09-28,7,7,272,2008\n2008-10-01,6,3,275,2008\n2008-10-02,10,4,276,2008\n2008-10-03,5,5,277,2008\n2008-10-04,3,6,278,2008\n2008-10-05,5,7,279,2008\n2008-10-06,11,1,280,2008\n2008-10-07,2,2,281,2008\n2008-10-08,6,3,282,2008\n2008-10-09,3,4,283,2008\n2008-10-10,4,5,284,2008\n2008-10-11,2,6,285,2008\n2008-10-12,4,7,286,2008\n2008-10-13,2,1,287,2008\n2008-10-14,1,2,288,2008\n2008-10-15,10,3,289,2008\n2008-10-16,1,4,290,2008\n2008-10-17,4,5,291,2008\n2008-10-18,2,6,292,2008\n2008-10-19,8,7,293,2008\n2008-10-21,1,2,295,2008\n2008-10-22,5,3,296,2008\n2008-10-23,4,4,297,2008\n2008-10-25,2,6,299,2008\n2008-10-26,3,7,300,2008\n2008-10-27,1,1,301,2008\n2008-10-28,2,2,302,2008\n2008-10-29,1,3,303,2008\n2008-10-30,5,4,304,2008\n2008-10-31,5,5,305,2008\n2008-11-01,9,6,306,2008\n2008-11-02,1,7,307,2008\n2008-11-03,7,1,308,2008\n2008-11-08,3,6,313,2008\n2008-11-09,5,7,314,2008\n2008-11-11,1,2,316,2008\n2008-11-12,5,3,317,2008\n2008-11-13,23,4,318,2008\n2008-11-14,2,5,319,2008\n2008-11-15,8,6,320,2008\n2008-11-16,17,7,321,2008\n2008-11-17,3,1,322,2008\n2008-11-18,25,2,323,2008\n2008-11-19,3,3,324,2008\n2008-11-20,2,4,325,2008\n2008-11-21,3,5,326,2008\n2008-11-22,3,6,327,2008\n2008-11-23,7,7,328,2008\n2008-11-24,1,1,329,2008\n2008-11-25,6,2,330,2008\n2008-11-26,3,3,331,2008\n2008-11-27,3,4,332,2008\n2008-11-28,5,5,333,2008\n2008-11-29,5,6,334,2008\n2008-11-30,5,7,335,2008\n2008-12-01,7,1,336,2008\n2008-12-02,3,2,337,2008\n2008-12-03,7,3,338,2008\n2008-12-04,5,4,339,2008\n2008-12-05,6,5,340,2008\n2008-12-07,3,7,342,2008\n2008-12-08,4,1,343,2008\n2008-12-09,1,2,344,2008\n2008-12-10,6,3,345,2008\n2008-12-11,5,4,346,2008\n2008-12-12,1,5,347,2008\n2008-12-13,12,6,348,2008\n2008-12-14,8,7,349,2008\n2008-12-15,9,1,350,2008\n2008-12-16,6,2,351,2008\n2008-12-17,6,3,352,2008\n2008-12-19,18,5,354,2008\n2008-12-20,1,6,355,2008\n2008-12-21,3,7,356,2008\n2008-12-22,3,1,357,2008\n2008-12-23,10,2,358,2008\n2008-12-24,3,3,359,2008\n2008-12-25,1,4,360,2008\n2008-12-26,1,5,361,2008\n2008-12-27,13,6,362,2008\n2008-12-28,13,7,363,2008\n2008-12-29,15,1,364,2008\n2008-12-30,10,2,365,2008\n2008-12-31,2,3,366,2008\n2012-01-01,56,7,1,2012\n2012-01-02,76,1,2,2012\n2012-01-03,60,2,3,2012\n2012-01-04,121,3,4,2012\n2012-01-05,77,4,5,2012\n2012-01-06,65,5,6,2012\n2012-01-07,60,6,7,2012\n2012-01-08,82,7,8,2012\n2012-01-09,72,1,9,2012\n2012-01-10,106,2,10,2012\n2012-01-11,96,3,11,2012\n2012-01-12,82,4,12,2012\n2012-01-13,145,5,13,2012\n2012-01-14,65,6,14,2012\n2012-01-15,64,7,15,2012\n2012-01-16,71,1,16,2012\n2012-01-17,79,2,17,2012\n2012-01-18,88,3,18,2012\n2012-01-19,82,4,19,2012\n2012-01-20,71,5,20,2012\n2012-01-21,57,6,21,2012\n2012-01-22,61,7,22,2012\n2012-01-23,69,1,23,2012\n2012-01-24,107,2,24,2012\n2012-01-25,99,3,25,2012\n2012-01-26,92,4,26,2012\n2012-01-27,70,5,27,2012\n2012-01-28,44,6,28,2012\n2012-01-29,84,7,29,2012\n2012-01-30,54,1,30,2012\n2012-01-31,71,2,31,2012\n2012-02-01,95,3,32,2012\n2012-02-02,86,4,33,2012\n2012-02-03,87,5,34,2012\n2012-02-04,82,6,35,2012\n2012-02-05,115,7,36,2012\n2012-02-06,189,1,37,2012\n2012-02-07,116,2,38,2012\n2012-02-08,122,3,39,2012\n2012-02-09,106,4,40,2012\n2012-02-10,97,5,41,2012\n2012-02-11,94,6,42,2012\n2012-02-12,90,7,43,2012\n2012-02-13,91,1,44,2012\n2012-02-14,106,2,45,2012\n2012-02-15,93,3,46,2012\n2012-02-16,119,4,47,2012\n2012-02-17,122,5,48,2012\n2012-02-18,109,6,49,2012\n2012-02-19,161,7,50,2012\n2012-02-20,120,1,51,2012\n2012-02-21,144,2,52,2012\n2012-02-22,117,3,53,2012\n2012-02-23,78,4,54,2012\n2012-02-24,68,5,55,2012\n2012-02-25,104,6,56,2012\n2012-02-26,81,7,57,2012\n2012-02-27,87,1,58,2012\n2012-02-28,110,2,59,2012\n2012-02-29,71,3,60,2012\n2012-03-01,83,4,61,2012\n2012-03-02,65,5,62,2012\n2012-03-03,82,6,63,2012\n2012-03-04,133,7,64,2012\n2012-03-05,139,1,65,2012\n2012-03-06,107,2,66,2012\n2012-03-07,77,3,67,2012\n2012-03-08,102,4,68,2012\n2012-03-09,111,5,69,2012\n2012-03-10,100,6,70,2012\n2012-03-11,166,7,71,2012\n2012-03-12,78,1,72,2012\n2012-03-13,100,2,73,2012\n2012-03-14,57,3,74,2012\n2012-03-15,53,4,75,2012\n2012-03-16,99,5,76,2012\n2012-03-17,88,6,77,2012\n2012-03-18,67,7,78,2012\n2012-03-19,160,1,79,2012\n2012-03-20,100,2,80,2012\n2012-03-21,101,3,81,2012\n2012-03-22,99,4,82,2012\n2012-03-23,123,5,83,2012\n2012-03-24,93,6,84,2012\n2012-03-25,142,7,85,2012\n2012-03-26,111,1,86,2012\n2012-03-27,84,2,87,2012\n2012-03-28,67,3,88,2012\n2012-03-29,76,4,89,2012\n2012-03-30,99,5,90,2012\n2012-03-31,109,6,91,2012\n2012-04-01,118,7,92,2012\n2012-04-02,124,1,93,2012\n2012-04-03,109,2,94,2012\n2012-04-04,175,3,95,2012\n2012-04-05,118,4,96,2012\n2012-04-06,92,5,97,2012\n2012-04-07,147,6,98,2012\n2012-04-08,80,7,99,2012\n2012-04-09,135,1,100,2012\n2012-04-10,120,2,101,2012\n2012-04-11,166,3,102,2012\n2012-04-12,77,4,103,2012\n2012-04-13,67,5,104,2012\n2012-04-14,94,6,105,2012\n2012-04-15,102,7,106,2012\n2012-04-16,123,1,107,2012\n2012-04-17,88,2,108,2012\n2012-04-18,130,3,109,2012\n2012-04-19,86,4,110,2012\n2012-04-20,108,5,111,2012\n2012-04-21,117,6,112,2012\n2012-04-22,95,7,113,2012\n2012-04-23,112,1,114,2012\n2012-04-24,106,2,115,2012\n2012-04-25,95,3,116,2012\n2012-04-26,85,4,117,2012\n2012-04-27,107,5,118,2012\n2012-04-28,120,6,119,2012\n2012-04-29,165,7,120,2012\n2012-04-30,127,1,121,2012\n2012-05-01,86,2,122,2012\n2012-05-02,107,3,123,2012\n2012-05-03,92,4,124,2012\n2012-05-04,107,5,125,2012\n2012-05-05,80,6,126,2012\n2012-05-06,103,7,127,2012\n2012-05-07,131,1,128,2012\n2012-05-08,73,2,129,2012\n2012-05-09,105,3,130,2012\n2012-05-10,81,4,131,2012\n2012-05-11,140,5,132,2012\n2012-05-12,83,6,133,2012\n2012-05-13,109,7,134,2012\n2012-05-14,67,1,135,2012\n2012-05-15,71,2,136,2012\n2012-05-16,86,3,137,2012\n2012-05-17,118,4,138,2012\n2012-05-18,101,5,139,2012\n2012-05-19,77,6,140,2012\n2012-05-20,110,7,141,2012\n2012-05-21,77,1,142,2012\n2012-05-22,81,2,143,2012\n2012-05-23,70,3,144,2012\n2012-05-24,79,4,145,2012\n2012-05-25,60,5,146,2012\n2012-05-26,65,6,147,2012\n2012-05-27,46,7,148,2012\n2012-05-28,74,1,149,2012\n2012-05-29,48,2,150,2012\n2012-05-30,90,3,151,2012\n2012-05-31,95,4,152,2012\n2012-06-01,84,5,153,2012\n2012-06-02,31,6,154,2012\n2012-06-03,73,7,155,2012\n2012-06-04,98,1,156,2012\n2012-06-05,63,2,157,2012\n2012-06-06,66,3,158,2012\n2012-06-07,57,4,159,2012\n2012-06-08,78,5,160,2012\n2012-06-09,54,6,161,2012\n2012-06-10,122,7,162,2012\n2012-06-11,95,1,163,2012\n2012-06-12,87,2,164,2012\n2012-06-13,113,3,165,2012\n2012-06-14,79,4,166,2012\n2012-06-15,107,5,167,2012\n2012-06-16,106,6,168,2012\n2012-06-17,96,7,169,2012\n2012-06-18,55,1,170,2012\n2012-06-19,79,2,171,2012\n2012-06-20,94,3,172,2012\n2012-06-21,54,4,173,2012\n2012-06-22,44,5,174,2012\n2012-06-23,67,6,175,2012\n2012-06-24,64,7,176,2012\n2012-06-25,96,1,177,2012\n2012-06-26,73,2,178,2012\n2012-06-27,108,3,179,2012\n2012-06-28,40,4,180,2012\n2012-06-29,87,5,181,2012\n2012-06-30,44,6,182,2012\n2012-07-01,61,7,183,2012\n2012-07-02,91,1,184,2012\n2012-07-03,118,2,185,2012\n2012-07-04,96,3,186,2012\n2012-07-05,123,4,187,2012\n2012-07-06,170,5,188,2012\n2012-07-07,117,6,189,2012\n2012-07-08,136,7,190,2012\n2012-07-09,81,1,191,2012\n2012-07-10,74,2,192,2012\n2012-07-11,96,3,193,2012\n2012-07-12,97,4,194,2012\n2012-07-13,74,5,195,2012\n2012-07-14,83,6,196,2012\n2012-07-15,64,7,197,2012\n2012-07-16,95,1,198,2012\n2012-07-17,119,2,199,2012\n2012-07-18,102,3,200,2012\n2012-07-19,70,4,201,2012\n2012-07-20,66,5,202,2012\n2012-07-21,102,6,203,2012\n2012-07-22,96,7,204,2012\n2012-07-23,107,1,205,2012\n2012-07-24,100,2,206,2012\n2012-07-25,76,3,207,2012\n2012-07-26,96,4,208,2012\n2012-07-27,81,5,209,2012\n2012-07-28,93,6,210,2012\n2012-07-29,74,7,211,2012\n2012-07-30,61,1,212,2012\n2012-07-31,73,2,213,2012\n2012-08-01,100,3,214,2012\n2012-08-02,128,4,215,2012\n2012-08-03,79,5,216,2012\n2012-08-04,106,6,217,2012\n2012-08-05,72,7,218,2012\n2012-08-06,109,1,219,2012\n2012-08-07,107,2,220,2012\n2012-08-08,119,3,221,2012\n2012-08-09,82,4,222,2012\n2012-08-10,100,5,223,2012\n2012-08-11,60,6,224,2012\n2012-08-12,77,7,225,2012\n2012-08-13,68,1,226,2012\n2012-08-14,88,2,227,2012\n2012-08-15,116,3,228,2012\n2012-08-16,94,4,229,2012\n2012-08-17,99,5,230,2012\n2012-08-18,141,6,231,2012\n2012-08-19,46,7,232,2012\n2012-08-20,61,1,233,2012\n2012-08-21,75,2,234,2012\n2012-08-22,84,3,235,2012\n2012-08-23,76,4,236,2012\n2012-08-24,102,5,237,2012\n2012-08-25,71,6,238,2012\n2012-08-26,83,7,239,2012\n2012-08-27,91,1,240,2012\n2012-08-28,80,2,241,2012\n2012-08-29,60,3,242,2012\n2012-08-30,78,4,243,2012\n2012-08-31,61,5,244,2012\n2012-09-01,57,6,245,2012\n2012-09-02,41,7,246,2012\n2012-09-03,86,1,247,2012\n2012-09-04,88,2,248,2012\n2012-09-05,73,3,249,2012\n2012-09-06,100,4,250,2012\n2012-09-07,106,5,251,2012\n2012-09-08,57,6,252,2012\n2012-09-09,40,7,253,2012\n2012-09-10,65,1,254,2012\n2012-09-11,59,2,255,2012\n2012-09-12,99,3,256,2012\n2012-09-13,55,4,257,2012\n2012-09-14,64,5,258,2012\n2012-09-15,40,6,259,2012\n2012-09-16,52,7,260,2012\n2012-09-17,57,1,261,2012\n2012-09-18,51,2,262,2012\n2012-09-19,38,3,263,2012\n2012-09-20,47,4,264,2012\n2012-09-21,53,5,265,2012\n2012-09-22,69,6,266,2012\n2012-09-23,40,7,267,2012\n2012-09-24,32,1,268,2012\n2012-09-25,45,2,269,2012\n2012-09-26,54,3,270,2012\n2012-09-27,59,4,271,2012\n2012-09-28,61,5,272,2012\n2012-09-29,58,6,273,2012\n2012-09-30,62,7,274,2012\n2012-10-01,96,1,275,2012\n2012-10-02,101,2,276,2012\n2012-10-03,98,3,277,2012\n2012-10-04,78,4,278,2012\n2012-10-05,107,5,279,2012\n2012-10-06,59,6,280,2012\n2012-10-07,48,7,281,2012\n2012-10-08,60,1,282,2012\n2012-10-09,76,2,283,2012\n2012-10-10,126,3,284,2012\n2012-10-11,58,4,285,2012\n2012-10-12,70,5,286,2012\n2012-10-13,37,6,287,2012\n2012-10-14,70,7,288,2012\n2012-10-15,79,1,289,2012\n2012-10-16,77,2,290,2012\n2012-10-17,33,3,291,2012\n2012-10-18,86,4,292,2012\n2012-10-19,62,5,293,2012\n2012-10-20,36,6,294,2012\n2012-10-21,18,7,295,2012\n2012-10-22,55,1,296,2012\n2012-10-23,78,2,297,2012\n2012-10-24,58,3,298,2012\n2012-10-25,87,4,299,2012\n2012-10-26,112,5,300,2012\n2012-10-27,48,6,301,2012\n2012-10-28,53,7,302,2012\n2012-10-29,38,1,303,2012\n2012-10-30,92,2,304,2012\n2012-10-31,54,3,305,2012\n2012-11-01,109,4,306,2012\n2012-11-02,119,5,307,2012\n2012-11-03,71,6,308,2012\n2012-11-04,59,7,309,2012\n2012-11-05,91,1,310,2012\n2012-11-06,39,2,311,2012\n2012-11-07,56,3,312,2012\n2012-11-08,66,4,313,2012\n2012-11-09,56,5,314,2012\n2012-11-10,39,6,315,2012\n2012-11-11,20,7,316,2012\n2012-11-12,68,1,317,2012\n2012-11-13,66,2,318,2012\n2012-11-14,58,3,319,2012\n2012-11-15,47,4,320,2012\n2012-11-16,70,5,321,2012\n2012-11-17,67,6,322,2012\n2012-11-18,38,7,323,2012\n2012-11-19,95,1,324,2012\n2012-11-20,114,2,325,2012\n2012-11-21,87,3,326,2012\n2012-11-22,49,4,327,2012\n2012-11-23,80,5,328,2012\n2012-11-24,70,6,329,2012\n2012-11-25,35,7,330,2012\n2012-11-26,71,1,331,2012\n2012-11-27,79,2,332,2012\n2012-11-28,64,3,333,2012\n2012-11-29,58,4,334,2012\n2012-11-30,36,5,335,2012\n2012-12-01,56,6,336,2012\n2012-12-02,58,7,337,2012\n2012-12-03,84,1,338,2012\n2012-12-04,103,2,339,2012\n2012-12-05,118,3,340,2012\n2012-12-06,46,4,341,2012\n2012-12-07,58,5,342,2012\n2012-12-08,98,6,343,2012\n2012-12-09,43,7,344,2012\n2012-12-10,70,1,345,2012\n2012-12-11,91,2,346,2012\n2012-12-12,74,3,347,2012\n2012-12-13,88,4,348,2012\n2012-12-14,74,5,349,2012\n2012-12-15,54,6,350,2012\n2012-12-16,81,7,351,2012\n2012-12-17,121,1,352,2012\n2012-12-18,111,2,353,2012\n2012-12-19,112,3,354,2012\n2012-12-20,103,4,355,2012\n2012-12-21,63,5,356,2012\n2012-12-22,59,6,357,2012\n2012-12-23,91,7,358,2012\n2012-12-24,99,1,359,2012\n2012-12-25,60,2,360,2012\n2012-12-26,61,3,361,2012\n2012-12-27,90,4,362,2012\n2012-12-28,59,5,363,2012\n2012-12-29,69,6,364,2012\n2012-12-30,66,7,365,2012\n2012-12-31,110,1,366,2012\n2016-01-01,131,5,1,2016\n2016-01-02,151,6,2,2016\n2016-01-03,154,7,3,2016\n2016-01-04,185,1,4,2016\n2016-01-05,226,2,5,2016\n2016-01-06,276,3,6,2016\n2016-01-07,310,4,7,2016\n2016-01-08,242,5,8,2016\n2016-01-09,197,6,9,2016\n2016-01-10,159,7,10,2016\n2016-01-11,330,1,11,2016\n2016-01-12,212,2,12,2016\n2016-01-13,261,3,13,2016\n2016-01-14,294,4,14,2016\n2016-01-15,263,5,15,2016\n2016-01-16,228,6,16,2016\n2016-01-17,200,7,17,2016\n2016-01-18,203,1,18,2016\n2016-01-19,214,2,19,2016\n2016-01-20,195,3,20,2016\n2016-01-21,268,4,21,2016\n2016-01-22,209,5,22,2016\n2016-01-23,169,6,23,2016\n2016-01-24,214,7,24,2016\n2016-01-25,251,1,25,2016\n2016-01-26,224,2,26,2016\n2016-01-27,211,3,27,2016\n2016-01-28,238,4,28,2016\n2016-01-29,316,5,29,2016\n2016-01-30,187,6,30,2016\n2016-01-31,253,7,31,2016\n2016-02-01,271,1,32,2016\n2016-02-02,307,2,33,2016\n2016-02-03,221,3,34,2016\n2016-02-04,209,4,35,2016\n2016-02-05,201,5,36,2016\n2016-02-06,165,6,37,2016\n2016-02-07,163,7,38,2016\n2016-02-08,191,1,39,2016\n2016-02-09,217,2,40,2016\n2016-02-10,192,3,41,2016\n2016-02-11,218,4,42,2016\n2016-02-12,291,5,43,2016\n2016-02-13,137,6,44,2016\n2016-02-14,145,7,45,2016\n2016-02-15,182,1,46,2016\n2016-02-16,249,2,47,2016\n2016-02-17,238,3,48,2016\n2016-02-18,224,4,49,2016\n2016-02-19,258,5,50,2016\n2016-02-20,136,6,51,2016\n2016-02-21,168,7,52,2016\n2016-02-22,211,1,53,2016\n2016-02-23,256,2,54,2016\n2016-02-24,232,3,55,2016\n2016-02-25,242,4,56,2016\n2016-02-26,232,5,57,2016\n2016-02-27,179,6,58,2016\n2016-02-28,246,7,59,2016\n2016-02-29,257,1,60,2016\n2016-03-01,208,2,61,2016\n2016-03-02,178,3,62,2016\n2016-03-03,208,4,63,2016\n2016-03-04,208,5,64,2016\n2016-03-05,127,6,65,2016\n2016-03-06,119,7,66,2016\n2016-03-07,208,1,67,2016\n2016-03-08,236,2,68,2016\n2016-03-09,191,3,69,2016\n2016-03-10,136,4,70,2016\n2016-03-11,180,5,71,2016\n2016-03-12,98,6,72,2016\n2016-03-13,97,7,73,2016\n2016-03-14,142,1,74,2016\n2016-03-15,180,2,75,2016\n2016-03-16,245,3,76,2016\n2016-03-17,232,4,77,2016\n2016-03-18,188,5,78,2016\n2016-03-19,119,6,79,2016\n2016-03-20,159,7,80,2016\n2016-03-21,216,1,81,2016\n2016-03-22,200,2,82,2016\n2016-03-23,228,3,83,2016\n2016-03-24,184,4,84,2016\n2016-03-25,272,5,85,2016\n2016-03-26,145,6,86,2016\n2016-03-27,101,7,87,2016\n2016-03-28,241,1,88,2016\n2016-03-29,204,2,89,2016\n2016-03-30,153,3,90,2016\n2016-03-31,177,4,91,2016\n2016-04-01,245,5,92,2016\n2016-04-02,175,6,93,2016\n2016-04-03,175,7,94,2016\n2016-04-04,206,1,95,2016\n2016-04-05,190,2,96,2016\n2016-04-06,189,3,97,2016\n2016-04-07,174,4,98,2016\n2016-04-08,182,5,99,2016\n2016-04-09,182,6,100,2016\n2016-04-10,124,7,101,2016\n2016-04-11,205,1,102,2016\n2016-04-12,218,2,103,2016\n2016-04-13,233,3,104,2016\n2016-04-14,231,4,105,2016\n2016-04-15,155,5,106,2016\n2016-04-16,174,6,107,2016\n2016-04-17,61,7,108,2016\n2016-04-18,197,1,109,2016\n2016-04-19,197,2,110,2016\n2016-04-20,270,3,111,2016\n2016-04-21,174,4,112,2016\n2016-04-22,166,5,113,2016\n2016-04-23,131,6,114,2016\n2016-04-24,89,7,115,2016\n2016-04-25,185,1,116,2016\n2016-04-26,155,2,117,2016\n2016-04-27,147,3,118,2016\n2016-04-28,129,4,119,2016\n2016-04-29,182,5,120,2016\n2016-04-30,114,6,121,2016\n2016-05-01,114,7,122,2016\n2016-05-02,268,1,123,2016\n2016-05-03,157,2,124,2016\n2016-05-04,184,3,125,2016\n2016-05-05,176,4,126,2016\n2016-05-06,162,5,127,2016\n2016-05-07,147,6,128,2016\n2016-05-08,80,7,129,2016\n2016-05-09,190,1,130,2016\n2016-05-10,205,2,131,2016\n2016-05-11,190,3,132,2016\n2016-05-12,140,4,133,2016\n2016-05-13,192,5,134,2016\n2016-05-14,77,6,135,2016\n2016-05-15,98,7,136,2016\n2016-05-16,187,1,137,2016\n2016-05-17,144,2,138,2016\n2016-05-18,160,3,139,2016\n2016-05-19,197,4,140,2016\n2016-05-20,183,5,141,2016\n2016-05-21,135,6,142,2016\n2016-05-22,150,7,143,2016\n2016-05-23,211,1,144,2016\n2016-05-24,236,2,145,2016\n2016-05-25,167,3,146,2016\n2016-05-26,160,4,147,2016\n2016-05-27,131,5,148,2016\n2016-05-28,97,6,149,2016\n2016-05-29,118,7,150,2016\n2016-05-30,162,1,151,2016\n2016-05-31,206,2,152,2016\n2016-06-01,280,3,153,2016\n2016-06-02,224,4,154,2016\n2016-06-03,255,5,155,2016\n2016-06-04,130,6,156,2016\n2016-06-05,119,7,157,2016\n2016-06-06,176,1,158,2016\n2016-06-07,212,2,159,2016\n2016-06-08,211,3,160,2016\n2016-06-09,209,4,161,2016\n2016-06-10,195,5,162,2016\n2016-06-11,178,6,163,2016\n2016-06-12,129,7,164,2016\n2016-06-13,292,1,165,2016\n2016-06-14,211,2,166,2016\n2016-06-15,196,3,167,2016\n2016-06-16,192,4,168,2016\n2016-06-17,199,5,169,2016\n2016-06-18,116,6,170,2016\n2016-06-19,103,7,171,2016\n2016-06-20,271,1,172,2016\n2016-06-21,233,2,173,2016\n2016-06-22,220,3,174,2016\n2016-06-23,175,4,175,2016\n2016-06-24,165,5,176,2016\n2016-06-25,67,6,177,2016\n2016-06-26,116,7,178,2016\n2016-06-27,271,1,179,2016\n2016-06-28,214,2,180,2016\n2016-06-29,193,3,181,2016\n2016-06-30,243,4,182,2016\n2016-07-01,218,5,183,2016\n2016-07-02,170,6,184,2016\n2016-07-03,102,7,185,2016\n2016-07-04,123,1,186,2016\n2016-07-05,148,2,187,2016\n2016-07-06,222,3,188,2016\n2016-07-07,204,4,189,2016\n2016-07-08,172,5,190,2016\n2016-07-09,115,6,191,2016\n2016-07-10,115,7,192,2016\n2016-07-11,175,1,193,2016\n2016-07-12,189,2,194,2016\n2016-07-13,168,3,195,2016\n2016-07-14,164,4,196,2016\n2016-07-15,159,5,197,2016\n2016-07-16,91,6,198,2016\n2016-07-17,135,7,199,2016\n2016-07-18,210,1,200,2016\n2016-07-19,208,2,201,2016\n2016-07-20,199,3,202,2016\n2016-07-21,155,4,203,2016\n2016-07-22,184,5,204,2016\n2016-07-23,98,6,205,2016\n2016-07-24,117,7,206,2016\n2016-07-25,200,1,207,2016\n2016-07-26,233,2,208,2016\n2016-07-27,221,3,209,2016\n2016-07-28,267,4,210,2016\n2016-07-29,219,5,211,2016\n2016-07-30,122,6,212,2016\n2016-07-31,133,7,213,2016\n2016-08-01,178,1,214,2016\n2016-08-02,170,2,215,2016\n2016-08-03,131,3,216,2016\n2016-08-04,230,4,217,2016\n2016-08-05,155,5,218,2016\n2016-08-06,71,6,219,2016\n2016-08-07,105,7,220,2016\n2016-08-08,158,1,221,2016\n2016-08-09,205,2,222,2016\n2016-08-10,227,3,223,2016\n2016-08-11,206,4,224,2016\n2016-08-12,194,5,225,2016\n2016-08-13,66,6,226,2016\n2016-08-14,97,7,227,2016\n2016-08-15,165,1,228,2016\n2016-08-16,170,2,229,2016\n2016-08-17,191,3,230,2016\n2016-08-18,180,4,231,2016\n2016-08-19,181,5,232,2016\n2016-08-20,126,6,233,2016\n2016-08-21,143,7,234,2016\n2016-08-22,207,1,235,2016\n2016-08-23,227,2,236,2016\n2016-08-24,206,3,237,2016\n2016-08-25,242,4,238,2016\n2016-08-26,140,5,239,2016\n2016-08-27,104,6,240,2016\n2016-08-28,111,7,241,2016\n2016-08-29,192,1,242,2016\n2016-08-30,182,2,243,2016\n2016-08-31,254,3,244,2016\n2016-09-01,265,4,245,2016\n2016-09-02,143,5,246,2016\n2016-09-03,74,6,247,2016\n2016-09-04,122,7,248,2016\n2016-09-05,130,1,249,2016\n2016-09-06,213,2,250,2016\n2016-09-07,177,3,251,2016\n2016-09-08,169,4,252,2016\n2016-09-09,174,5,253,2016\n2016-09-10,65,6,254,2016\n2016-09-11,89,7,255,2016\n2016-09-12,153,1,256,2016\n2016-09-13,119,2,257,2016\n2016-09-14,141,3,258,2016\n2016-09-15,179,4,259,2016\n2016-09-16,190,5,260,2016\n2016-09-17,85,6,261,2016\n2016-09-18,150,7,262,2016\n2016-09-19,200,1,263,2016\n2016-09-20,169,2,264,2016\n2016-09-21,149,3,265,2016\n2016-09-22,170,4,266,2016\n2016-09-23,131,5,267,2016\n2016-09-24,76,6,268,2016\n2016-09-25,78,7,269,2016\n2016-09-26,163,1,270,2016\n2016-09-27,187,2,271,2016\n2016-09-28,259,3,272,2016\n2016-09-29,154,4,273,2016\n2016-09-30,150,5,274,2016\n2016-10-01,98,6,275,2016\n2016-10-02,103,7,276,2016\n2016-10-03,199,1,277,2016\n2016-10-04,213,2,278,2016\n2016-10-05,263,3,279,2016\n2016-10-06,212,4,280,2016\n2016-10-07,226,5,281,2016\n2016-10-08,135,6,282,2016\n2016-10-09,93,7,283,2016\n2016-10-10,181,1,284,2016\n2016-10-11,166,2,285,2016\n2016-10-12,224,3,286,2016\n2016-10-13,221,4,287,2016\n2016-10-14,194,5,288,2016\n2016-10-15,97,6,289,2016\n2016-10-16,126,7,290,2016\n2016-10-17,141,1,291,2016\n2016-10-18,146,2,292,2016\n2016-10-19,114,3,293,2016\n2016-10-20,142,4,294,2016\n2016-10-21,182,5,295,2016\n2016-10-22,95,6,296,2016\n2016-10-23,124,7,297,2016\n2016-10-24,132,1,298,2016\n2016-10-25,140,2,299,2016\n2016-10-26,148,3,300,2016\n2016-10-27,124,4,301,2016\n2016-10-28,123,5,302,2016\n2016-10-29,156,6,303,2016\n2016-10-30,81,7,304,2016\n2016-10-31,82,1,305,2016\n2016-11-01,127,2,306,2016\n2016-11-02,146,3,307,2016\n2016-11-03,156,4,308,2016\n2016-11-04,148,5,309,2016\n2016-11-05,78,6,310,2016\n2016-11-06,103,7,311,2016\n2016-11-07,156,1,312,2016\n2016-11-08,120,2,313,2016\n2016-11-09,147,3,314,2016\n2016-11-10,148,4,315,2016\n2016-11-11,114,5,316,2016\n2016-11-12,76,6,317,2016\n2016-11-13,158,7,318,2016\n2016-11-14,179,1,319,2016\n2016-11-15,214,2,320,2016\n2016-11-16,142,3,321,2016\n2016-11-17,184,4,322,2016\n2016-11-18,162,5,323,2016\n2016-11-19,97,6,324,2016\n2016-11-20,103,7,325,2016\n2016-11-21,162,1,326,2016\n2016-11-22,112,2,327,2016\n2016-11-23,86,3,328,2016\n2016-11-24,75,4,329,2016\n2016-11-25,56,5,330,2016\n2016-11-26,47,6,331,2016\n2016-11-27,78,7,332,2016\n2016-11-28,141,1,333,2016\n2016-11-29,169,2,334,2016\n2016-11-30,138,3,335,2016\n2016-12-01,137,4,336,2016\n2016-12-02,147,5,337,2016\n2016-12-03,123,6,338,2016\n2016-12-04,107,7,339,2016\n2016-12-05,214,1,340,2016\n2016-12-06,213,2,341,2016\n2016-12-07,190,3,342,2016\n2016-12-08,168,4,343,2016\n2016-12-09,156,5,344,2016\n2016-12-10,87,6,345,2016\n2016-12-11,106,7,346,2016\n2016-12-12,210,1,347,2016\n2016-12-13,227,2,348,2016\n2016-12-14,192,3,349,2016\n2016-12-15,163,4,350,2016\n2016-12-16,89,5,351,2016\n2016-12-17,95,6,352,2016\n2016-12-18,90,7,353,2016\n2016-12-19,144,1,354,2016\n2016-12-20,138,2,355,2016\n2016-12-21,152,3,356,2016\n2016-12-22,159,4,357,2016\n2016-12-23,122,5,358,2016\n2016-12-24,67,6,359,2016\n2016-12-25,57,7,360,2016\n2016-12-26,75,1,361,2016\n2016-12-27,111,2,362,2016\n2016-12-28,117,3,363,2016\n2016-12-29,129,4,364,2016\n2016-12-30,75,5,365,2016\n2016-12-31,23,6,366,2016\n2020-01-01,99,3,1,2020\n2020-01-02,142,4,2,2020\n2020-01-03,95,5,3,2020\n2020-01-04,120,6,4,2020\n2020-01-05,81,7,5,2020\n2020-01-06,125,1,6,2020\n2020-01-07,161,2,7,2020\n2020-01-08,169,3,8,2020\n2020-01-09,192,4,9,2020\n2020-01-10,251,5,10,2020\n2020-01-11,164,6,11,2020\n2020-01-12,219,7,12,2020\n2020-01-13,224,1,13,2020\n2020-01-14,193,2,14,2020\n2020-01-15,248,3,15,2020\n2020-01-16,255,4,16,2020\n2020-01-17,247,5,17,2020\n2020-01-18,117,6,18,2020\n2020-01-19,187,7,19,2020\n2020-01-20,138,1,20,2020\n2020-01-21,139,2,21,2020\n2020-01-22,199,3,22,2020\n2020-01-23,185,4,23,2020\n2020-01-24,143,5,24,2020\n2020-01-25,64,6,25,2020\n2020-01-26,89,7,26,2020\n2020-01-27,135,1,27,2020\n2020-01-28,101,2,28,2020\n2020-01-29,114,3,29,2020\n2020-01-30,155,4,30,2020\n2020-01-31,165,5,31,2020\n2020-02-01,89,6,32,2020\n2020-02-02,111,7,33,2020\n2020-02-03,141,1,34,2020\n2020-02-04,120,2,35,2020\n2020-02-05,145,3,36,2020\n2020-02-06,133,4,37,2020\n2020-02-07,149,5,38,2020\n2020-02-08,137,6,39,2020\n2020-02-09,119,7,40,2020\n2020-02-10,152,1,41,2020\n2020-02-11,162,2,42,2020\n2020-02-12,134,3,43,2020\n2020-02-13,170,4,44,2020\n2020-02-14,149,5,45,2020\n2020-02-15,95,6,46,2020\n2020-02-16,66,7,47,2020\n2020-02-17,129,1,48,2020\n2020-02-18,175,2,49,2020\n2020-02-19,133,3,50,2020\n2020-02-20,128,4,51,2020\n2020-02-21,133,5,52,2020\n2020-02-22,99,6,53,2020\n2020-02-23,165,7,54,2020\n2020-02-24,153,1,55,2020\n2020-02-25,146,2,56,2020\n2020-02-26,155,3,57,2020\n2020-02-27,144,4,58,2020\n2020-02-28,115,5,59,2020\n2020-02-29,54,6,60,2020\n2020-03-01,56,7,61,2020\n2020-03-02,166,1,62,2020\n2020-03-03,142,2,63,2020\n2020-03-04,120,3,64,2020\n2020-03-05,148,4,65,2020\n2020-03-06,119,5,66,2020\n2020-03-07,92,6,67,2020\n2020-03-08,143,7,68,2020\n2020-03-09,112,1,69,2020\n2020-03-10,111,2,70,2020\n2020-03-11,110,3,71,2020\n2020-03-12,82,4,72,2020\n2020-03-13,76,5,73,2020\n2020-03-14,74,6,74,2020\n2020-03-15,80,7,75,2020\n2020-03-16,91,1,76,2020\n2020-03-17,106,2,77,2020\n2020-03-18,132,3,78,2020\n2020-03-19,92,4,79,2020\n2020-03-20,115,5,80,2020\n2020-03-21,89,6,81,2020\n2020-03-22,105,7,82,2020\n2020-03-23,138,1,83,2020\n2020-03-24,144,2,84,2020\n2020-03-25,116,3,85,2020\n2020-03-26,99,4,86,2020\n2020-03-27,112,5,87,2020\n2020-03-28,123,6,88,2020\n2020-03-29,76,7,89,2020\n2020-03-30,107,1,90,2020\n2020-03-31,177,2,91,2020\n2020-04-01,158,3,92,2020\n2020-04-02,160,4,93,2020\n2020-04-03,145,5,94,2020\n2020-04-04,130,6,95,2020\n2020-04-05,95,7,96,2020\n2020-04-06,152,1,97,2020\n2020-04-07,157,2,98,2020\n2020-04-08,217,3,99,2020\n2020-04-09,160,4,100,2020\n2020-04-10,151,5,101,2020\n2020-04-11,127,6,102,2020\n2020-04-12,106,7,103,2020\n2020-04-13,208,1,104,2020\n2020-04-14,175,2,105,2020\n2020-04-15,146,3,106,2020\n2020-04-16,149,4,107,2020\n2020-04-17,110,5,108,2020\n2020-04-18,148,6,109,2020\n2020-04-19,151,7,110,2020\n2020-04-20,155,1,111,2020\n2020-04-21,141,2,112,2020\n2020-04-22,147,3,113,2020\n2020-04-23,146,4,114,2020\n2020-04-24,136,5,115,2020\n2020-04-25,98,6,116,2020\n2020-04-26,139,7,117,2020\n2020-04-27,180,1,118,2020\n2020-04-28,118,2,119,2020\n2020-04-29,166,3,120,2020\n2020-04-30,220,4,121,2020\n2020-05-01,237,5,122,2020\n2020-05-02,133,6,123,2020\n2020-05-03,142,7,124,2020\n2020-05-04,132,1,125,2020\n2020-05-05,218,2,126,2020\n2020-05-06,229,3,127,2020\n2020-05-07,184,4,128,2020\n2020-05-08,160,5,129,2020\n2020-05-09,152,6,130,2020\n2020-05-10,160,7,131,2020\n2020-05-11,187,1,132,2020\n2020-05-12,158,2,133,2020\n2020-05-13,200,3,134,2020\n2020-05-14,212,4,135,2020\n2020-05-15,199,5,136,2020\n2020-05-16,113,6,137,2020\n2020-05-17,142,7,138,2020\n2020-05-18,161,1,139,2020\n2020-05-19,162,2,140,2020\n2020-05-20,186,3,141,2020\n2020-05-21,122,4,142,2020\n2020-05-22,142,5,143,2020\n2020-05-23,128,6,144,2020\n2020-05-24,134,7,145,2020\n2020-05-25,143,1,146,2020\n2020-05-26,220,2,147,2020\n2020-05-27,144,3,148,2020\n2020-05-28,100,4,149,2020\n2020-05-29,108,5,150,2020\n2020-05-30,213,6,151,2020\n2020-05-31,86,7,152,2020\n2020-06-01,115,1,153,2020\n2020-06-02,107,2,154,2020\n2020-06-03,169,3,155,2020\n2020-06-04,147,4,156,2020\n2020-06-05,115,5,157,2020\n2020-06-06,107,6,158,2020\n2020-06-07,93,7,159,2020\n2020-06-08,116,1,160,2020\n2020-06-09,143,2,161,2020\n2020-06-10,84,3,162,2020\n2020-06-11,177,4,163,2020\n2020-06-12,125,5,164,2020\n2020-06-13,61,6,165,2020\n2020-06-14,98,7,166,2020\n2020-06-15,191,1,167,2020\n2020-06-16,157,2,168,2020\n2020-06-17,199,3,169,2020\n2020-06-18,192,4,170,2020\n2020-06-19,146,5,171,2020\n2020-06-20,60,6,172,2020\n2020-06-21,116,7,173,2020\n2020-06-22,128,1,174,2020\n2020-06-23,130,2,175,2020\n2020-06-24,160,3,176,2020\n2020-06-25,120,4,177,2020\n2020-06-26,167,5,178,2020\n2020-06-27,164,6,179,2020\n2020-06-28,113,7,180,2020\n2020-06-29,167,1,181,2020\n2020-06-30,155,2,182,2020\n2020-07-01,126,3,183,2020\n2020-07-02,134,4,184,2020\n2020-07-03,146,5,185,2020\n2020-07-04,59,6,186,2020\n2020-07-05,102,7,187,2020\n2020-07-06,120,1,188,2020\n2020-07-07,124,2,189,2020\n2020-07-08,154,3,190,2020\n2020-07-09,164,4,191,2020\n2020-07-10,171,5,192,2020\n2020-07-11,79,6,193,2020\n2020-07-12,63,7,194,2020\n2020-07-13,85,1,195,2020\n2020-07-14,90,2,196,2020\n2020-07-15,129,3,197,2020\n2020-07-16,138,4,198,2020\n2020-07-17,95,5,199,2020\n2020-07-18,71,6,200,2020\n2020-07-19,99,7,201,2020\n2020-07-20,140,1,202,2020\n2020-07-21,189,2,203,2020\n2020-07-22,212,3,204,2020\n2020-07-23,176,4,205,2020\n2020-07-24,183,5,206,2020\n2020-07-25,55,6,207,2020\n2020-07-26,100,7,208,2020\n2020-07-27,137,1,209,2020\n2020-07-28,190,2,210,2020\n2020-07-29,118,3,211,2020\n2020-07-30,147,4,212,2020\n2020-07-31,139,5,213,2020\n2020-08-01,91,6,214,2020\n2020-08-02,99,7,215,2020\n2020-08-03,168,1,216,2020\n2020-08-04,168,2,217,2020\n2020-08-05,151,3,218,2020\n2020-08-06,167,4,219,2020\n2020-08-07,155,5,220,2020\n2020-08-08,78,6,221,2020\n2020-08-09,111,7,222,2020\n2020-08-10,126,1,223,2020\n2020-08-11,111,2,224,2020\n2020-08-12,121,3,225,2020\n2020-08-13,97,4,226,2020\n2020-08-14,137,5,227,2020\n2020-08-15,54,6,228,2020\n2020-08-16,56,7,229,2020\n2020-08-17,162,1,230,2020\n2020-08-18,98,2,231,2020\n2020-08-19,139,3,232,2020\n2020-08-20,128,4,233,2020\n2020-08-21,133,5,234,2020\n2020-08-22,93,6,235,2020\n2020-08-23,68,7,236,2020\n2020-08-24,149,1,237,2020\n2020-08-25,165,2,238,2020\n2020-08-26,201,3,239,2020\n2020-08-27,135,4,240,2020\n2020-08-28,108,5,241,2020\n2020-08-29,84,6,242,2020\n2020-08-30,42,7,243,2020\n2020-08-31,100,1,244,2020\n2020-09-01,134,2,245,2020\n2020-09-02,120,3,246,2020\n2020-09-03,100,4,247,2020\n2020-09-04,110,5,248,2020\n2020-09-05,53,6,249,2020\n2020-09-06,62,7,250,2020\n2020-09-07,129,1,251,2020\n2020-09-08,153,2,252,2020\n2020-09-09,148,3,253,2020\n2020-09-10,141,4,254,2020\n2020-09-11,162,5,255,2020\n2020-09-12,78,6,256,2020\n2020-09-13,58,7,257,2020\n2020-09-14,133,1,258,2020\n2020-09-15,126,2,259,2020\n2020-09-16,134,3,260,2020\n2020-09-17,94,4,261,2020\n2020-09-18,102,5,262,2020\n2020-09-19,65,6,263,2020\n2020-09-20,86,7,264,2020\n2020-09-21,136,1,265,2020\n2020-09-22,144,2,266,2020\n2020-09-23,137,3,267,2020\n2020-09-24,146,4,268,2020\n2020-09-25,148,5,269,2020\n2020-09-26,78,6,270,2020\n2020-09-27,82,7,271,2020\n2020-09-28,158,1,272,2020\n2020-09-29,122,2,273,2020\n2020-09-30,108,3,274,2020\n2020-10-01,146,4,275,2020\n2020-10-02,125,5,276,2020\n2020-10-03,55,6,277,2020\n2020-10-04,62,7,278,2020\n2020-10-05,99,1,279,2020\n2020-10-06,95,2,280,2020\n2020-10-07,137,3,281,2020\n2020-10-08,125,4,282,2020\n2020-10-09,117,5,283,2020\n2020-10-10,67,6,284,2020\n2020-10-11,62,7,285,2020\n2020-10-12,133,1,286,2020\n2020-10-13,80,2,287,2020\n2020-10-14,115,3,288,2020\n2020-10-15,104,4,289,2020\n2020-10-16,122,5,290,2020\n2020-10-17,106,6,291,2020\n2020-10-18,86,7,292,2020\n2020-10-19,156,1,293,2020\n2020-10-20,156,2,294,2020\n2020-10-21,162,3,295,2020\n2020-10-22,180,4,296,2020\n2020-10-23,93,5,297,2020\n2020-10-24,81,6,298,2020\n2020-10-25,75,7,299,2020\n2020-10-26,125,1,300,2020\n2020-10-27,130,2,301,2020\n2020-10-28,147,3,302,2020\n2020-10-29,118,4,303,2020\n2020-10-30,99,5,304,2020\n2020-10-31,93,6,305,2020\n2020-11-01,103,7,306,2020\n2020-11-02,103,1,307,2020\n2020-11-03,121,2,308,2020\n2020-11-04,86,3,309,2020\n2020-11-05,153,4,310,2020\n2020-11-06,129,5,311,2020\n2020-11-07,116,6,312,2020\n2020-11-08,51,7,313,2020\n2020-11-09,129,1,314,2020\n2020-11-10,150,2,315,2020\n2020-11-11,112,3,316,2020\n2020-11-12,117,4,317,2020\n2020-11-13,161,5,318,2020\n2020-11-14,69,6,319,2020\n2020-11-15,95,7,320,2020\n2020-11-16,144,1,321,2020\n2020-11-17,155,2,322,2020\n2020-11-18,142,3,323,2020\n2020-11-19,119,4,324,2020\n2020-11-20,113,5,325,2020\n2020-11-21,77,6,326,2020\n2020-11-22,69,7,327,2020\n2020-11-23,97,1,328,2020\n2020-11-24,107,2,329,2020\n2020-11-25,122,3,330,2020\n2020-11-26,131,4,331,2020\n2020-11-27,120,5,332,2020\n2020-11-28,93,6,333,2020\n2020-11-29,116,7,334,2020\n2020-11-30,139,1,335,2020\n2020-12-01,89,2,336,2020\n2020-12-02,152,3,337,2020\n2020-12-03,130,4,338,2020\n2020-12-04,128,5,339,2020\n2020-12-05,112,6,340,2020\n2020-12-06,69,7,341,2020\n2020-12-07,153,1,342,2020\n2020-12-08,110,2,343,2020\n2020-12-09,122,3,344,2020\n2020-12-10,124,4,345,2020\n2020-12-11,107,5,346,2020\n2020-12-12,78,6,347,2020\n2020-12-13,90,7,348,2020\n2020-12-14,87,1,349,2020\n2020-12-15,125,2,350,2020\n2020-12-16,143,3,351,2020\n2020-12-17,114,4,352,2020\n2020-12-18,96,5,353,2020\n2020-12-19,55,6,354,2020\n2020-12-20,66,7,355,2020\n2020-12-21,150,1,356,2020\n2020-12-22,129,2,357,2020\n2020-12-23,134,3,358,2020\n2020-12-24,70,4,359,2020\n2020-12-25,38,5,360,2020\n2020-12-26,68,6,361,2020\n2020-12-27,93,7,362,2020\n2020-12-28,116,1,363,2020\n2020-12-29,151,2,364,2020\n2020-12-30,114,3,365,2020\n2020-12-31,53,4,366,2020\n","format":{"type":"csv"}}});</script></div>

<div class="sourceClojure">
```clojure
(defn plot-average-by-temporal-field [tf template]
  (-> date-counts
      (tc/add-column tf (fn [ds]
                          (->> ds
                               :date
                               (datetime/long-temporal-field tf))))
      (tc/group-by [tf])
      (tc/aggregate {:n (fn [ds]
                          (-> ds
                              :n
                              fun/mean))})
      (tc/order-by [tf])
      (vis/hanami-plot template
                       {:X tf
                        :Y "n"})))
```
</div>

<div class="sourceClojure">
```clojure
(plot-average-by-temporal-field :day-of-week
                                ht/bar-chart)
```
</div>

<div><script>vegaEmbed(document.currentScript.parentElement, {"encoding":{"y":{"field":"n","type":"quantitative"},"x":{"field":"day-of-week","type":"quantitative"}},"mark":{"type":"bar","tooltip":true},"width":400,"background":"floralwhite","height":300,"data":{"values":"day-of-week,n\n1,113.39775561097257\n2,117.70931677018633\n3,119.55569306930693\n4,114.44389642416769\n5,112.25990099009901\n6,75.59062885326757\n7,77.97775030902349\n","format":{"type":"csv"}}});</script></div>

<div class="sourceClojure">
```clojure
(plot-average-by-temporal-field :day-of-year
                                ht/line-chart)
```
</div>

<div><script>vegaEmbed(document.currentScript.parentElement, {"encoding":{"y":{"field":"n","type":"quantitative"},"x":{"field":"day-of-year","type":"quantitative"}},"mark":{"type":"line","tooltip":true},"width":400,"background":"floralwhite","height":300,"data":{"values":"day-of-year,n\n1,58.8\n2,92.2\n3,95.53333333333333\n4,112.8\n5,96.13333333333334\n6,104.06666666666666\n7,112.86666666666666\n8,124.06666666666666\n9,119.6\n10,135.4\n11,129.4\n12,118.13333333333334\n13,136.2\n14,112.8\n15,119.33333333333333\n16,121.06666666666666\n17,125.46666666666667\n18,112.33333333333333\n19,126.0\n20,111.46666666666667\n21,109.06666666666666\n22,117.53333333333333\n23,123.6\n24,115.0\n25,112.46666666666667\n26,113.93333333333334\n27,107.6\n28,104.93333333333334\n29,112.26666666666667\n30,113.2\n31,116.46666666666667\n32,109.26666666666667\n33,120.33333333333333\n34,105.4\n35,104.66666666666667\n36,116.46666666666667\n37,120.86666666666666\n38,107.8\n39,116.46666666666667\n40,127.78571428571429\n41,121.6\n42,113.2\n43,111.26666666666667\n44,111.86666666666666\n45,106.33333333333333\n46,117.26666666666667\n47,110.33333333333333\n48,119.26666666666667\n49,128.26666666666668\n50,117.6\n51,114.4\n52,130.93333333333334\n53,122.0\n54,119.86666666666666\n55,125.46666666666667\n56,127.53333333333333\n57,110.93333333333334\n58,109.33333333333333\n59,120.2\n60,105.26666666666667\n61,103.93333333333334\n62,113.86666666666666\n63,104.33333333333333\n64,110.46666666666667\n65,113.2\n66,101.26666666666667\n67,103.53333333333333\n68,118.53333333333333\n69,118.53333333333333\n70,105.53333333333333\n71,101.73333333333333\n72,99.46666666666667\n73,110.13333333333334\n74,102.86666666666666\n75,107.53333333333333\n76,115.73333333333333\n77,104.33333333333333\n78,106.0\n79,104.66666666666667\n80,113.06666666666666\n81,113.26666666666667\n82,114.26666666666667\n83,124.33333333333333\n84,111.13333333333334\n85,124.6\n86,112.26666666666667\n87,109.66666666666667\n88,113.8\n89,108.1875\n90,111.4\n91,109.93333333333334\n92,114.53333333333333\n93,107.9375\n94,104.25\n95,117.86666666666666\n96,104.8125\n97,105.125\n98,111.875\n99,107.5\n100,117.2\n101,96.0625\n102,117.875\n103,111.4\n104,114.0\n105,114.33333333333333\n106,105.13333333333334\n107,101.3125\n108,106.26666666666667\n109,109.3125\n110,95.4375\n111,119.8\n112,99.8125\n113,102.33333333333333\n114,116.53333333333333\n115,101.1875\n116,107.4375\n117,106.13333333333334\n118,116.73333333333333\n119,113.6\n120,102.875\n121,104.5625\n122,106.5\n123,103.9375\n124,98.25\n125,109.93333333333334\n126,107.33333333333333\n127,111.93333333333334\n128,117.875\n129,99.9375\n130,116.2\n131,109.125\n132,106.4375\n133,95.3125\n134,109.3125\n135,96.1875\n136,98.1875\n137,106.3125\n138,106.0\n139,102.875\n140,107.53333333333333\n141,111.46666666666667\n142,109.33333333333333\n143,100.9375\n144,106.0\n145,96.75\n146,96.125\n147,98.0\n148,95.375\n149,85.1875\n150,96.8125\n151,112.86666666666666\n152,110.625\n153,105.875\n154,92.0625\n155,97.9375\n156,95.25\n157,92.8125\n158,103.3125\n159,103.875\n160,108.875\n161,107.4375\n162,95.0\n163,111.5625\n164,96.875\n165,111.125\n166,95.125\n167,116.3125\n168,120.33333333333333\n169,103.93333333333334\n170,99.5\n171,96.5625\n172,101.73333333333333\n173,110.0\n174,108.06666666666666\n175,103.53333333333333\n176,98.06666666666666\n177,102.73333333333333\n178,105.0\n179,105.9375\n180,94.1875\n181,104.5\n182,108.75\n183,104.2\n184,93.53333333333333\n185,80.1875\n186,88.0625\n187,104.26666666666667\n188,104.0\n189,95.125\n190,89.4375\n191,97.8125\n192,105.25\n193,98.875\n194,103.125\n195,97.25\n196,100.0625\n197,109.4375\n198,102.66666666666667\n199,115.06666666666666\n200,103.73333333333333\n201,117.26666666666667\n202,109.6\n203,109.26666666666667\n204,101.5\n205,110.8\n206,93.5625\n207,96.25\n208,99.6875\n209,105.0625\n210,111.6\n211,100.1875\n212,99.46666666666667\n213,115.2\n214,103.26666666666667\n215,107.53333333333333\n216,102.06666666666666\n217,118.4\n218,100.625\n219,102.0\n220,94.3125\n221,92.1875\n222,98.9375\n223,105.3125\n224,97.6875\n225,95.125\n226,94.5625\n227,89.125\n228,102.9375\n229,91.75\n230,98.4375\n231,105.53333333333333\n232,90.8125\n233,98.0\n234,90.75\n235,108.93333333333334\n236,107.6\n237,109.8\n238,104.875\n239,105.46666666666667\n240,101.625\n241,87.625\n242,97.9375\n243,99.6875\n244,112.33333333333333\n245,108.0\n246,90.3125\n247,87.75\n248,95.75\n249,85.6875\n250,107.4375\n251,99.0\n252,108.125\n253,98.125\n254,91.8125\n255,89.0\n256,94.375\n257,92.625\n258,100.4375\n259,96.5\n260,93.6875\n261,94.1875\n262,100.5625\n263,100.25\n264,99.75\n265,89.75\n266,99.4375\n267,109.33333333333333\n268,96.6875\n269,99.4375\n270,107.625\n271,116.3125\n272,113.25\n273,112.13333333333334\n274,112.8\n275,105.125\n276,107.125\n277,94.5\n278,96.125\n279,110.4375\n280,97.5625\n281,102.5\n282,111.6875\n283,98.0\n284,103.25\n285,100.46666666666667\n286,108.8125\n287,94.1875\n288,98.1875\n289,99.875\n290,96.625\n291,105.375\n292,106.9375\n293,103.06666666666666\n294,108.8\n295,86.3125\n296,89.875\n297,95.3125\n298,102.06666666666666\n299,88.25\n300,97.8125\n301,91.375\n302,98.75\n303,104.75\n304,102.125\n305,99.5625\n306,107.375\n307,101.1875\n308,111.875\n309,109.73333333333333\n310,112.06666666666666\n311,102.4\n312,112.8\n313,101.1875\n314,107.25\n315,122.13333333333334\n316,91.0625\n317,103.6875\n318,106.5625\n319,100.625\n320,111.0625\n321,112.6875\n322,104.0\n323,106.3125\n324,103.8125\n325,100.0\n326,99.25\n327,91.5625\n328,95.3125\n329,91.25\n330,80.9375\n331,89.25\n332,89.5\n333,97.75\n334,108.75\n335,104.75\n336,95.0625\n337,93.3125\n338,91.46666666666667\n339,107.93333333333334\n340,125.6\n341,110.64285714285714\n342,105.8\n343,105.06666666666666\n344,104.2\n345,98.0\n346,106.86666666666666\n347,105.53333333333333\n348,103.8\n349,108.8\n350,107.6\n351,105.26666666666667\n352,98.8\n353,110.78571428571429\n354,103.86666666666666\n355,99.2\n356,97.06666666666666\n357,80.0\n358,66.13333333333334\n359,40.46666666666667\n360,51.0\n361,82.26666666666667\n362,83.4\n363,84.2\n364,101.8\n365,77.26666666666667\n366,47.0\n","format":{"type":"csv"}}});</script></div>

<div class="sourceClojure">
```clojure
(-> processed-date-counts
    (tc/add-column :smoothed-n
                   (fn [ds]
                     (-> ds
                         :n
                         (rolling/fixed-rolling-window
                          30
                          fun/mean
                          {:relative-window-position :left}))))
    (vis/hanami-layers {}
                       [(vis/hanami-plot nil
                                         ht/line-chart
                                         {:X "date"
                                          :XTYPE "temporal"
                                          :Y "n"})
                        (vis/hanami-plot nil
                                         ht/line-chart
                                         {:X "date"
                                          :XTYPE "temporal"
                                          :Y "smoothed-n"
                                          :MCOLOR "brown"})]))
```
</div>

<div><script>vegaEmbed(document.currentScript.parentElement, {"width":400,"background":"floralwhite","layer":[{"encoding":{"y":{"field":"n","type":"quantitative"},"x":{"field":"date","type":"temporal"}},"mark":{"type":"line","tooltip":true},"width":400,"background":"floralwhite","height":300},{"encoding":{"y":{"field":"smoothed-n","type":"quantitative"},"x":{"field":"date","type":"temporal"}},"mark":{"type":"line","color":"brown","tooltip":true},"width":400,"background":"floralwhite","height":300}],"height":300,"data":{"values":"date,n,day-of-week,day-of-year,years,smoothed-n\n2008-03-29,7,6,89,2008,7.0\n2008-04-02,4,3,93,2008,6.9\n2008-04-03,9,4,94,2008,6.966666666666667\n2008-04-05,2,6,96,2008,6.8\n2008-04-06,1,7,97,2008,6.6\n2008-04-07,5,1,98,2008,6.533333333333333\n2008-04-08,9,2,99,2008,6.6\n2008-04-10,1,4,101,2008,6.4\n2008-04-11,1,5,102,2008,6.2\n2008-04-16,1,3,107,2008,6.0\n2008-04-18,1,5,109,2008,5.8\n2008-04-19,1,6,110,2008,5.6\n2008-04-21,1,1,112,2008,5.4\n2008-04-24,3,4,115,2008,5.266666666666667\n2008-04-25,1,5,116,2008,5.066666666666666\n2008-04-29,1,2,120,2008,4.866666666666666\n2008-04-30,4,3,121,2008,4.766666666666667\n2008-05-01,1,4,122,2008,4.566666666666666\n2008-05-02,2,5,123,2008,4.4\n2008-05-03,2,6,124,2008,4.233333333333333\n2008-05-07,3,3,128,2008,4.1\n2008-05-08,2,4,129,2008,3.933333333333333\n2008-05-10,6,6,131,2008,3.9\n2008-05-11,1,7,132,2008,3.7\n2008-05-12,1,1,133,2008,3.5\n2008-05-13,2,2,134,2008,3.3333333333333335\n2008-05-14,1,3,135,2008,3.1333333333333333\n2008-05-15,1,4,136,2008,2.933333333333333\n2008-05-16,2,5,137,2008,2.7666666666666666\n2008-05-17,9,6,138,2008,2.8333333333333335\n2008-05-18,12,7,139,2008,3.0\n2008-05-22,1,4,143,2008,2.9\n2008-05-24,3,6,145,2008,2.7\n2008-05-25,2,7,146,2008,2.7\n2008-05-26,1,1,147,2008,2.7\n2008-05-27,1,2,148,2008,2.566666666666667\n2008-05-28,1,3,149,2008,2.3\n2008-05-29,4,4,150,2008,2.4\n2008-05-31,1,6,152,2008,2.4\n2008-06-01,2,7,153,2008,2.433333333333333\n2008-06-02,1,1,154,2008,2.433333333333333\n2008-06-03,3,2,155,2008,2.5\n2008-06-04,2,3,156,2008,2.533333333333333\n2008-06-05,1,4,157,2008,2.466666666666667\n2008-06-06,1,5,158,2008,2.466666666666667\n2008-06-07,1,6,159,2008,2.466666666666667\n2008-06-08,5,7,160,2008,2.5\n2008-06-09,3,1,161,2008,2.566666666666667\n2008-06-11,2,3,163,2008,2.566666666666667\n2008-06-12,2,4,164,2008,2.566666666666667\n2008-06-13,2,5,165,2008,2.533333333333333\n2008-06-14,4,6,166,2008,2.6\n2008-06-15,5,7,167,2008,2.566666666666667\n2008-06-18,1,3,170,2008,2.566666666666667\n2008-06-19,3,4,171,2008,2.6333333333333333\n2008-06-27,1,5,179,2008,2.6\n2008-06-28,6,6,180,2008,2.7666666666666666\n2008-06-29,6,7,181,2008,2.933333333333333\n2008-06-30,2,1,182,2008,2.933333333333333\n2008-07-03,2,4,185,2008,2.7\n2008-07-04,1,5,186,2008,2.3333333333333335\n2008-07-06,5,7,188,2008,2.466666666666667\n2008-07-07,2,1,189,2008,2.433333333333333\n2008-07-08,1,2,190,2008,2.4\n2008-07-09,8,3,191,2008,2.6333333333333333\n2008-07-10,5,4,192,2008,2.7666666666666666\n2008-07-11,6,5,193,2008,2.933333333333333\n2008-07-12,4,6,194,2008,2.933333333333333\n2008-07-13,3,7,195,2008,3.0\n2008-07-14,1,1,196,2008,2.966666666666667\n2008-07-15,5,2,197,2008,3.1\n2008-07-21,1,1,203,2008,3.033333333333333\n2008-07-22,5,2,204,2008,3.1333333333333333\n2008-07-24,2,4,206,2008,3.1666666666666665\n2008-07-25,2,5,207,2008,3.2\n2008-07-26,7,6,208,2008,3.4\n2008-07-27,3,7,209,2008,3.3333333333333335\n2008-07-29,8,2,211,2008,3.5\n2008-08-05,1,2,218,2008,3.466666666666667\n2008-08-06,4,3,219,2008,3.533333333333333\n2008-08-07,7,4,220,2008,3.7\n2008-08-08,1,5,221,2008,3.6\n2008-08-09,1,6,222,2008,3.466666666666667\n2008-08-10,1,7,223,2008,3.466666666666667\n2008-08-11,1,1,224,2008,3.4\n2008-08-12,4,2,225,2008,3.5\n2008-08-13,1,3,226,2008,3.3333333333333335\n2008-08-14,1,4,227,2008,3.1666666666666665\n2008-08-15,3,5,228,2008,3.2\n2008-08-16,9,6,229,2008,3.433333333333333\n2008-08-17,4,7,230,2008,3.533333333333333\n2008-08-19,7,2,232,2008,3.6\n2008-08-20,3,3,233,2008,3.6333333333333333\n2008-08-21,1,4,234,2008,3.6333333333333333\n2008-08-25,2,1,238,2008,3.433333333333333\n2008-08-27,2,3,240,2008,3.3333333333333335\n2008-08-28,2,4,241,2008,3.2\n2008-08-29,5,5,242,2008,3.2333333333333334\n2008-08-30,8,6,243,2008,3.4\n2008-09-02,8,2,246,2008,3.6333333333333333\n2008-09-03,7,3,247,2008,3.7\n2008-09-04,1,4,248,2008,3.7\n2008-09-05,1,5,249,2008,3.566666666666667\n2008-09-06,2,6,250,2008,3.566666666666667\n2008-09-07,3,7,251,2008,3.6\n2008-09-08,29,1,252,2008,4.333333333333333\n2008-09-09,4,2,253,2008,4.366666666666666\n2008-09-10,3,3,254,2008,4.2\n2008-09-11,2,4,255,2008,4.233333333333333\n2008-09-12,3,5,256,2008,4.2\n2008-09-13,11,6,257,2008,4.333333333333333\n2008-09-14,2,7,258,2008,4.366666666666666\n2008-09-15,8,1,259,2008,4.6\n2008-09-16,11,2,260,2008,4.933333333333334\n2008-09-17,8,3,261,2008,5.166666666666667\n2008-09-18,1,4,262,2008,5.066666666666666\n2008-09-19,2,5,263,2008,5.1\n2008-09-20,8,6,264,2008,5.333333333333333\n2008-09-21,7,7,265,2008,5.466666666666667\n2008-09-22,3,1,266,2008,5.266666666666667\n2008-09-23,4,2,267,2008,5.266666666666667\n2008-09-24,4,3,268,2008,5.166666666666667\n2008-09-25,6,4,269,2008,5.266666666666667\n2008-09-26,4,5,270,2008,5.366666666666666\n2008-09-27,2,6,271,2008,5.366666666666666\n2008-09-28,7,7,272,2008,5.533333333333333\n2008-10-01,6,3,275,2008,5.666666666666667\n2008-10-02,10,4,276,2008,5.833333333333333\n2008-10-03,5,5,277,2008,5.733333333333333\n2008-10-04,3,6,278,2008,5.566666666666666\n2008-10-05,5,7,279,2008,5.5\n2008-10-06,11,1,280,2008,5.833333333333333\n2008-10-07,2,2,281,2008,5.866666666666666\n2008-10-08,6,3,282,2008,6.0\n2008-10-09,3,4,283,2008,6.0\n2008-10-10,4,5,284,2008,5.166666666666667\n2008-10-11,2,6,285,2008,5.1\n2008-10-12,4,7,286,2008,5.133333333333334\n2008-10-13,2,1,287,2008,5.133333333333334\n2008-10-14,1,2,288,2008,5.066666666666666\n2008-10-15,10,3,289,2008,5.033333333333333\n2008-10-16,1,4,290,2008,5.0\n2008-10-17,4,5,291,2008,4.866666666666666\n2008-10-18,2,6,292,2008,4.566666666666666\n2008-10-19,8,7,293,2008,4.566666666666666\n2008-10-21,1,2,295,2008,4.566666666666666\n2008-10-22,5,3,296,2008,4.666666666666667\n2008-10-23,4,4,297,2008,4.533333333333333\n2008-10-25,2,6,299,2008,4.366666666666666\n2008-10-26,3,7,300,2008,4.366666666666666\n2008-10-27,1,1,301,2008,4.266666666666667\n2008-10-28,2,2,302,2008,4.2\n2008-10-29,1,3,303,2008,4.033333333333333\n2008-10-30,5,4,304,2008,4.066666666666666\n2008-10-31,5,5,305,2008,4.166666666666667\n2008-11-01,9,6,306,2008,4.233333333333333\n2008-11-02,1,7,307,2008,4.066666666666666\n2008-11-03,7,1,308,2008,3.966666666666667\n2008-11-08,3,6,313,2008,3.9\n2008-11-09,5,7,314,2008,3.966666666666667\n2008-11-11,1,2,316,2008,3.8333333333333335\n2008-11-12,5,3,317,2008,3.6333333333333333\n2008-11-13,23,4,318,2008,4.333333333333333\n2008-11-14,2,5,319,2008,4.2\n2008-11-15,8,6,320,2008,4.366666666666666\n2008-11-16,17,7,321,2008,4.8\n2008-11-17,3,1,322,2008,4.833333333333333\n2008-11-18,25,2,323,2008,5.533333333333333\n2008-11-19,3,3,324,2008,5.566666666666666\n2008-11-20,2,4,325,2008,5.6\n2008-11-21,3,5,326,2008,5.366666666666666\n2008-11-22,3,6,327,2008,5.433333333333334\n2008-11-23,7,7,328,2008,5.533333333333333\n2008-11-24,1,1,329,2008,5.5\n2008-11-25,6,2,330,2008,5.433333333333334\n2008-11-26,3,3,331,2008,5.5\n2008-11-27,3,4,332,2008,5.433333333333334\n2008-11-28,5,5,333,2008,5.466666666666667\n2008-11-29,5,6,334,2008,5.566666666666666\n2008-11-30,5,7,335,2008,5.633333333333334\n2008-12-01,7,1,336,2008,5.833333333333333\n2008-12-02,3,2,337,2008,5.866666666666666\n2008-12-03,7,3,338,2008,6.066666666666666\n2008-12-04,5,4,339,2008,6.066666666666666\n2008-12-05,6,5,340,2008,6.1\n2008-12-07,3,7,342,2008,5.9\n2008-12-08,4,1,343,2008,6.0\n2008-12-09,1,2,344,2008,5.8\n2008-12-10,6,3,345,2008,5.9\n2008-12-11,5,4,346,2008,5.9\n2008-12-12,1,5,347,2008,5.9\n2008-12-13,12,6,348,2008,6.133333333333334\n2008-12-14,8,7,349,2008,5.633333333333334\n2008-12-15,9,1,350,2008,5.866666666666666\n2008-12-16,6,2,351,2008,5.8\n2008-12-17,6,3,352,2008,5.433333333333334\n2008-12-19,18,5,354,2008,5.933333333333334\n2008-12-20,1,6,355,2008,5.133333333333334\n2008-12-21,3,7,356,2008,5.133333333333334\n2008-12-22,3,1,357,2008,5.166666666666667\n2008-12-23,10,2,358,2008,5.4\n2008-12-24,3,3,359,2008,5.4\n2008-12-25,1,4,360,2008,5.2\n2008-12-26,1,5,361,2008,5.2\n2008-12-27,13,6,362,2008,5.433333333333334\n2008-12-28,13,7,363,2008,5.766666666666667\n2008-12-29,15,1,364,2008,6.166666666666667\n2008-12-30,10,2,365,2008,6.333333333333333\n2008-12-31,2,3,366,2008,6.233333333333333\n2009-01-01,4,4,1,2009,6.2\n2009-01-02,5,5,2,2009,6.133333333333334\n2009-01-03,2,6,3,2009,6.1\n2009-01-04,14,7,4,2009,6.333333333333333\n2009-01-05,3,1,5,2009,6.266666666666667\n2009-01-06,7,2,6,2009,6.3\n2009-01-07,1,3,7,2009,6.233333333333333\n2009-01-08,17,4,8,2009,6.666666666666667\n2009-01-09,4,5,9,2009,6.766666666666667\n2009-01-10,3,6,10,2009,6.666666666666667\n2009-01-11,9,7,11,2009,6.8\n2009-01-12,19,1,12,2009,7.4\n2009-01-13,5,2,13,2009,7.166666666666667\n2009-01-14,4,3,14,2009,7.033333333333333\n2009-01-15,11,4,15,2009,7.1\n2009-01-16,10,5,16,2009,7.233333333333333\n2009-01-17,36,6,17,2009,8.233333333333333\n2009-01-18,36,7,18,2009,8.833333333333334\n2009-01-19,22,1,19,2009,9.533333333333333\n2009-01-20,12,2,20,2009,9.833333333333334\n2009-01-21,16,3,21,2009,10.266666666666667\n2009-01-22,8,4,22,2009,10.2\n2009-01-23,13,5,23,2009,10.533333333333333\n2009-01-24,7,6,24,2009,10.733333333333333\n2009-01-25,15,7,25,2009,11.2\n2009-01-26,13,1,26,2009,11.2\n2009-01-27,6,2,27,2009,10.966666666666667\n2009-01-28,5,3,28,2009,10.633333333333333\n2009-01-29,7,4,29,2009,10.533333333333333\n2009-01-30,5,5,30,2009,10.633333333333333\n2009-01-31,2,6,31,2009,10.566666666666666\n2009-02-01,2,7,32,2009,10.466666666666667\n2009-02-02,4,1,33,2009,10.533333333333333\n2009-02-03,1,2,34,2009,10.1\n2009-02-04,3,3,35,2009,10.1\n2009-02-05,1,4,36,2009,9.9\n2009-02-06,8,5,37,2009,10.133333333333333\n2009-02-07,3,6,38,2009,9.666666666666666\n2009-02-08,3,7,39,2009,9.633333333333333\n2009-02-10,22,2,41,2009,10.266666666666667\n2009-02-11,14,3,42,2009,10.433333333333334\n2009-02-12,18,4,43,2009,10.4\n2009-02-13,6,5,44,2009,10.433333333333334\n2009-02-14,5,6,45,2009,10.466666666666667\n2009-02-15,9,7,46,2009,10.4\n2009-02-16,5,1,47,2009,10.233333333333333\n2009-02-17,8,2,48,2009,9.3\n2009-02-18,7,3,49,2009,8.333333333333334\n2009-02-19,4,4,50,2009,7.733333333333333\n2009-02-20,8,5,51,2009,7.6\n2009-02-21,10,6,52,2009,7.4\n2009-02-22,15,7,53,2009,7.633333333333334\n2009-02-23,12,1,54,2009,7.6\n2009-02-24,27,2,55,2009,8.266666666666667\n2009-02-25,7,3,56,2009,8.0\n2009-02-26,11,4,57,2009,7.933333333333334\n2009-02-27,19,5,58,2009,8.366666666666667\n2009-02-28,3,6,59,2009,8.3\n2009-03-01,5,7,60,2009,8.233333333333333\n2009-03-02,3,1,61,2009,8.166666666666666\n2009-03-03,8,2,62,2009,8.366666666666667\n2009-03-04,7,3,63,2009,8.533333333333333\n2009-03-05,2,4,64,2009,8.466666666666667\n2009-03-06,6,5,65,2009,8.633333333333333\n2009-03-07,3,6,66,2009,8.633333333333333\n2009-03-08,2,7,67,2009,8.666666666666666\n2009-03-09,6,1,68,2009,8.6\n2009-03-10,4,2,69,2009,8.633333333333333\n2009-03-11,7,3,70,2009,8.766666666666667\n2009-03-12,13,4,71,2009,8.466666666666667\n2009-03-13,4,5,72,2009,8.133333333333333\n2009-03-14,8,6,73,2009,7.8\n2009-03-15,7,7,74,2009,7.833333333333333\n2009-03-16,9,1,75,2009,7.966666666666667\n2009-03-17,9,2,76,2009,7.966666666666667\n2009-03-18,11,3,77,2009,8.166666666666666\n2009-03-19,13,4,78,2009,8.333333333333334\n2009-03-20,8,5,79,2009,8.366666666666667\n2009-03-21,13,6,80,2009,8.666666666666666\n2009-03-22,11,7,81,2009,8.766666666666667\n2009-03-23,8,1,82,2009,8.7\n2009-03-24,4,2,83,2009,8.333333333333334\n2009-03-25,4,3,84,2009,8.066666666666666\n2009-03-26,4,4,85,2009,7.3\n2009-03-27,9,5,86,2009,7.366666666666666\n2009-03-28,10,6,87,2009,7.333333333333333\n2009-03-29,17,7,88,2009,7.266666666666667\n2009-03-30,9,1,89,2009,7.466666666666667\n2009-03-31,7,2,90,2009,7.533333333333333\n2009-04-01,8,3,91,2009,7.7\n2009-04-02,1,4,92,2009,7.466666666666667\n2009-04-03,5,5,93,2009,7.4\n2009-04-04,8,6,94,2009,7.6\n2009-04-05,16,7,95,2009,7.933333333333334\n2009-04-06,15,1,96,2009,8.333333333333334\n2009-04-07,5,2,97,2009,8.433333333333334\n2009-04-08,7,3,98,2009,8.466666666666667\n2009-04-09,15,4,99,2009,8.833333333333334\n2009-04-10,5,5,100,2009,8.766666666666667\n2009-04-11,2,6,101,2009,8.4\n2009-04-12,3,7,102,2009,8.366666666666667\n2009-04-13,1,1,103,2009,8.133333333333333\n2009-04-14,9,2,104,2009,8.2\n2009-04-15,17,3,105,2009,8.466666666666667\n2009-04-16,9,4,106,2009,8.466666666666667\n2009-04-17,19,5,107,2009,8.733333333333333\n2009-04-18,8,6,108,2009,8.566666666666666\n2009-04-19,21,7,109,2009,9.0\n2009-04-20,28,1,110,2009,9.5\n2009-04-21,24,2,111,2009,9.933333333333334\n2009-04-22,13,3,112,2009,10.1\n2009-04-23,25,4,113,2009,10.8\n2009-04-24,8,5,114,2009,10.933333333333334\n2009-04-25,8,6,115,2009,11.066666666666666\n2009-04-26,6,7,116,2009,10.966666666666667\n2009-04-27,4,1,117,2009,10.766666666666667\n2009-04-28,2,2,118,2009,10.266666666666667\n2009-04-29,6,3,119,2009,10.166666666666666\n2009-04-30,1,4,120,2009,9.966666666666667\n2009-05-01,8,5,121,2009,9.966666666666667\n2009-05-02,1,6,122,2009,9.966666666666667\n2009-05-03,12,7,123,2009,10.2\n2009-05-04,17,1,124,2009,10.5\n2009-05-05,7,2,125,2009,10.2\n2009-05-06,9,3,126,2009,10.0\n2009-05-07,13,4,127,2009,10.266666666666667\n2009-05-08,4,5,128,2009,10.166666666666666\n2009-05-09,15,6,129,2009,10.166666666666666\n2009-05-10,7,7,130,2009,10.233333333333333\n2009-05-11,13,1,131,2009,10.6\n2009-05-12,3,2,132,2009,10.6\n2009-05-13,2,3,133,2009,10.633333333333333\n2009-05-14,17,4,134,2009,10.9\n2009-05-15,3,5,135,2009,10.433333333333334\n2009-05-16,5,6,136,2009,10.3\n2009-05-17,5,7,137,2009,9.833333333333334\n2009-05-18,6,1,138,2009,9.766666666666667\n2009-05-19,6,2,139,2009,9.266666666666667\n2009-05-20,3,3,140,2009,8.433333333333334\n2009-05-21,5,4,141,2009,7.8\n2009-05-22,2,5,142,2009,7.433333333333334\n2009-05-23,16,6,143,2009,7.133333333333334\n2009-05-24,7,7,144,2009,7.1\n2009-05-25,7,1,145,2009,7.066666666666666\n2009-05-26,4,2,146,2009,7.0\n2009-05-27,7,3,147,2009,7.1\n2009-05-28,3,4,148,2009,7.133333333333334\n2009-05-29,7,5,149,2009,7.166666666666667\n2009-05-30,4,6,150,2009,7.266666666666667\n2009-05-31,19,7,151,2009,7.633333333333334\n2009-06-01,7,1,152,2009,7.833333333333333\n2009-06-02,10,2,153,2009,7.766666666666667\n2009-06-03,13,3,154,2009,7.633333333333334\n2009-06-04,7,4,155,2009,7.633333333333334\n2009-06-05,17,5,156,2009,7.9\n2009-06-06,8,6,157,2009,7.733333333333333\n2009-06-07,1,7,158,2009,7.633333333333334\n2009-06-08,11,1,159,2009,7.5\n2009-06-09,8,2,160,2009,7.533333333333333\n2009-06-10,4,3,161,2009,7.233333333333333\n2009-06-11,7,4,162,2009,7.366666666666666\n2009-06-12,4,5,163,2009,7.433333333333334\n2009-06-13,12,6,164,2009,7.266666666666667\n2009-06-14,4,7,165,2009,7.3\n2009-06-15,3,1,166,2009,7.233333333333333\n2009-06-16,8,2,167,2009,7.333333333333333\n2009-06-17,2,3,168,2009,7.2\n2009-06-18,2,4,169,2009,7.066666666666666\n2009-06-19,18,5,170,2009,7.566666666666666\n2009-06-20,1,6,171,2009,7.433333333333334\n2009-06-21,1,7,172,2009,7.4\n2009-06-22,9,1,173,2009,7.166666666666667\n2009-06-23,13,2,174,2009,7.366666666666666\n2009-06-24,14,3,175,2009,7.6\n2009-06-25,3,4,176,2009,7.566666666666666\n2009-06-26,17,5,177,2009,7.9\n2009-06-27,7,6,178,2009,8.033333333333333\n2009-06-28,10,7,179,2009,8.133333333333333\n2009-06-29,3,1,180,2009,8.1\n2009-06-30,6,2,181,2009,7.666666666666667\n2009-07-01,9,3,182,2009,7.733333333333333\n2009-07-02,17,4,183,2009,7.966666666666667\n2009-07-03,6,5,184,2009,7.733333333333333\n2009-07-04,4,6,185,2009,7.633333333333334\n2009-07-05,6,7,186,2009,7.266666666666667\n2009-07-06,9,1,187,2009,7.3\n2009-07-07,7,2,188,2009,7.5\n2009-07-08,13,3,189,2009,7.566666666666666\n2009-07-09,7,4,190,2009,7.533333333333333\n2009-07-10,6,5,191,2009,7.6\n2009-07-11,2,6,192,2009,7.433333333333334\n2009-07-12,8,7,193,2009,7.566666666666666\n2009-07-13,5,1,194,2009,7.333333333333333\n2009-07-14,7,2,195,2009,7.433333333333334\n2009-07-15,5,3,196,2009,7.5\n2009-07-16,6,4,197,2009,7.433333333333334\n2009-07-17,3,5,198,2009,7.466666666666667\n2009-07-18,5,6,199,2009,7.566666666666666\n2009-07-19,4,7,200,2009,7.1\n2009-07-20,14,1,201,2009,7.533333333333333\n2009-07-21,6,2,202,2009,7.7\n2009-07-23,10,4,204,2009,7.733333333333333\n2009-07-24,11,5,205,2009,7.666666666666667\n2009-07-25,3,6,206,2009,7.3\n2009-07-26,9,7,207,2009,7.5\n2009-07-27,5,1,208,2009,7.1\n2009-07-28,21,2,209,2009,7.566666666666666\n2009-07-29,17,3,210,2009,7.8\n2009-07-30,5,4,211,2009,7.866666666666666\n2009-07-31,6,5,212,2009,7.866666666666666\n2009-08-01,11,6,213,2009,7.933333333333334\n2009-08-02,4,7,214,2009,7.5\n2009-08-03,9,1,215,2009,7.6\n2009-08-04,5,2,216,2009,7.633333333333334\n2009-08-05,3,3,217,2009,7.533333333333333\n2009-08-06,1,4,218,2009,7.266666666666667\n2009-08-07,7,5,219,2009,7.266666666666667\n2009-08-08,7,6,220,2009,7.066666666666666\n2009-08-09,1,7,221,2009,6.866666666666666\n2009-08-10,4,1,222,2009,6.8\n2009-08-11,2,2,223,2009,6.8\n2009-08-12,5,3,224,2009,6.7\n2009-08-13,4,4,225,2009,6.666666666666667\n2009-08-14,9,5,226,2009,6.733333333333333\n2009-08-15,6,6,227,2009,6.766666666666667\n2009-08-16,2,7,228,2009,6.633333333333334\n2009-08-17,21,1,229,2009,7.233333333333333\n2009-08-18,14,2,230,2009,7.533333333333333\n2009-08-19,13,3,231,2009,7.833333333333333\n2009-08-20,6,4,232,2009,7.566666666666666\n2009-08-21,5,5,233,2009,7.533333333333333\n2009-08-22,1,6,234,2009,7.233333333333333\n2009-08-23,9,7,235,2009,7.166666666666667\n2009-08-24,4,1,236,2009,7.2\n2009-08-25,10,2,237,2009,7.233333333333333\n2009-08-26,2,3,238,2009,7.133333333333334\n2009-08-27,5,4,239,2009,6.6\n2009-08-28,7,5,240,2009,6.266666666666667\n2009-08-29,8,6,241,2009,6.366666666666666\n2009-08-30,3,7,242,2009,6.266666666666667\n2009-08-31,4,1,243,2009,6.033333333333333\n2009-09-01,8,2,244,2009,6.166666666666667\n2009-09-02,8,3,245,2009,6.133333333333334\n2009-09-03,19,4,246,2009,6.6\n2009-09-04,22,5,247,2009,7.233333333333333\n2009-09-05,4,6,248,2009,7.333333333333333\n2009-09-06,8,7,249,2009,7.366666666666666\n2009-09-07,15,1,250,2009,7.633333333333334\n2009-09-08,16,2,251,2009,8.133333333333333\n2009-09-09,14,3,252,2009,8.466666666666667\n2009-09-10,16,4,253,2009,8.933333333333334\n2009-09-11,17,5,254,2009,9.333333333333334\n2009-09-12,4,6,255,2009,9.333333333333334\n2009-09-13,11,7,256,2009,9.4\n2009-09-14,12,1,257,2009,9.6\n2009-09-15,12,2,258,2009,9.933333333333334\n2009-09-16,10,3,259,2009,9.566666666666666\n2009-09-17,9,4,260,2009,9.4\n2009-09-18,19,5,261,2009,9.6\n2009-09-19,9,6,262,2009,9.7\n2009-09-20,4,7,263,2009,9.666666666666666\n2009-09-21,4,1,264,2009,9.766666666666667\n2009-09-22,6,2,265,2009,9.666666666666666\n2009-09-23,21,3,266,2009,10.233333333333333\n2009-09-25,2,5,268,2009,9.966666666666667\n2009-09-26,4,6,269,2009,10.033333333333333\n2009-09-27,2,7,270,2009,9.933333333333334\n2009-09-28,8,1,271,2009,9.966666666666667\n2009-09-29,5,2,272,2009,9.866666666666667\n2009-09-30,4,3,273,2009,9.9\n2009-10-01,6,4,274,2009,9.966666666666667\n2009-10-02,2,5,275,2009,9.766666666666667\n2009-10-03,12,6,276,2009,9.9\n2009-10-04,4,7,277,2009,9.4\n2009-10-05,10,1,278,2009,9.0\n2009-10-06,12,2,279,2009,9.266666666666667\n2009-10-07,30,3,280,2009,10.0\n2009-10-08,8,4,281,2009,9.766666666666667\n2009-10-09,9,5,282,2009,9.533333333333333\n2009-10-10,2,6,283,2009,9.133333333333333\n2009-10-11,1,7,284,2009,8.633333333333333\n2009-10-13,2,2,286,2009,8.133333333333333\n2009-10-14,5,3,287,2009,8.166666666666666\n2009-10-15,4,4,288,2009,7.933333333333334\n2009-10-16,8,5,289,2009,7.8\n2009-10-17,3,6,290,2009,7.5\n2009-10-18,5,7,291,2009,7.333333333333333\n2009-10-19,5,1,292,2009,7.2\n2009-10-21,4,3,294,2009,6.7\n2009-10-22,7,4,295,2009,6.633333333333334\n2009-10-23,15,5,296,2009,7.0\n2009-10-24,29,6,297,2009,7.833333333333333\n2009-10-25,7,7,298,2009,7.866666666666666\n2009-10-26,20,1,299,2009,7.833333333333333\n2009-10-27,9,2,300,2009,8.066666666666666\n2009-10-28,8,3,301,2009,8.2\n2009-10-29,5,4,302,2009,8.3\n2009-10-30,2,5,303,2009,8.1\n2009-10-31,10,6,304,2009,8.266666666666667\n2009-11-01,17,7,305,2009,8.7\n2009-11-02,24,1,306,2009,9.3\n2009-11-03,7,2,307,2009,9.466666666666667\n2009-11-04,9,3,308,2009,9.366666666666667\n2009-11-05,8,4,309,2009,9.5\n2009-11-06,10,5,310,2009,9.5\n2009-11-07,2,6,311,2009,9.166666666666666\n2009-11-08,16,7,312,2009,8.7\n2009-11-09,14,1,313,2009,8.9\n2009-11-10,13,2,314,2009,9.033333333333333\n2009-11-11,19,3,315,2009,9.6\n2009-11-12,12,4,316,2009,9.966666666666667\n2009-11-13,11,5,317,2009,10.266666666666667\n2009-11-14,33,6,318,2009,11.2\n2009-11-15,18,7,319,2009,11.666666666666666\n2009-11-16,13,1,320,2009,11.833333333333334\n2009-11-17,15,2,321,2009,12.233333333333333\n2009-11-18,23,3,322,2009,12.833333333333334\n2009-11-19,23,4,323,2009,13.433333333333334\n2009-11-20,20,5,324,2009,13.966666666666667\n2009-11-21,23,6,325,2009,14.5\n2009-11-22,18,7,326,2009,14.6\n2009-11-23,22,1,327,2009,14.366666666666667\n2009-11-24,12,2,328,2009,14.533333333333333\n2009-11-25,14,3,329,2009,14.333333333333334\n2009-11-26,17,4,330,2009,14.6\n2009-11-27,18,5,331,2009,14.933333333333334\n2009-11-28,26,6,332,2009,15.633333333333333\n2009-11-29,14,7,333,2009,16.033333333333335\n2009-11-30,16,1,334,2009,16.233333333333334\n2009-12-01,27,2,335,2009,16.566666666666666\n2009-12-02,22,3,336,2009,16.5\n2009-12-03,9,4,337,2009,16.566666666666666\n2009-12-04,8,5,338,2009,16.533333333333335\n2009-12-05,10,6,339,2009,16.6\n2009-12-06,10,7,340,2009,16.6\n2009-12-07,8,1,341,2009,16.8\n2009-12-08,9,2,342,2009,16.566666666666666\n2009-12-09,9,3,343,2009,16.4\n2009-12-10,22,4,344,2009,16.7\n2009-12-11,10,5,345,2009,16.4\n2009-12-12,9,6,346,2009,16.3\n2009-12-13,6,7,347,2009,16.133333333333333\n2009-12-14,8,1,348,2009,15.3\n2009-12-15,8,2,349,2009,14.966666666666667\n2009-12-16,14,3,350,2009,15.0\n2009-12-17,5,4,351,2009,14.666666666666666\n2009-12-18,15,5,352,2009,14.4\n2009-12-19,19,6,353,2009,14.266666666666667\n2009-12-20,12,7,354,2009,14.0\n2009-12-21,16,1,355,2009,13.766666666666667\n2009-12-22,12,2,356,2009,13.566666666666666\n2009-12-23,12,3,357,2009,13.233333333333333\n2009-12-24,13,4,358,2009,13.266666666666667\n2009-12-25,11,5,359,2009,13.166666666666666\n2009-12-26,32,6,360,2009,13.666666666666666\n2009-12-27,25,7,361,2009,13.9\n2009-12-28,16,1,362,2009,13.566666666666666\n2009-12-29,21,2,363,2009,13.8\n2009-12-30,27,3,364,2009,14.166666666666666\n2009-12-31,13,4,365,2009,13.7\n2010-01-01,14,5,1,2010,13.433333333333334\n2010-01-02,14,6,2,2010,13.6\n2010-01-03,12,7,3,2010,13.733333333333333\n2010-01-04,28,1,4,2010,14.333333333333334\n2010-01-05,7,2,5,2010,14.233333333333333\n2010-01-06,4,3,6,2010,14.1\n2010-01-07,5,4,7,2010,13.966666666666667\n2010-01-08,18,5,8,2010,14.266666666666667\n2010-01-09,10,6,9,2010,13.866666666666667\n2010-01-10,20,7,10,2010,14.2\n2010-01-11,39,1,11,2010,15.2\n2010-01-12,31,2,12,2010,16.033333333333335\n2010-01-13,21,3,13,2010,16.466666666666665\n2010-01-14,27,4,14,2010,17.1\n2010-01-15,6,5,15,2010,16.833333333333332\n2010-01-16,11,6,16,2010,17.033333333333335\n2010-01-17,7,7,17,2010,16.766666666666666\n2010-01-18,10,1,18,2010,16.466666666666665\n2010-01-19,11,2,19,2010,16.433333333333334\n2010-01-20,26,3,20,2010,16.766666666666666\n2010-01-21,16,4,21,2010,16.9\n2010-01-22,20,5,22,2010,17.166666666666668\n2010-01-23,8,6,23,2010,17.0\n2010-01-24,18,7,24,2010,17.233333333333334\n2010-01-25,24,1,25,2010,16.966666666666665\n2010-01-26,46,2,26,2010,17.666666666666668\n2010-01-27,32,3,27,2010,18.2\n2010-01-28,28,4,28,2010,18.433333333333334\n2010-01-29,30,5,29,2010,18.533333333333335\n2010-01-30,43,6,30,2010,19.533333333333335\n2010-01-31,20,7,31,2010,19.733333333333334\n2010-02-01,43,1,32,2010,20.7\n2010-02-02,14,2,33,2010,20.766666666666666\n2010-02-03,8,3,34,2010,20.1\n2010-02-04,6,4,35,2010,20.066666666666666\n2010-02-05,10,5,36,2010,20.266666666666666\n2010-02-06,16,6,37,2010,20.633333333333333\n2010-02-07,11,7,38,2010,20.4\n2010-02-08,13,1,39,2010,20.5\n2010-02-09,26,2,40,2010,20.7\n2010-02-10,26,3,41,2010,20.266666666666666\n2010-02-11,49,4,42,2010,20.866666666666667\n2010-02-12,38,5,43,2010,21.433333333333334\n2010-02-13,20,6,44,2010,21.2\n2010-02-14,9,7,45,2010,21.3\n2010-02-15,19,1,46,2010,21.566666666666666\n2010-02-16,24,2,47,2010,22.133333333333333\n2010-02-17,20,3,48,2010,22.466666666666665\n2010-02-18,25,4,49,2010,22.933333333333334\n2010-02-19,29,5,50,2010,23.033333333333335\n2010-02-20,24,6,51,2010,23.3\n2010-02-21,37,7,52,2010,23.866666666666667\n2010-02-22,44,1,53,2010,25.066666666666666\n2010-02-23,24,2,54,2010,25.266666666666666\n2010-02-24,14,3,55,2010,24.933333333333334\n2010-02-25,10,4,56,2010,23.733333333333334\n2010-02-26,12,5,57,2010,23.066666666666666\n2010-02-27,14,6,58,2010,22.6\n2010-02-28,45,7,59,2010,23.1\n2010-03-01,35,1,60,2010,22.833333333333332\n2010-03-02,50,2,61,2010,23.833333333333332\n2010-03-03,31,3,62,2010,23.433333333333334\n2010-03-04,20,4,63,2010,23.633333333333333\n2010-03-05,15,5,64,2010,23.866666666666667\n2010-03-06,20,6,65,2010,24.333333333333332\n2010-03-07,25,7,66,2010,24.833333333333332\n2010-03-08,24,1,67,2010,25.1\n2010-03-09,18,2,68,2010,25.333333333333332\n2010-03-10,24,3,69,2010,25.7\n2010-03-11,13,4,70,2010,25.266666666666666\n2010-03-12,17,5,71,2010,24.966666666666665\n2010-03-13,8,6,72,2010,23.6\n2010-03-14,35,7,73,2010,23.5\n2010-03-15,33,1,74,2010,23.933333333333334\n2010-03-16,23,2,75,2010,24.4\n2010-03-17,40,3,76,2010,25.1\n2010-03-18,25,4,77,2010,25.133333333333333\n2010-03-19,13,5,78,2010,24.9\n2010-03-20,6,6,79,2010,24.266666666666666\n2010-03-21,8,7,80,2010,23.566666666666666\n2010-03-22,9,1,81,2010,23.066666666666666\n2010-03-23,15,2,82,2010,22.333333333333332\n2010-03-24,28,3,83,2010,21.8\n2010-03-25,15,4,84,2010,21.5\n2010-03-26,40,5,85,2010,22.366666666666667\n2010-03-27,10,6,86,2010,22.366666666666667\n2010-03-28,20,7,87,2010,22.633333333333333\n2010-03-29,34,1,88,2010,23.3\n2010-03-30,39,2,89,2010,23.1\n2010-03-31,26,3,90,2010,22.8\n2010-04-01,39,4,91,2010,22.433333333333334\n2010-04-02,36,5,92,2010,22.6\n2010-04-03,20,6,93,2010,22.6\n2010-04-04,19,7,94,2010,22.733333333333334\n2010-04-05,31,1,95,2010,23.1\n2010-04-06,36,2,96,2010,23.466666666666665\n2010-04-07,31,3,97,2010,23.7\n2010-04-08,30,4,98,2010,24.1\n2010-04-09,22,5,99,2010,24.033333333333335\n2010-04-10,26,6,100,2010,24.466666666666665\n2010-04-11,26,7,101,2010,24.766666666666666\n2010-04-12,62,1,102,2010,26.566666666666666\n2010-04-13,41,2,103,2010,26.766666666666666\n2010-04-14,30,3,104,2010,26.666666666666668\n2010-04-15,36,4,105,2010,27.1\n2010-04-16,40,5,106,2010,27.1\n2010-04-17,34,6,107,2010,27.4\n2010-04-18,30,7,108,2010,27.966666666666665\n2010-04-19,21,1,109,2010,28.466666666666665\n2010-04-20,25,2,110,2010,29.033333333333335\n2010-04-21,23,3,111,2010,29.5\n2010-04-22,30,4,112,2010,30.0\n2010-04-23,32,5,113,2010,30.133333333333333\n2010-04-24,23,6,114,2010,30.4\n2010-04-25,44,7,115,2010,30.533333333333335\n2010-04-26,29,1,116,2010,31.166666666666668\n2010-04-27,22,2,117,2010,31.233333333333334\n2010-04-28,42,3,118,2010,31.5\n2010-04-29,55,4,119,2010,32.03333333333333\n2010-04-30,19,5,120,2010,31.8\n2010-05-01,21,6,121,2010,31.2\n2010-05-02,39,7,122,2010,31.3\n2010-05-03,42,1,123,2010,32.03333333333333\n2010-05-04,32,2,124,2010,32.46666666666667\n2010-05-05,42,3,125,2010,32.833333333333336\n2010-05-06,41,4,126,2010,33.0\n2010-05-07,16,5,127,2010,32.5\n2010-05-08,14,6,128,2010,31.966666666666665\n2010-05-09,13,7,129,2010,31.666666666666668\n2010-05-10,16,1,130,2010,31.333333333333332\n2010-05-11,21,2,131,2010,31.166666666666668\n2010-05-12,13,3,132,2010,29.533333333333335\n2010-05-13,17,4,133,2010,28.733333333333334\n2010-05-14,37,5,134,2010,28.966666666666665\n2010-05-15,25,6,135,2010,28.6\n2010-05-16,7,7,136,2010,27.5\n2010-05-17,18,1,137,2010,26.966666666666665\n2010-05-18,19,2,138,2010,26.6\n2010-05-19,15,3,139,2010,26.4\n2010-05-20,24,4,140,2010,26.366666666666667\n2010-05-21,28,5,141,2010,26.533333333333335\n2010-05-22,26,6,142,2010,26.4\n2010-05-23,6,7,143,2010,25.533333333333335\n2010-05-24,7,1,144,2010,25.0\n2010-05-25,19,2,145,2010,24.166666666666668\n2010-05-26,26,3,146,2010,24.066666666666666\n2010-05-27,22,4,147,2010,24.066666666666666\n2010-05-28,20,5,148,2010,23.333333333333332\n2010-05-29,24,6,149,2010,22.3\n2010-05-30,29,7,150,2010,22.633333333333333\n2010-05-31,4,1,151,2010,22.066666666666666\n2010-06-01,14,2,152,2010,21.233333333333334\n2010-06-02,20,3,153,2010,20.5\n2010-06-03,22,4,154,2010,20.166666666666668\n2010-06-04,14,5,155,2010,19.233333333333334\n2010-06-05,14,6,156,2010,18.333333333333332\n2010-06-06,20,7,157,2010,18.466666666666665\n2010-06-07,31,1,158,2010,19.033333333333335\n2010-06-08,30,2,159,2010,19.6\n2010-06-09,36,3,160,2010,20.266666666666666\n2010-06-10,20,4,161,2010,20.233333333333334\n2010-06-11,37,5,162,2010,21.033333333333335\n2010-06-12,34,6,163,2010,21.6\n2010-06-13,17,7,164,2010,20.933333333333334\n2010-06-14,27,1,165,2010,21.0\n2010-06-15,17,2,166,2010,21.333333333333332\n2010-06-16,34,3,167,2010,21.866666666666667\n2010-06-17,44,4,168,2010,22.7\n2010-06-18,30,5,169,2010,23.2\n2010-06-19,23,6,170,2010,23.166666666666668\n2010-06-20,24,7,171,2010,23.033333333333335\n2010-06-21,24,1,172,2010,22.966666666666665\n2010-06-22,25,2,173,2010,23.6\n2010-06-23,24,3,174,2010,24.166666666666668\n2010-06-24,29,4,175,2010,24.5\n2010-06-25,33,5,176,2010,24.733333333333334\n2010-06-26,28,6,177,2010,24.933333333333334\n2010-06-27,21,7,178,2010,24.966666666666665\n2010-06-28,13,1,179,2010,24.6\n2010-06-29,43,2,180,2010,25.066666666666666\n2010-06-30,36,3,181,2010,26.133333333333333\n2010-07-01,17,4,182,2010,26.233333333333334\n2010-07-02,14,5,183,2010,26.033333333333335\n2010-07-03,16,6,184,2010,25.833333333333332\n2010-07-04,21,7,185,2010,26.066666666666666\n2010-07-05,14,1,186,2010,26.066666666666666\n2010-07-06,24,2,187,2010,26.2\n2010-07-07,23,3,188,2010,25.933333333333334\n2010-07-08,42,4,189,2010,26.333333333333332\n2010-07-09,19,5,190,2010,25.766666666666666\n2010-07-10,19,6,191,2010,25.733333333333334\n2010-07-11,11,7,192,2010,24.866666666666667\n2010-07-12,22,1,193,2010,24.466666666666665\n2010-07-13,25,2,194,2010,24.733333333333334\n2010-07-14,28,3,195,2010,24.766666666666666\n2010-07-15,13,4,196,2010,24.633333333333333\n2010-07-16,38,5,197,2010,24.766666666666666\n2010-07-17,12,6,198,2010,23.7\n2010-07-18,12,7,199,2010,23.1\n2010-07-19,19,1,200,2010,22.966666666666665\n2010-07-20,41,2,201,2010,23.533333333333335\n2010-07-21,28,3,202,2010,23.666666666666668\n2010-07-22,24,4,203,2010,23.633333333333333\n2010-07-23,36,5,204,2010,24.033333333333335\n2010-07-24,21,6,205,2010,23.766666666666666\n2010-07-25,7,7,206,2010,22.9\n2010-07-26,46,1,207,2010,23.5\n2010-07-27,25,2,208,2010,23.633333333333333\n2010-07-28,32,3,209,2010,24.266666666666666\n2010-07-29,16,4,210,2010,23.366666666666667\n2010-07-30,25,5,211,2010,23.0\n2010-07-31,21,6,212,2010,23.133333333333333\n2010-08-01,40,7,213,2010,24.0\n2010-08-02,20,1,214,2010,24.133333333333333\n2010-08-03,20,2,215,2010,24.1\n2010-08-04,19,3,216,2010,24.266666666666666\n2010-08-05,28,4,217,2010,24.4\n2010-08-06,24,5,218,2010,24.433333333333334\n2010-08-07,34,6,219,2010,24.166666666666668\n2010-08-08,18,7,220,2010,24.133333333333333\n2010-08-09,28,1,221,2010,24.433333333333334\n2010-08-10,19,2,222,2010,24.7\n2010-08-11,23,3,223,2010,24.733333333333334\n2010-08-12,24,4,224,2010,24.7\n2010-08-13,32,5,225,2010,24.833333333333332\n2010-08-14,16,6,226,2010,24.933333333333334\n2010-08-15,30,7,227,2010,24.666666666666668\n2010-08-16,48,1,228,2010,25.866666666666667\n2010-08-17,38,2,229,2010,26.733333333333334\n2010-08-18,23,3,230,2010,26.866666666666667\n2010-08-19,36,4,231,2010,26.7\n2010-08-20,33,5,232,2010,26.866666666666667\n2010-08-21,60,6,233,2010,28.066666666666666\n2010-08-22,17,7,234,2010,27.433333333333334\n2010-08-23,50,1,235,2010,28.4\n2010-08-24,38,2,236,2010,29.433333333333334\n2010-08-25,47,3,237,2010,29.466666666666665\n2010-08-26,50,4,238,2010,30.3\n2010-08-27,51,5,239,2010,30.933333333333334\n2010-08-28,33,6,240,2010,31.5\n2010-08-29,43,7,241,2010,32.1\n2010-08-30,38,1,242,2010,32.666666666666664\n2010-08-31,29,2,243,2010,32.3\n2010-09-01,22,3,244,2010,32.36666666666667\n2010-09-02,20,4,245,2010,32.36666666666667\n2010-09-03,20,5,246,2010,32.4\n2010-09-04,16,6,247,2010,32.0\n2010-09-05,9,7,248,2010,31.5\n2010-09-06,29,1,249,2010,31.333333333333332\n2010-09-07,20,2,250,2010,31.4\n2010-09-08,15,3,251,2010,30.966666666666665\n2010-09-09,40,4,252,2010,31.666666666666668\n2010-09-10,34,5,253,2010,32.03333333333333\n2010-09-11,32,6,254,2010,32.3\n2010-09-12,18,7,255,2010,31.833333333333332\n2010-09-13,42,1,256,2010,32.7\n2010-09-14,39,2,257,2010,33.0\n2010-09-15,24,3,258,2010,32.2\n2010-09-16,33,4,259,2010,32.03333333333333\n2010-09-17,59,5,260,2010,33.233333333333334\n2010-09-18,35,6,261,2010,33.2\n2010-09-19,37,7,262,2010,33.333333333333336\n2010-09-20,61,1,263,2010,33.36666666666667\n2010-09-21,43,2,264,2010,34.233333333333334\n2010-09-22,49,3,265,2010,34.2\n2010-09-23,44,4,266,2010,34.4\n2010-09-24,53,5,267,2010,34.6\n2010-09-25,33,6,268,2010,34.03333333333333\n2010-09-26,21,7,269,2010,33.03333333333333\n2010-09-27,42,1,270,2010,33.333333333333336\n2010-09-28,22,2,271,2010,32.63333333333333\n2010-09-29,49,3,272,2010,33.0\n2010-09-30,28,4,273,2010,32.96666666666667\n2010-10-01,33,5,274,2010,33.333333333333336\n2010-10-02,32,6,275,2010,33.733333333333334\n2010-10-03,30,7,276,2010,34.06666666666667\n2010-10-04,41,1,277,2010,34.9\n2010-10-05,40,2,278,2010,35.93333333333333\n2010-10-06,30,3,279,2010,35.96666666666667\n2010-10-07,26,4,280,2010,36.166666666666664\n2010-10-08,24,5,281,2010,36.46666666666667\n2010-10-09,10,6,282,2010,35.46666666666667\n2010-10-10,18,7,283,2010,34.93333333333333\n2010-10-11,31,1,284,2010,34.9\n2010-10-12,47,2,285,2010,35.86666666666667\n2010-10-13,24,3,286,2010,35.266666666666666\n2010-10-14,33,4,287,2010,35.06666666666667\n2010-10-15,34,5,288,2010,35.4\n2010-10-16,31,6,289,2010,35.333333333333336\n2010-10-17,18,7,290,2010,33.96666666666667\n2010-10-18,25,1,291,2010,33.63333333333333\n2010-10-19,47,2,292,2010,33.96666666666667\n2010-10-20,42,3,293,2010,33.333333333333336\n2010-10-21,13,4,294,2010,32.333333333333336\n2010-10-22,13,5,295,2010,31.133333333333333\n2010-10-23,17,6,296,2010,30.233333333333334\n2010-10-24,31,7,297,2010,29.5\n2010-10-25,38,1,298,2010,29.666666666666668\n2010-10-26,41,2,299,2010,30.333333333333332\n2010-10-27,34,3,300,2010,30.066666666666666\n2010-10-28,37,4,301,2010,30.566666666666666\n2010-10-29,36,5,302,2010,30.133333333333333\n2010-10-30,15,6,303,2010,29.7\n2010-10-31,26,7,304,2010,29.466666666666665\n2010-11-01,32,1,305,2010,29.466666666666665\n2010-11-02,42,2,306,2010,29.866666666666667\n2010-11-03,33,3,307,2010,29.6\n2010-11-04,32,4,308,2010,29.333333333333332\n2010-11-05,18,5,309,2010,28.933333333333334\n2010-11-06,39,6,310,2010,29.366666666666667\n2010-11-07,37,7,311,2010,29.8\n2010-11-08,45,1,312,2010,30.966666666666665\n2010-11-09,60,2,313,2010,32.36666666666667\n2010-11-10,62,3,314,2010,33.4\n2010-11-11,47,4,315,2010,33.4\n2010-11-12,45,5,316,2010,34.1\n2010-11-13,24,6,317,2010,33.8\n2010-11-14,43,7,318,2010,34.1\n2010-11-15,38,1,319,2010,34.333333333333336\n2010-11-16,46,2,320,2010,35.266666666666666\n2010-11-17,33,3,321,2010,35.53333333333333\n2010-11-18,51,4,322,2010,35.666666666666664\n2010-11-19,48,5,323,2010,35.86666666666667\n2010-11-20,50,6,324,2010,37.1\n2010-11-21,13,7,325,2010,37.1\n2010-11-22,43,1,326,2010,37.96666666666667\n2010-11-23,60,2,327,2010,38.93333333333333\n2010-11-24,50,3,328,2010,39.333333333333336\n2010-11-25,35,4,329,2010,39.13333333333333\n2010-11-26,60,5,330,2010,40.0\n2010-11-27,76,6,331,2010,41.3\n2010-11-28,61,7,332,2010,42.13333333333333\n2010-11-29,28,1,333,2010,42.56666666666667\n2010-11-30,47,2,334,2010,43.266666666666666\n2010-12-01,40,3,335,2010,43.53333333333333\n2010-12-02,36,4,336,2010,43.333333333333336\n2010-12-03,32,5,337,2010,43.3\n2010-12-04,34,6,338,2010,43.36666666666667\n2010-12-05,38,7,339,2010,44.03333333333333\n2010-12-06,37,1,340,2010,43.96666666666667\n2010-12-07,28,2,341,2010,43.666666666666664\n2010-12-08,24,3,342,2010,42.96666666666667\n2010-12-09,26,4,343,2010,41.833333333333336\n2010-12-10,5,5,344,2010,39.93333333333333\n2010-12-11,17,6,345,2010,38.93333333333333\n2010-12-12,21,7,346,2010,38.13333333333333\n2010-12-13,32,1,347,2010,38.4\n2010-12-14,23,2,348,2010,37.733333333333334\n2010-12-15,29,3,349,2010,37.43333333333333\n2010-12-16,33,4,350,2010,37.0\n2010-12-17,30,5,351,2010,36.9\n2010-12-18,38,6,352,2010,36.46666666666667\n2010-12-19,19,7,353,2010,35.5\n2010-12-20,18,1,354,2010,34.43333333333333\n2010-12-21,35,2,355,2010,35.166666666666664\n2010-12-22,42,3,356,2010,35.13333333333333\n2010-12-23,34,4,357,2010,34.266666666666666\n2010-12-24,17,5,358,2010,33.166666666666664\n2010-12-25,13,6,359,2010,32.43333333333333\n2010-12-26,15,7,360,2010,30.933333333333334\n2010-12-27,31,1,361,2010,29.433333333333334\n2010-12-28,40,2,362,2010,28.733333333333334\n2010-12-29,67,3,363,2010,30.033333333333335\n2010-12-30,18,4,364,2010,29.066666666666666\n2010-12-31,19,5,365,2010,28.366666666666667\n2011-01-01,20,6,1,2011,27.833333333333332\n2011-01-02,39,7,2,2011,28.066666666666666\n2011-01-03,50,1,3,2011,28.6\n2011-01-04,38,2,4,2011,28.6\n2011-01-05,45,3,5,2011,28.866666666666667\n2011-01-06,37,4,6,2011,29.166666666666668\n2011-01-07,27,5,7,2011,29.266666666666666\n2011-01-08,38,6,8,2011,29.666666666666668\n2011-01-09,33,7,9,2011,30.6\n2011-01-10,43,1,10,2011,31.466666666666665\n2011-01-11,49,2,11,2011,32.4\n2011-01-12,45,3,12,2011,32.833333333333336\n2011-01-13,53,4,13,2011,33.833333333333336\n2011-01-14,45,5,14,2011,34.36666666666667\n2011-01-15,29,6,15,2011,34.233333333333334\n2011-01-16,7,7,16,2011,33.46666666666667\n2011-01-17,35,1,17,2011,33.36666666666667\n2011-01-18,26,2,18,2011,33.6\n2011-01-19,31,3,19,2011,34.03333333333333\n2011-01-20,26,4,20,2011,33.733333333333334\n2011-01-21,57,5,21,2011,34.233333333333334\n2011-01-22,28,6,22,2011,34.03333333333333\n2011-01-23,44,7,23,2011,34.93333333333333\n2011-01-24,26,1,24,2011,35.36666666666667\n2011-01-25,26,2,25,2011,35.733333333333334\n2011-01-26,38,3,26,2011,35.96666666666667\n2011-01-27,39,4,27,2011,35.93333333333333\n2011-01-28,23,5,28,2011,34.46666666666667\n2011-01-29,21,6,29,2011,34.56666666666667\n2011-01-30,52,7,30,2011,35.666666666666664\n2011-01-31,19,1,31,2011,35.63333333333333\n2011-02-01,28,2,32,2011,35.266666666666666\n2011-02-02,31,3,33,2011,34.63333333333333\n2011-02-03,24,4,34,2011,34.166666666666664\n2011-02-04,16,5,35,2011,33.2\n2011-02-05,34,6,36,2011,33.1\n2011-02-06,16,7,37,2011,32.733333333333334\n2011-02-07,36,1,38,2011,32.666666666666664\n2011-02-08,24,2,39,2011,32.36666666666667\n2011-02-09,22,3,40,2011,31.666666666666668\n2011-02-10,38,4,41,2011,31.3\n2011-02-11,19,5,42,2011,30.433333333333334\n2011-02-12,17,6,43,2011,29.233333333333334\n2011-02-13,28,7,44,2011,28.666666666666668\n2011-02-14,20,1,45,2011,28.366666666666667\n2011-02-15,20,2,46,2011,28.8\n2011-02-16,30,3,47,2011,28.633333333333333\n2011-02-17,39,4,48,2011,29.066666666666666\n2011-02-18,31,5,49,2011,29.066666666666666\n2011-02-19,12,6,50,2011,28.6\n2011-02-20,11,7,51,2011,27.066666666666666\n2011-02-21,29,1,52,2011,27.1\n2011-02-22,29,2,53,2011,26.6\n2011-02-23,24,3,54,2011,26.533333333333335\n2011-02-24,49,4,55,2011,27.3\n2011-02-25,12,5,56,2011,26.433333333333334\n2011-02-26,10,6,57,2011,25.466666666666665\n2011-02-27,19,7,58,2011,25.333333333333332\n2011-02-28,14,1,59,2011,25.1\n2011-03-01,22,2,60,2011,24.1\n2011-03-02,48,3,61,2011,25.066666666666666\n2011-03-03,28,4,62,2011,25.066666666666666\n2011-03-04,26,5,63,2011,24.9\n2011-03-05,13,6,64,2011,24.533333333333335\n2011-03-06,19,7,65,2011,24.633333333333333\n2011-03-07,15,1,66,2011,24.0\n2011-03-08,30,2,67,2011,24.466666666666665\n2011-03-09,28,3,68,2011,24.2\n2011-03-10,22,4,69,2011,24.133333333333333\n2011-03-11,31,5,70,2011,24.433333333333334\n2011-03-12,24,6,71,2011,23.966666666666665\n2011-03-13,26,7,72,2011,24.2\n2011-03-14,71,1,73,2011,26.0\n2011-03-15,54,2,74,2011,26.866666666666667\n2011-03-16,22,3,75,2011,26.933333333333334\n2011-03-17,42,4,76,2011,27.666666666666668\n2011-03-18,22,5,77,2011,27.4\n2011-03-19,17,6,78,2011,26.666666666666668\n2011-03-20,31,7,79,2011,26.666666666666668\n2011-03-21,46,1,80,2011,27.8\n2011-03-22,44,2,81,2011,28.9\n2011-03-23,55,3,82,2011,29.766666666666666\n2011-03-24,44,4,83,2011,30.266666666666666\n2011-03-25,40,5,84,2011,30.8\n2011-03-26,22,6,85,2011,29.9\n2011-03-27,28,7,86,2011,30.433333333333334\n2011-03-28,18,1,87,2011,30.7\n2011-03-29,44,2,88,2011,31.533333333333335\n2011-03-30,48,3,89,2011,32.666666666666664\n2011-03-31,51,4,90,2011,33.63333333333333\n2011-04-01,49,5,91,2011,33.666666666666664\n2011-04-02,30,6,92,2011,33.733333333333334\n2011-04-03,10,7,93,2011,33.2\n2011-04-04,29,1,94,2011,33.733333333333334\n2011-04-05,27,2,95,2011,34.0\n2011-04-06,8,3,96,2011,33.766666666666666\n2011-04-07,20,4,97,2011,33.43333333333333\n2011-04-08,27,5,98,2011,33.4\n2011-04-09,36,6,99,2011,33.86666666666667\n2011-04-10,14,7,100,2011,33.3\n2011-04-11,25,1,101,2011,33.333333333333336\n2011-04-12,40,2,102,2011,33.8\n2011-04-13,16,3,103,2011,31.966666666666665\n2011-04-14,15,4,104,2011,30.666666666666668\n2011-04-15,50,5,105,2011,31.6\n2011-04-16,28,6,106,2011,31.133333333333333\n2011-04-17,47,7,107,2011,31.966666666666665\n2011-04-18,65,1,108,2011,33.56666666666667\n2011-04-19,56,2,109,2011,34.4\n2011-04-20,47,3,110,2011,34.43333333333333\n2011-04-21,80,4,111,2011,35.63333333333333\n2011-04-22,59,5,112,2011,35.766666666666666\n2011-04-23,65,6,113,2011,36.46666666666667\n2011-04-24,50,7,114,2011,36.8\n2011-04-25,62,1,115,2011,38.13333333333333\n2011-04-26,51,2,116,2011,38.9\n2011-04-27,37,3,117,2011,39.53333333333333\n2011-04-28,59,4,118,2011,40.03333333333333\n2011-04-29,54,5,119,2011,40.233333333333334\n2011-04-30,23,6,120,2011,39.3\n2011-05-01,50,7,121,2011,39.333333333333336\n2011-05-02,49,1,122,2011,39.96666666666667\n2011-05-03,27,2,123,2011,40.53333333333333\n2011-05-04,52,3,124,2011,41.3\n2011-05-05,33,4,125,2011,41.5\n2011-05-06,26,5,126,2011,42.1\n2011-05-07,12,6,127,2011,41.833333333333336\n2011-05-08,18,7,128,2011,41.53333333333333\n2011-05-09,22,1,129,2011,41.06666666666667\n2011-05-10,26,2,130,2011,41.46666666666667\n2011-05-11,34,3,131,2011,41.766666666666666\n2011-05-12,27,4,132,2011,41.333333333333336\n2011-05-13,29,5,133,2011,41.766666666666666\n2011-05-14,12,6,134,2011,41.666666666666664\n2011-05-15,22,7,135,2011,40.733333333333334\n2011-05-16,25,1,136,2011,40.63333333333333\n2011-05-17,20,2,137,2011,39.733333333333334\n2011-05-18,36,3,138,2011,38.766666666666666\n2011-05-19,27,4,139,2011,37.8\n2011-05-20,14,5,140,2011,36.7\n2011-05-21,14,6,141,2011,34.5\n2011-05-22,45,7,142,2011,34.03333333333333\n2011-05-23,37,1,143,2011,33.1\n2011-05-24,43,2,144,2011,32.86666666666667\n2011-05-25,23,3,145,2011,31.566666666666666\n2011-05-26,18,4,146,2011,30.466666666666665\n2011-05-27,42,5,147,2011,30.633333333333333\n2011-05-28,79,6,148,2011,31.3\n2011-05-29,57,7,149,2011,31.4\n2011-05-30,39,1,150,2011,31.933333333333334\n2011-05-31,73,2,151,2011,32.7\n2011-06-01,37,3,152,2011,32.3\n2011-06-02,58,4,153,2011,33.333333333333336\n2011-06-03,30,5,154,2011,32.6\n2011-06-04,58,6,155,2011,33.43333333333333\n2011-06-05,10,7,156,2011,32.9\n2011-06-06,30,1,157,2011,33.5\n2011-06-07,44,2,158,2011,34.36666666666667\n2011-06-08,59,3,159,2011,35.6\n2011-06-09,38,4,160,2011,36.0\n2011-06-10,101,5,161,2011,38.233333333333334\n2011-06-11,36,6,162,2011,38.53333333333333\n2011-06-12,46,7,163,2011,39.1\n2011-06-13,50,1,164,2011,40.36666666666667\n2011-06-14,83,2,165,2011,42.4\n2011-06-15,37,3,166,2011,42.8\n2011-06-16,67,4,167,2011,44.36666666666667\n2011-06-17,116,5,168,2011,47.03333333333333\n2011-06-18,98,6,169,2011,49.4\n2011-06-19,49,7,170,2011,50.56666666666667\n2011-06-20,71,1,171,2011,52.46666666666667\n2011-06-21,68,2,172,2011,53.233333333333334\n2011-06-22,82,3,173,2011,54.733333333333334\n2011-06-23,58,4,174,2011,55.233333333333334\n2011-06-24,131,5,175,2011,58.833333333333336\n2011-06-25,51,6,176,2011,59.93333333333333\n2011-06-26,34,7,177,2011,59.666666666666664\n2011-06-27,88,1,178,2011,59.96666666666667\n2011-06-28,81,2,179,2011,60.766666666666666\n2011-06-29,64,3,180,2011,61.6\n2011-06-30,46,4,181,2011,60.7\n2011-07-01,118,5,182,2011,63.4\n2011-07-02,49,6,183,2011,63.1\n2011-07-03,37,7,184,2011,63.333333333333336\n2011-07-04,29,1,185,2011,62.36666666666667\n2011-07-05,50,2,186,2011,63.7\n2011-07-06,42,3,187,2011,64.1\n2011-07-07,44,4,188,2011,64.1\n2011-07-08,57,5,189,2011,64.03333333333333\n2011-07-09,27,6,190,2011,63.666666666666664\n2011-07-10,26,7,191,2011,61.166666666666664\n2011-07-11,40,1,192,2011,61.3\n2011-07-12,56,2,193,2011,61.63333333333333\n2011-07-13,55,3,194,2011,61.8\n2011-07-14,59,4,195,2011,61.0\n2011-07-15,160,5,196,2011,65.1\n2011-07-16,83,6,197,2011,65.63333333333334\n2011-07-17,147,7,198,2011,66.66666666666667\n2011-07-18,139,1,199,2011,68.03333333333333\n2011-07-19,105,2,200,2011,69.9\n2011-07-20,94,3,201,2011,70.66666666666667\n2011-07-21,28,4,202,2011,69.33333333333333\n2011-07-22,80,5,203,2011,69.26666666666667\n2011-07-23,35,6,204,2011,68.5\n2011-07-24,28,7,205,2011,65.06666666666666\n2011-07-25,42,1,206,2011,64.76666666666667\n2011-07-26,73,2,207,2011,66.06666666666666\n2011-07-27,76,3,208,2011,65.66666666666667\n2011-07-28,48,4,209,2011,64.56666666666666\n2011-07-29,66,5,210,2011,64.63333333333334\n2011-07-30,66,6,211,2011,65.3\n2011-07-31,54,7,212,2011,63.166666666666664\n2011-08-01,45,1,213,2011,63.03333333333333\n2011-08-02,79,2,214,2011,64.43333333333334\n2011-08-03,70,3,215,2011,65.8\n2011-08-04,81,4,216,2011,66.83333333333333\n2011-08-05,85,5,217,2011,68.26666666666667\n2011-08-06,36,6,218,2011,68.0\n2011-08-07,73,7,219,2011,68.53333333333333\n2011-08-08,25,1,220,2011,68.46666666666667\n2011-08-09,60,2,221,2011,69.6\n2011-08-10,58,3,222,2011,70.2\n2011-08-11,73,4,223,2011,70.76666666666667\n2011-08-12,70,5,224,2011,71.26666666666667\n2011-08-13,34,6,225,2011,70.43333333333334\n2011-08-14,38,7,226,2011,66.36666666666666\n2011-08-15,37,1,227,2011,64.83333333333333\n2011-08-16,66,2,228,2011,62.13333333333333\n2011-08-17,35,3,229,2011,58.666666666666664\n2011-08-18,51,4,230,2011,56.86666666666667\n2011-08-19,54,5,231,2011,55.53333333333333\n2011-08-20,36,6,232,2011,55.8\n2011-08-21,65,7,233,2011,55.3\n2011-08-22,40,1,234,2011,55.46666666666667\n2011-08-23,33,2,235,2011,55.63333333333333\n2011-08-24,65,3,236,2011,56.4\n2011-08-25,59,4,237,2011,55.93333333333333\n2011-08-26,66,5,238,2011,55.6\n2011-08-27,31,6,239,2011,55.03333333333333\n2011-08-28,45,7,240,2011,54.333333333333336\n2011-08-29,32,1,241,2011,53.2\n2011-08-30,38,2,242,2011,52.666666666666664\n2011-08-31,60,3,243,2011,53.166666666666664\n2011-09-01,58,4,244,2011,52.46666666666667\n2011-09-02,63,5,245,2011,52.233333333333334\n2011-09-03,38,6,246,2011,50.8\n2011-09-04,25,7,247,2011,48.8\n2011-09-05,40,1,248,2011,48.93333333333333\n2011-09-06,62,2,249,2011,48.56666666666667\n2011-09-07,91,3,250,2011,50.766666666666666\n2011-09-08,64,4,251,2011,50.9\n2011-09-09,56,5,252,2011,50.833333333333336\n2011-09-10,37,6,253,2011,49.63333333333333\n2011-09-11,51,7,254,2011,49.0\n2011-09-12,41,1,255,2011,49.233333333333334\n2011-09-13,40,2,256,2011,49.3\n2011-09-14,69,3,257,2011,50.36666666666667\n2011-09-15,48,4,258,2011,49.766666666666666\n2011-09-16,44,5,259,2011,50.06666666666667\n2011-09-17,14,6,260,2011,48.833333333333336\n2011-09-18,28,7,261,2011,47.96666666666667\n2011-09-19,29,1,262,2011,47.733333333333334\n2011-09-20,59,2,263,2011,47.53333333333333\n2011-09-21,48,3,264,2011,47.8\n2011-09-22,34,4,265,2011,47.833333333333336\n2011-09-23,104,5,266,2011,49.13333333333333\n2011-09-24,84,6,267,2011,49.96666666666667\n2011-09-25,48,7,268,2011,49.36666666666667\n2011-09-26,81,1,269,2011,51.03333333333333\n2011-09-27,84,2,270,2011,52.333333333333336\n2011-09-28,88,3,271,2011,54.2\n2011-09-29,99,4,272,2011,56.233333333333334\n2011-09-30,110,5,273,2011,57.9\n2011-10-01,120,6,274,2011,59.96666666666667\n2011-10-02,90,7,275,2011,60.86666666666667\n2011-10-03,57,1,276,2011,61.5\n2011-10-04,67,2,277,2011,62.9\n2011-10-05,71,3,278,2011,63.93333333333333\n2011-10-06,48,4,279,2011,63.46666666666667\n2011-10-07,74,5,280,2011,62.9\n2011-10-08,59,6,281,2011,62.733333333333334\n2011-10-09,69,7,282,2011,63.166666666666664\n2011-10-10,75,1,283,2011,64.43333333333334\n2011-10-11,78,2,284,2011,65.33333333333333\n2011-10-12,72,3,285,2011,66.36666666666666\n2011-10-13,82,4,286,2011,67.76666666666667\n2011-10-14,61,5,287,2011,67.5\n2011-10-15,61,6,288,2011,67.93333333333334\n2011-10-16,41,7,289,2011,67.83333333333333\n2011-10-17,104,1,290,2011,70.83333333333333\n2011-10-18,62,2,291,2011,71.96666666666667\n2011-10-19,88,3,292,2011,73.93333333333334\n2011-10-20,75,4,293,2011,74.46666666666667\n2011-10-21,97,5,294,2011,76.1\n2011-10-22,45,6,295,2011,76.46666666666667\n2011-10-23,58,7,296,2011,74.93333333333334\n2011-10-24,85,1,297,2011,74.96666666666667\n2011-10-25,109,2,298,2011,77.0\n2011-10-26,66,3,299,2011,76.5\n2011-10-27,45,4,300,2011,75.2\n2011-10-28,57,5,301,2011,74.16666666666667\n2011-10-29,53,6,302,2011,72.63333333333334\n2011-10-30,50,7,303,2011,70.63333333333334\n2011-10-31,89,1,304,2011,69.6\n2011-11-01,61,2,305,2011,68.63333333333334\n2011-11-02,78,3,306,2011,69.33333333333333\n2011-11-03,74,4,307,2011,69.56666666666666\n2011-11-04,85,5,308,2011,70.03333333333333\n2011-11-05,84,6,309,2011,71.23333333333333\n2011-11-06,43,7,310,2011,70.2\n2011-11-07,53,1,311,2011,70.0\n2011-11-08,48,2,312,2011,69.3\n2011-11-09,71,3,313,2011,69.16666666666667\n2011-11-10,35,4,314,2011,67.73333333333333\n2011-11-11,72,5,315,2011,67.73333333333333\n2011-11-12,115,6,316,2011,68.83333333333333\n2011-11-13,90,7,317,2011,69.8\n2011-11-14,64,1,318,2011,69.9\n2011-11-15,49,2,319,2011,70.16666666666667\n2011-11-16,45,3,320,2011,68.2\n2011-11-17,79,4,321,2011,68.76666666666667\n2011-11-18,61,5,322,2011,67.86666666666666\n2011-11-19,81,6,323,2011,68.06666666666666\n2011-11-20,78,7,324,2011,67.43333333333334\n2011-11-21,63,1,325,2011,68.03333333333333\n2011-11-22,89,2,326,2011,69.06666666666666\n2011-11-23,66,3,327,2011,68.43333333333334\n2011-11-24,60,4,328,2011,66.8\n2011-11-25,90,5,329,2011,67.6\n2011-11-26,108,6,330,2011,69.7\n2011-11-27,88,7,331,2011,70.73333333333333\n2011-11-28,64,1,332,2011,71.1\n2011-11-29,72,2,333,2011,71.83333333333333\n2011-11-30,76,3,334,2011,71.4\n2011-12-01,107,4,335,2011,72.93333333333334\n2011-12-02,73,5,336,2011,72.76666666666667\n2011-12-03,65,6,337,2011,72.46666666666667\n2011-12-04,74,7,338,2011,72.1\n2011-12-05,80,1,339,2011,71.96666666666667\n2011-12-06,116,2,340,2011,74.4\n2011-12-07,96,3,341,2011,75.83333333333333\n2011-12-08,113,4,342,2011,78.0\n2011-12-09,101,5,343,2011,79.0\n2011-12-10,88,6,344,2011,80.76666666666667\n2011-12-11,142,7,345,2011,83.1\n2011-12-12,121,1,346,2011,83.3\n2011-12-13,118,2,347,2011,84.23333333333333\n2011-12-14,86,3,348,2011,84.96666666666667\n2011-12-15,73,4,349,2011,85.76666666666667\n2011-12-16,84,5,350,2011,87.06666666666666\n2011-12-17,56,6,351,2011,86.3\n2011-12-18,86,7,352,2011,87.13333333333334\n2011-12-19,85,1,353,2011,87.26666666666667\n2011-12-20,89,2,354,2011,87.63333333333334\n2011-12-21,58,3,355,2011,87.46666666666667\n2011-12-22,46,4,356,2011,86.03333333333333\n2011-12-23,65,5,357,2011,86.0\n2011-12-24,39,6,358,2011,85.3\n2011-12-25,18,7,359,2011,82.9\n2011-12-26,39,1,360,2011,80.6\n2011-12-27,55,2,361,2011,79.5\n2011-12-28,47,3,362,2011,78.93333333333334\n2011-12-29,71,4,363,2011,78.9\n2011-12-30,85,5,364,2011,79.2\n2011-12-31,73,6,365,2011,78.06666666666666\n2012-01-01,56,7,1,2012,77.5\n2012-01-02,76,1,2,2012,77.86666666666666\n2012-01-03,60,2,3,2012,77.4\n2012-01-04,121,3,4,2012,78.76666666666667\n2012-01-05,77,4,5,2012,77.46666666666667\n2012-01-06,65,5,6,2012,76.43333333333334\n2012-01-07,60,6,7,2012,74.66666666666667\n2012-01-08,82,7,8,2012,74.03333333333333\n2012-01-09,72,1,9,2012,73.5\n2012-01-10,106,2,10,2012,72.3\n2012-01-11,96,3,11,2012,71.46666666666667\n2012-01-12,82,4,12,2012,70.26666666666667\n2012-01-13,145,5,13,2012,72.23333333333333\n2012-01-14,65,6,14,2012,71.96666666666667\n2012-01-15,64,7,15,2012,71.3\n2012-01-16,71,1,16,2012,71.8\n2012-01-17,79,2,17,2012,71.56666666666666\n2012-01-18,88,3,18,2012,71.66666666666667\n2012-01-19,82,4,19,2012,71.43333333333334\n2012-01-20,71,5,20,2012,71.86666666666666\n2012-01-21,57,6,21,2012,72.23333333333333\n2012-01-22,61,7,22,2012,72.1\n2012-01-23,69,1,23,2012,73.1\n2012-01-24,107,2,24,2012,76.06666666666666\n2012-01-25,99,3,25,2012,78.06666666666666\n2012-01-26,92,4,26,2012,79.3\n2012-01-27,70,5,27,2012,80.06666666666666\n2012-01-28,44,6,28,2012,79.16666666666667\n2012-01-29,84,7,29,2012,79.13333333333334\n2012-01-30,54,1,30,2012,78.5\n2012-01-31,71,2,31,2012,79.0\n2012-02-01,95,3,32,2012,79.63333333333334\n2012-02-02,86,4,33,2012,80.5\n2012-02-03,87,5,34,2012,79.36666666666666\n2012-02-04,82,6,35,2012,79.53333333333333\n2012-02-05,115,7,36,2012,81.2\n2012-02-06,189,1,37,2012,85.5\n2012-02-07,116,2,38,2012,86.63333333333334\n2012-02-08,122,3,39,2012,88.3\n2012-02-09,106,4,40,2012,88.3\n2012-02-10,97,5,41,2012,88.33333333333333\n2012-02-11,94,6,42,2012,88.73333333333333\n2012-02-12,90,7,43,2012,86.9\n2012-02-13,91,1,44,2012,87.76666666666667\n2012-02-14,106,2,45,2012,89.16666666666667\n2012-02-15,93,3,46,2012,89.9\n2012-02-16,119,4,47,2012,91.23333333333333\n2012-02-17,122,5,48,2012,92.36666666666666\n2012-02-18,109,6,49,2012,93.26666666666667\n2012-02-19,161,7,50,2012,96.26666666666667\n2012-02-20,120,1,51,2012,98.36666666666666\n2012-02-21,144,2,52,2012,101.13333333333334\n2012-02-22,117,3,53,2012,102.73333333333333\n2012-02-23,78,4,54,2012,101.76666666666667\n2012-02-24,68,5,55,2012,100.73333333333333\n2012-02-25,104,6,56,2012,101.13333333333334\n2012-02-26,81,7,57,2012,101.5\n2012-02-27,87,1,58,2012,102.93333333333334\n2012-02-28,110,2,59,2012,103.8\n2012-02-29,71,3,60,2012,104.36666666666666\n2012-03-01,83,4,61,2012,104.76666666666667\n2012-03-02,65,5,62,2012,103.76666666666667\n2012-03-03,82,6,63,2012,103.63333333333334\n2012-03-04,133,7,64,2012,105.16666666666667\n2012-03-05,139,1,65,2012,107.06666666666666\n2012-03-06,107,2,66,2012,106.8\n2012-03-07,77,3,67,2012,103.06666666666666\n2012-03-08,102,4,68,2012,102.6\n2012-03-09,111,5,69,2012,102.23333333333333\n2012-03-10,100,6,70,2012,102.03333333333333\n2012-03-11,166,7,71,2012,104.33333333333333\n2012-03-12,78,1,72,2012,103.8\n2012-03-13,100,2,73,2012,104.13333333333334\n2012-03-14,57,3,74,2012,103.0\n2012-03-15,53,4,75,2012,101.23333333333333\n2012-03-16,99,5,76,2012,101.43333333333334\n2012-03-17,88,6,77,2012,100.4\n2012-03-18,67,7,78,2012,98.56666666666666\n2012-03-19,160,1,79,2012,100.26666666666667\n2012-03-20,100,2,80,2012,98.23333333333333\n2012-03-21,101,3,81,2012,97.6\n2012-03-22,99,4,82,2012,96.1\n2012-03-23,123,5,83,2012,96.3\n2012-03-24,93,6,84,2012,96.8\n2012-03-25,142,7,85,2012,99.26666666666667\n2012-03-26,111,1,86,2012,99.5\n2012-03-27,84,2,87,2012,99.6\n2012-03-28,67,3,88,2012,98.93333333333334\n2012-03-29,76,4,89,2012,97.8\n2012-03-30,99,5,90,2012,98.73333333333333\n2012-03-31,109,6,91,2012,99.6\n2012-04-01,118,7,92,2012,101.36666666666666\n2012-04-02,124,1,93,2012,102.76666666666667\n2012-04-03,109,2,94,2012,101.96666666666667\n2012-04-04,175,3,95,2012,103.16666666666667\n2012-04-05,118,4,96,2012,103.53333333333333\n2012-04-06,92,5,97,2012,104.03333333333333\n2012-04-07,147,6,98,2012,105.53333333333333\n2012-04-08,80,7,99,2012,104.5\n2012-04-09,135,1,100,2012,105.66666666666667\n2012-04-10,120,2,101,2012,104.13333333333334\n2012-04-11,166,3,102,2012,107.06666666666666\n2012-04-12,77,4,103,2012,106.3\n2012-04-13,67,5,104,2012,106.63333333333334\n2012-04-14,94,6,105,2012,108.0\n2012-04-15,102,7,106,2012,108.1\n2012-04-16,123,1,107,2012,109.26666666666667\n2012-04-17,88,2,108,2012,109.96666666666667\n2012-04-18,130,3,109,2012,108.96666666666667\n2012-04-19,86,4,110,2012,108.5\n2012-04-20,108,5,111,2012,108.73333333333333\n2012-04-21,117,6,112,2012,109.33333333333333\n2012-04-22,95,7,113,2012,108.4\n2012-04-23,112,1,114,2012,109.03333333333333\n2012-04-24,106,2,115,2012,107.83333333333333\n2012-04-25,95,3,116,2012,107.3\n2012-04-26,85,4,117,2012,107.33333333333333\n2012-04-27,107,5,118,2012,108.66666666666667\n2012-04-28,120,6,119,2012,110.13333333333334\n2012-04-29,165,7,120,2012,112.33333333333333\n2012-04-30,127,1,121,2012,112.93333333333334\n2012-05-01,86,2,122,2012,111.86666666666666\n2012-05-02,107,3,123,2012,111.3\n2012-05-03,92,4,124,2012,110.73333333333333\n2012-05-04,107,5,125,2012,108.46666666666667\n2012-05-05,80,6,126,2012,107.2\n2012-05-06,103,7,127,2012,107.56666666666666\n2012-05-07,131,1,128,2012,107.03333333333333\n2012-05-08,73,2,129,2012,106.8\n2012-05-09,105,3,130,2012,105.8\n2012-05-10,81,4,131,2012,104.5\n2012-05-11,140,5,132,2012,103.63333333333334\n2012-05-12,83,6,133,2012,103.83333333333333\n2012-05-13,109,7,134,2012,105.23333333333333\n2012-05-14,67,1,135,2012,104.33333333333333\n2012-05-15,71,2,136,2012,103.3\n2012-05-16,86,3,137,2012,102.06666666666666\n2012-05-17,118,4,138,2012,103.06666666666666\n2012-05-18,101,5,139,2012,102.1\n2012-05-19,77,6,140,2012,101.8\n2012-05-20,110,7,141,2012,101.86666666666666\n2012-05-21,77,1,142,2012,100.53333333333333\n2012-05-22,81,2,143,2012,100.06666666666666\n2012-05-23,70,3,144,2012,98.66666666666667\n2012-05-24,79,4,145,2012,97.76666666666667\n2012-05-25,60,5,146,2012,96.6\n2012-05-26,65,6,147,2012,95.93333333333334\n2012-05-27,46,7,148,2012,93.9\n2012-05-28,74,1,149,2012,92.36666666666666\n2012-05-29,48,2,150,2012,88.46666666666667\n2012-05-30,90,3,151,2012,87.23333333333333\n2012-05-31,95,4,152,2012,87.53333333333333\n2012-06-01,84,5,153,2012,86.76666666666667\n2012-06-02,31,6,154,2012,84.73333333333333\n2012-06-03,73,7,155,2012,83.6\n2012-06-04,98,1,156,2012,84.2\n2012-06-05,63,2,157,2012,82.86666666666666\n2012-06-06,66,3,158,2012,80.7\n2012-06-07,57,4,159,2012,80.16666666666667\n2012-06-08,78,5,160,2012,79.26666666666667\n2012-06-09,54,6,161,2012,78.36666666666666\n2012-06-10,122,7,162,2012,77.76666666666667\n2012-06-11,95,1,163,2012,78.16666666666667\n2012-06-12,87,2,164,2012,77.43333333333334\n2012-06-13,113,3,165,2012,78.96666666666667\n2012-06-14,79,4,166,2012,79.23333333333333\n2012-06-15,107,5,167,2012,79.93333333333334\n2012-06-16,106,6,168,2012,79.53333333333333\n2012-06-17,96,7,169,2012,79.36666666666666\n2012-06-18,55,1,170,2012,78.63333333333334\n2012-06-19,79,2,171,2012,77.6\n2012-06-20,94,3,172,2012,78.16666666666667\n2012-06-21,54,4,173,2012,77.26666666666667\n2012-06-22,44,5,174,2012,76.4\n2012-06-23,67,6,175,2012,76.0\n2012-06-24,64,7,176,2012,76.13333333333334\n2012-06-25,96,1,177,2012,77.16666666666667\n2012-06-26,73,2,178,2012,78.06666666666666\n2012-06-27,108,3,179,2012,79.2\n2012-06-28,40,4,180,2012,78.93333333333334\n2012-06-29,87,5,181,2012,78.83333333333333\n2012-06-30,44,6,182,2012,77.13333333333334\n2012-07-01,61,7,183,2012,76.36666666666666\n2012-07-02,91,1,184,2012,78.36666666666666\n2012-07-03,118,2,185,2012,79.86666666666666\n2012-07-04,96,3,186,2012,79.8\n2012-07-05,123,4,187,2012,81.8\n2012-07-06,170,5,188,2012,85.26666666666667\n2012-07-07,117,6,189,2012,87.26666666666667\n2012-07-08,136,7,190,2012,89.2\n2012-07-09,81,1,191,2012,90.1\n2012-07-10,74,2,192,2012,88.5\n2012-07-11,96,3,193,2012,88.53333333333333\n2012-07-12,97,4,194,2012,88.86666666666666\n2012-07-13,74,5,195,2012,87.56666666666666\n2012-07-14,83,6,196,2012,87.7\n2012-07-15,64,7,197,2012,86.26666666666667\n2012-07-16,95,1,198,2012,85.9\n2012-07-17,119,2,199,2012,86.66666666666667\n2012-07-18,102,3,200,2012,88.23333333333333\n2012-07-19,70,4,201,2012,87.93333333333334\n2012-07-20,66,5,202,2012,87.0\n2012-07-21,102,6,203,2012,88.6\n2012-07-22,96,7,204,2012,90.33333333333333\n2012-07-23,107,1,205,2012,91.66666666666667\n2012-07-24,100,2,206,2012,92.86666666666666\n2012-07-25,76,3,207,2012,92.2\n2012-07-26,96,4,208,2012,92.96666666666667\n2012-07-27,81,5,209,2012,92.06666666666666\n2012-07-28,93,6,210,2012,93.83333333333333\n2012-07-29,74,7,211,2012,93.4\n2012-07-30,61,1,212,2012,93.96666666666667\n2012-07-31,73,2,213,2012,94.36666666666666\n2012-08-01,100,3,214,2012,94.66666666666667\n2012-08-02,128,4,215,2012,95.0\n2012-08-03,79,5,216,2012,94.43333333333334\n2012-08-04,106,6,217,2012,93.86666666666666\n2012-08-05,72,7,218,2012,90.6\n2012-08-06,109,1,219,2012,90.33333333333333\n2012-08-07,107,2,220,2012,89.36666666666666\n2012-08-08,119,3,221,2012,90.63333333333334\n2012-08-09,82,4,222,2012,90.9\n2012-08-10,100,5,223,2012,91.03333333333333\n2012-08-11,60,6,224,2012,89.8\n2012-08-12,77,7,225,2012,89.9\n2012-08-13,68,1,226,2012,89.4\n2012-08-14,88,2,227,2012,90.2\n2012-08-15,116,3,228,2012,90.9\n2012-08-16,94,4,229,2012,90.06666666666666\n2012-08-17,99,5,230,2012,89.96666666666667\n2012-08-18,141,6,231,2012,92.33333333333333\n2012-08-19,46,7,232,2012,91.66666666666667\n2012-08-20,61,1,233,2012,90.3\n2012-08-21,75,2,234,2012,89.6\n2012-08-22,84,3,235,2012,88.83333333333333\n2012-08-23,76,4,236,2012,88.03333333333333\n2012-08-24,102,5,237,2012,88.9\n2012-08-25,71,6,238,2012,88.06666666666666\n2012-08-26,83,7,239,2012,88.13333333333334\n2012-08-27,91,1,240,2012,88.06666666666666\n2012-08-28,80,2,241,2012,88.26666666666667\n2012-08-29,60,3,242,2012,88.23333333333333\n2012-08-30,78,4,243,2012,88.4\n2012-08-31,61,5,244,2012,87.1\n2012-09-01,57,6,245,2012,84.73333333333333\n2012-09-02,41,7,246,2012,83.46666666666667\n2012-09-03,86,1,247,2012,82.8\n2012-09-04,88,2,248,2012,83.33333333333333\n2012-09-05,73,3,249,2012,82.13333333333334\n2012-09-06,100,4,250,2012,81.9\n2012-09-07,106,5,251,2012,81.46666666666667\n2012-09-08,57,6,252,2012,80.63333333333334\n2012-09-09,40,7,253,2012,78.63333333333334\n2012-09-10,65,1,254,2012,78.8\n2012-09-11,59,2,255,2012,78.2\n2012-09-12,99,3,256,2012,79.23333333333333\n2012-09-13,55,4,257,2012,78.13333333333334\n2012-09-14,64,5,258,2012,76.4\n2012-09-15,40,6,259,2012,74.6\n2012-09-16,52,7,260,2012,73.03333333333333\n2012-09-17,57,1,261,2012,70.23333333333333\n2012-09-18,51,2,262,2012,70.4\n2012-09-19,38,3,263,2012,69.63333333333334\n2012-09-20,47,4,264,2012,68.7\n2012-09-21,53,5,265,2012,67.66666666666667\n2012-09-22,69,6,266,2012,67.43333333333334\n2012-09-23,40,7,267,2012,65.36666666666666\n2012-09-24,32,1,268,2012,64.06666666666666\n2012-09-25,45,2,269,2012,62.8\n2012-09-26,54,3,270,2012,61.56666666666667\n2012-09-27,59,4,271,2012,60.86666666666667\n2012-09-28,61,5,272,2012,60.9\n2012-09-29,58,6,273,2012,60.233333333333334\n2012-09-30,62,7,274,2012,60.266666666666666\n2012-10-01,96,1,275,2012,61.56666666666667\n2012-10-02,101,2,276,2012,63.56666666666667\n2012-10-03,98,3,277,2012,63.96666666666667\n2012-10-04,78,4,278,2012,63.63333333333333\n2012-10-05,107,5,279,2012,64.76666666666667\n2012-10-06,59,6,280,2012,63.4\n2012-10-07,48,7,281,2012,61.46666666666667\n2012-10-08,60,1,282,2012,61.56666666666667\n2012-10-09,76,2,283,2012,62.766666666666666\n2012-10-10,126,3,284,2012,64.8\n2012-10-11,58,4,285,2012,64.76666666666667\n2012-10-12,70,5,286,2012,63.8\n2012-10-13,37,6,287,2012,63.2\n2012-10-14,70,7,288,2012,63.4\n2012-10-15,79,1,289,2012,64.7\n2012-10-16,77,2,290,2012,65.53333333333333\n2012-10-17,33,3,291,2012,64.73333333333333\n2012-10-18,86,4,292,2012,65.9\n2012-10-19,62,5,293,2012,66.7\n2012-10-20,36,6,294,2012,66.33333333333333\n2012-10-21,18,7,295,2012,65.16666666666667\n2012-10-22,55,1,296,2012,64.7\n2012-10-23,78,2,297,2012,65.96666666666667\n2012-10-24,58,3,298,2012,66.83333333333333\n2012-10-25,87,4,299,2012,68.23333333333333\n2012-10-26,112,5,300,2012,70.16666666666667\n2012-10-27,48,6,301,2012,69.8\n2012-10-28,53,7,302,2012,69.53333333333333\n2012-10-29,38,1,303,2012,68.86666666666666\n2012-10-30,92,2,304,2012,69.86666666666666\n2012-10-31,54,3,305,2012,68.46666666666667\n2012-11-01,109,4,306,2012,68.73333333333333\n2012-11-02,119,5,307,2012,69.43333333333334\n2012-11-03,71,6,308,2012,69.2\n2012-11-04,59,7,309,2012,67.6\n2012-11-05,91,1,310,2012,68.66666666666667\n2012-11-06,39,2,311,2012,68.36666666666666\n2012-11-07,56,3,312,2012,68.23333333333333\n2012-11-08,66,4,313,2012,67.9\n2012-11-09,56,5,314,2012,65.56666666666666\n2012-11-10,39,6,315,2012,64.93333333333334\n2012-11-11,20,7,316,2012,63.266666666666666\n2012-11-12,68,1,317,2012,64.3\n2012-11-13,66,2,318,2012,64.16666666666667\n2012-11-14,58,3,319,2012,63.46666666666667\n2012-11-15,47,4,320,2012,62.46666666666667\n2012-11-16,70,5,321,2012,63.7\n2012-11-17,67,6,322,2012,63.06666666666667\n2012-11-18,38,7,323,2012,62.266666666666666\n2012-11-19,95,1,324,2012,64.23333333333333\n2012-11-20,114,2,325,2012,67.43333333333334\n2012-11-21,87,3,326,2012,68.5\n2012-11-22,49,4,327,2012,67.53333333333333\n2012-11-23,80,5,328,2012,68.26666666666667\n2012-11-24,70,6,329,2012,67.7\n2012-11-25,35,7,330,2012,65.13333333333334\n2012-11-26,71,1,331,2012,65.9\n2012-11-27,79,2,332,2012,66.76666666666667\n2012-11-28,64,3,333,2012,67.63333333333334\n2012-11-29,58,4,334,2012,66.5\n2012-11-30,36,5,335,2012,65.9\n2012-12-01,56,6,336,2012,64.13333333333334\n2012-12-02,58,7,337,2012,62.1\n2012-12-03,84,1,338,2012,62.53333333333333\n2012-12-04,103,2,339,2012,64.0\n2012-12-05,118,3,340,2012,64.9\n2012-12-06,46,4,341,2012,65.13333333333334\n2012-12-07,58,5,342,2012,65.2\n2012-12-08,98,6,343,2012,66.26666666666667\n2012-12-09,43,7,344,2012,65.83333333333333\n2012-12-10,70,1,345,2012,66.86666666666666\n2012-12-11,91,2,346,2012,69.23333333333333\n2012-12-12,74,3,347,2012,69.43333333333334\n2012-12-13,88,4,348,2012,70.16666666666667\n2012-12-14,74,5,349,2012,70.7\n2012-12-15,54,6,350,2012,70.93333333333334\n2012-12-16,81,7,351,2012,71.3\n2012-12-17,121,1,352,2012,73.1\n2012-12-18,111,2,353,2012,75.53333333333333\n2012-12-19,112,3,354,2012,76.1\n2012-12-20,103,4,355,2012,75.73333333333333\n2012-12-21,63,5,356,2012,74.93333333333334\n2012-12-22,59,6,357,2012,75.26666666666667\n2012-12-23,91,7,358,2012,75.63333333333334\n2012-12-24,99,1,359,2012,76.6\n2012-12-25,60,2,360,2012,77.43333333333334\n2012-12-26,61,3,361,2012,77.1\n2012-12-27,90,4,362,2012,77.46666666666667\n2012-12-28,59,5,363,2012,77.3\n2012-12-29,69,6,364,2012,77.66666666666667\n2012-12-30,66,7,365,2012,78.66666666666667\n2012-12-31,110,1,366,2012,80.46666666666667\n2013-01-01,65,2,1,2013,80.7\n2013-01-02,73,3,2,2013,80.33333333333333\n2013-01-03,92,4,3,2013,79.96666666666667\n2013-01-04,93,5,4,2013,79.13333333333334\n2013-01-05,67,6,5,2013,79.83333333333333\n2013-01-06,84,7,6,2013,80.7\n2013-01-07,176,1,7,2013,83.3\n2013-01-08,150,2,8,2013,86.86666666666666\n2013-01-09,143,3,9,2013,89.3\n2013-01-10,195,4,10,2013,92.76666666666667\n2013-01-11,158,5,11,2013,95.56666666666666\n2013-01-12,99,6,12,2013,95.93333333333334\n2013-01-13,156,7,13,2013,98.66666666666667\n2013-01-14,114,1,14,2013,100.66666666666667\n2013-01-15,109,2,15,2013,101.6\n2013-01-16,182,3,16,2013,103.63333333333334\n2013-01-17,139,4,17,2013,104.56666666666666\n2013-01-18,96,5,18,2013,104.03333333333333\n2013-01-19,82,6,19,2013,103.33333333333333\n2013-01-20,116,7,20,2013,105.1\n2013-01-21,88,1,21,2013,106.06666666666666\n2013-01-22,124,2,22,2013,107.16666666666667\n2013-01-23,120,3,23,2013,107.86666666666666\n2013-01-24,108,4,24,2013,109.46666666666667\n2013-01-25,143,5,25,2013,112.2\n2013-01-26,142,6,26,2013,113.93333333333334\n2013-01-27,69,7,27,2013,114.26666666666667\n2013-01-28,128,1,28,2013,116.23333333333333\n2013-01-29,114,2,29,2013,117.83333333333333\n2013-01-30,120,3,30,2013,118.16666666666667\n2013-01-31,96,4,31,2013,119.2\n2013-02-01,69,5,32,2013,119.06666666666666\n2013-02-02,104,6,33,2013,119.46666666666667\n2013-02-03,156,7,34,2013,121.56666666666666\n2013-02-04,90,1,35,2013,122.33333333333333\n2013-02-05,116,2,36,2013,123.4\n2013-02-06,81,3,37,2013,120.23333333333333\n2013-02-07,117,4,38,2013,119.13333333333334\n2013-02-08,123,5,39,2013,118.46666666666667\n2013-02-09,80,6,40,2013,114.63333333333334\n2013-02-10,83,7,41,2013,112.13333333333334\n2013-02-11,100,1,42,2013,112.16666666666667\n2013-02-12,115,2,43,2013,110.8\n2013-02-13,153,3,44,2013,112.1\n2013-02-14,95,4,45,2013,111.63333333333334\n2013-02-15,146,5,46,2013,110.43333333333334\n2013-02-16,111,6,47,2013,109.5\n2013-02-17,118,7,48,2013,110.23333333333333\n2013-02-18,119,1,49,2013,111.46666666666667\n2013-02-19,120,2,50,2013,111.6\n2013-02-20,142,3,51,2013,113.4\n2013-02-21,127,4,52,2013,113.5\n2013-02-22,120,5,53,2013,113.5\n2013-02-23,44,6,54,2013,111.36666666666666\n2013-02-24,131,7,55,2013,110.96666666666667\n2013-02-25,166,1,56,2013,111.76666666666667\n2013-02-26,176,2,57,2013,115.33333333333333\n2013-02-27,113,3,58,2013,114.83333333333333\n2013-02-28,116,4,59,2013,114.9\n2013-03-01,112,5,60,2013,114.63333333333334\n2013-03-02,72,6,61,2013,113.83333333333333\n2013-03-03,73,7,62,2013,113.96666666666667\n2013-03-04,123,1,63,2013,114.6\n2013-03-05,111,2,64,2013,113.1\n2013-03-06,102,3,65,2013,113.5\n2013-03-07,99,4,66,2013,112.93333333333334\n2013-03-08,109,5,67,2013,113.86666666666666\n2013-03-09,108,6,68,2013,113.56666666666666\n2013-03-10,130,7,69,2013,113.8\n2013-03-11,121,1,70,2013,115.16666666666667\n2013-03-12,85,2,71,2013,115.23333333333333\n2013-03-13,118,3,72,2013,115.83333333333333\n2013-03-14,129,4,73,2013,116.3\n2013-03-15,139,5,74,2013,115.83333333333333\n2013-03-16,98,6,75,2013,115.93333333333334\n2013-03-17,137,7,76,2013,115.63333333333334\n2013-03-18,139,1,77,2013,116.56666666666666\n2013-03-19,114,2,78,2013,116.43333333333334\n2013-03-20,153,3,79,2013,117.56666666666666\n2013-03-21,88,4,80,2013,116.5\n2013-03-22,151,5,81,2013,116.8\n2013-03-23,105,6,82,2013,116.06666666666666\n2013-03-24,107,7,83,2013,115.63333333333334\n2013-03-25,137,1,84,2013,118.73333333333333\n2013-03-26,114,2,85,2013,118.16666666666667\n2013-03-27,107,3,86,2013,116.2\n2013-03-28,113,4,87,2013,114.1\n2013-03-29,150,5,88,2013,115.33333333333333\n2013-03-30,83,6,89,2013,114.23333333333333\n2013-03-31,95,7,90,2013,113.66666666666667\n2013-04-01,124,1,91,2013,115.4\n2013-04-02,132,2,92,2013,117.36666666666666\n2013-04-03,116,3,93,2013,117.13333333333334\n2013-04-04,135,4,94,2013,117.93333333333334\n2013-04-05,80,5,95,2013,117.2\n2013-04-06,90,6,96,2013,116.9\n2013-04-07,149,7,97,2013,118.23333333333333\n2013-04-08,178,1,98,2013,120.56666666666666\n2013-04-09,164,2,99,2013,121.7\n2013-04-10,120,3,100,2013,121.66666666666667\n2013-04-11,111,4,101,2013,122.53333333333333\n2013-04-12,138,5,102,2013,123.2\n2013-04-13,123,6,103,2013,123.0\n2013-04-14,132,7,104,2013,122.76666666666667\n2013-04-15,111,1,105,2013,123.2\n2013-04-16,90,2,106,2013,121.63333333333334\n2013-04-17,90,3,107,2013,120.0\n2013-04-18,143,4,108,2013,120.96666666666667\n2013-04-19,110,5,109,2013,119.53333333333333\n2013-04-20,79,6,110,2013,119.23333333333333\n2013-04-21,128,7,111,2013,118.46666666666667\n2013-04-22,93,1,112,2013,118.06666666666666\n2013-04-23,70,2,113,2013,116.83333333333333\n2013-04-24,85,3,114,2013,115.1\n2013-04-25,83,4,115,2013,114.06666666666666\n2013-04-26,95,5,116,2013,113.66666666666667\n2013-04-27,82,6,117,2013,112.63333333333334\n2013-04-28,96,7,118,2013,110.83333333333333\n2013-04-29,91,1,119,2013,111.1\n2013-04-30,130,2,120,2013,112.26666666666667\n2013-05-01,117,3,121,2013,112.03333333333333\n2013-05-02,134,4,122,2013,112.1\n2013-05-03,113,5,123,2013,112.0\n2013-05-04,97,6,124,2013,110.73333333333333\n2013-05-05,124,7,125,2013,112.2\n2013-05-06,120,1,126,2013,113.2\n2013-05-07,98,2,127,2013,111.5\n2013-05-08,112,3,128,2013,109.3\n2013-05-09,155,4,129,2013,109.0\n2013-05-10,135,5,130,2013,109.5\n2013-05-11,88,6,131,2013,108.73333333333333\n2013-05-12,64,7,132,2013,106.26666666666667\n2013-05-13,123,1,133,2013,106.26666666666667\n2013-05-14,88,2,134,2013,104.8\n2013-05-15,162,3,135,2013,106.5\n2013-05-16,136,4,136,2013,108.03333333333333\n2013-05-17,143,5,137,2013,109.8\n2013-05-18,71,6,138,2013,107.4\n2013-05-19,89,7,139,2013,106.7\n2013-05-20,106,1,140,2013,107.6\n2013-05-21,136,2,141,2013,107.86666666666666\n2013-05-22,115,3,142,2013,108.6\n2013-05-23,91,4,143,2013,109.3\n2013-05-24,148,5,144,2013,111.4\n2013-05-25,70,6,145,2013,110.96666666666667\n2013-05-26,47,7,146,2013,109.36666666666666\n2013-05-27,135,1,147,2013,111.13333333333334\n2013-05-28,150,2,148,2013,112.93333333333334\n2013-05-29,88,3,149,2013,112.83333333333333\n2013-05-30,102,4,150,2013,111.9\n2013-05-31,82,5,151,2013,110.73333333333333\n2013-06-01,79,6,152,2013,108.9\n2013-06-02,86,7,153,2013,108.0\n2013-06-03,87,1,154,2013,107.66666666666667\n2013-06-04,100,2,155,2013,106.86666666666666\n2013-06-05,121,3,156,2013,106.9\n2013-06-06,112,4,157,2013,107.36666666666666\n2013-06-07,167,5,158,2013,109.2\n2013-06-08,103,6,159,2013,107.46666666666667\n2013-06-09,83,7,160,2013,105.73333333333333\n2013-06-10,132,1,161,2013,107.2\n2013-06-11,109,2,162,2013,108.7\n2013-06-12,133,3,163,2013,109.03333333333333\n2013-06-13,160,4,164,2013,111.43333333333334\n2013-06-14,178,5,165,2013,111.96666666666667\n2013-06-15,73,6,166,2013,109.86666666666666\n2013-06-16,138,7,167,2013,109.7\n2013-06-17,111,1,168,2013,111.03333333333333\n2013-06-18,112,2,169,2013,111.8\n2013-06-19,126,3,170,2013,112.46666666666667\n2013-06-20,146,4,171,2013,112.8\n2013-06-21,122,5,172,2013,113.03333333333333\n2013-06-22,122,6,173,2013,114.06666666666666\n2013-06-23,98,7,174,2013,112.4\n2013-06-24,86,1,175,2013,112.93333333333334\n2013-06-25,128,2,176,2013,115.63333333333334\n2013-06-26,160,3,177,2013,116.46666666666667\n2013-06-27,115,4,178,2013,115.3\n2013-06-28,103,5,179,2013,115.8\n2013-06-29,57,6,180,2013,114.3\n2013-06-30,76,7,181,2013,114.1\n2013-07-01,132,1,182,2013,115.86666666666666\n2013-07-02,134,2,183,2013,117.46666666666667\n2013-07-03,95,3,184,2013,117.73333333333333\n2013-07-04,62,4,185,2013,116.46666666666667\n2013-07-05,131,5,186,2013,116.8\n2013-07-06,90,6,187,2013,116.06666666666666\n2013-07-07,92,7,188,2013,113.56666666666666\n2013-07-08,142,1,189,2013,114.86666666666666\n2013-07-09,111,2,190,2013,115.8\n2013-07-10,109,3,191,2013,115.03333333333333\n2013-07-11,93,4,192,2013,114.5\n2013-07-12,104,5,193,2013,113.53333333333333\n2013-07-13,39,6,194,2013,109.5\n2013-07-14,67,7,195,2013,105.8\n2013-07-15,125,1,196,2013,107.53333333333333\n2013-07-16,114,2,197,2013,106.73333333333333\n2013-07-17,89,3,198,2013,106.0\n2013-07-18,107,4,199,2013,105.83333333333333\n2013-07-19,142,5,200,2013,106.36666666666666\n2013-07-20,50,6,201,2013,103.16666666666667\n2013-07-21,58,7,202,2013,101.03333333333333\n2013-07-22,166,1,203,2013,102.5\n2013-07-23,149,2,204,2013,104.2\n2013-07-24,169,3,205,2013,106.96666666666667\n2013-07-25,71,4,206,2013,105.06666666666666\n2013-07-26,160,5,207,2013,105.06666666666666\n2013-07-27,50,6,208,2013,102.9\n2013-07-28,100,7,209,2013,102.8\n2013-07-29,108,1,210,2013,104.5\n2013-07-30,206,2,211,2013,108.83333333333333\n2013-07-31,181,3,212,2013,110.46666666666667\n2013-08-01,188,4,213,2013,112.26666666666667\n2013-08-02,143,5,214,2013,113.86666666666666\n2013-08-03,121,6,215,2013,115.83333333333333\n2013-08-04,82,7,216,2013,114.2\n2013-08-05,117,1,217,2013,115.1\n2013-08-06,138,2,218,2013,116.63333333333334\n2013-08-07,143,3,219,2013,116.66666666666667\n2013-08-08,117,4,220,2013,116.86666666666666\n2013-08-09,149,5,221,2013,118.2\n2013-08-10,69,6,222,2013,117.4\n2013-08-11,68,7,223,2013,116.2\n2013-08-12,98,1,224,2013,118.16666666666667\n2013-08-13,89,2,225,2013,118.9\n2013-08-14,108,3,226,2013,118.33333333333333\n2013-08-15,115,4,227,2013,118.36666666666666\n2013-08-16,99,5,228,2013,118.7\n2013-08-17,62,6,229,2013,117.2\n2013-08-18,66,7,230,2013,114.66666666666667\n2013-08-19,157,1,231,2013,118.23333333333333\n2013-08-20,126,2,232,2013,120.5\n2013-08-21,129,3,233,2013,119.26666666666667\n2013-08-22,122,4,234,2013,118.36666666666666\n2013-08-23,124,5,235,2013,116.86666666666666\n2013-08-24,101,6,236,2013,117.86666666666666\n2013-08-25,109,7,237,2013,116.16666666666667\n2013-08-26,175,1,238,2013,120.33333333333333\n2013-08-27,155,2,239,2013,122.16666666666667\n2013-08-28,159,3,240,2013,123.86666666666666\n2013-08-29,116,4,241,2013,120.86666666666666\n2013-08-30,133,5,242,2013,119.26666666666667\n2013-08-31,112,6,243,2013,116.73333333333333\n2013-09-01,103,7,244,2013,115.4\n2013-09-02,83,1,245,2013,114.13333333333334\n2013-09-03,115,2,246,2013,115.23333333333333\n2013-09-04,141,3,247,2013,116.03333333333333\n2013-09-05,115,4,248,2013,115.26666666666667\n2013-09-06,77,5,249,2013,113.06666666666666\n2013-09-07,99,6,250,2013,112.46666666666667\n2013-09-08,93,7,251,2013,110.6\n2013-09-09,163,1,252,2013,113.73333333333333\n2013-09-10,143,2,253,2013,116.23333333333333\n2013-09-11,86,3,254,2013,115.83333333333333\n2013-09-12,137,4,255,2013,117.43333333333334\n2013-09-13,135,5,256,2013,118.33333333333333\n2013-09-14,83,6,257,2013,117.26666666666667\n2013-09-15,130,7,258,2013,118.3\n2013-09-16,123,1,259,2013,120.33333333333333\n2013-09-17,104,2,260,2013,121.6\n2013-09-18,97,3,261,2013,119.6\n2013-09-19,104,4,262,2013,118.86666666666666\n2013-09-20,118,5,263,2013,118.5\n2013-09-21,88,6,264,2013,117.36666666666666\n2013-09-22,88,7,265,2013,116.16666666666667\n2013-09-23,105,1,266,2013,116.3\n2013-09-24,219,2,267,2013,119.96666666666667\n2013-09-25,188,3,268,2013,120.4\n2013-09-26,108,4,269,2013,118.83333333333333\n2013-09-27,123,5,270,2013,117.63333333333334\n2013-09-28,81,6,271,2013,116.46666666666667\n2013-09-29,83,7,272,2013,114.8\n2013-09-30,99,1,273,2013,114.36666666666666\n2013-10-01,80,2,274,2013,113.6\n2013-10-02,99,3,275,2013,114.13333333333334\n2013-10-03,110,4,276,2013,113.96666666666667\n2013-10-04,121,5,277,2013,113.3\n2013-10-05,103,6,278,2013,112.9\n2013-10-06,141,7,279,2013,115.03333333333333\n2013-10-07,155,1,280,2013,116.9\n2013-10-08,177,2,281,2013,119.7\n2013-10-09,117,3,282,2013,118.16666666666667\n2013-10-10,117,4,283,2013,117.3\n2013-10-11,202,5,284,2013,121.16666666666667\n2013-10-12,103,6,285,2013,120.03333333333333\n2013-10-13,140,7,286,2013,120.2\n2013-10-14,94,1,287,2013,120.56666666666666\n2013-10-15,128,2,288,2013,120.5\n2013-10-16,167,3,289,2013,121.96666666666667\n2013-10-17,96,4,290,2013,121.7\n2013-10-18,148,5,291,2013,123.4\n2013-10-19,89,6,292,2013,122.9\n2013-10-20,122,7,293,2013,123.03333333333333\n2013-10-21,113,1,294,2013,123.86666666666666\n2013-10-22,160,2,295,2013,126.26666666666667\n2013-10-23,115,3,296,2013,126.6\n2013-10-24,134,4,297,2013,123.76666666666667\n2013-10-25,160,5,298,2013,122.83333333333333\n2013-10-26,58,6,299,2013,121.16666666666667\n2013-10-27,110,7,300,2013,120.73333333333333\n2013-10-28,191,1,301,2013,124.4\n2013-10-29,113,2,302,2013,125.4\n2013-10-30,169,3,303,2013,127.73333333333333\n2013-10-31,143,4,304,2013,129.83333333333334\n2013-11-01,119,5,305,2013,130.5\n2013-11-02,161,6,306,2013,132.2\n2013-11-03,104,7,307,2013,131.63333333333333\n2013-11-04,103,1,308,2013,131.63333333333333\n2013-11-05,165,2,309,2013,132.43333333333334\n2013-11-06,121,3,310,2013,131.3\n2013-11-07,156,4,311,2013,130.6\n2013-11-08,177,5,312,2013,132.6\n2013-11-09,85,6,313,2013,131.53333333333333\n2013-11-10,81,7,314,2013,127.5\n2013-11-11,194,1,315,2013,130.53333333333333\n2013-11-12,180,2,316,2013,131.86666666666667\n2013-11-13,163,3,317,2013,134.16666666666666\n2013-11-14,106,4,318,2013,133.43333333333334\n2013-11-15,97,5,319,2013,131.1\n2013-11-16,39,6,320,2013,129.2\n2013-11-17,93,7,321,2013,127.36666666666666\n2013-11-18,127,1,322,2013,128.63333333333333\n2013-11-19,129,2,323,2013,128.86666666666667\n2013-11-20,126,3,324,2013,129.3\n2013-11-21,109,4,325,2013,127.6\n2013-11-22,124,5,326,2013,127.9\n2013-11-23,76,6,327,2013,125.96666666666667\n2013-11-24,107,7,328,2013,124.2\n2013-11-25,117,1,329,2013,126.16666666666667\n2013-11-26,136,2,330,2013,127.03333333333333\n2013-11-27,233,3,331,2013,128.43333333333334\n2013-11-28,37,4,332,2013,125.9\n2013-11-29,94,5,333,2013,123.4\n2013-11-30,138,6,334,2013,123.23333333333333\n2013-12-01,94,7,335,2013,122.4\n2013-12-02,119,1,336,2013,121.0\n2013-12-03,152,2,337,2013,122.6\n2013-12-04,123,3,338,2013,123.26666666666667\n2013-12-05,141,4,339,2013,122.46666666666667\n2013-12-06,144,5,340,2013,123.23333333333333\n2013-12-07,131,6,341,2013,122.4\n2013-12-08,116,7,342,2013,120.36666666666666\n2013-12-09,149,1,343,2013,122.5\n2013-12-10,173,2,344,2013,125.56666666666666\n2013-12-11,118,3,345,2013,123.03333333333333\n2013-12-12,172,4,346,2013,122.76666666666667\n2013-12-13,165,5,347,2013,122.83333333333333\n2013-12-14,115,6,348,2013,123.13333333333334\n2013-12-15,84,7,349,2013,122.7\n2013-12-16,122,1,350,2013,125.46666666666667\n2013-12-17,150,2,351,2013,127.36666666666666\n2013-12-18,124,3,352,2013,127.26666666666667\n2013-12-19,140,4,353,2013,127.63333333333334\n2013-12-20,131,5,354,2013,127.8\n2013-12-21,106,6,355,2013,127.7\n2013-12-22,97,7,356,2013,126.8\n2013-12-23,82,1,357,2013,127.0\n2013-12-24,50,2,358,2013,125.1\n2013-12-25,29,3,359,2013,122.16666666666667\n2013-12-26,60,4,360,2013,119.63333333333334\n2013-12-27,98,5,361,2013,115.13333333333334\n2013-12-28,66,6,362,2013,116.1\n2013-12-29,106,7,363,2013,116.5\n2013-12-30,173,1,364,2013,117.66666666666667\n2013-12-31,167,2,365,2013,120.1\n2014-01-01,74,3,1,2014,118.6\n2014-01-02,128,4,2,2014,117.8\n2014-01-03,201,5,3,2014,120.4\n2014-01-04,123,6,4,2014,119.8\n2014-01-05,93,7,5,2014,118.1\n2014-01-06,155,1,6,2014,118.9\n2014-01-07,237,2,7,2014,122.93333333333334\n2014-01-08,207,3,8,2014,124.86666666666666\n2014-01-09,202,4,9,2014,125.83333333333333\n2014-01-10,203,5,10,2014,128.66666666666666\n2014-01-11,124,6,11,2014,127.06666666666666\n2014-01-12,112,7,12,2014,125.3\n2014-01-13,135,1,13,2014,125.96666666666667\n2014-01-14,119,2,14,2014,127.13333333333334\n2014-01-15,186,3,15,2014,129.26666666666668\n2014-01-16,197,4,16,2014,130.83333333333334\n2014-01-17,107,5,17,2014,130.26666666666668\n2014-01-18,105,6,18,2014,129.1\n2014-01-19,150,7,19,2014,129.73333333333332\n2014-01-20,116,1,20,2014,130.06666666666666\n2014-01-21,143,2,21,2014,131.6\n2014-01-22,164,3,22,2014,134.33333333333334\n2014-01-23,234,4,23,2014,140.46666666666667\n2014-01-24,136,5,24,2014,144.03333333333333\n2014-01-25,123,6,25,2014,146.13333333333333\n2014-01-26,102,7,26,2014,146.26666666666668\n2014-01-27,164,1,27,2014,149.53333333333333\n2014-01-28,166,2,28,2014,151.53333333333333\n2014-01-29,151,3,29,2014,150.8\n2014-01-30,136,4,30,2014,149.76666666666668\n2014-01-31,102,5,31,2014,150.7\n2014-02-01,73,6,32,2014,148.86666666666667\n2014-02-02,122,7,33,2014,146.23333333333332\n2014-02-03,167,1,34,2014,147.7\n2014-02-04,209,2,35,2014,151.56666666666666\n2014-02-05,171,3,36,2014,152.1\n2014-02-06,161,4,37,2014,149.56666666666666\n2014-02-07,137,5,38,2014,147.23333333333332\n2014-02-08,131,6,39,2014,144.86666666666667\n2014-02-09,208,7,40,2014,145.03333333333333\n2014-02-10,227,1,41,2014,148.46666666666667\n2014-02-11,212,2,42,2014,151.8\n2014-02-12,135,3,43,2014,151.8\n2014-02-13,191,4,44,2014,154.2\n2014-02-14,161,5,45,2014,153.36666666666667\n2014-02-15,152,6,46,2014,151.86666666666667\n2014-02-16,101,7,47,2014,151.66666666666666\n2014-02-17,184,1,48,2014,154.3\n2014-02-18,307,2,49,2014,159.53333333333333\n2014-02-19,165,3,50,2014,161.16666666666666\n2014-02-20,122,4,51,2014,160.46666666666667\n2014-02-21,183,5,52,2014,161.1\n2014-02-22,139,6,53,2014,157.93333333333334\n2014-02-23,132,7,54,2014,157.8\n2014-02-24,181,1,55,2014,159.73333333333332\n2014-02-25,173,2,56,2014,162.1\n2014-02-26,161,3,57,2014,162.0\n2014-02-27,113,4,58,2014,160.23333333333332\n2014-02-28,103,5,59,2014,158.63333333333333\n2014-03-01,91,6,60,2014,157.13333333333333\n2014-03-02,76,7,61,2014,156.26666666666668\n2014-03-03,126,1,62,2014,158.03333333333333\n2014-03-04,109,2,63,2014,157.6\n2014-03-05,119,3,64,2014,156.0\n2014-03-06,162,4,65,2014,154.43333333333334\n2014-03-07,115,5,66,2014,152.56666666666666\n2014-03-08,118,6,67,2014,151.13333333333333\n2014-03-09,73,7,68,2014,149.0\n2014-03-10,140,1,69,2014,149.3\n2014-03-11,158,2,70,2014,147.63333333333333\n2014-03-12,119,3,71,2014,144.03333333333333\n2014-03-13,139,4,72,2014,141.6\n2014-03-14,179,5,73,2014,143.06666666666666\n2014-03-15,110,6,74,2014,140.36666666666667\n2014-03-16,80,7,75,2014,137.66666666666666\n2014-03-17,144,1,76,2014,137.4\n2014-03-18,114,2,77,2014,137.83333333333334\n2014-03-19,132,3,78,2014,136.1\n2014-03-20,129,4,79,2014,130.16666666666666\n2014-03-21,167,5,80,2014,130.23333333333332\n2014-03-22,132,6,81,2014,130.56666666666666\n2014-03-23,102,7,82,2014,127.86666666666666\n2014-03-24,187,1,83,2014,129.46666666666667\n2014-03-25,148,2,84,2014,130.0\n2014-03-26,223,3,85,2014,131.4\n2014-03-27,145,4,86,2014,130.46666666666667\n2014-03-28,162,5,87,2014,130.5\n2014-03-29,91,6,88,2014,129.76666666666668\n2014-03-30,100,7,89,2014,129.66666666666666\n2014-03-31,147,1,90,2014,131.53333333333333\n2014-04-01,116,2,91,2014,132.86666666666667\n2014-04-02,154,3,92,2014,133.8\n2014-04-03,107,4,93,2014,133.73333333333332\n2014-04-04,103,5,94,2014,133.2\n2014-04-05,94,6,95,2014,130.93333333333334\n2014-04-06,86,7,96,2014,129.96666666666667\n2014-04-07,140,1,97,2014,130.7\n2014-04-08,142,2,98,2014,133.0\n2014-04-09,118,3,99,2014,132.26666666666668\n2014-04-10,149,4,100,2014,131.96666666666667\n2014-04-11,90,5,101,2014,131.0\n2014-04-12,82,6,102,2014,129.1\n2014-04-13,94,7,103,2014,126.26666666666667\n2014-04-14,76,1,104,2014,125.13333333333334\n2014-04-15,128,2,105,2014,126.73333333333333\n2014-04-16,157,3,106,2014,127.16666666666667\n2014-04-17,130,4,107,2014,127.7\n2014-04-18,191,5,108,2014,129.66666666666666\n2014-04-19,127,6,109,2014,129.6\n2014-04-20,102,7,110,2014,127.43333333333334\n2014-04-21,179,1,111,2014,129.0\n2014-04-22,153,2,112,2014,130.7\n2014-04-23,167,3,113,2014,130.03333333333333\n2014-04-24,232,4,114,2014,132.83333333333334\n2014-04-25,182,5,115,2014,131.46666666666667\n2014-04-26,119,6,116,2014,130.6\n2014-04-27,187,7,117,2014,131.43333333333334\n2014-04-28,161,1,118,2014,133.76666666666668\n2014-04-29,190,2,119,2014,136.76666666666668\n2014-04-30,170,3,120,2014,137.53333333333333\n2014-05-01,137,4,121,2014,138.23333333333332\n2014-05-02,151,5,122,2014,138.13333333333333\n2014-05-03,107,6,123,2014,138.13333333333333\n2014-05-04,108,7,124,2014,138.3\n2014-05-05,193,1,125,2014,141.6\n2014-05-06,114,2,126,2014,142.53333333333333\n2014-05-07,247,3,127,2014,146.1\n2014-05-08,230,4,128,2014,149.03333333333333\n2014-05-09,264,5,129,2014,153.9\n2014-05-10,175,6,130,2014,154.76666666666668\n2014-05-11,127,7,131,2014,156.0\n2014-05-12,144,1,132,2014,158.06666666666666\n2014-05-13,212,2,133,2014,162.0\n2014-05-14,236,3,134,2014,167.33333333333334\n2014-05-15,154,4,135,2014,168.2\n2014-05-16,155,5,136,2014,168.13333333333333\n2014-05-17,140,6,137,2014,168.46666666666667\n2014-05-18,105,7,138,2014,165.6\n2014-05-19,145,1,139,2014,166.2\n2014-05-20,156,2,140,2014,168.0\n2014-05-21,149,3,141,2014,167.0\n2014-05-22,190,4,142,2014,168.23333333333332\n2014-05-23,196,5,143,2014,169.2\n2014-05-24,102,6,144,2014,164.86666666666667\n2014-05-25,150,7,145,2014,163.8\n2014-05-26,124,1,146,2014,163.96666666666667\n2014-05-27,134,2,147,2014,162.2\n2014-05-28,148,3,148,2014,161.76666666666668\n2014-05-29,185,4,149,2014,161.6\n2014-05-30,174,5,150,2014,161.73333333333332\n2014-05-31,119,6,151,2014,161.13333333333333\n2014-06-01,182,7,152,2014,162.16666666666666\n2014-06-02,133,1,153,2014,163.03333333333333\n2014-06-03,195,2,154,2014,165.93333333333334\n2014-06-04,183,3,155,2014,165.6\n2014-06-05,186,4,156,2014,168.0\n2014-06-06,254,5,157,2014,168.23333333333332\n2014-06-07,141,6,158,2014,165.26666666666668\n2014-06-08,130,7,159,2014,160.8\n2014-06-09,230,1,160,2014,162.63333333333333\n2014-06-10,203,2,161,2014,165.16666666666666\n2014-06-11,164,3,162,2014,165.83333333333334\n2014-06-12,212,4,163,2014,165.83333333333334\n2014-06-13,169,5,164,2014,163.6\n2014-06-14,129,6,165,2014,162.76666666666668\n2014-06-15,100,7,166,2014,160.93333333333334\n2014-06-16,160,1,167,2014,161.6\n2014-06-17,175,2,168,2014,163.93333333333334\n2014-06-18,154,3,169,2014,164.23333333333332\n2014-06-19,197,4,170,2014,165.6\n2014-06-20,146,5,171,2014,165.5\n2014-06-21,44,6,172,2014,160.63333333333333\n2014-06-22,81,7,173,2014,156.8\n2014-06-23,153,1,174,2014,158.5\n2014-06-24,174,2,175,2014,159.3\n2014-06-25,174,3,176,2014,160.96666666666667\n2014-06-26,168,4,177,2014,162.1\n2014-06-27,173,5,178,2014,162.93333333333334\n2014-06-28,96,6,179,2014,159.96666666666667\n2014-06-29,63,7,180,2014,156.26666666666668\n2014-06-30,139,1,181,2014,156.93333333333334\n2014-07-01,160,2,182,2014,156.2\n2014-07-02,179,3,183,2014,157.73333333333332\n2014-07-03,151,4,184,2014,156.26666666666668\n2014-07-04,134,5,185,2014,154.63333333333333\n2014-07-05,121,6,186,2014,152.46666666666667\n2014-07-06,144,7,187,2014,148.8\n2014-07-07,151,1,188,2014,149.13333333333333\n2014-07-08,120,2,189,2014,148.8\n2014-07-09,160,3,190,2014,146.46666666666667\n2014-07-10,160,4,191,2014,145.03333333333333\n2014-07-11,198,5,192,2014,146.16666666666666\n2014-07-12,116,6,193,2014,142.96666666666667\n2014-07-13,125,7,194,2014,141.5\n2014-07-14,115,1,195,2014,141.03333333333333\n2014-07-15,127,2,196,2014,141.93333333333334\n2014-07-16,242,3,197,2014,144.66666666666666\n2014-07-17,147,4,198,2014,143.73333333333332\n2014-07-18,199,5,199,2014,145.23333333333332\n2014-07-19,103,6,200,2014,142.1\n2014-07-20,159,7,201,2014,142.53333333333333\n2014-07-21,241,1,202,2014,149.1\n2014-07-22,189,2,203,2014,152.7\n2014-07-23,149,3,204,2014,152.56666666666666\n2014-07-24,215,4,205,2014,153.93333333333334\n2014-07-25,136,5,206,2014,152.66666666666666\n2014-07-26,105,6,207,2014,150.56666666666666\n2014-07-27,80,7,208,2014,147.46666666666667\n2014-07-28,126,1,209,2014,148.46666666666667\n2014-07-29,176,2,210,2014,152.23333333333332\n2014-07-30,145,3,211,2014,152.43333333333334\n2014-07-31,108,4,212,2014,150.7\n2014-08-01,205,5,213,2014,151.56666666666666\n2014-08-02,125,6,214,2014,150.7\n2014-08-03,93,7,215,2014,149.33333333333334\n2014-08-04,162,1,216,2014,150.7\n2014-08-05,229,2,217,2014,153.53333333333333\n2014-08-06,253,3,218,2014,156.93333333333334\n2014-08-07,226,4,219,2014,160.46666666666667\n2014-08-08,177,5,220,2014,161.03333333333333\n2014-08-09,133,6,221,2014,160.13333333333333\n2014-08-10,164,7,222,2014,159.0\n2014-08-11,199,1,223,2014,161.76666666666668\n2014-08-12,162,2,224,2014,163.0\n2014-08-13,121,3,225,2014,163.2\n2014-08-14,157,4,226,2014,164.2\n2014-08-15,123,5,227,2014,160.23333333333332\n2014-08-16,96,6,228,2014,158.53333333333333\n2014-08-17,139,7,229,2014,156.53333333333333\n2014-08-18,134,1,230,2014,157.56666666666666\n2014-08-19,186,2,231,2014,158.46666666666667\n2014-08-20,182,3,232,2014,156.5\n2014-08-21,205,4,233,2014,157.03333333333333\n2014-08-22,169,5,234,2014,157.7\n2014-08-23,93,6,235,2014,153.63333333333333\n2014-08-24,110,7,236,2014,152.76666666666668\n2014-08-25,167,1,237,2014,154.83333333333334\n2014-08-26,218,2,238,2014,159.43333333333334\n2014-08-27,188,3,239,2014,161.5\n2014-08-28,237,4,240,2014,163.53333333333333\n2014-08-29,163,5,241,2014,164.13333333333333\n2014-08-30,126,6,242,2014,164.73333333333332\n2014-08-31,127,7,243,2014,162.13333333333333\n2014-09-01,181,1,244,2014,164.0\n2014-09-02,240,2,245,2014,168.9\n2014-09-03,195,3,246,2014,170.0\n2014-09-04,181,4,247,2014,168.4\n2014-09-05,177,5,248,2014,165.86666666666667\n2014-09-06,119,6,249,2014,162.3\n2014-09-07,226,7,250,2014,163.93333333333334\n2014-09-08,133,1,251,2014,163.93333333333334\n2014-09-09,176,2,252,2014,164.33333333333334\n2014-09-10,207,3,253,2014,164.6\n2014-09-11,146,4,254,2014,164.06666666666666\n2014-09-12,145,5,255,2014,164.86666666666667\n2014-09-13,89,6,256,2014,162.6\n2014-09-14,125,7,257,2014,162.66666666666666\n2014-09-15,205,1,258,2014,166.3\n2014-09-16,189,2,259,2014,167.96666666666667\n2014-09-17,165,3,260,2014,169.0\n2014-09-18,232,4,261,2014,170.53333333333333\n2014-09-19,192,5,262,2014,170.86666666666667\n2014-09-20,124,6,263,2014,168.16666666666666\n2014-09-21,143,7,264,2014,167.3\n2014-09-22,171,1,265,2014,169.9\n2014-09-23,192,2,266,2014,172.63333333333333\n2014-09-24,207,3,267,2014,173.96666666666667\n2014-09-25,183,4,268,2014,172.8\n2014-09-26,201,5,269,2014,173.23333333333332\n2014-09-27,268,6,270,2014,174.26666666666668\n2014-09-28,335,7,271,2014,180.0\n2014-09-29,244,1,272,2014,183.93333333333334\n2014-09-30,256,2,273,2014,188.23333333333332\n2014-10-01,321,3,274,2014,192.9\n2014-10-02,241,4,275,2014,192.93333333333334\n2014-10-03,199,5,276,2014,193.06666666666666\n2014-10-04,118,6,277,2014,190.96666666666667\n2014-10-05,93,7,278,2014,188.16666666666666\n2014-10-06,149,1,279,2014,189.16666666666666\n2014-10-07,173,2,280,2014,187.4\n2014-10-08,169,3,281,2014,188.6\n2014-10-09,337,4,282,2014,193.96666666666667\n2014-10-10,155,5,283,2014,192.23333333333332\n2014-10-11,89,6,284,2014,190.33333333333334\n2014-10-12,128,7,285,2014,189.76666666666668\n2014-10-13,135,1,286,2014,191.3\n2014-10-14,190,2,287,2014,193.46666666666667\n2014-10-15,181,3,288,2014,192.66666666666666\n2014-10-16,187,4,289,2014,192.6\n2014-10-17,181,5,290,2014,193.13333333333333\n2014-10-18,139,6,291,2014,190.03333333333333\n2014-10-19,157,7,292,2014,188.86666666666667\n2014-10-20,155,1,293,2014,189.9\n2014-10-21,165,2,294,2014,190.63333333333333\n2014-10-22,124,3,295,2014,189.06666666666666\n2014-10-23,211,4,296,2014,189.7\n2014-10-24,158,5,297,2014,188.06666666666666\n2014-10-25,86,6,298,2014,184.83333333333334\n2014-10-26,65,7,299,2014,180.3\n2014-10-27,132,1,300,2014,175.76666666666668\n2014-10-28,135,2,301,2014,169.1\n2014-10-29,227,3,302,2014,168.53333333333333\n2014-10-30,187,4,303,2014,166.23333333333332\n2014-10-31,156,5,304,2014,160.73333333333332\n2014-11-01,86,6,305,2014,155.56666666666666\n2014-11-02,81,7,306,2014,151.63333333333333\n2014-11-03,152,1,307,2014,152.76666666666668\n2014-11-04,167,2,308,2014,155.23333333333332\n2014-11-05,216,3,309,2014,157.46666666666667\n2014-11-06,229,4,310,2014,159.33333333333334\n2014-11-07,208,5,311,2014,160.63333333333333\n2014-11-08,110,6,312,2014,153.06666666666666\n2014-11-09,205,7,313,2014,154.73333333333332\n2014-11-10,161,1,314,2014,157.13333333333333\n2014-11-11,165,2,315,2014,158.36666666666667\n2014-11-12,167,3,316,2014,159.43333333333334\n2014-11-13,162,4,317,2014,158.5\n2014-11-14,174,5,318,2014,158.26666666666668\n2014-11-15,153,6,319,2014,157.13333333333333\n2014-11-16,204,7,320,2014,157.9\n2014-11-17,219,1,321,2014,160.56666666666666\n2014-11-18,203,2,322,2014,162.1\n2014-11-19,180,3,323,2014,162.93333333333334\n2014-11-20,129,4,324,2014,161.73333333333332\n2014-11-21,168,5,325,2014,163.2\n2014-11-22,164,6,326,2014,161.63333333333333\n2014-11-23,141,7,327,2014,161.06666666666666\n2014-11-24,109,1,328,2014,161.83333333333334\n2014-11-25,176,2,329,2014,165.53333333333333\n2014-11-26,154,3,330,2014,166.26666666666668\n2014-11-27,90,4,331,2014,164.76666666666668\n2014-11-28,93,5,332,2014,160.3\n2014-11-29,116,6,333,2014,157.93333333333334\n2014-11-30,174,7,334,2014,158.53333333333333\n2014-12-01,184,1,335,2014,161.8\n2014-12-02,134,2,336,2014,163.56666666666666\n2014-12-03,129,3,337,2014,162.8\n2014-12-04,126,4,338,2014,161.43333333333334\n2014-12-05,132,5,339,2014,158.63333333333333\n2014-12-06,115,6,340,2014,154.83333333333334\n2014-12-07,160,7,341,2014,153.23333333333332\n2014-12-08,218,1,342,2014,156.83333333333334\n2014-12-09,158,2,343,2014,155.26666666666668\n2014-12-10,127,3,344,2014,154.13333333333333\n2014-12-11,140,4,345,2014,153.3\n2014-12-12,135,5,346,2014,152.23333333333332\n2014-12-13,136,6,347,2014,151.36666666666667\n2014-12-14,160,7,348,2014,150.9\n2014-12-15,196,1,349,2014,152.33333333333334\n2014-12-16,191,2,350,2014,151.9\n2014-12-17,138,3,351,2014,149.2\n2014-12-18,104,4,352,2014,145.9\n2014-12-19,205,5,353,2014,146.73333333333332\n2014-12-20,147,6,354,2014,147.33333333333334\n2014-12-21,196,7,355,2014,148.26666666666668\n2014-12-22,194,1,356,2014,149.26666666666668\n2014-12-23,142,2,357,2014,149.3\n2014-12-24,124,3,358,2014,149.8\n2014-12-25,70,4,359,2014,146.26666666666668\n2014-12-26,112,5,360,2014,144.86666666666667\n2014-12-27,193,6,361,2014,148.3\n2014-12-28,118,7,362,2014,149.13333333333333\n2014-12-29,103,1,363,2014,148.7\n2014-12-30,214,2,364,2014,150.03333333333333\n2014-12-31,134,3,365,2014,148.36666666666667\n2015-01-01,103,4,1,2015,147.33333333333334\n2015-01-02,223,5,2,2015,150.46666666666667\n2015-01-03,188,6,3,2015,152.53333333333333\n2015-01-04,143,7,4,2015,152.9\n2015-01-05,165,1,5,2015,154.56666666666666\n2015-01-06,185,2,6,2015,155.4\n2015-01-07,145,3,7,2015,152.96666666666667\n2015-01-08,252,4,8,2015,156.1\n2015-01-09,185,5,9,2015,158.03333333333333\n2015-01-10,160,6,10,2015,158.7\n2015-01-11,146,7,11,2015,159.06666666666666\n2015-01-12,186,1,12,2015,160.73333333333332\n2015-01-13,211,2,13,2015,162.43333333333334\n2015-01-14,131,3,14,2015,160.26666666666668\n2015-01-15,187,4,15,2015,160.13333333333333\n2015-01-16,174,5,16,2015,161.33333333333334\n2015-01-17,164,6,17,2015,163.33333333333334\n2015-01-18,148,7,18,2015,161.43333333333334\n2015-01-19,233,1,19,2015,164.3\n2015-01-20,139,2,20,2015,162.4\n2015-01-21,160,3,21,2015,161.26666666666668\n2015-01-22,156,4,22,2015,161.73333333333332\n2015-01-23,161,5,23,2015,162.96666666666667\n2015-01-24,134,6,24,2015,165.1\n2015-01-25,120,7,25,2015,165.36666666666667\n2015-01-26,174,1,26,2015,164.73333333333332\n2015-01-27,166,2,27,2015,166.33333333333334\n2015-01-28,217,3,28,2015,170.13333333333333\n2015-01-29,146,4,29,2015,167.86666666666667\n2015-01-30,205,5,30,2015,170.23333333333332\n2015-01-31,185,6,31,2015,172.96666666666667\n2015-02-01,126,7,32,2015,169.73333333333332\n2015-02-02,182,1,33,2015,169.53333333333333\n2015-02-03,148,2,34,2015,169.7\n2015-02-04,182,3,35,2015,170.26666666666668\n2015-02-05,256,4,36,2015,172.63333333333333\n2015-02-06,218,5,37,2015,175.06666666666666\n2015-02-07,156,6,38,2015,171.86666666666667\n2015-02-08,163,7,39,2015,171.13333333333333\n2015-02-09,195,1,40,2015,172.3\n2015-02-10,232,2,41,2015,175.16666666666666\n2015-02-11,189,3,42,2015,175.26666666666668\n2015-02-12,216,4,43,2015,175.43333333333334\n2015-02-13,152,5,44,2015,176.13333333333333\n2015-02-14,120,6,45,2015,173.9\n2015-02-15,203,7,46,2015,174.86666666666667\n2015-02-16,177,1,47,2015,175.3\n2015-02-17,215,2,48,2015,177.53333333333333\n2015-02-18,241,3,49,2015,177.8\n2015-02-19,190,4,50,2015,179.5\n2015-02-20,266,5,51,2015,183.03333333333333\n2015-02-21,285,6,52,2015,187.33333333333334\n2015-02-22,205,7,53,2015,188.8\n2015-02-23,278,1,54,2015,193.6\n2015-02-24,197,2,55,2015,196.16666666666666\n2015-02-25,285,3,56,2015,199.86666666666667\n2015-02-26,175,4,57,2015,200.16666666666666\n2015-02-27,185,5,58,2015,199.1\n2015-02-28,158,6,59,2015,199.5\n2015-03-01,140,7,60,2015,197.33333333333334\n2015-03-02,247,1,61,2015,199.4\n2015-03-03,230,2,62,2015,202.86666666666667\n2015-03-04,200,3,63,2015,203.46666666666667\n2015-03-05,283,4,64,2015,207.96666666666667\n2015-03-06,271,5,65,2015,210.93333333333334\n2015-03-07,187,6,66,2015,208.63333333333333\n2015-03-08,117,7,67,2015,205.26666666666668\n2015-03-09,267,1,68,2015,208.96666666666667\n2015-03-10,204,2,69,2015,210.33333333333334\n2015-03-11,276,3,70,2015,213.03333333333333\n2015-03-12,207,4,71,2015,212.2\n2015-03-13,289,5,72,2015,215.53333333333333\n2015-03-14,155,6,73,2015,213.5\n2015-03-15,107,7,74,2015,212.0\n2015-03-16,314,1,75,2015,218.46666666666667\n2015-03-17,266,2,76,2015,220.56666666666666\n2015-03-18,183,3,77,2015,220.76666666666668\n2015-03-19,228,4,78,2015,221.2\n2015-03-20,217,5,79,2015,220.4\n2015-03-21,101,6,80,2015,217.43333333333334\n2015-03-22,187,7,81,2015,214.8\n2015-03-23,230,1,82,2015,212.96666666666667\n2015-03-24,254,2,83,2015,214.6\n2015-03-25,222,3,84,2015,212.73333333333332\n2015-03-26,228,4,85,2015,213.76666666666668\n2015-03-27,242,5,86,2015,212.33333333333334\n2015-03-28,157,6,87,2015,211.73333333333332\n2015-03-29,108,7,88,2015,209.16666666666666\n2015-03-30,252,1,89,2015,212.3\n2015-03-31,201,2,90,2015,214.33333333333334\n2015-04-01,237,3,91,2015,214.0\n2015-04-02,175,4,92,2015,212.16666666666666\n2015-04-03,222,5,93,2015,212.9\n2015-04-04,107,6,94,2015,207.03333333333333\n2015-04-05,91,7,95,2015,201.03333333333333\n2015-04-06,198,1,96,2015,201.4\n2015-04-07,181,2,97,2015,203.53333333333333\n2015-04-08,215,3,98,2015,201.8\n2015-04-09,230,4,99,2015,202.66666666666666\n2015-04-10,141,5,100,2015,198.16666666666666\n2015-04-11,94,6,101,2015,194.4\n2015-04-12,72,7,102,2015,187.16666666666666\n2015-04-13,180,1,103,2015,188.0\n2015-04-14,258,2,104,2015,193.03333333333333\n2015-04-15,178,3,105,2015,188.5\n2015-04-16,210,4,106,2015,186.63333333333333\n2015-04-17,246,5,107,2015,188.73333333333332\n2015-04-18,111,6,108,2015,184.83333333333334\n2015-04-19,139,7,109,2015,182.23333333333332\n2015-04-20,134,1,110,2015,183.33333333333334\n2015-04-21,160,2,111,2015,182.43333333333334\n2015-04-22,186,3,112,2015,180.96666666666667\n2015-04-23,193,4,113,2015,178.93333333333334\n2015-04-24,201,5,114,2015,178.23333333333332\n2015-04-25,140,6,115,2015,175.3\n2015-04-26,164,7,116,2015,172.7\n2015-04-27,185,1,117,2015,173.63333333333333\n2015-04-28,176,2,118,2015,175.9\n2015-04-29,262,3,119,2015,176.23333333333332\n2015-04-30,216,4,120,2015,176.73333333333332\n2015-05-01,188,5,121,2015,175.1\n2015-05-02,142,6,122,2015,174.0\n2015-05-03,172,7,123,2015,172.33333333333334\n2015-05-04,173,1,124,2015,174.53333333333333\n2015-05-05,174,2,125,2015,177.3\n2015-05-06,202,3,126,2015,177.43333333333334\n2015-05-07,157,4,127,2015,176.63333333333333\n2015-05-08,222,5,128,2015,176.86666666666667\n2015-05-09,97,6,129,2015,172.43333333333334\n2015-05-10,97,7,130,2015,170.96666666666667\n2015-05-11,193,1,131,2015,174.26666666666668\n2015-05-12,176,2,132,2015,177.73333333333332\n2015-05-13,193,3,133,2015,178.16666666666666\n2015-05-14,235,4,134,2015,177.4\n2015-05-15,190,5,135,2015,177.8\n2015-05-16,116,6,136,2015,174.66666666666666\n2015-05-17,159,7,137,2015,171.76666666666668\n2015-05-18,312,1,138,2015,178.46666666666667\n2015-05-19,185,2,139,2015,180.0\n2015-05-20,242,3,140,2015,183.6\n2015-05-21,164,4,141,2015,183.73333333333332\n2015-05-22,175,5,142,2015,183.36666666666667\n2015-05-23,75,6,143,2015,179.43333333333334\n2015-05-24,115,7,144,2015,176.56666666666666\n2015-05-25,141,1,145,2015,176.6\n2015-05-26,248,2,146,2015,179.4\n2015-05-27,220,3,147,2015,180.56666666666666\n2015-05-28,196,4,148,2015,181.23333333333332\n2015-05-29,196,5,149,2015,179.03333333333333\n2015-05-30,152,6,150,2015,176.9\n2015-05-31,109,7,151,2015,174.26666666666668\n2015-06-01,193,1,152,2015,175.96666666666667\n2015-06-02,184,2,153,2015,176.36666666666667\n2015-06-03,155,3,154,2015,175.76666666666668\n2015-06-04,134,4,155,2015,174.43333333333334\n2015-06-05,128,5,156,2015,171.96666666666667\n2015-06-06,98,6,157,2015,170.0\n2015-06-07,105,7,158,2015,166.1\n2015-06-08,161,1,159,2015,168.23333333333332\n2015-06-09,176,2,160,2015,170.86666666666667\n2015-06-10,267,3,161,2015,173.33333333333334\n2015-06-11,184,4,162,2015,173.6\n2015-06-12,195,5,163,2015,173.66666666666666\n2015-06-13,108,6,164,2015,169.43333333333334\n2015-06-14,182,7,165,2015,169.16666666666666\n2015-06-15,195,1,166,2015,171.8\n2015-06-16,240,2,167,2015,174.5\n2015-06-17,279,3,168,2015,173.4\n2015-06-18,209,4,169,2015,174.2\n2015-06-19,251,5,170,2015,174.5\n2015-06-20,157,6,171,2015,174.26666666666668\n2015-06-21,120,7,172,2015,172.43333333333334\n2015-06-22,282,1,173,2015,179.33333333333334\n2015-06-23,223,2,174,2015,182.93333333333334\n2015-06-24,233,3,175,2015,186.0\n2015-06-25,199,4,176,2015,184.36666666666667\n2015-06-26,156,5,177,2015,182.23333333333332\n2015-06-27,79,6,178,2015,178.33333333333334\n2015-06-28,159,7,179,2015,177.1\n2015-06-29,192,1,180,2015,178.43333333333334\n2015-06-30,234,2,181,2015,182.6\n2015-07-01,252,3,182,2015,184.56666666666666\n2015-07-02,191,4,183,2015,184.8\n2015-07-03,151,5,184,2015,184.66666666666666\n2015-07-04,107,6,185,2015,183.76666666666668\n2015-07-05,95,7,186,2015,182.66666666666666\n2015-07-06,155,1,187,2015,184.56666666666666\n2015-07-07,224,2,188,2015,188.53333333333333\n2015-07-08,201,3,189,2015,189.86666666666667\n2015-07-09,189,4,190,2015,190.3\n2015-07-10,184,5,191,2015,187.53333333333333\n2015-07-11,142,6,192,2015,186.13333333333333\n2015-07-12,130,7,193,2015,183.96666666666667\n2015-07-13,149,1,194,2015,185.33333333333334\n2015-07-14,235,2,195,2015,187.1\n2015-07-15,166,3,196,2015,186.13333333333333\n2015-07-16,258,4,197,2015,186.73333333333332\n2015-07-17,236,5,198,2015,185.3\n2015-07-18,182,6,199,2015,184.4\n2015-07-19,129,7,200,2015,180.33333333333334\n2015-07-20,237,1,201,2015,183.0\n2015-07-21,275,2,202,2015,188.16666666666666\n2015-07-22,151,3,203,2015,183.8\n2015-07-23,211,4,204,2015,183.4\n2015-07-24,181,5,205,2015,181.66666666666666\n2015-07-25,125,6,206,2015,179.2\n2015-07-26,125,7,207,2015,178.16666666666666\n2015-07-27,206,1,208,2015,182.4\n2015-07-28,211,2,209,2015,184.13333333333333\n2015-07-29,176,3,210,2015,183.6\n2015-07-30,194,4,211,2015,182.26666666666668\n2015-07-31,210,5,212,2015,180.86666666666667\n2015-08-01,134,6,213,2015,178.96666666666667\n2015-08-02,153,7,214,2015,179.03333333333333\n2015-08-03,224,1,215,2015,182.93333333333334\n2015-08-04,221,2,216,2015,187.13333333333333\n2015-08-05,221,3,217,2015,189.33333333333334\n2015-08-06,217,4,218,2015,189.1\n2015-08-07,163,5,219,2015,187.83333333333334\n2015-08-08,99,6,220,2015,184.83333333333334\n2015-08-09,95,7,221,2015,181.86666666666667\n2015-08-10,197,1,222,2015,183.7\n2015-08-11,186,2,223,2015,185.56666666666666\n2015-08-12,225,3,224,2015,188.1\n2015-08-13,203,4,225,2015,187.03333333333333\n2015-08-14,170,5,226,2015,187.16666666666666\n2015-08-15,93,6,227,2015,181.66666666666666\n2015-08-16,108,7,228,2015,177.4\n2015-08-17,112,1,229,2015,175.06666666666666\n2015-08-18,170,2,230,2015,176.43333333333334\n2015-08-19,198,3,231,2015,175.13333333333333\n2015-08-20,200,4,232,2015,172.63333333333333\n2015-08-21,203,5,233,2015,174.36666666666667\n2015-08-22,105,6,234,2015,170.83333333333334\n2015-08-23,119,7,235,2015,168.76666666666668\n2015-08-24,186,1,236,2015,170.8\n2015-08-25,175,2,237,2015,172.46666666666667\n2015-08-26,205,3,238,2015,172.43333333333334\n2015-08-27,197,4,239,2015,171.96666666666667\n2015-08-28,174,5,240,2015,171.9\n2015-08-29,52,6,241,2015,167.16666666666666\n2015-08-30,127,7,242,2015,164.4\n2015-08-31,199,1,243,2015,166.56666666666666\n2015-09-01,219,2,244,2015,168.76666666666668\n2015-09-02,168,3,245,2015,166.9\n2015-09-03,251,4,246,2015,167.9\n2015-09-04,153,5,247,2015,165.63333333333333\n2015-09-05,102,6,248,2015,161.8\n2015-09-06,82,7,249,2015,159.1\n2015-09-07,114,1,250,2015,159.6\n2015-09-08,196,2,251,2015,162.96666666666667\n2015-09-09,272,3,252,2015,165.46666666666667\n2015-09-10,228,4,253,2015,166.86666666666667\n2015-09-11,241,5,254,2015,167.4\n2015-09-12,136,6,255,2015,165.16666666666666\n2015-09-13,108,7,256,2015,163.1\n2015-09-14,216,1,257,2015,167.2\n2015-09-15,197,2,258,2015,170.16666666666666\n2015-09-16,199,3,259,2015,173.06666666666666\n2015-09-17,209,4,260,2015,174.36666666666667\n2015-09-18,197,5,261,2015,174.33333333333334\n2015-09-19,148,6,262,2015,172.6\n2015-09-20,168,7,263,2015,171.43333333333334\n2015-09-21,194,1,264,2015,174.4\n2015-09-22,193,2,265,2015,176.86666666666667\n2015-09-23,152,3,266,2015,175.73333333333332\n2015-09-24,166,4,267,2015,175.43333333333334\n2015-09-25,215,5,268,2015,175.76666666666668\n2015-09-26,161,6,269,2015,174.56666666666666\n2015-09-27,139,7,270,2015,173.4\n2015-09-28,253,1,271,2015,180.1\n2015-09-29,211,2,272,2015,182.9\n2015-09-30,232,3,273,2015,184.0\n2015-10-01,206,4,274,2015,183.56666666666666\n2015-10-02,213,5,275,2015,185.06666666666666\n2015-10-03,129,6,276,2015,181.0\n2015-10-04,114,7,277,2015,179.7\n2015-10-05,227,1,278,2015,183.86666666666667\n2015-10-06,240,2,279,2015,189.13333333333333\n2015-10-07,202,3,280,2015,192.06666666666666\n2015-10-08,219,4,281,2015,192.83333333333334\n2015-10-09,190,5,282,2015,190.1\n2015-10-10,146,6,283,2015,187.36666666666667\n2015-10-11,145,7,284,2015,184.16666666666666\n2015-10-12,125,1,285,2015,183.8\n2015-10-13,218,2,286,2015,187.46666666666667\n2015-10-14,247,3,287,2015,188.5\n2015-10-15,201,4,288,2015,188.63333333333333\n2015-10-16,184,5,289,2015,188.13333333333333\n2015-10-17,112,6,290,2015,184.9\n2015-10-18,128,7,291,2015,182.6\n2015-10-19,276,1,292,2015,186.86666666666667\n2015-10-20,190,2,293,2015,187.6\n2015-10-21,251,3,294,2015,189.5\n2015-10-22,145,4,295,2015,187.9\n2015-10-23,159,5,296,2015,188.13333333333333\n2015-10-24,104,6,297,2015,186.06666666666666\n2015-10-25,153,7,298,2015,184.0\n2015-10-26,212,1,299,2015,185.7\n2015-10-27,179,2,300,2015,187.03333333333333\n2015-10-28,166,3,301,2015,184.13333333333333\n2015-10-29,180,4,302,2015,183.1\n2015-10-30,188,5,303,2015,181.63333333333333\n2015-10-31,65,6,304,2015,176.93333333333334\n2015-11-01,120,7,305,2015,173.83333333333334\n2015-11-02,225,1,306,2015,177.03333333333333\n2015-11-03,196,2,307,2015,179.76666666666668\n2015-11-04,229,3,308,2015,179.83333333333334\n2015-11-05,211,4,309,2015,178.86666666666667\n2015-11-06,181,5,310,2015,178.16666666666666\n2015-11-07,111,6,311,2015,174.56666666666666\n2015-11-08,167,7,312,2015,173.8\n2015-11-09,213,1,313,2015,176.03333333333333\n2015-11-10,256,2,314,2015,179.73333333333332\n2015-11-11,297,3,315,2015,185.46666666666667\n2015-11-12,223,4,316,2015,185.63333333333333\n2015-11-13,290,5,317,2015,187.06666666666666\n2015-11-14,139,6,318,2015,185.0\n2015-11-15,131,7,319,2015,183.23333333333332\n2015-11-16,221,1,320,2015,186.86666666666667\n2015-11-17,189,2,321,2015,188.9\n2015-11-18,208,3,322,2015,186.63333333333333\n2015-11-19,235,4,323,2015,188.13333333333333\n2015-11-20,245,5,324,2015,187.93333333333334\n2015-11-21,160,6,325,2015,188.43333333333334\n2015-11-22,150,7,326,2015,188.13333333333333\n2015-11-23,212,1,327,2015,191.73333333333332\n2015-11-24,316,2,328,2015,197.16666666666666\n2015-11-25,256,3,329,2015,198.63333333333333\n2015-11-26,96,4,330,2015,195.86666666666667\n2015-11-27,114,5,331,2015,194.13333333333333\n2015-11-28,152,6,332,2015,193.2\n2015-11-29,160,7,333,2015,192.26666666666668\n2015-11-30,165,1,334,2015,195.6\n2015-12-01,230,2,335,2015,199.26666666666668\n2015-12-02,192,3,336,2015,198.16666666666666\n2015-12-03,171,4,337,2015,197.33333333333334\n2015-12-04,149,5,338,2015,194.66666666666666\n2015-12-05,138,6,339,2015,192.23333333333332\n2015-12-06,268,7,340,2015,195.13333333333333\n2015-12-07,200,1,341,2015,198.1\n2015-12-08,248,2,342,2015,200.8\n2015-12-09,213,3,343,2015,200.8\n2015-12-10,241,4,344,2015,200.3\n2015-12-11,232,5,345,2015,198.13333333333333\n2015-12-12,95,6,346,2015,193.86666666666667\n2015-12-13,120,7,347,2015,188.2\n2015-12-14,220,1,348,2015,190.9\n2015-12-15,258,2,349,2015,195.13333333333333\n2015-12-16,230,3,350,2015,195.43333333333334\n2015-12-17,262,4,351,2015,197.86666666666667\n2015-12-18,226,5,352,2015,198.46666666666667\n2015-12-19,163,6,353,2015,196.06666666666666\n2015-12-20,131,7,354,2015,192.26666666666668\n2015-12-21,220,1,355,2015,194.26666666666668\n2015-12-22,153,2,356,2015,194.36666666666667\n2015-12-23,192,3,357,2015,193.7\n2015-12-24,96,4,358,2015,186.36666666666667\n2015-12-25,53,5,359,2015,179.6\n2015-12-26,85,6,360,2015,179.23333333333332\n2015-12-27,96,7,361,2015,178.63333333333333\n2015-12-28,139,1,362,2015,178.2\n2015-12-29,166,2,363,2015,178.4\n2015-12-30,156,3,364,2015,178.1\n2015-12-31,128,4,365,2015,174.7\n2016-01-01,131,5,1,2016,172.66666666666666\n2016-01-02,151,6,2,2016,172.0\n2016-01-03,154,7,3,2016,172.16666666666666\n2016-01-04,185,1,4,2016,173.73333333333332\n2016-01-05,226,2,5,2016,172.33333333333334\n2016-01-06,276,3,6,2016,174.86666666666667\n2016-01-07,310,4,7,2016,176.93333333333334\n2016-01-08,242,5,8,2016,177.9\n2016-01-09,197,6,9,2016,176.43333333333334\n2016-01-10,159,7,10,2016,174.0\n2016-01-11,330,1,11,2016,181.83333333333334\n2016-01-12,212,2,12,2016,184.9\n2016-01-13,261,3,13,2016,186.26666666666668\n2016-01-14,294,4,14,2016,187.46666666666667\n2016-01-15,263,5,15,2016,188.56666666666666\n2016-01-16,228,6,16,2016,187.43333333333334\n2016-01-17,200,7,17,2016,186.56666666666666\n2016-01-18,203,1,18,2016,187.9\n2016-01-19,214,2,19,2016,190.66666666666666\n2016-01-20,195,3,20,2016,189.83333333333334\n2016-01-21,268,4,21,2016,193.66666666666666\n2016-01-22,209,5,22,2016,194.23333333333332\n2016-01-23,169,6,23,2016,196.66666666666666\n2016-01-24,214,7,24,2016,202.03333333333333\n2016-01-25,251,1,25,2016,207.56666666666666\n2016-01-26,224,2,26,2016,211.83333333333334\n2016-01-27,211,3,27,2016,214.23333333333332\n2016-01-28,238,4,28,2016,216.63333333333333\n2016-01-29,316,5,29,2016,221.96666666666667\n2016-01-30,187,6,30,2016,223.93333333333334\n2016-01-31,253,7,31,2016,228.0\n2016-02-01,271,1,32,2016,232.0\n2016-02-02,307,2,33,2016,237.1\n2016-02-03,221,3,34,2016,238.3\n2016-02-04,209,4,35,2016,237.73333333333332\n2016-02-05,201,5,36,2016,235.23333333333332\n2016-02-06,165,6,37,2016,230.4\n2016-02-07,163,7,38,2016,227.76666666666668\n2016-02-08,191,1,39,2016,227.56666666666666\n2016-02-09,217,2,40,2016,229.5\n2016-02-10,192,3,41,2016,224.9\n2016-02-11,218,4,42,2016,225.1\n2016-02-12,291,5,43,2016,226.1\n2016-02-13,137,6,44,2016,220.86666666666667\n2016-02-14,145,7,45,2016,216.93333333333334\n2016-02-15,182,1,46,2016,215.4\n2016-02-16,249,2,47,2016,217.03333333333333\n2016-02-17,238,3,48,2016,218.2\n2016-02-18,224,4,49,2016,218.53333333333333\n2016-02-19,258,5,50,2016,220.63333333333333\n2016-02-20,136,6,51,2016,216.23333333333332\n2016-02-21,168,7,52,2016,214.86666666666667\n2016-02-22,211,1,53,2016,216.26666666666668\n2016-02-23,256,2,54,2016,217.66666666666666\n2016-02-24,232,3,55,2016,217.03333333333333\n2016-02-25,242,4,56,2016,217.63333333333333\n2016-02-26,232,5,57,2016,218.33333333333334\n2016-02-27,179,6,58,2016,216.36666666666667\n2016-02-28,246,7,59,2016,214.03333333333333\n2016-02-29,257,1,60,2016,216.36666666666667\n2016-03-01,208,2,61,2016,214.86666666666667\n2016-03-02,178,3,62,2016,211.76666666666668\n2016-03-03,208,4,63,2016,208.46666666666667\n2016-03-04,208,5,64,2016,208.03333333333333\n2016-03-05,127,6,65,2016,205.3\n2016-03-06,119,7,66,2016,202.56666666666666\n2016-03-07,208,1,67,2016,204.0\n2016-03-08,236,2,68,2016,206.43333333333334\n2016-03-09,191,3,69,2016,206.43333333333334\n2016-03-10,136,4,70,2016,203.73333333333332\n2016-03-11,180,5,71,2016,203.33333333333334\n2016-03-12,98,6,72,2016,199.33333333333334\n2016-03-13,97,7,73,2016,192.86666666666667\n2016-03-14,142,1,74,2016,193.03333333333333\n2016-03-15,180,2,75,2016,194.2\n2016-03-16,245,3,76,2016,196.3\n2016-03-17,232,4,77,2016,195.73333333333332\n2016-03-18,188,5,78,2016,194.06666666666666\n2016-03-19,119,6,79,2016,190.56666666666666\n2016-03-20,159,7,80,2016,187.26666666666668\n2016-03-21,216,1,81,2016,189.93333333333334\n2016-03-22,200,2,82,2016,191.0\n2016-03-23,228,3,83,2016,191.56666666666666\n2016-03-24,184,4,84,2016,189.16666666666666\n2016-03-25,272,5,85,2016,190.5\n2016-03-26,145,6,86,2016,187.26666666666668\n2016-03-27,101,7,87,2016,182.9\n2016-03-28,241,1,88,2016,184.96666666666667\n2016-03-29,204,2,89,2016,183.56666666666666\n2016-03-30,153,3,90,2016,180.1\n2016-03-31,177,4,91,2016,179.06666666666666\n2016-04-01,245,5,92,2016,181.3\n2016-04-02,175,6,93,2016,180.2\n2016-04-03,175,7,94,2016,179.1\n2016-04-04,206,1,95,2016,181.73333333333332\n2016-04-05,190,2,96,2016,184.1\n2016-04-06,189,3,97,2016,183.46666666666667\n2016-04-07,174,4,98,2016,181.4\n2016-04-08,182,5,99,2016,181.1\n2016-04-09,182,6,100,2016,182.63333333333333\n2016-04-10,124,7,101,2016,180.76666666666668\n2016-04-11,205,1,102,2016,184.33333333333334\n2016-04-12,218,2,103,2016,188.36666666666667\n2016-04-13,233,3,104,2016,191.4\n2016-04-14,231,4,105,2016,193.1\n2016-04-15,155,5,106,2016,190.1\n2016-04-16,174,6,107,2016,188.16666666666666\n2016-04-17,61,7,108,2016,183.93333333333334\n2016-04-18,197,1,109,2016,186.53333333333333\n2016-04-19,197,2,110,2016,187.8\n2016-04-20,270,3,111,2016,189.6\n2016-04-21,174,4,112,2016,188.73333333333332\n2016-04-22,166,5,113,2016,186.66666666666666\n2016-04-23,131,6,114,2016,184.9\n2016-04-24,89,7,115,2016,178.8\n2016-04-25,185,1,116,2016,180.13333333333333\n2016-04-26,155,2,117,2016,181.93333333333334\n2016-04-27,147,3,118,2016,178.8\n2016-04-28,129,4,119,2016,176.3\n2016-04-29,182,5,120,2016,177.26666666666668\n2016-04-30,114,6,121,2016,175.16666666666666\n2016-05-01,114,7,122,2016,170.8\n2016-05-02,268,1,123,2016,173.9\n2016-05-03,157,2,124,2016,173.3\n2016-05-04,184,3,125,2016,172.56666666666666\n2016-05-05,176,4,126,2016,172.1\n2016-05-06,162,5,127,2016,171.2\n2016-05-07,147,6,128,2016,170.3\n2016-05-08,80,7,129,2016,166.9\n2016-05-09,190,1,130,2016,167.16666666666666\n2016-05-10,205,2,131,2016,169.86666666666667\n2016-05-11,190,3,132,2016,169.36666666666667\n2016-05-12,140,4,133,2016,166.76666666666668\n2016-05-13,192,5,134,2016,165.4\n2016-05-14,77,6,135,2016,160.26666666666668\n2016-05-15,98,7,136,2016,158.36666666666667\n2016-05-16,187,1,137,2016,158.8\n2016-05-17,144,2,138,2016,161.56666666666666\n2016-05-18,160,3,139,2016,160.33333333333334\n2016-05-19,197,4,140,2016,160.33333333333334\n2016-05-20,183,5,141,2016,157.43333333333334\n2016-05-21,135,6,142,2016,156.13333333333333\n2016-05-22,150,7,143,2016,155.6\n2016-05-23,211,1,144,2016,158.26666666666668\n2016-05-24,236,2,145,2016,163.16666666666666\n2016-05-25,167,3,146,2016,162.56666666666666\n2016-05-26,160,4,147,2016,162.73333333333332\n2016-05-27,131,5,148,2016,162.2\n2016-05-28,97,6,149,2016,161.13333333333333\n2016-05-29,118,7,150,2016,159.0\n2016-05-30,162,1,151,2016,160.6\n2016-05-31,206,2,152,2016,163.66666666666666\n2016-06-01,280,3,153,2016,164.06666666666666\n2016-06-02,224,4,154,2016,166.3\n2016-06-03,255,5,155,2016,168.66666666666666\n2016-06-04,130,6,156,2016,167.13333333333333\n2016-06-05,119,7,157,2016,165.7\n2016-06-06,176,1,158,2016,166.66666666666666\n2016-06-07,212,2,159,2016,171.06666666666666\n2016-06-08,211,3,160,2016,171.76666666666668\n2016-06-09,209,4,161,2016,171.9\n2016-06-10,195,5,162,2016,172.06666666666666\n2016-06-11,178,6,163,2016,173.33333333333334\n2016-06-12,129,7,164,2016,171.23333333333332\n2016-06-13,292,1,165,2016,178.4\n2016-06-14,211,2,166,2016,182.16666666666666\n2016-06-15,196,3,167,2016,182.46666666666667\n2016-06-16,192,4,168,2016,184.06666666666666\n2016-06-17,199,5,169,2016,185.36666666666667\n2016-06-18,116,6,170,2016,182.66666666666666\n2016-06-19,103,7,171,2016,180.0\n2016-06-20,271,1,172,2016,184.53333333333333\n2016-06-21,233,2,173,2016,187.3\n2016-06-22,220,3,174,2016,187.6\n2016-06-23,175,4,175,2016,185.56666666666666\n2016-06-24,165,5,176,2016,185.5\n2016-06-25,67,6,177,2016,182.4\n2016-06-26,116,7,178,2016,181.9\n2016-06-27,271,1,179,2016,187.7\n2016-06-28,214,2,180,2016,190.9\n2016-06-29,193,3,181,2016,191.93333333333334\n2016-06-30,243,4,182,2016,193.16666666666666\n2016-07-01,218,5,183,2016,191.1\n2016-07-02,170,6,184,2016,189.3\n2016-07-03,102,7,185,2016,184.2\n2016-07-04,123,1,186,2016,183.96666666666667\n2016-07-05,148,2,187,2016,184.93333333333334\n2016-07-06,222,3,188,2016,186.46666666666667\n2016-07-07,204,4,189,2016,186.2\n2016-07-08,172,5,190,2016,184.9\n2016-07-09,115,6,191,2016,181.76666666666668\n2016-07-10,115,7,192,2016,179.1\n2016-07-11,175,1,193,2016,179.0\n2016-07-12,189,2,194,2016,181.0\n2016-07-13,168,3,195,2016,176.86666666666667\n2016-07-14,164,4,196,2016,175.3\n2016-07-15,159,5,197,2016,174.06666666666666\n2016-07-16,91,6,198,2016,170.7\n2016-07-17,135,7,199,2016,168.56666666666666\n2016-07-18,210,1,200,2016,171.7\n2016-07-19,208,2,201,2016,175.2\n2016-07-20,199,3,202,2016,172.8\n2016-07-21,155,4,203,2016,170.2\n2016-07-22,184,5,204,2016,169.0\n2016-07-23,98,6,205,2016,166.43333333333334\n2016-07-24,117,7,206,2016,164.83333333333334\n2016-07-25,200,1,207,2016,169.26666666666668\n2016-07-26,233,2,208,2016,173.16666666666666\n2016-07-27,221,3,209,2016,171.5\n2016-07-28,267,4,210,2016,173.26666666666668\n2016-07-29,219,5,211,2016,174.13333333333333\n2016-07-30,122,6,212,2016,170.1\n2016-07-31,133,7,213,2016,167.26666666666668\n2016-08-01,178,1,214,2016,167.53333333333333\n2016-08-02,170,2,215,2016,169.8\n2016-08-03,131,3,216,2016,170.06666666666666\n2016-08-04,230,4,217,2016,172.8\n2016-08-05,155,5,218,2016,170.56666666666666\n2016-08-06,71,6,219,2016,166.13333333333333\n2016-08-07,105,7,220,2016,163.9\n2016-08-08,158,1,221,2016,165.33333333333334\n2016-08-09,205,2,222,2016,168.33333333333334\n2016-08-10,227,3,223,2016,170.06666666666666\n2016-08-11,206,4,224,2016,170.63333333333333\n2016-08-12,194,5,225,2016,171.5\n2016-08-13,66,6,226,2016,168.23333333333332\n2016-08-14,97,7,227,2016,166.16666666666666\n2016-08-15,165,1,228,2016,168.63333333333333\n2016-08-16,170,2,229,2016,169.8\n2016-08-17,191,3,230,2016,169.16666666666666\n2016-08-18,180,4,231,2016,168.23333333333332\n2016-08-19,181,5,232,2016,167.63333333333333\n2016-08-20,126,6,233,2016,166.66666666666666\n2016-08-21,143,7,234,2016,165.3\n2016-08-22,207,1,235,2016,168.93333333333334\n2016-08-23,227,2,236,2016,172.6\n2016-08-24,206,3,237,2016,172.8\n2016-08-25,242,4,238,2016,173.1\n2016-08-26,140,5,239,2016,170.4\n2016-08-27,104,6,240,2016,164.96666666666667\n2016-08-28,111,7,241,2016,161.36666666666667\n2016-08-29,192,1,242,2016,163.7\n2016-08-30,182,2,243,2016,165.33333333333334\n2016-08-31,254,3,244,2016,167.86666666666667\n2016-09-01,265,4,245,2016,171.03333333333333\n2016-09-02,143,5,246,2016,171.43333333333334\n2016-09-03,74,6,247,2016,166.23333333333332\n2016-09-04,122,7,248,2016,165.13333333333333\n2016-09-05,130,1,249,2016,167.1\n2016-09-06,213,2,250,2016,170.7\n2016-09-07,177,3,251,2016,171.33333333333334\n2016-09-08,169,4,252,2016,170.13333333333333\n2016-09-09,174,5,253,2016,168.36666666666667\n2016-09-10,65,6,254,2016,163.66666666666666\n2016-09-11,89,7,255,2016,160.16666666666666\n2016-09-12,153,1,256,2016,163.06666666666666\n2016-09-13,119,2,257,2016,163.8\n2016-09-14,141,3,258,2016,163.0\n2016-09-15,179,4,259,2016,163.3\n2016-09-16,190,5,260,2016,163.26666666666668\n2016-09-17,85,6,261,2016,160.1\n2016-09-18,150,7,262,2016,159.06666666666666\n2016-09-19,200,1,263,2016,161.53333333333333\n2016-09-20,169,2,264,2016,162.4\n2016-09-21,149,3,265,2016,160.46666666666667\n2016-09-22,170,4,266,2016,158.56666666666666\n2016-09-23,131,5,267,2016,156.06666666666666\n2016-09-24,76,6,268,2016,150.53333333333333\n2016-09-25,78,7,269,2016,148.46666666666667\n2016-09-26,163,1,270,2016,150.43333333333334\n2016-09-27,187,2,271,2016,152.96666666666667\n2016-09-28,259,3,272,2016,155.2\n2016-09-29,154,4,273,2016,154.26666666666668\n2016-09-30,150,5,274,2016,150.8\n2016-10-01,98,6,275,2016,145.23333333333332\n2016-10-02,103,7,276,2016,143.9\n2016-10-03,199,1,277,2016,148.06666666666666\n2016-10-04,213,2,278,2016,151.1\n2016-10-05,263,3,279,2016,155.53333333333333\n2016-10-06,212,4,280,2016,155.5\n2016-10-07,226,5,281,2016,157.13333333333333\n2016-10-08,135,6,282,2016,156.0\n2016-10-09,93,7,283,2016,153.3\n2016-10-10,181,1,284,2016,157.16666666666666\n2016-10-11,166,2,285,2016,159.73333333333332\n2016-10-12,224,3,286,2016,162.1\n2016-10-13,221,4,287,2016,165.5\n2016-10-14,194,5,288,2016,167.26666666666668\n2016-10-15,97,6,289,2016,164.53333333333333\n2016-10-16,126,7,290,2016,162.4\n2016-10-17,141,1,291,2016,164.26666666666668\n2016-10-18,146,2,292,2016,164.13333333333333\n2016-10-19,114,3,293,2016,161.26666666666668\n2016-10-20,142,4,294,2016,160.36666666666667\n2016-10-21,182,5,295,2016,161.46666666666667\n2016-10-22,95,6,296,2016,158.96666666666667\n2016-10-23,124,7,297,2016,158.73333333333332\n2016-10-24,132,1,298,2016,160.6\n2016-10-25,140,2,299,2016,162.66666666666666\n2016-10-26,148,3,300,2016,162.16666666666666\n2016-10-27,124,4,301,2016,160.06666666666666\n2016-10-28,123,5,302,2016,155.53333333333333\n2016-10-29,156,6,303,2016,155.6\n2016-10-30,81,7,304,2016,153.3\n2016-10-31,82,1,305,2016,152.76666666666668\n2016-11-01,127,2,306,2016,153.56666666666666\n2016-11-02,146,3,307,2016,151.8\n2016-11-03,156,4,308,2016,149.9\n2016-11-04,148,5,309,2016,146.06666666666666\n2016-11-05,78,6,310,2016,141.6\n2016-11-06,103,7,311,2016,137.5\n2016-11-07,156,1,312,2016,138.2\n2016-11-08,120,2,313,2016,139.1\n2016-11-09,147,3,314,2016,137.96666666666667\n2016-11-10,148,4,315,2016,137.36666666666667\n2016-11-11,114,5,316,2016,133.7\n2016-11-12,76,6,317,2016,128.86666666666667\n2016-11-13,158,7,318,2016,127.66666666666667\n2016-11-14,179,1,319,2016,130.4\n2016-11-15,214,2,320,2016,133.33333333333334\n2016-11-16,142,3,321,2016,133.36666666666667\n2016-11-17,184,4,322,2016,134.63333333333333\n2016-11-18,162,5,323,2016,136.23333333333332\n2016-11-19,97,6,324,2016,134.73333333333332\n2016-11-20,103,7,325,2016,132.1\n2016-11-21,162,1,326,2016,134.33333333333334\n2016-11-22,112,2,327,2016,133.93333333333334\n2016-11-23,86,3,328,2016,132.4\n2016-11-24,75,4,329,2016,130.23333333333332\n2016-11-25,56,5,330,2016,127.16666666666667\n2016-11-26,47,6,331,2016,124.6\n2016-11-27,78,7,332,2016,123.1\n2016-11-28,141,1,333,2016,122.6\n2016-11-29,169,2,334,2016,125.53333333333333\n2016-11-30,138,3,335,2016,127.4\n2016-12-01,137,4,336,2016,127.73333333333333\n2016-12-02,147,5,337,2016,127.76666666666667\n2016-12-03,123,6,338,2016,126.66666666666667\n2016-12-04,107,7,339,2016,125.3\n2016-12-05,214,1,340,2016,129.83333333333334\n2016-12-06,213,2,341,2016,133.5\n2016-12-07,190,3,342,2016,134.63333333333333\n2016-12-08,168,4,343,2016,136.23333333333332\n2016-12-09,156,5,344,2016,136.53333333333333\n2016-12-10,87,6,345,2016,134.5\n2016-12-11,106,7,346,2016,134.23333333333332\n2016-12-12,210,1,347,2016,138.7\n2016-12-13,227,2,348,2016,141.0\n2016-12-14,192,3,349,2016,141.43333333333334\n2016-12-15,163,4,350,2016,139.73333333333332\n2016-12-16,89,5,351,2016,137.96666666666667\n2016-12-17,95,6,352,2016,135.0\n2016-12-18,90,7,353,2016,132.6\n2016-12-19,144,1,354,2016,134.16666666666666\n2016-12-20,138,2,355,2016,135.33333333333334\n2016-12-21,152,3,356,2016,135.0\n2016-12-22,159,4,357,2016,136.56666666666666\n2016-12-23,122,5,358,2016,137.76666666666668\n2016-12-24,67,6,359,2016,137.5\n2016-12-25,57,7,360,2016,137.53333333333333\n2016-12-26,75,1,361,2016,138.46666666666667\n2016-12-27,111,2,362,2016,139.56666666666666\n2016-12-28,117,3,363,2016,138.76666666666668\n2016-12-29,129,4,364,2016,137.43333333333334\n2016-12-30,75,5,365,2016,135.33333333333334\n2016-12-31,23,6,366,2016,131.53333333333333\n2017-01-01,59,7,1,2017,128.6\n2017-01-02,113,1,2,2017,128.26666666666668\n2017-01-03,135,2,3,2017,129.2\n2017-01-04,186,3,4,2017,128.26666666666668\n2017-01-05,137,4,5,2017,125.73333333333333\n2017-01-06,141,5,6,2017,124.1\n2017-01-07,80,6,7,2017,121.16666666666667\n2017-01-08,85,7,8,2017,118.8\n2017-01-09,111,1,9,2017,119.6\n2017-01-10,183,2,10,2017,122.16666666666667\n2017-01-11,188,3,11,2017,121.43333333333334\n2017-01-12,157,4,12,2017,119.1\n2017-01-13,159,5,13,2017,118.0\n2017-01-14,58,6,14,2017,114.5\n2017-01-15,124,7,15,2017,115.66666666666667\n2017-01-16,116,1,16,2017,116.36666666666666\n2017-01-17,156,2,17,2017,118.56666666666666\n2017-01-18,146,3,18,2017,118.63333333333334\n2017-01-19,98,4,19,2017,117.3\n2017-01-20,103,5,20,2017,115.66666666666667\n2017-01-21,83,6,21,2017,113.13333333333334\n2017-01-22,133,7,22,2017,113.5\n2017-01-23,166,1,23,2017,116.8\n2017-01-24,133,2,24,2017,119.33333333333333\n2017-01-25,183,3,25,2017,122.93333333333334\n2017-01-26,135,4,26,2017,123.73333333333333\n2017-01-27,153,5,27,2017,124.93333333333334\n2017-01-28,103,6,28,2017,124.06666666666666\n2017-01-29,65,7,29,2017,123.73333333333333\n2017-01-30,154,1,30,2017,128.1\n2017-01-31,137,2,31,2017,130.7\n2017-02-01,125,3,32,2017,131.1\n2017-02-02,143,4,33,2017,131.36666666666667\n2017-02-03,102,5,34,2017,128.56666666666666\n2017-02-04,73,6,35,2017,126.43333333333334\n2017-02-05,100,7,36,2017,125.06666666666666\n2017-02-06,131,1,37,2017,126.76666666666667\n2017-02-07,134,2,38,2017,128.4\n2017-02-08,128,3,39,2017,128.96666666666667\n2017-02-09,122,4,40,2017,126.93333333333334\n2017-02-10,114,5,41,2017,124.46666666666667\n2017-02-11,65,6,42,2017,121.4\n2017-02-12,116,7,43,2017,119.96666666666667\n2017-02-13,110,1,44,2017,121.7\n2017-02-14,151,2,45,2017,122.6\n2017-02-15,109,3,46,2017,122.36666666666666\n2017-02-16,84,4,47,2017,119.96666666666667\n2017-02-17,96,5,48,2017,118.3\n2017-02-18,86,6,49,2017,117.9\n2017-02-19,72,7,50,2017,116.86666666666666\n2017-02-20,109,1,51,2017,117.73333333333333\n2017-02-21,153,2,52,2017,118.4\n2017-02-22,166,3,53,2017,118.4\n2017-02-23,109,4,54,2017,117.6\n2017-02-24,117,5,55,2017,115.4\n2017-02-25,58,6,56,2017,112.83333333333333\n2017-02-26,77,7,57,2017,110.3\n2017-02-27,102,1,58,2017,110.26666666666667\n2017-02-28,125,2,59,2017,112.26666666666667\n2017-03-01,135,3,60,2017,111.63333333333334\n2017-03-02,144,4,61,2017,111.86666666666666\n2017-03-03,139,5,62,2017,112.33333333333333\n2017-03-04,85,6,63,2017,110.4\n2017-03-05,71,7,64,2017,109.36666666666666\n2017-03-06,149,1,65,2017,111.9\n2017-03-07,132,2,66,2017,112.96666666666667\n2017-03-08,149,3,67,2017,113.56666666666666\n2017-03-09,152,4,68,2017,114.16666666666667\n2017-03-10,194,5,69,2017,116.36666666666666\n2017-03-11,118,6,70,2017,116.23333333333333\n2017-03-12,63,7,71,2017,114.53333333333333\n2017-03-13,102,1,72,2017,115.76666666666667\n2017-03-14,204,2,73,2017,118.7\n2017-03-15,166,3,74,2017,120.56666666666666\n2017-03-16,131,4,75,2017,119.9\n2017-03-17,147,5,76,2017,121.16666666666667\n2017-03-18,83,6,77,2017,121.13333333333334\n2017-03-19,96,7,78,2017,121.13333333333334\n2017-03-20,111,1,79,2017,121.96666666666667\n2017-03-21,201,2,80,2017,126.26666666666667\n2017-03-22,140,3,81,2017,127.3\n2017-03-23,163,4,82,2017,127.63333333333334\n2017-03-24,157,5,83,2017,127.33333333333333\n2017-03-25,102,6,84,2017,127.1\n2017-03-26,79,7,85,2017,125.83333333333333\n2017-03-27,146,1,86,2017,128.76666666666668\n2017-03-28,157,2,87,2017,131.43333333333334\n2017-03-29,130,3,88,2017,132.36666666666667\n2017-03-30,121,4,89,2017,132.23333333333332\n2017-03-31,105,5,90,2017,131.23333333333332\n2017-04-01,35,6,91,2017,127.6\n2017-04-02,109,7,92,2017,126.6\n2017-04-03,149,1,93,2017,128.73333333333332\n2017-04-04,129,2,94,2017,130.66666666666666\n2017-04-05,201,3,95,2017,132.4\n2017-04-06,169,4,96,2017,133.63333333333333\n2017-04-07,195,5,97,2017,135.16666666666666\n2017-04-08,63,6,98,2017,132.2\n2017-04-09,66,7,99,2017,127.93333333333334\n2017-04-10,152,1,100,2017,129.06666666666666\n2017-04-11,168,2,101,2017,132.56666666666666\n2017-04-12,235,3,102,2017,137.0\n2017-04-13,189,4,103,2017,136.5\n2017-04-14,148,5,104,2017,135.9\n2017-04-15,93,6,105,2017,134.63333333333333\n2017-04-16,69,7,106,2017,132.03333333333333\n2017-04-17,128,1,107,2017,133.53333333333333\n2017-04-18,196,2,108,2017,136.86666666666667\n2017-04-19,198,3,109,2017,139.76666666666668\n2017-04-20,133,4,110,2017,137.5\n2017-04-21,131,5,111,2017,137.2\n2017-04-22,81,6,112,2017,134.46666666666667\n2017-04-23,66,7,113,2017,131.43333333333334\n2017-04-24,148,1,114,2017,132.96666666666667\n2017-04-25,171,2,115,2017,136.03333333333333\n2017-04-26,176,3,116,2017,137.03333333333333\n2017-04-27,169,4,117,2017,137.43333333333334\n2017-04-28,181,5,118,2017,139.13333333333333\n2017-04-29,84,6,119,2017,137.9\n2017-04-30,81,7,120,2017,137.1\n2017-05-01,161,1,121,2017,141.3\n2017-05-02,182,2,122,2017,143.73333333333332\n2017-05-03,193,3,123,2017,145.2\n2017-05-04,170,4,124,2017,146.56666666666666\n2017-05-05,122,5,125,2017,143.93333333333334\n2017-05-06,80,6,126,2017,140.96666666666667\n2017-05-07,127,7,127,2017,138.7\n2017-05-08,180,1,128,2017,142.6\n2017-05-09,152,2,129,2017,145.46666666666667\n2017-05-10,225,3,130,2017,147.9\n2017-05-11,111,4,131,2017,146.0\n2017-05-12,173,5,132,2017,143.93333333333334\n2017-05-13,76,6,133,2017,140.16666666666666\n2017-05-14,52,7,134,2017,136.96666666666667\n2017-05-15,134,1,135,2017,138.33333333333334\n2017-05-16,140,2,136,2017,140.7\n2017-05-17,161,3,137,2017,141.8\n2017-05-18,175,4,138,2017,141.1\n2017-05-19,199,5,139,2017,141.13333333333333\n2017-05-20,102,6,140,2017,140.1\n2017-05-21,145,7,141,2017,140.56666666666666\n2017-05-22,170,1,142,2017,143.53333333333333\n2017-05-23,158,2,143,2017,146.6\n2017-05-24,141,3,144,2017,146.36666666666667\n2017-05-25,132,4,145,2017,145.06666666666666\n2017-05-26,181,5,146,2017,145.23333333333332\n2017-05-27,83,6,147,2017,142.36666666666667\n2017-05-28,78,7,148,2017,138.93333333333334\n2017-05-29,122,1,149,2017,140.2\n2017-05-30,188,2,150,2017,143.76666666666668\n2017-05-31,152,3,151,2017,143.46666666666667\n2017-06-01,189,4,152,2017,143.7\n2017-06-02,201,5,153,2017,143.96666666666667\n2017-06-03,91,6,154,2017,141.33333333333334\n2017-06-04,113,7,155,2017,141.03333333333333\n2017-06-05,164,1,156,2017,143.83333333333334\n2017-06-06,125,2,157,2017,143.76666666666668\n2017-06-07,175,3,158,2017,143.6\n2017-06-08,184,4,159,2017,144.66666666666666\n2017-06-09,154,5,160,2017,142.3\n2017-06-10,86,6,161,2017,141.46666666666667\n2017-06-11,37,7,162,2017,136.93333333333334\n2017-06-12,179,1,163,2017,140.36666666666667\n2017-06-13,196,2,164,2017,145.16666666666666\n2017-06-14,169,3,165,2017,146.33333333333334\n2017-06-15,178,4,166,2017,147.6\n2017-06-16,149,5,167,2017,147.2\n2017-06-17,89,6,168,2017,144.33333333333334\n2017-06-18,85,7,169,2017,140.53333333333333\n2017-06-19,142,1,170,2017,141.86666666666667\n2017-06-20,157,2,171,2017,142.26666666666668\n2017-06-21,143,3,172,2017,141.36666666666667\n2017-06-22,152,4,173,2017,141.16666666666666\n2017-06-23,185,5,174,2017,142.63333333333333\n2017-06-24,77,6,175,2017,140.8\n2017-06-25,77,7,176,2017,137.33333333333334\n2017-06-26,140,1,177,2017,139.23333333333332\n2017-06-27,239,2,178,2017,144.6\n2017-06-28,181,3,179,2017,146.56666666666666\n2017-06-29,154,4,180,2017,145.43333333333334\n2017-06-30,162,5,181,2017,145.76666666666668\n2017-07-01,99,6,182,2017,142.76666666666668\n2017-07-02,108,7,183,2017,139.66666666666666\n2017-07-03,129,1,184,2017,140.93333333333334\n2017-07-04,162,2,185,2017,142.56666666666666\n2017-07-05,180,3,186,2017,143.1\n2017-07-06,174,4,187,2017,144.73333333333332\n2017-07-07,153,5,188,2017,144.0\n2017-07-08,67,6,189,2017,140.1\n2017-07-09,76,7,190,2017,137.5\n2017-07-10,184,1,191,2017,140.76666666666668\n2017-07-11,200,2,192,2017,146.2\n2017-07-12,204,3,193,2017,147.03333333333333\n2017-07-13,181,4,194,2017,146.53333333333333\n2017-07-14,157,5,195,2017,146.13333333333333\n2017-07-15,115,6,196,2017,144.03333333333333\n2017-07-16,152,7,197,2017,144.13333333333333\n2017-07-17,117,1,198,2017,145.06666666666666\n2017-07-18,168,2,199,2017,147.83333333333334\n2017-07-19,162,3,200,2017,148.5\n2017-07-20,150,4,201,2017,148.26666666666668\n2017-07-21,142,5,202,2017,148.23333333333332\n2017-07-22,86,6,203,2017,146.03333333333333\n2017-07-23,94,7,204,2017,143.0\n2017-07-24,113,1,205,2017,144.2\n2017-07-25,152,2,206,2017,146.7\n2017-07-26,97,3,207,2017,145.26666666666668\n2017-07-27,139,4,208,2017,141.93333333333334\n2017-07-28,108,5,209,2017,139.5\n2017-07-29,55,6,210,2017,136.2\n2017-07-30,60,7,211,2017,132.8\n2017-07-31,132,1,212,2017,133.9\n2017-08-01,175,2,213,2017,136.13333333333333\n2017-08-02,107,3,214,2017,135.4\n2017-08-03,118,4,215,2017,133.93333333333334\n2017-08-04,123,5,216,2017,132.03333333333333\n2017-08-05,88,6,217,2017,129.16666666666666\n2017-08-06,36,7,218,2017,125.26666666666667\n2017-08-07,131,1,219,2017,127.4\n2017-08-08,119,2,220,2017,128.83333333333334\n2017-08-09,116,3,221,2017,126.56666666666666\n2017-08-10,71,4,222,2017,122.26666666666667\n2017-08-11,116,5,223,2017,119.33333333333333\n2017-08-12,40,6,224,2017,114.63333333333334\n2017-08-13,94,7,225,2017,112.53333333333333\n2017-08-14,149,1,226,2017,113.66666666666667\n2017-08-15,146,2,227,2017,113.46666666666667\n2017-08-16,167,3,228,2017,115.13333333333334\n2017-08-17,160,4,229,2017,114.86666666666666\n2017-08-18,120,5,230,2017,113.46666666666667\n2017-08-19,31,6,231,2017,109.5\n2017-08-20,43,7,232,2017,106.2\n2017-08-21,84,1,233,2017,106.13333333333334\n2017-08-22,114,2,234,2017,106.8\n2017-08-23,144,3,235,2017,107.83333333333333\n2017-08-24,114,4,236,2017,106.56666666666666\n2017-08-25,118,5,237,2017,107.26666666666667\n2017-08-26,48,6,238,2017,104.23333333333333\n2017-08-27,85,7,239,2017,103.46666666666667\n2017-08-28,94,1,240,2017,104.76666666666667\n2017-08-29,119,2,241,2017,106.73333333333333\n2017-08-30,135,3,242,2017,106.83333333333333\n2017-08-31,127,4,243,2017,105.23333333333333\n2017-09-01,116,5,244,2017,105.53333333333333\n2017-09-02,84,6,245,2017,104.4\n2017-09-03,70,7,246,2017,102.63333333333334\n2017-09-04,113,1,247,2017,103.46666666666667\n2017-09-05,123,2,248,2017,106.36666666666666\n2017-09-06,99,3,249,2017,105.3\n2017-09-07,95,4,250,2017,104.5\n2017-09-08,125,5,251,2017,104.8\n2017-09-09,59,6,252,2017,104.4\n2017-09-10,56,7,253,2017,102.4\n2017-09-11,148,1,254,2017,106.0\n2017-09-12,127,2,255,2017,107.1\n2017-09-13,123,3,256,2017,106.23333333333333\n2017-09-14,120,4,257,2017,105.36666666666666\n2017-09-15,129,5,258,2017,104.1\n2017-09-16,40,6,259,2017,100.1\n2017-09-17,59,7,260,2017,98.06666666666666\n2017-09-18,115,1,261,2017,100.86666666666666\n2017-09-19,137,2,262,2017,104.0\n2017-09-20,98,3,263,2017,104.46666666666667\n2017-09-21,96,4,264,2017,103.86666666666666\n2017-09-22,82,5,265,2017,101.8\n2017-09-23,88,6,266,2017,100.93333333333334\n2017-09-24,86,7,267,2017,99.86666666666666\n2017-09-25,144,1,268,2017,103.06666666666666\n2017-09-26,169,2,269,2017,105.86666666666666\n2017-09-27,150,3,270,2017,107.73333333333333\n2017-09-28,151,4,271,2017,108.8\n2017-09-29,136,5,272,2017,108.83333333333333\n2017-09-30,95,6,273,2017,107.76666666666667\n2017-10-01,94,7,274,2017,107.03333333333333\n2017-10-02,138,1,275,2017,108.83333333333333\n2017-10-03,165,2,276,2017,112.0\n2017-10-04,114,3,277,2017,112.03333333333333\n2017-10-05,95,4,278,2017,111.1\n2017-10-06,113,5,279,2017,111.56666666666666\n2017-10-07,80,6,280,2017,111.06666666666666\n2017-10-08,95,7,281,2017,110.06666666666666\n2017-10-09,122,1,282,2017,112.16666666666667\n2017-10-10,175,2,283,2017,116.13333333333334\n2017-10-11,127,3,284,2017,115.43333333333334\n2017-10-12,179,4,285,2017,117.16666666666667\n2017-10-13,197,5,286,2017,119.63333333333334\n2017-10-14,84,6,287,2017,118.43333333333334\n2017-10-15,60,7,288,2017,116.13333333333334\n2017-10-16,128,1,289,2017,119.06666666666666\n2017-10-17,138,2,290,2017,121.7\n2017-10-18,154,3,291,2017,123.0\n2017-10-19,163,4,292,2017,123.86666666666666\n2017-10-20,118,5,293,2017,124.53333333333333\n2017-10-21,82,6,294,2017,124.06666666666666\n2017-10-22,49,7,295,2017,122.96666666666667\n2017-10-23,96,1,296,2017,123.23333333333333\n2017-10-24,142,2,297,2017,125.1\n2017-10-25,160,3,298,2017,125.63333333333334\n2017-10-26,140,4,299,2017,124.66666666666667\n2017-10-27,126,5,300,2017,123.86666666666666\n2017-10-28,81,6,301,2017,121.53333333333333\n2017-10-29,99,7,302,2017,120.3\n2017-10-30,107,1,303,2017,120.7\n2017-10-31,157,2,304,2017,122.8\n2017-11-01,157,3,305,2017,123.43333333333334\n2017-11-02,151,4,306,2017,122.96666666666667\n2017-11-03,136,5,307,2017,123.7\n2017-11-04,76,6,308,2017,123.06666666666666\n2017-11-05,63,7,309,2017,121.4\n2017-11-06,155,1,310,2017,123.9\n2017-11-07,124,2,311,2017,124.86666666666666\n2017-11-08,153,3,312,2017,125.9\n2017-11-09,119,4,313,2017,124.03333333333333\n2017-11-10,90,5,314,2017,122.8\n2017-11-11,57,6,315,2017,118.73333333333333\n2017-11-12,86,7,316,2017,115.03333333333333\n2017-11-13,129,1,317,2017,116.53333333333333\n2017-11-14,142,2,318,2017,119.26666666666667\n2017-11-15,137,3,319,2017,119.56666666666666\n2017-11-16,147,4,320,2017,119.86666666666666\n2017-11-17,175,5,321,2017,120.56666666666666\n2017-11-18,101,6,322,2017,118.5\n2017-11-19,95,7,323,2017,117.73333333333333\n2017-11-20,140,1,324,2017,119.66666666666667\n2017-11-21,127,2,325,2017,122.26666666666667\n2017-11-22,152,3,326,2017,124.13333333333334\n2017-11-23,98,4,327,2017,122.66666666666667\n2017-11-24,90,5,328,2017,120.33333333333333\n2017-11-25,59,6,329,2017,117.63333333333334\n2017-11-26,86,7,330,2017,116.3\n2017-11-27,95,1,331,2017,116.76666666666667\n2017-11-28,114,2,332,2017,117.26666666666667\n2017-11-29,140,3,333,2017,118.36666666666666\n2017-11-30,131,4,334,2017,117.5\n2017-12-01,129,5,335,2017,116.56666666666666\n2017-12-02,98,6,336,2017,114.8\n2017-12-03,85,7,337,2017,113.1\n2017-12-04,150,1,338,2017,115.56666666666666\n2017-12-05,174,2,339,2017,119.26666666666667\n2017-12-06,168,3,340,2017,119.7\n2017-12-07,96,4,341,2017,118.76666666666667\n2017-12-08,71,5,342,2017,116.03333333333333\n2017-12-09,96,6,343,2017,115.26666666666667\n2017-12-10,106,7,344,2017,115.8\n2017-12-11,133,1,345,2017,118.33333333333333\n2017-12-12,144,2,346,2017,120.26666666666667\n2017-12-13,96,3,347,2017,119.16666666666667\n2017-12-14,102,4,348,2017,117.83333333333333\n2017-12-15,133,5,349,2017,117.7\n2017-12-16,69,6,350,2017,115.1\n2017-12-17,80,7,351,2017,111.93333333333334\n2017-12-18,83,1,352,2017,111.33333333333333\n2017-12-19,107,2,353,2017,111.73333333333333\n2017-12-20,137,3,354,2017,111.63333333333334\n2017-12-21,132,4,355,2017,111.8\n2017-12-22,108,5,356,2017,110.33333333333333\n2017-12-23,34,6,357,2017,108.2\n2017-12-24,37,7,358,2017,106.43333333333334\n2017-12-25,35,1,359,2017,105.63333333333334\n2017-12-26,65,2,360,2017,104.93333333333334\n2017-12-27,129,3,361,2017,106.06666666666666\n2017-12-28,126,4,362,2017,106.46666666666667\n2017-12-29,86,5,363,2017,104.66666666666667\n2017-12-30,99,6,364,2017,103.6\n2017-12-31,60,7,365,2017,101.3\n2018-01-01,44,1,1,2018,99.5\n2018-01-02,79,2,2,2018,99.3\n2018-01-03,93,3,3,2018,97.4\n2018-01-04,131,4,4,2018,95.96666666666667\n2018-01-05,125,5,5,2018,94.53333333333333\n2018-01-06,63,6,6,2018,93.43333333333334\n2018-01-07,58,7,7,2018,93.0\n2018-01-08,180,1,8,2018,95.8\n2018-01-09,139,2,9,2018,96.9\n2018-01-10,155,3,10,2018,97.63333333333334\n2018-01-11,128,4,11,2018,97.1\n2018-01-12,92,5,12,2018,96.96666666666667\n2018-01-13,116,6,13,2018,97.43333333333334\n2018-01-14,92,7,14,2018,96.06666666666666\n2018-01-15,116,1,15,2018,97.63333333333334\n2018-01-16,106,2,16,2018,98.5\n2018-01-17,145,3,17,2018,100.56666666666666\n2018-01-18,152,4,18,2018,102.06666666666666\n2018-01-19,110,5,19,2018,101.16666666666667\n2018-01-20,100,6,20,2018,100.1\n2018-01-21,67,7,21,2018,98.73333333333333\n2018-01-22,154,1,22,2018,102.73333333333333\n2018-01-23,122,2,23,2018,105.56666666666666\n2018-01-24,215,3,24,2018,111.56666666666666\n2018-01-25,153,4,25,2018,114.5\n2018-01-26,127,5,26,2018,114.43333333333334\n2018-01-27,81,6,27,2018,112.93333333333334\n2018-01-28,79,7,28,2018,112.7\n2018-01-29,146,1,29,2018,114.26666666666667\n2018-01-30,159,2,30,2018,117.56666666666666\n2018-01-31,128,3,31,2018,120.36666666666666\n2018-02-01,97,4,32,2018,120.96666666666667\n2018-02-02,124,5,33,2018,122.0\n2018-02-03,51,6,34,2018,119.33333333333333\n2018-02-04,82,7,35,2018,117.9\n2018-02-05,163,1,36,2018,121.23333333333333\n2018-02-06,183,2,37,2018,125.4\n2018-02-07,107,3,38,2018,122.96666666666667\n2018-02-08,151,4,39,2018,123.36666666666666\n2018-02-09,174,5,40,2018,124.0\n2018-02-10,147,6,41,2018,124.63333333333334\n2018-02-11,131,7,42,2018,125.93333333333334\n2018-02-12,135,1,43,2018,126.56666666666666\n2018-02-13,170,2,44,2018,129.16666666666666\n2018-02-14,151,3,45,2018,130.33333333333334\n2018-02-15,180,4,46,2018,132.8\n2018-02-16,128,5,47,2018,132.23333333333332\n2018-02-17,125,6,48,2018,131.33333333333334\n2018-02-18,119,7,49,2018,131.63333333333333\n2018-02-19,128,1,50,2018,132.56666666666666\n2018-02-20,148,2,51,2018,135.26666666666668\n2018-02-21,115,3,52,2018,133.96666666666667\n2018-02-22,136,4,53,2018,134.43333333333334\n2018-02-23,173,5,54,2018,133.03333333333333\n2018-02-24,103,6,55,2018,131.36666666666667\n2018-02-25,115,7,56,2018,130.96666666666667\n2018-02-26,168,1,57,2018,133.86666666666667\n2018-02-27,174,2,58,2018,137.03333333333333\n2018-02-28,180,3,59,2018,138.16666666666666\n2018-03-01,126,4,60,2018,137.06666666666666\n2018-03-02,168,5,61,2018,138.4\n2018-03-03,112,6,62,2018,138.9\n2018-03-04,76,7,63,2018,137.3\n2018-03-05,164,1,64,2018,141.06666666666666\n2018-03-06,123,2,65,2018,142.43333333333334\n2018-03-07,143,3,66,2018,141.76666666666668\n2018-03-08,138,4,67,2018,140.26666666666668\n2018-03-09,161,5,68,2018,142.06666666666666\n2018-03-10,111,6,69,2018,140.73333333333332\n2018-03-11,47,7,70,2018,136.5\n2018-03-12,175,1,71,2018,137.43333333333334\n2018-03-13,146,2,72,2018,137.93333333333334\n2018-03-14,175,3,73,2018,139.26666666666668\n2018-03-15,141,4,74,2018,138.3\n2018-03-16,135,5,75,2018,137.76666666666668\n2018-03-17,116,6,76,2018,135.63333333333333\n2018-03-18,113,7,77,2018,135.13333333333333\n2018-03-19,199,1,78,2018,137.6\n2018-03-20,199,2,79,2018,140.26666666666668\n2018-03-21,207,3,80,2018,142.9\n2018-03-22,147,4,81,2018,142.86666666666667\n2018-03-23,145,5,82,2018,143.86666666666667\n2018-03-24,161,6,83,2018,144.7\n2018-03-25,131,7,84,2018,143.3\n2018-03-26,170,1,85,2018,145.53333333333333\n2018-03-27,160,2,86,2018,147.03333333333333\n2018-03-28,192,3,87,2018,147.83333333333334\n2018-03-29,142,4,88,2018,146.76666666666668\n2018-03-30,152,5,89,2018,145.83333333333334\n2018-03-31,151,6,90,2018,146.66666666666666\n2018-04-01,79,7,91,2018,143.7\n2018-04-02,150,1,92,2018,144.96666666666667\n2018-04-03,229,2,93,2018,150.06666666666666\n2018-04-04,209,3,94,2018,151.56666666666666\n2018-04-05,157,4,95,2018,152.7\n2018-04-06,221,5,96,2018,155.3\n2018-04-07,119,6,97,2018,154.66666666666666\n2018-04-08,162,7,98,2018,154.7\n2018-04-09,133,1,99,2018,155.43333333333334\n2018-04-10,196,2,100,2018,160.4\n2018-04-11,172,3,101,2018,160.3\n2018-04-12,206,4,102,2018,162.3\n2018-04-13,203,5,103,2018,163.23333333333332\n2018-04-14,89,6,104,2018,161.5\n2018-04-15,98,7,105,2018,160.26666666666668\n2018-04-16,176,1,106,2018,162.26666666666668\n2018-04-17,166,2,107,2018,164.03333333333333\n2018-04-18,155,3,108,2018,162.56666666666666\n2018-04-19,140,4,109,2018,160.6\n2018-04-20,166,5,110,2018,159.23333333333332\n2018-04-21,69,6,111,2018,156.63333333333333\n2018-04-22,91,7,112,2018,154.83333333333334\n2018-04-23,156,1,113,2018,154.66666666666666\n2018-04-24,199,2,114,2018,156.93333333333334\n2018-04-25,144,3,115,2018,156.06666666666666\n2018-04-26,148,4,116,2018,155.66666666666666\n2018-04-27,147,5,117,2018,154.16666666666666\n2018-04-28,145,6,118,2018,154.26666666666668\n2018-04-29,161,7,119,2018,154.56666666666666\n2018-04-30,176,1,120,2018,155.4\n2018-05-01,197,2,121,2018,159.33333333333334\n2018-05-02,175,3,122,2018,160.16666666666666\n2018-05-03,162,4,123,2018,157.93333333333334\n2018-05-04,159,5,124,2018,156.26666666666668\n2018-05-05,68,6,125,2018,153.3\n2018-05-06,100,7,126,2018,149.26666666666668\n2018-05-07,184,1,127,2018,151.43333333333334\n2018-05-08,227,2,128,2018,153.6\n2018-05-09,129,3,129,2018,153.46666666666667\n2018-05-10,150,4,130,2018,151.93333333333334\n2018-05-11,195,5,131,2018,152.7\n2018-05-12,115,6,132,2018,149.66666666666666\n2018-05-13,77,7,133,2018,145.46666666666667\n2018-05-14,178,1,134,2018,148.43333333333334\n2018-05-15,194,2,135,2018,151.63333333333333\n2018-05-16,178,3,136,2018,151.7\n2018-05-17,198,4,137,2018,152.76666666666668\n2018-05-18,150,5,138,2018,152.6\n2018-05-19,70,6,139,2018,150.26666666666668\n2018-05-20,68,7,140,2018,147.0\n2018-05-21,140,1,141,2018,149.36666666666667\n2018-05-22,175,2,142,2018,152.16666666666666\n2018-05-23,186,3,143,2018,153.16666666666666\n2018-05-24,152,4,144,2018,151.6\n2018-05-25,116,5,145,2018,150.66666666666666\n2018-05-26,38,6,146,2018,147.0\n2018-05-27,136,7,147,2018,146.63333333333333\n2018-05-28,169,1,148,2018,147.43333333333334\n2018-05-29,149,2,149,2018,147.03333333333333\n2018-05-30,170,3,150,2018,146.83333333333334\n2018-05-31,229,4,151,2018,147.9\n2018-06-01,228,5,152,2018,149.66666666666666\n2018-06-02,50,6,153,2018,145.93333333333334\n2018-06-03,58,7,154,2018,142.56666666666666\n2018-06-04,167,1,155,2018,145.86666666666667\n2018-06-05,150,2,156,2018,147.53333333333333\n2018-06-06,200,3,157,2018,148.06666666666666\n2018-06-07,232,4,158,2018,148.23333333333332\n2018-06-08,186,5,159,2018,150.13333333333333\n2018-06-09,113,6,160,2018,148.9\n2018-06-10,88,7,161,2018,145.33333333333334\n2018-06-11,160,1,162,2018,146.83333333333334\n2018-06-12,216,2,163,2018,151.46666666666667\n2018-06-13,180,3,164,2018,151.53333333333333\n2018-06-14,176,4,165,2018,150.93333333333334\n2018-06-15,178,5,166,2018,150.93333333333334\n2018-06-16,93,6,167,2018,147.43333333333334\n2018-06-17,51,7,168,2018,144.13333333333333\n2018-06-18,112,1,169,2018,145.53333333333333\n2018-06-19,117,2,170,2018,147.16666666666666\n2018-06-20,135,3,171,2018,147.0\n2018-06-21,126,4,172,2018,145.36666666666667\n2018-06-22,94,5,173,2018,142.3\n2018-06-23,73,6,174,2018,139.66666666666666\n2018-06-24,81,7,175,2018,138.5\n2018-06-25,137,1,176,2018,141.8\n2018-06-26,208,2,177,2018,144.2\n2018-06-27,127,3,178,2018,142.8\n2018-06-28,152,4,179,2018,142.9\n2018-06-29,134,5,180,2018,141.7\n2018-06-30,98,6,181,2018,137.33333333333334\n2018-07-01,128,7,182,2018,134.0\n2018-07-02,144,1,183,2018,137.13333333333333\n2018-07-03,127,2,184,2018,139.43333333333334\n2018-07-04,70,3,185,2018,136.2\n2018-07-05,117,4,186,2018,135.1\n2018-07-06,115,5,187,2018,132.26666666666668\n2018-07-07,55,6,188,2018,126.36666666666666\n2018-07-08,57,7,189,2018,122.06666666666666\n2018-07-09,104,1,190,2018,121.76666666666667\n2018-07-10,133,2,191,2018,123.26666666666667\n2018-07-11,238,3,192,2018,125.86666666666666\n2018-07-12,135,4,193,2018,123.16666666666667\n2018-07-13,244,5,194,2018,125.3\n2018-07-14,74,6,195,2018,121.9\n2018-07-15,86,7,196,2018,118.83333333333333\n2018-07-16,178,1,197,2018,121.66666666666667\n2018-07-17,195,2,198,2018,126.46666666666667\n2018-07-18,201,3,199,2018,129.43333333333334\n2018-07-19,149,4,200,2018,130.5\n2018-07-20,122,5,201,2018,130.06666666666666\n2018-07-21,89,6,202,2018,128.83333333333334\n2018-07-22,81,7,203,2018,128.4\n2018-07-23,169,1,204,2018,131.6\n2018-07-24,170,2,205,2018,134.56666666666666\n2018-07-25,134,3,206,2018,134.46666666666667\n2018-07-26,136,4,207,2018,132.06666666666666\n2018-07-27,135,5,208,2018,132.33333333333334\n2018-07-28,79,6,209,2018,129.9\n2018-07-29,79,7,210,2018,128.06666666666666\n2018-07-30,144,1,211,2018,129.6\n2018-07-31,120,2,212,2018,129.33333333333334\n2018-08-01,150,3,213,2018,129.53333333333333\n2018-08-02,108,4,214,2018,128.9\n2018-08-03,147,5,215,2018,131.46666666666667\n2018-08-04,65,6,216,2018,129.73333333333332\n2018-08-05,66,7,217,2018,128.1\n2018-08-06,139,1,218,2018,130.9\n2018-08-07,142,2,219,2018,133.73333333333332\n2018-08-08,117,3,220,2018,134.16666666666666\n2018-08-09,79,4,221,2018,132.36666666666667\n2018-08-10,84,5,222,2018,127.23333333333333\n2018-08-11,103,6,223,2018,126.16666666666667\n2018-08-12,79,7,224,2018,120.66666666666667\n2018-08-13,99,1,225,2018,121.5\n2018-08-14,126,2,226,2018,122.83333333333333\n2018-08-15,106,3,227,2018,120.43333333333334\n2018-08-16,143,4,228,2018,118.7\n2018-08-17,114,5,229,2018,115.8\n2018-08-18,68,6,230,2018,113.1\n2018-08-19,34,7,231,2018,110.16666666666667\n2018-08-20,83,1,232,2018,109.96666666666667\n2018-08-21,75,2,233,2018,109.76666666666667\n2018-08-22,108,3,234,2018,107.73333333333333\n2018-08-23,84,4,235,2018,104.86666666666666\n2018-08-24,85,5,236,2018,103.23333333333333\n2018-08-25,66,6,237,2018,100.9\n2018-08-26,50,7,238,2018,98.06666666666666\n2018-08-27,77,1,239,2018,98.0\n2018-08-28,114,2,240,2018,99.16666666666667\n2018-08-29,116,3,241,2018,98.23333333333333\n2018-08-30,112,4,242,2018,97.96666666666667\n2018-08-31,114,5,243,2018,96.76666666666667\n2018-09-01,99,6,244,2018,96.46666666666667\n2018-09-02,92,7,245,2018,94.63333333333334\n2018-09-03,87,1,246,2018,95.36666666666666\n2018-09-04,122,2,247,2018,97.23333333333333\n2018-09-05,125,3,248,2018,96.76666666666667\n2018-09-06,104,4,249,2018,95.5\n2018-09-07,139,5,250,2018,96.23333333333333\n2018-09-08,83,6,251,2018,96.36666666666666\n2018-09-09,57,7,252,2018,95.46666666666667\n2018-09-10,110,1,253,2018,95.7\n2018-09-11,119,2,254,2018,97.03333333333333\n2018-09-12,119,3,255,2018,97.7\n2018-09-13,141,4,256,2018,98.2\n2018-09-14,147,5,257,2018,99.56666666666666\n2018-09-15,67,6,258,2018,97.03333333333333\n2018-09-16,96,7,259,2018,96.43333333333334\n2018-09-17,104,1,260,2018,97.63333333333334\n2018-09-18,140,2,261,2018,101.16666666666667\n2018-09-19,159,3,262,2018,103.7\n2018-09-20,125,4,263,2018,105.36666666666666\n2018-09-21,141,5,264,2018,106.46666666666667\n2018-09-22,61,6,265,2018,105.7\n2018-09-23,84,7,266,2018,105.66666666666667\n2018-09-24,139,1,267,2018,108.1\n2018-09-25,107,2,268,2018,110.0\n2018-09-26,159,3,269,2018,112.73333333333333\n2018-09-27,140,4,270,2018,113.6\n2018-09-28,144,5,271,2018,114.53333333333333\n2018-09-29,92,6,272,2018,113.86666666666666\n2018-09-30,95,7,273,2018,113.23333333333333\n2018-10-01,157,1,274,2018,115.16666666666667\n2018-10-02,173,2,275,2018,117.86666666666666\n2018-10-03,186,3,276,2018,121.16666666666667\n2018-10-04,155,4,277,2018,122.26666666666667\n2018-10-05,137,5,278,2018,122.66666666666667\n2018-10-06,65,6,279,2018,121.36666666666666\n2018-10-07,35,7,280,2018,117.9\n2018-10-08,135,1,281,2018,119.63333333333334\n2018-10-09,172,2,282,2018,123.46666666666667\n2018-10-10,130,3,283,2018,124.13333333333334\n2018-10-11,119,4,284,2018,124.13333333333334\n2018-10-12,140,5,285,2018,124.83333333333333\n2018-10-13,98,6,286,2018,123.4\n2018-10-14,66,7,287,2018,120.7\n2018-10-15,134,1,288,2018,122.93333333333334\n2018-10-16,183,2,289,2018,125.83333333333333\n2018-10-17,152,3,290,2018,127.43333333333334\n2018-10-18,150,4,291,2018,127.76666666666667\n2018-10-19,146,5,292,2018,127.33333333333333\n2018-10-20,125,6,293,2018,127.33333333333333\n2018-10-21,146,7,294,2018,127.5\n2018-10-22,134,1,295,2018,129.93333333333334\n2018-10-23,152,2,296,2018,132.2\n2018-10-24,143,3,297,2018,132.33333333333334\n2018-10-25,116,4,298,2018,132.63333333333333\n2018-10-26,133,5,299,2018,131.76666666666668\n2018-10-27,136,6,300,2018,131.63333333333333\n2018-10-28,119,7,301,2018,130.8\n2018-10-29,149,1,302,2018,132.7\n2018-10-30,250,2,303,2018,137.86666666666667\n2018-10-31,174,3,304,2018,138.43333333333334\n2018-11-01,251,4,305,2018,141.03333333333333\n2018-11-02,168,5,306,2018,140.43333333333334\n2018-11-03,86,6,307,2018,138.13333333333333\n2018-11-04,287,7,308,2018,143.13333333333333\n2018-11-05,157,1,309,2018,146.2\n2018-11-06,161,2,310,2018,150.4\n2018-11-07,157,3,311,2018,151.13333333333333\n2018-11-08,158,4,312,2018,150.66666666666666\n2018-11-09,205,5,313,2018,153.16666666666666\n2018-11-10,168,6,314,2018,154.8\n2018-11-11,235,7,315,2018,157.96666666666667\n2018-11-12,95,1,316,2018,157.86666666666667\n2018-11-13,179,2,317,2018,161.63333333333333\n2018-11-14,205,3,318,2018,164.0\n2018-11-15,171,4,319,2018,163.6\n2018-11-16,171,5,320,2018,164.23333333333332\n2018-11-17,86,6,321,2018,162.1\n2018-11-18,67,7,322,2018,159.46666666666667\n2018-11-19,160,1,323,2018,160.63333333333333\n2018-11-20,210,2,324,2018,162.76666666666668\n2018-11-21,177,3,325,2018,164.2\n2018-11-22,80,4,326,2018,161.8\n2018-11-23,121,5,327,2018,161.06666666666666\n2018-11-24,92,6,328,2018,160.26666666666668\n2018-11-25,117,7,329,2018,159.73333333333332\n2018-11-26,131,1,330,2018,159.56666666666666\n2018-11-27,143,2,331,2018,160.36666666666667\n2018-11-28,169,3,332,2018,161.03333333333333\n2018-11-29,177,4,333,2018,158.6\n2018-11-30,188,5,334,2018,159.06666666666666\n2018-12-01,153,6,335,2018,155.8\n2018-12-02,128,7,336,2018,154.46666666666667\n2018-12-03,225,1,337,2018,159.1\n2018-12-04,137,2,338,2018,154.1\n2018-12-05,219,3,339,2018,156.16666666666666\n2018-12-06,163,4,340,2018,156.23333333333332\n2018-12-07,169,5,341,2018,156.63333333333333\n2018-12-08,59,6,342,2018,153.33333333333334\n2018-12-09,67,7,343,2018,148.73333333333332\n2018-12-10,142,1,344,2018,147.86666666666667\n2018-12-11,130,2,345,2018,144.36666666666667\n2018-12-12,214,3,346,2018,148.33333333333334\n2018-12-13,115,4,347,2018,146.2\n2018-12-14,110,5,348,2018,143.03333333333333\n2018-12-15,95,6,349,2018,140.5\n2018-12-16,91,7,350,2018,137.83333333333334\n2018-12-17,166,1,351,2018,140.5\n2018-12-18,164,2,352,2018,143.73333333333332\n2018-12-19,129,3,353,2018,142.7\n2018-12-20,132,4,354,2018,140.1\n2018-12-21,82,5,355,2018,136.93333333333334\n2018-12-22,75,6,356,2018,136.76666666666668\n2018-12-23,49,7,357,2018,134.36666666666667\n2018-12-24,47,1,358,2018,132.86666666666667\n2018-12-25,25,2,359,2018,129.8\n2018-12-26,50,3,360,2018,127.1\n2018-12-27,103,4,361,2018,125.76666666666667\n2018-12-28,123,5,362,2018,124.23333333333333\n2018-12-29,60,6,363,2018,120.33333333333333\n2018-12-30,92,7,364,2018,117.13333333333334\n2018-12-31,91,1,365,2018,115.06666666666666\n2019-01-01,73,2,1,2019,113.23333333333333\n2019-01-02,133,3,2,2019,110.16666666666667\n2019-01-03,107,4,3,2019,109.16666666666667\n2019-01-04,141,5,4,2019,106.56666666666666\n2019-01-05,74,6,5,2019,103.6\n2019-01-06,93,7,6,2019,101.06666666666666\n2019-01-07,113,1,7,2019,102.86666666666666\n2019-01-08,137,2,8,2019,105.2\n2019-01-09,178,3,9,2019,106.4\n2019-01-10,185,4,10,2019,108.23333333333333\n2019-01-11,133,5,11,2019,105.53333333333333\n2019-01-12,102,6,12,2019,105.1\n2019-01-13,83,7,13,2019,104.2\n2019-01-14,155,1,14,2019,106.2\n2019-01-15,149,2,15,2019,108.13333333333334\n2019-01-16,139,3,16,2019,107.23333333333333\n2019-01-17,175,4,17,2019,107.6\n2019-01-18,165,5,18,2019,108.8\n2019-01-19,158,6,19,2019,109.66666666666667\n2019-01-20,101,7,20,2019,110.3\n2019-01-21,181,1,21,2019,113.83333333333333\n2019-01-22,186,2,22,2019,118.4\n2019-01-23,164,3,23,2019,122.3\n2019-01-24,173,4,24,2019,127.23333333333333\n2019-01-25,112,5,25,2019,129.3\n2019-01-26,61,6,26,2019,127.9\n2019-01-27,83,7,27,2019,126.56666666666666\n2019-01-28,103,1,28,2019,128.0\n2019-01-29,106,2,29,2019,128.46666666666667\n2019-01-30,117,3,30,2019,129.33333333333334\n2019-01-31,93,4,31,2019,130.0\n2019-02-01,156,5,32,2019,130.76666666666668\n2019-02-02,101,6,33,2019,130.56666666666666\n2019-02-03,101,7,34,2019,129.23333333333332\n2019-02-04,138,1,35,2019,131.36666666666667\n2019-02-05,129,2,36,2019,132.56666666666666\n2019-02-06,158,3,37,2019,134.06666666666666\n2019-02-07,123,4,38,2019,133.6\n2019-02-08,175,5,39,2019,133.5\n2019-02-09,115,6,40,2019,131.16666666666666\n2019-02-10,122,7,41,2019,130.8\n2019-02-11,150,1,42,2019,132.4\n2019-02-12,151,2,43,2019,134.66666666666666\n2019-02-13,157,3,44,2019,134.73333333333332\n2019-02-14,150,4,45,2019,134.76666666666668\n2019-02-15,164,5,46,2019,135.6\n2019-02-16,178,6,47,2019,135.7\n2019-02-17,100,7,48,2019,133.53333333333333\n2019-02-18,161,1,49,2019,133.63333333333333\n2019-02-19,226,2,50,2019,137.8\n2019-02-20,197,3,51,2019,138.33333333333334\n2019-02-21,217,4,52,2019,139.36666666666667\n2019-02-22,154,5,53,2019,139.03333333333333\n2019-02-23,91,6,54,2019,136.3\n2019-02-24,102,7,55,2019,135.96666666666667\n2019-02-25,241,1,56,2019,141.96666666666667\n2019-02-26,164,2,57,2019,144.66666666666666\n2019-02-27,191,3,58,2019,147.6\n2019-02-28,164,4,59,2019,149.53333333333333\n2019-03-01,117,5,60,2019,149.53333333333333\n2019-03-02,43,6,61,2019,147.86666666666667\n2019-03-03,150,7,62,2019,147.66666666666666\n2019-03-04,129,1,63,2019,148.6\n2019-03-05,128,2,64,2019,149.5\n2019-03-06,113,3,65,2019,148.66666666666666\n2019-03-07,109,4,66,2019,148.0\n2019-03-08,101,5,67,2019,146.1\n2019-03-09,109,6,68,2019,145.63333333333333\n2019-03-10,85,7,69,2019,142.63333333333333\n2019-03-11,133,1,70,2019,143.23333333333332\n2019-03-12,110,2,71,2019,142.83333333333334\n2019-03-13,104,3,72,2019,141.3\n2019-03-14,104,4,73,2019,139.73333333333332\n2019-03-15,116,5,74,2019,138.36666666666667\n2019-03-16,66,6,75,2019,135.56666666666666\n2019-03-17,66,7,76,2019,132.3\n2019-03-18,105,1,77,2019,129.86666666666667\n2019-03-19,119,2,78,2019,130.5\n2019-03-20,137,3,79,2019,129.7\n2019-03-21,172,4,80,2019,127.9\n2019-03-22,118,5,81,2019,125.26666666666667\n2019-03-23,73,6,82,2019,120.46666666666667\n2019-03-24,50,7,83,2019,117.0\n2019-03-25,141,1,84,2019,118.66666666666667\n2019-03-26,159,2,85,2019,120.56666666666666\n2019-03-27,107,3,86,2019,116.1\n2019-03-28,100,4,87,2019,113.96666666666667\n2019-03-29,134,5,88,2019,112.06666666666666\n2019-03-30,92,6,89,2019,109.66666666666667\n2019-03-31,70,7,90,2019,108.1\n2019-04-01,139,1,91,2019,111.3\n2019-04-02,114,2,92,2019,110.1\n2019-04-03,113,3,93,2019,109.56666666666666\n2019-04-04,97,4,94,2019,108.53333333333333\n2019-04-05,125,5,95,2019,108.93333333333334\n2019-04-06,37,6,96,2019,106.53333333333333\n2019-04-07,51,7,97,2019,104.86666666666666\n2019-04-08,154,1,98,2019,106.36666666666666\n2019-04-09,118,2,99,2019,107.46666666666667\n2019-04-10,169,3,100,2019,108.66666666666667\n2019-04-11,109,4,101,2019,108.63333333333334\n2019-04-12,107,5,102,2019,108.73333333333333\n2019-04-13,42,6,103,2019,106.66666666666667\n2019-04-14,58,7,104,2019,104.73333333333333\n2019-04-15,108,1,105,2019,106.13333333333334\n2019-04-16,135,2,106,2019,108.43333333333334\n2019-04-17,73,3,107,2019,107.36666666666666\n2019-04-18,114,4,108,2019,107.2\n2019-04-19,112,5,109,2019,106.36666666666666\n2019-04-20,92,6,110,2019,103.7\n2019-04-21,70,7,111,2019,102.1\n2019-04-22,104,1,112,2019,103.13333333333334\n2019-04-23,84,2,113,2019,104.26666666666667\n2019-04-24,100,3,114,2019,102.9\n2019-04-25,115,4,115,2019,101.43333333333334\n2019-04-26,92,5,116,2019,100.93333333333334\n2019-04-27,44,6,117,2019,99.06666666666666\n2019-04-28,106,7,118,2019,98.13333333333334\n2019-04-29,126,1,119,2019,99.26666666666667\n2019-04-30,78,2,120,2019,99.53333333333333\n2019-05-01,116,3,121,2019,98.76666666666667\n2019-05-02,108,4,122,2019,98.56666666666666\n2019-05-03,78,5,123,2019,97.4\n2019-05-04,50,6,124,2019,95.83333333333333\n2019-05-05,85,7,125,2019,94.5\n2019-05-06,115,1,126,2019,97.1\n2019-05-07,120,2,127,2019,99.4\n2019-05-08,124,3,128,2019,98.4\n2019-05-09,108,4,129,2019,98.06666666666666\n2019-05-10,80,5,130,2019,95.1\n2019-05-11,86,6,131,2019,94.33333333333333\n2019-05-12,69,7,132,2019,93.06666666666666\n2019-05-13,102,1,133,2019,95.06666666666666\n2019-05-14,100,2,134,2019,96.46666666666667\n2019-05-15,80,3,135,2019,95.53333333333333\n2019-05-16,99,4,136,2019,94.33333333333333\n2019-05-17,89,5,137,2019,94.86666666666666\n2019-05-18,51,6,138,2019,92.76666666666667\n2019-05-19,64,7,139,2019,91.16666666666667\n2019-05-20,111,1,140,2019,91.8\n2019-05-21,97,2,141,2019,92.7\n2019-05-22,103,3,142,2019,92.66666666666667\n2019-05-23,141,4,143,2019,94.56666666666666\n2019-05-24,68,5,144,2019,93.5\n2019-05-25,75,6,145,2019,92.16666666666667\n2019-05-26,75,7,146,2019,91.6\n2019-05-27,58,1,147,2019,92.06666666666666\n2019-05-28,79,2,148,2019,91.16666666666667\n2019-05-29,87,3,149,2019,89.86666666666666\n2019-05-30,125,4,150,2019,91.43333333333334\n2019-05-31,120,5,151,2019,91.56666666666666\n2019-06-01,69,6,152,2019,90.26666666666667\n2019-06-02,74,7,153,2019,90.13333333333334\n2019-06-03,112,1,154,2019,92.2\n2019-06-04,88,2,155,2019,92.3\n2019-06-05,138,3,156,2019,93.06666666666666\n2019-06-06,95,4,157,2019,92.23333333333333\n2019-06-07,100,5,158,2019,91.43333333333334\n2019-06-08,95,6,159,2019,91.0\n2019-06-09,88,7,160,2019,91.26666666666667\n2019-06-10,97,1,161,2019,91.63333333333334\n2019-06-11,105,2,162,2019,92.83333333333333\n2019-06-12,104,3,163,2019,92.9\n2019-06-13,88,4,164,2019,92.5\n2019-06-14,89,5,165,2019,92.8\n2019-06-15,47,6,166,2019,91.06666666666666\n2019-06-16,41,7,167,2019,89.46666666666667\n2019-06-17,78,1,168,2019,90.36666666666666\n2019-06-18,68,2,169,2019,90.5\n2019-06-19,115,3,170,2019,90.63333333333334\n2019-06-20,86,4,171,2019,90.26666666666667\n2019-06-21,105,5,172,2019,90.33333333333333\n2019-06-22,43,6,173,2019,87.06666666666666\n2019-06-23,79,7,174,2019,87.43333333333334\n2019-06-24,92,1,175,2019,88.0\n2019-06-25,77,2,176,2019,88.06666666666666\n2019-06-26,97,3,177,2019,89.36666666666666\n2019-06-27,96,4,178,2019,89.93333333333334\n2019-06-28,92,5,179,2019,90.1\n2019-06-29,92,6,180,2019,89.0\n2019-06-30,57,7,181,2019,86.9\n2019-07-01,94,1,182,2019,87.73333333333333\n2019-07-02,164,2,183,2019,90.73333333333333\n2019-07-03,101,3,184,2019,90.36666666666666\n2019-07-04,89,4,185,2019,90.4\n2019-07-05,67,5,186,2019,88.03333333333333\n2019-07-06,59,6,187,2019,86.83333333333333\n2019-07-07,37,7,188,2019,84.73333333333333\n2019-07-08,146,1,189,2019,86.43333333333334\n2019-07-09,89,2,190,2019,86.46666666666667\n2019-07-10,136,3,191,2019,87.76666666666667\n2019-07-11,109,4,192,2019,87.9\n2019-07-12,102,5,193,2019,87.83333333333333\n2019-07-13,78,6,194,2019,87.5\n2019-07-14,65,7,195,2019,86.7\n2019-07-15,122,1,196,2019,89.2\n2019-07-16,142,2,197,2019,92.56666666666666\n2019-07-17,113,3,198,2019,93.73333333333333\n2019-07-18,103,4,199,2019,94.9\n2019-07-19,102,5,200,2019,94.46666666666667\n2019-07-20,92,6,201,2019,94.66666666666667\n2019-07-21,41,7,202,2019,92.53333333333333\n2019-07-22,97,1,203,2019,94.33333333333333\n2019-07-23,82,2,204,2019,94.43333333333334\n2019-07-24,124,3,205,2019,95.5\n2019-07-25,129,4,206,2019,97.23333333333333\n2019-07-26,115,5,207,2019,97.83333333333333\n2019-07-27,70,6,208,2019,96.96666666666667\n2019-07-28,67,7,209,2019,96.13333333333334\n2019-07-29,98,1,210,2019,96.33333333333333\n2019-07-30,106,2,211,2019,97.96666666666667\n2019-07-31,76,3,212,2019,97.36666666666666\n2019-08-01,140,4,213,2019,96.56666666666666\n2019-08-02,77,5,214,2019,95.76666666666667\n2019-08-03,86,6,215,2019,95.66666666666667\n2019-08-04,52,7,216,2019,95.16666666666667\n2019-08-05,159,1,217,2019,98.5\n2019-08-06,132,2,218,2019,101.66666666666667\n2019-08-07,142,3,219,2019,101.53333333333333\n2019-08-08,154,4,220,2019,103.7\n2019-08-09,146,5,221,2019,104.03333333333333\n2019-08-10,87,6,222,2019,103.3\n2019-08-11,124,7,223,2019,104.03333333333333\n2019-08-12,175,1,224,2019,107.26666666666667\n2019-08-13,204,2,225,2019,111.9\n2019-08-14,204,3,226,2019,114.63333333333334\n2019-08-15,150,4,227,2019,114.9\n2019-08-16,149,5,228,2019,116.1\n2019-08-17,114,6,229,2019,116.46666666666667\n2019-08-18,116,7,230,2019,116.93333333333334\n2019-08-19,131,1,231,2019,118.23333333333333\n2019-08-20,135,2,232,2019,121.36666666666666\n2019-08-21,162,3,233,2019,123.53333333333333\n2019-08-22,159,4,234,2019,126.1\n2019-08-23,193,5,235,2019,128.4\n2019-08-24,115,6,236,2019,127.93333333333334\n2019-08-25,59,7,237,2019,126.06666666666666\n2019-08-26,106,1,238,2019,127.26666666666667\n2019-08-27,123,2,239,2019,129.13333333333333\n2019-08-28,172,3,240,2019,131.6\n2019-08-29,120,4,241,2019,132.06666666666666\n2019-08-30,86,5,242,2019,132.4\n2019-08-31,75,6,243,2019,130.23333333333332\n2019-09-01,87,7,244,2019,130.56666666666666\n2019-09-02,126,1,245,2019,131.9\n2019-09-03,125,2,246,2019,134.33333333333334\n2019-09-04,144,3,247,2019,133.83333333333334\n2019-09-05,145,4,248,2019,134.26666666666668\n2019-09-06,109,5,249,2019,133.16666666666666\n2019-09-07,105,6,250,2019,131.53333333333333\n2019-09-08,111,7,251,2019,130.36666666666667\n2019-09-09,131,1,252,2019,131.83333333333334\n2019-09-10,144,2,253,2019,132.5\n2019-09-11,148,3,254,2019,131.6\n2019-09-12,104,4,255,2019,128.26666666666668\n2019-09-13,161,5,256,2019,126.83333333333333\n2019-09-14,83,6,257,2019,124.6\n2019-09-15,75,7,258,2019,122.13333333333334\n2019-09-16,109,1,259,2019,121.96666666666667\n2019-09-17,131,2,260,2019,122.46666666666667\n2019-09-18,112,3,261,2019,121.83333333333333\n2019-09-19,115,4,262,2019,121.16666666666667\n2019-09-20,167,5,263,2019,121.33333333333333\n2019-09-21,156,6,264,2019,121.23333333333333\n2019-09-22,94,7,265,2019,117.93333333333334\n2019-09-23,107,1,266,2019,117.66666666666667\n2019-09-24,147,2,267,2019,120.6\n2019-09-25,138,3,268,2019,121.66666666666667\n2019-09-26,144,4,269,2019,122.36666666666666\n2019-09-27,119,5,270,2019,120.6\n2019-09-28,29,6,271,2019,117.56666666666666\n2019-09-29,50,7,272,2019,116.36666666666666\n2019-09-30,132,1,273,2019,118.26666666666667\n2019-10-01,142,2,274,2019,120.1\n2019-10-02,138,3,275,2019,120.5\n2019-10-03,129,4,276,2019,120.63333333333334\n2019-10-04,110,5,277,2019,119.5\n2019-10-05,74,6,278,2019,117.13333333333334\n2019-10-06,70,7,279,2019,115.83333333333333\n2019-10-07,118,1,280,2019,116.26666666666667\n2019-10-08,114,2,281,2019,116.36666666666666\n2019-10-09,123,3,282,2019,116.1\n2019-10-10,137,4,283,2019,115.86666666666666\n2019-10-11,151,5,284,2019,115.96666666666667\n2019-10-12,70,6,285,2019,114.83333333333333\n2019-10-13,103,7,286,2019,112.9\n2019-10-14,110,1,287,2019,113.8\n2019-10-15,170,2,288,2019,116.96666666666667\n2019-10-16,130,3,289,2019,117.66666666666667\n2019-10-17,122,4,290,2019,117.36666666666666\n2019-10-18,122,5,291,2019,117.7\n2019-10-19,85,6,292,2019,116.7\n2019-10-20,51,7,293,2019,112.83333333333333\n2019-10-21,126,1,294,2019,111.83333333333333\n2019-10-22,94,2,295,2019,111.83333333333333\n2019-10-23,99,3,296,2019,111.56666666666666\n2019-10-24,154,4,297,2019,111.8\n2019-10-25,119,5,298,2019,111.16666666666667\n2019-10-26,105,6,299,2019,109.86666666666666\n2019-10-27,92,7,300,2019,108.96666666666667\n2019-10-28,130,1,301,2019,112.33333333333333\n2019-10-29,146,2,302,2019,115.53333333333333\n2019-10-30,164,3,303,2019,116.6\n2019-10-31,184,4,304,2019,118.0\n2019-11-01,187,5,305,2019,119.63333333333334\n2019-11-02,102,6,306,2019,118.73333333333333\n2019-11-03,110,7,307,2019,118.73333333333333\n2019-11-04,121,1,308,2019,120.3\n2019-11-05,161,2,309,2019,123.33333333333333\n2019-11-06,123,3,310,2019,123.5\n2019-11-07,119,4,311,2019,123.66666666666667\n2019-11-08,131,5,312,2019,123.93333333333334\n2019-11-09,71,6,313,2019,121.73333333333333\n2019-11-10,85,7,314,2019,119.53333333333333\n2019-11-11,113,1,315,2019,120.96666666666667\n2019-11-12,94,2,316,2019,120.66666666666667\n2019-11-13,124,3,317,2019,121.13333333333334\n2019-11-14,100,4,318,2019,118.8\n2019-11-15,135,5,319,2019,118.96666666666667\n2019-11-16,97,6,320,2019,118.13333333333334\n2019-11-17,110,7,321,2019,117.73333333333333\n2019-11-18,111,1,322,2019,118.6\n2019-11-19,123,2,323,2019,121.0\n2019-11-20,127,3,324,2019,121.03333333333333\n2019-11-21,115,4,325,2019,121.73333333333333\n2019-11-22,92,5,326,2019,121.5\n2019-11-23,61,6,327,2019,118.4\n2019-11-24,90,7,328,2019,117.43333333333334\n2019-11-25,114,1,329,2019,117.73333333333333\n2019-11-26,106,2,330,2019,118.2\n2019-11-27,144,3,331,2019,118.66666666666667\n2019-11-28,114,4,332,2019,117.6\n2019-11-29,115,5,333,2019,115.96666666666667\n2019-11-30,116,6,334,2019,113.7\n2019-12-01,103,7,335,2019,110.9\n2019-12-02,141,1,336,2019,112.2\n2019-12-03,116,2,337,2019,112.4\n2019-12-04,122,3,338,2019,112.43333333333334\n2019-12-05,157,4,339,2019,112.3\n2019-12-06,158,5,340,2019,113.46666666666667\n2019-12-07,95,6,341,2019,112.66666666666667\n2019-12-08,110,7,342,2019,111.96666666666667\n2019-12-09,113,1,343,2019,113.36666666666666\n2019-12-10,149,2,344,2019,115.5\n2019-12-11,142,3,345,2019,116.46666666666667\n2019-12-12,187,4,346,2019,119.56666666666666\n2019-12-13,145,5,347,2019,120.26666666666667\n2019-12-14,113,6,348,2019,120.7\n2019-12-15,89,7,349,2019,119.16666666666667\n2019-12-16,125,1,350,2019,120.1\n2019-12-17,201,2,351,2019,123.13333333333334\n2019-12-18,176,3,352,2019,125.3\n2019-12-19,206,4,353,2019,128.06666666666666\n2019-12-20,180,5,354,2019,129.83333333333334\n2019-12-21,90,6,355,2019,129.0\n2019-12-22,79,7,356,2019,128.56666666666666\n2019-12-23,49,1,357,2019,128.16666666666666\n2019-12-24,110,2,358,2019,128.83333333333334\n2019-12-25,76,3,359,2019,127.56666666666666\n2019-12-26,78,4,360,2019,126.63333333333334\n2019-12-27,124,5,361,2019,125.96666666666667\n2019-12-28,117,6,362,2019,126.06666666666666\n2019-12-29,60,7,363,2019,124.23333333333333\n2019-12-30,156,1,364,2019,125.56666666666666\n2019-12-31,126,2,365,2019,126.33333333333333\n2020-01-01,99,3,1,2020,124.93333333333334\n2020-01-02,142,4,2,2020,125.8\n2020-01-03,95,5,3,2020,124.9\n2020-01-04,120,6,4,2020,123.66666666666667\n2020-01-05,81,7,5,2020,121.1\n2020-01-06,125,1,6,2020,122.1\n2020-01-07,161,2,7,2020,123.8\n2020-01-08,169,3,8,2020,125.66666666666667\n2020-01-09,192,4,9,2020,127.1\n2020-01-10,251,5,10,2020,130.73333333333332\n2020-01-11,164,6,11,2020,129.96666666666667\n2020-01-12,219,7,12,2020,132.43333333333334\n2020-01-13,224,1,13,2020,136.13333333333333\n2020-01-14,193,2,14,2020,139.6\n2020-01-15,248,3,15,2020,143.7\n2020-01-16,255,4,16,2020,145.5\n2020-01-17,247,5,17,2020,147.86666666666667\n2020-01-18,117,6,18,2020,144.9\n2020-01-19,187,7,19,2020,145.13333333333333\n2020-01-20,138,1,20,2020,146.73333333333332\n2020-01-21,139,2,21,2020,148.73333333333332\n2020-01-22,199,3,22,2020,153.73333333333332\n2020-01-23,185,4,23,2020,156.23333333333332\n2020-01-24,143,5,24,2020,158.46666666666667\n2020-01-25,64,6,25,2020,158.0\n2020-01-26,89,7,26,2020,156.83333333333334\n2020-01-27,135,1,27,2020,157.43333333333334\n2020-01-28,101,2,28,2020,158.8\n2020-01-29,114,3,29,2020,157.4\n2020-01-30,155,4,30,2020,158.36666666666667\n2020-01-31,165,5,31,2020,160.56666666666666\n2020-02-01,89,6,32,2020,158.8\n2020-02-02,111,7,33,2020,159.33333333333334\n2020-02-03,141,1,34,2020,160.03333333333333\n2020-02-04,120,2,35,2020,161.33333333333334\n2020-02-05,145,3,36,2020,162.0\n2020-02-06,133,4,37,2020,161.06666666666666\n2020-02-07,149,5,38,2020,160.4\n2020-02-08,137,6,39,2020,158.56666666666666\n2020-02-09,119,7,40,2020,154.16666666666666\n2020-02-10,152,1,41,2020,153.76666666666668\n2020-02-11,162,2,42,2020,151.86666666666667\n2020-02-12,134,3,43,2020,148.86666666666667\n2020-02-13,170,4,44,2020,148.1\n2020-02-14,149,5,45,2020,144.8\n2020-02-15,95,6,46,2020,139.46666666666667\n2020-02-16,66,7,47,2020,133.43333333333334\n2020-02-17,129,1,48,2020,133.83333333333334\n2020-02-18,175,2,49,2020,133.43333333333334\n2020-02-19,133,3,50,2020,133.26666666666668\n2020-02-20,128,4,51,2020,132.9\n2020-02-21,133,5,52,2020,130.7\n2020-02-22,99,6,53,2020,127.83333333333333\n2020-02-23,165,7,54,2020,128.56666666666666\n2020-02-24,153,1,55,2020,131.53333333333333\n2020-02-25,146,2,56,2020,133.43333333333334\n2020-02-26,155,3,57,2020,134.1\n2020-02-27,144,4,58,2020,135.53333333333333\n2020-02-28,115,5,59,2020,135.56666666666666\n2020-02-29,54,6,60,2020,132.2\n2020-03-01,56,7,61,2020,128.56666666666666\n2020-03-02,166,1,62,2020,131.13333333333333\n2020-03-03,142,2,63,2020,132.16666666666666\n2020-03-04,120,3,64,2020,131.46666666666667\n2020-03-05,148,4,65,2020,132.4\n2020-03-06,119,5,66,2020,131.53333333333333\n2020-03-07,92,6,67,2020,130.16666666666666\n2020-03-08,143,7,68,2020,129.96666666666667\n2020-03-09,112,1,69,2020,129.13333333333333\n2020-03-10,111,2,70,2020,128.86666666666667\n2020-03-11,110,3,71,2020,127.46666666666667\n2020-03-12,82,4,72,2020,124.8\n2020-03-13,76,5,73,2020,122.86666666666666\n2020-03-14,74,6,74,2020,119.66666666666667\n2020-03-15,80,7,75,2020,117.36666666666666\n2020-03-16,91,1,76,2020,117.23333333333333\n2020-03-17,106,2,77,2020,118.56666666666666\n2020-03-18,132,3,78,2020,118.66666666666667\n2020-03-19,92,4,79,2020,115.9\n2020-03-20,115,5,80,2020,115.3\n2020-03-21,89,6,81,2020,114.0\n2020-03-22,105,7,82,2020,113.06666666666666\n2020-03-23,138,1,83,2020,114.36666666666666\n2020-03-24,144,2,84,2020,113.66666666666667\n2020-03-25,116,3,85,2020,112.43333333333334\n2020-03-26,99,4,86,2020,110.86666666666666\n2020-03-27,112,5,87,2020,109.43333333333334\n2020-03-28,123,6,88,2020,108.73333333333333\n2020-03-29,76,7,89,2020,107.43333333333334\n2020-03-30,107,1,90,2020,109.2\n2020-03-31,177,2,91,2020,113.23333333333333\n2020-04-01,158,3,92,2020,112.96666666666667\n2020-04-02,160,4,93,2020,113.56666666666666\n2020-04-03,145,5,94,2020,114.4\n2020-04-04,130,6,95,2020,113.8\n2020-04-05,95,7,96,2020,113.0\n2020-04-06,152,1,97,2020,115.0\n2020-04-07,157,2,98,2020,115.46666666666667\n2020-04-08,217,3,99,2020,118.96666666666667\n2020-04-09,160,4,100,2020,120.6\n2020-04-10,151,5,101,2020,121.96666666666667\n2020-04-11,127,6,102,2020,123.46666666666667\n2020-04-12,106,7,103,2020,124.46666666666667\n2020-04-13,208,1,104,2020,128.93333333333334\n2020-04-14,175,2,105,2020,132.1\n2020-04-15,146,3,106,2020,133.93333333333334\n2020-04-16,149,4,107,2020,135.36666666666667\n2020-04-17,110,5,108,2020,134.63333333333333\n2020-04-18,148,6,109,2020,136.5\n2020-04-19,151,7,110,2020,137.7\n2020-04-20,155,1,111,2020,139.9\n2020-04-21,141,2,112,2020,141.1\n2020-04-22,147,3,113,2020,141.4\n2020-04-23,146,4,114,2020,141.46666666666667\n2020-04-24,136,5,115,2020,142.13333333333333\n2020-04-25,98,6,116,2020,142.1\n2020-04-26,139,7,117,2020,143.0\n2020-04-27,180,1,118,2020,144.9\n2020-04-28,118,2,119,2020,146.3\n2020-04-29,166,3,120,2020,148.26666666666668\n2020-04-30,220,4,121,2020,149.7\n2020-05-01,237,5,122,2020,152.33333333333334\n2020-05-02,133,6,123,2020,151.43333333333334\n2020-05-03,142,7,124,2020,151.33333333333334\n2020-05-04,132,1,125,2020,151.4\n2020-05-05,218,2,126,2020,155.5\n2020-05-06,229,3,127,2020,158.06666666666666\n2020-05-07,184,4,128,2020,158.96666666666667\n2020-05-08,160,5,129,2020,157.06666666666666\n2020-05-09,152,6,130,2020,156.8\n2020-05-10,160,7,131,2020,157.1\n2020-05-11,187,1,132,2020,159.1\n2020-05-12,158,2,133,2020,160.83333333333334\n2020-05-13,200,3,134,2020,160.56666666666666\n2020-05-14,212,4,135,2020,161.8\n2020-05-15,199,5,136,2020,163.56666666666666\n2020-05-16,113,6,137,2020,162.36666666666667\n2020-05-17,142,7,138,2020,163.43333333333334\n2020-05-18,161,1,139,2020,163.86666666666667\n2020-05-19,162,2,140,2020,164.23333333333332\n2020-05-20,186,3,141,2020,165.26666666666668\n2020-05-21,122,4,142,2020,164.63333333333333\n2020-05-22,142,5,143,2020,164.46666666666667\n2020-05-23,128,6,144,2020,163.86666666666667\n2020-05-24,134,7,145,2020,163.8\n2020-05-25,143,1,146,2020,165.3\n2020-05-26,220,2,147,2020,168.0\n2020-05-27,144,3,148,2020,166.8\n2020-05-28,100,4,149,2020,166.2\n2020-05-29,108,5,150,2020,164.26666666666668\n2020-05-30,213,6,151,2020,164.03333333333333\n2020-05-31,86,7,152,2020,159.0\n2020-06-01,115,1,153,2020,158.4\n2020-06-02,107,2,154,2020,157.23333333333332\n2020-06-03,169,3,155,2020,158.46666666666667\n2020-06-04,147,4,156,2020,156.1\n2020-06-05,115,5,157,2020,152.3\n2020-06-06,107,6,158,2020,149.73333333333332\n2020-06-07,93,7,159,2020,147.5\n2020-06-08,116,1,160,2020,146.3\n2020-06-09,143,2,161,2020,145.73333333333332\n2020-06-10,84,3,162,2020,142.3\n2020-06-11,177,4,163,2020,142.93333333333334\n2020-06-12,125,5,164,2020,140.43333333333334\n2020-06-13,61,6,165,2020,135.4\n2020-06-14,98,7,166,2020,132.03333333333333\n2020-06-15,191,1,167,2020,134.63333333333333\n2020-06-16,157,2,168,2020,135.13333333333333\n2020-06-17,199,3,169,2020,136.4\n2020-06-18,192,4,170,2020,137.4\n2020-06-19,146,5,171,2020,136.06666666666666\n2020-06-20,60,6,172,2020,134.0\n2020-06-21,116,7,173,2020,133.13333333333333\n2020-06-22,128,1,174,2020,133.13333333333333\n2020-06-23,130,2,175,2020,133.0\n2020-06-24,160,3,176,2020,133.56666666666666\n2020-06-25,120,4,177,2020,130.23333333333332\n2020-06-26,167,5,178,2020,131.0\n2020-06-27,164,6,179,2020,133.13333333333333\n2020-06-28,113,7,180,2020,133.3\n2020-06-29,167,1,181,2020,131.76666666666668\n2020-06-30,155,2,182,2020,134.06666666666666\n2020-07-01,126,3,183,2020,134.43333333333334\n2020-07-02,134,4,184,2020,135.33333333333334\n2020-07-03,146,5,185,2020,134.56666666666666\n2020-07-04,59,6,186,2020,131.63333333333333\n2020-07-05,102,7,187,2020,131.2\n2020-07-06,120,1,188,2020,131.63333333333333\n2020-07-07,124,2,189,2020,132.66666666666666\n2020-07-08,154,3,190,2020,133.93333333333334\n2020-07-09,164,4,191,2020,134.63333333333333\n2020-07-10,171,5,192,2020,137.53333333333333\n2020-07-11,79,6,193,2020,134.26666666666668\n2020-07-12,63,7,194,2020,132.2\n2020-07-13,85,1,195,2020,133.0\n2020-07-14,90,2,196,2020,132.73333333333332\n2020-07-15,129,3,197,2020,130.66666666666666\n2020-07-16,138,4,198,2020,130.03333333333333\n2020-07-17,95,5,199,2020,126.56666666666666\n2020-07-18,71,6,200,2020,122.53333333333333\n2020-07-19,99,7,201,2020,120.96666666666667\n2020-07-20,140,1,202,2020,123.63333333333334\n2020-07-21,189,2,203,2020,126.06666666666666\n2020-07-22,212,3,204,2020,128.86666666666667\n2020-07-23,176,4,205,2020,130.4\n2020-07-24,183,5,206,2020,131.16666666666666\n2020-07-25,55,6,207,2020,129.0\n2020-07-26,100,7,208,2020,126.76666666666667\n2020-07-27,137,1,209,2020,125.86666666666666\n2020-07-28,190,2,210,2020,128.43333333333334\n2020-07-29,118,3,211,2020,126.8\n2020-07-30,147,4,212,2020,126.53333333333333\n2020-07-31,139,5,213,2020,126.96666666666667\n2020-08-01,91,6,214,2020,125.53333333333333\n2020-08-02,99,7,215,2020,123.96666666666667\n2020-08-03,168,1,216,2020,127.6\n2020-08-04,168,2,217,2020,129.8\n2020-08-05,151,3,218,2020,130.83333333333334\n2020-08-06,167,4,219,2020,132.26666666666668\n2020-08-07,155,5,220,2020,132.3\n2020-08-08,78,6,221,2020,129.43333333333334\n2020-08-09,111,7,222,2020,127.43333333333334\n2020-08-10,126,1,223,2020,129.0\n2020-08-11,111,2,224,2020,130.6\n2020-08-12,121,3,225,2020,131.8\n2020-08-13,97,4,226,2020,132.03333333333333\n2020-08-14,137,5,227,2020,132.3\n2020-08-15,54,6,228,2020,129.5\n2020-08-16,56,7,229,2020,128.2\n2020-08-17,162,1,230,2020,131.23333333333332\n2020-08-18,98,2,231,2020,131.2\n2020-08-19,139,3,232,2020,131.16666666666666\n2020-08-20,128,4,233,2020,129.13333333333333\n2020-08-21,133,5,234,2020,126.5\n2020-08-22,93,6,235,2020,123.73333333333333\n2020-08-23,68,7,236,2020,119.9\n2020-08-24,149,1,237,2020,123.03333333333333\n2020-08-25,165,2,238,2020,125.2\n2020-08-26,201,3,239,2020,127.33333333333333\n2020-08-27,135,4,240,2020,125.5\n2020-08-28,108,5,241,2020,125.16666666666667\n2020-08-29,84,6,242,2020,123.06666666666666\n2020-08-30,42,7,243,2020,119.83333333333333\n2020-08-31,100,1,244,2020,120.13333333333334\n2020-09-01,134,2,245,2020,121.3\n2020-09-02,120,3,246,2020,119.7\n2020-09-03,100,4,247,2020,117.43333333333334\n2020-09-04,110,5,248,2020,116.06666666666666\n2020-09-05,53,6,249,2020,112.26666666666667\n2020-09-06,62,7,250,2020,109.16666666666667\n2020-09-07,129,1,251,2020,110.86666666666666\n2020-09-08,153,2,252,2020,112.26666666666667\n2020-09-09,148,3,253,2020,113.0\n2020-09-10,141,4,254,2020,114.0\n2020-09-11,162,5,255,2020,115.36666666666666\n2020-09-12,78,6,256,2020,114.73333333333333\n2020-09-13,58,7,257,2020,112.1\n2020-09-14,133,1,258,2020,114.73333333333333\n2020-09-15,126,2,259,2020,117.06666666666666\n2020-09-16,134,3,260,2020,116.13333333333334\n2020-09-17,94,4,261,2020,116.0\n2020-09-18,102,5,262,2020,114.76666666666667\n2020-09-19,65,6,263,2020,112.66666666666667\n2020-09-20,86,7,264,2020,111.1\n2020-09-21,136,1,265,2020,112.53333333333333\n2020-09-22,144,2,266,2020,115.06666666666666\n2020-09-23,137,3,267,2020,114.66666666666667\n2020-09-24,146,4,268,2020,114.03333333333333\n2020-09-25,148,5,269,2020,112.26666666666667\n2020-09-26,78,6,270,2020,110.36666666666666\n2020-09-27,82,7,271,2020,109.5\n2020-09-28,158,1,272,2020,111.96666666666667\n2020-09-29,122,2,273,2020,114.63333333333334\n2020-09-30,108,3,274,2020,114.9\n2020-10-01,146,4,275,2020,115.3\n2020-10-02,125,5,276,2020,115.46666666666667\n2020-10-03,55,6,277,2020,113.96666666666667\n2020-10-04,62,7,278,2020,112.36666666666666\n2020-10-05,99,1,279,2020,113.9\n2020-10-06,95,2,280,2020,115.0\n2020-10-07,137,3,281,2020,115.26666666666667\n2020-10-08,125,4,282,2020,114.33333333333333\n2020-10-09,117,5,283,2020,113.3\n2020-10-10,67,6,284,2020,110.83333333333333\n2020-10-11,62,7,285,2020,107.5\n2020-10-12,133,1,286,2020,109.33333333333333\n2020-10-13,80,2,287,2020,110.06666666666666\n2020-10-14,115,3,288,2020,109.46666666666667\n2020-10-15,104,4,289,2020,108.73333333333333\n2020-10-16,122,5,290,2020,108.33333333333333\n2020-10-17,106,6,291,2020,108.73333333333333\n2020-10-18,86,7,292,2020,108.2\n2020-10-19,156,1,293,2020,111.23333333333333\n2020-10-20,156,2,294,2020,113.56666666666666\n2020-10-21,162,3,295,2020,114.43333333333334\n2020-10-22,180,4,296,2020,115.63333333333334\n2020-10-23,93,5,297,2020,114.16666666666667\n2020-10-24,81,6,298,2020,112.0\n2020-10-25,75,7,299,2020,109.56666666666666\n2020-10-26,125,1,300,2020,111.13333333333334\n2020-10-27,130,2,301,2020,112.73333333333333\n2020-10-28,147,3,302,2020,112.36666666666666\n2020-10-29,118,4,303,2020,112.23333333333333\n2020-10-30,99,5,304,2020,111.93333333333334\n2020-10-31,93,6,305,2020,110.16666666666667\n2020-11-01,103,7,306,2020,109.43333333333334\n2020-11-02,103,1,307,2020,111.03333333333333\n2020-11-03,121,2,308,2020,113.0\n2020-11-04,86,3,309,2020,112.56666666666666\n2020-11-05,153,4,310,2020,114.5\n2020-11-06,129,5,311,2020,114.23333333333333\n2020-11-07,116,6,312,2020,113.93333333333334\n2020-11-08,51,7,313,2020,111.73333333333333\n2020-11-09,129,1,314,2020,113.8\n2020-11-10,150,2,315,2020,116.73333333333333\n2020-11-11,112,3,316,2020,116.03333333333333\n2020-11-12,117,4,317,2020,117.26666666666667\n2020-11-13,161,5,318,2020,118.8\n2020-11-14,69,6,319,2020,117.63333333333334\n2020-11-15,95,7,320,2020,116.73333333333333\n2020-11-16,144,1,321,2020,118.0\n2020-11-17,155,2,322,2020,120.3\n2020-11-18,142,3,323,2020,119.83333333333333\n2020-11-19,119,4,324,2020,118.6\n2020-11-20,113,5,325,2020,116.96666666666667\n2020-11-21,77,6,326,2020,113.53333333333333\n2020-11-22,69,7,327,2020,112.73333333333333\n2020-11-23,97,1,328,2020,113.26666666666667\n2020-11-24,107,2,329,2020,114.33333333333333\n2020-11-25,122,3,330,2020,114.23333333333333\n2020-11-26,131,4,331,2020,114.26666666666667\n2020-11-27,120,5,332,2020,113.36666666666666\n2020-11-28,93,6,333,2020,112.53333333333333\n2020-11-29,116,7,334,2020,113.1\n2020-11-30,139,1,335,2020,114.63333333333334\n2020-12-01,89,2,336,2020,114.16666666666667\n2020-12-02,152,3,337,2020,115.8\n2020-12-03,130,4,338,2020,116.1\n2020-12-04,128,5,339,2020,117.5\n2020-12-05,112,6,340,2020,116.13333333333334\n2020-12-06,69,7,341,2020,114.13333333333334\n2020-12-07,153,1,342,2020,115.36666666666666\n2020-12-08,110,2,343,2020,117.33333333333333\n2020-12-09,122,3,344,2020,117.1\n2020-12-10,124,4,345,2020,116.23333333333333\n2020-12-11,107,5,346,2020,116.06666666666666\n2020-12-12,78,6,347,2020,114.76666666666667\n2020-12-13,90,7,348,2020,112.4\n2020-12-14,87,1,349,2020,113.0\n2020-12-15,125,2,350,2020,114.0\n2020-12-16,143,3,351,2020,113.96666666666667\n2020-12-17,114,4,352,2020,112.6\n2020-12-18,96,5,353,2020,111.06666666666666\n2020-12-19,55,6,354,2020,108.93333333333334\n2020-12-20,66,7,355,2020,107.36666666666666\n2020-12-21,150,1,356,2020,109.8\n2020-12-22,129,2,357,2020,111.8\n2020-12-23,134,3,358,2020,113.03333333333333\n2020-12-24,70,4,359,2020,111.8\n2020-12-25,38,5,360,2020,109.0\n2020-12-26,68,6,361,2020,106.9\n2020-12-27,93,7,362,2020,106.0\n2020-12-28,116,1,363,2020,106.76666666666667\n2020-12-29,151,2,364,2020,107.93333333333334\n2020-12-30,114,3,365,2020,107.1\n2020-12-31,53,4,366,2020,105.9\n2021-01-01,67,5,1,2021,103.06666666666666\n2021-01-02,70,6,2,2021,101.06666666666666\n2021-01-03,60,7,3,2021,98.8\n2021-01-04,113,1,4,2021,98.83333333333333\n2021-01-05,108,2,5,2021,100.13333333333334\n2021-01-06,106,3,6,2021,98.56666666666666\n2021-01-07,133,4,7,2021,99.33333333333333\n2021-01-08,133,5,8,2021,99.7\n2021-01-09,99,6,9,2021,98.86666666666666\n2021-01-10,114,7,10,2021,99.1\n2021-01-11,115,1,11,2021,100.33333333333333\n2021-01-12,119,2,12,2021,101.3\n2021-01-13,167,3,13,2021,103.96666666666667\n2021-01-14,178,4,14,2021,105.73333333333333\n2021-01-15,184,5,15,2021,107.1\n2021-01-16,116,6,16,2021,107.16666666666667\n2021-01-17,130,7,17,2021,108.3\n2021-01-18,115,1,18,2021,110.3\n2021-01-19,155,2,19,2021,113.26666666666667\n2021-01-20,204,3,20,2021,115.06666666666666\n2021-01-21,137,4,21,2021,115.33333333333333\n2021-01-22,141,5,22,2021,115.56666666666666\n2021-01-23,130,6,23,2021,117.56666666666666\n2021-01-24,67,7,24,2021,118.53333333333333\n2021-01-25,132,1,25,2021,120.66666666666667\n2021-01-26,180,2,26,2021,123.56666666666666\n2021-01-27,161,3,27,2021,125.06666666666666\n2021-01-28,162,4,28,2021,125.43333333333334\n2021-01-29,207,5,29,2021,128.53333333333333\n2021-01-30,95,6,30,2021,129.93333333333334\n2021-01-31,149,7,31,2021,132.66666666666666\n2021-02-01,176,1,32,2021,136.2\n2021-02-02,157,2,33,2021,139.43333333333334\n2021-02-03,191,3,34,2021,142.03333333333333\n2021-02-04,156,4,35,2021,143.63333333333333\n2021-02-05,214,5,36,2021,147.23333333333332\n2021-02-06,159,6,37,2021,148.1\n2021-02-07,94,7,38,2021,146.8\n2021-02-08,121,1,39,2021,147.53333333333333\n2021-02-09,134,2,40,2021,148.2\n2021-02-10,126,3,41,2021,148.56666666666666\n2021-02-11,138,4,42,2021,149.2\n2021-02-12,115,5,43,2021,147.46666666666667\n2021-02-13,121,6,44,2021,145.56666666666666\n2021-02-14,115,7,45,2021,143.26666666666668\n2021-02-15,148,1,46,2021,144.33333333333334\n2021-02-16,127,2,47,2021,144.23333333333332\n2021-02-17,149,3,48,2021,145.36666666666667\n2021-02-18,153,4,49,2021,145.3\n2021-02-19,130,5,50,2021,142.83333333333334\n2021-02-20,120,6,51,2021,142.26666666666668\n2021-02-21,98,7,52,2021,140.83333333333334\n2021-02-22,160,1,53,2021,141.83333333333334\n2021-02-23,150,2,54,2021,144.6\n2021-02-24,185,3,55,2021,146.36666666666667\n2021-02-25,171,4,56,2021,146.06666666666666\n2021-02-26,162,5,57,2021,146.1\n2021-02-27,130,6,58,2021,145.03333333333333\n2021-02-28,189,7,59,2021,144.43333333333334\n2021-03-01,218,1,60,2021,148.53333333333333\n2021-03-02,138,2,61,2021,148.16666666666666\n2021-03-03,172,3,62,2021,148.03333333333333\n2021-03-04,202,4,63,2021,149.53333333333333\n2021-03-05,152,5,64,2021,148.23333333333332\n2021-03-06,128,6,65,2021,147.3\n2021-03-07,115,7,66,2021,144.0\n2021-03-08,195,1,67,2021,145.2\n2021-03-09,161,2,68,2021,147.43333333333334\n2021-03-10,190,3,69,2021,149.73333333333332\n2021-03-11,172,4,70,2021,151.0\n2021-03-12,149,5,71,2021,151.76666666666668\n2021-03-13,141,6,72,2021,151.86666666666667\n2021-03-14,77,7,73,2021,150.6\n2021-03-15,167,1,74,2021,152.13333333333333\n2021-03-16,161,2,75,2021,153.66666666666666\n2021-03-17,115,3,76,2021,152.56666666666666\n2021-03-18,144,4,77,2021,153.13333333333333\n2021-03-19,145,5,78,2021,153.0\n2021-03-20,88,6,79,2021,150.83333333333334\n2021-03-21,150,7,80,2021,151.5\n2021-03-22,110,1,81,2021,151.16666666666666\n2021-03-23,157,2,82,2021,153.13333333333333\n2021-03-24,145,3,83,2021,152.63333333333333\n2021-03-25,116,4,84,2021,151.5\n2021-03-26,126,5,85,2021,149.53333333333333\n2021-03-27,153,6,86,2021,148.93333333333334\n2021-03-28,111,7,87,2021,147.23333333333332\n2021-03-29,178,1,88,2021,148.83333333333334\n2021-03-30,178,2,89,2021,148.46666666666667\n2021-03-31,170,3,90,2021,146.86666666666667\n2021-04-01,170,4,91,2021,147.93333333333334\n2021-04-02,152,5,92,2021,147.26666666666668\n2021-04-03,100,6,93,2021,143.86666666666667\n2021-04-04,119,7,94,2021,142.76666666666668\n2021-04-05,129,1,95,2021,142.8\n2021-04-06,200,2,96,2021,145.63333333333333\n2021-04-07,164,3,97,2021,144.6\n2021-04-08,120,4,98,2021,143.23333333333332\n2021-04-09,178,5,99,2021,142.83333333333334\n2021-04-10,136,6,100,2021,141.63333333333333\n2021-04-11,106,7,101,2021,140.2\n2021-04-12,155,1,102,2021,140.66666666666666\n2021-04-13,114,2,103,2021,141.9\n2021-04-14,138,3,104,2021,140.93333333333334\n2021-04-15,175,4,105,2021,141.4\n2021-04-16,130,5,106,2021,141.9\n2021-04-17,100,6,107,2021,140.43333333333334\n2021-04-18,80,7,108,2021,138.26666666666668\n2021-04-19,137,1,109,2021,139.9\n2021-04-20,107,2,110,2021,138.46666666666667\n2021-04-21,174,3,111,2021,140.6\n2021-04-22,136,4,112,2021,139.9\n2021-04-23,141,5,113,2021,139.76666666666668\n2021-04-24,94,6,114,2021,139.03333333333333\n2021-04-25,124,7,115,2021,138.96666666666667\n2021-04-26,190,1,116,2021,140.2\n2021-04-27,96,2,117,2021,139.7\n2021-04-28,131,3,118,2021,138.13333333333333\n2021-04-29,135,4,119,2021,136.7\n2021-04-30,166,5,120,2021,136.56666666666666\n2021-05-01,80,6,121,2021,133.56666666666666\n2021-05-02,94,7,122,2021,131.63333333333333\n2021-05-03,71,1,123,2021,130.66666666666666\n2021-05-04,96,2,124,2021,129.9\n2021-05-05,117,3,125,2021,129.5\n2021-05-06,162,4,126,2021,128.23333333333332\n2021-05-07,133,5,127,2021,127.2\n2021-05-08,131,6,128,2021,127.56666666666666\n2021-05-09,94,7,129,2021,124.76666666666667\n2021-05-10,151,1,130,2021,125.26666666666667\n2021-05-11,159,2,131,2021,127.03333333333333\n2021-05-12,156,3,132,2021,127.06666666666666\n2021-05-13,172,4,133,2021,129.0\n2021-05-14,181,5,134,2021,130.43333333333334\n2021-05-15,75,6,135,2021,127.1\n2021-05-16,94,7,136,2021,125.9\n2021-05-17,144,1,137,2021,127.36666666666666\n2021-05-18,117,2,138,2021,128.6\n2021-05-19,157,3,139,2021,129.26666666666668\n2021-05-20,195,4,140,2021,132.2\n2021-05-21,160,5,141,2021,131.73333333333332\n2021-05-22,104,6,142,2021,130.66666666666666\n2021-05-23,75,7,143,2021,128.46666666666667\n2021-05-24,141,1,144,2021,130.03333333333333\n2021-05-25,177,2,145,2021,131.8\n2021-05-26,186,3,146,2021,131.66666666666666\n2021-05-27,141,4,147,2021,133.16666666666666\n2021-05-28,146,5,148,2021,133.66666666666666\n2021-05-29,66,6,149,2021,131.36666666666667\n2021-05-30,83,7,150,2021,128.6\n2021-05-31,120,1,151,2021,129.93333333333334\n2021-06-01,151,2,152,2021,131.83333333333334\n2021-06-02,132,3,153,2021,133.86666666666667\n2021-06-03,171,4,154,2021,136.36666666666667\n2021-06-04,98,5,155,2021,135.73333333333332\n2021-06-05,63,6,156,2021,132.43333333333334\n2021-06-06,64,7,157,2021,130.13333333333333\n2021-06-07,133,1,158,2021,130.2\n2021-06-08,117,2,159,2021,130.96666666666667\n2021-06-09,181,3,160,2021,131.96666666666667\n2021-06-10,166,4,161,2021,132.2\n2021-06-11,100,5,162,2021,130.33333333333334\n2021-06-12,56,6,163,2021,126.46666666666667\n2021-06-13,54,7,164,2021,122.23333333333333\n2021-06-14,89,1,165,2021,122.7\n2021-06-15,95,2,166,2021,122.73333333333333\n2021-06-16,134,3,167,2021,122.4\n2021-06-17,149,4,168,2021,123.46666666666667\n2021-06-18,116,5,169,2021,122.1\n2021-06-19,48,6,170,2021,117.2\n2021-06-20,60,7,171,2021,113.86666666666666\n2021-06-21,115,1,172,2021,114.23333333333333\n2021-06-22,122,2,173,2021,115.8\n2021-06-23,136,3,174,2021,115.63333333333334\n2021-06-24,116,4,175,2021,113.6\n2021-06-25,110,5,176,2021,111.06666666666666\n2021-06-26,68,6,177,2021,108.63333333333334\n2021-06-27,104,7,178,2021,107.23333333333333\n2021-06-28,118,1,179,2021,108.96666666666667\n2021-06-29,110,2,180,2021,109.86666666666666\n2021-06-30,129,3,181,2021,110.16666666666667\n2021-07-01,154,4,182,2021,110.26666666666667\n2021-07-02,91,5,183,2021,108.9\n2021-07-03,49,6,184,2021,104.83333333333333\n2021-07-04,81,7,185,2021,104.26666666666667\n2021-07-05,139,1,186,2021,106.8\n2021-07-06,141,2,187,2021,109.36666666666666\n2021-07-07,113,3,188,2021,108.7\n2021-07-08,96,4,189,2021,108.0\n2021-07-09,118,5,190,2021,105.9\n2021-07-10,111,6,191,2021,104.06666666666666\n2021-07-11,104,7,192,2021,104.2\n2021-07-12,131,1,193,2021,106.7\n2021-07-13,126,2,194,2021,109.1\n2021-07-14,174,3,195,2021,111.93333333333334\n2021-07-15,140,4,196,2021,113.43333333333334\n2021-07-16,102,5,197,2021,112.36666666666666\n2021-07-17,68,6,198,2021,109.66666666666667\n2021-07-18,55,7,199,2021,107.63333333333334\n2021-07-19,87,1,200,2021,108.93333333333334\n2021-07-20,145,2,201,2021,111.76666666666667\n2021-07-21,125,3,202,2021,112.1\n2021-07-22,118,4,203,2021,111.96666666666667\n2021-07-23,107,5,204,2021,111.0\n2021-07-24,37,6,205,2021,108.36666666666666\n2021-07-25,45,7,206,2021,106.2\n2021-07-26,80,1,207,2021,106.6\n2021-07-27,152,2,208,2021,108.2\n2021-07-28,145,3,209,2021,109.1\n2021-07-29,148,4,210,2021,110.36666666666666\n2021-07-30,118,5,211,2021,110.0\n2021-07-31,98,6,212,2021,108.13333333333334\n2021-08-01,81,7,213,2021,107.8\n2021-08-02,137,1,214,2021,110.73333333333333\n2021-08-03,129,2,215,2021,112.33333333333333\n2021-08-04,129,3,216,2021,112.0\n2021-08-05,103,4,217,2021,110.73333333333333\n2021-08-06,143,5,218,2021,111.73333333333333\n2021-08-07,112,6,219,2021,112.26666666666667\n2021-08-08,94,7,220,2021,111.46666666666667\n2021-08-09,102,1,221,2021,111.16666666666667\n2021-08-10,106,2,222,2021,111.23333333333333\n2021-08-11,140,3,223,2021,111.53333333333333\n2021-08-12,113,4,224,2021,111.1\n2021-08-13,113,5,225,2021,109.06666666666666\n2021-08-14,109,6,226,2021,108.03333333333333\n2021-08-15,71,7,227,2021,107.0\n2021-08-16,124,1,228,2021,108.86666666666666\n2021-08-17,95,2,229,2021,110.2\n2021-08-18,148,3,230,2021,112.23333333333333\n2021-08-19,124,4,231,2021,111.53333333333333\n2021-08-20,93,5,232,2021,110.46666666666667\n2021-08-21,94,6,233,2021,109.66666666666667\n2021-08-22,80,7,234,2021,108.76666666666667\n2021-08-23,156,1,235,2021,112.73333333333333\n2021-08-24,135,2,236,2021,115.73333333333333\n2021-08-25,170,3,237,2021,118.73333333333333\n2021-08-26,125,4,238,2021,117.83333333333333\n2021-08-27,122,5,239,2021,117.06666666666666\n2021-08-28,89,6,240,2021,115.1\n2021-08-29,80,7,241,2021,113.83333333333333\n2021-08-30,125,1,242,2021,114.73333333333333\n2021-08-31,160,2,243,2021,117.36666666666666\n2021-09-01,158,3,244,2021,118.06666666666666\n2021-09-02,106,4,245,2021,117.3\n2021-09-03,125,5,246,2021,117.16666666666667\n2021-09-04,96,6,247,2021,116.93333333333334\n2021-09-05,89,7,248,2021,115.13333333333334\n2021-09-06,128,1,249,2021,115.66666666666667\n2021-09-07,124,2,250,2021,116.66666666666667\n2021-09-08,99,3,251,2021,116.56666666666666\n2021-09-09,145,4,252,2021,117.86666666666666\n2021-09-10,132,5,253,2021,117.6\n2021-09-11,80,6,254,2021,116.5\n2021-09-12,82,7,255,2021,115.46666666666667\n2021-09-13,148,1,256,2021,116.76666666666667\n2021-09-14,129,2,257,2021,118.7\n2021-09-15,146,3,258,2021,119.43333333333334\n2021-09-16,183,4,259,2021,122.36666666666666\n2021-09-17,157,5,260,2021,122.66666666666667\n2021-09-18,103,6,261,2021,121.96666666666667\n2021-09-19,133,7,262,2021,123.3\n2021-09-20,133,1,263,2021,124.6\n2021-09-21,146,2,264,2021,126.8\n2021-09-22,146,3,265,2021,126.46666666666667\n2021-09-23,145,4,266,2021,126.8\n2021-09-24,84,5,267,2021,123.93333333333334\n2021-09-25,108,6,268,2021,123.36666666666666\n2021-09-26,64,7,269,2021,121.43333333333334\n2021-09-27,122,1,270,2021,122.53333333333333\n2021-09-28,170,2,271,2021,125.53333333333333\n2021-09-29,143,3,272,2021,126.13333333333334\n2021-09-30,142,4,273,2021,125.53333333333333\n2021-10-01,104,5,274,2021,123.73333333333333\n2021-10-02,73,6,275,2021,122.63333333333334\n2021-10-03,123,7,276,2021,122.56666666666666\n2021-10-04,96,1,277,2021,122.56666666666666\n2021-10-05,141,2,278,2021,124.3\n2021-10-06,190,3,279,2021,126.36666666666666\n2021-10-07,152,4,280,2021,127.3\n2021-10-08,118,5,281,2021,127.93333333333334\n2021-10-09,158,6,282,2021,128.36666666666667\n2021-10-10,118,7,283,2021,127.9\n2021-10-11,137,1,284,2021,129.8\n2021-10-12,139,2,285,2021,131.7\n2021-10-13,110,3,286,2021,130.43333333333334\n2021-10-14,100,4,287,2021,129.46666666666667\n2021-10-15,114,5,288,2021,128.4\n2021-10-16,87,6,289,2021,125.2\n2021-10-17,51,7,290,2021,121.66666666666667\n2021-10-18,152,1,291,2021,123.3\n2021-10-19,135,2,292,2021,123.36666666666666\n2021-10-20,126,3,293,2021,123.13333333333334\n2021-10-21,156,4,294,2021,123.46666666666667\n2021-10-22,137,5,295,2021,123.16666666666667\n2021-10-23,68,6,296,2021,120.6\n2021-10-24,63,7,297,2021,119.9\n2021-10-25,126,1,298,2021,120.5\n2021-10-26,92,2,299,2021,121.43333333333334\n2021-10-27,130,3,300,2021,121.7\n2021-10-28,112,4,301,2021,119.76666666666667\n2021-10-29,135,5,302,2021,119.5\n2021-10-30,70,6,303,2021,117.1\n2021-10-31,86,7,304,2021,116.5\n2021-11-01,116,1,305,2021,117.93333333333334\n2021-11-02,129,2,306,2021,118.13333333333334\n2021-11-03,163,3,307,2021,120.36666666666666\n2021-11-04,188,4,308,2021,121.93333333333334\n2021-11-05,145,5,309,2021,120.43333333333334\n2021-11-06,116,6,310,2021,119.23333333333333\n2021-11-07,57,7,311,2021,117.2\n2021-11-08,143,1,312,2021,116.7\n2021-11-09,128,2,313,2021,117.03333333333333\n2021-11-10,178,3,314,2021,118.4\n2021-11-11,140,4,315,2021,118.43333333333334\n2021-11-12,98,5,316,2021,118.03333333333333\n2021-11-13,54,6,317,2021,116.5\n2021-11-14,63,7,318,2021,114.8\n2021-11-15,136,1,319,2021,116.43333333333334\n2021-11-16,146,2,320,2021,119.6\n2021-11-17,175,3,321,2021,120.36666666666666\n2021-11-18,133,4,322,2021,120.3\n2021-11-19,148,5,323,2021,121.03333333333333\n2021-11-20,38,6,324,2021,117.1\n2021-11-21,72,7,325,2021,114.93333333333334\n2021-11-22,114,1,326,2021,116.46666666666667\n2021-11-23,135,2,327,2021,118.86666666666666\n2021-11-24,120,3,328,2021,118.66666666666667\n2021-11-25,81,4,329,2021,118.3\n2021-11-26,96,5,330,2021,117.16666666666667\n2021-11-27,67,6,331,2021,115.66666666666667\n2021-11-28,72,7,332,2021,113.56666666666666\n2021-11-29,128,1,333,2021,115.5\n2021-11-30,101,2,334,2021,116.0\n2021-12-01,124,3,335,2021,116.26666666666667\n2021-12-02,100,4,336,2021,115.3\n2021-12-03,79,5,337,2021,112.5\n2021-12-04,49,6,338,2021,107.86666666666666\n2021-12-05,64,7,339,2021,105.16666666666667\n2021-12-06,122,1,340,2021,105.36666666666666\n2021-12-07,113,2,341,2021,107.23333333333333\n2021-12-08,104,3,342,2021,105.93333333333334\n2021-12-09,125,4,343,2021,105.83333333333333\n2021-12-10,122,5,344,2021,103.96666666666667\n2021-12-11,38,6,345,2021,100.56666666666666\n2021-12-12,94,7,346,2021,100.43333333333334\n2021-12-13,144,1,347,2021,103.43333333333334\n2021-12-14,107,2,348,2021,104.9\n2021-12-15,143,3,349,2021,105.13333333333334\n2021-12-16,145,4,350,2021,105.1\n2021-12-17,102,5,351,2021,102.66666666666667\n2021-12-18,62,6,352,2021,100.3\n2021-12-19,88,7,353,2021,98.3\n2021-12-20,120,1,354,2021,101.03333333333333\n2021-12-21,126,2,355,2021,102.83333333333333\n2021-12-22,163,3,356,2021,104.46666666666667\n2021-12-23,107,4,357,2021,103.53333333333333\n2021-12-24,48,5,358,2021,101.13333333333334\n2021-12-25,16,6,359,2021,98.96666666666667\n2021-12-26,29,7,360,2021,96.73333333333333\n2021-12-27,111,1,361,2021,98.2\n2021-12-28,95,2,362,2021,98.96666666666667\n2021-12-29,133,3,363,2021,99.13333333333334\n2021-12-30,88,4,364,2021,98.7\n2021-12-31,44,5,365,2021,96.03333333333333\n2022-01-01,33,6,1,2022,93.8\n2022-01-02,44,7,2,2022,92.63333333333334\n2022-01-03,85,1,3,2022,93.83333333333333\n2022-01-04,147,2,4,2022,96.6\n2022-01-05,120,3,5,2022,96.53333333333333\n2022-01-06,110,4,6,2022,96.43333333333334\n2022-01-07,127,5,7,2022,97.2\n2022-01-08,95,6,8,2022,96.2\n2022-01-09,85,7,9,2022,94.96666666666667\n2022-01-10,150,1,10,2022,98.7\n2022-01-11,118,2,11,2022,99.5\n2022-01-12,180,3,12,2022,100.7\n2022-01-13,153,4,13,2022,102.23333333333333\n2022-01-14,157,5,14,2022,102.7\n2022-01-15,74,6,15,2022,100.33333333333333\n2022-01-16,86,7,16,2022,99.8\n2022-01-17,151,1,17,2022,102.76666666666667\n2022-01-18,150,2,18,2022,104.83333333333333\n2022-01-19,182,3,19,2022,106.9\n2022-01-20,156,4,20,2022,107.9\n2022-01-21,156,5,21,2022,107.66666666666667\n2022-01-22,92,6,22,2022,107.16666666666667\n2022-01-23,115,7,23,2022,109.4\n2022-01-24,141,1,24,2022,113.56666666666666\n2022-01-25,125,2,25,2022,116.76666666666667\n2022-01-26,158,3,26,2022,118.33333333333333\n2022-01-27,109,4,27,2022,118.8\n2022-01-28,130,5,28,2022,118.7\n2022-01-29,103,6,29,2022,119.2\n2022-01-30,67,7,30,2022,119.96666666666667\n2022-01-31,152,1,31,2022,123.93333333333334\n2022-02-01,153,2,32,2022,127.56666666666666\n2022-02-02,160,3,33,2022,130.06666666666666\n2022-02-03,104,4,34,2022,128.63333333333333\n2022-02-04,124,5,35,2022,128.76666666666668\n2022-02-05,38,6,36,2022,126.36666666666666\n2022-02-06,70,7,37,2022,124.46666666666667\n2022-02-07,129,1,38,2022,125.6\n2022-02-08,133,2,39,2022,127.2\n2022-02-09,152,3,40,2022,127.26666666666667\n2022-02-10,144,4,41,2022,128.13333333333333\n2022-02-11,110,5,42,2022,125.8\n2022-02-12,48,6,43,2022,122.3\n2022-02-13,43,7,44,2022,118.5\n2022-02-14,108,1,45,2022,119.63333333333334\n2022-02-15,138,2,46,2022,121.36666666666666\n2022-02-16,140,3,47,2022,121.0\n2022-02-17,171,4,48,2022,121.7\n2022-02-18,108,5,49,2022,119.23333333333333\n2022-02-19,63,6,50,2022,116.13333333333334\n2022-02-20,82,7,51,2022,113.66666666666667\n2022-02-21,165,1,52,2022,116.1\n2022-02-22,120,2,53,2022,116.26666666666667\n2022-02-23,163,3,54,2022,117.0\n2022-02-24,168,4,55,2022,118.43333333333334\n2022-02-25,122,5,56,2022,117.23333333333333\n2022-02-26,37,6,57,2022,114.83333333333333\n2022-02-27,42,7,58,2022,111.9\n2022-02-28,97,1,59,2022,111.7\n2022-03-01,90,2,60,2022,112.46666666666667\n2022-03-02,122,3,61,2022,111.46666666666667\n2022-03-03,103,4,62,2022,109.8\n2022-03-04,112,5,63,2022,108.2\n2022-03-05,92,6,64,2022,107.8\n2022-03-06,62,7,65,2022,105.73333333333333\n2022-03-07,118,1,66,2022,108.4\n2022-03-08,96,2,67,2022,109.26666666666667\n2022-03-09,110,3,68,2022,108.63333333333334\n2022-03-10,131,4,69,2022,108.56666666666666\n2022-03-11,101,5,70,2022,106.86666666666666\n2022-03-12,54,6,71,2022,103.86666666666666\n2022-03-13,53,7,72,2022,101.96666666666667\n2022-03-14,133,1,73,2022,104.8\n2022-03-15,121,2,74,2022,107.4\n2022-03-16,129,3,75,2022,108.1\n2022-03-17,123,4,76,2022,107.6\n2022-03-18,121,5,77,2022,106.96666666666667\n2022-03-19,62,6,78,2022,103.33333333333333\n2022-03-20,49,7,79,2022,101.36666666666666\n2022-03-21,94,1,80,2022,102.4\n2022-03-22,126,2,81,2022,103.86666666666666\n2022-03-23,152,3,82,2022,103.43333333333334\n2022-03-24,152,4,83,2022,104.5\n2022-03-25,145,5,84,2022,103.9\n2022-03-26,133,6,85,2022,102.73333333333333\n2022-03-27,103,7,86,2022,102.1\n2022-03-28,170,1,87,2022,106.53333333333333\n2022-03-29,125,2,88,2022,109.3\n2022-03-30,146,3,89,2022,110.93333333333334\n2022-03-31,152,4,90,2022,113.0\n2022-04-01,123,5,91,2022,113.03333333333333\n2022-04-02,92,6,92,2022,112.66666666666667\n2022-04-03,77,7,93,2022,111.5\n2022-04-04,174,1,94,2022,114.23333333333333\n2022-04-05,182,2,95,2022,118.23333333333333\n2022-04-06,118,3,96,2022,118.23333333333333\n2022-04-07,114,4,97,2022,118.83333333333333\n2022-04-08,156,5,98,2022,120.36666666666666\n2022-04-09,105,6,99,2022,119.5\n2022-04-10,93,7,100,2022,119.23333333333333\n2022-04-11,135,1,101,2022,121.93333333333334\n2022-04-12,153,2,102,2022,125.26666666666667\n2022-04-13,118,3,103,2022,124.76666666666667\n2022-04-14,149,4,104,2022,125.7\n2022-04-15,169,5,105,2022,127.03333333333333\n2022-04-16,70,6,106,2022,125.26666666666667\n2022-04-17,55,7,107,2022,123.06666666666666\n2022-04-18,125,1,108,2022,125.16666666666667\n2022-04-19,126,2,109,2022,127.73333333333333\n2022-04-20,108,3,110,2022,128.2\n2022-04-21,147,4,111,2022,128.9\n2022-04-22,173,5,112,2022,129.6\n2022-04-23,90,6,113,2022,127.53333333333333\n2022-04-24,100,7,114,2022,126.03333333333333\n2022-04-25,101,1,115,2022,124.96666666666667\n2022-04-26,150,2,116,2022,126.53333333333333\n2022-04-27,158,3,117,2022,126.13333333333334\n2022-04-28,134,4,118,2022,126.43333333333334\n2022-04-29,107,5,119,2022,125.13333333333334\n2022-04-30,50,6,120,2022,121.73333333333333\n2022-05-01,62,7,121,2022,119.7\n2022-05-02,106,1,122,2022,120.16666666666667\n2022-05-03,104,2,123,2022,121.06666666666666\n2022-05-04,133,3,124,2022,119.7\n2022-05-05,167,4,125,2022,119.2\n2022-05-06,140,5,126,2022,119.93333333333334\n2022-05-07,44,6,127,2022,117.6\n2022-05-08,71,7,128,2022,114.76666666666667\n2022-05-09,147,1,129,2022,116.16666666666667\n2022-05-10,119,2,130,2022,117.03333333333333\n2022-05-11,151,3,131,2022,117.56666666666666\n2022-05-12,132,4,132,2022,116.86666666666666\n2022-05-13,112,5,133,2022,116.66666666666667\n2022-05-14,96,6,134,2022,114.9\n2022-05-15,47,7,135,2022,110.83333333333333\n2022-05-16,149,1,136,2022,113.46666666666667\n2022-05-17,141,2,137,2022,116.33333333333333\n2022-05-18,117,3,138,2022,116.06666666666666\n2022-05-19,158,4,139,2022,117.13333333333334\n2022-05-20,112,5,140,2022,117.26666666666667\n2022-05-21,85,6,141,2022,115.2\n2022-05-22,83,7,142,2022,112.2\n2022-05-23,154,1,143,2022,114.33333333333333\n2022-05-24,149,2,144,2022,115.96666666666667\n2022-05-25,113,3,145,2022,116.36666666666666\n2022-05-26,113,4,146,2022,115.13333333333334\n2022-05-27,102,5,147,2022,113.26666666666667\n2022-05-28,99,6,148,2022,112.1\n2022-05-29,56,7,149,2022,110.4\n2022-05-30,107,1,150,2022,112.3\n2022-05-31,106,2,151,2022,113.76666666666667\n2022-06-01,130,3,152,2022,114.56666666666666\n2022-06-02,173,4,153,2022,116.86666666666666\n2022-06-03,134,5,154,2022,116.9\n2022-06-04,70,6,155,2022,113.66666666666667\n2022-06-05,88,7,156,2022,111.93333333333334\n2022-06-06,110,1,157,2022,114.13333333333334\n2022-06-07,113,2,158,2022,115.53333333333333\n2022-06-08,139,3,159,2022,115.26666666666667\n2022-06-09,142,4,160,2022,116.03333333333333\n2022-06-10,109,5,161,2022,114.63333333333334\n2022-06-11,56,6,162,2022,112.1\n2022-06-12,76,7,163,2022,110.9\n2022-06-13,95,1,164,2022,110.86666666666666\n2022-06-14,94,2,165,2022,112.43333333333334\n2022-06-15,127,3,166,2022,111.7\n2022-06-16,200,4,167,2022,113.66666666666667\n2022-06-17,202,5,168,2022,116.5\n2022-06-18,49,6,169,2022,112.86666666666666\n2022-06-19,62,7,170,2022,111.2\n2022-06-20,116,1,171,2022,112.23333333333333\n2022-06-21,112,2,172,2022,113.2\n2022-06-22,131,3,173,2022,112.43333333333334\n2022-06-23,114,4,174,2022,111.26666666666667\n2022-06-24,110,5,175,2022,111.16666666666667\n2022-06-25,52,6,176,2022,109.13333333333334\n2022-06-26,56,7,177,2022,107.6\n2022-06-27,76,1,178,2022,106.83333333333333\n2022-06-28,72,2,179,2022,107.36666666666666\n2022-06-29,120,3,180,2022,107.8\n2022-06-30,118,4,181,2022,108.2\n2022-07-01,87,5,182,2022,106.76666666666667\n2022-07-02,32,6,183,2022,102.06666666666666\n2022-07-03,39,7,184,2022,98.9\n2022-07-04,89,1,185,2022,99.53333333333333\n2022-07-05,120,2,186,2022,100.6\n2022-07-06,106,3,187,2022,100.46666666666667\n2022-07-07,138,4,188,2022,101.3\n2022-07-08,109,5,189,2022,100.3\n2022-07-09,42,6,190,2022,96.96666666666667\n2022-07-10,44,7,191,2022,94.8\n2022-07-11,92,1,192,2022,96.0\n2022-07-12,135,2,193,2022,97.96666666666667\n2022-07-13,157,3,194,2022,100.03333333333333\n2022-07-14,124,4,195,2022,101.03333333333333\n2022-07-15,177,5,196,2022,102.7\n2022-07-16,56,6,197,2022,97.9\n2022-07-17,41,7,198,2022,92.53333333333333\n2022-07-18,143,1,199,2022,95.66666666666667\n2022-07-19,119,2,200,2022,97.56666666666666\n2022-07-20,211,3,201,2022,100.73333333333333\n2022-07-21,128,4,202,2022,101.26666666666667\n2022-07-22,162,5,203,2022,102.3\n2022-07-23,44,6,204,2022,99.96666666666667\n2022-07-24,99,7,205,2022,99.6\n2022-07-25,152,1,206,2022,102.93333333333334\n2022-07-26,158,2,207,2022,106.33333333333333\n2022-07-27,133,3,208,2022,108.23333333333333\n2022-07-28,214,4,209,2022,112.96666666666667\n2022-07-29,133,5,210,2022,113.4\n2022-07-30,79,6,211,2022,112.1\n2022-07-31,70,7,212,2022,111.53333333333333\n2022-08-01,130,1,213,2022,114.8\n2022-08-02,123,2,214,2022,117.6\n2022-08-03,105,3,215,2022,118.13333333333334\n2022-08-04,125,4,216,2022,118.3\n2022-08-05,127,5,217,2022,119.0\n2022-08-06,71,6,218,2022,116.76666666666667\n2022-08-07,55,7,219,2022,114.96666666666667\n2022-08-08,108,1,220,2022,117.16666666666667\n2022-08-09,99,2,221,2022,119.0\n2022-08-10,145,3,222,2022,120.76666666666667\n2022-08-11,115,4,223,2022,120.1\n2022-08-12,110,5,224,2022,118.53333333333333\n2022-08-13,100,6,225,2022,117.73333333333333\n2022-08-14,95,7,226,2022,115.0\n2022-08-15,133,1,227,2022,117.56666666666666\n2022-08-16,130,2,228,2022,120.53333333333333\n2022-08-17,131,3,229,2022,120.13333333333334\n2022-08-18,116,4,230,2022,120.03333333333333\n2022-08-19,143,5,231,2022,117.76666666666667\n2022-08-20,82,6,232,2022,116.23333333333333\n2022-08-21,78,7,233,2022,113.43333333333334\n2022-08-22,110,1,234,2022,115.63333333333334\n2022-08-23,130,2,235,2022,116.66666666666667\n2022-08-24,181,3,236,2022,117.63333333333334\n2022-08-25,132,4,237,2022,116.76666666666667\n2022-08-26,105,5,238,2022,115.83333333333333\n2022-08-27,79,6,239,2022,111.33333333333333\n2022-08-28,66,7,240,2022,109.1\n2022-08-29,173,1,241,2022,112.23333333333333\n2022-08-30,174,2,242,2022,115.7\n2022-08-31,174,3,243,2022,117.16666666666667\n2022-09-01,137,4,244,2022,117.63333333333334\n2022-09-02,135,5,245,2022,118.63333333333334\n2022-09-03,63,6,246,2022,116.56666666666666\n2022-09-04,81,7,247,2022,115.03333333333333\n2022-09-05,168,1,248,2022,118.26666666666667\n2022-09-06,207,2,249,2022,123.33333333333333\n2022-09-07,194,3,250,2022,126.2\n2022-09-08,162,4,251,2022,128.3\n2022-09-09,135,5,252,2022,127.96666666666667\n2022-09-10,54,6,253,2022,125.93333333333334\n2022-09-11,53,7,254,2022,124.03333333333333\n2022-09-12,123,1,255,2022,124.8\n2022-09-13,120,2,256,2022,125.63333333333334\n2022-09-14,137,3,257,2022,125.76666666666667\n2022-09-15,158,4,258,2022,126.7\n2022-09-16,110,5,259,2022,126.0\n2022-09-17,77,6,260,2022,124.7\n2022-09-18,86,7,261,2022,122.8\n2022-09-19,165,1,262,2022,125.56666666666666\n2022-09-20,177,2,263,2022,128.86666666666667\n2022-09-21,156,3,264,2022,130.4\n2022-09-22,113,4,265,2022,129.83333333333334\n2022-09-23,137,5,266,2022,128.36666666666667\n2022-09-24,94,6,267,2022,127.1\n2022-09-25,44,7,268,2022,125.06666666666666\n2022-09-26,122,1,269,2022,126.5\n2022-09-27,121,2,270,2022,128.33333333333334\n2022-09-28,159,3,271,2022,127.86666666666666\n2022-09-29,139,4,272,2022,126.7\n2022-09-30,122,5,273,2022,124.96666666666667\n2022-10-01,81,6,274,2022,123.1\n2022-10-02,57,7,275,2022,120.5\n2022-10-03,155,1,276,2022,123.56666666666666\n2022-10-04,127,2,277,2022,125.1\n2022-10-05,114,3,278,2022,123.3\n2022-10-06,139,4,279,2022,121.03333333333333\n2022-10-07,88,5,280,2022,117.5\n2022-10-08,61,6,281,2022,114.13333333333334\n2022-10-09,66,7,282,2022,111.83333333333333\n2022-10-10,130,1,283,2022,114.36666666666666\n2022-10-11,138,2,284,2022,117.2\n2022-10-12,126,3,285,2022,117.3\n2022-10-13,128,4,286,2022,117.56666666666666\n2022-10-14,147,5,287,2022,117.9\n2022-10-15,70,6,288,2022,114.96666666666667\n2022-10-16,65,7,289,2022,113.46666666666667\n2022-10-17,125,1,290,2022,115.06666666666666\n2022-10-18,192,2,291,2022,118.6\n2022-10-19,107,3,292,2022,116.66666666666667\n2022-10-20,121,4,293,2022,114.8\n2022-10-21,90,5,294,2022,112.6\n2022-10-22,66,6,295,2022,111.03333333333333\n2022-10-23,60,7,296,2022,108.46666666666667\n2022-10-24,99,1,297,2022,108.63333333333334\n2022-10-25,84,2,298,2022,109.96666666666667\n2022-10-26,79,3,299,2022,108.53333333333333\n2022-10-27,95,4,300,2022,107.66666666666667\n2022-10-28,79,5,301,2022,105.0\n2022-10-29,43,6,302,2022,101.8\n2022-10-30,34,7,303,2022,98.86666666666666\n2022-10-31,133,1,304,2022,100.6\n2022-11-01,101,2,305,2022,102.06666666666666\n2022-11-02,124,3,306,2022,101.03333333333333\n2022-11-03,105,4,307,2022,100.3\n2022-11-04,86,5,308,2022,99.36666666666666\n2022-11-05,62,6,309,2022,96.8\n2022-11-06,67,7,310,2022,96.1\n2022-11-07,101,1,311,2022,97.43333333333334\n2022-11-08,117,2,312,2022,99.13333333333334\n2022-11-09,109,3,313,2022,98.43333333333334\n2022-11-10,152,4,314,2022,98.9\n2022-11-11,89,5,315,2022,97.66666666666667\n2022-11-12,43,6,316,2022,94.83333333333333\n2022-11-13,67,7,317,2022,92.16666666666667\n2022-11-14,122,1,318,2022,93.9\n2022-11-15,115,2,319,2022,95.56666666666666\n2022-11-16,171,3,320,2022,97.1\n2022-11-17,149,4,321,2022,95.66666666666667\n2022-11-18,123,5,322,2022,96.2\n2022-11-19,77,6,323,2022,94.73333333333333\n2022-11-20,100,7,324,2022,95.06666666666666\n2022-11-21,133,1,325,2022,97.3\n2022-11-22,142,2,326,2022,100.03333333333333\n2022-11-23,154,3,327,2022,101.86666666666666\n2022-11-24,114,4,328,2022,102.86666666666666\n2022-11-25,116,5,329,2022,104.1\n2022-11-26,32,6,330,2022,102.0\n2022-11-27,42,7,331,2022,100.76666666666667\n2022-11-28,155,1,332,2022,104.5\n2022-11-29,130,2,333,2022,107.7\n2022-11-30,149,3,334,2022,108.23333333333333\n2022-12-01,124,4,335,2022,109.0\n2022-12-02,157,5,336,2022,110.1\n2022-12-03,51,6,337,2022,108.3\n2022-12-04,56,7,338,2022,107.3\n2022-12-05,123,1,339,2022,109.33333333333333\n2022-12-06,133,2,340,2022,111.53333333333333\n2022-12-07,125,3,341,2022,112.33333333333333\n2022-12-08,111,4,342,2022,112.13333333333334\n2022-12-09,139,5,343,2022,113.13333333333334\n2022-12-10,66,6,344,2022,110.26666666666667\n2022-12-11,81,7,345,2022,110.0\n2022-12-12,102,1,346,2022,111.96666666666667\n2022-12-13,143,2,347,2022,114.5\n2022-12-14,96,3,348,2022,113.63333333333334\n2022-12-15,163,4,349,2022,115.23333333333333\n2022-12-16,159,5,350,2022,114.83333333333333\n2022-12-17,70,6,351,2022,112.2\n2022-12-18,68,7,352,2022,110.36666666666666\n2022-12-19,93,1,353,2022,110.9\n2022-12-20,132,2,354,2022,111.96666666666667\n2022-12-21,119,3,355,2022,111.5\n2022-12-22,119,4,356,2022,110.73333333333333\n2022-12-23,84,5,357,2022,108.4\n2022-12-24,54,6,358,2022,106.4\n2022-12-25,22,7,359,2022,103.26666666666667\n2022-12-26,44,1,360,2022,103.66666666666667\n2022-12-27,64,2,361,2022,104.4\n2022-12-28,57,3,362,2022,101.13333333333334\n2022-12-29,85,4,363,2022,99.63333333333334\n2022-12-30,55,5,364,2022,96.5\n2022-12-31,39,6,365,2022,93.66666666666667\n2023-01-01,40,7,1,2023,89.76666666666667\n2023-01-02,93,1,2,2023,91.16666666666667\n2023-01-03,99,2,3,2023,92.6\n2023-01-04,109,3,4,2023,92.13333333333334\n2023-01-05,114,4,5,2023,91.5\n2023-01-06,110,5,6,2023,91.0\n2023-01-07,60,6,7,2023,89.3\n2023-01-08,56,7,8,2023,86.53333333333333\n2023-01-09,144,1,9,2023,89.13333333333334\n2023-01-10,104,2,10,2023,89.9\n2023-01-11,144,3,11,2023,91.3\n2023-01-12,117,4,12,2023,90.43333333333334\n2023-01-13,154,5,13,2023,92.36666666666666\n2023-01-14,60,6,14,2023,88.93333333333334\n2023-01-15,40,7,15,2023,84.96666666666667\n2023-01-16,118,1,16,2023,86.56666666666666\n2023-01-17,111,2,17,2023,88.0\n2023-01-18,128,3,18,2023,89.16666666666667\n2023-01-19,175,4,19,2023,90.6\n2023-01-20,169,5,20,2023,92.26666666666667\n2023-01-21,68,6,21,2023,90.56666666666666\n2023-01-22,88,7,22,2023,90.7\n2023-01-23,154,1,23,2023,94.03333333333333\n2023-01-24,103,2,24,2023,96.73333333333333\n2023-01-25,117,3,25,2023,99.16666666666667\n2023-01-26,128,4,26,2023,101.3\n2023-01-27,135,5,27,2023,103.9\n2023-01-28,47,6,28,2023,102.63333333333334\n2023-01-29,74,7,29,2023,103.26666666666667\n2023-01-30,149,1,30,2023,106.93333333333334\n2023-01-31,175,2,31,2023,111.43333333333334\n2023-02-01,136,3,32,2023,112.86666666666666\n2023-02-02,159,4,33,2023,114.86666666666666\n2023-02-03,79,5,34,2023,113.86666666666666\n2023-02-04,80,6,35,2023,112.73333333333333\n2023-02-05,54,7,36,2023,110.86666666666666\n2023-02-06,125,1,37,2023,113.03333333333333\n2023-02-07,142,2,38,2023,115.9\n2023-02-08,132,3,39,2023,115.5\n2023-02-09,119,4,40,2023,116.0\n2023-02-10,102,5,41,2023,114.6\n2023-02-11,47,6,42,2023,112.26666666666667\n2023-02-12,50,7,43,2023,108.8\n2023-02-13,129,1,44,2023,111.1\n2023-02-14,110,2,45,2023,113.43333333333334\n2023-02-15,101,3,46,2023,112.86666666666666\n2023-02-16,116,4,47,2023,113.03333333333333\n2023-02-17,75,5,48,2023,111.26666666666667\n2023-02-18,59,6,49,2023,107.4\n2023-02-19,73,7,50,2023,104.2\n2023-02-20,103,1,51,2023,105.36666666666666\n2023-02-21,100,2,52,2023,105.76666666666667\n2023-02-22,115,3,53,2023,104.46666666666667\n2023-02-23,99,4,54,2023,104.33333333333333\n2023-02-24,155,5,55,2023,105.6\n2023-02-25,61,6,56,2023,103.36666666666666\n2023-02-26,43,7,57,2023,100.3\n2023-02-27,128,1,58,2023,103.0\n2023-02-28,138,2,59,2023,105.13333333333334\n2023-03-01,106,3,60,2023,103.7\n2023-03-02,101,4,61,2023,101.23333333333333\n2023-03-03,127,5,62,2023,100.93333333333334\n2023-03-04,44,6,63,2023,97.1\n2023-03-05,46,7,64,2023,96.0\n2023-03-06,129,1,65,2023,97.63333333333334\n2023-03-07,113,2,66,2023,99.6\n2023-03-08,97,3,67,2023,98.66666666666667\n2023-03-09,104,4,68,2023,97.4\n2023-03-10,129,5,69,2023,97.3\n2023-03-11,59,6,70,2023,95.3\n2023-03-12,54,7,71,2023,93.7\n2023-03-13,104,1,72,2023,95.6\n2023-03-14,109,2,73,2023,97.56666666666666\n2023-03-15,109,3,74,2023,96.9\n2023-03-16,132,4,75,2023,97.63333333333334\n2023-03-17,96,5,76,2023,97.46666666666667\n2023-03-18,79,6,77,2023,96.23333333333333\n2023-03-19,65,7,78,2023,95.9\n2023-03-20,71,1,79,2023,96.3\n2023-03-21,75,2,80,2023,96.36666666666666\n2023-03-22,118,3,81,2023,96.86666666666666\n2023-03-23,105,4,82,2023,97.03333333333333\n2023-03-24,87,5,83,2023,96.1\n2023-03-25,45,6,84,2023,94.3\n2023-03-26,41,7,85,2023,90.5\n2023-03-27,119,1,86,2023,92.43333333333334\n2023-03-28,138,2,87,2023,95.6\n2023-03-29,123,3,88,2023,95.43333333333334\n2023-03-30,148,4,89,2023,95.76666666666667\n2023-03-31,137,5,90,2023,96.8\n2023-04-01,67,6,91,2023,95.66666666666667\n2023-04-02,52,7,92,2023,93.16666666666667\n2023-04-03,116,1,93,2023,95.56666666666666\n2023-04-04,101,2,94,2023,97.4\n2023-04-05,124,3,95,2023,97.23333333333333\n2023-04-06,94,4,96,2023,96.6\n2023-04-07,79,5,97,2023,96.0\n2023-04-08,53,6,98,2023,94.3\n2023-04-09,47,7,99,2023,91.56666666666666\n2023-04-10,80,1,100,2023,92.26666666666667\n2023-04-11,103,2,101,2023,93.9\n2023-04-12,134,3,102,2023,94.9\n2023-04-13,149,4,103,2023,96.23333333333333\n2023-04-14,100,5,104,2023,95.93333333333334\n2023-04-15,52,6,105,2023,93.26666666666667\n2023-04-16,60,7,106,2023,92.06666666666666\n2023-04-17,86,1,107,2023,92.3\n2023-04-18,117,2,108,2023,94.03333333333333\n2023-04-19,86,3,109,2023,94.53333333333333\n2023-04-20,71,4,110,2023,94.4\n2023-04-21,79,5,111,2023,93.1\n2023-04-22,45,6,112,2023,91.1\n2023-04-23,38,7,113,2023,89.46666666666667\n2023-04-24,119,1,114,2023,91.93333333333334\n2023-04-25,111,2,115,2023,94.26666666666667\n2023-04-26,120,3,116,2023,94.3\n2023-04-27,82,4,117,2023,92.43333333333334\n2023-04-28,84,5,118,2023,91.13333333333334\n2023-04-29,66,6,119,2023,88.4\n2023-04-30,22,7,120,2023,84.56666666666666\n2023-05-01,71,1,121,2023,84.7\n2023-05-02,85,2,122,2023,85.8\n2023-05-03,72,3,123,2023,84.33333333333333\n2023-05-04,92,4,124,2023,84.03333333333333\n2023-05-05,94,5,125,2023,83.03333333333333\n2023-05-06,27,6,126,2023,80.8\n2023-05-07,34,7,127,2023,79.3\n2023-05-08,88,1,128,2023,80.46666666666667\n2023-05-09,88,2,129,2023,81.83333333333333\n2023-05-10,115,3,130,2023,83.0\n2023-05-11,116,4,131,2023,83.43333333333334\n2023-05-12,113,5,132,2023,82.73333333333333\n2023-05-13,28,6,133,2023,78.7\n2023-05-14,14,7,134,2023,75.83333333333333\n2023-05-15,96,1,135,2023,77.3\n2023-05-16,98,2,136,2023,78.56666666666666\n2023-05-17,95,3,137,2023,78.86666666666666\n2023-05-18,124,4,138,2023,79.1\n2023-05-19,97,5,139,2023,79.46666666666667\n2023-05-20,44,6,140,2023,78.56666666666666\n2023-05-21,70,7,141,2023,78.26666666666667\n2023-05-22,118,1,142,2023,80.7\n2023-05-23,106,2,143,2023,82.96666666666667\n2023-05-24,108,3,144,2023,82.6\n2023-05-25,73,4,145,2023,81.33333333333333\n2023-05-26,106,5,146,2023,80.86666666666666\n2023-05-27,42,6,147,2023,79.53333333333333\n2023-05-28,37,7,148,2023,77.96666666666667\n2023-05-29,54,1,149,2023,77.56666666666666\n2023-05-30,98,2,150,2023,80.1\n2023-05-31,95,3,151,2023,80.9\n2023-06-01,103,4,152,2023,81.5\n2023-06-02,92,5,153,2023,82.16666666666667\n2023-06-03,42,6,154,2023,80.5\n2023-06-04,35,7,155,2023,78.53333333333333\n2023-06-05,68,1,156,2023,79.9\n2023-06-06,71,2,157,2023,81.13333333333334\n2023-06-07,61,3,158,2023,80.23333333333333\n2023-06-08,84,4,159,2023,80.1\n2023-06-09,83,5,160,2023,79.03333333333333\n2023-06-10,37,6,161,2023,76.4\n2023-06-11,29,7,162,2023,73.6\n2023-06-12,78,1,163,2023,75.26666666666667\n2023-06-13,78,2,164,2023,77.4\n2023-06-14,90,3,165,2023,77.2\n2023-06-15,80,4,166,2023,76.6\n2023-06-16,98,5,167,2023,76.7\n2023-06-17,54,6,168,2023,74.36666666666666\n2023-06-18,30,7,169,2023,72.13333333333334\n2023-06-19,80,1,170,2023,73.33333333333333\n2023-06-20,115,2,171,2023,74.83333333333333\n2023-06-21,121,3,172,2023,74.93333333333334\n2023-06-22,104,4,173,2023,74.86666666666666\n2023-06-23,73,5,174,2023,73.7\n2023-06-24,38,6,175,2023,72.53333333333333\n2023-06-25,41,7,176,2023,70.36666666666666\n2023-06-26,126,1,177,2023,73.16666666666667\n2023-06-27,94,2,178,2023,75.06666666666666\n2023-06-28,74,3,179,2023,75.73333333333333\n2023-06-29,102,4,180,2023,75.86666666666666\n2023-06-30,118,5,181,2023,76.63333333333334\n2023-07-01,46,6,182,2023,74.73333333333333\n2023-07-02,35,7,183,2023,72.83333333333333\n2023-07-03,107,1,184,2023,75.0\n2023-07-04,67,2,185,2023,76.06666666666666\n2023-07-05,90,3,186,2023,76.8\n2023-07-06,132,4,187,2023,78.83333333333333\n2023-07-07,110,5,188,2023,80.46666666666667\n2023-07-08,25,6,189,2023,78.5\n2023-07-09,26,7,190,2023,76.6\n2023-07-10,85,1,191,2023,78.2\n2023-07-11,90,2,192,2023,80.23333333333333\n2023-07-12,83,3,193,2023,80.4\n2023-07-13,113,4,194,2023,81.56666666666666\n2023-07-14,121,5,195,2023,82.6\n2023-07-15,27,6,196,2023,80.83333333333333\n2023-07-16,23,7,197,2023,78.33333333333333\n2023-07-17,48,1,198,2023,78.13333333333334\n2023-07-18,63,2,199,2023,79.23333333333333\n2023-07-19,52,3,200,2023,78.3\n2023-07-20,67,4,201,2023,76.7\n2023-07-21,78,5,202,2023,75.26666666666667\n2023-07-22,38,6,203,2023,73.06666666666666\n2023-07-23,41,7,204,2023,72.0\n2023-07-24,113,1,205,2023,74.5\n2023-07-25,99,2,206,2023,76.43333333333334\n2023-07-26,103,3,207,2023,75.66666666666667\n2023-07-27,88,4,208,2023,75.46666666666667\n2023-07-28,88,5,209,2023,75.93333333333334\n2023-07-29,52,6,210,2023,74.26666666666667\n2023-07-30,36,7,211,2023,71.53333333333333\n2023-07-31,86,1,212,2023,72.86666666666666\n2023-08-01,84,2,213,2023,74.5\n2023-08-02,104,3,214,2023,74.4\n2023-08-03,94,4,215,2023,75.3\n2023-08-04,89,5,216,2023,75.26666666666667\n2023-08-05,46,6,217,2023,72.4\n2023-08-06,41,7,218,2023,70.1\n2023-08-07,53,1,219,2023,71.03333333333333\n2023-08-08,100,2,220,2023,73.5\n2023-08-09,111,3,221,2023,74.36666666666666\n2023-08-10,180,4,222,2023,77.36666666666666\n2023-08-11,82,5,223,2023,77.33333333333333\n2023-08-12,84,6,224,2023,76.36666666666666\n2023-08-13,33,7,225,2023,73.43333333333334\n2023-08-14,100,1,226,2023,75.86666666666666\n2023-08-15,93,2,227,2023,78.2\n2023-08-16,177,3,228,2023,82.5\n2023-08-17,118,4,229,2023,84.33333333333333\n2023-08-18,93,5,230,2023,85.7\n2023-08-19,57,6,231,2023,85.36666666666666\n2023-08-20,61,7,232,2023,84.8\n2023-08-21,90,1,233,2023,86.53333333333333\n2023-08-22,75,2,234,2023,87.66666666666667\n2023-08-23,115,3,235,2023,87.73333333333333\n2023-08-24,109,4,236,2023,88.06666666666666\n2023-08-25,78,5,237,2023,87.23333333333333\n2023-08-26,48,6,238,2023,85.9\n2023-08-27,45,7,239,2023,84.46666666666667\n2023-08-28,104,1,240,2023,86.2\n2023-08-29,79,2,241,2023,87.63333333333334\n2023-08-30,129,3,242,2023,89.06666666666666\n2023-08-31,104,4,243,2023,89.73333333333333\n2023-09-01,82,5,244,2023,89.0\n2023-09-02,39,6,245,2023,87.16666666666667\n2023-09-03,25,7,246,2023,85.03333333333333\n2023-09-04,43,1,247,2023,84.93333333333334\n2023-09-05,114,2,248,2023,87.36666666666666\n2023-09-06,90,3,249,2023,88.6\n2023-09-07,120,4,250,2023,89.26666666666667\n2023-09-08,72,5,251,2023,87.96666666666667\n2023-09-09,74,6,252,2023,84.43333333333334\n2023-09-10,43,7,253,2023,83.13333333333334\n2023-09-11,74,1,254,2023,82.8\n2023-09-12,76,2,255,2023,84.23333333333333\n2023-09-13,59,3,256,2023,82.86666666666666\n2023-09-14,79,4,257,2023,82.4\n2023-09-15,76,5,258,2023,79.03333333333333\n2023-09-16,55,6,259,2023,76.93333333333334\n2023-09-17,24,7,260,2023,74.63333333333334\n2023-09-18,99,1,261,2023,76.03333333333333\n2023-09-19,77,2,262,2023,76.56666666666666\n2023-09-20,65,3,263,2023,75.73333333333333\n2023-09-21,71,4,264,2023,75.6\n2023-09-22,54,5,265,2023,73.56666666666666\n2023-09-23,26,6,266,2023,70.8\n2023-09-24,49,7,267,2023,69.83333333333333\n2023-09-25,79,1,268,2023,70.86666666666666\n2023-09-26,80,2,269,2023,72.03333333333333\n2023-09-27,113,3,270,2023,72.33333333333333\n2023-09-28,91,4,271,2023,72.73333333333333\n2023-09-29,76,5,272,2023,70.96666666666667\n2023-09-30,33,6,273,2023,68.6\n2023-10-01,28,7,274,2023,66.8\n2023-10-02,80,1,275,2023,68.16666666666667\n2023-10-03,80,2,276,2023,70.0\n2023-10-04,88,3,277,2023,71.5\n2023-10-05,77,4,278,2023,70.26666666666667\n2023-10-06,96,5,279,2023,70.46666666666667\n2023-10-07,51,6,280,2023,68.16666666666667\n2023-10-08,48,7,281,2023,67.36666666666666\n2023-10-09,88,1,282,2023,67.83333333333333\n2023-10-10,76,2,283,2023,68.93333333333334\n2023-10-11,56,3,284,2023,68.33333333333333\n2023-10-12,90,4,285,2023,68.8\n2023-10-13,73,5,286,2023,69.26666666666667\n2023-10-14,30,6,287,2023,67.63333333333334\n2023-10-15,34,7,288,2023,66.23333333333333\n2023-10-16,97,1,289,2023,67.63333333333334\n2023-10-17,118,2,290,2023,70.76666666666667\n2023-10-18,125,3,291,2023,71.63333333333334\n2023-10-19,93,4,292,2023,72.16666666666667\n2023-10-20,81,5,293,2023,72.7\n2023-10-21,55,6,294,2023,72.16666666666667\n2023-10-22,44,7,295,2023,71.83333333333333\n2023-10-23,53,1,296,2023,72.73333333333333\n2023-10-24,84,2,297,2023,73.9\n2023-10-25,102,3,298,2023,74.66666666666667\n2023-10-26,97,4,299,2023,75.23333333333333\n2023-10-27,89,5,300,2023,74.43333333333334\n2023-10-28,44,6,301,2023,72.86666666666666\n2023-10-29,69,7,302,2023,72.63333333333334\n2023-10-30,127,1,303,2023,75.76666666666667\n2023-10-31,134,2,304,2023,79.3\n2023-11-01,112,3,305,2023,80.36666666666666\n2023-11-02,85,4,306,2023,80.53333333333333\n2023-11-03,84,5,307,2023,80.4\n2023-11-04,52,6,308,2023,79.56666666666666\n2023-11-05,63,7,309,2023,78.46666666666667\n2023-11-06,114,1,310,2023,80.56666666666666\n2023-11-07,140,2,311,2023,83.63333333333334\n2023-11-08,99,3,312,2023,84.0\n2023-11-09,99,4,313,2023,84.76666666666667\n2023-11-10,98,5,314,2023,86.16666666666667\n2023-11-11,67,6,315,2023,85.4\n2023-11-12,52,7,316,2023,84.7\n2023-11-13,100,1,317,2023,87.03333333333333\n2023-11-14,106,2,318,2023,89.43333333333334\n2023-11-15,122,3,319,2023,90.26666666666667\n2023-11-16,113,4,320,2023,90.1\n2023-11-17,107,5,321,2023,89.5\n2023-11-18,47,6,322,2023,87.96666666666667\n2023-11-19,35,7,323,2023,86.43333333333334\n2023-11-20,84,1,324,2023,87.4\n2023-11-21,108,2,325,2023,89.53333333333333\n2023-11-22,91,3,326,2023,90.8\n2023-11-23,86,4,327,2023,90.86666666666666\n2023-11-24,95,5,328,2023,90.63333333333334\n2023-11-25,32,6,329,2023,88.46666666666667\n2023-11-26,54,7,330,2023,87.3\n2023-11-27,66,1,331,2023,88.03333333333333\n2023-11-28,95,2,332,2023,88.9\n2023-11-29,87,3,333,2023,87.56666666666666\n2023-11-30,91,4,334,2023,86.13333333333334\n2023-12-01,43,5,335,2023,83.83333333333333\n2023-12-02,32,6,336,2023,82.06666666666666\n2023-12-03,19,7,337,2023,79.9\n","format":{"type":"csv"}}});</script></div>

<div style="background-color:grey;height:2px;width:100%;"></div>

<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/london-clojurians-talk/projects/london-clojurians-dec-2023-talk/src/notebooks/commits.clj">projects/london-clojurians-dec-2023-talk/src/notebooks/commits.clj</a></small></small></pre></div>