### 2023-12-03 projects/london-clojurians-dec-2023-talk/docs/notebooks.explore-revealjs.qmd

---
format:
  revealjs:
    theme: [default, src/notebooks/custom.scss]
    include-in-header: [text, <link rel="shortcut icon" href="Cookbook.svg" />]
    navigation-mode: vertical
    transition: slide
    background-transition: fade
    incremental: true
    output-file: notebooks.explore-revealjs.html
highlight-style: solarized

---
<style>table {
  border-style: thin;
}
th, td {
  padding: 6px;
}
td {
  text-align: left;
}
th {
  text-align: center;
  background-color: #ddd;
}
tr:nth-child(even) {
  background-color: #f6f6f6;
}
</style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="notebooks.explore-revealjs.html_files/md-default0.js" type="text/javascript"></script><script src="notebooks.explore-revealjs.html_files/md-default1.js" type="text/javascript"></script>

<div class="sourceClojure">
```clojure
(ns notebooks.explore
  (:require [scicloj.clay.v2.config :as config]))
```
</div>

<div class="sourceClojure">
```clojure
(+ 1 2)
```
</div>

<div class="printedClojure">
```clojure
3

```
</div>

# A1

## B1

## B2

# A2

<div style="background-color:grey;height:2px;width:100%;"></div>

<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/london-clojurians-talk//home/daslu/projects/scicloj/clojure-data-scrapbook/projects/london-clojurians-dec-2023-talk/src/notebooks/explore.clj">/home/daslu/projects/scicloj/clojure-data-scrapbook/projects/london-clojurians-dec-2023-talk/src/notebooks/explore.clj</a></small></small></pre></div>