### 2023-12-03 projects/london-clojurians-dec-2023-talk/docs/notebooks.explore.qmd

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: notebooks.explore.html}
highlight-style: solarized

---
<style>table {
  border-style: thin;
}
th, td {
  padding: 6px;
}
td {
  text-align: left;
}
th {
  text-align: center;
  background-color: #ddd;
}
tr:nth-child(even) {
  background-color: #f6f6f6;
}
</style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="notebooks.explore_files/md-default0.js" type="text/javascript"></script><script src="notebooks.explore_files/md-default1.js" type="text/javascript"></script>

<div class="sourceClojure">
```clojure
(ns notebooks.explore
  (:require [data.generate-dataset]
            [scicloj.kindly.v4.kind :as kind]
            [scicloj.noj.v1.vis :as vis]
            [tablecloth.api :as tc]
            [clojure.string :as string]))
```
</div>

<div class="sourceClojure">
```clojure
(def processed-data
  (-> (->> data.generate-dataset/data
           (mapcat (comp :items :items))
           (map #(select-keys % [:html_url :open_issues_count :watchers :forks_count])))
      tc/dataset
      (tc/map-columns :user [:html_url]
                      (fn [html-url]
                        (-> html-url
                            (string/split #"/")
                            (->> (drop 3)
                                 first))))))
```
</div>

<div class="sourceClojure">
```clojure
(-> processed-data
    (tc/group-by [:user])
    (tc/aggregate {:n tc/row-count})
    (tc/order-by [:n] :desc))
```
</div>

_unnamed [580 2]:

|              :user |  :n |
|--------------------|----:|
|            clojure |  30 |
|        clj-commons |  26 |
|        weavejester |  22 |
|              oakes |  13 |
|            metosin |  12 |
|           taoensso |  11 |
|               juxt |  10 |
|           babashka |   9 |
|        swannodette |   9 |
|            yogthos |   8 |
|                ... | ... |
|         congomongo |   1 |
|          ReactiveX |   1 |
|        ingesolvoll |   1 |
|         hashobject |   1 |
|         jsa-aerial |   1 |
|          jonromero |   1 |
| alexander-yakushev |   1 |
|               l3nz |   1 |
|                gcv |   1 |
|          MastodonC |   1 |
|          gnarroway |   1 |

<div style="background-color:grey;height:2px;width:100%;"></div>

<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/london-clojurians-talk/projects/london-clojurians-dec-2023-talk/src/notebooks/explore.clj">projects/london-clojurians-dec-2023-talk/src/notebooks/explore.clj</a></small></small></pre></div>