### 2023-12-22 docs/projects/london-clojurians-dec-2023-talk/notebooks.combining_tools.qmd

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: notebooks.combining_tools.html}
highlight-style: solarized

---
<style>table {
  border-style: thin;
}
th, td {
  padding: 6px;
}
td {
  text-align: left;
}
th {
  text-align: center;
  background-color: #ddd;
}
tr:nth-child(even) {
  background-color: #f6f6f6;
}
</style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="notebooks.combining_tools_files/md-default0.js" type="text/javascript"></script><script src="notebooks.combining_tools_files/md-default1.js" type="text/javascript"></script>
# Combining tools for data exploration

##
![](https://scicloj.github.io/sci-cloj-logo-transparent.png)

[scicloj.org](http://scicloj.org)

##
![](https://raw.githubusercontent.com/practicalli-john/london-clojurians-logo/master/london-clojurians-logo.svg)

##
![](https://sessionize.com/image/68c6-1140o400o3-Boo2pA28xQfvKskyXqwcmd.jpg)

## Plan
- Dynamic report making with slow processes
- Data processing & visualization in harmony

## Use case: analysing commit logs
- Which projects are encouraging community conributions?

## Data pipeline
- take the 1000 most popular Clojure repos (according to Github's API)
- .. and also the Python ones
- clone them
- look into the commit logs
- process & visualize

## Limitations
- a biased data sample
- descriptive statistics, not statistical inference
- interpretation is not obvious

# Dynamic report making with slow processes

## Slow processes
- collecting remote data
- fetching API responses
- applying slow AI models

# Data processing & visualization in harmony

## Lots of existing functionality

- dtype-next - array processing
- tech.ml.dataset - high-performance datasets
- Tablecloth - dataset ergonomics
- Fasthmath - lots of math & stats
- scicloj.ml - machine learning, data modelling
- Hanami - data visualization
- Libpython-clj, ClojisR - interop
- Visual tools
- etc.

## Noj
- collecting the relevant dependencies
- composing them for common use cases

# Thanks

<div style="background-color:grey;height:2px;width:100%;"></div>

<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/london-clojurians-talk/projects/london-clojurians-dec-2023-talk/src/notebooks/combining_tools.clj">projects/london-clojurians-dec-2023-talk/src/notebooks/combining_tools.clj</a></small></small></pre></div>