### 2023-12-22 docs/projects/london-clojurians-dec-2023-talk/notebooks.commit_details_gradually.qmd

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: notebooks.commit_details_gradually.html}
highlight-style: solarized

---
<style>table {
  border-style: thin;
}
th, td {
  padding: 6px;
}
td {
  text-align: left;
}
th {
  text-align: center;
  background-color: #ddd;
}
tr:nth-child(even) {
  background-color: #f6f6f6;
}
</style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="notebooks.commit_details_gradually_files/md-default1.js" type="text/javascript"></script><script src="notebooks.commit_details_gradually_files/md-default2.js" type="text/javascript"></script><script src="notebooks.commit_details_gradually_files/vega3.js" type="text/javascript"></script><script src="notebooks.commit_details_gradually_files/vega4.js" type="text/javascript"></script><script src="notebooks.commit_details_gradually_files/vega5.js" type="text/javascript"></script>

<div class="sourceClojure">
```clojure
(ns notebooks.commit-details-gradually
  (:require [tablecloth.api :as tc]
            [scicloj.noj.v1.vis.hanami :as hanami]
            [aerial.hanami.templates :as ht]
            [tech.v3.datatype.datetime :as datetime]
            [tech.v3.datatype.rolling :as rolling]
            [tech.v3.datatype.functional :as fun]
            [tech.v3.dataset.print :as print]
            [data.generate-dataset]
            [clojure.math :as math]
            [scicloj.kindly.v4.kind :as kind]
            [scicloj.noj.v1.stats :as stats]
            [scicloj.clay.v2.api :as clay]
            [clojure.string :as string]
            [clojure.java.shell :as shell]
            [scicloj.kind-portal.v1.api :as kind-portal]
            [portal.api :as portal]
            [clojure.java.io :as io]
            [scicloj.kind-portal.v1.api :as kind-portal]))
```
</div>

<div class="sourceClojure">
```clojure
(defn url->clone-path [url]
  (-> url
      (string/replace #"^https://github.com/" "")
      (string/replace #"/" "__")
      (->> (str "/workspace/clones-for-gradual-analysis/"))))
```
</div>

<div class="sourceClojure">
```clojure
(defn clone! [{:keys [html_url clone-path]}]
  (io/make-parents clone-path)
  (when-not (-> clone-path
                io/file
                (.exists))
    (prn [:git-clone html_url clone-path])
    (shell/sh "git"
              "clone"
              html_url
              clone-path)))
```
</div>

<div class="sourceClojure">
```clojure
(defn clone-all! [])
```
</div>

<div class="sourceClojure">
```clojure
(defn git-log [{:as repo
                :keys [html_url clone-path]}]
  (prn [:git-log clone-path])
  (-> clone-path
      (->> (format "--git-dir=%s/.git"))
      (#(shell/sh "git" %
                  "log"
                  "--pretty=format:\"%ad,%ae\""
                  "--date=short"))
      :out
      (string/replace #"\"" "")
      string/split-lines
      (->> (map (fn [line]
                  (-> line
                      (string/split #",")
                      ((fn [[date email]]
                         (merge repo
                                {:date date
                                 :email email})))))))
      tc/dataset
      (tc/set-dataset-name html_url)))
```
</div>

<div class="sourceClojure">
```clojure
(defonce *stop (atom false))
```
</div>

<div class="sourceClojure">
```clojure
(defonce *raw-commit-logs (atom []))
```
</div>

<div class="sourceClojure">
```clojure
(defonce *portal-commit-logs (atom []))
```
</div>

<div class="sourceClojure">
```clojure
(defn collect! [limit]
  (prn [:DBG1])
  (reset! *stop false)
  (prn [:DBG2])
  (reset! *raw-commit-logs [])
  (prn [:DBG2])
  (reset! *portal-commit-logs [])
  (prn [:DBG3])
  (-> data.generate-dataset/repos-ds
      (tc/map-columns :clone-path [:html_url]
                      url->clone-path)
      (tc/select-rows #(and (-> % :size (< 100000))
                            (-> % :language (= "Clojure"))))
      (tc/rows :as-maps)
      (->> (take limit)
           (mapv (fn [repo]
                   (prn repo)
                   (when-not @*stop
                     (clone! repo)
                     (let [log (git-log repo)]
                       (swap! *raw-commit-logs
                              conj log)
                       (swap! *portal-commit-logs
                              #(-> %
                                   (conj (kind-portal/prepare
                                          {:value log}))
                                   reverse))))))))
  (prn [:stopped]))
```
</div>

<div class="sourceClojure">
```clojure
(comment
  (future (collect! 220))

  (reset! *stop true)

  (portal/open)
  (portal/submit *portal-commit-logs))
```
</div>

<div class="sourceClojure">
```clojure
(defn report []
  (let [raw-commit-logs @*raw-commit-logs
        n (count raw-commit-logs)]
    [(kind/hiccup
      [:h1 n " repos"])
     (when (> n 10)
       (-> raw-commit-logs
           (->> (apply tc/concat))
           (tc/group-by [:language :html_url])
           (tc/aggregate {:n-contributors (fn [ds]
                                            (-> ds
                                                :email
                                                distinct
                                                count))})
           (tc/order-by [:n-contributors] :desc)
           (tc/left-join data.generate-dataset/repos-ds [:html_url])
           (tc/select-rows #(-> % :n-contributors (> 20)))
           (tc/add-columns {:log-stargazers #(-> %
                                                 :stargazers_count
                                                 fun/log)
                            :log-contributors #(-> %
                                                   :n-contributors
                                                   fun/log)})
           (stats/add-predictions :log-stargazers
                                  [:log-contributors]
                                  {:model-type :smile.regression/ordinary-least-square})
           (hanami/linear-regression-plot
            :log-contributors
            :log-stargazers
            {:XSCALE {:zero false
                      :domain [5 12]}
             :YSCALE {:zero false}
             :point-options {:MSIZE 100}
             :line-options {:MCOLOR "brown"
                            :MSIZE 10
                            :OPACITY 0.5}})
           (assoc-in [:encoding :tooltip]
                     {:field "html_url"})))]))
```
</div>

<div class="sourceClojure">
```clojure
(def *report (atom nil))
```
</div>

<div class="sourceClojure">
```clojure
(defn update-report! []
  (prn [:update-report])
  (reset! *report (kind-portal/prepare
                   {:value (report)})))
```
</div>

<div class="sourceClojure">
```clojure
(comment

  (report)

  (update-report!)

  (add-watch *raw-commit-logs
             :report
             (fn [_ _ _ _]
               (update-report!)))

  (portal/submit *report))
```
</div>

<div class="sourceClojure">
```clojure
(report)
```
</div>

<div><p>\[</p><div style="margin-left:10%;width:110%;"><h1>194 repos</h1><div><script>vegaEmbed(document.currentScript.parentElement, {"encoding":{"tooltip":{"field":"html_url"}},"width":400,"background":"floralwhite","title":{"text":"R^2 = 0.213"},"layer":[{"encoding":{"y":{"scale":{"zero":false},"field":"log-contributors","type":"quantitative"},"x":{"scale":{"zero":false,"domain":[5,12]},"field":"log-stargazers","type":"quantitative"}},"mark":{"type":"circle","size":100,"tooltip":true},"width":400,"background":"floralwhite","title":{"text":"R^2 = 0.213"},"height":300},{"encoding":{"y":{"scale":{"zero":false},"field":"log-contributors-prediction","type":"quantitative","axis":{"title":"log-contributors"}},"opacity":{"value":0.5},"x":{"scale":{"zero":false,"domain":[5,12]},"field":"log-stargazers","type":"quantitative"}},"mark":{"type":"line","size":10,"color":"brown","tooltip":true},"width":400,"background":"floralwhite","title":{"text":"R^2 = 0.213"},"height":300}],"height":300,"data":{"url":"notebooks.commit_details_gradually_files\/0.csv","format":{"type":"csv"}}});</script></div></div><p>\]</p></div>

<div style="background-color:grey;height:2px;width:100%;"></div>

<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/london-clojurians-talk/projects/london-clojurians-dec-2023-talk/src/notebooks/commit_details_gradually.clj">projects/london-clojurians-dec-2023-talk/src/notebooks/commit_details_gradually.clj</a></small></small></pre></div>