### 2023-12-22 docs/projects/london-clojurians-dec-2023-talk/notebooks.exploring_commit_logs.qmd

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: notebooks.exploring_commit_logs.html}
highlight-style: solarized

---
<style>table {
  border-style: thin;
}
th, td {
  padding: 6px;
}
td {
  text-align: left;
}
th {
  text-align: center;
  background-color: #ddd;
}
tr:nth-child(even) {
  background-color: #f6f6f6;
}
</style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="notebooks.exploring_commit_logs_files/md-default0.js" type="text/javascript"></script><script src="notebooks.exploring_commit_logs_files/md-default1.js" type="text/javascript"></script>
AA

<div style="background-color:grey;height:2px;width:100%;"></div>

<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/london-clojurians-talk/projects/london-clojurians-dec-2023-talk/src/notebooks/exploring_commit_logs.clj">projects/london-clojurians-dec-2023-talk/src/notebooks/exploring_commit_logs.clj</a></small></small></pre></div>