### 2024-01-01 docs/projects/visual-tools/clay-calva-demo-20231216/index.qmd

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: index.html}
highlight-style: solarized
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

---
<style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="index_files/md-default1.js" type="text/javascript"></script><script src="index_files/md-default2.js" type="text/javascript"></script><script src="https://cdn.datatables.net/1.13.6/js/jquery.dataTables.min.js" type="text/javascript"></script><script src="index_files/vega3.js" type="text/javascript"></script><script src="index_files/vega4.js" type="text/javascript"></script><script src="index_files/vega5.js" type="text/javascript"></script><link href="https://cdn.datatables.net/1.13.6/css/jquery.dataTables.min.css" rel="stylesheet" type="text/css" />
<table><tr><td>This is part of the Scicloj <a href="https://scicloj.github.io/clojure-data-scrapbook/">Clojure Data Scrapbook</a>.</td><a href="https://scicloj.github.io/clojure-data-scrapbook/"><img align="left" alt="SciCloj logo" src="https://scicloj.github.io/sci-cloj-logo-transparent.png" width="40" /></a></tr></table>

# Clay & Noj demo: data visualization

<div class="sourceClojure">
```clojure
(ns index
  (:require [scicloj.noj.v1.datasets
             :as datasets]
            [scicloj.noj.v1.vis.hanami
             :as hanami]
            [scicloj.kindly.v4.kind
             :as kind]))
```
</div>

## Arithmetic

<div class="sourceClojure">
```clojure
(+ 1 2)
```
</div>

<div class="printedClojure">
```clojure
3

```
</div>

## Datasets

<div class="sourceClojure">
```clojure
datasets/iris
```
</div>

_unnamed [150 5]:

| :sepal-length | :sepal-width | :petal-length | :petal-width |  :species |
|--------------:|-------------:|--------------:|-------------:|-----------|
|           5.1 |          3.5 |           1.4 |          0.2 |    setosa |
|           4.9 |          3.0 |           1.4 |          0.2 |    setosa |
|           4.7 |          3.2 |           1.3 |          0.2 |    setosa |
|           4.6 |          3.1 |           1.5 |          0.2 |    setosa |
|           5.0 |          3.6 |           1.4 |          0.2 |    setosa |
|           5.4 |          3.9 |           1.7 |          0.4 |    setosa |
|           4.6 |          3.4 |           1.4 |          0.3 |    setosa |
|           5.0 |          3.4 |           1.5 |          0.2 |    setosa |
|           4.4 |          2.9 |           1.4 |          0.2 |    setosa |
|           4.9 |          3.1 |           1.5 |          0.1 |    setosa |
|           ... |          ... |           ... |          ... |       ... |
|           6.9 |          3.1 |           5.4 |          2.1 | virginica |
|           6.7 |          3.1 |           5.6 |          2.4 | virginica |
|           6.9 |          3.1 |           5.1 |          2.3 | virginica |
|           5.8 |          2.7 |           5.1 |          1.9 | virginica |
|           6.8 |          3.2 |           5.9 |          2.3 | virginica |
|           6.7 |          3.3 |           5.7 |          2.5 | virginica |
|           6.7 |          3.0 |           5.2 |          2.3 | virginica |
|           6.3 |          2.5 |           5.0 |          1.9 | virginica |
|           6.5 |          3.0 |           5.2 |          2.0 | virginica |
|           6.2 |          3.4 |           5.4 |          2.3 | virginica |
|           5.9 |          3.0 |           5.1 |          1.8 | virginica |

## Tables

<div class="sourceClojure">
```clojure
(-> datasets/iris
    kind/table)
```
</div>

<table class="table"><thead><tr><th>sepal-length</th><th>sepal-width</th><th>petal-length</th><th>petal-width</th><th>species</th></tr></thead><tbody><tr><td><div><p>5.1</p></div></td><td><div><p>3.5</p></div></td><td><div><p>1.4</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.9</p></div></td><td><div><p>3.0</p></div></td><td><div><p>1.4</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.7</p></div></td><td><div><p>3.2</p></div></td><td><div><p>1.3</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.6</p></div></td><td><div><p>3.1</p></div></td><td><div><p>1.5</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.0</p></div></td><td><div><p>3.6</p></div></td><td><div><p>1.4</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.4</p></div></td><td><div><p>3.9</p></div></td><td><div><p>1.7</p></div></td><td><div><p>0.4</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.6</p></div></td><td><div><p>3.4</p></div></td><td><div><p>1.4</p></div></td><td><div><p>0.3</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.0</p></div></td><td><div><p>3.4</p></div></td><td><div><p>1.5</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.4</p></div></td><td><div><p>2.9</p></div></td><td><div><p>1.4</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.9</p></div></td><td><div><p>3.1</p></div></td><td><div><p>1.5</p></div></td><td><div><p>0.1</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.4</p></div></td><td><div><p>3.7</p></div></td><td><div><p>1.5</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.8</p></div></td><td><div><p>3.4</p></div></td><td><div><p>1.6</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.8</p></div></td><td><div><p>3.0</p></div></td><td><div><p>1.4</p></div></td><td><div><p>0.1</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.3</p></div></td><td><div><p>3.0</p></div></td><td><div><p>1.1</p></div></td><td><div><p>0.1</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.8</p></div></td><td><div><p>4.0</p></div></td><td><div><p>1.2</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.7</p></div></td><td><div><p>4.4</p></div></td><td><div><p>1.5</p></div></td><td><div><p>0.4</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.4</p></div></td><td><div><p>3.9</p></div></td><td><div><p>1.3</p></div></td><td><div><p>0.4</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.1</p></div></td><td><div><p>3.5</p></div></td><td><div><p>1.4</p></div></td><td><div><p>0.3</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.7</p></div></td><td><div><p>3.8</p></div></td><td><div><p>1.7</p></div></td><td><div><p>0.3</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.1</p></div></td><td><div><p>3.8</p></div></td><td><div><p>1.5</p></div></td><td><div><p>0.3</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.4</p></div></td><td><div><p>3.4</p></div></td><td><div><p>1.7</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.1</p></div></td><td><div><p>3.7</p></div></td><td><div><p>1.5</p></div></td><td><div><p>0.4</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.6</p></div></td><td><div><p>3.6</p></div></td><td><div><p>1.0</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.1</p></div></td><td><div><p>3.3</p></div></td><td><div><p>1.7</p></div></td><td><div><p>0.5</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.8</p></div></td><td><div><p>3.4</p></div></td><td><div><p>1.9</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.0</p></div></td><td><div><p>3.0</p></div></td><td><div><p>1.6</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.0</p></div></td><td><div><p>3.4</p></div></td><td><div><p>1.6</p></div></td><td><div><p>0.4</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.2</p></div></td><td><div><p>3.5</p></div></td><td><div><p>1.5</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.2</p></div></td><td><div><p>3.4</p></div></td><td><div><p>1.4</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.7</p></div></td><td><div><p>3.2</p></div></td><td><div><p>1.6</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.8</p></div></td><td><div><p>3.1</p></div></td><td><div><p>1.6</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.4</p></div></td><td><div><p>3.4</p></div></td><td><div><p>1.5</p></div></td><td><div><p>0.4</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.2</p></div></td><td><div><p>4.1</p></div></td><td><div><p>1.5</p></div></td><td><div><p>0.1</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.5</p></div></td><td><div><p>4.2</p></div></td><td><div><p>1.4</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.9</p></div></td><td><div><p>3.1</p></div></td><td><div><p>1.5</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.0</p></div></td><td><div><p>3.2</p></div></td><td><div><p>1.2</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.5</p></div></td><td><div><p>3.5</p></div></td><td><div><p>1.3</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.9</p></div></td><td><div><p>3.6</p></div></td><td><div><p>1.4</p></div></td><td><div><p>0.1</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.4</p></div></td><td><div><p>3.0</p></div></td><td><div><p>1.3</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.1</p></div></td><td><div><p>3.4</p></div></td><td><div><p>1.5</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.0</p></div></td><td><div><p>3.5</p></div></td><td><div><p>1.3</p></div></td><td><div><p>0.3</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.5</p></div></td><td><div><p>2.3</p></div></td><td><div><p>1.3</p></div></td><td><div><p>0.3</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.4</p></div></td><td><div><p>3.2</p></div></td><td><div><p>1.3</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.0</p></div></td><td><div><p>3.5</p></div></td><td><div><p>1.6</p></div></td><td><div><p>0.6</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.1</p></div></td><td><div><p>3.8</p></div></td><td><div><p>1.9</p></div></td><td><div><p>0.4</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.8</p></div></td><td><div><p>3.0</p></div></td><td><div><p>1.4</p></div></td><td><div><p>0.3</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.1</p></div></td><td><div><p>3.8</p></div></td><td><div><p>1.6</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>4.6</p></div></td><td><div><p>3.2</p></div></td><td><div><p>1.4</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.3</p></div></td><td><div><p>3.7</p></div></td><td><div><p>1.5</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>5.0</p></div></td><td><div><p>3.3</p></div></td><td><div><p>1.4</p></div></td><td><div><p>0.2</p></div></td><td><div><p>setosa</p></div></td></tr><tr><td><div><p>7.0</p></div></td><td><div><p>3.2</p></div></td><td><div><p>4.7</p></div></td><td><div><p>1.4</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.4</p></div></td><td><div><p>3.2</p></div></td><td><div><p>4.5</p></div></td><td><div><p>1.5</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.9</p></div></td><td><div><p>3.1</p></div></td><td><div><p>4.9</p></div></td><td><div><p>1.5</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.5</p></div></td><td><div><p>2.3</p></div></td><td><div><p>4.0</p></div></td><td><div><p>1.3</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.5</p></div></td><td><div><p>2.8</p></div></td><td><div><p>4.6</p></div></td><td><div><p>1.5</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.7</p></div></td><td><div><p>2.8</p></div></td><td><div><p>4.5</p></div></td><td><div><p>1.3</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.3</p></div></td><td><div><p>3.3</p></div></td><td><div><p>4.7</p></div></td><td><div><p>1.6</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>4.9</p></div></td><td><div><p>2.4</p></div></td><td><div><p>3.3</p></div></td><td><div><p>1.0</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.6</p></div></td><td><div><p>2.9</p></div></td><td><div><p>4.6</p></div></td><td><div><p>1.3</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.2</p></div></td><td><div><p>2.7</p></div></td><td><div><p>3.9</p></div></td><td><div><p>1.4</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.0</p></div></td><td><div><p>2.0</p></div></td><td><div><p>3.5</p></div></td><td><div><p>1.0</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.9</p></div></td><td><div><p>3.0</p></div></td><td><div><p>4.2</p></div></td><td><div><p>1.5</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.0</p></div></td><td><div><p>2.2</p></div></td><td><div><p>4.0</p></div></td><td><div><p>1.0</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.1</p></div></td><td><div><p>2.9</p></div></td><td><div><p>4.7</p></div></td><td><div><p>1.4</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.6</p></div></td><td><div><p>2.9</p></div></td><td><div><p>3.6</p></div></td><td><div><p>1.3</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.7</p></div></td><td><div><p>3.1</p></div></td><td><div><p>4.4</p></div></td><td><div><p>1.4</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.6</p></div></td><td><div><p>3.0</p></div></td><td><div><p>4.5</p></div></td><td><div><p>1.5</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.8</p></div></td><td><div><p>2.7</p></div></td><td><div><p>4.1</p></div></td><td><div><p>1.0</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.2</p></div></td><td><div><p>2.2</p></div></td><td><div><p>4.5</p></div></td><td><div><p>1.5</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.6</p></div></td><td><div><p>2.5</p></div></td><td><div><p>3.9</p></div></td><td><div><p>1.1</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.9</p></div></td><td><div><p>3.2</p></div></td><td><div><p>4.8</p></div></td><td><div><p>1.8</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.1</p></div></td><td><div><p>2.8</p></div></td><td><div><p>4.0</p></div></td><td><div><p>1.3</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.3</p></div></td><td><div><p>2.5</p></div></td><td><div><p>4.9</p></div></td><td><div><p>1.5</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.1</p></div></td><td><div><p>2.8</p></div></td><td><div><p>4.7</p></div></td><td><div><p>1.2</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.4</p></div></td><td><div><p>2.9</p></div></td><td><div><p>4.3</p></div></td><td><div><p>1.3</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.6</p></div></td><td><div><p>3.0</p></div></td><td><div><p>4.4</p></div></td><td><div><p>1.4</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.8</p></div></td><td><div><p>2.8</p></div></td><td><div><p>4.8</p></div></td><td><div><p>1.4</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.7</p></div></td><td><div><p>3.0</p></div></td><td><div><p>5.0</p></div></td><td><div><p>1.7</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.0</p></div></td><td><div><p>2.9</p></div></td><td><div><p>4.5</p></div></td><td><div><p>1.5</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.7</p></div></td><td><div><p>2.6</p></div></td><td><div><p>3.5</p></div></td><td><div><p>1.0</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.5</p></div></td><td><div><p>2.4</p></div></td><td><div><p>3.8</p></div></td><td><div><p>1.1</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.5</p></div></td><td><div><p>2.4</p></div></td><td><div><p>3.7</p></div></td><td><div><p>1.0</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.8</p></div></td><td><div><p>2.7</p></div></td><td><div><p>3.9</p></div></td><td><div><p>1.2</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.0</p></div></td><td><div><p>2.7</p></div></td><td><div><p>5.1</p></div></td><td><div><p>1.6</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.4</p></div></td><td><div><p>3.0</p></div></td><td><div><p>4.5</p></div></td><td><div><p>1.5</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.0</p></div></td><td><div><p>3.4</p></div></td><td><div><p>4.5</p></div></td><td><div><p>1.6</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.7</p></div></td><td><div><p>3.1</p></div></td><td><div><p>4.7</p></div></td><td><div><p>1.5</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.3</p></div></td><td><div><p>2.3</p></div></td><td><div><p>4.4</p></div></td><td><div><p>1.3</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.6</p></div></td><td><div><p>3.0</p></div></td><td><div><p>4.1</p></div></td><td><div><p>1.3</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.5</p></div></td><td><div><p>2.5</p></div></td><td><div><p>4.0</p></div></td><td><div><p>1.3</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.5</p></div></td><td><div><p>2.6</p></div></td><td><div><p>4.4</p></div></td><td><div><p>1.2</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.1</p></div></td><td><div><p>3.0</p></div></td><td><div><p>4.6</p></div></td><td><div><p>1.4</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.8</p></div></td><td><div><p>2.6</p></div></td><td><div><p>4.0</p></div></td><td><div><p>1.2</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.0</p></div></td><td><div><p>2.3</p></div></td><td><div><p>3.3</p></div></td><td><div><p>1.0</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.6</p></div></td><td><div><p>2.7</p></div></td><td><div><p>4.2</p></div></td><td><div><p>1.3</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.7</p></div></td><td><div><p>3.0</p></div></td><td><div><p>4.2</p></div></td><td><div><p>1.2</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.7</p></div></td><td><div><p>2.9</p></div></td><td><div><p>4.2</p></div></td><td><div><p>1.3</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.2</p></div></td><td><div><p>2.9</p></div></td><td><div><p>4.3</p></div></td><td><div><p>1.3</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.1</p></div></td><td><div><p>2.5</p></div></td><td><div><p>3.0</p></div></td><td><div><p>1.1</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>5.7</p></div></td><td><div><p>2.8</p></div></td><td><div><p>4.1</p></div></td><td><div><p>1.3</p></div></td><td><div><p>versicolor</p></div></td></tr><tr><td><div><p>6.3</p></div></td><td><div><p>3.3</p></div></td><td><div><p>6.0</p></div></td><td><div><p>2.5</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>5.8</p></div></td><td><div><p>2.7</p></div></td><td><div><p>5.1</p></div></td><td><div><p>1.9</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>7.1</p></div></td><td><div><p>3.0</p></div></td><td><div><p>5.9</p></div></td><td><div><p>2.1</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.3</p></div></td><td><div><p>2.9</p></div></td><td><div><p>5.6</p></div></td><td><div><p>1.8</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.5</p></div></td><td><div><p>3.0</p></div></td><td><div><p>5.8</p></div></td><td><div><p>2.2</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>7.6</p></div></td><td><div><p>3.0</p></div></td><td><div><p>6.6</p></div></td><td><div><p>2.1</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>4.9</p></div></td><td><div><p>2.5</p></div></td><td><div><p>4.5</p></div></td><td><div><p>1.7</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>7.3</p></div></td><td><div><p>2.9</p></div></td><td><div><p>6.3</p></div></td><td><div><p>1.8</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.7</p></div></td><td><div><p>2.5</p></div></td><td><div><p>5.8</p></div></td><td><div><p>1.8</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>7.2</p></div></td><td><div><p>3.6</p></div></td><td><div><p>6.1</p></div></td><td><div><p>2.5</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.5</p></div></td><td><div><p>3.2</p></div></td><td><div><p>5.1</p></div></td><td><div><p>2.0</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.4</p></div></td><td><div><p>2.7</p></div></td><td><div><p>5.3</p></div></td><td><div><p>1.9</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.8</p></div></td><td><div><p>3.0</p></div></td><td><div><p>5.5</p></div></td><td><div><p>2.1</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>5.7</p></div></td><td><div><p>2.5</p></div></td><td><div><p>5.0</p></div></td><td><div><p>2.0</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>5.8</p></div></td><td><div><p>2.8</p></div></td><td><div><p>5.1</p></div></td><td><div><p>2.4</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.4</p></div></td><td><div><p>3.2</p></div></td><td><div><p>5.3</p></div></td><td><div><p>2.3</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.5</p></div></td><td><div><p>3.0</p></div></td><td><div><p>5.5</p></div></td><td><div><p>1.8</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>7.7</p></div></td><td><div><p>3.8</p></div></td><td><div><p>6.7</p></div></td><td><div><p>2.2</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>7.7</p></div></td><td><div><p>2.6</p></div></td><td><div><p>6.9</p></div></td><td><div><p>2.3</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.0</p></div></td><td><div><p>2.2</p></div></td><td><div><p>5.0</p></div></td><td><div><p>1.5</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.9</p></div></td><td><div><p>3.2</p></div></td><td><div><p>5.7</p></div></td><td><div><p>2.3</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>5.6</p></div></td><td><div><p>2.8</p></div></td><td><div><p>4.9</p></div></td><td><div><p>2.0</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>7.7</p></div></td><td><div><p>2.8</p></div></td><td><div><p>6.7</p></div></td><td><div><p>2.0</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.3</p></div></td><td><div><p>2.7</p></div></td><td><div><p>4.9</p></div></td><td><div><p>1.8</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.7</p></div></td><td><div><p>3.3</p></div></td><td><div><p>5.7</p></div></td><td><div><p>2.1</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>7.2</p></div></td><td><div><p>3.2</p></div></td><td><div><p>6.0</p></div></td><td><div><p>1.8</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.2</p></div></td><td><div><p>2.8</p></div></td><td><div><p>4.8</p></div></td><td><div><p>1.8</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.1</p></div></td><td><div><p>3.0</p></div></td><td><div><p>4.9</p></div></td><td><div><p>1.8</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.4</p></div></td><td><div><p>2.8</p></div></td><td><div><p>5.6</p></div></td><td><div><p>2.1</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>7.2</p></div></td><td><div><p>3.0</p></div></td><td><div><p>5.8</p></div></td><td><div><p>1.6</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>7.4</p></div></td><td><div><p>2.8</p></div></td><td><div><p>6.1</p></div></td><td><div><p>1.9</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>7.9</p></div></td><td><div><p>3.8</p></div></td><td><div><p>6.4</p></div></td><td><div><p>2.0</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.4</p></div></td><td><div><p>2.8</p></div></td><td><div><p>5.6</p></div></td><td><div><p>2.2</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.3</p></div></td><td><div><p>2.8</p></div></td><td><div><p>5.1</p></div></td><td><div><p>1.5</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.1</p></div></td><td><div><p>2.6</p></div></td><td><div><p>5.6</p></div></td><td><div><p>1.4</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>7.7</p></div></td><td><div><p>3.0</p></div></td><td><div><p>6.1</p></div></td><td><div><p>2.3</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.3</p></div></td><td><div><p>3.4</p></div></td><td><div><p>5.6</p></div></td><td><div><p>2.4</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.4</p></div></td><td><div><p>3.1</p></div></td><td><div><p>5.5</p></div></td><td><div><p>1.8</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.0</p></div></td><td><div><p>3.0</p></div></td><td><div><p>4.8</p></div></td><td><div><p>1.8</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.9</p></div></td><td><div><p>3.1</p></div></td><td><div><p>5.4</p></div></td><td><div><p>2.1</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.7</p></div></td><td><div><p>3.1</p></div></td><td><div><p>5.6</p></div></td><td><div><p>2.4</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.9</p></div></td><td><div><p>3.1</p></div></td><td><div><p>5.1</p></div></td><td><div><p>2.3</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>5.8</p></div></td><td><div><p>2.7</p></div></td><td><div><p>5.1</p></div></td><td><div><p>1.9</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.8</p></div></td><td><div><p>3.2</p></div></td><td><div><p>5.9</p></div></td><td><div><p>2.3</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.7</p></div></td><td><div><p>3.3</p></div></td><td><div><p>5.7</p></div></td><td><div><p>2.5</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.7</p></div></td><td><div><p>3.0</p></div></td><td><div><p>5.2</p></div></td><td><div><p>2.3</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.3</p></div></td><td><div><p>2.5</p></div></td><td><div><p>5.0</p></div></td><td><div><p>1.9</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.5</p></div></td><td><div><p>3.0</p></div></td><td><div><p>5.2</p></div></td><td><div><p>2.0</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>6.2</p></div></td><td><div><p>3.4</p></div></td><td><div><p>5.4</p></div></td><td><div><p>2.3</p></div></td><td><div><p>virginica</p></div></td></tr><tr><td><div><p>5.9</p></div></td><td><div><p>3.0</p></div></td><td><div><p>5.1</p></div></td><td><div><p>1.8</p></div></td><td><div><p>virginica</p></div></td></tr></tbody><script>new DataTable(document.currentScript.parentElement, {"sPagunationType":"full_numbers","order":[]});</script></table>

## Visualization

<div class="sourceClojure">
```clojure
(-> datasets/iris
    (hanami/linear-regression-plot
     :petal-length
     :petal-width
     {:line-options {:MCOLOR "brown"
                     :MSIZE 10
                     :OPACITY 0.5}
      :point-options {:MSIZE 100}}))
```
</div>

<div><script>vegaEmbed(document.currentScript.parentElement, {"width":400,"background":"floralwhite","title":{"text":"R^2 = 0.927"},"layer":[{"encoding":{"y":{"field":"petal-length","type":"quantitative"},"x":{"field":"petal-width","type":"quantitative"}},"mark":{"type":"circle","size":100,"tooltip":true},"width":400,"background":"floralwhite","title":{"text":"R^2 = 0.927"},"height":300},{"encoding":{"y":{"field":"petal-length-prediction","type":"quantitative","axis":{"title":"petal-length"}},"opacity":{"value":0.5},"x":{"field":"petal-width","type":"quantitative"}},"mark":{"type":"line","size":10,"color":"brown","tooltip":true},"width":400,"background":"floralwhite","title":{"text":"R^2 = 0.927"},"height":300}],"height":300,"data":{"url":"index_files\/0.csv","format":{"type":"csv"}}});</script></div>

<div style="background-color:grey;height:2px;width:100%;"></div>

<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/main/projects/visual-tools/clay-calva-demo-20231216/notebooks/index.clj">projects/visual-tools/clay-calva-demo-20231216/notebooks/index.clj</a></small></small></pre></div>