### 2024-01-20 projects/geography/seattle-parks/docs/index.qmd
Ranking Seattle neighborhoods by the proportion of area covered by parks. WIP

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: index.html}
highlight-style: solarized
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="index_files/md-default0.js" type="text/javascript"></script><script src="index_files/md-default1.js" type="text/javascript"></script><script src="index_files/portal2.js" type="text/javascript"></script>

# Seattle Parks & Neighborhoods

::: {.sourceClojure}
```clojure
(ns index
  (:require [geo
             [geohash :as geohash]
             [jts :as jts]
             [spatial :as spatial]
             [io :as geoio]
             [crs :as crs]]
            [tech.v3.datatype.functional :as fun]
            [tablecloth.api :as tc]
            [scicloj.kindly.v4.kind :as kind]
            [scicloj.kindly.v4.api :as kindly])
  (:import (org.locationtech.jts.index.strtree STRtree)
           (org.locationtech.jts.geom Geometry Point Polygon)
           (org.locationtech.jts.geom.prep PreparedGeometry
                                           PreparedLineString
                                           PreparedPolygon
                                           PreparedGeometryFactory)
           (java.util TreeMap)))
```
:::

In this analysis we rank Seattle neighborhoods
by the presence of of Parks in them
(in terms of area proportion).

## Raw geo data

::: {.sourceClojure}
```clojure
(defn read-geogson-gz
  "Read a gzipped GeoJSON file."
  [path]
  ;; See [Reading and Writing Compressed Files](Reading and Writing Compressed Files) by John Cromartie.
  (with-open [in (java.util.zip.GZIPInputStream.
                  (clojure.java.io/input-stream
                   path))]
    (-> in
        slurp
        geoio/read-geojson)))
```
:::

::: {.sourceClojure}
```clojure
(def neighborhoods-geojson
  (read-geogson-gz
   "data/Seattle/Neighborhood_Map_Atlas_Neighborhoods.geojson.gz"))
```
:::

::: {.sourceClojure}
```clojure
(count neighborhoods-geojson)
```
:::

::: {.printedClojure}
```clojure
94

```
:::

::: {.sourceClojure}
```clojure
(->> neighborhoods-geojson
     (take 5)
     kind/portal)
```
:::

<div style="max-height:600px;overflow-y:auto;"><div><script>portal_api.embed().renderOutputItem(
                {'mime': 'x-application/edn',
                 'text': (() => "({^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :properties]] {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OBJECTID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 27]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :L_HOOD]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"Ballard\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :S_HOOD]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"Loyal Heights\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :S_HOOD_ALT_NAMES]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :Shape__Area]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 2.13206555455933E7]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :Shape__Length]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 18831.00959637]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geometry]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector #object[org.locationtech.jts.geom.Polygon 0x1f3c397d \"POLYGON ((-122.376336564723 47.6759176989664, -122.376707907517 47.6759982290459, -122.377903057631 47.6759978849021, -122.379988625303 47.6759893429385, -122.38182788443 47.675992079457, -122.383602752477 47.675979810424, -122.385435721425 47.6759587956618, -122.387087434206 47.6759521343769, -122.387702391216 47.6759476875929, -122.388955759168 47.6759424488678, -122.390449346263 47.6759259868868, -122.391773109273 47.6759158131504, -122.392827172066 47.6759153143353, -122.392956550545 47.6759141859541, -122.392956303464 47.6763287099834, -122.392959618409 47.676929558447, -122.392961263359 47.6775085790216, -122.392971502986 47.6781034200237, -122.392975009807 47.6792517755972, -122.392983771129 47.6798142464259, -122.39300195772 47.6811303942788, -122.393014530791 47.6824285808909, -122.393016158048 47.6837341155494, -122.393022827056 47.685021586284, -122.393030577589 47.6863450742964, -122.393032746621 47.687668638629, -122.393046187318 47.6889956475812, -122.393010665051 47.6902980913372, -122.393002415356 47.6905813205566, -122.392432373275 47.6905826938111, -122.390347910813 47.6905631204492, -122.388192979124 47.6905805029145, -122.385832831602 47.6905766041982, -122.383275316007 47.690593164597, -122.383201215024 47.6905936585447, -122.381453020656 47.6905978829044, -122.379876313408 47.6905988541749, -122.379103841332 47.6906041929746, -122.377117910432 47.6906029639028, -122.376810344766 47.6906051591463, -122.376810264834 47.6905341242701, -122.376810989127 47.6900117143006, -122.376803067488 47.688976146724, -122.376794751014 47.6879443550321, -122.376800402808 47.6870068457338, -122.376784403749 47.6860923366759, -122.376785260609 47.6851813698577, -122.376780535885 47.6842704786057, -122.376787310842 47.6833707854178, -122.376777423242 47.6822860591882, -122.376767820487 47.6811938310584, -122.376770442503 47.680154307228, -122.376770065639 47.6792017979916, -122.376769882216 47.6780678404081, -122.376759750881 47.6775387646826, -122.376761985459 47.676862137338, -122.376772171763 47.6764361697665, -122.376772451106 47.6764114022001, -122.376770700102 47.6763866189683, -122.37676692258 47.6763619488292, -122.376761120573 47.6763374777451, -122.376753802121 47.6763131984825, -122.376744969247 47.6762891966531, -122.376733608705 47.6762655291531, -122.376721246367 47.6762423037704, -122.376706358898 47.6762194979699, -122.376690473449 47.6761972626929, -122.376672575492 47.6761756116819, -122.376653693551 47.6761550023704, -122.376620811506 47.6761243856628, -122.37659791092 47.6761051588887, -122.376573006221 47.6760868163511, -122.376547112434 47.6760693435982, -122.376519722058 47.6760527482053, -122.376336564723 47.6759176989664))\"]]]} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :properties]] {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OBJECTID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 28]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :L_HOOD]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"Ballard\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :S_HOOD]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"Ballard\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :S_HOOD_ALT_NAMES]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"Adams\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :Shape__Area]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 2.25521329700012E7]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :Shape__Length]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 29926.3393002841]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geometry]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector #object[org.locationtech.jts.geom.Polygon 0x7ee462dc \"POLYGON ((-122.376218981706 47.6681795106282, -122.376214914328 47.6671023311431, -122.376222341407 47.6660363069762, -122.376218608455 47.6649703907559, -122.376224015951 47.6638363569551, -122.376213335479 47.6628953406738, -122.376215127819 47.6622035949171, -122.376226087158 47.6616326743118, -122.376232028288 47.6610807160121, -122.376228293294 47.6609890799218, -122.376311034007 47.6610075816866, -122.376347589125 47.6610084573967, -122.376398615454 47.6610011680257, -122.376399768145 47.6609545379595, -122.376409994362 47.6608888474507, -122.376480021595 47.6608892699844, -122.376649883965 47.6608858972706, -122.376802044838 47.6608843923845, -122.376904962586 47.6608818836445, -122.377020085186 47.6608803234087, -122.377118045713 47.6608817375375, -122.377232905674 47.6608884069341, -122.377236922833 47.66093826616, -122.377241607037 47.6610105668296, -122.377247426981 47.6611040169469, -122.377305369317 47.6611246108013, -122.377431782777 47.6611615428936, -122.377533178917 47.661176106273, -122.377584412861 47.6611416500979, -122.377614865595 47.6610908096894, -122.377654375344 47.6610203519417, -122.377677258636 47.6609879511404, -122.377715398852 47.6609397055094, -122.377726544603 47.6609050217262, -122.377743040356 47.6608625536466, -122.377758604553 47.6608228828779, -122.377777465054 47.6607746413824, -122.37779127635 47.6607442918602, -122.377820881952 47.6607161699336, -122.377887950662 47.6607024071657, -122.377930778536 47.6607094956517, -122.377982464813 47.6607244326554, -122.378037380882 47.6607456241841, -122.378088199365 47.6607655003304, -122.378144441121 47.6607800757394, -122.378223937077 47.6608089028157, -122.378285257703 47.6608236665358, -122.378343627844 47.6608415550177, -122.378411485271 47.6608543445299, -122.378488168664 47.6608738696264, -122.378572163043 47.6609004074106, -122.378624823197 47.6609139603348, -122.378697367829 47.6609307992309, -122.378751686634 47.6609489565742, -122.378820681289 47.6609658435608, -122.378862952393 47.6609883202932, -122.378922441567 47.6610097066181, -122.379081035255 47.6610539119557, -122.379265109385 47.6611018840283, -122.379389299035 47.6611322460587, -122.379421555758 47.661107986769, -122.379478983191 47.6611282867812, -122.379642682521 47.6611735357023, -122.379800016148 47.6612264976409, -122.379946110899 47.6612768697639, -122.380065852996 47.6613111901579, -122.380130945811 47.6613333136908, -122.380153431187 47.6613558013855, -122.380174148089 47.6613870535744, -122.380188835134 47.6614202727708, -122.380211038217 47.6614844088921, -122.380243167414 47.6615070228645, -122.380303354114 47.6615177305899, -122.380350754978 47.6615250129872, -122.380395375702 47.6615411592479, -122.380479446062 47.6615701795669, -122.380560130101 47.6616047302561, -122.380614654221 47.6616297389514, -122.380663593942 47.6616546093877, -122.380701872643 47.6616792817304, -122.380715579541 47.6617306802593, -122.380690732307 47.6617822596553, -122.380664357784 47.6618166364496, -122.380648035454 47.6618648436206, -122.38068448469 47.6618791728427, -122.380717946765 47.6618954274041, -122.380752187784 47.6619037455183, -122.3807905011 47.6619125221318, -122.380829216037 47.6619347892012, -122.380866227641 47.6619680050916, -122.380895420488 47.6619942145727, -122.380932407944 47.6620266164946, -122.380968643252 47.6620508026807, -122.381020641587 47.6620761452478, -122.381098871967 47.6621134707754, -122.381167450224 47.6621503707618, -122.381238114463 47.6621891699481, -122.381288223574 47.6622192082262, -122.381338421512 47.6622522447606, -122.3813897159 47.6622880078852, -122.381432789931 47.6623203269265, -122.381503551814 47.6623794330982, -122.381572734045 47.6624365895899, -122.38161407395 47.6624788294554, -122.381656280718 47.6625160872786, -122.381699821145 47.6625470286, -122.381806787089 47.662629206954, -122.381897197478 47.6627006417839, -122.38202451515 47.6628017373534, -122.382159570742 47.6629070976345, -122.382234956653 47.6629680680843, -122.382331552118 47.6630427176998, -122.382358242278 47.6630700320529, -122.382342968803 47.6631023303298, -122.382300682104 47.6631816104811, -122.382248627965 47.663256610552, -122.382205916571 47.663321629404, -122.38216549549 47.6633783907977, -122.382130138578 47.6634348263973, -122.382111305681 47.6634668729497, -122.382088660645 47.6635071976301, -122.382072994312 47.6635433570874, -122.382079921409 47.6635714972306, -122.382102317187 47.6636079530782, -122.382132615196 47.6636371467198, -122.382152009826 47.6636580053991, -122.382179632041 47.6636995745085, -122.382226640061 47.6637617023103, -122.382265013459 47.6637894576431, -122.382287032988 47.6638132797402, -122.382294565904 47.6638447105061, -122.382324691125 47.6638851315305, -122.382351104041 47.6639201617707, -122.382392788521 47.6639568697887, -122.38242584513 47.6639764710789, -122.38247158058 47.6639788482028, -122.382526667558 47.6639545348073, -122.382589068071 47.6639202245904, -122.382688434965 47.6638664743225, -122.382803320873 47.6638057006604, -122.382905199031 47.663751101759, -122.382941717148 47.6637336385167, -122.382966133764 47.6637527577184, -122.383031165255 47.6638067137846, -122.383132900026 47.6639004872976, -122.383239585043 47.663990079865, -122.383357582155 47.664084702746, -122.383387571069 47.6640694706022, -122.383426541835 47.6640661980484, -122.383483321407 47.6641156388281, -122.383559518957 47.6641867091696, -122.383607936888 47.6642279949219, -122.38366498611 47.6642694629143, -122.383705216851 47.6642573901608, -122.38374678886 47.6642902858182, -122.383815084444 47.6643515662311, -122.383891529058 47.6644138925092, -122.383928986632 47.6644449161712, -122.383947792141 47.6644800067992, -122.383881817267 47.6645135521425, -122.383814999715 47.6645528502698, -122.38387902594 47.664607076829, -122.3839775665 47.6646786995044, -122.384135832697 47.6645923552987, -122.384242145277 47.664533325113, -122.384331352798 47.6644791975459, -122.384412812772 47.6644375145094, -122.384489691936 47.6643953371378, -122.384583062228 47.6643447088215, -122.384661052548 47.6643057722524, -122.384713096475 47.6642814988847, -122.384746725629 47.6642692587552, -122.384788936095 47.6643065154152, -122.384849380778 47.6643597190268, -122.384916058885 47.6644177644245, -122.384963881728 47.6644560585874, -122.385017250443 47.66451017243, -122.38495372199 47.6645406434711, -122.38488642209 47.6645807627338, -122.384922310085 47.6646101792558, -122.385012204179 47.6646810620896, -122.385091014123 47.664737571192, -122.385159582752 47.6647909208934, -122.385201908298 47.6648150228112, -122.385244683332 47.6647862053257, -122.385303491814 47.6647505716182, -122.385335942888 47.664766838992, -122.385364999663 47.6647883794225, -122.385420652472 47.6648339787411, -122.385451231037 47.6648554984145, -122.385494792308 47.6648869950637, -122.385533657102 47.6648631581812, -122.385581912972 47.6648309670042, -122.385653628924 47.6647858172476, -122.385701427821 47.6648232971616, -122.385777171524 47.6648790340154, -122.385893932504 47.6649660019066, -122.385994659925 47.6650427770733, -122.386107642287 47.6651221273146, -122.386185644498 47.6651855020668, -122.386271942321 47.6652548044143, -122.386383902301 47.6653338685933, -122.386499667119 47.6654214064798, -122.386605559289 47.6655011526632, -122.386735763139 47.6655964196413, -122.386825130185 47.665666494009, -122.386949062028 47.6657555910979, -122.387060182093 47.6658404499214, -122.387166631533 47.6659218162079, -122.387306671535 47.666023546488, -122.387401610193 47.6660932446749, -122.387475179069 47.6661439966965, -122.387516634605 47.6661728654544, -122.387661780928 47.6662756396602, -122.387757343215 47.6663491849378, -122.387916210759 47.6664695947904, -122.387977069777 47.6665364583886, -122.387952685858 47.6665694384435, -122.387879792562 47.6666430976967, -122.387801925444 47.6667201233013, -122.387757098252 47.6667651647428, -122.387756642093 47.6668008177192, -122.387783862165 47.666845689913, -122.387817905054 47.6668981382935, -122.387853754042 47.6669261407119, -122.387899028582 47.6669639114684, -122.387925044609 47.6669854927731, -122.387989093289 47.6670062977376, -122.388036952435 47.6670286949155, -122.388068309967 47.6670422342458, -122.388102023645 47.6670667095868, -122.388061205643 47.6671100682211, -122.388074746639 47.6671386747837, -122.38812121483 47.6671654609862, -122.388179258012 47.6671720807621, -122.388207496189 47.667149202303, -122.388257608331 47.6670942340291, -122.388295448462 47.6670531011843, -122.388323114251 47.6670280449451, -122.388366114046 47.6670406541589, -122.38840365641 47.667057407844, -122.388461732815 47.6670821074944, -122.388539763869 47.6671123185078, -122.388634327555 47.6671524151722, -122.388685173013 47.6671728433664, -122.38881245377 47.6672210192078, -122.388873688193 47.6672494885957, -122.38891982302 47.6672820632114, -122.388953218756 47.6673128406666, -122.388986123599 47.6673272153655, -122.389022855146 47.6673338260221, -122.389090731696 47.66732989987, -122.389113865738 47.6672212730366, -122.389136699377 47.6671365149558, -122.389173242354 47.6671368297695, -122.389173484642 47.6672805268316, -122.389171263858 47.6674266996656, -122.389202123689 47.6674575116872, -122.389281412655 47.6674619127757, -122.389330063777 47.6674598772914, -122.389379189758 47.6674567641031, -122.389482438853 47.6674479412747, -122.389543221697 47.6674443690314, -122.389608953719 47.6674366159831, -122.389712620413 47.6674247884508, -122.3897598728 47.6674268846986, -122.389807306022 47.6674350195555, -122.389904659175 47.6674495849077, -122.390051923575 47.6674703227754, -122.390137799552 47.6674913424451, -122.390203337111 47.6674940457067, -122.390247520863 47.6674953269465, -122.390283174704 47.6674998095369, -122.390300447147 47.6675343634824, -122.390353770916 47.6675358623119, -122.390433026738 47.6675391490383, -122.390475250473 47.6675596943012, -122.390473342253 47.6676145613099, -122.390477772294 47.6676608156235, -122.390514276394 47.6676598024281, -122.390635008225 47.6676586661427, -122.390746079771 47.6676568906986, -122.390827781605 47.6676571445704, -122.39089434969 47.6676603475363, -122.390898510793 47.6676298710342, -122.39090934373 47.6675849930683, -122.390949953969 47.6675855517923, -122.390993547426 47.6675671324767, -122.39099884901 47.6675069918754, -122.391040432475 47.6675061663225, -122.391084025833 47.6674877469723, -122.391108680302 47.6674638452388, -122.391143028365 47.6674416960582, -122.391201627742 47.6673821978372, -122.391240126961 47.667380043104, -122.391290771825 47.6673598990265, -122.391335198315 47.6673354270804, -122.391379477649 47.6673060300438, -122.391412940626 47.6672712968426, -122.391423768106 47.6672431712659, -122.391428443478 47.6672129446238, -122.391447849834 47.6671663214747, -122.391436925275 47.6671404217219, -122.391413510072 47.6671210333067, -122.391450270666 47.6670947382224, -122.39150153965 47.6670954078681, -122.391596984383 47.6670971013938, -122.391693353066 47.667095783058, -122.391784275357 47.6670989520519, -122.391847118954 47.6670964215749, -122.391887213677 47.6670967297431, -122.39192374287 47.6671134961467, -122.392030845721 47.667114772877, -122.392088872647 47.6671208338811, -122.392091136672 47.6671624904442, -122.392096235224 47.6672309720358, -122.392099602433 47.6673263407844, -122.392123114103 47.6673827879395, -122.392134561906 47.6674091946602, -122.392171081212 47.6674086949117, -122.392252243641 47.667407884292, -122.392389680333 47.6674054463277, -122.392477951447 47.6674047951721, -122.392540287934 47.6674022709074, -122.392632321467 47.6673917139587, -122.392675287915 47.6674031651836, -122.392708553001 47.6674126068037, -122.392806337896 47.6674077119961, -122.392873990148 47.6673963317975, -122.392919826064 47.667385035803, -122.392979481849 47.6673946302089, -122.393020401485 47.6673885858434, -122.393097774332 47.6673798574582, -122.393167294004 47.667379976715, -122.393225074576 47.6673778141995, -122.393221955442 47.6674091336357, -122.393222832524 47.6674721885578, -122.393238436297 47.6675355991731, -122.39329750546 47.6675594685231, -122.39335375596 47.6675570271599, -122.393399964841 47.6675243468383, -122.393444656135 47.6675256198886, -122.393447697804 47.6675593395581, -122.393489845271 47.667560390512, -122.393527436354 47.6675618035958, -122.393566974398 47.6675604481272, -122.393614160218 47.6675603160078, -122.393645262763 47.6675484076, -122.393680391047 47.6675015686923, -122.393720254174 47.6674434394496, -122.393758752289 47.6673905560685, -122.393757692749 47.6673214625349, -122.393758475899 47.6672123268719, -122.393761223881 47.6671179455581, -122.393761680877 47.667014855468, -122.393760758474 47.6669334205727, -122.393761456334 47.6668552627146, -122.393760920936 47.6667867186574, -122.393765722004 47.6666931234954, -122.393763843119 47.6666136296473, -122.393767269073 47.6665587417421, -122.393796709599 47.6665421862331, -122.393833203324 47.6665408720148, -122.393910313456 47.6665403298084, -122.394029980183 47.6665376194448, -122.394117717169 47.6665361172178, -122.394193764847 47.6665340039903, -122.394245997497 47.6665329885134, -122.394307860673 47.6665316265894, -122.394382484335 47.6665327891635, -122.394451454801 47.6665315438599, -122.394526529294 47.6665308150352, -122.394590436569 47.6665299391312, -122.3945952516 47.6665551084125, -122.394635468656 47.6665594842731, -122.394716744218 47.6665624834001, -122.394813536912 47.6665584143733, -122.394843759846 47.6665341357291, -122.394841147025 47.6664809587645, -122.394834886514 47.6664245753894, -122.394832801734 47.6663889575007, -122.394849524551 47.6663544523763, -122.394885346531 47.6663476632794, -122.394935394624 47.6663414930711, -122.395069633444 47.6663171592766, -122.395206433201 47.6662936044866, -122.395308991403 47.6662787878676, -122.395398394388 47.6662652224049, -122.395458038067 47.6662575494403, -122.395482297078 47.6662374655706, -122.395511564436 47.6661814089142, -122.395580500711 47.6660608844761, -122.395639755034 47.6659558738869, -122.395696400198 47.6658484569148, -122.395770194014 47.6657207534141, -122.395814140519 47.6656466724319, -122.395837925516 47.6655939472982, -122.395801046284 47.6655824137779, -122.395729887252 47.665561453444, -122.395570832898 47.6655197621848, -122.395426186168 47.6654847281626, -122.395205449697 47.6654334710868, -122.395101200575 47.6654088509248, -122.394958059481 47.6653732390149, -122.394867536756 47.6653495016551, -122.394823231201 47.665344110738, -122.394783768728 47.6653479077655, -122.394711747893 47.6653488950151, -122.394649838534 47.6653486724238, -122.394611634832 47.6653437119756, -122.394571254065 47.6653338542645, -122.394512428063 47.6653349178054, -122.394489224569 47.6653563579404, -122.394421250033 47.6653400233256, -122.394273061951 47.6653052936606, -122.394154612135 47.6652808672428, -122.394109115682 47.6652526992279, -122.394142815643 47.6651921692967, -122.394150283421 47.6651367117978, -122.394063403634 47.665115965855, -122.393934023035 47.6650823489977, -122.393773467102 47.6650412328498, -122.393637085721 47.6650110104485, -122.393577692627 47.664993186853, -122.393446811582 47.6649771563125, -122.393294488555 47.6649567491072, -122.393088231659 47.6649313388824, -122.39292273862 47.6649116685336, -122.392737956194 47.6648917050166, -122.392595402171 47.6648755763921, -122.392420804653 47.6648568438781, -122.392196290103 47.6648316823013, -122.392034348038 47.6648119620737, -122.391865820573 47.6647925886425, -122.391679468479 47.6647710170423, -122.391507873986 47.664750871118, -122.391337343463 47.6647323810584, -122.391230366599 47.6647182062387, -122.391190126095 47.6647129724218, -122.391156789332 47.6647010892321, -122.391116662995 47.6646827435971, -122.391090137256 47.6646441606977, -122.391132274605 47.6646449553595, -122.391252961814 47.6646595001083, -122.391423533134 47.6646793603173, -122.391584933773 47.6646979751453, -122.39174836492 47.6647166044343, -122.391919967094 47.66473700681, -122.392068598633 47.6647527961854, -122.392237607771 47.6647713059764, -122.392384277691 47.6647893494728, -122.392543119103 47.6648071837026, -122.392703498522 47.6648255536658, -122.392838967999 47.6648423365095, -122.393000910464 47.6648620555932, -122.393189301041 47.6648838972179, -122.393332362163 47.6649000181571, -122.393533391983 47.6649203151783, -122.393601740175 47.6649322322938, -122.393684023226 47.6649519707042, -122.393836414988 47.6649915706156, -122.393974481457 47.6650272109698, -122.39413602018 47.6650672421714, -122.39427651762 47.6650993355217, -122.394383548187 47.6651321442783, -122.394521615404 47.6651677839692, -122.394626878068 47.6651923904079, -122.394748535505 47.6652222564645, -122.394858520011 47.6652520251458, -122.394941932347 47.665275560252, -122.395007135528 47.6653009728056, -122.395042412106 47.6653098291457, -122.395098606394 47.6653225118381, -122.395187899967 47.6653391110931, -122.395185307563 47.6653034998161, -122.395182532559 47.6652786879627, -122.395179178089 47.6652514850901, -122.395162586765 47.6652059118399, -122.395194553866 47.6651890640388, -122.395224023521 47.6651735357711, -122.395290858212 47.6651350445393, -122.395204106523 47.6651017012969, -122.395119909873 47.6650858468054, -122.395046830492 47.6650684681901, -122.39491744166 47.6650345956364, -122.394787129257 47.6650037343916, -122.394669611865 47.6649765538592, -122.394623315925 47.6649555521337, -122.394630722072 47.664931843196, -122.39466863731 47.6649272103267, -122.394781561663 47.6649366305867, -122.394908810353 47.6649499675211, -122.395034604815 47.6649655519374, -122.395104542044 47.6649796744066, -122.395233389999 47.6650124404972, -122.395414408425 47.6650590578252, -122.395551871431 47.6650914481458, -122.395691469559 47.6651273655221, -122.395839806131 47.6651670184029, -122.396042201092 47.6652157840011, -122.396171549386 47.6652482849121, -122.396296251292 47.6652781075585, -122.396459688551 47.6653136965464, -122.396570246818 47.6653456409346, -122.396718427527 47.6653800676452, -122.396855483979 47.665415761231, -122.396963980581 47.6654466195237, -122.397073450918 47.6654760934093, -122.397141178563 47.6654842038667, -122.397149422473 47.6654377757309, -122.397177066384 47.6654121605609, -122.397138272521 47.6654044671828, -122.397102876745 47.6653747903804, -122.397134701875 47.6653195124396, -122.397163384799 47.6652609787434, -122.397222202334 47.6651584582248, -122.397276785874 47.665050125973, -122.397423157647 47.665075022853, -122.397399749934 47.6651233727284, -122.39737971908 47.6651658927397, -122.397359414468 47.6651993330766, -122.397339259321 47.6652377416216, -122.397319095638 47.6652758501897, -122.397296012182 47.6653181120077, -122.397274685418 47.6653513094429, -122.397295673685 47.6653743286203, -122.397342829988 47.6653901334534, -122.397400122238 47.6654055421706, -122.397475746638 47.6654231409141, -122.397556484426 47.6654420406776, -122.397623962211 47.665458680349, -122.397699669262 47.665479020038, -122.397774279557 47.6654966325135, -122.397832552849 47.6655109135237, -122.397880558751 47.6655212223654, -122.397927550357 47.6655315451191, -122.397966319421 47.6655384246432, -122.397999728422 47.6655694994146, -122.39802676278 47.6655910644734, -122.398084211448 47.6656116547174, -122.398134246127 47.6656219355846, -122.398174556359 47.6656294365165, -122.398212799301 47.6656356808877, -122.398278140071 47.6656487502093, -122.398360834154 47.665665180397, -122.398431287859 47.665679550672, -122.398513964754 47.6656954239765, -122.398583380244 47.6657089944956, -122.398645670001 47.6657218489658, -122.398687293904 47.6657561095398, -122.398612765463 47.665741238405, -122.398356595061 47.6656929164311, -122.398235140748 47.6656699064362, -122.398104491423 47.6656448377607, -122.3980167349 47.6656287770187, -122.397948259732 47.6656127082847, -122.397885804754 47.6655943716031, -122.39782650029 47.6655795481232, -122.397781530071 47.6655689402946, -122.397723263948 47.6655549164294, -122.397678260268 47.6655431953524, -122.397616894538 47.6655273285234, -122.397555520348 47.6655111620389, -122.397495259751 47.6654982794665, -122.397433836139 47.6654804853879, -122.397378589189 47.6654656059629, -122.397333561173 47.6654530708063, -122.397278355512 47.6654395617044, -122.397239495894 47.6654296840457, -122.397190129026 47.6654415869513, -122.397181717127 47.665499285341, -122.397245120336 47.6655154241672, -122.397316687075 47.6655330790791, -122.397409730533 47.6655562228905, -122.397492513889 47.6655756516231, -122.397578422737 47.6655977792514, -122.397664290935 47.665618536436, -122.39772258882 47.6656336310949, -122.39777682947 47.6656487814479, -122.397851432305 47.6656661370596, -122.397910720126 47.6656804040982, -122.397981183276 47.6656950746105, -122.39804053691 47.6657115255453, -122.398085524056 47.6657226896999, -122.398160185648 47.6657419726038, -122.398198014074 47.6657513071111, -122.398264560144 47.6657707010876, -122.398336103604 47.6657875417595, -122.398407605837 47.6658030116642, -122.398480171431 47.6658200947712, -122.398531336782 47.6658342161785, -122.398603977261 47.6658537831363, -122.398692287664 47.6658882164346, -122.39877946725 47.66591880896, -122.398835952565 47.6659410831018, -122.398867753365 47.6659524281792, -122.398911932974 47.6659704585057, -122.399006965293 47.6659922027588, -122.399079482332 47.6660076582998, -122.399163281345 47.6660270715246, -122.399245966763 47.6660432014615, -122.39934098258 47.6660643886569, -122.399419676086 47.6660827578667, -122.399488136071 47.6660982689299, -122.399581040329 47.6661167430826, -122.399672938066 47.6661354879204, -122.399775013722 47.6661552063895, -122.399864882943 47.66617397897, -122.399947619147 47.6661917783836, -122.400028293199 47.6662084924321, -122.400089635415 47.6662235440262, -122.40013665262 47.6662346794603, -122.400172438128 47.6662435272319, -122.400203530533 47.6662650359713, -122.400199574332 47.6663021081989, -122.400159847648 47.6662971705439, -122.400125135044 47.666290235959, -122.400029162253 47.6662709905185, -122.399946434937 47.6662534911302, -122.399868770265 47.6662356216171, -122.39976559026 47.6662129192681, -122.39966444622 47.6661904457502, -122.399572531612 47.6661711441265, -122.399477549058 47.6661510706823, -122.399381494707 47.6661290839471, -122.399274256775 47.6661064365975, -122.399172099348 47.6660839765795, -122.399061843144 47.666062184803, -122.399020912783 47.6660509649151, -122.398963990868 47.6660478914062, -122.398916850764 47.6660326440166, -122.39882696536 47.666013313909, -122.398762522226 47.6659963763111, -122.398690961584 47.6659789794789, -122.398613307306 47.6659614090685, -122.398548839807 47.6659436577576, -122.398501814173 47.665932221702, -122.398459845214 47.6659202019576, -122.398420572908 47.6659303388334, -122.398360483097 47.6659568711198, -122.398348008176 47.6659806927817, -122.398288513526 47.6659932923167, -122.398303533746 47.6660371300684, -122.398320785868 47.6660877069077, -122.398361478046 47.6661247222789, -122.398396248342 47.6661335849026, -122.398431092144 47.6661448882923, -122.398493440936 47.6661596699889, -122.39856190796 47.6661754384288, -122.398684601054 47.666205843122, -122.398793076407 47.6662358860725, -122.398907679134 47.6662672156447, -122.399026356089 47.6662990461122, -122.399149222623 47.6663351889532, -122.39924339749 47.6663621288211, -122.39932849276 47.666390864466, -122.399431921207 47.666421790131, -122.399521931429 47.6664452312246, -122.399587472287 47.6664648951874, -122.399670365025 47.6664878770214, -122.399712450875 47.6665037506191, -122.399744244471 47.6665148386387, -122.399740807532 47.6665691699517, -122.399739268132 47.6666191050841, -122.399739055292 47.6666457146117, -122.399792844176 47.6666795074254, -122.399849372052 47.6667031514349, -122.399940892961 47.6667430237498, -122.400033394773 47.6667817681315, -122.400128141528 47.6668276370482, -122.400275810269 47.6668785189333, -122.400401302962 47.6669343755368, -122.400542819837 47.6669831564955, -122.400688345804 47.6670302542602, -122.400786334216 47.6670826329956, -122.400912759638 47.6671357343367, -122.400982205724 47.667183878683, -122.401043987268 47.6672134481104, -122.40117032336 47.6672635512539, -122.401287627703 47.6673170346491, -122.401420841206 47.6673596307978, -122.401540018899 47.6674079041331, -122.401615187632 47.6674438870858, -122.401685123866 47.667440996262, -122.401763005473 47.6674659731992, -122.401847320935 47.6675023870714, -122.401937968462 47.6675636492733, -122.402004341665 47.6678461950895, -122.401941481828 47.6680499727374, -122.401012947588 47.6676816094831, -122.3983273256 47.6674845638036, -122.398337852112 47.6681206608513, -122.398327335285 47.668563432689, -122.398305959458 47.6694545076039, -122.398302731334 47.6695663760642, -122.398337579049 47.671094419885, -122.398343274343 47.6720250537468, -122.398373497892 47.6740903885558, -122.39836900529 47.6758286109207, -122.398264017499 47.6758644570828, -122.395673683783 47.6758796144233, -122.392956550545 47.6759141859541, -122.392827172066 47.6759153143353, -122.391773109273 47.6759158131504, -122.390449346263 47.6759259868868, -122.388955759168 47.6759424488678, -122.387702391216 47.6759476875929, -122.387087434206 47.6759521343769, -122.385435721425 47.6759587956618, -122.383602752477 47.675979810424, -122.38182788443 47.675992079457, -122.379988625303 47.6759893429385, -122.377903057631 47.6759978849021, -122.376707907517 47.6759982290459, -122.376336564723 47.6759176989664, -122.376214150095 47.6757431382831, -122.376206970115 47.6747491618888, -122.376206910876 47.6736189728118, -122.376208757484 47.6725529816933, -122.376215401015 47.6714605332274, -122.376216126032 47.6703567685299, -122.376212003553 47.6692606527246, -122.376218981706 47.6681795106282))\"]]]} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :properties]] {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OBJECTID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 29]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :L_HOOD]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"Ballard\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :S_HOOD]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"Whittier Heights\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :S_HOOD_ALT_NAMES]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :Shape__Area]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 1.41956874919434E7]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :Shape__Length]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 15934.4392663199]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geometry]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector #object[org.locationtech.jts.geom.Polygon 0x11cff25b \"POLYGON ((-122.376336564723 47.6759176989664, -122.376519722058 47.6760527482053, -122.376547112434 47.6760693435982, -122.376573006221 47.6760868163511, -122.37659791092 47.6761051588887, -122.376620811506 47.6761243856628, -122.376653693551 47.6761550023704, -122.376672575492 47.6761756116819, -122.376690473449 47.6761972626929, -122.376706358898 47.6762194979699, -122.376721246367 47.6762423037704, -122.376733608705 47.6762655291531, -122.376744969247 47.6762891966531, -122.376753802121 47.6763131984825, -122.376761120573 47.6763374777451, -122.37676692258 47.6763619488292, -122.376770700102 47.6763866189683, -122.376772451106 47.6764114022001, -122.376772171763 47.6764361697665, -122.376761985459 47.676862137338, -122.376759750881 47.6775387646826, -122.376769882216 47.6780678404081, -122.376770065639 47.6792017979916, -122.376770442503 47.680154307228, -122.376767820487 47.6811938310584, -122.376777423242 47.6822860591882, -122.376787310842 47.6833707854178, -122.376780535885 47.6842704786057, -122.376785260609 47.6851813698577, -122.376784403749 47.6860923366759, -122.376800402808 47.6870068457338, -122.376794751014 47.6879443550321, -122.376803067488 47.688976146724, -122.376810989127 47.6900117143006, -122.376810264834 47.6905341242701, -122.376810344766 47.6906051591463, -122.375110107987 47.6906173771311, -122.373230534841 47.6906248812211, -122.37140439251 47.6906367315375, -122.369524666912 47.6906390778747, -122.367607389199 47.6906418559576, -122.366004446548 47.6906481724782, -122.366003399739 47.6904922338054, -122.36599385014 47.6887402933599, -122.366005835095 47.6862900886837, -122.36599586678 47.6845239294054, -122.366032959089 47.6832519011367, -122.366038176946 47.6830846527711, -122.366006789592 47.6805923152563, -122.366049650322 47.6763326418883, -122.366053675503 47.6760045296047, -122.366299114137 47.6760152803027, -122.367446257021 47.6759532251209, -122.368513687945 47.6759415886584, -122.369529736837 47.6759772497078, -122.370612103444 47.6759215200542, -122.371140989063 47.6759582978223, -122.371714110737 47.6759807230408, -122.373288186248 47.6759402633333, -122.375585947496 47.6759338606538, -122.376080171672 47.6758970068068, -122.376336564723 47.6759176989664))\"]]]} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :properties]] {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OBJECTID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 30]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :L_HOOD]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"Ballard\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :S_HOOD]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"West Woodland\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :S_HOOD_ALT_NAMES]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :Shape__Area]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 2.21993668730164E7]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :Shape__Length]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 21789.6510873493]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geometry]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector #object[org.locationtech.jts.geom.Polygon 0x41b6d8d1 \"POLYGON ((-122.376336564723 47.6759176989664, -122.376080171672 47.6758970068068, -122.375585947496 47.6759338606538, -122.373288186248 47.6759402633333, -122.371714110737 47.6759807230408, -122.371140989063 47.6759582978223, -122.370612103444 47.6759215200542, -122.369529736837 47.6759772497078, -122.368513687945 47.6759415886584, -122.367446257021 47.6759532251209, -122.366299114137 47.6760152803027, -122.366053675503 47.6760045296047, -122.365227601151 47.6759917757154, -122.364351081075 47.6759893337515, -122.363647800755 47.6759930842157, -122.36271708327 47.6759941851371, -122.362245131082 47.6759948642237, -122.36094806215 47.6759980385768, -122.360691453297 47.6760043068067, -122.36077646407 47.6758922436911, -122.360771594157 47.6752948902842, -122.360775562234 47.6745665726628, -122.360777556465 47.6739264976242, -122.360786418032 47.673226562611, -122.360788381154 47.6724300211048, -122.360793230899 47.671593595961, -122.360789714591 47.6707487566245, -122.360797202559 47.66986396762, -122.36080519647 47.6689791289769, -122.360803810493 47.6680859329074, -122.360802846427 47.6671898180224, -122.360802670245 47.6669075629066, -122.360801382733 47.6662767003006, -122.360803809514 47.6653578322292, -122.360802011097 47.6644333226362, -122.360799877982 47.6634974208525, -122.360790125255 47.6626647898188, -122.360789123417 47.6625616346537, -122.360789702508 47.662374140783, -122.36079141909 47.6614656503796, -122.36214579117 47.6614566051262, -122.363439573738 47.661457526372, -122.364137181521 47.6614573212743, -122.366155922999 47.6614736585099, -122.366120224218 47.6593510369868, -122.365384555672 47.6584891840664, -122.36504049713 47.6581417180235, -122.365039920817 47.6577605839202, -122.365898909139 47.6571615480529, -122.36609426973 47.6570254164477, -122.366796991946 47.6565298464815, -122.366826844662 47.6565443969658, -122.366857901565 47.6565653580207, -122.366891472649 47.6565855569296, -122.366931099591 47.6566046887304, -122.366984914141 47.6566404245544, -122.36696633103 47.6566639819699, -122.366941865149 47.6566944739385, -122.366945853884 47.6567265106697, -122.366920737647 47.656769350089, -122.366914591904 47.6568017802057, -122.366910524252 47.6568358533544, -122.366929434489 47.6568577922272, -122.366954358163 47.6568772076512, -122.366989609651 47.6569027823278, -122.3670234701 47.6568812040272, -122.367058714568 47.6569065218601, -122.367084805348 47.6569311055126, -122.367136503061 47.6569638706548, -122.367176745134 47.6569866362092, -122.367208199216 47.657021044885, -122.367256821766 47.6570527373811, -122.367316047732 47.6570999253012, -122.36736372904 47.657134115369, -122.367407280992 47.6571658761019, -122.367396828116 47.6571898810789, -122.367342953188 47.6571864936152, -122.367260663289 47.6571829321847, -122.36712573072 47.6571833781758, -122.367123844393 47.6572226062806, -122.367068701234 47.657245027876, -122.367084706917 47.6572716756283, -122.367124722507 47.6572867749918, -122.367208951699 47.6572872688121, -122.367317651982 47.657291845841, -122.367357287779 47.6572940538329, -122.367410165606 47.6572980117273, -122.367491586888 47.6573065121605, -122.367593130693 47.6573092570451, -122.36765110338 47.6573139601153, -122.367750869179 47.6573252123496, -122.367747961565 47.6573641969622, -122.367744851179 47.6573963293169, -122.367735882061 47.6574362095165, -122.367731058077 47.6574790329932, -122.367936013957 47.6574790133695, -122.36811967029 47.6574792807636, -122.368143685487 47.657502264309, -122.368179766803 47.6575215294621, -122.368221128874 47.6575478353438, -122.368257128799 47.6575643593633, -122.368282158263 47.6575873292113, -122.368331017204 47.6576269870795, -122.368369279504 47.6576514067545, -122.36839535663 47.6576754764687, -122.368436678466 47.6577004118894, -122.368470690691 47.6577183335634, -122.368499655604 47.6577371375356, -122.368533877812 47.6577621682691, -122.368560912329 47.6577842971086, -122.368597146509 47.6578087440486, -122.368627142653 47.6578280910938, -122.368654079328 47.6578469220186, -122.368707076446 47.6578892661541, -122.368734199128 47.6579143928902, -122.368793128141 47.6579514298714, -122.36882312391 47.657970776521, -122.36885318474 47.6579923074877, -122.368876064888 47.6580111934047, -122.368919610383 47.6580426963992, -122.368944478452 47.6580601845718, -122.36898805566 47.6580927583049, -122.369025330251 47.6581180478279, -122.36905969088 47.6581477470354, -122.369033975504 47.6581702869537, -122.36900709337 47.6581876586224, -122.368999691683 47.65821187988, -122.369018464588 47.6582634682636, -122.369043721192 47.6582941041555, -122.36907469147 47.6583120666742, -122.369109790614 47.6583324156284, -122.369156891474 47.6583468617493, -122.369212098743 47.6583609412648, -122.369318009446 47.6583740372247, -122.369358310498 47.6583987288221, -122.369406708617 47.6584227545456, -122.369440827205 47.6584442306362, -122.369473898213 47.6584646068355, -122.369514255849 47.6584912259011, -122.369572211863 47.658529646547, -122.36960323056 47.6585492364699, -122.369632132312 47.658565855815, -122.369664043734 47.6585813205437, -122.369695987565 47.6585978988375, -122.369736191513 47.6586192927535, -122.369776192963 47.6586338340435, -122.369816583052 47.6586615237593, -122.369826797328 47.6586981462173, -122.369867901674 47.6587156720736, -122.369905201188 47.6587417749067, -122.369942483773 47.6587673213019, -122.369968447737 47.6587875363313, -122.36999861459 47.6588126217082, -122.370029609556 47.6588313975654, -122.370067898867 47.6588566734205, -122.370117927743 47.6589014987462, -122.370168841769 47.658941942404, -122.370195999136 47.658968182, -122.370216908268 47.6589890221901, -122.370246158652 47.6590174187339, -122.370277517982 47.6590485290195, -122.370311839933 47.6590768571086, -122.370336830517 47.659098456122, -122.370367021263 47.6591243550065, -122.37039624775 47.6591519379133, -122.370416312519 47.6591785304467, -122.37044639117 47.6592006177754, -122.370473353282 47.6592202622183, -122.370502724993 47.6592527701992, -122.370517288468 47.6592991883743, -122.370540243237 47.6593205575537, -122.370541017091 47.6594496797416, -122.370642889997 47.6594633860064, -122.370668935633 47.6594863416389, -122.370703095449 47.6595091877392, -122.370724046049 47.6595313982104, -122.370741182515 47.6595618864373, -122.370756054473 47.659584435873, -122.370772022238 47.6596097127233, -122.370788980152 47.6596341619315, -122.370825412255 47.6596652035978, -122.370858696122 47.6596927316213, -122.370910463566 47.659727679378, -122.370957379969 47.6597358289275, -122.370966433314 47.6598017724136, -122.370995075217 47.6598438873314, -122.37109454691 47.6598793049694, -122.371137413708 47.6598877660403, -122.371168305238 47.6598687118902, -122.371211132648 47.6598415272093, -122.37124000278 47.6598227572507, -122.371275908778 47.6598017070751, -122.371311854781 47.6597820276275, -122.371350762806 47.6597595659607, -122.37142619989 47.6598057186435, -122.371463427483 47.6598293373683, -122.371491518562 47.6598528222021, -122.371540422629 47.6598938490718, -122.37156744254 47.6599154204198, -122.371596499027 47.6599372215243, -122.371639131162 47.6599720350715, -122.371672301585 47.6599957081771, -122.371750903628 47.6600459313941, -122.371792245425 47.6600714220276, -122.371820280267 47.6600929796089, -122.371851340784 47.6601139393066, -122.371913698707 47.6601638248364, -122.371986215918 47.660214129744, -122.372040239984 47.6602567154385, -122.372114688276 47.6603036953151, -122.372164590997 47.6603441514304, -122.372243445585 47.6604028537521, -122.372274482371 47.6604229997362, -122.372276324586 47.660485227525, -122.37227074266 47.6605365874913, -122.37225504101 47.6605888984651, -122.372252063055 47.6606253992408, -122.372300935393 47.660631036882, -122.372337587815 47.660635211775, -122.37240070095 47.6606420278766, -122.372440477734 47.6606489023762, -122.372484156738 47.6606504973298, -122.372563572134 47.6606595783311, -122.372624730427 47.6606689056685, -122.37273364879 47.6606805869493, -122.372806816049 47.6606842681617, -122.372842543097 47.6606914541774, -122.37289961108 47.660699722995, -122.37296059064 47.6607030118459, -122.373007484147 47.6607103469529, -122.373209839634 47.660724877436, -122.373245437373 47.6607276953685, -122.373330945994 47.6607369503739, -122.373385815625 47.6607395073355, -122.373419895432 47.6607253365243, -122.373466983151 47.660739223728, -122.373496885087 47.6607552715937, -122.373551113066 47.6607704334901, -122.373591847227 47.6607753667089, -122.373628337943 47.6607740593064, -122.373632702775 47.660750136064, -122.373666450031 47.6607247442675, -122.373705682622 47.6607132455431, -122.373771253658 47.6607002765727, -122.373804034818 47.6606765259987, -122.373804998654 47.6606063340659, -122.373874968362 47.6606048306868, -122.374082995739 47.6606053156307, -122.374297091976 47.6606051614314, -122.374502100296 47.6606065007312, -122.374690879977 47.6606080589335, -122.374897876649 47.6606079996829, -122.375106909387 47.6606081694125, -122.375206342681 47.6606079373933, -122.375281407609 47.6606071780974, -122.375395605626 47.6606086739619, -122.37542406779 47.6606274826827, -122.375423424091 47.6606570969644, -122.375407829398 47.6606786873881, -122.375378139928 47.6607040247744, -122.375332531732 47.6607400316276, -122.375302401399 47.6607676029867, -122.375283362005 47.6607927961628, -122.3752817205 47.660822980606, -122.375298776239 47.6608677362913, -122.375322319322 47.6609088480108, -122.375367789142 47.6609536903689, -122.375414915253 47.6609859567612, -122.375480757744 47.6610163415218, -122.375549698916 47.661048569405, -122.375616507521 47.6610773131594, -122.37567132516 47.6610952082863, -122.375713866021 47.6611097131508, -122.375766573924 47.6611248945907, -122.375816044728 47.6611335649309, -122.375873685334 47.6611440093711, -122.375928012797 47.6611624680215, -122.375974541099 47.6611574674372, -122.375994709938 47.6611190630255, -122.376028531523 47.6610791024086, -122.376045764347 47.6610443363587, -122.376053339804 47.6609918779637, -122.376029447109 47.6609389890523, -122.376000962688 47.6609023572787, -122.376035494866 47.6608692423874, -122.376094954205 47.6608554981113, -122.376142250917 47.6608593131576, -122.376170357689 47.660900320073, -122.37622552108 47.6609298638619, -122.376227046633 47.6609641613863, -122.376228293294 47.6609890799218, -122.376232028288 47.6610807160121, -122.376226087158 47.6616326743118, -122.376215127819 47.6622035949171, -122.376213335479 47.6628953406738, -122.376224015951 47.6638363569551, -122.376218608455 47.6649703907559, -122.376222341407 47.6660363069762, -122.376214914328 47.6671023311431, -122.376218981706 47.6681795106282, -122.376212003553 47.6692606527246, -122.376216126032 47.6703567685299, -122.376215401015 47.6714605332274, -122.376208757484 47.6725529816933, -122.376206910876 47.6736189728118, -122.376206970115 47.6747491618888, -122.376214150095 47.6757431382831, -122.376336564723 47.6759176989664))\"]]]} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :properties]] {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OBJECTID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 31]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :L_HOOD]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"North Central\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :S_HOOD]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"Phinney Ridge\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :S_HOOD_ALT_NAMES]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"Woodland Park\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :Shape__Area]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 3.21231178777466E7]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :Shape__Length]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 27120.7859379681]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geometry]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector #object[org.locationtech.jts.geom.Polygon 0x3fbfe64b \"POLYGON ((-122.347307566662 47.6650052913031, -122.348867895321 47.6650090912859, -122.349792869513 47.6649918965054, -122.349989420364 47.6650332375445, -122.350304163556 47.6650388147716, -122.351421221038 47.6650141642968, -122.352495514378 47.6650193785027, -122.353450405799 47.6650188929375, -122.354365227588 47.6650188912667, -122.355056521311 47.6650267528834, -122.355414376827 47.6650317400371, -122.355416519216 47.6643598303278, -122.355424977054 47.6636610158142, -122.355424140879 47.6630086831036, -122.355427805224 47.6625273664197, -122.355436677765 47.662206602761, -122.355438314261 47.6621412005632, -122.355573420679 47.6621809112223, -122.355864863892 47.6622576482863, -122.356113413965 47.6623422844097, -122.356200383141 47.6623840944536, -122.356311534565 47.6624373631027, -122.356476832915 47.6625377642513, -122.356569133827 47.6625883712575, -122.357013669094 47.6628335347683, -122.357378460614 47.6630582987986, -122.357703834377 47.6633056110634, -122.358267789163 47.6637988671757, -122.359051229737 47.664492005053, -122.359500905186 47.6648939854014, -122.359642871567 47.6650460222814, -122.3598308178 47.6652951273317, -122.360257228516 47.6660101362634, -122.360375131079 47.6662064525472, -122.360743692216 47.666817181635, -122.360802670245 47.6669075629066, -122.360802846427 47.6671898180224, -122.360803810493 47.6680859329074, -122.36080519647 47.6689791289769, -122.360797202559 47.66986396762, -122.360789714591 47.6707487566245, -122.360793230899 47.671593595961, -122.360788381154 47.6724300211048, -122.360786418032 47.673226562611, -122.360777556465 47.6739264976242, -122.360775562234 47.6745665726628, -122.360771594157 47.6752948902842, -122.36077646407 47.6758922436911, -122.360691453297 47.6760043068067, -122.36094806215 47.6759980385768, -122.362245131082 47.6759948642237, -122.36271708327 47.6759941851371, -122.363647800755 47.6759930842157, -122.364351081075 47.6759893337515, -122.365227601151 47.6759917757154, -122.366053675503 47.6760045296047, -122.366049650322 47.6763326418883, -122.366006789592 47.6805923152563, -122.366038176946 47.6830846527711, -122.366032959089 47.6832519011367, -122.365072292044 47.6832400978159, -122.360853796868 47.6832540116804, -122.356720447386 47.6832808551648, -122.352902763328 47.6833062957672, -122.350224039798 47.6832962725287, -122.346173568009 47.683293139086, -122.344486258114 47.6833050746089, -122.34448184051 47.682507422038, -122.344493201644 47.6813797044772, -122.344498539691 47.6813538417466, -122.344504896624 47.6813280939604, -122.344511764863 47.6813024678503, -122.344520157279 47.6812769071635, -122.34452906298 47.6812515541157, -122.344538988795 47.6812263584634, -122.34454993299 47.6812012784631, -122.344561392211 47.6811764481954, -122.344574379345 47.6811518121086, -122.344587880268 47.6811273833018, -122.344613552414 47.6810851841658, -122.344629611324 47.6810614500549, -122.344646692321 47.6810379593003, -122.344664290317 47.6810148042395, -122.344682911652 47.6809919356879, -122.344702049261 47.6809693600197, -122.344982950426 47.6806457607991, -122.345002598574 47.6806233071122, -122.345021223875 47.6806005669062, -122.34503984117 47.6805775698876, -122.345057435099 47.6805542863574, -122.345074008684 47.6805308022656, -122.345090066683 47.6805070677453, -122.345105103095 47.6804830898613, -122.345119626422 47.6804589475059, -122.345132679663 47.6804365809458, -122.345145671323 47.6804121160041, -122.345159168049 47.6803875587207, -122.345171645889 47.6803628864969, -122.345183615423 47.680338178197, -122.345195581199 47.6803133411366, -122.345207037435 47.6802884255483, -122.34521798359 47.6802634307376, -122.345228927245 47.6802383503202, -122.345239360831 47.6802131910313, -122.345249286114 47.6801879956669, -122.345258700074 47.6801626782776, -122.345267605743 47.6801373251639, -122.345276510157 47.6801119292468, -122.345284904496 47.6800864541079, -122.34529278877 47.6800609000981, -122.345300164744 47.6800353100135, -122.345307539466 47.6800096771256, -122.345314404644 47.6799839653602, -122.34532076225 47.6799582603297, -122.345326609795 47.6799324764289, -122.345331949043 47.6799066564535, -122.345337286522 47.679880793682, -122.345342116214 47.6798548944784, -122.345346436583 47.6798289595651, -122.345350249894 47.6798030310293, -122.345388766947 47.6794348332565, -122.345489574102 47.678591097451, -122.34550729821 47.6783281431494, -122.345510616247 47.6783026497212, -122.345521037373 47.6782770623701, -122.345527907924 47.6782515646106, -122.345534778989 47.6782260668437, -122.345542157085 47.6782005626977, -122.345549536409 47.6781751010031, -122.345556915727 47.678149639308, -122.345564296292 47.6781242207661, -122.345572183868 47.678098795143, -122.345580071965 47.6780733698634, -122.345587960771 47.678047987042, -122.345595850098 47.678022604564, -122.345603740134 47.6779972645442, -122.345612138237 47.6779719181311, -122.345620537059 47.6779466145269, -122.345628935861 47.6779213105713, -122.345637334666 47.6778960069659, -122.345646242253 47.6778707394257, -122.345654642277 47.6778454782708, -122.345663550572 47.6778202535388, -122.345672966923 47.6777950220623, -122.345681876447 47.6777698401315, -122.345691294025 47.6777446514562, -122.34570071232 47.6777195055896, -122.345710130596 47.6776943593714, -122.345719550118 47.677669256306, -122.345729477184 47.6776441468533, -122.345738897934 47.677619086589, -122.345748826217 47.6775940195866, -122.345759262031 47.6775689458458, -122.345769192266 47.6775439644539, -122.345779629306 47.677518933514, -122.34590637354 47.6773574430743, -122.346596886857 47.6764789695799, -122.346617048988 47.6764568084887, -122.346636187836 47.6764343612505, -122.346654809675 47.6764115779707, -122.346672408949 47.6763885510035, -122.3466889862 47.6763652810441, -122.346705555936 47.6763417535642, -122.346720594874 47.6763179470316, -122.346735120546 47.6762939328673, -122.346749132953 47.6762697110715, -122.346761616528 47.6762452954857, -122.346774094379 47.676220665878, -122.346785043921 47.6761958424742, -122.346795482175 47.6761708970522, -122.346804899651 47.6761457510927, -122.34681380637 47.6761204834594, -122.346821694796 47.6760951005437, -122.346828564421 47.6760696027034, -122.346834923303 47.6760439835405, -122.346840265133 47.6760182915476, -122.346844589422 47.6759925277843, -122.346848405454 47.6759667279532, -122.346850697451 47.6759408627296, -122.346852986424 47.6759149115564, -122.34685325133 47.6758891589, -122.346853447502 47.6758436130604, -122.346851677106 47.6758176729946, -122.346849400445 47.6757917828246, -122.346846618018 47.6757659418417, -122.346842313553 47.6757401214293, -122.346837504581 47.6757143933574, -122.346831683051 47.6756887640242, -122.346824849504 47.6756632341244, -122.346817510208 47.675637753762, -122.34680916087 47.675612458445, -122.346800307021 47.675587255117, -122.346789935404 47.6755622007595, -122.346779060742 47.6755373243608, -122.346767684594 47.6755126259001, -122.34675479089 47.6754881192253, -122.346741395172 47.6754637901441, -122.346731081703 47.6754407485692, -122.346719703106 47.675415964498, -122.346707818984 47.6753912731306, -122.346695937874 47.6753666670096, -122.346683549257 47.6753420676288, -122.346671164401 47.6753175970057, -122.346658273283 47.6752931759253, -122.346644876414 47.6752688040298, -122.346631482061 47.675244518089, -122.346617583215 47.6752203244863, -122.346603685108 47.6751961736916, -122.346589282508 47.6751721152348, -122.34657437366 47.6751481066708, -122.346559468555 47.6751242261621, -122.346544565239 47.6751003887981, -122.346528649385 47.6750766501685, -122.346513243046 47.675052990759, -122.34649682522 47.6750294304204, -122.346480916897 47.6750059489513, -122.346463996579 47.6749825669103, -122.346447078758 47.6749592701215, -122.346430163445 47.6749360589356, -122.346412743654 47.6749129404369, -122.346394818336 47.6748899142881, -122.346376896045 47.674866973735, -122.346358977509 47.6748441615873, -122.346340046971 47.6748214485152, -122.3463216247 47.6747987718606, -122.346302700426 47.6747762727961, -122.34628326991 47.6747538236216, -122.346263336154 47.6747315095848, -122.346243913149 47.67470931722, -122.346223478154 47.6746872242802, -122.346203045659 47.6746652165915, -122.34618261693 47.6746433376581, -122.346161683706 47.6746215507078, -122.346140753729 47.6745998925198, -122.346119319278 47.6745783270162, -122.34609738158 47.674556896298, -122.346075954121 47.6745355876097, -122.346053516422 47.6745144204382, -122.346031081244 47.6744933392187, -122.346008649822 47.6744723864029, -122.345985713408 47.6744515262775, -122.345962274995 47.6744308437389, -122.345939346093 47.6744102404205, -122.345915407691 47.6743898217777, -122.345891471799 47.6743694887354, -122.345867539664 47.6743492840963, -122.345843104304 47.6743292149421, -122.345818672689 47.6743092738401, -122.345793737331 47.6742894682293, -122.345768806985 47.674269834175, -122.34574337163 47.6742502921071, -122.345717940041 47.6742308787927, -122.345692006464 47.6742116434137, -122.345666076642 47.674192536437, -122.345640150577 47.6741735578627, -122.344897905857 47.6736916881744, -122.344393544584 47.6731878075112, -122.344369438002 47.673161563858, -122.344350015147 47.6731393283662, -122.344330081596 47.6731169711869, -122.344311160051 47.6730945149437, -122.344292234788 47.6730719302889, -122.344273812777 47.673049210148, -122.344255388293 47.6730264043986, -122.344237468833 47.67300350631, -122.344219545657 47.6729804798103, -122.34420212574 47.6729573181759, -122.344185212081 47.6729341066548, -122.344168294704 47.6729107667229, -122.34415188236 47.6728873348036, -122.344135974508 47.6728638102023, -122.344120062938 47.6728401571904, -122.344104148884 47.6728164182198, -122.3440892486 47.6727926233352, -122.344074344068 47.6727686996965, -122.344059438315 47.6727447336037, -122.344045543303 47.6727206252973, -122.344031647579 47.6726964741796, -122.344017747616 47.6726721946586, -122.344004353905 47.6726478652528, -122.343991465222 47.6726234438612, -122.343979082281 47.6725989729427, -122.343966696853 47.6725744160666, -122.343954308948 47.6725497735835, -122.343942934287 47.6725250751955, -122.343931557139 47.6725002908499, -122.343920177513 47.6724754208975, -122.343909812881 47.6724505374858, -122.343899443997 47.6724255253209, -122.343889075134 47.6724005135058, -122.343879718775 47.6723754026261, -122.343870360651 47.672350248599, -122.343861001821 47.6723250521123, -122.343852655688 47.6722997990272, -122.343843800105 47.6722744670578, -122.34383595899 47.672249121637, -122.343828115396 47.6722236906101, -122.343820779308 47.6721982532048, -122.343813947688 47.6721727227699, -122.343807116606 47.6721471926785, -122.343800790735 47.672121613069, -122.343794464146 47.6720959906495, -122.34378915202 47.6720703547793, -122.343783837412 47.6720446333033, -122.343778524042 47.672018954279, -122.343773715688 47.6719931832711, -122.343769414305 47.6719674055414, -122.343765619884 47.6719416207391, -122.343761824753 47.6719157934778, -122.343758536582 47.6718899591439, -122.343755754667 47.6718640756295, -122.343752972744 47.6718381917642, -122.343751739629 47.67181323009, -122.343751492865 47.6717872701443, -122.343749724879 47.6717613728357, -122.343746942965 47.6717354889702, -122.343743147858 47.671709661708, -122.343738847758 47.6716839267797, -122.343733028208 47.6716582972839, -122.343725689748 47.6716327739155, -122.343717847554 47.671607386034, -122.343708994642 47.671582139659, -122.343699133522 47.6715571210975, -122.343688261696 47.6715322443929, -122.343676381135 47.6715075951573, -122.343663492889 47.6714831737275, -122.34364959591 47.671458979766, -122.343634693207 47.6714350988712, -122.34361877909 47.6714113173727, -122.343603376406 47.6713877003691, -122.343586963543 47.6713642252133, -122.343570046946 47.6713408851911, -122.343552627868 47.6713177234559, -122.343535211803 47.6712946469658, -122.343517294501 47.6712717915651, -122.343498364753 47.6712490355727, -122.343479441749 47.6712264932332, -122.343460522503 47.6712040796497, -122.343440593288 47.6711818507283, -122.343420162116 47.6711598000854, -122.343399736445 47.6711379202922, -122.343378807777 47.671116218791, -122.343357377142 47.6710946952168, -122.343335445275 47.6710733927298, -122.343313010197 47.6710522253667, -122.343290582599 47.6710313148156, -122.343267144313 47.6710105461142, -122.343243712997 47.6709900345824, -122.343219778473 47.6709696581733, -122.343195851958 47.6709495389198, -122.343171423671 47.6709296400571, -122.343145986483 47.6709099265396, -122.343121062478 47.6708904199607, -122.343095131009 47.670871183995, -122.343069205052 47.6708521192276, -122.343042779628 47.6708333183402, -122.343015852443 47.6708147381926, -122.342988425058 47.6707963787639, -122.342961004632 47.6707782761521, -122.342933590462 47.6707603882493, -122.342905169328 47.6707427702483, -122.342876754441 47.6707253666049, -122.342847840075 47.6707082264886, -122.342818932678 47.6706913435395, -122.342789017086 47.6706746880384, -122.342759617689 47.6706583254248, -122.342729210107 47.6706421906094, -122.342698811245 47.6706263554054, -122.342668418626 47.670610734558, -122.342637527254 47.6705954200455, -122.342606136412 47.6705803694086, -122.342574245574 47.670565582303, -122.342542870937 47.6705510884359, -122.342510488818 47.6705368644717, -122.341536132237 47.6702371791038, -122.340924268323 47.6700484103442, -122.340890520367 47.6700395599196, -122.34085726302 47.6700301031724, -122.340824497725 47.6700201257224, -122.340792222002 47.6700095419644, -122.340759930663 47.669998401403, -122.340728637606 47.669986690622, -122.340697329137 47.6699744658547, -122.340666513773 47.6699617207234, -122.340636697406 47.669948447833, -122.340606864917 47.6699346184983, -122.340578033715 47.6699203045453, -122.340549187104 47.6699054766074, -122.34052134178 47.6698901640522, -122.340493989757 47.6698743736014, -122.340467132805 47.669858148402, -122.340440769155 47.6698414453077, -122.340415406792 47.6698242575984, -122.340390540223 47.6698066779514, -122.340366170679 47.6697887488192, -122.340340290972 47.6697712253839, -122.340315919499 47.6697532106285, -122.340288929275 47.6697324026858, -122.340264500721 47.6697124179372, -122.340241075418 47.6696920341884, -122.340217639676 47.6696713076683, -122.340195210196 47.6696502680984, -122.340173785216 47.6696288726834, -122.340152857771 47.6696071274278, -122.340132428613 47.669585076194, -122.340113006425 47.6695627543711, -122.340094083021 47.6695401258623, -122.340076164836 47.6695171843201, -122.340058747205 47.6694939792394, -122.340017468094 47.6694703611258, -122.340018094829 47.669439462271, -122.340018846127 47.6693953660748, -122.340021771514 47.669198716564, -122.340026409396 47.6687810557474, -122.340032268235 47.668755614529, -122.340037618902 47.6687301375634, -122.340042462098 47.6687046669585, -122.340047304048 47.6686791535506, -122.340051637817 47.6686536040447, -122.340055971591 47.6686280548895, -122.340059796656 47.6686024692924, -122.340063114263 47.6685768904071, -122.340066431144 47.668551268712, -122.340069747504 47.6685256470237, -122.340072555684 47.6684999892376, -122.340074856928 47.6684743381566, -122.34007715641 47.6684486442795, -122.340082471836 47.667914353136, -122.340086900669 47.6670869356763, -122.340087969201 47.6664061318327, -122.340099824506 47.6657298978586, -122.340202545541 47.6650376379612, -122.340449279385 47.6650425998776, -122.342697946464 47.6650225937497, -122.343218691111 47.6650108080465, -122.344520287628 47.6649984316009, -122.345091959432 47.6650105968045, -122.346067549327 47.6650414741391, -122.346093435108 47.6650243357179, -122.346125026488 47.66501149119, -122.346159298506 47.6650035812492, -122.346230329133 47.6650034517696, -122.346449839817 47.664996980288, -122.347307566662 47.6650052913031))\"]]]})")}
                , document.currentScript.parentElement);</script></div></div>

::: {.sourceClojure}
```clojure
(def parks-geojson
  (read-geogson-gz
   "data/Seattle/Park_Boundary_(details).geojson.gz"))
```
:::

::: {.sourceClojure}
```clojure
(count parks-geojson)
```
:::

::: {.printedClojure}
```clojure
2809

```
:::

::: {.sourceClojure}
```clojure
(->> parks-geojson
     (take 5)
     kind/portal)
```
:::

<div style="max-height:600px;overflow-y:auto;"><div><script>portal_api.embed().renderOutputItem(
                {'mime': 'x-application/edn',
                 'text': (() => "({^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :properties]] {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PMA]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 442]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SE_ANNO_CAD_DATA]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SDQL]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"QL-D1\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OBJECTID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 207882]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GIS_CRT_DT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2024-01-19T14:07:23Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PIN]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"4006000485\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ADDRESS]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \" \"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SHAPE_Area]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 1.4691943312522366E-6]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PMA_NAME]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"East Duwamish GS: S Chicago St\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :NAMEFLAG]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 9]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GIS_EDT_DT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2024-01-19T14:07:23Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SUBPARCEL]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 9851]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OWNER]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"DPR\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :NAME]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"EAST DUWAMISH GREENBELT\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :USE_]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GlobalID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"{D4B7025E-1135-4232-88AB-CBF72932E6AE}\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :AMWOID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"PROPERTY-EDUWSC\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :MAINT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"DPR\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :REVIEW_DATE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2004-04-08T00:00:00Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :LEASE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"N\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ACQ_DATE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"1997-10-21T00:00:00Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SHAPE_Length]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.005215887160444538]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geometry]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector #object[org.locationtech.jts.geom.MultiPolygon 0x40d3f919 \"MULTIPOLYGON (((-122.28235899499998 47.525165833000074, -122.28168484899999 47.525164389000054, -122.28168307199996 47.52516438600003, -122.28145050399996 47.52516388600003, -122.281450582 47.525147437000044, -122.28092621899998 47.525146310000025, -122.28092721899998 47.52433074500004, -122.28129411599997 47.52433142700005, -122.28247185299995 47.52433360800006, -122.28270306299999 47.52433403400005, -122.28270162899997 47.52516656600005, -122.28235899499998 47.525165833000074)))\"]]]} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :properties]] {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PMA]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 443]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SE_ANNO_CAD_DATA]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SDQL]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"QL-D1\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OBJECTID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 207883]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GIS_CRT_DT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2024-01-19T14:07:23Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PIN]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"0603001220\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ADDRESS]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \" \"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SHAPE_Area]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 1.460570191594512E-7]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PMA_NAME]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"East Duwamish GS: S Chicago St\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :NAMEFLAG]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 9]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GIS_EDT_DT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2024-01-19T14:07:23Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SUBPARCEL]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 4779]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OWNER]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"DPR\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :NAME]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"EAST DUWAMISH GREENBELT\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :USE_]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GlobalID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"{26B4F4BC-F9AE-42D4-B062-51387C6D9FFF}\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :AMWOID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"PROPERTY-EDUWC\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :MAINT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"DPR\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :REVIEW_DATE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2004-04-08T00:00:00Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :LEASE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"N\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ACQ_DATE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"1994-10-06T00:00:00Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SHAPE_Length]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.0015461198840460832]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geometry]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector #object[org.locationtech.jts.geom.MultiPolygon 0x58cfc71a \"MULTIPOLYGON (((-122.28304862799996 47.52231724600006, -122.28304832999999 47.52242690400004, -122.28304803199995 47.52253656500005, -122.28304773399998 47.52264622200005, -122.28260370699996 47.52264502600008, -122.28260405999998 47.522588839000036, -122.28260414399995 47.52253536900008, -122.28260449199996 47.522316051000075, -122.28304862799996 47.52231724600006)))\"]]]} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :properties]] {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PMA]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 443]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SE_ANNO_CAD_DATA]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SDQL]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"QL-D1\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OBJECTID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 207884]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GIS_CRT_DT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2024-01-19T14:07:23Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PIN]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"0603001220\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ADDRESS]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \" \"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SHAPE_Area]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 9.743577447048599E-8]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PMA_NAME]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"East Duwamish GS: S Chicago St\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :NAMEFLAG]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 9]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GIS_EDT_DT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2024-01-19T14:07:23Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SUBPARCEL]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 4711]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OWNER]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"DPR\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :NAME]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"EAST DUWAMISH GREENBELT\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :USE_]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GlobalID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"{E97D7F7F-2666-4008-A905-2F416A66C2A1}\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :AMWOID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"PROPERTY-EDUWC\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :MAINT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"DPR\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :REVIEW_DATE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2004-04-08T00:00:00Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :LEASE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"N\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ACQ_DATE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"1994-10-06T00:00:00Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SHAPE_Length]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.0013271653095071236]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geometry]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector #object[org.locationtech.jts.geom.MultiPolygon 0x8cfa074 \"MULTIPOLYGON (((-122.28304922499996 47.52209792800005, -122.28304892699998 47.52220758500005, -122.28304862799996 47.52231724600006, -122.28260449199996 47.522316051000075, -122.28260466499995 47.522206389000075, -122.28260483899999 47.522096730000044, -122.28304922499996 47.52209792800005)))\"]]]} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :properties]] {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PMA]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 443]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SE_ANNO_CAD_DATA]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SDQL]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"QL-D1\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OBJECTID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 207885]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GIS_CRT_DT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2024-01-19T14:07:23Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PIN]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"0603001220\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ADDRESS]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \" \"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SHAPE_Area]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 4.87386722719263E-8]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PMA_NAME]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"East Duwamish GS: S Chicago St\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :NAMEFLAG]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 9]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GIS_EDT_DT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2024-01-19T14:07:23Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SUBPARCEL]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 4779]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OWNER]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"DPR\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :NAME]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"EAST DUWAMISH GREENBELT\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :USE_]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GlobalID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"{C28A4700-BB4F-4C25-928E-57022F6D001B}\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :AMWOID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"PROPERTY-EDUWC\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :MAINT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"DPR\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :REVIEW_DATE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2004-04-08T00:00:00Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :LEASE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"N\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ACQ_DATE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"1994-10-06T00:00:00Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SHAPE_Length]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.0011082207640129814]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geometry]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector #object[org.locationtech.jts.geom.MultiPolygon 0x640bf559 \"MULTIPOLYGON (((-122.28304952299999 47.52198826600005, -122.28304922499996 47.52209792800005, -122.28260483899999 47.522096730000044, -122.28260501299997 47.52198707100007, -122.28304952299999 47.52198826600005)))\"]]]} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :properties]] {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PMA]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 443]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SE_ANNO_CAD_DATA]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SDQL]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"QL-D1\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OBJECTID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 207886]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GIS_CRT_DT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2024-01-19T14:07:23Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PIN]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"0603001220\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ADDRESS]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \" \"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SHAPE_Area]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 9.757207624044397E-8]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :PMA_NAME]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"East Duwamish GS: S Chicago St\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :NAMEFLAG]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 9]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GIS_EDT_DT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2024-01-19T14:07:23Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SUBPARCEL]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 4712]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :OWNER]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"DPR\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :NAME]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"EAST DUWAMISH GREENBELT\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :USE_]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :GlobalID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"{4DB6A932-327A-4E2E-AC8A-C068D114AA61}\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :AMWOID]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"PROPERTY-EDUWC\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :MAINT]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"DPR\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :REVIEW_DATE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"2004-04-08T00:00:00Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :LEASE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"N\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ACQ_DATE]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"1994-10-06T00:00:00Z\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :SHAPE_Length]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.0013284113092379031]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geometry]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector #object[org.locationtech.jts.geom.MultiPolygon 0x3b4fd3f0 \"MULTIPOLYGON (((-122.28305071699998 47.52154962900005, -122.28305041899995 47.52165928900007, -122.28305012099997 47.521768948000044, -122.28260536 47.52176775000004, -122.28260570799995 47.52154843200003, -122.28305071699998 47.52154962900005)))\"]]]})")}
                , document.currentScript.parentElement);</script></div></div>

## Coordinate conversions

Both datasets use the [WGS84](https://en.wikipedia.org/wiki/World_Geodetic_System)
coordinate system.
For area computations, etc., we need to convert them to a coordinate system
which is locally correct in terms of distances in a region around Seattle.

[EPSG:4326](https://epsg.io/4326)

::: {.sourceClojure}
```clojure
(def wgs84-crs (crs/create-crs 4326))
```
:::

[EPSG:2285](https://epsg.io/2285)

United States (USA) - Oregon and Washington. [1 3]:

|      Center coordinates |                                Projected bounds |                      WGS84 bounds |
|-------------------------|-------------------------------------------------|-----------------------------------|
| [1692592.39 -541752.55] | [[559165.71 -1834123.81] [2832684.98 777497.1]] | [[-124.79 41.98] [-116.47 49.05]] |

::: {.sourceClojure}
```clojure
(def Seattle-crs (crs/create-crs 2285))
```
:::

::: {.sourceClojure}
```clojure
(def crs-transform
  (crs/create-transform wgs84-crs Seattle-crs))
```
:::

::: {.sourceClojure}
```clojure
(defn wgs84->Seattle
  "Transforming latitude-longitude coordinates
  to local Eucledian coordinates around Seattle."
  [geometry]
  (jts/transform-geom geometry crs-transform))
```
:::

## Some geometrical functions

::: {.sourceClojure}
```clojure
(defn area [^Geometry geometry]
  (.getArea geometry))
```
:::

::: {.sourceClojure}
```clojure
(defn buffer [^Geometry geometry radius]
  (.buffer geometry radius))
```
:::

::: {.sourceClojure}
```clojure
(defn intersection [^Geometry geometry1
                    ^Geometry geometry2]
  (.intersection geometry1
                 geometry2))
```
:::

## Preproessed datasets

::: {.sourceClojure}
```clojure
(defn geojson->dataset [geojson dataset-name]
  (-> geojson
      (->> (map (fn [{:keys [geometry properties]}]
                  (assoc properties :geometry geometry))))
      tc/dataset
      (tc/map-columns :geometry [:geometry]
                      wgs84->Seattle)
      (tc/set-dataset-name dataset-name)))
```
:::

::: {.sourceClojure}
```clojure
(def neighborhoods
  (-> neighborhoods-geojson
      (geojson->dataset "Seattle neighborhoods")))
```
:::

::: {.sourceClojure}
```clojure
(-> neighborhoods
    (tc/drop-columns [:geometry])
    kind/table)
```
:::

<div style="max-height:600px;overflow-y:auto;"><table class="table"><thead><tr><th>OBJECTID</th><th>L_HOOD</th><th>S_HOOD</th><th>S_HOOD_ALT_NAMES</th><th>Shape__Area</th><th>Shape__Length</th></tr></thead><tbody><tr><td><div><p>27</p></div></td><td><div><p>Ballard</p></div></td><td><div><p>Loyal Heights</p></div></td><td><div></div></td><td><div><p>2.13206555455933E7</p></div></td><td><div><p>18831.00959637</p></div></td></tr><tr><td><div><p>28</p></div></td><td><div><p>Ballard</p></div></td><td><div><p>Ballard</p></div></td><td><div><p>Adams</p></div></td><td><div><p>2.25521329700012E7</p></div></td><td><div><p>29926.3393002841</p></div></td></tr><tr><td><div><p>29</p></div></td><td><div><p>Ballard</p></div></td><td><div><p>Whittier Heights</p></div></td><td><div></div></td><td><div><p>1.41956874919434E7</p></div></td><td><div><p>15934.4392663199</p></div></td></tr><tr><td><div><p>30</p></div></td><td><div><p>Ballard</p></div></td><td><div><p>West Woodland</p></div></td><td><div></div></td><td><div><p>2.21993668730164E7</p></div></td><td><div><p>21789.6510873493</p></div></td></tr><tr><td><div><p>31</p></div></td><td><div><p>North Central</p></div></td><td><div><p>Phinney Ridge</p></div></td><td><div><p>Woodland Park</p></div></td><td><div><p>3.21231178777466E7</p></div></td><td><div><p>27120.7859379681</p></div></td></tr><tr><td><div><p>32</p></div></td><td><div><p>North Central</p></div></td><td><div><p>Wallingford</p></div></td><td><div><p>Meridian, Tangle Town, Northlake</p></div></td><td><div><p>4.20683426137695E7</p></div></td><td><div><p>33040.2469861198</p></div></td></tr><tr><td><div><p>33</p></div></td><td><div><p>North Central</p></div></td><td><div><p>Fremont</p></div></td><td><div></div></td><td><div><p>2.65037565718689E7</p></div></td><td><div><p>25999.2849357187</p></div></td></tr><tr><td><div><p>34</p></div></td><td><div><p>North Central</p></div></td><td><div><p>Green Lake</p></div></td><td><div></div></td><td><div><p>3.73146297898865E7</p></div></td><td><div><p>25387.7074962486</p></div></td></tr><tr><td><div><p>35</p></div></td><td><div><p>Northeast</p></div></td><td><div><p>View Ridge</p></div></td><td><div></div></td><td><div><p>2.56129799440918E7</p></div></td><td><div><p>22657.0394732799</p></div></td></tr><tr><td><div><p>36</p></div></td><td><div><p>Northeast</p></div></td><td><div><p>Ravenna</p></div></td><td><div><p>University Village</p></div></td><td><div><p>3.67833915366516E7</p></div></td><td><div><p>31163.6208763331</p></div></td></tr><tr><td><div><p>37</p></div></td><td><div><p>Northeast</p></div></td><td><div><p>Sand Point</p></div></td><td><div></div></td><td><div><p>1.92498571397095E7</p></div></td><td><div><p>21396.6958008648</p></div></td></tr><tr><td><div><p>38</p></div></td><td><div><p>Northeast</p></div></td><td><div><p>Bryant</p></div></td><td><div></div></td><td><div><p>1.54850780333252E7</p></div></td><td><div><p>20928.0017243062</p></div></td></tr><tr><td><div><p>39</p></div></td><td><div><p>Northeast</p></div></td><td><div><p>Windermere</p></div></td><td><div><p>Hawthorne Hills</p></div></td><td><div><p>2.23639666230164E7</p></div></td><td><div><p>33332.8291968304</p></div></td></tr><tr><td><div><p>40</p></div></td><td><div><p>Northeast</p></div></td><td><div><p>Laurelhurst</p></div></td><td><div></div></td><td><div><p>1.90517768183594E7</p></div></td><td><div><p>26489.2422321737</p></div></td></tr><tr><td><div><p>41</p></div></td><td><div><p>Northeast</p></div></td><td><div><p>Roosevelt</p></div></td><td><div><p>Fairview</p></div></td><td><div><p>2.1491594324707E7</p></div></td><td><div><p>22866.0797374364</p></div></td></tr><tr><td><div><p>42</p></div></td><td><div><p>University District</p></div></td><td><div><p>University of Washington</p></div></td><td><div></div></td><td><div><p>2.55236751180115E7</p></div></td><td><div><p>30983.2126641678</p></div></td></tr><tr><td><div><p>43</p></div></td><td><div><p>Queen Anne</p></div></td><td><div><p>East Queen Anne</p></div></td><td><div></div></td><td><div><p>1.93546995965576E7</p></div></td><td><div><p>21488.397378299</p></div></td></tr><tr><td><div><p>44</p></div></td><td><div><p>Queen Anne</p></div></td><td><div><p>West Queen Anne</p></div></td><td><div></div></td><td><div><p>1.79749859528503E7</p></div></td><td><div><p>20631.7057292923</p></div></td></tr><tr><td><div><p>45</p></div></td><td><div><p>Queen Anne</p></div></td><td><div><p>Lower Queen Anne</p></div></td><td><div><p>Uptown, Seattle Center</p></div></td><td><div><p>1.75156244566956E7</p></div></td><td><div><p>20496.7289222394</p></div></td></tr><tr><td><div><p>46</p></div></td><td><div><p>Queen Anne</p></div></td><td><div><p>North Queen Anne</p></div></td><td><div></div></td><td><div><p>2.9604288333252E7</p></div></td><td><div><p>26753.5542817429</p></div></td></tr><tr><td><div><p>47</p></div></td><td><div><p>Cascade</p></div></td><td><div><p>Westlake</p></div></td><td><div></div></td><td><div><p>3645184.07894897</p></div></td><td><div><p>11017.3231739615</p></div></td></tr><tr><td><div><p>48</p></div></td><td><div><p>Cascade</p></div></td><td><div><p>Eastlake</p></div></td><td><div></div></td><td><div><p>7426907.96139526</p></div></td><td><div><p>16672.5117837403</p></div></td></tr><tr><td><div><p>49</p></div></td><td><div><p>Cascade</p></div></td><td><div><p>South Lake Union</p></div></td><td><div></div></td><td><div><p>1.6107105419342E7</p></div></td><td><div><p>29323.5462843847</p></div></td></tr><tr><td><div><p>50</p></div></td><td><div><p>Magnolia</p></div></td><td><div><p>Lawton Park</p></div></td><td><div></div></td><td><div><p>5.5091525290863E7</p></div></td><td><div><p>38766.0673349107</p></div></td></tr><tr><td><div><p>51</p></div></td><td><div><p>Magnolia</p></div></td><td><div><p>Briarcliff</p></div></td><td><div><p>Carleton Park</p></div></td><td><div><p>2.65509365782471E7</p></div></td><td><div><p>26609.7912261135</p></div></td></tr><tr><td><div><p>52</p></div></td><td><div><p>Magnolia</p></div></td><td><div><p>Southeast Magnolia</p></div></td><td><div></div></td><td><div><p>1.84692638882751E7</p></div></td><td><div><p>20223.1801501522</p></div></td></tr><tr><td><div><p>53</p></div></td><td><div><p>Central Area</p></div></td><td><div><p>Madrona</p></div></td><td><div></div></td><td><div><p>1.38454828085938E7</p></div></td><td><div><p>17052.9427067693</p></div></td></tr><tr><td><div><p>54</p></div></td><td><div><p>Central Area</p></div></td><td><div><p>Harrison/Denny-Blaine</p></div></td><td><div></div></td><td><div><p>1.36033956731873E7</p></div></td><td><div><p>17241.2743191595</p></div></td></tr><tr><td><div><p>55</p></div></td><td><div><p>Central Area</p></div></td><td><div><p>Minor</p></div></td><td><div><p>Central District, Squire Park</p></div></td><td><div><p>1.7933144708252E7</p></div></td><td><div><p>18241.5487859812</p></div></td></tr><tr><td><div><p>56</p></div></td><td><div><p>Central Area</p></div></td><td><div><p>Leschi</p></div></td><td><div></div></td><td><div><p>1.80917413764954E7</p></div></td><td><div><p>20413.0560516882</p></div></td></tr><tr><td><div><p>57</p></div></td><td><div><p>Central Area</p></div></td><td><div><p>Mann</p></div></td><td><div><p>Central District, Garfield</p></div></td><td><div><p>1.11338288620605E7</p></div></td><td><div><p>17843.105758736</p></div></td></tr><tr><td><div><p>58</p></div></td><td><div><p>Central Area</p></div></td><td><div><p>Atlantic</p></div></td><td><div><p>Judkins Park, Jackson Place, Coleman</p></div></td><td><div><p>2.06649862726135E7</p></div></td><td><div><p>25426.2311626675</p></div></td></tr><tr><td><div><p>59</p></div></td><td><div><p>Downtown</p></div></td><td><div><p>Pike-Market</p></div></td><td><div><p>Pike Place Market</p></div></td><td><div><p>2343359.32580566</p></div></td><td><div><p>7326.58173024624</p></div></td></tr><tr><td><div><p>60</p></div></td><td><div><p>Downtown</p></div></td><td><div><p>Belltown</p></div></td><td><div><p>Denny Regrade</p></div></td><td><div><p>9845864.78594971</p></div></td><td><div><p>17836.7271544608</p></div></td></tr><tr><td><div><p>61</p></div></td><td><div><p>Downtown</p></div></td><td><div><p>International District</p></div></td><td><div><p>Chinatown, ID, Little Saigon</p></div></td><td><div><p>6169375.81655884</p></div></td><td><div><p>11772.3871410021</p></div></td></tr><tr><td><div><p>62</p></div></td><td><div><p>Downtown</p></div></td><td><div><p>Central Business District</p></div></td><td><div><p>Commercial Core, West Edge</p></div></td><td><div><p>9477398.61834717</p></div></td><td><div><p>18209.4628624588</p></div></td></tr><tr><td><div><p>63</p></div></td><td><div><p>Downtown</p></div></td><td><div><p>First Hill</p></div></td><td><div></div></td><td><div><p>9517568.47363281</p></div></td><td><div><p>13225.2335159542</p></div></td></tr><tr><td><div><p>64</p></div></td><td><div><p>Downtown</p></div></td><td><div><p>Yesler Terrace</p></div></td><td><div></div></td><td><div><p>5348937.24771118</p></div></td><td><div><p>9611.35354641661</p></div></td></tr><tr><td><div><p>65</p></div></td><td><div><p>Downtown</p></div></td><td><div><p>Pioneer Square</p></div></td><td><div></div></td><td><div><p>7545979.0869751</p></div></td><td><div><p>17297.5867148564</p></div></td></tr><tr><td><div><p>66</p></div></td><td><div><p>Interbay</p></div></td><td><div><p>Interbay</p></div></td><td><div></div></td><td><div><p>3.35039410921631E7</p></div></td><td><div><p>64207.5867611718</p></div></td></tr><tr><td><div><p>67</p></div></td><td><div><p>Greater Duwamish</p></div></td><td><div><p>SODO</p></div></td><td><div></div></td><td><div><p>4.96031675905914E7</p></div></td><td><div><p>38520.9885110863</p></div></td></tr><tr><td><div><p>68</p></div></td><td><div><p>Greater Duwamish</p></div></td><td><div><p>Georgetown</p></div></td><td><div></div></td><td><div><p>5.13050726461182E7</p></div></td><td><div><p>45963.6656769607</p></div></td></tr><tr><td><div><p>69</p></div></td><td><div><p>Greater Duwamish</p></div></td><td><div><p>South Park</p></div></td><td><div></div></td><td><div><p>2.94234521522369E7</p></div></td><td><div><p>30433.4048686311</p></div></td></tr><tr><td><div><p>70</p></div></td><td><div><p>Greater Duwamish</p></div></td><td><div><p>Harbor Island</p></div></td><td><div></div></td><td><div><p>1.78323007354431E7</p></div></td><td><div><p>24381.2966280601</p></div></td></tr><tr><td><div><p>71</p></div></td><td><div><p>West Seattle</p></div></td><td><div><p>Seaview</p></div></td><td><div></div></td><td><div><p>1.8463061797821E7</p></div></td><td><div><p>23122.2640032568</p></div></td></tr><tr><td><div><p>72</p></div></td><td><div><p>West Seattle</p></div></td><td><div><p>Gatewood</p></div></td><td><div><p>Morgan Junction</p></div></td><td><div><p>2.28217074689484E7</p></div></td><td><div><p>23801.2647005122</p></div></td></tr><tr><td><div><p>73</p></div></td><td><div><p>West Seattle</p></div></td><td><div><p>Arbor Heights</p></div></td><td><div></div></td><td><div><p>2.61973140573883E7</p></div></td><td><div><p>28162.4351386166</p></div></td></tr><tr><td><div><p>74</p></div></td><td><div><p>West Seattle</p></div></td><td><div><p>Alki</p></div></td><td><div></div></td><td><div><p>2.45834691841736E7</p></div></td><td><div><p>45568.1378536906</p></div></td></tr><tr><td><div><p>75</p></div></td><td><div><p>West Seattle</p></div></td><td><div><p>North Admiral</p></div></td><td><div><p>Belvidere</p></div></td><td><div><p>4.85814163202972E7</p></div></td><td><div><p>30687.7844700837</p></div></td></tr><tr><td><div><p>76</p></div></td><td><div><p>West Seattle</p></div></td><td><div><p>Fairmount Park</p></div></td><td><div></div></td><td><div><p>1.61539000976257E7</p></div></td><td><div><p>24955.7128414259</p></div></td></tr><tr><td><div><p>77</p></div></td><td><div><p>West Seattle</p></div></td><td><div><p>Genesee</p></div></td><td><div><p>Alaska Junction, West Seattle Junction</p></div></td><td><div><p>2.11649765813446E7</p></div></td><td><div><p>19870.8696089241</p></div></td></tr><tr><td><div><p>78</p></div></td><td><div><p>West Seattle</p></div></td><td><div><p>Fauntleroy</p></div></td><td><div><p>Endolyne, Arroyo Heights, Brace Point</p></div></td><td><div><p>3.38818638706665E7</p></div></td><td><div><p>42276.5317877774</p></div></td></tr><tr><td><div><p>79</p></div></td><td><div><p>Beacon Hill</p></div></td><td><div><p>North Beacon Hill</p></div></td><td><div><p>Jefferson Park</p></div></td><td><div><p>4.84759144557953E7</p></div></td><td><div><p>33865.4891739901</p></div></td></tr><tr><td><div><p>80</p></div></td><td><div><p>Beacon Hill</p></div></td><td><div><p>Mid-Beacon Hill</p></div></td><td><div></div></td><td><div><p>5.26204707883301E7</p></div></td><td><div><p>40522.1829275323</p></div></td></tr><tr><td><div><p>81</p></div></td><td><div><p>Beacon Hill</p></div></td><td><div><p>South Beacon Hill</p></div></td><td><div></div></td><td><div><p>3.84488268989563E7</p></div></td><td><div><p>38996.7151070849</p></div></td></tr><tr><td><div><p>82</p></div></td><td><div><p>Beacon Hill</p></div></td><td><div><p>Holly Park</p></div></td><td><div><p>New Holly</p></div></td><td><div><p>8444235.97859192</p></div></td><td><div><p>15848.3624185113</p></div></td></tr><tr><td><div><p>83</p></div></td><td><div><p>Rainier Valley</p></div></td><td><div><p>Brighton</p></div></td><td><div></div></td><td><div><p>1.8366938275177E7</p></div></td><td><div><p>18911.3293490837</p></div></td></tr><tr><td><div><p>84</p></div></td><td><div><p>Rainier Valley</p></div></td><td><div><p>Dunlap</p></div></td><td><div></div></td><td><div><p>1.96016012150879E7</p></div></td><td><div><p>20847.6531922997</p></div></td></tr><tr><td><div><p>85</p></div></td><td><div><p>Rainier Valley</p></div></td><td><div><p>Rainier Beach</p></div></td><td><div></div></td><td><div><p>3.31282361414948E7</p></div></td><td><div><p>34965.6132100475</p></div></td></tr><tr><td><div><p>86</p></div></td><td><div><p>Rainier Valley</p></div></td><td><div><p>Rainier View</p></div></td><td><div><p>Lake Ridge</p></div></td><td><div><p>1.82485661794434E7</p></div></td><td><div><p>34791.8218147105</p></div></td></tr><tr><td><div><p>87</p></div></td><td><div><p>Rainier Valley</p></div></td><td><div><p>Mount Baker</p></div></td><td><div><p>North Rainier</p></div></td><td><div><p>2.89869952343292E7</p></div></td><td><div><p>26318.9819054871</p></div></td></tr><tr><td><div><p>88</p></div></td><td><div><p>Rainier Valley</p></div></td><td><div><p>Columbia City</p></div></td><td><div><p>Columbia Heights</p></div></td><td><div><p>3.74065764927979E7</p></div></td><td><div><p>35177.0608039675</p></div></td></tr><tr><td><div><p>89</p></div></td><td><div><p>Delridge</p></div></td><td><div><p>Highland Park</p></div></td><td><div></div></td><td><div><p>3.98103334158325E7</p></div></td><td><div><p>31111.0630837714</p></div></td></tr><tr><td><div><p>90</p></div></td><td><div><p>Delridge</p></div></td><td><div><p>North Delridge</p></div></td><td><div><p>Avalon, Luna Park, Pigeon Point</p></div></td><td><div><p>2.94546565672607E7</p></div></td><td><div><p>23549.9313666304</p></div></td></tr><tr><td><div><p>91</p></div></td><td><div><p>Delridge</p></div></td><td><div><p>Riverview</p></div></td><td><div></div></td><td><div><p>3.30382671929626E7</p></div></td><td><div><p>25917.1176391015</p></div></td></tr><tr><td><div><p>92</p></div></td><td><div><p>Delridge</p></div></td><td><div><p>High Point</p></div></td><td><div></div></td><td><div><p>2.33430083368683E7</p></div></td><td><div><p>23180.3963596714</p></div></td></tr><tr><td><div><p>93</p></div></td><td><div><p>Delridge</p></div></td><td><div><p>South Delridge</p></div></td><td><div><p>White Center</p></div></td><td><div><p>1.79233162238464E7</p></div></td><td><div><p>23416.3405359794</p></div></td></tr><tr><td><div><p>94</p></div></td><td><div><p>Delridge</p></div></td><td><div><p>Roxhill</p></div></td><td><div><p>Westwood Village</p></div></td><td><div><p>1.87810598441162E7</p></div></td><td><div><p>22360.0472289145</p></div></td></tr><tr><td><div><p>95</p></div></td><td><div><p>Seward Park</p></div></td><td><div><p>Seward Park</p></div></td><td><div><p>Lakewood, Hillman City</p></div></td><td><div><p>4.38352790059662E7</p></div></td><td><div><p>52212.7868103531</p></div></td></tr><tr><td><div><p>96</p></div></td><td><div><p>Northeast</p></div></td><td><div><p>Wedgwood</p></div></td><td><div></div></td><td><div><p>3.05997136261597E7</p></div></td><td><div><p>23840.2187274324</p></div></td></tr><tr><td><div><p>97</p></div></td><td><div><p>Capitol Hill</p></div></td><td><div><p>Portage Bay</p></div></td><td><div><p>Roanoke Park</p></div></td><td><div><p>4421635.26721191</p></div></td><td><div><p>10013.2739563989</p></div></td></tr><tr><td><div><p>98</p></div></td><td><div><p>Capitol Hill</p></div></td><td><div><p>Montlake</p></div></td><td><div><p>North Capitol Hill</p></div></td><td><div><p>2.1474786741394E7</p></div></td><td><div><p>46165.83098345</p></div></td></tr><tr><td><div><p>99</p></div></td><td><div><p>Capitol Hill</p></div></td><td><div><p>Madison Park</p></div></td><td><div><p>Broadmoor, Washington Park</p></div></td><td><div><p>2.46809502223511E7</p></div></td><td><div><p>26230.4347480002</p></div></td></tr><tr><td><div><p>100</p></div></td><td><div><p>Capitol Hill</p></div></td><td><div><p>Broadway</p></div></td><td><div><p>Pike/Pine</p></div></td><td><div><p>3.02392405212402E7</p></div></td><td><div><p>26851.9474418256</p></div></td></tr><tr><td><div><p>101</p></div></td><td><div><p>Capitol Hill</p></div></td><td><div><p>Stevens</p></div></td><td><div><p>Interlaken Park, Miller Park, Madison-Miller</p></div></td><td><div><p>2.60395920790405E7</p></div></td><td><div><p>25522.3456368768</p></div></td></tr><tr><td><div><p>102</p></div></td><td><div><p>Lake City</p></div></td><td><div><p>Victory Heights</p></div></td><td><div></div></td><td><div><p>2.15127314890137E7</p></div></td><td><div><p>20753.0018992847</p></div></td></tr><tr><td><div><p>103</p></div></td><td><div><p>Lake City</p></div></td><td><div><p>Matthews Beach</p></div></td><td><div></div></td><td><div><p>3.09809783014832E7</p></div></td><td><div><p>31369.8925404597</p></div></td></tr><tr><td><div><p>104</p></div></td><td><div><p>Lake City</p></div></td><td><div><p>Meadowbrook</p></div></td><td><div></div></td><td><div><p>1.70348487789307E7</p></div></td><td><div><p>18770.2820324215</p></div></td></tr><tr><td><div><p>105</p></div></td><td><div><p>Lake City</p></div></td><td><div><p>Olympic Hills</p></div></td><td><div></div></td><td><div><p>2.58367421641541E7</p></div></td><td><div><p>19993.8160570763</p></div></td></tr><tr><td><div><p>106</p></div></td><td><div><p>Lake City</p></div></td><td><div><p>Cedar Park</p></div></td><td><div></div></td><td><div><p>2.04414858487244E7</p></div></td><td><div><p>22788.760382662</p></div></td></tr><tr><td><div><p>110</p></div></td><td><div><p>Northwest</p></div></td><td><div><p>Broadview</p></div></td><td><div></div></td><td><div><p>4.81134497303162E7</p></div></td><td><div><p>34640.309665739</p></div></td></tr><tr><td><div><p>111</p></div></td><td><div><p>Northwest</p></div></td><td><div><p>Bitter Lake</p></div></td><td><div></div></td><td><div><p>2.76903364438171E7</p></div></td><td><div><p>26446.2030540938</p></div></td></tr><tr><td><div><p>112</p></div></td><td><div><p>Northgate</p></div></td><td><div><p>Haller Lake</p></div></td><td><div></div></td><td><div><p>4.72030810796204E7</p></div></td><td><div><p>29413.5534484151</p></div></td></tr><tr><td><div><p>113</p></div></td><td><div><p>Northgate</p></div></td><td><div><p>Pinehurst</p></div></td><td><div><p>Jackson Park</p></div></td><td><div><p>2.94388508770142E7</p></div></td><td><div><p>26171.2717608103</p></div></td></tr><tr><td><div><p>114</p></div></td><td><div><p>Northwest</p></div></td><td><div><p>North Beach/Blue Ridge</p></div></td><td><div></div></td><td><div><p>3.10622649915771E7</p></div></td><td><div><p>32947.8059119899</p></div></td></tr><tr><td><div><p>115</p></div></td><td><div><p>Northgate</p></div></td><td><div><p>Licton Springs</p></div></td><td><div><p>North College Park</p></div></td><td><div><p>2.17047062611389E7</p></div></td><td><div><p>20132.8982458333</p></div></td></tr><tr><td><div><p>116</p></div></td><td><div><p>Northgate</p></div></td><td><div><p>Maple Leaf</p></div></td><td><div><p>Olympic View</p></div></td><td><div><p>3.43804707826233E7</p></div></td><td><div><p>25503.0574082691</p></div></td></tr><tr><td><div><p>117</p></div></td><td><div><p>Northwest</p></div></td><td><div><p>Crown Hill</p></div></td><td><div></div></td><td><div><p>1.58657504196777E7</p></div></td><td><div><p>18997.8080379411</p></div></td></tr><tr><td><div><p>118</p></div></td><td><div><p>Northwest</p></div></td><td><div><p>Greenwood</p></div></td><td><div></div></td><td><div><p>4.20445636731262E7</p></div></td><td><div><p>26501.5265911049</p></div></td></tr><tr><td><div><p>119</p></div></td><td><div><p>Ballard</p></div></td><td><div><p>Sunset Hill</p></div></td><td><div><p>Golden Gardens, Shilshole</p></div></td><td><div><p>2.43356204689941E7</p></div></td><td><div><p>29416.7300633571</p></div></td></tr><tr><td><div><p>121</p></div></td><td><div><p>University District</p></div></td><td><div><p>University District</p></div></td><td><div></div></td><td><div><p>1.36862548569946E7</p></div></td><td><div><p>21581.6459789299</p></div></td></tr><tr><td><div><p>123</p></div></td><td><div><p>University District</p></div></td><td><div><p>University Heights</p></div></td><td><div><p>Cowen Park</p></div></td><td><div><p>1.03757480522461E7</p></div></td><td><div><p>15163.9069096811</p></div></td></tr><tr><td><div><p>124</p></div></td><td><div><p>Downtown</p></div></td><td><div><p>Denny Triangle</p></div></td><td><div><p>Denny Regrade</p></div></td><td><div><p>5128172.31045532</p></div></td><td><div><p>10574.1779826691</p></div></td></tr><tr><td><div><p>126</p></div></td><td><div><p>Greater Duwamish</p></div></td><td><div><p>Industrial District</p></div></td><td><div></div></td><td><div><p>6.24341310114746E7</p></div></td><td><div><p>88995.7569839975</p></div></td></tr></tbody></table></div>

::: {.sourceClojure}
```clojure
(def parks
  (-> parks-geojson
      (geojson->dataset "Seattle parks")
      (tc/map-columns :geometry
                      [:geometry]
                      ;; avoiding some [linestring pathologies](https://gis.stackexchange.com/questions/50399/fixing-non-noded-intersection-problem-using-postgis)
                      #(buffer % 1))))
```
:::

::: {.sourceClojure}
```clojure
(delay
  (-> parks
      (tc/drop-columns [:geometry])
      (tc/head 20)
      kind/table))
```
:::

<div style="max-height:600px;overflow-y:auto;"><table class="table"><thead><tr><th>SE_ANNO_CAD_DATA</th><th>NAMEFLAG</th><th>ADDRESS</th><th>MAINT</th><th>LEASE</th><th>PMA_NAME</th><th>SUBPARCEL</th><th>PMA</th><th>REVIEW_DATE</th><th>GlobalID</th><th>OBJECTID</th><th>AMWOID</th><th>SDQL</th><th>USE_</th><th>PIN</th><th>GIS_EDT_DT</th><th>SHAPE_Area</th><th>SHAPE_Length</th><th>GIS_CRT_DT</th><th>NAME</th><th>OWNER</th><th>ACQ_DATE</th></tr></thead><tbody><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>East Duwamish GS: S Chicago St</p></div></td><td><div><p>9851</p></div></td><td><div><p>442</p></div></td><td><div><p>2004-04-08T00:00:00Z</p></div></td><td><div><p>{D4B7025E-1135-4232-88AB-CBF72932E6AE}</p></div></td><td><div><p>207882</p></div></td><td><div><p>PROPERTY-EDUWSC</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>4006000485</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>1.4691943312522366E-6</p></div></td><td><div><p>0.005215887160444538</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>EAST DUWAMISH GREENBELT</p></div></td><td><div><p>DPR</p></div></td><td><div><p>1997-10-21T00:00:00Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>East Duwamish GS: S Chicago St</p></div></td><td><div><p>4779</p></div></td><td><div><p>443</p></div></td><td><div><p>2004-04-08T00:00:00Z</p></div></td><td><div><p>{26B4F4BC-F9AE-42D4-B062-51387C6D9FFF}</p></div></td><td><div><p>207883</p></div></td><td><div><p>PROPERTY-EDUWC</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>0603001220</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>1.460570191594512E-7</p></div></td><td><div><p>0.0015461198840460832</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>EAST DUWAMISH GREENBELT</p></div></td><td><div><p>DPR</p></div></td><td><div><p>1994-10-06T00:00:00Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>East Duwamish GS: S Chicago St</p></div></td><td><div><p>4711</p></div></td><td><div><p>443</p></div></td><td><div><p>2004-04-08T00:00:00Z</p></div></td><td><div><p>{E97D7F7F-2666-4008-A905-2F416A66C2A1}</p></div></td><td><div><p>207884</p></div></td><td><div><p>PROPERTY-EDUWC</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>0603001220</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>9.743577447048599E-8</p></div></td><td><div><p>0.0013271653095071236</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>EAST DUWAMISH GREENBELT</p></div></td><td><div><p>DPR</p></div></td><td><div><p>1994-10-06T00:00:00Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>East Duwamish GS: S Chicago St</p></div></td><td><div><p>4779</p></div></td><td><div><p>443</p></div></td><td><div><p>2004-04-08T00:00:00Z</p></div></td><td><div><p>{C28A4700-BB4F-4C25-928E-57022F6D001B}</p></div></td><td><div><p>207885</p></div></td><td><div><p>PROPERTY-EDUWC</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>0603001220</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>4.87386722719263E-8</p></div></td><td><div><p>0.0011082207640129814</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>EAST DUWAMISH GREENBELT</p></div></td><td><div><p>DPR</p></div></td><td><div><p>1994-10-06T00:00:00Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>East Duwamish GS: S Chicago St</p></div></td><td><div><p>4712</p></div></td><td><div><p>443</p></div></td><td><div><p>2004-04-08T00:00:00Z</p></div></td><td><div><p>{4DB6A932-327A-4E2E-AC8A-C068D114AA61}</p></div></td><td><div><p>207886</p></div></td><td><div><p>PROPERTY-EDUWC</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>0603001220</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>9.757207624044397E-8</p></div></td><td><div><p>0.0013284113092379031</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>EAST DUWAMISH GREENBELT</p></div></td><td><div><p>DPR</p></div></td><td><div><p>1994-10-06T00:00:00Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>Genesee Park and Playfield</p></div></td><td><div><p>4618</p></div></td><td><div><p>409</p></div></td><td><div><p>1899-12-30T00:00:01Z</p></div></td><td><div><p>{AD2E47C9-323B-4DDE-8186-98A4FC79C3F8}</p></div></td><td><div><p>207887</p></div></td><td><div><p>PROPERTY-GENPK</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>4154300585</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>2.892289652506793E-6</p></div></td><td><div><p>0.007000440212562059</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>GENESEE PARK AND PLAYFIELD</p></div></td><td><div><p>DPR</p></div></td><td><div><p>1996-06-03T00:00:00Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div><p>5244 SW JACOBSEN ROAD</p></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>Wolf Creek Ravine Natural Area</p></div></td><td><div><p>16397</p></div></td><td><div><p>278</p></div></td><td><div><p>2004-04-19T00:00:00Z</p></div></td><td><div><p>{A7E54352-0DC3-47E4-B562-DAA0F18B78E2}</p></div></td><td><div><p>207888</p></div></td><td><div><p>PROPERTY-WLFCRAV</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>5344200125</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>6.25557675764001E-8</p></div></td><td><div><p>0.0010504796990982068</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>WOLF CREEK RAVINE NATURAL AREA</p></div></td><td><div><p>DPR</p></div></td><td><div><p>2004-04-07T00:00:00Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>Jefferson Park</p></div></td><td><div><p>15550</p></div></td><td><div><p>114</p></div></td><td><div><p>2004-04-08T00:00:00Z</p></div></td><td><div><p>{B82B7BD2-DCD9-49A2-BE5D-30C1D9BF4392}</p></div></td><td><div><p>207889</p></div></td><td><div><p>PROPERTY-JEFFPK</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>1624049270</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>3.2670040661028276E-6</p></div></td><td><div><p>0.018380332716286137</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>JEFFERSON PARK</p></div></td><td><div><p>DPR</p></div></td><td><div><p>1988-07-25T00:00:00Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>Queen Anne Boulevard</p></div></td><td><div><p>753</p></div></td><td><div><p>328</p></div></td><td><div><p>2004-04-09T00:00:00Z</p></div></td><td><div><p>{3DA31C7B-C33D-4DD9-9092-71B1DA135BC8}</p></div></td><td><div><p>207890</p></div></td><td><div><p>PROPERTY-QABL</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>2533300180</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>5.790553187521671E-9</p></div></td><td><div><p>4.111515370853996E-4</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>QUEEN ANNE BOULEVARD</p></div></td><td><div><p>DPR</p></div></td><td><div><p>1910-07-25T00:00:01Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>Queen Anne Boulevard</p></div></td><td><div><p>4688</p></div></td><td><div><p>328</p></div></td><td><div><p>2004-04-09T00:00:00Z</p></div></td><td><div><p>{D3DA12EE-31E0-4476-95CE-1739E8FF8550}</p></div></td><td><div><p>207891</p></div></td><td><div><p>PROPERTY-QABL</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>2533300180</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>5.139665582905777E-9</p></div></td><td><div><p>3.9228813174494234E-4</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>QUEEN ANNE BOULEVARD</p></div></td><td><div><p>DPR</p></div></td><td><div><p>1910-07-25T00:00:01Z</p></div></td></tr><tr><td><div></div></td><td><div><p>7</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>Ravenna Woods</p></div></td><td><div><p>15807</p></div></td><td><div><p>4411</p></div></td><td><div><p>2004-05-05T00:00:00Z</p></div></td><td><div><p>{B1A2F3C6-C43A-4600-8064-A3DEFAE61DDC}</p></div></td><td><div><p>207892</p></div></td><td><div><p>PROPERTY-RAVWOOD</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>0925049168</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>5.585829095854119E-7</p></div></td><td><div><p>0.0030915009702507513</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>RAVENNA WOODS</p></div></td><td><div><p>DPR</p></div></td><td><div><p>2002-07-02T00:00:00Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>Montlake Playfield</p></div></td><td><div><p>1507</p></div></td><td><div><p>376</p></div></td><td><div><p>1998-10-15T00:00:00Z</p></div></td><td><div><p>{BF38EE1A-DBAB-400C-B083-B3F8B6A1EB39}</p></div></td><td><div><p>207893</p></div></td><td><div><p>PROPERTY-MTLKPF</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>4089400080</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>3.917576385699785E-7</p></div></td><td><div><p>0.0026768636765498316</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>MONTLAKE PLAYFIELD</p></div></td><td><div><p>DPR</p></div></td><td><div><p>1933-12-31T00:00:01Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>Montlake Playfield</p></div></td><td><div><p>1507</p></div></td><td><div><p>376</p></div></td><td><div><p>2004-03-30T00:00:00Z</p></div></td><td><div><p>{4CC47610-4927-465F-A346-04F66008D480}</p></div></td><td><div><p>207894</p></div></td><td><div><p>PROPERTY-MTLKPF</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>6788202280</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>1.1273410584671711E-6</p></div></td><td><div><p>0.004459900867682873</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>MONTLAKE PLAYFIELD</p></div></td><td><div><p>DPR</p></div></td><td><div><p>1933-12-31T00:00:01Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>Montlake Playfield</p></div></td><td><div><p>1507</p></div></td><td><div><p>376</p></div></td><td><div><p>2004-03-30T00:00:00Z</p></div></td><td><div><p>{10310B09-822A-4344-95CC-2D5D90C2466C}</p></div></td><td><div><p>207895</p></div></td><td><div><p>PROPERTY-MTLKPF</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>6788202280</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>1.4591960880701082E-6</p></div></td><td><div><p>0.006312257575513224</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>MONTLAKE PLAYFIELD</p></div></td><td><div><p>DPR</p></div></td><td><div><p>1933-12-31T00:00:01Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div><p>4357 PALATINE AVE N</p></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>Fremont Peak Park</p></div></td><td><div><p>16454</p></div></td><td><div><p>4424</p></div></td><td><div><p>2004-10-06T00:00:00Z</p></div></td><td><div><p>{A13D75B6-B355-4BDF-84CF-4D2838170201}</p></div></td><td><div><p>207896</p></div></td><td><div><p>PROPERTY-FRMPKPK</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>6610000810</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>6.118914081332531E-8</p></div></td><td><div><p>0.0011669443524576426</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>FREMONT PEAK PARK</p></div></td><td><div><p>DPR</p></div></td><td><div><p>2004-09-30T00:00:00Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>SW Queen Anne Greenbelt</p></div></td><td><div><p>16459</p></div></td><td><div><p>234</p></div></td><td><div><p>2005-01-06T00:00:00Z</p></div></td><td><div><p>{A13C70F6-F0AE-4E55-A12C-990BBC545F8D}</p></div></td><td><div><p>207897</p></div></td><td><div><p>PROPERTY-SWQAGRB</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>7705100037</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>5.565195258199712E-9</p></div></td><td><div><p>3.556012201607765E-4</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>SW QUEEN ANNE GREENBELT</p></div></td><td><div><p>DPR</p></div></td><td><div><p>2004-11-04T00:00:00Z</p></div></td></tr><tr><td><div></div></td><td><div><p>3</p></div></td><td><div><p>9026 4TH AVE S</p></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>Marra-Desimone Park</p></div></td><td><div><p>16455</p></div></td><td><div><p>4447</p></div></td><td><div><p>2007-02-21T00:00:00Z</p></div></td><td><div><p>{E641A30D-FE4B-4B1D-9358-7F0F71324691}</p></div></td><td><div><p>207898</p></div></td><td><div><p>PROPERTY-MARDESPK</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>3224049044</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>2.211413981358282E-6</p></div></td><td><div><p>0.006388002898489655</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>MARRA-DESIMONE PARK</p></div></td><td><div><p>DPR</p></div></td><td><div><p>2004-12-28T00:00:00Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div><p>9026 4TH AVE S</p></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>Marra-Desimone Park</p></div></td><td><div><p>16455</p></div></td><td><div><p>4447</p></div></td><td><div><p>2007-02-21T00:00:00Z</p></div></td><td><div><p>{6A440787-F078-4DB9-9187-DA0E029A8E18}</p></div></td><td><div><p>207899</p></div></td><td><div><p>PROPERTY-MARDESPK</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div><p>3224049035</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>1.981420659558925E-6</p></div></td><td><div><p>0.009774073825241834</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>MARRA-DESIMONE PARK</p></div></td><td><div><p>DPR</p></div></td><td><div><p>2004-12-28T00:00:00Z</p></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>Y</p></div></td><td><div><p>Meadowbrook Playfield</p></div></td><td><div><p>803</p></div></td><td><div><p>352</p></div></td><td><div></div></td><td><div><p>{49C14F14-94FF-4080-9DE8-56EDF006145A}</p></div></td><td><div><p>207900</p></div></td><td><div><p>PROPERTY-MBPF</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>5.268695375108406E-7</p></div></td><td><div><p>0.00550990095454217</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>MEADOWBROOK PLAYFIELD</p></div></td><td><div><p>SSD1</p></div></td><td><div></div></td></tr><tr><td><div></div></td><td><div><p>9</p></div></td><td><div></div></td><td><div><p>DPR</p></div></td><td><div><p>N</p></div></td><td><div><p>Meadowbrook Teen Life Center</p></div></td><td><div></div></td><td><div><p>352</p></div></td><td><div></div></td><td><div><p>{D099D793-CFEA-44E9-A57B-D8A1BA4525AF}</p></div></td><td><div><p>207901</p></div></td><td><div><p>PROPERTY-MBPF</p></div></td><td><div><p>QL-D1</p></div></td><td><div></div></td><td><div></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>8.504523207167994E-8</p></div></td><td><div><p>0.0011865488770908574</p></div></td><td><div><p>2024-01-19T14:07:23Z</p></div></td><td><div><p>MEADOWBROOK TEEN LIFE CENTER</p></div></td><td><div><p>SSD1</p></div></td><td><div></div></td></tr></tbody></table></div>

## A Spatial index structure

We need an index structure to quickly match between the two sets of geometries.

See the JTS [SearchUsingPreparedGeometryIndex tutorial](https://github.com/locationtech/jts/blob/master/modules/example/src/main/java/org/locationtech/jtsexample/technique/SearchUsingPreparedGeometryIndex.java).

::: {.sourceClojure}
```clojure
(defn make-spatial-index [dataset & {:keys [geometry-column]
                                     :or   {geometry-column :geometry}}]
  (let [tree ^STRtree (STRtree.)]
    (doseq [row (tc/rows dataset :as-maps)]
      (let [geometry ^Geometry (row geometry-column)]
        (.insert tree
                 (.getEnvelopeInternal geometry)
                 (assoc row
                        :prepared-geometry
                        (PreparedGeometryFactory/prepare geometry)))))
    tree))
```
:::

::: {.sourceClojure}
```clojure
(def parks-index
  (make-spatial-index parks))
```
:::

::: {.sourceClojure}
```clojure
(defn intersecting-places [^Geometry region spatial-index]
  (->> (.query ^STRtree spatial-index
               ^Envelope (.getEnvelopeInternal ^Geometry region))
       (filter (fn [row]
                 (.intersects ^PreparedGeometry (:prepared-geometry row)
                              region)))
       tc/dataset))
```
:::

For example, let us find the parks intersecting with the first neighborhood:

::: {.sourceClojure}
```clojure
(delay
  (-> neighborhoods
      :geometry
      first
      (intersecting-places parks-index)
      (tc/select-columns [:PMA_NAME :NAME])))
```
:::

_unnamed [2 2]:

|               :PMA_NAME |                   :NAME |
|-------------------------|-------------------------|
|         Salmon Bay Park |         SALMON BAY PARK |
| Loyal Heights Playfield | LOYAL HEIGHTS PLAYFIELD |

## A Joined dataset

We compute [a spatial join](https://en.wikipedia.org/wiki/Spatial_join)
of the two datasets.

Note that even though many parks will appear as intersecting many neighbourhoods, this is not too memory-heavy, since they are references to the same map.

::: {.sourceClojure}
```clojure
(def neighborhoods-with-parks
  (-> neighborhoods
      (tc/map-columns :parks
                      [:geometry]
                      #(intersecting-places % parks-index))))
```
:::

::: {.sourceClojure}
```clojure
(delay
  (-> neighborhoods-with-parks
      (tc/map-columns :n-parks
                      [:parks]
                      tc/row-count)
      (tc/select-columns [:L_HOOD :S_HOOD :S_HOOD_ALT_NAMES
                          :n-parks])))
```
:::

Seattle neighborhoods [94 4]:

|             :L_HOOD |                :S_HOOD |                :S_HOOD_ALT_NAMES | :n-parks |
|---------------------|------------------------|----------------------------------|---------:|
|             Ballard |          Loyal Heights |                                  |        2 |
|             Ballard |                Ballard |                            Adams |        7 |
|             Ballard |       Whittier Heights |                                  |        4 |
|             Ballard |          West Woodland |                                  |        5 |
|       North Central |          Phinney Ridge |                    Woodland Park |       11 |
|       North Central |            Wallingford | Meridian, Tangle Town, Northlake |       27 |
|       North Central |                Fremont |                                  |       15 |
|       North Central |             Green Lake |                                  |       19 |
|           Northeast |             View Ridge |                                  |       15 |
|           Northeast |                Ravenna |               University Village |       12 |
|                 ... |                    ... |                              ... |      ... |
|           Northgate |              Pinehurst |                     Jackson Park |        8 |
|           Northwest | North Beach/Blue Ridge |                                  |       32 |
|           Northgate |         Licton Springs |               North College Park |        6 |
|           Northgate |             Maple Leaf |                     Olympic View |       18 |
|           Northwest |             Crown Hill |                                  |        5 |
|           Northwest |              Greenwood |                                  |       36 |
|             Ballard |            Sunset Hill |        Golden Gardens, Shilshole |        9 |
| University District |    University District |                                  |       11 |
| University District |     University Heights |                       Cowen Park |       21 |
|            Downtown |         Denny Triangle |                    Denny Regrade |        3 |
|    Greater Duwamish |    Industrial District |                                  |        5 |

## Computing areas
For evey neigbourhood, we will compute the proportion of its area covered by parks.

::: {.sourceClojure}
```clojure
(delay
  (-> neighborhoods-with-parks
      (tc/map-columns :neighborhood-area
                      [:geometry]
                      area)
      (tc/map-columns :intersection-area
                      [:geometry :parks]
                      (fn [^Geometry neigh-geometry parks]
                        (->> parks
                             :geometry
                             (map (fn [^Geometry park-geometry]
                                    (area
                                     (.intersection (.buffer park-geometry 1)
                                                    neigh-geometry))))
                             fun/sum)))
      (tc/map-columns :park-names
                      [:parks]
                      (fn [parks]
                        (->> parks
                             :PMA_NAME
                             distinct
                             vec)))
      (tc/add-column :area-proportion
                     #(fun// (:intersection-area %)
                             (:neighborhood-area %)))
      (tc/select-columns [:L_HOOD :S_HOOD :S_HOOD_ALT_NAMES
                          :park-names
                          :neighborhood-area
                          :intersection-area
                          :area-proportion])
      (tc/order-by [:area-proportion] :desc)
      kind/table))
```
:::

<div style="max-height:600px;overflow-y:auto;"><table class="table"><thead><tr><th>L_HOOD</th><th>S_HOOD</th><th>S_HOOD_ALT_NAMES</th><th>park-names</th><th>neighborhood-area</th><th>intersection-area</th><th>area-proportion</th></tr></thead><tbody><tr><td><div><p>Northeast</p></div></td><td><div><p>Sand Point</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Warren G. Magnuson Park"]
</code></pre></div></div></td><td><div><p>1.924985714007309E7</p></div></td><td><div><p>1.2694407322392896E7</p></div></td><td><div><p>0.6594546250406457</p></div></td></tr><tr><td><div><p>Magnolia</p></div></td><td><div><p>Lawton Park</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Magnolia Manor Park"
 "Lawton Park"
 "Discovery Park"
 "Kiwanis Memorial Preserve Park"
 "Discovery Park Boulevard"
 "Discovery Park Tidelands"
 "Kiwanis Ravine Overlook"
 "Daybreak Star"
 "Commodore Park"]
</code></pre></div></div></td><td><div><p>5.5091525292193405E7</p></div></td><td><div><p>2.5768911594667178E7</p></div></td><td><div><p>0.46774728886148104</p></div></td></tr><tr><td><div><p>Delridge</p></div></td><td><div><p>North Delridge</p></div></td><td><div><p>Avalon, Luna Park, Pigeon Point</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Greg Davis Park"
 "Cottage Grove Park"
 "Rotary Viewpoint"
 "Camp Long"
 "West Seattle Golf Course"
 "West Seattle Stadium"
 "Longfellow Creek GS: North"
 "Puget Boulevard"
 "West Duwamish Greenbelt"
 "Pigeon Point"
 "Delridge Playfield"
 "Westbridge Shops"
 "WEST DUWAMISH GREENBELT"]
</code></pre></div></div></td><td><div><p>2.945465656812994E7</p></div></td><td><div><p>1.3245569878227614E7</p></div></td><td><div><p>0.44969357723081976</p></div></td></tr><tr><td><div><p>North Central</p></div></td><td><div><p>Green Lake</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Ravenna Boulevard"
 "Green Lake Park"
 "Green Lake"
 "Crescent Place"
 "North Park Shops"]
</code></pre></div></div></td><td><div><p>3.731462978987473E7</p></div></td><td><div><p>1.5314645167249074E7</p></div></td><td><div><p>0.41041932489987293</p></div></td></tr><tr><td><div><p>Seward Park</p></div></td><td><div><p>Seward Park</p></div></td><td><div><p>Lakewood, Hillman City</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Martha Washington Park"
 "Seward Park"
 "Lake Washington Boulevard"
 "Lakewood Playground"
 "Lakewood Moorage"
 "Genesee Park and Playfield"
 "Stan Sayres Memorial Park"]
</code></pre></div></div></td><td><div><p>4.3835279006813824E7</p></div></td><td><div><p>1.3252664479684925E7</p></div></td><td><div><p>0.30232873566573876</p></div></td></tr><tr><td><div><p>Delridge</p></div></td><td><div><p>Riverview</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Delridge and Myrtle"
 "West Duwamish GS: Riverview"
 "Riverview Playfield"
 "West Duwamish GS: Puget Park"
 "Puget Creek Edible Park"
 "Puget Creek Greenspace"
 "Puget Ridge Playground"
 "Delridge Wetland"
 "Puget Park"
 "Puget Boulevard"
 "Pigeon Point"]
</code></pre></div></div></td><td><div><p>3.303826719332341E7</p></div></td><td><div><p>8769575.924383609</p></div></td><td><div><p>0.2654369211638261</p></div></td></tr><tr><td><div><p>North Central</p></div></td><td><div><p>Phinney Ridge</p></div></td><td><div><p>Woodland Park</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Cascade Place"
 "Sunset Place"
 "Rainier Place"
 "Woodland Park Zoo"
 "Woodland Park"
 "Linden Orchard Park"
 "Green Lake Park"]
</code></pre></div></div></td><td><div><p>3.2123117878649313E7</p></div></td><td><div><p>8349752.958056866</p></div></td><td><div><p>0.25992971758219474</p></div></td></tr><tr><td><div><p>Northgate</p></div></td><td><div><p>Pinehurst</p></div></td><td><div><p>Jackson Park</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Hubbard Homestead Park"
 "Victory Creek Park"
 "Pinehurst Playground"
 "Flicker Haven Natural Area on Thornton Creek"
 "Licorice Fern Natural Area on Thornton Creek"
 "Jackson Park Golf Course"]
</code></pre></div></div></td><td><div><p>2.9438850877818238E7</p></div></td><td><div><p>7646861.188494857</p></div></td><td><div><p>0.2597540651376668</p></div></td></tr><tr><td><div><p>Capitol Hill</p></div></td><td><div><p>Montlake</p></div></td><td><div><p>North Capitol Hill</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Washington Park and Arboretum"
 "Interlaken Park"
 "Montlake Boulevard"
 "Lake Washington Boulevard"
 "East Montlake Park"
 "Grand Army Cemetery"
 "Montlake Playfield"
 "Harvard-Miller/Roanoke Annex"
 "West Montlake Park"]
</code></pre></div></div></td><td><div><p>2.14747867418963E7</p></div></td><td><div><p>5452986.107546446</p></div></td><td><div><p>0.25392504116969533</p></div></td></tr><tr><td><div><p>West Seattle</p></div></td><td><div><p>Fauntleroy</p></div></td><td><div><p>Endolyne, Arroyo Heights, Brace Point</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Fauntleroy Park"
 "Arroyos Natural Area"
 "Stim Bullitt Natural Area"
 "Endolyne Park"
 "Fauntleroy Creek Ravine"
 "Lincoln Park"
 "Kilbourne Park"
 "Solstice Park"]
</code></pre></div></div></td><td><div><p>3.3881863871407755E7</p></div></td><td><div><p>7399817.180664994</p></div></td><td><div><p>0.2184005345381709</p></div></td></tr><tr><td><div><p>Beacon Hill</p></div></td><td><div><p>North Beacon Hill</p></div></td><td><div><p>Jefferson Park</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Cheasty GS: Cheasty Blvd"
 "Jefferson Park Golf Course"
 "Cheasty Boulevard"
 "Jefferson Park"
 "East Duwamish GS: Beacon Ave S"
 "Twelfth Avenue South Viewpoint"
 "Stevens Place"
 "Cheasty Greenspace: McClellan"
 "East Duwamish GS: I-90"
 "Beacon Hill Playground"
 "Future Sister City Park"
 "Katie Black's Garden"
 "Daejeon Park"
 "Lewis Park"
 "Dr. Jose Rizal Park"
 "Sturgus Park"]
</code></pre></div></div></td><td><div><p>4.847591445665426E7</p></div></td><td><div><p>1.046069634381901E7</p></div></td><td><div><p>0.2157916247907537</p></div></td></tr><tr><td><div><p>Delridge</p></div></td><td><div><p>Highland Park</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["West Duwamish GS: Othello St."
 "Westcrest Park"
 "Highland Park Playground"]
</code></pre></div></div></td><td><div><p>3.981033341691431E7</p></div></td><td><div><p>8327022.199101201</p></div></td><td><div><p>0.20916735642217155</p></div></td></tr><tr><td><div><p>Capitol Hill</p></div></td><td><div><p>Madison Park</p></div></td><td><div><p>Broadmoor, Washington Park</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Lambert Place"
 "Hyde Place"
 "Madison Park"
 "Madison Park North Beach"
 "Washington Park and Arboretum"
 "Lake Washington Boulevard"
 " Lake Washington Boulevard"]
</code></pre></div></div></td><td><div><p>2.4680950222860403E7</p></div></td><td><div><p>4675806.696395554</p></div></td><td><div><p>0.18945002741688</p></div></td></tr><tr><td><div><p>Rainier Valley</p></div></td><td><div><p>Mount Baker</p></div></td><td><div><p>North Rainier</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["S Charlestown & 35th Ave S"
 "Stan Sayres Memorial Park"
 "Hunter Boulevard"
 "Horton Hil Corridor"
 "Lakewood Triangle"
 "Trudys Triangle"
 "Mount Baker Park"
 "Lake Washington Boulevard"
 "Landing Parkway"
 "Sierra Place"
 "Mount Claire Park"
 "Colman Park"
 "East Portal Viewpoint"
 "South Day Street Park"
 "York Playground"
 "Mount Baker Boulevard"
 "Martin Luther King JR Memorial"
 "College Street Park"
 "Amy Yee Tennis Center"
 "Bradner Gardens Park"
 "Mt Baker Ridge Viewpoint"
 "Sam Smith Park"]
</code></pre></div></div></td><td><div><p>2.8986995235084333E7</p></div></td><td><div><p>4682493.1951638255</p></div></td><td><div><p>0.16153772259555838</p></div></td></tr><tr><td><div><p>Delridge</p></div></td><td><div><p>High Point</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Walt Hundley Playfield"
 "Longfellow Creek GS: South"
 "Longfellow Creek GS: Central"
 "Camp Long"]
</code></pre></div></div></td><td><div><p>2.3343008337069765E7</p></div></td><td><div><p>3379054.5025715176</p></div></td><td><div><p>0.14475659922570583</p></div></td></tr><tr><td><div><p>Northwest</p></div></td><td><div><p>Broadview</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Carkeek Park"
 "Pipers Creek Natural Area"
 "Llandover Woods Greenspace"]
</code></pre></div></div></td><td><div><p>4.8113449731243916E7</p></div></td><td><div><p>6789015.469879707</p></div></td><td><div><p>0.14110431714629385</p></div></td></tr><tr><td><div><p>Central Area</p></div></td><td><div><p>Leschi</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Lake Washington Boulevard"
 "Frink Park"
 "Flo Ware Park"
 "Leschi Park"
 "Peppi's Playground"
 "Leschi-Lake Dell Natural Area"
 "Leschi Park and South Moorage"
 "North Leschi Moorage"
 "Madrona Park"
 "Powell Barnett Park"]
</code></pre></div></div></td><td><div><p>1.8091741377063915E7</p></div></td><td><div><p>2488664.0659542754</p></div></td><td><div><p>0.13755801689213387</p></div></td></tr><tr><td><div><p>West Seattle</p></div></td><td><div><p>Alki</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Me-Kwa-Mooks Natural Area"
 "Me-Kwa-Mooks Park"
 "Emma Schmitz Memorial Overlook"
 "Andover Place"
 "Cormorant Cove"
 "Charles Richey Sr Viewpoint"
 "Schmitz Preserve Park"
 "Schmitz Boulevard"
 "Duwamish Head Greenbelt"
 "Alki Beach Park"
 "Bar-S Playground"
 "Alki Playground"
 "Hamilton Viewpoint Park"]
</code></pre></div></div></td><td><div><p>2.45834691846126E7</p></div></td><td><div><p>3368610.236147073</p></div></td><td><div><p>0.13702745576102698</p></div></td></tr><tr><td><div><p>Capitol Hill</p></div></td><td><div><p>Stevens</p></div></td><td><div><p>Interlaken Park, Miller Park, Madison-Miller</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Miller Triangle"
 "Seven Hills Park"
 "Washington Park - Japanese Gdn"
 " Lake Washington Boulevard"
 "Washington Park and Arboretum"
 "Washington Park"
 "Lake Washington Boulevard"
 "Pendleton Miller Playfield"
 "Prentis I. Frazier Park"
 "Interlaken Park"
 "Boren Park"]
</code></pre></div></div></td><td><div><p>2.6039592079779927E7</p></div></td><td><div><p>3257296.7380544366</p></div></td><td><div><p>0.12509015994086056</p></div></td></tr><tr><td><div><p>Ballard</p></div></td><td><div><p>Sunset Hill</p></div></td><td><div><p>Golden Gardens, Shilshole</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["NW 60th Street Viewpoint"
 "Webster Park"
 "Sunset Hill Park"
 "Golden Gardens Park"]
</code></pre></div></div></td><td><div><p>2.4335620469794728E7</p></div></td><td><div><p>2999062.204514957</p></div></td><td><div><p>0.12323754835991878</p></div></td></tr><tr><td><div><p>West Seattle</p></div></td><td><div><p>North Admiral</p></div></td><td><div><p>Belvidere</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Hiawatha Playfield"
 "Schmitz Preserve Park"
 "Schmitz Boulevard"
 "Duwamish Head Greenbelt"
 "Fairmount Park"
 "Belvedere Park"
 "Nantes Park"
 "College Street Ravine"
 "California Place"
 "Hamilton Viewpoint Park"]
</code></pre></div></div></td><td><div><p>4.858141632113909E7</p></div></td><td><div><p>5960598.691665671</p></div></td><td><div><p>0.12269297898324247</p></div></td></tr><tr><td><div><p>Rainier Valley</p></div></td><td><div><p>Columbia City</p></div></td><td><div><p>Columbia Heights</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Hitt's Hill Park"
 "Brighton Playfield"
 "Hillman P-Patch"
 "Genesee Park and Playfield"
 "Lake People Park (Xacua'bs)"
 "Cheasty GS: Cheasty Blvd"
 "Cheasty Boulevard"
 "York Park"
 "Columbia Park"
 "Rainier Playfield"]
</code></pre></div></div></td><td><div><p>3.7406576493219726E7</p></div></td><td><div><p>4344458.806638756</p></div></td><td><div><p>0.11614157760270384</p></div></td></tr><tr><td><div><p>Central Area</p></div></td><td><div><p>Atlantic</p></div></td><td><div><p>Judkins Park, Jackson Place, Coleman</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Colman Playground"
 "Martin Luther King JR Memorial"
 "Amy Yee Tennis Center"
 "Jimi Hendrix Park"
 "Atlantic Street Park"
 "Sam Smith Park"
 "Benvenuto Viewpoint"
 "Judkins Park and Playfield"
 "Judge Charles M. Stokes Overlook"
 "Pratt Park"
 "Blue Dog Pond"
 "Dr. Blanche Lavizzo Park"
 "Langston Hughes Cultural Arts"]
</code></pre></div></div></td><td><div><p>2.0664986273336366E7</p></div></td><td><div><p>2395983.464898887</p></div></td><td><div><p>0.11594411112628604</p></div></td></tr><tr><td><div><p>Queen Anne</p></div></td><td><div><p>West Queen Anne</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Kinnear Park"
 "Marshall Park"
 "Parsons Gardens"
 "Queen Anne Boulevard"
 "SW Queen Anne Greenbelt"
 "West Queen Anne Playfield"
 "Queen Anne Swimming Pool"
 "West Central Grounds Maintenance"
 "12th West & West Howe Park"]
</code></pre></div></div></td><td><div><p>1.797498595288394E7</p></div></td><td><div><p>1964455.447449051</p></div></td><td><div><p>0.10928828832457976</p></div></td></tr><tr><td><div><p>Capitol Hill</p></div></td><td><div><p>Broadway</p></div></td><td><div><p>Pike/Pine</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Plymouth Pillars Park"
 "Bellevue Place"
 "Lakeview Place"
 "Cal Anderson Park"
 "Thomas Street Gardens"
 "Tashkent Park"
 "Williams Place"
 "Broadway Hill Park"
 "Volunteer Parkway"
 "Summit Slope Park"
 "Thomas Street Mini Park"
 "Belmont Place"
 "Summit Place"
 "Bullitt Life Estate"
 "St. Marks Greenbelt"
 "Streissguth Gardens"
 "I-5 Colonnade"
 "Volunteer Park"]
</code></pre></div></div></td><td><div><p>3.0239240521690477E7</p></div></td><td><div><p>3304530.7655596808</p></div></td><td><div><p>0.10927955558901538</p></div></td></tr><tr><td><div><p>Downtown</p></div></td><td><div><p>Pike-Market</p></div></td><td><div><p>Pike Place Market</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Seattle Aquarium"
 "Piers 62 and 63"
 "Waterfront Park"
 "Victor Steinbrueck Park"]
</code></pre></div></div></td><td><div><p>2343359.3260714533</p></div></td><td><div><p>255005.04346191516</p></div></td><td><div><p>0.10882029086397978</p></div></td></tr><tr><td><div><p>Magnolia</p></div></td><td><div><p>Briarcliff</p></div></td><td><div><p>Carleton Park</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Magnolia Tidelands Park"
 "Magnolia Park"
 "Magnolia Boulevard"
 "Parkmont Place"
 "Eastmont Place"
 "Carleton Center"
 "Carleton Highlands"
 "ARAPAHOE TRIANGLE"
 "Discovery Park"]
</code></pre></div></div></td><td><div><p>2.6550936578728132E7</p></div></td><td><div><p>2713294.302010858</p></div></td><td><div><p>0.10219203733041468</p></div></td></tr><tr><td><div><p>Interbay</p></div></td><td><div><p>Interbay</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Myrtle Edwards Park"
 "Ursula Judkins Viewpoint"
 "Smith Cove Park"
 "Magnolia Boulevard"
 "Magnolia Greenbelt"
 "Thorndyke Park"
 "Open Water Park"
 "Interbay P-patch"
 "Interbay Golf"
 "Interbay Athletic Field"
 "Commodore Park"]
</code></pre></div></div></td><td><div><p>3.350394109364654E7</p></div></td><td><div><p>3234156.7950585634</p></div></td><td><div><p>0.09653063757540652</p></div></td></tr><tr><td><div><p>Rainier Valley</p></div></td><td><div><p>Dunlap</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Chinook Beach Park"
 "Atlantic City Boat Ramp"
 "Rainier Beach Urban Farm and Wetlands"
 "Beer Sheva Park"
 "Rainier Beach Community Center"
 "Pritchard Island Beach"
 "East Duwamish GS: MLK Way"
 "Rainier Beach Playfield"]
</code></pre></div></div></td><td><div><p>1.9601601215604942E7</p></div></td><td><div><p>1782788.7703944917</p></div></td><td><div><p>0.0909511805073967</p></div></td></tr><tr><td><div><p>Central Area</p></div></td><td><div><p>Madrona</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Nora's Woods"
 "Madrona Playground"
 "Madrona Park"
 "Lake Washington Boulevard"
 "Madrona Ravine"
 "Alvin Larkins Park"
 "Madrona Briar Patch"
 "Howell Park"
 "William Grose Park"
 "Harrison Ridge Greenbelt"]
</code></pre></div></div></td><td><div><p>1.3845482808694465E7</p></div></td><td><div><p>1211682.0145605851</p></div></td><td><div><p>0.08751460901021757</p></div></td></tr><tr><td><div><p>Magnolia</p></div></td><td><div><p>Southeast Magnolia</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Magnolia Boulevard"
 "Condon Way Centerstrip"
 "Magnolia Park"
 "West Magnolia Playfield"
 " "
 "Lowery C. \"Pop\" Mounger Pool"
 "Ella Bailey Park"
 "Bayview Playground"]
</code></pre></div></div></td><td><div><p>1.8469263888735063E7</p></div></td><td><div><p>1466342.6454210556</p></div></td><td><div><p>0.07939367016762483</p></div></td></tr><tr><td><div><p>Beacon Hill</p></div></td><td><div><p>Holly Park</p></div></td><td><div><p>New Holly</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["John C. Little, Sr. Park" "Van Asselt Playground"]
</code></pre></div></div></td><td><div><p>8444235.979074124</p></div></td><td><div><p>656778.4801970643</p></div></td><td><div><p>0.07777831905984671</p></div></td></tr><tr><td><div><p>Beacon Hill</p></div></td><td><div><p>South Beacon Hill</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["East Duwamish GS: S Chicago St" "Benefit Playground"]
</code></pre></div></div></td><td><div><p>3.844882690015128E7</p></div></td><td><div><p>2989598.1245943466</p></div></td><td><div><p>0.07775524939572562</p></div></td></tr><tr><td><div><p>Rainier Valley</p></div></td><td><div><p>Rainier View</p></div></td><td><div><p>Lake Ridge</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Lakeridge Park"
 "Taylor Creek Headwaters"
 "Lakeridge Playground"
 "Thayer Place"]
</code></pre></div></div></td><td><div><p>1.8248566179783218E7</p></div></td><td><div><p>1398373.180214077</p></div></td><td><div><p>0.07662920836834146</p></div></td></tr><tr><td><div><p>Queen Anne</p></div></td><td><div><p>East Queen Anne</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Queen Anne Boulevard"
 "Bhy Kracke Park"
 "Northeast Queen Anne Greenbelt"
 "Highland Place"
 "Highland Drive Parkway"
 "Observatory Courts"
 "East Queen Anne Playground"
 "Wolf Creek Ravine Natural Area"
 "Trolley Hill Park"
 "MacLean Park"
 "Westlake Greenbelt"
 "Thomas C. Wales Park"]
</code></pre></div></div></td><td><div><p>1.9354699596494563E7</p></div></td><td><div><p>1448114.7496615453</p></div></td><td><div><p>0.07481979983423878</p></div></td></tr><tr><td><div><p>Rainier Valley</p></div></td><td><div><p>Rainier Beach</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Kubota Gardens Natural Area"
 "Lakeridge Park"
 "Lakeridge Playground"
 "Thayer Place"
 "Kubota Gardens"
 "Fred Hutchinson Playground"
 "Sturtevant Ravine"
 "East Duwamish GS: MLK Way"
 "Fletcher Place"
 "Chinook Beach Park"]
</code></pre></div></div></td><td><div><p>3.312823614223316E7</p></div></td><td><div><p>2442620.8848146023</p></div></td><td><div><p>0.07373229514325559</p></div></td></tr><tr><td><div><p>Central Area</p></div></td><td><div><p>Harrison/Denny-Blaine</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Denny Blaine Lake Park"
 "Harrison Ridge Greenbelt"
 "Viretta Park"
 "Stevens Triangle"
 "Denny Blaine Park"
 "Lakeview Park"
 "McGilvra Boulevard"
 "Lake Washington Boulevard"
 "3001 E Madison"]
</code></pre></div></div></td><td><div><p>1.360339567318952E7</p></div></td><td><div><p>946560.209820138</p></div></td><td><div><p>0.06958264190504156</p></div></td></tr><tr><td><div><p>Northeast</p></div></td><td><div><p>Ravenna</p></div></td><td><div><p>University Village</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Blakeley Crescent Park"
 "Ravenna Ravine"
 "Burke-Gilman Trail"
 "Ravenna Park"
 "Ravenna-Eckstein Park"]
</code></pre></div></div></td><td><div><p>3.678339153768188E7</p></div></td><td><div><p>2537736.5238547004</p></div></td><td><div><p>0.06899136859783514</p></div></td></tr><tr><td><div><p>Lake City</p></div></td><td><div><p>Meadowbrook</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Thornton Creek Natural Area: 25th/95th"
 "Meadowbrook Playfield"
 "Meadowbrook Teen Life Center"
 "LaVilla Meadows Natural Area on Thornton Creek"]
</code></pre></div></div></td><td><div><p>1.703484877942584E7</p></div></td><td><div><p>1162909.7193452865</p></div></td><td><div><p>0.06826651262967551</p></div></td></tr><tr><td><div><p>Northwest</p></div></td><td><div><p>North Beach/Blue Ridge</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Soundview Playfield"
 "Blue Ridge Places"
 "Carkeek Park"
 "Blue Ridge Circle"
 "North Beach Park"
 "North Beach Ravine: 30th NW Natural Area"
 "Golden Gardens Park"]
</code></pre></div></div></td><td><div><p>3.1062264992039643E7</p></div></td><td><div><p>1942086.5722656162</p></div></td><td><div><p>0.06252237474515514</p></div></td></tr><tr><td><div><p>Northeast</p></div></td><td><div><p>Roosevelt</p></div></td><td><div><p>Fairview</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Ravenna Boulevard"
 "Ravenna Park"
 "Cowen Park"
 "Roosevelt Land-Bank Site"
 "Banner Way Triangles"
 "Froula Playground"
 "Rainbow Point"
 "Maple Leaf Reservoir Park"]
</code></pre></div></div></td><td><div><p>2.1491594324983705E7</p></div></td><td><div><p>1309062.5613024468</p></div></td><td><div><p>0.06091044440480054</p></div></td></tr><tr><td><div><p>Lake City</p></div></td><td><div><p>Matthews Beach</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Mock Creek Ravine"
 "Little Brook Creek Natural Area"
 "Chinook Passage Natural Area on Thornton Creek"
 "Burke-Gilman Trail"
 "Matthews Beach Park"]
</code></pre></div></div></td><td><div><p>3.0980978301755752E7</p></div></td><td><div><p>1770241.0308238121</p></div></td><td><div><p>0.05713961042745668</p></div></td></tr><tr><td><div><p>Northeast</p></div></td><td><div><p>Bryant</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Burke-Gilman Trail"
 "Burke-Gilman Playground Park"
 "Bryant Playground"]
</code></pre></div></div></td><td><div><p>1.5485078033882016E7</p></div></td><td><div><p>829201.5932133237</p></div></td><td><div><p>0.053548428454735265</p></div></td></tr><tr><td><div><p>Delridge</p></div></td><td><div><p>Roxhill</p></div></td><td><div><p>Westwood Village</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Roxhill Park"
 "S.W. Community Center and Pool"
 "E.C. Hughes Playground"]
</code></pre></div></div></td><td><div><p>1.87810598448122E7</p></div></td><td><div><p>995987.7545642204</p></div></td><td><div><p>0.053031498903366586</p></div></td></tr><tr><td><div><p>Northeast</p></div></td><td><div><p>View Ridge</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Burke-Gilman Trail" "View Ridge Playfield" "Inverness Ravine Park"]
</code></pre></div></div></td><td><div><p>2.561297994427941E7</p></div></td><td><div><p>1288763.708935325</p></div></td><td><div><p>0.05031682028951758</p></div></td></tr><tr><td><div><p>Cascade</p></div></td><td><div><p>South Lake Union</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Denny Park"
 "Lake Union Park"
 "Fairview Walkway"
 "Cascade Playground"
 "I-5 Colonnade"
 "Terry Pettus Park"
 "Lynn Street Mini Park"]
</code></pre></div></div></td><td><div><p>1.6107105419750454E7</p></div></td><td><div><p>800866.192857775</p></div></td><td><div><p>0.04972129827099516</p></div></td></tr><tr><td><div><p>Queen Anne</p></div></td><td><div><p>North Queen Anne</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Queen Anne Boulevard"
 "Coe Play Park"
 "Queen Anne Bowl Playfield"
 "Mayfair Park"
 "David Rodgers Park"
 "West Ewing Mini Park"
 "Soundview Terrace"]
</code></pre></div></div></td><td><div><p>2.960428833366627E7</p></div></td><td><div><p>1355843.879844315</p></div></td><td><div><p>0.04579890131330861</p></div></td></tr><tr><td><div><p>North Central</p></div></td><td><div><p>Wallingford</p></div></td><td><div><p>Meridian, Tangle Town, Northlake</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Gas Works Park"
 "Harbor Patrol Office and Shop"
 "Waterway 19"
 "Wallingford Steps"
 "Burke-Gilman Trail"
 "Sunnyside Ave N Boat Ramp"
 "Wallingford Playfield"
 "Demonstration Garden"
 "North Passage Point Park"
 "NE 60th Street Park"
 "Meridian Playground"
 "Corliss Place"
 "Keystone Place"
 "Green Lake Park"]
</code></pre></div></div></td><td><div><p>4.2068342614797205E7</p></div></td><td><div><p>1777393.4021316427</p></div></td><td><div><p>0.04225014088162481</p></div></td></tr><tr><td><div><p>Lake City</p></div></td><td><div><p>Victory Heights</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Homewood Natural Area on Thornton Creek"
 "Kingfisher Natural Area on Thornton Creek"
 "Virgil Flaim Park"
 "Lake City Park"
 "Victory Heights Playground"
 "Pinehurst Pocket Park"]
</code></pre></div></div></td><td><div><p>2.151273148967749E7</p></div></td><td><div><p>900485.900698212</p></div></td><td><div><p>0.04185827825398623</p></div></td></tr><tr><td><div><p>Beacon Hill</p></div></td><td><div><p>Mid-Beacon Hill</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Dearborn Park"
 "Cheasty Greenspace: Mt View"
 "Maple School Ravine"
 "Maple Wood Playfield"
 "Cleveland Playfield"
 "East Duwamish GS: S Spokane St"]
</code></pre></div></div></td><td><div><p>5.262047078955701E7</p></div></td><td><div><p>2161142.3873948674</p></div></td><td><div><p>0.041070373468109767</p></div></td></tr><tr><td><div><p>Central Area</p></div></td><td><div><p>Mann</p></div></td><td><div><p>Central District, Garfield</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Garfield Teen Life Center"
 "Garfield Playfield"
 "Homer Harris Park"
 "Gerber Park"
 "Plum Tree Park"
 "Julia Lee's Park"]
</code></pre></div></div></td><td><div><p>1.113382886245558E7</p></div></td><td><div><p>443902.819098556</p></div></td><td><div><p>0.03986973615118534</p></div></td></tr><tr><td><div><p>Cascade</p></div></td><td><div><p>Eastlake</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["I-5 Colonnade"
 "Fairview Park"
 "South Passage Point Park"
 "Rogers Playground"
 "Roanoke Street Mini Park"
 "Terry Pettus Park"
 "Lynn Street Mini Park"]
</code></pre></div></div></td><td><div><p>7426907.961769626</p></div></td><td><div><p>293104.7959594849</p></div></td><td><div><p>0.039465252224513386</p></div></td></tr><tr><td><div><p>Northgate</p></div></td><td><div><p>Maple Leaf</p></div></td><td><div><p>Olympic View</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Sacajawea Playground"
 "Maple Leaf Community Garden"
 "Beaver Pond Natural Area on Thornton Creek"
 "Maple Leaf Reservoir Park"
 "Kingfisher Natural Area on Thornton Creek"
 "Victory Creek Confluence Natural Area on TC"
 "Northgate Park"
 "Northgate Community Center"]
</code></pre></div></div></td><td><div><p>3.4380470783073254E7</p></div></td><td><div><p>1130968.2364097128</p></div></td><td><div><p>0.03289565880425725</p></div></td></tr><tr><td><div><p>Northeast</p></div></td><td><div><p>Laurelhurst</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Union Bay Boglands #2"
 "Union Bay Boglands #1"
 "LAURELHURST TRIANGLE"
 "Belvoir Place"
 "Laurelhurst Playfield"
 "Laurelhurst Recreation Center"
 "Sun Park"]
</code></pre></div></div></td><td><div><p>1.905177681936156E7</p></div></td><td><div><p>622689.0135556267</p></div></td><td><div><p>0.03268403884108136</p></div></td></tr><tr><td><div><p>Northeast</p></div></td><td><div><p>Wedgwood</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Dahl (Waldo J.) Playfield"
 "Picardo (Rainie) P-Patch"
 "Thornton Creek Natural Area: 25th/95th"
 "Wedgwood Square"
 "8605 35th Ave NE"
 "Inverness Ravine Park"
 "Maple Springs Natural Area"]
</code></pre></div></div></td><td><div><p>3.059971362703256E7</p></div></td><td><div><p>968985.0098010622</p></div></td><td><div><p>0.031666473144540686</p></div></td></tr><tr><td><div><p>Downtown</p></div></td><td><div><p>Yesler Terrace</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Yesler Terrace Park" "Horiuchi Park" "Boren Place"]
</code></pre></div></div></td><td><div><p>5348937.247801803</p></div></td><td><div><p>159300.53646585828</p></div></td><td><div><p>0.02978171720584764</p></div></td></tr><tr><td><div><p>Downtown</p></div></td><td><div><p>Belltown</p></div></td><td><div><p>Denny Regrade</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Myrtle Edwards Park"
 "Bell Street Park Boulevard"
 "Belltown Cottage Park"
 "Belltown P-patch"
 "Regrade Park"
 "Alaskan Way Boulevard"
 "Tilikum Place"]
</code></pre></div></div></td><td><div><p>9845864.786277873</p></div></td><td><div><p>279689.8893769705</p></div></td><td><div><p>0.028406838347686095</p></div></td></tr><tr><td><div><p>Capitol Hill</p></div></td><td><div><p>Portage Bay</p></div></td><td><div><p>Roanoke Park</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Roanoke Park" "Bagley Viewpoint"]
</code></pre></div></div></td><td><div><p>4421635.267367266</p></div></td><td><div><p>121026.6277003076</p></div></td><td><div><p>0.027371463357349548</p></div></td></tr><tr><td><div><p>Greater Duwamish</p></div></td><td><div><p>South Park</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["South Park Meadow"
 "Marra-Desimone Park"
 "Cesar Chavez Park"
 "South Park Playground"
 "Duwamish Waterway Park"]
</code></pre></div></div></td><td><div><p>2.9423452152724065E7</p></div></td><td><div><p>788757.1095861244</p></div></td><td><div><p>0.0268070893072654</p></div></td></tr><tr><td><div><p>West Seattle</p></div></td><td><div><p>Arbor Heights</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Seola Park" "Seola Beach Greenbelt" "Fauntleroy Park"]
</code></pre></div></div></td><td><div><p>2.619731405785146E7</p></div></td><td><div><p>661354.2410716552</p></div></td><td><div><p>0.0252451163356361</p></div></td></tr><tr><td><div><p>Lake City</p></div></td><td><div><p>Cedar Park</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Lake City Memorial Triangle"
 "Burke-Gilman Trail"
 "Lake City Urban Hub Village"
 "Cedar Park"
 "University Lake Shore Place"
 "NE 130th Street End"]
</code></pre></div></div></td><td><div><p>2.04414858490784E7</p></div></td><td><div><p>500421.66464288835</p></div></td><td><div><p>0.02448068933626221</p></div></td></tr><tr><td><div><p>University District</p></div></td><td><div><p>University District</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Burke-Gilman Trail"
 "Ravenna Woods"
 "Christie Park"
 "17th Avenue NE Centerstrip"
 "North Passage Point Park"
 "Northlake Park"
 "University Playground"]
</code></pre></div></div></td><td><div><p>1.3686254858800735E7</p></div></td><td><div><p>324977.58184178185</p></div></td><td><div><p>0.023744814428383235</p></div></td></tr><tr><td><div><p>Northgate</p></div></td><td><div><p>Licton Springs</p></div></td><td><div><p>North College Park</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Mineral Springs Park" "Licton Springs Park"]
</code></pre></div></div></td><td><div><p>2.1704706262052845E7</p></div></td><td><div><p>513945.6292462934</p></div></td><td><div><p>0.023678994916639065</p></div></td></tr><tr><td><div><p>Northgate</p></div></td><td><div><p>Haller Lake</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Haller Lake Community St End" "Northacres Park" "Helene Madison Pool"]
</code></pre></div></div></td><td><div><p>4.720308108008111E7</p></div></td><td><div><p>1112625.044392759</p></div></td><td><div><p>0.023571025851155034</p></div></td></tr><tr><td><div><p>Northeast</p></div></td><td><div><p>Windermere</p></div></td><td><div><p>Hawthorne Hills</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Burke-Gilman Trail" "University Circle" "Warren G. Magnuson Park"]
</code></pre></div></div></td><td><div><p>2.2363966623912185E7</p></div></td><td><div><p>500567.5192937572</p></div></td><td><div><p>0.022382769913389885</p></div></td></tr><tr><td><div><p>Ballard</p></div></td><td><div><p>Loyal Heights</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Salmon Bay Park" "Loyal Heights Playfield"]
</code></pre></div></div></td><td><div><p>2.1320655545884836E7</p></div></td><td><div><p>421263.7501266854</p></div></td><td><div><p>0.019758480184629914</p></div></td></tr><tr><td><div><p>Downtown</p></div></td><td><div><p>International District</p></div></td><td><div><p>Chinatown, ID, Little Saigon</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Donnie Chin International Children's Park"
 "Hing Hay Park Addition"
 "ID/Chinatown Community Center"
 "Beacon Place"
 "Chinatown/ID Hub Urban Village"
 "Hing Hay Park"
 "Kobe Terrace"]
</code></pre></div></div></td><td><div><p>6169375.816795917</p></div></td><td><div><p>121130.99413208748</p></div></td><td><div><p>0.019634238167548886</p></div></td></tr><tr><td><div><p>Rainier Valley</p></div></td><td><div><p>Brighton</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Othello Park"]
</code></pre></div></div></td><td><div><p>1.8366938275487844E7</p></div></td><td><div><p>342712.9789050467</p></div></td><td><div><p>0.01865923289797433</p></div></td></tr><tr><td><div><p>Downtown</p></div></td><td><div><p>Central Business District</p></div></td><td><div><p>Commercial Core, West Edge</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Waterfront Park" "Jim Ellis Freeway Park" "Westlake Park"]
</code></pre></div></div></td><td><div><p>9477398.618861396</p></div></td><td><div><p>176559.42602316735</p></div></td><td><div><p>0.018629524105041706</p></div></td></tr><tr><td><div><p>North Central</p></div></td><td><div><p>Fremont</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["A. B. Ernst Park"
 "Fremont Canal Park"
 "Burke-Gilman Trail"
 "Fremont Peak Park"
 "Troll's Knoll Park"
 "B.F. Day Playground"
 "Ross Playground"]
</code></pre></div></div></td><td><div><p>2.650375657209128E7</p></div></td><td><div><p>443605.6587439842</p></div></td><td><div><p>0.016737463519081116</p></div></td></tr><tr><td><div><p>Northwest</p></div></td><td><div><p>Bitter Lake</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Bitter Lake Playfield" "Bitter Lake Open Space Park"]
</code></pre></div></div></td><td><div><p>2.769033644488096E7</p></div></td><td><div><p>463326.8142235621</p></div></td><td><div><p>0.016732437149900217</p></div></td></tr><tr><td><div><p>Northwest</p></div></td><td><div><p>Greenwood</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Alice Ball Park"
 "Greenwood Park"
 "6th Avenue NW Pocket Park"
 "Sandel Playground"
 "Carkeek Park"]
</code></pre></div></div></td><td><div><p>4.204456367454267E7</p></div></td><td><div><p>689596.1664853458</p></div></td><td><div><p>0.01640155364254346</p></div></td></tr><tr><td><div><p>University District</p></div></td><td><div><p>University Heights</p></div></td><td><div><p>Cowen Park</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["17th Avenue NE Centerstrip"
 "Ravenna Boulevard"
 "Park Home Circle"
 "University Heights"
 "Ravenna Park"]
</code></pre></div></div></td><td><div><p>1.037574805292391E7</p></div></td><td><div><p>166849.97798407465</p></div></td><td><div><p>0.016080766141681316</p></div></td></tr><tr><td><div><p>West Seattle</p></div></td><td><div><p>Gatewood</p></div></td><td><div><p>Morgan Junction</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Orchard Street Ravine"
 "Myrtle Reservoir Park"
 "Lincoln Park"
 "Lowman Beach Park"
 "Pelly Place Natural Area"
 "Watton Life Estate"
 nil]
</code></pre></div></div></td><td><div><p>2.2821707469523415E7</p></div></td><td><div><p>355085.68614532915</p></div></td><td><div><p>0.015559120044787095</p></div></td></tr><tr><td><div><p>West Seattle</p></div></td><td><div><p>Fairmount Park</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Fairmount Playground"]
</code></pre></div></div></td><td><div><p>1.6153900097573552E7</p></div></td><td><div><p>237376.26008061314</p></div></td><td><div><p>0.014694671791134142</p></div></td></tr><tr><td><div><p>Downtown</p></div></td><td><div><p>First Hill</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Jim Ellis Freeway Park" "First Hill Park" "Boylston Place"]
</code></pre></div></div></td><td><div><p>9517568.473671103</p></div></td><td><div><p>126769.23408250557</p></div></td><td><div><p>0.0133194979824094</p></div></td></tr><tr><td><div><p>Downtown</p></div></td><td><div><p>Pioneer Square</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Union Station Square"
 "Occidental Square"
 "City Hall Park"
 "Prefontaine Place"
 "Pioneer Square"]
</code></pre></div></div></td><td><div><p>7545979.087288558</p></div></td><td><div><p>95347.44250189843</p></div></td><td><div><p>0.0126355296508195</p></div></td></tr><tr><td><div><p>Ballard</p></div></td><td><div><p>Ballard</p></div></td><td><div><p>Adams</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Bergen Place"
 "Ballard Corners Park"
 "Marvins Garden"
 "Ballard Commons Park"
 "Thyme Patch Park"
 "Ballard Playground"]
</code></pre></div></div></td><td><div><p>2.255213297047855E7</p></div></td><td><div><p>244344.9612760165</p></div></td><td><div><p>0.01083467189537556</p></div></td></tr><tr><td><div><p>Cascade</p></div></td><td><div><p>Westlake</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Lake Union Park"]
</code></pre></div></div></td><td><div><p>3645184.079110308</p></div></td><td><div><p>37961.61985636592</p></div></td><td><div><p>0.010414184587800388</p></div></td></tr><tr><td><div><p>Ballard</p></div></td><td><div><p>West Woodland</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Greenwood Triangle"
 "14th Avenue NW Boat Ramp"
 "Gilman Playground"
 "Gemenskap Park"]
</code></pre></div></div></td><td><div><p>2.2199366873893585E7</p></div></td><td><div><p>226099.72709449235</p></div></td><td><div><p>0.010184962858575269</p></div></td></tr><tr><td><div><p>Central Area</p></div></td><td><div><p>Minor</p></div></td><td><div><p>Central District, Squire Park</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Spruce Street Mini Park"
 "Garfield Playfield"
 "Cayton Corner Park"
 "Langston Hughes Annex"
 "12th Ave Square Park"
 "Firehouse Mini Park"
 "Spring Street Mini Park"
 "T.T. Minor Playground"
 "McGilvra Place"]
</code></pre></div></div></td><td><div><p>1.7933144708720244E7</p></div></td><td><div><p>172158.54773067622</p></div></td><td><div><p>0.009600019992420052</p></div></td></tr><tr><td><div><p>Greater Duwamish</p></div></td><td><div><p>Industrial District</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Herrings House Park (Tulaltx)"
 "Alki Beach Park"
 "Seacrest Park"
 "Don Armeni Boat Ramp"]
</code></pre></div></div></td><td><div><p>6.243413101303743E7</p></div></td><td><div><p>559378.0064992205</p></div></td><td><div><p>0.00895949054504197</p></div></td></tr><tr><td><div><p>Ballard</p></div></td><td><div><p>Whittier Heights</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Kirke Park" "Ballard Swimming Pool" "Baker Park on Crown Hill"]
</code></pre></div></div></td><td><div><p>1.4195687492422596E7</p></div></td><td><div><p>108946.50009691996</p></div></td><td><div><p>0.007674619503639655</p></div></td></tr><tr><td><div><p>Delridge</p></div></td><td><div><p>South Delridge</p></div></td><td><div><p>White Center</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Longfellow Creek GS: South"]
</code></pre></div></div></td><td><div><p>1.792331622408078E7</p></div></td><td><div><p>132742.1724476857</p></div></td><td><div><p>0.007406116746929937</p></div></td></tr><tr><td><div><p>Queen Anne</p></div></td><td><div><p>Lower Queen Anne</p></div></td><td><div><p>Uptown, Seattle Center</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Blaine Place"
 "Counterbalance Park"
 "Kinnear Place"
 "Kerry Park (Franklin Place)"
 "Kerry Park and Viewpoint"
 "Ward Springs Park"]
</code></pre></div></div></td><td><div><p>1.7515624457317874E7</p></div></td><td><div><p>106952.95177903834</p></div></td><td><div><p>0.006106145518229259</p></div></td></tr><tr><td><div><p>Northwest</p></div></td><td><div><p>Crown Hill</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Crown Hill Glen" "Crown Hill Park"]
</code></pre></div></div></td><td><div><p>1.586575041985613E7</p></div></td><td><div><p>94943.40486711139</p></div></td><td><div><p>0.005984173603808166</p></div></td></tr><tr><td><div><p>University District</p></div></td><td><div><p>University of Washington</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Union Bay Boglands #2"
 "Union Bay Boglands #1"
 "Fritz Hedges Waterway Park"]
</code></pre></div></div></td><td><div><p>2.5523675124439575E7</p></div></td><td><div><p>142125.03524102044</p></div></td><td><div><p>0.005568360925614982</p></div></td></tr><tr><td><div><p>Greater Duwamish</p></div></td><td><div><p>Georgetown</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Georgetown Pump Station" "Oxbow Park" "Georgetown Playfield"]
</code></pre></div></div></td><td><div><p>5.130507264703818E7</p></div></td><td><div><p>281438.2271119965</p></div></td><td><div><p>0.005485582859383083</p></div></td></tr><tr><td><div><p>Lake City</p></div></td><td><div><p>Olympic Hills</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Thornton Creek Addition: North Branch"
 "Albert Davis Park"
 "Little Brook Park"]
</code></pre></div></div></td><td><div><p>2.5836742164596498E7</p></div></td><td><div><p>141195.81351923634</p></div></td><td><div><p>0.005464923271662082</p></div></td></tr><tr><td><div><p>West Seattle</p></div></td><td><div><p>Genesee</p></div></td><td><div><p>Alaska Junction, West Seattle Junction</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["West Seattle Junction"
 "Junction Plaza Park"
 "Ercolini Park"
 "Dakota Place Park"
 "48th Ave SW/SW Charlestown St"
 "Fauntleroy Place"]
</code></pre></div></div></td><td><div><p>2.1164976581792817E7</p></div></td><td><div><p>91891.1780685815</p></div></td><td><div><p>0.00434166216595916</p></div></td></tr><tr><td><div><p>West Seattle</p></div></td><td><div><p>Seaview</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["Lowman Beach Park" "Morgan Junction Park"]
</code></pre></div></div></td><td><div><p>1.846306179837895E7</p></div></td><td><div><p>51860.33722001134</p></div></td><td><div><p>0.002808869828110777</p></div></td></tr><tr><td><div><p>Downtown</p></div></td><td><div><p>Denny Triangle</p></div></td><td><div><p>Denny Regrade</p></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">["McGraw Square" "Westlake Square" "Urban Triangle Park"]
</code></pre></div></div></td><td><div><p>5128172.310711999</p></div></td><td><div><p>11668.172603196248</p></div></td><td><div><p>0.002275308218256073</p></div></td></tr><tr><td><div><p>Greater Duwamish</p></div></td><td><div><p>Harbor Island</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">[]
</code></pre></div></div></td><td><div><p>1.7832300735864956E7</p></div></td><td><div><p>0.0</p></div></td><td><div><p>0.0</p></div></td></tr><tr><td><div><p>Greater Duwamish</p></div></td><td><div><p>SODO</p></div></td><td><div></div></td><td><div style="max-height:600px;overflow-y:auto;"><div><pre><code class="language-clojure">[]
</code></pre></div></div></td><td><div><p>4.9603167591402024E7</p></div></td><td><div><p>0.0</p></div></td><td><div><p>0.0</p></div></td></tr></tbody></table></div>

<div style="background-color:grey;height:2px;width:100%;"></div>

<div><pre><small><small>source: projects/geography/seattle-parks/notebooks/index.clj</small></small></pre></div>