### 2024-01-25 docs/projects/math/wolframite/index.qmd
Wolframite - rendered doc

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: index.html}
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="index_files/md-default0.js" type="text/javascript"></script><script src="index_files/md-default1.js" type="text/javascript"></script>
<div style="max-height:600px;overflow-y:auto;"><table><tr><td>This is part of the Scicloj <a href="https://scicloj.github.io/clojure-data-scrapbook/">Clojure Data Scrapbook</a>.</td><a href="https://scicloj.github.io/clojure-data-scrapbook/"><img align="left" alt="SciCloj logo" src="https://scicloj.github.io/sci-cloj-logo-transparent.png" width="40" /></a></tr></table></div>

::: {.sourceClojure}
```clojure
(ns index
  (:refer-clojure
   ;; Exclude symbols also used by Wolfram:
   :exclude [Byte Character Integer Number Short String Thread])
  (:require
   [wolframite.core :as wl :refer [wl]]
   [wolframite.tools.hiccup]
   [wolframite.base.parse :as parse]
   [wolframite.jlink]
   [scicloj.kindly.v4.kind :as kind]
   [scicloj.kindly.v4.api :as kindly])
  (:import (com.wolfram.jlink MathCanvas KernelLink)
           (java.awt Color Frame)
           (java.awt.event WindowAdapter ActionEvent)))
```
:::

::: {.sourceClojure}
```clojure
(def md
  (comp kindly/hide-code kind/md))
```
:::

# Wolframite

This notebook demonstrates basic usage of [Wolframite](https://github.com/scicloj/wolframite/),

It is mostly copied and adpated from [the official demo](https://github.com/scicloj/wolframite/blob/main/dev/demo.clj).

## Init (base example)

::: {.sourceClojure}
```clojure
(wl/eval '(Dot [1 2 3] [4 5 6]))
```
:::

::: {.printedClojure}
```clojure
32

```
:::

## Strings of WL code

::: {.sourceClojure}
```clojure
(wl/eval "{1 , 2, 3} . {4, 5, 6}")
```
:::

::: {.printedClojure}
```clojure
32

```
:::

## Def // intern WL fns, i.e. effectively define WL fns as clojure fns:

::: {.sourceClojure}
```clojure
(def W:Plus (parse/parse-fn 'Plus {:kernel/link @wl/kernel-link-atom}))
```
:::

::: {.sourceClojure}
```clojure
(W:Plus 1 2 3)
```
:::

::: {.printedClojure}
```clojure
6

```
:::

... and call it

::: {.sourceClojure}
```clojure
(wl/clj-intern 'Plus {})
```
:::

::: {.printedClojure}
```clojure
#'index/Plus

```
:::

a simpler way to do the same -> fn Plus in this ns

::: {.sourceClojure}
```clojure
(map wl/clj-intern ['Dot 'Plus])
```
:::

::: {.printedClojure}
```clojure
(#'index/Dot #'index/Plus)

```
:::

 Call interned Wl functions:

::: {.sourceClojure}
```clojure
(Dot [1 2 3] [4 5 6])
```
:::

::: {.printedClojure}
```clojure
32

```
:::

::: {.sourceClojure}
```clojure
(Plus 1 2 3)
```
:::

::: {.printedClojure}
```clojure
6

```
:::

::: {.sourceClojure}
```clojure
(Plus [1 2] [3 4])
```
:::

::: {.printedClojure}
```clojure
[4 6]

```
:::

::: {.sourceClojure}
```clojure
(def greetings
  (wl/eval
   '(Function [x] (StringJoin "Hello, " x "! This is a Mathematica function's output."))))
```
:::

::: {.sourceClojure}
```clojure
(greetings "Stephen")
```
:::

::: {.printedClojure}
```clojure
"Hello, Stephen! This is a Mathematica function's output."

```
:::

 ## Bidirectional translation
(Somewhat experimental, especially in the wl->clj direction)

::: {.sourceClojure}
```clojure
(wl/->clj! "GridGraph[{5, 5}]")
```
:::

::: {.printedClojure}
```clojure
(GridGraph [5 5])

```
:::

::: {.sourceClojure}
```clojure
(wl/->wl! '(GridGraph [5 5]) {:output-fn str})
```
:::

::: {.printedClojure}
```clojure
"GridGraph[{5, 5}]"

```
:::

## Graphics

### A helper function to view things in tools supporting [Kindly](https://scicloj.github.io/kindly-noted/kindly)

::: {.sourceClojure}
```clojure
(defn view
  ([form]
   (view form nil))
  ([form {:keys [folded?]}]
   (-> form
       (wolframite.tools.hiccup/view* folded?)
       kind/hiccup)))
```
:::

 ### Draw Something!

::: {.sourceClojure}
```clojure
(view
 '(GridGraph [5 5]))
```
:::

<div style="max-height:600px;overflow-y:auto;"><div><details open="open"><summary><h5 style="cursor:pointer;display:inline;padding:0.5rem 1rem 0 1rem;">GridGraph[{5, 5}]</h5></summary><div class="wl-results"><hr /><img src="data:image/jpeg;base64,R0lGODlh4AHgAff4ADI2PjU5Pzc7QDY8Rjg7QTo9Qjw/Qzo+RTw/RDtART5BRT5DSDpDUDpEUj1GUj1GVDpHWj1HWj1IWT9LXD1MY0BDRkJER0BDSUFESkVHSUJGTUZIS0NIT0ZJTUlLTUpMT0VKUUdNVUtNUU5PUkhOVkBLW0lPWE5QUktRWkxSW0xTXE5VX1FSVVJVV1VXWVZYWldaXVhZW1pbXVtcXkJQZk9WYEVWb1BXYVBYYlJaZFRbZlRcZ1peY11fYFVdaUtbc0pcd0pceF5gY1dgbFliblpjcVtkcl1mc15ndV9odmFiY2FjZGNkZmJlaWZnaWlqa2tsbWxtbm1ucWtvdmBqeGJtemNtfGRvfmZwf3Bxc3NzdXN0dnV2d3Z3eHN3fHd4eXV4fXl5enp7fHt8fX1+f1Flg1dngF9wi1lvkGZxgGhygml0g2x3h2RziWZ0im56ind9hH9/gHl+hnJ+jm5+lmF4nGh7mXN/kXB/l2V9on6BhHSAknWClHiFlnmGmHuImn2KnX+Mn2eApmuDpm6GqnGGpn2LoHqLpXeNr4CAgYCChYOEhIaHh4eHiIOHjYeIiISIjomJioqLjIuMjo2NjoaMlo+PkIONnoSPno+QkIWQn5KSk5CTlpOUlJWWlpGUmJKXnpqamp2enoGOoYKQo4ORpIWTpoaUp4eVqYmXq4qZro2brpWdp56foJOcqZSdqpGdroKVsoSWso2bsY6dsomdupCftJGhtpOiuJSjuZWkupOjvJenvpqnu5movp+rvqGhoaOjpKampqOlqqenqKmpqa2traOosaKqtKGrurGxsbKytLa2trS3vLq6ur29vZipw5uswpysw5urxJ6uxp+wx6GyyqWzyKO0zKi2yam3za+6yqy6zrG5xLnE1bzH2MLCwsXFxcXGyMnJyc7OzsDI1cXP3cbQ3sjR3tLS0tXV1dnZ2d7e3srT4NTb5eHh4ebm5urq6u3t7eXp8PHx8fX19fb3+vr6+v///////////////////////////////yH/C05FVFNDQVBFMi4wAwEAAAAh+QQECgD/ACwAAAAA4AHgAQAI/wDxCRxIsKDBgwgTKlzIsKHDhxAjSpxIsaLFixgzatzIsaPHjyBDihxJsqTJkyhTqlzJsqXLlzBjypxJs6bNmzhz6tzJs6fPn0CDCh1KtKjRo0iTKl3KtKnTp1CjSp1KtarVq1izat3KtWa9VlFexBHX1eE9YllciHFW1qGyLi66KGvb0JkYF1mI3aO7UFycF1GA1aMIT0YINXOMKNjEFyE9JxyuzEmyIFHjg/W0YKgypwoGLYMvF4yzIMmcKxyc0BNdcJMCI3PUhJgBb2KUHdSw6Va14BnrgYlURNONDZcGYr8FegrBizivEJ6S4yOGARdxaSriSH+mQBVxajuiSP8kh6A5cWxqlCR/p2DVeWx3TiSvZyHQ+1EWQrM+cef9KgXvJLfEGu/xYgA5ERFDwnvYlBKALLVEKOGEFFZo4YUYZqjhhhxq6IYC1rxHiwCHdGjiiSimqKKEeAggzXvRCIDHijTWaOOJhwhAC4MKuHHjj0AGKUsApTBIAnIQKcMBg4EgcI02UEYp5ZRUVmnllVhmqeWWWQYCon8G9MLlmGSWaeaZUa5igHnOGbAKmnDGKSeZvRjg3XnWHHDJnHz26ec1Btj3HghzQbSOAra8l4R4v8VjwSjvrRFDcvd40N95d3hgT3IyqHGfBfEkF0USIiqwjkSLoGAdNtbcYUE60oX/wsGOuo2iADjSKYMBKsShgkGhv4GjAKS60cJBK9KlowAfIRZnwiIT0UOaDkaQ0AFb0t0jiQU4GKGCBUjGqkANRtSgQCjSCUSdCkbcYEEke0nHTAcmGKGDAnGsRhE4cJRATIDpCjQOJBK0wk7AAqnDSQScqIMwPuy0EoEj4zwMDzESwEHWRegQ8vBA7giy6cfzCDLPx/jYI4g7KONDCDoZmeMxyiGP/HDJJ3+sMssov5xRxy3XjDLOKO/css8YyRy0yEObXPTKR8OMEdA0M02y0zpD3bPUFyldtc0IE501zx8jzfHMHwt9dc4PG711zGg/rPbNWLetddlcW0R12lbT/802wm7jDffSYAcstt1kP2y23nEjPHfYdQN+t+J5V+Q134Wne7jkiSO8eEV7y9035H8HHDjlg3/ddOnpnu555RSF7vjohkdu+uSvp4756k93HvDnsTce8OO1sy6d67/DPtHlomcu3ea3+54u8BPJPjztmtveOu7J694872NH/bPw6RKfvfHJIT+98hIxP7vzyUG/vfTSUS+R9eVj/7z2x3O/vvfvAx/ixDc18knHfPtD32/UVz/2RcR914Pfb+TXP/olx34Rwd8B9Rc//qXPfw0EYAQFyDkCno1wJIyeCS0CwfxJkDUU/KAFf4NBiGgwOQjsoAJZw8ALOhAiLdzgC/9FE8MFgtCH40Ph2nq3QssZEIccnKAHjThD1tTwITf8TQ6luEPR9JCGP3xIEKE4xMsUkYdHBGMSVbfE8L0taU/UYhRhOEU0VlE0V3RIFlmzRTp28TJftGIYHTJGOZaxMWf0YhoFucbdtXGAb+xaHPk4RyLWUZF3vEweG7JH0fTRkn9sTCDxOMiGFJKSh+RLIgG5SFI28nuPLGEkWThJT1bSjJdkZSYbs0mGdPIyn8RlKPkySk2WkiGntGUq6bJKUbbSmK8MYCxVOEsnKtFvTKxm8K5JumwKDo7cLJ43UVfAcJ5vnLkDJxux6cZvnnCd3WwnOSVpzgSis3vlhKc45Zn/Tnrq85z8xOc7HclOSLqTlvXU4T3/l0+CxtOg80ToP+0ZUIYOFJYFleVBrTlRhVY0hA3F6EM1GlGOOnSfEO0n4xLKxYWC1J8nBWhKBbrSjrb0o0hUZ0wpOlOL1nSnHu3pSyUK1JsKNacXlWZGqblRiiQTmLdEZC6duUu+9HIhv2xMMKU6TLoUk5fHXMhTtRpVVU6VmM8EazRHOM35NdGptYTqMtvSTLRWlS5XVUhW+bJVs3a1LV+1algVMla+lpWZZ/VqWgW7Vhem0K3aXF5cyTrXstRVsXdtS14Tsle69BWxfy1LYPE62IQU1rOHpWtiAbtY0jZWiI+t4Fsly1I//7oUqT8VKUpJqlKT6lamvKUp6CZr2Mp25bKszWxZNouQ07bls6oNbVdGq9nSIqSzz02tZVcr2tZWV4SObatsI3s/4qLWuFxBbneV2xXmHsS5ZYHudqXLFeou17oHwW58tXtc7k7Xu/cFL2zFK8PZlre2oLytGnX6W54G16fDRbAwFcxIBit1pEwt6TZtalucLhimDQ7qg4caYQ4n2MMVrsg9lDGFB3jiVAh7BhgaIAkEIUwciWiAIjYWMHJIogFg8A3C1uGJBkxBGfGSiDygoAEjXAEH4EpXPb6AASJcwQcKiE62FqGAIahhCBZYRJJ/4wkF+OAKRFhAGPTzG2Jw6/8KRtBAFOQxkTCsYDi68VLFkmOJEOiCOKq4ADOkQwwN0AobtOhAMOSFgVQQRxchsIR0xjEs4kSjBmGQSDoMsCriWMEJyYGHAnh1nje0YD4Z+MN7/pAB9FKlBW94DyoUUJvfOMEK78GFARwGEWOEgEGlWNM0hk3sYhv72MhOtrKXzexmO5vZfFAAg2hBgFk8+9rYzra2tz1sagRiTTASwCi4Te5ym/va0ZgFAQ5NnGWd+93wjjcvCECs84DAGBHxNYNGIQBcQOPf0YBGwAcu8IIT/OAGTzjCF67whjP84Q6P+MPp8KXzUDsVAJe4xiHO8Y17vOMgb7ghBIBnSwvAEBn/D7nKP87ylbu84apY93usYQA6tPzmL885zj++CwEU6T0hwDdENt1p3Xw61KN+j6lRrerz/AEDbL4MC2J9HlRYoNasufV7cmEAWEXkCzUoeSkUsOffUMIEbFrFAgadHGBoYFXWwAUHgMFo9+hGFyighHTEoYCfY+PSmZZIPJ7QZDXca9HSqQcXMGAENWCZMVtWABHUQAQFiDldrvGBGoyAAS5EXTTBsIAO4KyBJ9B5IiueggM2weuAOcMLDKjxw8ChBwboAVcI83EDvICtgKljEw448pirZ97sulor6v0ve7niXoPotyvy7S99t2Lf9uLX+cXf7/Gzkvz6Atj6AiZj/2wLTN4MZh/6/E2vf72//K00vyDwRf/2sdJ96n+f+dcvyPO5En31T18r1Yd/4WdI40dFBmZ+EsZVFOZKFsZWSwVZTUV8CehXCwhNDRheDzheEXhgJjZhKMaAIHZhu5VhvbVhRdVhR/VhRBViRjViuFViJ3hiKZhiKyiCwEWCwmWCLIiCLqiCvmWDDoaDEKaDQChiQkhicDWBoFWBahVSRdiCR/iCSdiBCviBFphUDohhEKhhtEWFFGiFTYiFGKiFGsiF7XN+/Jd+W1F/AHh/7pd/BLF/W9F/a7h+9td+WvF+BBF/aTh/V8GGWRGAb/ha4kdgBlh+QISGc6iGyGeHbf+Ih1mhhwMhh1pBh434f4HohnkIhwPBh4voh1YBiFghiJtIiARoiHZ0gImohNHFhIzlhFk4gltYgl0Ygx44gyCYW0/Ig1Hog1Noi1WIi1eoi7F4g7OYg7W4gzLYgzQIg8p4i8yYiz9YjEF4jEMogV64hGD4iiFIjUZojUiIjcD4hcIYhjXojVAIjlIojs8YjNE4jNM4hrJYhrTIgeOojeXIjecoj8ZIj8hoj+1Iju9ojvE4YBlIfhuIgNnYitvoWhdokGSIkGaokPfIkPnokN3Ij9Xoj9cIkLu4jL3YjL8YkPg4kProjB8JjSEpjSOZku64kvBIhOjIi+roi8nokgL/CZMEKZMa+Y0cGY5nyIrz5YoYSYw9mY4/uY5BuZBD2ZDfBYtHSZNJaZNLWZFNeZFPKYYQOY8SWY8PpIiVyIjc54iZCIlYIYkCQYlZYYljiYmjqImRyIkC4YlhCYpVIYpXQYpxaYqoVICpiIhiBJZrKZb0R5ZvaZZXgZb4oJZYwZaF6ZZ5CZdnKZf4QJeDaZdUgZdWoZeTyZfK5JeYpIqBKZTSR5RZaZRb2Y9d+Y9fSZr+Z5oBBpWpuZGr2ZGtyZSl6ZSxqZWFeJCHmJCriJuvqZvgJ5u9GZG/OZHBaZW5iZW7iZrHyZXJ6ZXLSZIWaZJFiZIzCZI1KZI3uZ0q2Z0s/8mOOFmSOnmSLQmeLymeMUme6pmT7LmT3xmV3DmV3umR72me8Yme8zmbPlmbQEmR1nmV2HmaBRmdqjmdrGlDgtmYhPmHhhmZiGkVimmZDoqZU6GZVcGZiUmZjHkVjgmhkLmZktmhGCEOcDABxYB16UIOkAABBvMw6/AJEPAJMBYwEQMBkGBjAXMxE6AxFVEPcaAAOjAEIeABvfcb90AJCrACRYACGBAuyREK40IE5oIs6UIdKUAEN6AAkzB8ouEMHkACRHAvcfB5DfEIIdBpd6AArccardABd4INgWABPMYaumIKxGEKvyIdwiIo2DALHYClv5EOFnApxRECkSARh/8yC4rCKKzhKH7HKmogA5RiKe+RKWDKF53yHoGAAaHyG1GABKVyo26xJJ5qANjADazaqq76qrAaq7I6q7Raq7Z6q7X6bc0CaAbwC7j6q8AarMI6rK2qJr5QIAYAC8S6rMzarMD6C3YycwegCc5ardZ6rdmAAPVGHBwALA5BDCgAbAIQC0FSruaaIm2QALtaLCRyru76rhiCBwHwIucRDQEwI/Car/maI4mCJwrQBvoasO4aC0RiJFLaEORgAMd6HmugHr/BHuwGHyKQHPRQaedRp2hKFyKwB6UCMKwhBARyHrygADxqFlHgA7tKCwqQpKIRB2H3aMchHZ0QAguLDb7/EAKdQGgakAuWVgPakRzOoACHVg0+AAWbuhDwMAMhsAZ0gAQKkAnpIg9M0AFXcAdVsABxcLRelQUaYAV3YAUYkAUZSxf3MKRVcAdqwAFMcHp8ZgFGQAdrQAIywKIQUQ+iMAMHEAe4ly5n0QQDwAUsmxzGIAUDIAXeKi9cMABNoBcIEw5kcAA80Ar6op30GZ72OZ74WbnrebmaZA6eKVegqUuiSUgNCqIPGooRSqITWhWE4LnGeYq++ZfAOZrCWYcjuqElSqGuy5uwi5yyq5y0y5zD6ZzttbvQ2bvS+bvUGbwD2pwFulzGS7n+iZQAqpS3Kby2C5vF+7mUFbpUNbqm/1S6VhGiqHu7VMGhusu9xeW9dgW+yCS+VUG+d5m6uLu6VNG66nte7ItZ7itW8EsV8puZ9Hu+ucu60cuT0yuV1UuV19u8w/u82/u6fYmKoQmYpOua2UuczHfA7qm58Mm58lmVDpzBxLvB+Wt8+5tc/UtY/zsVAZyhAzwV6GvAJ6x9KbxeK2xaLSwVLywVGkrA9jsV+CvBn0nBomvB4YvBl6i9JkzEoGvE34vE76vEbcnE7sfBmZvA9bnA99nA5Xmd54lXWCygX0ygYWyg6enB+gnCVjXGDErFj2nFpfiQCEqbCjo9boxFOxwVPRwVPyzDBXy/lGmhpouhPhzDUjHD9/+bx3q0x1DRx1Dxx4kcyEI8yI78FJD8FJIcFYosxIzMSZfsFJnsFJsMFZ0sFRUayk0xyk1Ryk9xylExxLw7wbFbwbN7wbW7xBp8xTUsfzesfDncXKrMFKzMFK7sFLAMFbJ8vLTsu7YMvLiMvbpcwrzsxN0Lxe0rxf4LxyIqx5H4yb40zEtRzEtxzE2RzE+Rytxcvt58luCMVeKsFOSsFObMFOjsFOqcy1W8y3n4znoVz0kxz0lRz0txz02Rz9K8z9Tcz73ch7/MfsH8XgCNFAKNFAStFAbNFMssvXX8n3ccoNVZxs57xtDb0J/40HcY0QZByON7uvNrvoAcxKjsz5z/NdFHUdFHcdFJkdFLgdAjPM0Q3MSzXMS1fMS3nMT6HMf8/M0mXZco/YgqDX82bRQ4bRQ6jRQ8rRQbjcAdTb0fbb0hnZ9gvJ9i3NSX+dRlGdV7ONVFUdVFcdVHkdVJsdUdrMWWy8WY68VibcZkrVk0fV1sTRRuTRRwbRRyjRQ+LdIPTNIRPNRPXNRRfNRTnNTdvNTubNYXitaHqdadGNhDMdhDUdhFcdhHQddZ3NUK/NUMHNZqPNZsXNbWvL7YzL/azMLr/NLtnJh/nV+eLRSgLRSiTRSkbRSJvdcj3dclHdv6O9sqXNs6fNsCDNOTLNOxvNvYB90wLN2cTMmobMnY/33I2m3K3F3dmF3Imi2hnD2XvR0Uvx0UwT0Uw10Upk3Gxr3YyF2cGWnXm4vXnVveLW3IfozI203dyuzdlM3Olq3b/h2/Lh3duU2hBp7QSr3QTK3cKMzcOOzcwvzdAR7erzzeymzd+rfeQNHeQPHeQhHfRIGW61AJNPAMk5su7AAKFKAModqjw0ABw0C3yREPy0ABoHAwCEMPzkADlWCqEXEPrkECK7AALVB2bWcBIIADGiACQpYrG8ABOMABHSB02yECWs4BGAAMWksX48ACF3ADJLAYZZ4QngACjkapGSDkbaYBfjcHCuB1v/EMFvAHzfIHtyId5GABc0AcpaABXv/OGuyQAWrQLKnAAegSEexBarphDUTQBckhD6n2HlWwBJQiAlRX6W9wAm3OFUtQBe/RBxvAtqLBBUQgawASEcyAqhdrAd7wDbie67q+67ze677+68Ae7MI+7MHuCgdQDbJmANtA7Mze7M7+7NCe68lgAH92HrpgAMkQ7dq+7dze7NtgAJSuG9VwAK7Q7eZ+7ujuDfWxrtjAAWz3EOAKbABQCIhQ7/Z+7/ie7/q+7/ze7/7+7wDv72awAOxuCwJgBwGf8Aq/8Azf8PZ+BgGQG+dBDQFwBg5/8Rif8QpvBwBQdLqxAGag8SI/8iRfCAX7HkcSEZRWcrrxBjOQHOygAB7/3wcdUOpaUbF6eh6lYAGs3hiV0ge5ZgB0LhoyEOq6IQ1kFxH2oARGYO0YkOiikQU6gOy6EQ2QLh2pIvHYQA0qAC1TygElZw06kAXSYQwaUO26YQRKYPMFsQ4ncAN3EAhrgAFjkC7w8AIoMAeBwAYcEAVsf/NKAAJvEAhvEAJKEOOicQ9RoAGDPwco4AI87vOksQaBcAc4cAJI/hDx8AgiAABOcLi/QQ+ZwAIAIATB8PfUFwowAAAyEApjyxf3EAxCAAAssAmI/xvK4AQAIAKRcOOxMwhBkwcAHsl54OHInAcE/hSDIOJxSOI/YeI/geJBoeJDoZgf+t/nrbrpvZjO/+8T0O8T0g8U1C8U1t/9PfH9PRH+PzH+QVH+HB7JAi7eyY/PHmr+PIH+PKH+PsH+QOH+Bw4Q+AQOJFjQ4EGECRUuZHhwnqB5DSVOpFjRYkV7gtxd5NjR40eDhNCBxIeOEEmUKVUWdCfI3kqYMTk+jCjT5k2FGTfi5NlToEiSJn0OvdnyJVGkKmkmZUpSZ1OoHoGCFBrVakWjV7U2XLrVK8KnX8UWnPqx6li0WdGO7br2a1i3X8t6PBt3q1q7WtvmtQqXr9W5Hev+hYqXcNO9h5P6VZw0MMfBjYkaljw0ceWejDH3fHwx8maclEHfvDxapmbTMjtb/JwapmjXK0vHTv+JmnbK1RVb3yYJmzfI2b892hYudSTVk8VV+lZ+MXhzjBqho8xNcfd0isyxS3y+nSFx7w2rT7wefqF28wm7pz8Inj3C8RLLvzeInj7B9fcFutc/MH7D+frDx77+8tOPPwH/YyjA/gjUz8D7EOxPwYUY1M/B+yCkT0L9KFTIwvswpE/D9zi8z8OEQKRPxPdIZM9E+lBESMX3WGTPxfRgfE/Gg2hkz8b0cDRPR/Z4NMjH9IA0T8jwiEzPyIKQNE/J8Jj0zknzoCRIyvCo9M7K7bAMT8uBuPTOy+3AxE5M78gUyMzt0MROzenY3M7NkpITUCE5p6MTOjuxwxNO7PqE7s//5gKdblA990TI0OYQVU5R6Bh1lE+XLkVI0uIobc5STQ+CVDlOhfNUOVBDZSlTVfGDqFWCTi0uVVgHZLXWUn+TVThaYR21uFx52/W3Xlv9VbhgbxuWt2JVPfa3ZGlb9rZmQ32Wt2hjm5a2ajW99rZsXds2tm4v/Za2cFMb17VyHT03tnRNWze1dvd817V4R5vXtHoFvDe1fEHbd7R+G7wV1oA3Gxi0gi88uNWEMVt4s4ZDfFjViCubGLOKV7w41Iwl27iyjmv8WNOQGxtZspJ/PPnSlBVbubGWk3zZ0ZgPm1mxmqe8ec+cCdv5sJ67/FnAoP8amrCizzy6wFdrxWfp/7+ajvPpB6OulWq+rC4U6wy1hpXrvLye7l/TkuaLbLvMhg7t0dTOi+243G4ObtDktotut+xWDu/N9I6L77X8Lg5wzAR3i3C0DBcO8coUX4vxsRz/DXLJJEeLcrEs5w3zxjQfi3O5jjOrUV/BHlHsVkn3yvPbQFdMdLFc3wp22mQ/jPa3pJMaH9xj050w3r2yXavgXRv+r+K3Ov6q5FNbnq/mtXoeMNPpQt1Y1VtkXdXro4retOnzqv6q8KEaf7Ty7Tq/L9+lXh+09uN6P6r0m5p/s/rduh+q/DFlf5jp31r+15QAOiZ7gtmes7p3o++FKoFIGWBlCoiWAzJlgkSpoP9kLsiWCGpqg0PpYGM+KJYMLiZ+tSqhYk74lRQiZYQ+aeFhXuiVGBJlhpxZIGQaaK0HBimEl9ohT2pImBtuJYdDKSJOjviXJOpliI5q4k2eyJcoXmWJPqmiTa6YlyxaZYuZWSGsvmiXMEZljDzpomp66JkfeiuIS5rintoYkzPGJY1QWSNO7giTPLplj4ipo4D+uJJArmWQTOnjTQ6pkkSmZY5VKmR/HombN7Imjuaa5JcqeaAytiqSY1lkUhppk0tSJ5O62aS7OpmmT0YolKoapVhKiZRTnmaWoXJTPFjxA3Lco1byOIYNxFGPWtGjGTZoBj1qVQ9x2OAY8qjVPcj/8QNWxMMjxMDAAjhgACGso1XM8IACOKCAF5CjVeBgAQJCgAAWgKNV5IiBAc7pAWa0ah1LsOcCMEAMjnATENjARjSM4AFtasoZFriDNbBBjStkQJyaGocF3uDQarxBAerU1DoycAVqEPQOGHBGqODhASNEAxvWCMQ/LQIPCwSCoDPVQRw0RQ8P0GGmBDVCFELlgivsFBtXiEGoomAEoc7BA8i8VBx04NCZBsIC8KiIMzQA1Zn+AQPo4GpXvfpVsIZVrGMla1nNelazHsMA1RCqKRRQDrTGVa5zpWtdu9oNA+RCqLgwQDfs+lfABlau5TAAKoRKDQMgQ7CLZWxj0YGB/z8I1RoaKClFgJECoWLDFAGogyA8+1nQhla0oyVtaU17WtSm9rRAWEBmcREANKhWtrOlbW1t+9kyBCCkO6VGAMpwW+AGV7izRUMAcCHZBQRhuMtlLnMHUQcAmCKzKAhGRcShALbudA4v6KgCdCHUQGxAmI6ShwJUIdRUKICajrpHBga6U10oYKKOcsEchFoNBYijIvV4QVBnGg0OtEJT93BCEXZaDRJQIlRfwAFWrYGDMISKEiTILkGJ4ITxOqoVHFDpTInKVIqQIwNDGIUq7sCBKGTYUewQAQ4CsYo+kKAHIHZUPFqQAkCo4g81aEFCL0WPHoSgD6oIhA48wI5Q3YIjChzYgypGQYQMcNQi7OBCBizAglao+FLxiMMGLCCCTdD4x4vwgAU8wAhnqqoemziBATYQBx8PuBUsUEAGtPCOj9QjzrC6B1V/h494aLlVgf7dPeJxlD8nWtGLZnSjHf1oSEda0pOmdKUtfWlMZ1rTm+Z0pz39aVCHWtSjJnWpoRMQADs=" style="margin-top:1rem;" /></div></details></div></div>

::: {.sourceClojure}
```clojure
(view
 '(GridGraph [5 5])
 {:folded? true})
```
:::

<div style="max-height:600px;overflow-y:auto;"><div><details><summary><h5 style="cursor:pointer;display:inline;padding:0.5rem 1rem 0 1rem;">GridGraph[{5, 5}]</h5></summary><div class="wl-results"><hr /><img src="data:image/jpeg;base64,R0lGODlh4AHgAff4ADI2PjU5Pzc7QDY8Rjg7QTo9Qjw/Qzo+RTw/RDtART5BRT5DSDpDUDpEUj1GUj1GVDpHWj1HWj1IWT9LXD1MY0BDRkJER0BDSUFESkVHSUJGTUZIS0NIT0ZJTUlLTUpMT0VKUUdNVUtNUU5PUkhOVkBLW0lPWE5QUktRWkxSW0xTXE5VX1FSVVJVV1VXWVZYWldaXVhZW1pbXVtcXkJQZk9WYEVWb1BXYVBYYlJaZFRbZlRcZ1peY11fYFVdaUtbc0pcd0pceF5gY1dgbFliblpjcVtkcl1mc15ndV9odmFiY2FjZGNkZmJlaWZnaWlqa2tsbWxtbm1ucWtvdmBqeGJtemNtfGRvfmZwf3Bxc3NzdXN0dnV2d3Z3eHN3fHd4eXV4fXl5enp7fHt8fX1+f1Flg1dngF9wi1lvkGZxgGhygml0g2x3h2RziWZ0im56ind9hH9/gHl+hnJ+jm5+lmF4nGh7mXN/kXB/l2V9on6BhHSAknWClHiFlnmGmHuImn2KnX+Mn2eApmuDpm6GqnGGpn2LoHqLpXeNr4CAgYCChYOEhIaHh4eHiIOHjYeIiISIjomJioqLjIuMjo2NjoaMlo+PkIONnoSPno+QkIWQn5KSk5CTlpOUlJWWlpGUmJKXnpqamp2enoGOoYKQo4ORpIWTpoaUp4eVqYmXq4qZro2brpWdp56foJOcqZSdqpGdroKVsoSWso2bsY6dsomdupCftJGhtpOiuJSjuZWkupOjvJenvpqnu5movp+rvqGhoaOjpKampqOlqqenqKmpqa2traOosaKqtKGrurGxsbKytLa2trS3vLq6ur29vZipw5uswpysw5urxJ6uxp+wx6GyyqWzyKO0zKi2yam3za+6yqy6zrG5xLnE1bzH2MLCwsXFxcXGyMnJyc7OzsDI1cXP3cbQ3sjR3tLS0tXV1dnZ2d7e3srT4NTb5eHh4ebm5urq6u3t7eXp8PHx8fX19fb3+vr6+v///////////////////////////////yH/C05FVFNDQVBFMi4wAwEAAAAh+QQECgD/ACwAAAAA4AHgAQAI/wDxCRxIsKDBgwgTKlzIsKHDhxAjSpxIsaLFixgzatzIsaPHjyBDihxJsqTJkyhTqlzJsqXLlzBjypxJs6bNmzhz6tzJs6fPn0CDCh1KtKjRo0iTKl3KtKnTp1CjSp1KtarVq1izat3KtWa9VlFexBHX1eE9YllciHFW1qGyLi66KGvb0JkYF1mI3aO7UFycF1GA1aMIT0YINXOMKNjEFyE9JxyuzEmyIFHjg/W0YKgypwoGLYMvF4yzIMmcKxyc0BNdcJMCI3PUhJgBb2KUHdSw6Va14BnrgYlURNONDZcGYr8FegrBizivEJ6S4yOGARdxaSriSH+mQBVxajuiSP8kh6A5cWxqlCR/p2DVeWx3TiSvZyHQ+1EWQrM+cef9KgXvJLfEGu/xYgA5ERFDwnvYlBKALLVEKOGEFFZo4YUYZqjhhhxq6IYC1rxHiwCHdGjiiSimqKKEeAggzXvRCIDHijTWaOOJhwhAC4MKuHHjj0AGKUsApTBIAnIQKcMBg4EgcI02UEYp5ZRUVmnllVhmqeWWWQYCon8G9MLlmGSWaeaZUa5igHnOGbAKmnDGKSeZvRjg3XnWHHDJnHz26ec1Btj3HghzQbSOAra8l4R4v8VjwSjvrRFDcvd40N95d3hgT3IyqHGfBfEkF0USIiqwjkSLoGAdNtbcYUE60oX/wsGOuo2iADjSKYMBKsShgkGhv4GjAKS60cJBK9KlowAfIRZnwiIT0UOaDkaQ0AFb0t0jiQU4GKGCBUjGqkANRtSgQCjSCUSdCkbcYEEke0nHTAcmGKGDAnGsRhE4cJRATIDpCjQOJBK0wk7AAqnDSQScqIMwPuy0EoEj4zwMDzESwEHWRegQ8vBA7giy6cfzCDLPx/jYI4g7KONDCDoZmeMxyiGP/HDJJ3+sMssov5xRxy3XjDLOKO/css8YyRy0yEObXPTKR8OMEdA0M02y0zpD3bPUFyldtc0IE501zx8jzfHMHwt9dc4PG711zGg/rPbNWLetddlcW0R12lbT/802wm7jDffSYAcstt1kP2y23nEjPHfYdQN+t+J5V+Q134Wne7jkiSO8eEV7y9035H8HHDjlg3/ddOnpnu555RSF7vjohkdu+uSvp4756k93HvDnsTce8OO1sy6d67/DPtHlomcu3ea3+54u8BPJPjztmtveOu7J694872NH/bPw6RKfvfHJIT+98hIxP7vzyUG/vfTSUS+R9eVj/7z2x3O/vvfvAx/ixDc18knHfPtD32/UVz/2RcR914Pfb+TXP/olx34Rwd8B9Rc//qXPfw0EYAQFyDkCno1wJIyeCS0CwfxJkDUU/KAFf4NBiGgwOQjsoAJZw8ALOhAiLdzgC/9FE8MFgtCH40Ph2nq3QssZEIccnKAHjThD1tTwITf8TQ6luEPR9JCGP3xIEKE4xMsUkYdHBGMSVbfE8L0taU/UYhRhOEU0VlE0V3RIFlmzRTp28TJftGIYHTJGOZaxMWf0YhoFucbdtXGAb+xaHPk4RyLWUZF3vEweG7JH0fTRkn9sTCDxOMiGFJKSh+RLIgG5SFI28nuPLGEkWThJT1bSjJdkZSYbs0mGdPIyn8RlKPkySk2WkiGntGUq6bJKUbbSmK8MYCxVOEsnKtFvTKxm8K5JumwKDo7cLJ43UVfAcJ5vnLkDJxux6cZvnnCd3WwnOSVpzgSis3vlhKc45Zn/Tnrq85z8xOc7HclOSLqTlvXU4T3/l0+CxtOg80ToP+0ZUIYOFJYFleVBrTlRhVY0hA3F6EM1GlGOOnSfEO0n4xLKxYWC1J8nBWhKBbrSjrb0o0hUZ0wpOlOL1nSnHu3pSyUK1JsKNacXlWZGqblRiiQTmLdEZC6duUu+9HIhv2xMMKU6TLoUk5fHXMhTtRpVVU6VmM8EazRHOM35NdGptYTqMtvSTLRWlS5XVUhW+bJVs3a1LV+1algVMla+lpWZZ/VqWgW7Vhem0K3aXF5cyTrXstRVsXdtS14Tsle69BWxfy1LYPE62IQU1rOHpWtiAbtY0jZWiI+t4Fsly1I//7oUqT8VKUpJqlKT6lamvKUp6CZr2Mp25bKszWxZNouQ07bls6oNbVdGq9nSIqSzz02tZVcr2tZWV4SObatsI3s/4qLWuFxBbneV2xXmHsS5ZYHudqXLFeou17oHwW58tXtc7k7Xu/cFL2zFK8PZlre2oLytGnX6W54G16fDRbAwFcxIBit1pEwt6TZtalucLhimDQ7qg4caYQ4n2MMVrsg9lDGFB3jiVAh7BhgaIAkEIUwciWiAIjYWMHJIogFg8A3C1uGJBkxBGfGSiDygoAEjXAEH4EpXPb6AASJcwQcKiE62FqGAIahhCBZYRJJ/4wkF+OAKRFhAGPTzG2Jw6/8KRtBAFOQxkTCsYDi68VLFkmOJEOiCOKq4ADOkQwwN0AobtOhAMOSFgVQQRxchsIR0xjEs4kSjBmGQSDoMsCriWMEJyYGHAnh1nje0YD4Z+MN7/pAB9FKlBW94DyoUUJvfOMEK78GFARwGEWOEgEGlWNM0hk3sYhv72MhOtrKXzexmO5vZfFAAg2hBgFk8+9rYzra2tz1sagRiTTASwCi4Te5ym/va0ZgFAQ5NnGWd+93wjjcvCECs84DAGBHxNYNGIQBcQOPf0YBGwAcu8IIT/OAGTzjCF67whjP84Q6P+MPp8KXzUDsVAJe4xiHO8Y17vOMgb7ghBIBnSwvAEBn/D7nKP87ylbu84apY93usYQA6tPzmL885zj++CwEU6T0hwDdENt1p3Xw61KN+j6lRrerz/AEDbL4MC2J9HlRYoNasufV7cmEAWEXkCzUoeSkUsOffUMIEbFrFAgadHGBoYFXWwAUHgMFo9+hGFyighHTEoYCfY+PSmZZIPJ7QZDXca9HSqQcXMGAENWCZMVtWABHUQAQFiDldrvGBGoyAAS5EXTTBsIAO4KyBJ9B5IiueggM2weuAOcMLDKjxw8ChBwboAVcI83EDvICtgKljEw448pirZ97sulor6v0ve7niXoPotyvy7S99t2Lf9uLX+cXf7/Gzkvz6Atj6AiZj/2wLTN4MZh/6/E2vf72//K00vyDwRf/2sdJ96n+f+dcvyPO5En31T18r1Yd/4WdI40dFBmZ+EsZVFOZKFsZWSwVZTUV8CehXCwhNDRheDzheEXhgJjZhKMaAIHZhu5VhvbVhRdVhR/VhRBViRjViuFViJ3hiKZhiKyiCwEWCwmWCLIiCLqiCvmWDDoaDEKaDQChiQkhicDWBoFWBahVSRdiCR/iCSdiBCviBFphUDohhEKhhtEWFFGiFTYiFGKiFGsiF7XN+/Jd+W1F/AHh/7pd/BLF/W9F/a7h+9td+WvF+BBF/aTh/V8GGWRGAb/ha4kdgBlh+QISGc6iGyGeHbf+Ih1mhhwMhh1pBh434f4HohnkIhwPBh4voh1YBiFghiJtIiARoiHZ0gImohNHFhIzlhFk4gltYgl0Ygx44gyCYW0/Ig1Hog1Noi1WIi1eoi7F4g7OYg7W4gzLYgzQIg8p4i8yYiz9YjEF4jEMogV64hGD4iiFIjUZojUiIjcD4hcIYhjXojVAIjlIojs8YjNE4jNM4hrJYhrTIgeOojeXIjecoj8ZIj8hoj+1Iju9ojvE4YBlIfhuIgNnYitvoWhdokGSIkGaokPfIkPnokN3Ij9Xoj9cIkLu4jL3YjL8YkPg4kProjB8JjSEpjSOZku64kvBIhOjIi+roi8nokgL/CZMEKZMa+Y0cGY5nyIrz5YoYSYw9mY4/uY5BuZBD2ZDfBYtHSZNJaZNLWZFNeZFPKYYQOY8SWY8PpIiVyIjc54iZCIlYIYkCQYlZYYljiYmjqImRyIkC4YlhCYpVIYpXQYpxaYqoVICpiIhiBJZrKZb0R5ZvaZZXgZb4oJZYwZaF6ZZ5CZdnKZf4QJeDaZdUgZdWoZeTyZfK5JeYpIqBKZTSR5RZaZRb2Y9d+Y9fSZr+Z5oBBpWpuZGr2ZGtyZSl6ZSxqZWFeJCHmJCriJuvqZvgJ5u9GZG/OZHBaZW5iZW7iZrHyZXJ6ZXLSZIWaZJFiZIzCZI1KZI3uZ0q2Z0s/8mOOFmSOnmSLQmeLymeMUme6pmT7LmT3xmV3DmV3umR72me8Yme8zmbPlmbQEmR1nmV2HmaBRmdqjmdrGlDgtmYhPmHhhmZiGkVimmZDoqZU6GZVcGZiUmZjHkVjgmhkLmZktmhGCEOcDABxYB16UIOkAABBvMw6/AJEPAJMBYwEQMBkGBjAXMxE6AxFVEPcaAAOjAEIeABvfcb90AJCrACRYACGBAuyREK40IE5oIs6UIdKUAEN6AAkzB8ouEMHkACRHAvcfB5DfEIIdBpd6AArccardABd4INgWABPMYaumIKxGEKvyIdwiIo2DALHYClv5EOFnApxRECkSARh/8yC4rCKKzhKH7HKmogA5RiKe+RKWDKF53yHoGAAaHyG1GABKVyo26xJJ5qANjADazaqq76qrAaq7I6q7Raq7Z6q7X6bc0CaAbwC7j6q8AarMI6rK2qJr5QIAYAC8S6rMzarMD6C3YycwegCc5ardZ6rdmAAPVGHBwALA5BDCgAbAIQC0FSruaaIm2QALtaLCRyru76rhiCBwHwIucRDQEwI/Car/maI4mCJwrQBvoasO4aC0RiJFLaEORgAMd6HmugHr/BHuwGHyKQHPRQaedRp2hKFyKwB6UCMKwhBARyHrygADxqFlHgA7tKCwqQpKIRB2H3aMchHZ0QAguLDb7/EAKdQGgakAuWVgPakRzOoACHVg0+AAWbuhDwMAMhsAZ0gAQKkAnpIg9M0AFXcAdVsABxcLRelQUaYAV3YAUYkAUZSxf3MKRVcAdqwAFMcHp8ZgFGQAdrQAIywKIQUQ+iMAMHEAe4ly5n0QQDwAUsmxzGIAUDIAXeKi9cMABNoBcIEw5kcAA80Ar6op30GZ72OZ74WbnrebmaZA6eKVegqUuiSUgNCqIPGooRSqITWhWE4LnGeYq++ZfAOZrCWYcjuqElSqGuy5uwi5yyq5y0y5zD6ZzttbvQ2bvS+bvUGbwD2pwFulzGS7n+iZQAqpS3Kby2C5vF+7mUFbpUNbqm/1S6VhGiqHu7VMGhusu9xeW9dgW+yCS+VUG+d5m6uLu6VNG66nte7ItZ7itW8EsV8puZ9Hu+ucu60cuT0yuV1UuV19u8w/u82/u6fYmKoQmYpOua2UuczHfA7qm58Mm58lmVDpzBxLvB+Wt8+5tc/UtY/zsVAZyhAzwV6GvAJ6x9KbxeK2xaLSwVLywVGkrA9jsV+CvBn0nBomvB4YvBl6i9JkzEoGvE34vE76vEbcnE7sfBmZvA9bnA99nA5Xmd54lXWCygX0ygYWyg6enB+gnCVjXGDErFj2nFpfiQCEqbCjo9boxFOxwVPRwVPyzDBXy/lGmhpouhPhzDUjHD9/+bx3q0x1DRx1Dxx4kcyEI8yI78FJD8FJIcFYosxIzMSZfsFJnsFJsMFZ0sFRUayk0xyk1Ryk9xylExxLw7wbFbwbN7wbW7xBp8xTUsfzesfDncXKrMFKzMFK7sFLAMFbJ8vLTsu7YMvLiMvbpcwrzsxN0Lxe0rxf4LxyIqx5H4yb40zEtRzEtxzE2RzE+Rytxcvt58luCMVeKsFOSsFObMFOjsFOqcy1W8y3n4znoVz0kxz0lRz0txz02Rz9K8z9Tcz73ch7/MfsH8XgCNFAKNFAStFAbNFMssvXX8n3ccoNVZxs57xtDb0J/40HcY0QZByON7uvNrvoAcxKjsz5z/NdFHUdFHcdFJkdFLgdAjPM0Q3MSzXMS1fMS3nMT6HMf8/M0mXZco/YgqDX82bRQ4bRQ6jRQ8rRQbjcAdTb0fbb0hnZ9gvJ9i3NSX+dRlGdV7ONVFUdVFcdVHkdVJsdUdrMWWy8WY68VibcZkrVk0fV1sTRRuTRRwbRRyjRQ+LdIPTNIRPNRPXNRRfNRTnNTdvNTubNYXitaHqdadGNhDMdhDUdhFcdhHQddZ3NUK/NUMHNZqPNZsXNbWvL7YzL/azMLr/NLtnJh/nV+eLRSgLRSiTRSkbRSJvdcj3dclHdv6O9sqXNs6fNsCDNOTLNOxvNvYB90wLN2cTMmobMnY/33I2m3K3F3dmF3Imi2hnD2XvR0Uvx0UwT0Uw10Upk3Gxr3YyF2cGWnXm4vXnVveLW3IfozI203dyuzdlM3Olq3b/h2/Lh3duU2hBp7QSr3QTK3cKMzcOOzcwvzdAR7erzzeymzd+rfeQNHeQPHeQhHfRIGW61AJNPAMk5su7AAKFKAModqjw0ABw0C3yREPy0ABoHAwCEMPzkADlWCqEXEPrkECK7AALVB2bWcBIIADGiACQpYrG8ABOMABHSB02yECWs4BGAAMWksX48ACF3ADJLAYZZ4QngACjkapGSDkbaYBfjcHCuB1v/EMFvAHzfIHtyId5GABc0AcpaABXv/OGuyQAWrQLKnAAegSEexBarphDUTQBckhD6n2HlWwBJQiAlRX6W9wAm3OFUtQBe/RBxvAtqLBBUQgawASEcyAqhdrAd7wDbie67q+67ze677+68Ae7MI+7MHuCgdQDbJmANtA7Mze7M7+7NCe68lgAH92HrpgAMkQ7dq+7dze7NtgAJSuG9VwAK7Q7eZ+7ujuDfWxrtjAAWz3EOAKbABQCIhQ7/Z+7/ie7/q+7/ze7/7+7wDv72awAOxuCwJgBwGf8Aq/8Azf8PZ+BgGQG+dBDQFwBg5/8Rif8QpvBwBQdLqxAGag8SI/8iRfCAX7HkcSEZRWcrrxBjOQHOygAB7/3wcdUOpaUbF6eh6lYAGs3hiV0ge5ZgB0LhoyEOq6IQ1kFxH2oARGYO0YkOiikQU6gOy6EQ2QLh2pIvHYQA0qAC1TygElZw06kAXSYQwaUO26YQRKYPMFsQ4ncAN3EAhrgAFjkC7w8AIoMAeBwAYcEAVsf/NKAAJvEAhvEAJKEOOicQ9RoAGDPwco4AI87vOksQaBcAc4cAJI/hDx8AgiAABOcLi/QQ+ZwAIAIATB8PfUFwowAAAyEApjyxf3EAxCAAAssAmI/xvK4AQAIAKRcOOxMwhBkwcAHsl54OHInAcE/hSDIOJxSOI/YeI/geJBoeJDoZgf+t/nrbrpvZjO/+8T0O8T0g8U1C8U1t/9PfH9PRH+PzH+QVH+HB7JAi7eyY/PHmr+PIH+PKH+PsH+QOH+Bw4Q+AQOJFjQ4EGECRUuZHhwnqB5DSVOpFjRYkV7gtxd5NjR40eDhNCBxIeOEEmUKVUWdCfI3kqYMTk+jCjT5k2FGTfi5NlToEiSJn0OvdnyJVGkKmkmZUpSZ1OoHoGCFBrVakWjV7U2XLrVK8KnX8UWnPqx6li0WdGO7br2a1i3X8t6PBt3q1q7WtvmtQqXr9W5Hev+hYqXcNO9h5P6VZw0MMfBjYkaljw0ceWejDH3fHwx8maclEHfvDxapmbTMjtb/JwapmjXK0vHTv+JmnbK1RVb3yYJmzfI2b892hYudSTVk8VV+lZ+MXhzjBqho8xNcfd0isyxS3y+nSFx7w2rT7wefqF28wm7pz8Inj3C8RLLvzeInj7B9fcFutc/MH7D+frDx77+8tOPPwH/YyjA/gjUz8D7EOxPwYUY1M/B+yCkT0L9KFTIwvswpE/D9zi8z8OEQKRPxPdIZM9E+lBESMX3WGTPxfRgfE/Gg2hkz8b0cDRPR/Z4NMjH9IA0T8jwiEzPyIKQNE/J8Jj0zknzoCRIyvCo9M7K7bAMT8uBuPTOy+3AxE5M78gUyMzt0MROzenY3M7NkpITUCE5p6MTOjuxwxNO7PqE7s//5gKdblA990TI0OYQVU5R6Bh1lE+XLkVI0uIobc5STQ+CVDlOhfNUOVBDZSlTVfGDqFWCTi0uVVgHZLXWUn+TVThaYR21uFx52/W3Xlv9VbhgbxuWt2JVPfa3ZGlb9rZmQ32Wt2hjm5a2ajW99rZsXds2tm4v/Za2cFMb17VyHT03tnRNWze1dvd817V4R5vXtHoFvDe1fEHbd7R+G7wV1oA3Gxi0gi88uNWEMVt4s4ZDfFjViCubGLOKV7w41Iwl27iyjmv8WNOQGxtZspJ/PPnSlBVbubGWk3zZ0ZgPm1mxmqe8ec+cCdv5sJ67/FnAoP8amrCizzy6wFdrxWfp/7+ajvPpB6OulWq+rC4U6wy1hpXrvLye7l/TkuaLbLvMhg7t0dTOi+243G4ObtDktotut+xWDu/N9I6L77X8Lg5wzAR3i3C0DBcO8coUX4vxsRz/DXLJJEeLcrEs5w3zxjQfi3O5jjOrUV/BHlHsVkn3yvPbQFdMdLFc3wp22mQ/jPa3pJMaH9xj050w3r2yXavgXRv+r+K3Ov6q5FNbnq/mtXoeMNPpQt1Y1VtkXdXro4retOnzqv6q8KEaf7Ty7Tq/L9+lXh+09uN6P6r0m5p/s/rduh+q/DFlf5jp31r+15QAOiZ7gtmes7p3o++FKoFIGWBlCoiWAzJlgkSpoP9kLsiWCGpqg0PpYGM+KJYMLiZ+tSqhYk74lRQiZYQ+aeFhXuiVGBJlhpxZIGQaaK0HBimEl9ohT2pImBtuJYdDKSJOjviXJOpliI5q4k2eyJcoXmWJPqmiTa6YlyxaZYuZWSGsvmiXMEZljDzpomp66JkfeiuIS5rintoYkzPGJY1QWSNO7giTPLplj4ipo4D+uJJArmWQTOnjTQ6pkkSmZY5VKmR/HombN7Imjuaa5JcqeaAytiqSY1lkUhppk0tSJ5O62aS7OpmmT0YolKoapVhKiZRTnmaWoXJTPFjxA3Lco1byOIYNxFGPWtGjGTZoBj1qVQ9x2OAY8qjVPcj/8QNWxMMjxMDAAjhgACGso1XM8IACOKCAF5CjVeBgAQJCgAAWgKNV5IiBAc7pAWa0ah1LsOcCMEAMjnATENjARjSM4AFtasoZFriDNbBBjStkQJyaGocF3uDQarxBAerU1DoycAVqEPQOGHBGqODhASNEAxvWCMQ/LQIPCwSCoDPVQRw0RQ8P0GGmBDVCFELlgivsFBtXiEGoomAEoc7BA8i8VBx04NCZBsIC8KiIMzQA1Zn+AQPo4GpXvfpVsIZVrGMla1nNelazHsMA1RCqKRRQDrTGVa5zpWtdu9oNA+RCqLgwQDfs+lfABlau5TAAKoRKDQMgQ7CLZWxj0YGB/z8I1RoaKClFgJECoWLDFAGogyA8+1nQhla0oyVtaU17WtSm9rRAWEBmcREANKhWtrOlbW1t+9kyBCCkO6VGAMpwW+AGV7izRUMAcCHZBQRhuMtlLnMHUQcAmCKzKAhGRcShALbudA4v6KgCdCHUQGxAmI6ShwJUIdRUKICajrpHBga6U10oYKKOcsEchFoNBYijIvV4QVBnGg0OtEJT93BCEXZaDRJQIlRfwAFWrYGDMISKEiTILkGJ4ITxOqoVHFDpTInKVIqQIwNDGIUq7sCBKGTYUewQAQ4CsYo+kKAHIHZUPFqQAkCo4g81aEFCL0WPHoSgD6oIhA48wI5Q3YIjChzYgypGQYQMcNQi7OBCBizAglao+FLxiMMGLCCCTdD4x4vwgAU8wAhnqqoemziBATYQBx8PuBUsUEAGtPCOj9QjzrC6B1V/h494aLlVgf7dPeJxlD8nWtGLZnSjHf1oSEda0pOmdKUtfWlMZ1rTm+Z0pz39aVCHWtSjJnWpoRMQADs=" style="margin-top:1rem;" /></div></details></div></div>

::: {.sourceClojure}
```clojure
(view
 '(ChemicalData "Ethanol" "StructureDiagram"))
```
:::

<div style="max-height:600px;overflow-y:auto;"><div><details open="open"><summary><h5 style="cursor:pointer;display:inline;padding:0.5rem 1rem 0 1rem;">ChemicalData["Ethanol", "StructureDiagram"]</h5></summary><div class="wl-results"><hr /><img src="data:image/jpeg;base64,R0lGODlhgwArAPdXABUVFTc3N4CAgAMDAxAQEFVVVZCQkK2trRwcHJeXl3p6ekZGRggICAcHB/X19URERC8vL4+Pj/39/ZKSkllZWcrKypycnP7+/vb29lhYWIyMjE5OTlJSUtvb246Ojri4uJSUlKenp5OTk29vb8XFxYiIiGtra8jIyKqqqtfX14qKiu7u7vT09LKysh0dHVZWVqOjoz4+PuDg4OHh4b6+vry8vGlpaezs7EVFRcTExL29vfLy8rq6uoSEhF1dXaioqJqamjU1NXh4eDQ0NJ+fn19fX5aWlqWlpfr6+tXV1W1tbd/f36urq5ubm7W1tYmJifj4+H9/f6GhoVRUVJ2dndnZ2R4eHv///8vLywAAAMfHx2RkZPv7+9DQ0JmZmbS0tNTU1OLi4kdHR+vr683NzXNzc6ysrElJSYKCgsPDw7GxsaamplBQUHx8fDExMXl5eUhISK6urujo6N3d3Q8PD3R0dOXl5dPT02FhYVxcXOPj49zc3H5+fjIyMufn58DAwGJiYvz8/DY2NsbGxpGRkXFxcZiYmLCwsHd3dwQEBEpKSrOzs9HR0cLCwp6eng4ODubm5s7OzlFRUfDw8IGBgfPz82VlZYeHh4aGhjg4OAEBAUxMTAYGBmdnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZ2dnZyH/C05FVFNDQVBFMi4wAwEAAAAh+QQECgD/ACwAAAAAgwArAAAI/wCvCBxIsKDBgwgTKlzI0CCShhAjSpxIsaLEFTFyWNzIsaPHiUReMCnUY9LHkyhTTqzB5sjAP1OkqJxJc2YVJSVYGITBgUbNiGCyHCrIZsjPkxhUmEiRcAemER2OLgw6lGBRqR1DbODRMIkNDVCwHqRK1KhYimkyUJn4YdOPswTJWjULt6GMOpRuWLRAgURduQOv1l0oYcKWLh3HREEU5mzQLJAjZ6E7+GAcMS1QkgEkQmzQMmZChw5CufJALEWM0FSDo+rPoGbKmiYop80bO0dB4CHzOkvsubMFeslzQqweRHz20IQt23QgQUwGD3LzZCZz4JXvWCI029CmDymvB/+mrGiR1Ek9Cs0JfgWKBhtJ6l4gZIlRTSlT/rAn2GEEph11LVEGGjOkRAMHMOx30IEJ1tWIFY581N9/CiaEn351wSCJDhu5B1+FC6GnXl2VXDJCJQRBkskjmnDiQgQHofAdiBDd0UkEXNSVRCQDrZGIJhAU8QAdWQhSUBRA0DiRE2f8NpgDDTCgxUASLJCFEAStoGRFhvhQXF0vZGGIQY8ksqVHfrShAG5nWcHAQZ1kUcGZHlVQRAJnMYDAQY5kgQmdHx0AxxdY6XnQEVmUAehHgRgA3lFuHmRDFg0u6pFJRxWQRRMGETCABJaGKhCUDMw5kCJZCCDqqj5q0scWihAmkEUAq9aqIouJAOBBrbz26uuvwAYr7LDEFmvsscgmq+yyzDYLYkAAOw==" style="margin-top:1rem;" /></div></details></div></div>

## More Working Examples

::: {.sourceClojure}
```clojure
(wl/eval '(GeoNearest (Entity "Ocean") Here))
```
:::

::: {.printedClojure}
```clojure
[:entity.ocean/mediterranean-sea-eastern-basin
 :entity.ocean/mediterranean-sea]

```
:::

 TODO: Make this work with `view` as well.

::: {.sourceClojure}
```clojure
(view '(TextStructure "The cat sat on the mat."))
```
:::

<div style="max-height:600px;overflow-y:auto;"><div><details open="open"><summary><h5 style="cursor:pointer;display:inline;padding:0.5rem 1rem 0 1rem;">TextStructure["The cat sat on the mat."]</h5></summary><div class="wl-results"><hr /><img src="data:image/jpeg;base64,R0lGODlhpAGaAPd8ANzc3Kenp8zMzNLS0peXl/Hx8fT09M/Pz9PT0xwcHC0tLePj44+PjxkZGZqamjU1NcHBwbq6upubm7W1tYmJidHR0ba2tpmZmampqTY2Nl5eXlhYWBISEqCgoBEREW1tbcPDw7GxseDg4GlpaaWlpdbW1v7+/urq6gwMDO7u7ltbW/X19fr6+pGRkSwsLD8/P52dnVRUVJiYmHFxcTw8PKqqqnl5eUNDQ0JCQkRERBAQEIqKivn5+djY2KKiokdHR42NjdDQ0GNjY+bm5s7OzikpKQ4ODgkJCeXl5XR0dA8PD/Dw8PPz8woKCoaGhjg4OKSkpB0dHbKysjk5OWZmZsrKypycnC4uLoeHh2VlZZKSkllZWYuLiw0NDf39/RsbGxoaGtvb27i4uI6OjpSUlC8vLwAAAIKCgklJSRMTE5OTk3t7e25uboODg0hISN3d3a6urujo6HV1dRYWFqampnBwcGtra/f398jIyMXFxaioqPb29v///8vLy9fX11NTU7u7u7e3twcHB9/f3wgICNXV1QYGBu/v76GhoX9/f/j4+BUVFSsrK7CwsHd3dwQEBEtLS8LCwoiIiAsLCwMDA0ZGRiAgIH19fR8fH2FhYbm5uXx8fFBQUEVFRSMjI+zs7N7e3oyMjD09PZWVlaysrJ+fnzQ0NHh4eIWFhQUFBYCAgGRkZPv7+8fHx8bGxlZWVqurqzExMcnJyenp6SQkJK2trbOzswEBAcTExCYmJjo6OqOjoyoqKrS0tNTU1OLi4hQUFDAwMJ6enuTk5DMzM2pqaq+vr+3t7RcXF9nZ2dra2pCQkOHh4b6+vr29vfLy8sDAwGJiYvz8/Ofn5yUlJWhoaM3Nzevr6wICAr+/v19fX5aWlry8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vLy8vCH/C05FVFNDQVBFMi4wAwEAAAAh+QQECgD/ACwAAAAApAGaAAAI/wD5CBxIsKDBgwgTKlzIsKHDhxAjSpxIsaLFixgzatzIsaPHjyBDihxJsqTJkyhTqlzJsqXLlzBjypxJs6bNmzhz6tzJs6fPn0CDCh1KtKhRomoEHV3KVGLSplAtBjJDtSrVNnyeBmWV6JhKrl6jjgwDxKBWoGDFZvwEqC0aT20BDcqqNGgcM2FU3s2rNqSeL2brAt3bd6OQMgXVEBJaCK/KxnwLexwFOPHioJAlZzycWIkiNmkWZdlDENEVQ1E2sfII5ckRI39+DQwSLUqTMnAG2rFatU9EWT8WETKlhcfs2rdzC9zN24zvksy26GiSQ9ZANVMOZdIhmrRMRs0nCf9U4xl099KnU6/eSAmKpTSa6kziJE0gbdu4dTd3rpkiZ4JqGCEKFaScgg0qA11iiCqu7AJGHR3JYYY2sEgBSSUCzULIFrXY8sotz8XhByJmaOLHiXdA9IkSr0TSChm5PKchhx6CKJCIJJqIYklxfMGLD6TkIAge4wETjDDDGHMEgjKB4ocdyJzoRzLjCUiggUzyoSCDDkLIXi5+aNPFMnncEggfM3b4YYgjlihliv1J9N91t+wyECS6CKQMJVoM1EwqcWzUhxlOEPTMQNMM5AU11RCUjWMT5WGGdQclKtCijQ70aGQmFaNEWNIEk2dW2BAp0BajzpRIZQDWeeeoe/b/KdCfgWpEyTZ80JGLQHPIaikfmDoKaZz+IQagYHycYolAOxDiHbCGcLORDYYYl5AyNUiSC4aaDhsRDzSAscMbCGGrLbcCbZoSJpkQNIoZn9BF0BrL0rRqYAQpy6yzikbLnp2wECPQF1wQZO62wnJKbERzjofsvXzMgI0SFFN8CxQbZUJNQoh4gkwMaxDzQsIVsWIFDo+40AILA3X8ccgjd6swSYQkQpAUZhQi70AQq8rqdQ9XJnHFFmNsK8AC80GwQC6DLDLJC0/U8M4CQYwFIW8MorXWK2zkyBH1GUQKNsJ4IZAKMQskqR8YpdACCkIINHbZZ6fNx9opRRH3QNuY/3EC1Xz0LBMqyODLc2VXZ731IF0fzUfAAxc8t9l8oE0Q3i6ZEPZNU58VeGUlpEJCSLiYMQZBlo4QC0Gc2C2CGSFQlMIhBCWBiUCqs+467Cl94EEKlxJDg8MECR5TB5QAf2zxoIseEiVIR85H7gO1TtDrsbf0TC4c+IJT50EPhIogcgCSzQ4bdCTELavA0Ys2yCjCBwyTRGJCGBpMYjcfGdCCQTYXqABEHAEMCriiAiRAxggEQj/74U9/Benf/wJYkkN8gRG7qEUnBCFAwBkPJp9oghssEAEskMZzPRtf+c6Xvn89LmlLa+D98rc/CQKwgyiBAFVOxzljAY15BLHFA/8I0YVKQKMjJthGMAThgRh4jw8m2EQDmvCEGjBgfwtwwxGUoAtnQEQaHcCBDo5QhDNYK4pTrOIVC5LFLXbRJCLYwHRuYCoP/kwmkVCAIZABiQXYMYhDLOIRXQg5pRUMjVS0Iha1yEUvpoQJsUhAD6JGyUpa8pKYzKQmN8nJTnryk6AMpShHScpSmvKUqEylKlfZnwm48pWwjKUsZ0nLWtoSln9LCBFuyUtZSqsh0uilMF9JpYaIYJjITOYtddaQBSjzmdCcwC8REsxo0rKYCBmENbeZTEAgBAHgDKc4x0nOcprznOgUJxMU8ot0upOcbGuIF95Jz3DWiiHHqKc+95n/zl8tJAX8DKhAETDJhMxzoOa850E+gdCG7jOerIyoRCdK0Ypa9KIYzahGN8rRjnr0oyANqUhHStKSmvSkKE2pSlfK0pKqAQhjWAYJZIOQSFzDIstgGUdWEIoDDCQMsDiJLGpBEDFkYyFWoJ1FRgHTMTgAhxwZgp38UAI+SNUjL43pTBNiU5zqdCM89alAgCoSpo7BqVCNSFcdQlQ+5PQiaogXC0qghlkgBBZ+rMjmwAoEMhhgrEE1SQEYACcTjAIJSFVqRUYRKGmUAAjx8kh98OAKgexVI3Hlw1zrete8UuSyGVlBX//KB7KGhLF8cCxkKYLXh/AQtE6JLB9AYAs+/7ACEATwgc5a0QIJDOO2uWWmHsJAgBVgYABWkAAoGkEAW5iADyvrgTV6cQFjHAq4uhXIcIu7kBW0IBu1LW1Q72CLUcAgnmN4rltNcAhN4EECW50IHQYgkAU4oL4+2IYtUpSNHpQiGVaoAgy2gQuKoFYgEvBDCjSBCzstIL/75YMxSiAMApjqGnRQAwnsaoIIjGIUBY7DLgaxjW0IQMR8IK9544mBN/jACmJ9SGYFQlvb4ja7vPUtdoVLXOMiV7nMdS50WSBd6lrXxsHVbo+7+93wklXF5xVIegWyDPa6F740ZciB+ZDgYQRAIEP4MjSSoQcZAOK5JyABGUjxjBwPA/8WzJBbnJmBiBZYIS/GCEUHpLEyPiQDBqMIAWnGXOYzK2TGfFgADGbbCFakgABDeJwfQdDoR0daDY2g3SiaYYJshGIaipCAbFZ2gBbEgRW1KDClHQ3prGSaId5lwQXmQtZABEIasxiFV6a83k8sQwB8aMWXJ3KAwDbjiHtQwzC8kA1j8KERMJiGNKwQAEUswQHYhAhqvZCMUMziBC2AxgqSvexm88EKerC2A97gBStUwASFcAC3fXCHAmghBSimLB9QbGtc65oPo5iAFxYwBtgmBNGKZjSrI93aVVva1ZrmtKdBLeohl/rUqVb4wzGt2ITEetbi5UO/c71r9VbZ18D/FrZDtt3tWXgZzF+WQgf2sIJtMMMLDlAGK7iRm9bCWc58cAACvFAIYUhZICs7BBmkHYEJ8EHmNLf5oWV7jBYEHbF8cMYRW+sArGs9K5EG+E0HgQiBWMB7pDbaAbzYdYF8XQ1hZzIflCEBVpBVC4fiAyCsw+uTj8E4n+gARVbAANI4IFEICOwK/t4ISlkBFAIRgNMlMgoGLGMM2yACH8D93MQLZPE8eHzkAxEHKwxEGEP4hQMWoN58VxbFeBfI3gEOeYA/qyGIrvrV3b51P7Y960eEu0BGMfay8+HsFlc7270e/LgrxLtzr/vd8z77vpvgE3/nQ+BXbnnMa/7lVo05/wQEIgVlLMBOKVaGpB8XZz6QIs7rZEUhCHD0IfdBE59fxtPH/3T1H1y2CQcEpdABBFhgrSWABNgBBaYGjTMKXcMMgRUIaMcCBxABAlEBXoSABZgVjSN3EgYCQMUCZSEQsuBFvMYA17dofHAMgkcRpFABcSABAuEKo5CAHWAAjVBQVhBZphURWyYQJ2B0fECDNmgAOwhYb4ABA0EK6vdeZOAM0uB6+7YLIjgQJQhwpEUAt8cQCLdoGqiA6/eFC9iADxiBE1iBF5iBA7iBDNgQ0PeBITiCfHCFJ5iCAsGCK6dQfAB+YfZ08dQLylAC2TMQPtd+78cHCAADpRAC9McHPP+0MtBQWVS2B1Lwh/6HEIiGC7W1DevEB0twKK3FiQLxiVnhHQ7IBxAoEBJocd7EBxjIB6LoiYeiBluIEG9YAGSAB0GlBZ3IDUQyBmZzDKFwfUKIhxThCzUAApFgHxYILIjVCP5nBXPBB9ZwJpSnhyfQgkHQjF6AWKJHjRNQegNRCkjwDH91DIiQDFIIe71IJKfIB1oIEZm4iZ1IiqFYj7NoimWoimfYiq8Yi6RIi25odXyAi7rIB7woEL7oiME4jJ9QjC2oZXo4DHRwgTFXUIAoAqPDB3vgGz4nAgJBAMxgAMsQLwXQiI/IAvcnEHuAglKAkZd4EJnFCj3AWdwQCKz/cAgE4EewMBc3mZM7WYrDt48i149pyAc/qZN+JJCwRpB80ActEFSBoAlecA2jEC9kQCXQAATEeIcRKRE80AIXYFd88AzbMATSgAtKCI0CgW48UAASMI0+iI0taJZoqZbnpgdvKQHK0G4IwAcl4ADS0AewwAR3AAW+kG/LyG9UaZXx8o7xKGPxQpM2iZNK+Tg+aZlByZSnmIpFyYpHmZSbWYsH8YZPGZUi15hXyQdZyQdb2ZUr+JUK8YN8cA3L8AysgAEXKRCAyAIEAArSYG49yQchwGlvAATMcA1qcAcm4AxkIGX1kXRkMAteAAi19ZK8GZNmYXktEF88YAHbIAPW/yAQVUAAw/Cd4TmeQglwRLmKpOaPXoSe4jkepGkQpukFPjBeUjAKVlBVfOALpSAMuwSbxkgRIWB6A/EGpaAGsFAAzxaNuOAAMlAFBkaXCbqgDXpuuGAFBJAHQEgHZLBhfMADxqAF2wAIXoBiv0AAfYBid7Cf/TmUIVmfmMid3gme88kH5XmeOKqenNmeRumK8dmj9DmQipKfKQaj/gmgAjoBBCqbCUGbrnkBu5AMR4WdfACIVrULaoABDrqj06AHDuBedtUMo4AICLALRxQCF8BnLNMDgCZoTweTLQUTSVWneFpSd5qnfApSb/BVfRqogjqohFqohsoRv5CoirqojP/aqI76qJAaqZI6qZRKqVgXUgZQqZq6qZzaqZ7KqEugFnQwqqRaqqZ6qqiaqqq6qqzaqq7qqs4mUm/wqrRaq7Z6q7hqqmnFFCbQq776q8AarMI6rMRarMZ6rMiarCSVrMzarM76rNAKrIc6rdRardZ6rdiardq6rdzard76reAaruI6ruRaruYaEmH1U4F1EWaFVtoHpef6EOkKWBnRrk/1ri0lWn5FrxiBWqr1CdsXrxOhr6TVg0vVWI8FsPBqUt4FXvQKZehlclb2XvE1EFuWYJ9QChAAX3HWX/9VAKRABqMgiUFwAZjGMs9ACqNQA/7UrQ3rZONVXlHmiBLbXhT/m2XDd08Yq7EcywcemwwgK7Ika7KNgLIqy7IS9XG0JpW3RnI0S2XX92vBNmwWGyjc5m2fEApBMIRfBm3SFggRwArHsAwG4F3HwAN0IEB0IAv3cwHy861KG3Ij92/Wh3JTWxAsh7Vay7XPFm3SALZiS7Zmi7Zqy7Zh4LYRBX10Z3e7SH18J7HYB3iyWXmXl3naxwAsk43PRil7wAomMAtacAwsoAYlsB5LoAXr0QHkArdWt7jTJ3uPC7WRi6/u0n2W+wmYu3mC13gs6bmgK7qka7qoKxCqm7gEaQwgCAtVSIImqF4o+AkqWKA5WxAZKxDbl4MCMQ1QIAHGQAZeoQxQ/7AMjfAMC7AMNggA4PqGyBuHVti8AvG80Tu5eli9+Iq9fKC93Ou9cxe+41u+52u8AmGQu9iODLmCDgmReDu/LXi9/gcDf8llx6AIwLMH0nQMKogm1uKtt5iLA6yQv9iQT5rA1LvAu9vADywBETzBFXzBs5DBqmSaUCmVqomVWsmVD+mVIkwQAcvAwxdnPRAK1xAHBHAC0nCTJtAB7/YGagCo3ArDqDmVVbmarfmaNxybOTwQO1zCPcwHPxzEQ1zEgXDESbzEAHwpSPqi/LmkATqgVSy9AKfA1qvFfDB/wgABEYAxIDAKakAK65QCekAGHdB+rHuk+pnGAsGkbIzA7v8Cx/Xrf3Rsx3isx3zMB34MyIIssJicyZq8yZzcyZ78yaAcyqI8yqRcyh4hAraWyqq8yqzcyq78yrAcy7I8y7Rcy7Z8y7j8ytppyitxDRXwy8AczMI8zMRczMZ8zMiczMq8zMzczM78zMXcsrw8zdRczdZ8zdiczdq8zdzczd78zeAczuI8zuRczgbWVPfqEVdFVVtaEY3AKfYqQAFrzh3lrwn7EZMliQbXEO+8yKmVsPNMzxt1sQrGYA4GYSkyYRV2YRkmoh32YSE2YiV2YnYCsQLRYi8mVtbgAFpAAl7Rz1U7EDu7sSTQsf4FtCE7svZBtEa7stIs0JiUt98WbuP/pmzM5mzopm7s5m7wJm/JQG/2hm92om/81rT/FnADV3AFsAzXIA2AYIEgPb1XOwtZu7Wu0LV+C7hjW7YtcLZpmyuGi7gwvUmU632b1wKdp3h/940CQHoIygeop3qsdyND/Xp2Ent6Zx2jUHujsAfS0DV34Ay98GzwbLuah7uZu7uc67uhO7qlO4rDywfFO9aaRJtBOIM1mIBGyIOwkIRL2IQS8IRRWNdTuLxz6EWjkIV74AXQcAFQgAGDHdVvTL2lEMfPVlD4273fy7/ka74JiL6UnUmWrY3c6I21Zw3h+NbkaI4rmI7reNcEDJl74Ac+QBqFENvwzMg8PD8nnMIc/7nCA9HCwS3cFlqWZ5mWSuiWcNmXVvCXgTmYhXmYiTnUi2knUOyY7DmjRFADXqAIdBA7sk2bWfygW9zFQkzERozEJqDETDzeUTPcF8qgDmoFG9qhHxqidkWiJoqiKsqiLqqkMgqPe6AIsLANAVACBKAMAa7dcvzIdzxbktzHfxzIDq5Ke1rj1XzjOD7Nf7rjPv7jQB7kIWECw1DkRn7kSJ7kSr7kTN7kTv7kUB7lUj7lVF7lVn7lUu6gvMwKudrlXv7lYB7mYj7mZK6quzrKXpDmar7mbN7mbv7mcB7ncj7ndF7ndn7neJ7ner7ndq5eQv7ngB7ogj7ohF7ohn7oiP+e6Iq+6IxuSkNVVEelEDqOt+gszwvb6CI1WIV1WIkVpQgLWQGN6SY1X/V1X4mG0D570gE2YAV2xSPdsz8btCrNByV7smV5tC8t6qNUbLOCbDZtbl47bdV2bdk221Nd1Xwb7FotuF1NuGDdtm+r66ZEeIaHeGrNA7x7brUneXhr2Jeb2Jvbu5/b2MEL2am7utJuSi8Yg5hdhPZ7hCHnzzpc2/WL29ur2/srvr39v+luSsiojMx4Kc8YjdNYjVds243MQN0twd/NDRYc3i7c76IUlmMpEHaJ3gSel3spl9OrwySc8aPgw0B84GAsxgtOxhJfSgdKEAoq4RlP4RJKoQf/T7vb7eKRvMcybskpv/M83/M+//NAH/RCP/QYpQlGf/RIn/RKv/RM3/RO//RQH/VSP/VUX/VWf/VYn/Vav/Vc3/Ve//VgH/XZcABkX/Zmf/Zon/Zqv/Zs3/Zu//ZwH/dyP/d0X/d2f/d4n/d6v/d83/d+//YPTPSCP/iEX/iGf/iIn/iKv/iM3/iOH9yfoLLb0Aih6hBr9fgXZQIOIAB3IA14YHwN0VqYf1FMEApnJAUs82D6lSKEZmYm4GY7dtEuBmMC4QcDFranvvqjr0kdoAfKEO3kdtNPN3M1F2cNV2mthtQEJw3HQADHsAIk0AfBb267j0ksgAckwAB6MAyI/3jtUsB/5bd+v/d1ez18e+AKkf4JSOB5fAB61a9JitAHY1AARKjZlZidYbiGYJjaM6oJ6gkQfPi4GtXBYAcDAhUuZNjQ4UOIESVOpFjR4kWMGTVu5NjR40eQITX6qrUwQKEgEQR6QcJHSg+BvZTxgbWAzzYmApc84zNqhUACe0BEEhgnWcqVLUUuZdrU6VOoUaVOpYrRAJkgPLyAGlXg2bYh0nBhcAmTj0yag/hwC8TqEAGbPoHuiSPj0J4AAr6GHVvV71/AgQUPJuw3BZxtagKIEPimlBpYBcrGnFmFwDAeFrbJsCZQLp+gfBBIGCVGGh/HkCUXZt3a9WvYsWXPpitd2/Zt3Ll17+bd2/dv4MGFDyde3Phx5MmVL2fe3Plz6NGlT6de3fp13wEBADs=" style="margin-top:1rem;" /></div></details></div></div>

## Wolfram Alpha

::: {.sourceClojure}
```clojure
(wl/eval '(WolframAlpha "number of moons of Saturn" "Result"))
```
:::

::: {.printedClojure}
```clojure
(Quantity 145 (IndependentUnit "moons"))

```
:::

::: {.sourceClojure}
```clojure
(view '(WolframAlpha "number of moons of Saturn" "Result"))
```
:::

<div style="max-height:600px;overflow-y:auto;"><div><details open="open"><summary><h5 style="cursor:pointer;display:inline;padding:0.5rem 1rem 0 1rem;">WolframAlpha["number of moons of Saturn", "Result"]</h5></summary><div class="wl-results"><hr /><img src="data:image/jpeg;base64,R0lGODlhYwAqAPdeAPHx8dvb29/f3+7u7tra2rqvoquej+bi3ZuLebuwpOjk4JWEcfDu7N/a1ZiHdZ2OfKudjrGklo58Z8zEu+np6d3Y0dnTzM3FvINwWdPMxNDIwPf29O7r6M/Iv8O5r+rn4/X08srBuLCjlY99adXOx66hkqCRgIJuV9bPyIVxW8/Hv/7+/oFtVuTg29zW0LWqnKyfkJqKeN/a1OTk5NvVz5uMeoVyXPr5+MG4rZF/a417Ztzc3L+1qZB+aq+ilNPNxY9+afX19Yt5ZIZzXePe2fn5+fr6+tnZ2eXl5YRxW6eZiresn/Lw7pmJdrarnvLy8ol3YYh1X8K4rdHKwtDJwce+tLyypvT09O3t7dTNxbKmmOPf2ritoIl2Yf////b29vv7+6WXhuDb1YFtV+fn59bQyfz8/O3q55+QfuHc1+Le2aqdjt3d3ejo6PDw8PPz8760qPTz8fDu68jAturn5O3r6Obm5vr6+aGSgbSpm/j4+H9rVJmJd4d0X4JuWIp4Y6GSgrWpnP39/fv6+svDuo17Z5aGc6KUg6OVhPb19JaFcoRwWunl4d7Z0/f399jSy9XPx4d0XpKAbevo5Ovr69zX0eLd2LmuofPy8Me/tefk38W8sYNvWaKTgufj3/39/NnTzbqvo+Dg4LSomqaYiJKBbaeZifn496+jlMC2q97e3uXh3IBsVZWFcpyNe9LLw5SDcPj39tHKwZOCbrernuPj48rCueHd1+/t6u7s6baqnfb185eHdJyMetTOxuzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5uzp5iH/C05FVFNDQVBFMi4wAwEAAAAh+QQECgD/ACwAAAAAYwAqAAAI/wC9CBxIsKDBgwgTKlzIsKHDhxAjSpxIsaLFixgzatzIsaPHjyBDihxZsMibJyhTqlzJsqXLlyivgCGpMIiqIzhz6tzJs6fPnzxrGaFpMMgRMkVWEE244ouoHWaWDlRlRypDQaraWDVyRI9VhlhUWb1yRNDXhW+OKF1Kdu1ZhG2lxn0LV+1BJlQO7TmDUAyrQAS17Bk8GHDBuXQNIibIQweUvQdv5LBhWOAaNDQy06Cj2G7izm4NyoBs0EmMMJW9HHqxcPFnga4JjuZbMM2YFnhSx8DR2i6LKUC6uFiSQolZL3Jg9EkCqAXBQaEK+TE0haDwMooWGWpE8BWfIUMQPf8iGHvg7IJ3JF3ygia1pF6SosR4dTAui1KTfAzJZIkVDS+Y6AALFZAAwskWAq3QSSRVgDLKHpkM1EUMpGhShylRnCJQJaykskojcLQSy0DlCXQeQVzMMgh7qaFCSyO36LLHJqB5wYItXrxSikB/zOFFHpHsIpAgvLgikC97cCeQLikkIlAXpKz1yx6VCCTFEJ/UFVpBJwqkxglECNReQiVAsaV9snhBQhMC6UDjCCIQFMIeIHjhQw8EfbAHCk96MNAne3QgEAdCNKGBkId5dlCXg8wSykBjIpTBHpgkqhQLaa7ZJo0Y8EDQI3t84AUiaBAUyx4h9EnQHhcMlEgqhoz/YYIF5CkqGmlqEqbrHhUcNMEeG1hqY6ZseuGmFxJEQJAte8ThRQk5EETHHiSoOhCrBuUiY5qw2colrht4Iq64TUDgiYaM1DHQCg84UCOmahZ7rC5ROOnFCk08IBAKezQwEBdBWisQtl4wcJxAfPhAordegJBGGnPs8UMalhwUqRcJDGEFDVmYcNu7xG7qxQY6LCALCp1wosBAYfSRCSiB7EGIhH5e26oXCCiiwgEHeMDCDwtvmeuunFhc2QpU9DLEIodoUp9vIRtLoxe4GLAcHggOdEcBOkxHhXU1D3zzBjg4kEQkNdAXtFwMf1YiSG+/FbdHc39VN0d3sy20SI4cVTHUawS5sYNVKxAwAOAEzTDDV1gQ8AbiK1ByhCNfrUDBEQIgofnmnHfu+eegh775DGwEAABdjmBhh+ist+46EmQMMBPitNdu++2456777rz37jtIAQEAOw==" style="margin-top:1rem;" /></div></details></div></div>

<div style="background-color:grey;height:2px;width:100%;"></div>

<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/main/projects/math/wolframite/notebooks/index.clj">projects/math/wolframite/notebooks/index.clj</a></small></small></pre></div>