### 2024-02-01 docs/projects/noj/ggplot.qmd
noj - rerendered book

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: ggplot.html}
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="ggplot_files/md-default0.js" type="text/javascript"></script><script src="ggplot_files/md-default1.js" type="text/javascript"></script>
<div style="max-height:600px;overflow-y:auto;"><table><tr><td>This is part of the Scicloj <a href="https://scicloj.github.io/clojure-data-scrapbook/">Clojure Data Scrapbook</a>.</td><a href="https://scicloj.github.io/clojure-data-scrapbook/"><img align="left" alt="SciCloj logo" src="https://scicloj.github.io/sci-cloj-logo-transparent.png" width="40" /></a></tr></table></div>

# Exploring ggplot

::: {.sourceClojure}
```clojure
(ns ggplot
  (:require [clojisr.v1.r :as r :refer [r]]
            [clojisr.v1.applications.plotting :as plotting]
            [scicloj.kindly.v4.kind :as kind]))
```
:::

::: {.sourceClojure}
```clojure
(r/library "ggplot2")
```
:::

::: {.printedClojure}
```clojure
[1] "ggplot2"   "Rserve"    "stats"     "graphics"  "grDevices" "utils"    
[7] "datasets"  "methods"   "base"

```
:::

::: {.sourceClojure}
```clojure
(def plot
  (r "(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method=\"lm\")
         + facet_wrap(~cyl))"))
```
:::

::: {.sourceClojure}
```clojure
(-> plot
    plotting/plot->svg
    kind/html)
```
:::

<?xml version="1.0" encoding="UTF-8"?>
<svg xmlns="http://www.w3.org/2000/svg" xmlns:xlink="http://www.w3.org/1999/xlink" width="504pt" height="504pt" viewBox="0 0 504 504" version="1.1">
<defs>
<g>
<symbol overflow="visible" id="glyph0-0">
<path style="stroke:none;" d=""/>
</symbol>
<symbol overflow="visible" id="glyph0-1">
<path style="stroke:none;" d="M 4.375 -4.703125 C 4.234375 -5.75 3.5625 -6.359375 2.609375 -6.359375 C 1.921875 -6.359375 1.3125 -6.03125 0.9375 -5.453125 C 0.546875 -4.828125 0.375 -4.0625 0.375 -2.90625 C 0.375 -1.828125 0.53125 -1.15625 0.90625 -0.59375 C 1.234375 -0.078125 1.78125 0.203125 2.46875 0.203125 C 3.65625 0.203125 4.515625 -0.6875 4.515625 -1.921875 C 4.515625 -3.09375 3.71875 -3.921875 2.609375 -3.921875 C 1.984375 -3.921875 1.5 -3.6875 1.171875 -3.234375 C 1.171875 -4.796875 1.671875 -5.671875 2.5625 -5.671875 C 3.109375 -5.671875 3.484375 -5.328125 3.609375 -4.703125 Z M 2.5 -3.234375 C 3.25 -3.234375 3.71875 -2.703125 3.71875 -1.859375 C 3.71875 -1.0625 3.1875 -0.484375 2.484375 -0.484375 C 1.765625 -0.484375 1.21875 -1.09375 1.21875 -1.90625 C 1.21875 -2.6875 1.734375 -3.234375 2.5 -3.234375 Z M 2.5 -3.234375 "/>
</symbol>
<symbol overflow="visible" id="glyph0-2">
<path style="stroke:none;" d="M 3.4375 -3.359375 C 4.09375 -3.75 4.296875 -4.0625 4.296875 -4.65625 C 4.296875 -5.671875 3.53125 -6.359375 2.421875 -6.359375 C 1.3125 -6.359375 0.546875 -5.671875 0.546875 -4.671875 C 0.546875 -4.0625 0.75 -3.75 1.390625 -3.359375 C 0.671875 -2.984375 0.328125 -2.46875 0.328125 -1.765625 C 0.328125 -0.609375 1.171875 0.203125 2.421875 0.203125 C 3.65625 0.203125 4.515625 -0.609375 4.515625 -1.765625 C 4.515625 -2.46875 4.15625 -2.984375 3.4375 -3.359375 Z M 2.421875 -5.671875 C 3.078125 -5.671875 3.5 -5.28125 3.5 -4.640625 C 3.5 -4.046875 3.0625 -3.65625 2.421875 -3.65625 C 1.765625 -3.65625 1.34375 -4.046875 1.34375 -4.65625 C 1.34375 -5.28125 1.765625 -5.671875 2.421875 -5.671875 Z M 2.421875 -3 C 3.1875 -3 3.71875 -2.484375 3.71875 -1.75 C 3.71875 -0.984375 3.203125 -0.484375 2.40625 -0.484375 C 1.640625 -0.484375 1.125 -1 1.125 -1.734375 C 1.125 -2.5 1.640625 -3 2.421875 -3 Z M 2.421875 -3 "/>
</symbol>
<symbol overflow="visible" id="glyph0-3">
<path style="stroke:none;" d="M 2.875 -1.546875 L 2.875 0 L 3.65625 0 L 3.65625 -1.546875 L 4.578125 -1.546875 L 4.578125 -2.234375 L 3.65625 -2.234375 L 3.65625 -6.359375 L 3.078125 -6.359375 L 0.25 -2.359375 L 0.25 -1.546875 Z M 2.875 -2.234375 L 0.921875 -2.234375 L 2.875 -5.046875 Z M 2.875 -2.234375 "/>
</symbol>
<symbol overflow="visible" id="glyph0-4">
<path style="stroke:none;" d="M 4.1875 -6.234375 L 0.96875 -6.234375 L 0.5 -2.84375 L 1.21875 -2.84375 C 1.578125 -3.265625 1.875 -3.421875 2.359375 -3.421875 C 3.203125 -3.421875 3.71875 -2.84375 3.71875 -1.921875 C 3.71875 -1.03125 3.203125 -0.484375 2.359375 -0.484375 C 1.6875 -0.484375 1.265625 -0.828125 1.078125 -1.53125 L 0.3125 -1.53125 C 0.40625 -1.015625 0.5 -0.78125 0.6875 -0.546875 C 1.03125 -0.078125 1.671875 0.203125 2.375 0.203125 C 3.640625 0.203125 4.515625 -0.71875 4.515625 -2.03125 C 4.515625 -3.265625 3.6875 -4.109375 2.5 -4.109375 C 2.0625 -4.109375 1.703125 -4 1.34375 -3.734375 L 1.59375 -5.46875 L 4.1875 -5.46875 Z M 4.1875 -6.234375 "/>
</symbol>
<symbol overflow="visible" id="glyph0-5">
<path style="stroke:none;" d="M 2.28125 -4.53125 L 2.28125 0 L 3.046875 0 L 3.046875 -6.359375 L 2.546875 -6.359375 C 2.265625 -5.390625 2.09375 -5.25 0.890625 -5.09375 L 0.890625 -4.53125 Z M 2.28125 -4.53125 "/>
</symbol>
<symbol overflow="visible" id="glyph0-6">
<path style="stroke:none;" d="M 2.421875 -6.359375 C 1.84375 -6.359375 1.3125 -6.09375 0.984375 -5.671875 C 0.578125 -5.109375 0.375 -4.25 0.375 -3.078125 C 0.375 -0.9375 1.078125 0.203125 2.421875 0.203125 C 3.734375 0.203125 4.453125 -0.9375 4.453125 -3.03125 C 4.453125 -4.25 4.265625 -5.078125 3.859375 -5.671875 C 3.53125 -6.109375 3.015625 -6.359375 2.421875 -6.359375 Z M 2.421875 -5.671875 C 3.25 -5.671875 3.671875 -4.828125 3.671875 -3.09375 C 3.671875 -1.28125 3.265625 -0.4375 2.40625 -0.4375 C 1.578125 -0.4375 1.171875 -1.3125 1.171875 -3.0625 C 1.171875 -4.828125 1.578125 -5.671875 2.421875 -5.671875 Z M 2.421875 -5.671875 "/>
</symbol>
<symbol overflow="visible" id="glyph0-7">
<path style="stroke:none;" d="M 4.453125 -0.765625 L 1.171875 -0.765625 C 1.25 -1.296875 1.53125 -1.625 2.296875 -2.09375 L 3.171875 -2.59375 C 4.046875 -3.0625 4.5 -3.71875 4.5 -4.5 C 4.5 -5.03125 4.28125 -5.53125 3.921875 -5.875 C 3.546875 -6.203125 3.09375 -6.359375 2.5 -6.359375 C 1.703125 -6.359375 1.125 -6.078125 0.78125 -5.53125 C 0.546875 -5.203125 0.453125 -4.796875 0.4375 -4.15625 L 1.21875 -4.15625 C 1.234375 -4.59375 1.296875 -4.84375 1.40625 -5.0625 C 1.59375 -5.4375 2 -5.6875 2.46875 -5.6875 C 3.171875 -5.6875 3.703125 -5.171875 3.703125 -4.484375 C 3.703125 -3.96875 3.40625 -3.53125 2.859375 -3.21875 L 2.046875 -2.75 C 0.75 -2 0.375 -1.40625 0.296875 -0.015625 L 4.453125 -0.015625 Z M 4.453125 -0.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph0-8">
<path style="stroke:none;" d="M 1.9375 -2.921875 L 2.359375 -2.921875 C 3.203125 -2.921875 3.65625 -2.53125 3.65625 -1.765625 C 3.65625 -0.96875 3.171875 -0.484375 2.375 -0.484375 C 1.515625 -0.484375 1.109375 -0.921875 1.0625 -1.84375 L 0.28125 -1.84375 C 0.3125 -1.34375 0.40625 -1 0.546875 -0.71875 C 0.875 -0.109375 1.484375 0.203125 2.34375 0.203125 C 3.625 0.203125 4.453125 -0.578125 4.453125 -1.78125 C 4.453125 -2.59375 4.140625 -3.03125 3.390625 -3.296875 C 3.96875 -3.53125 4.265625 -3.96875 4.265625 -4.609375 C 4.265625 -5.703125 3.546875 -6.359375 2.359375 -6.359375 C 1.109375 -6.359375 0.4375 -5.65625 0.40625 -4.3125 L 1.1875 -4.3125 C 1.203125 -4.703125 1.234375 -4.921875 1.328125 -5.109375 C 1.5 -5.46875 1.890625 -5.6875 2.375 -5.6875 C 3.0625 -5.6875 3.46875 -5.265625 3.46875 -4.578125 C 3.46875 -4.140625 3.3125 -3.859375 2.96875 -3.71875 C 2.765625 -3.625 2.484375 -3.59375 1.9375 -3.578125 Z M 1.9375 -2.921875 "/>
</symbol>
<symbol overflow="visible" id="glyph1-0">
<path style="stroke:none;" d=""/>
</symbol>
<symbol overflow="visible" id="glyph1-1">
<path style="stroke:none;" d="M 5.1875 -3.828125 C 5.140625 -4.390625 5.015625 -4.75 4.796875 -5.078125 C 4.40625 -5.609375 3.703125 -5.921875 2.90625 -5.921875 C 1.34375 -5.921875 0.34375 -4.703125 0.34375 -2.78125 C 0.34375 -0.921875 1.328125 0.25 2.890625 0.25 C 4.265625 0.25 5.140625 -0.578125 5.25 -1.984375 L 4.328125 -1.984375 C 4.171875 -1.0625 3.703125 -0.59375 2.921875 -0.59375 C 1.90625 -0.59375 1.296875 -1.421875 1.296875 -2.78125 C 1.296875 -4.21875 1.890625 -5.078125 2.890625 -5.078125 C 3.65625 -5.078125 4.140625 -4.625 4.25 -3.828125 Z M 5.1875 -3.828125 "/>
</symbol>
<symbol overflow="visible" id="glyph1-2">
<path style="stroke:none;" d="M 2.796875 -5.765625 L 1.84375 -5.765625 L 1.84375 -7.34375 L 0.9375 -7.34375 L 0.9375 -5.765625 L 0.15625 -5.765625 L 0.15625 -5.015625 L 0.9375 -5.015625 L 0.9375 -0.65625 C 0.9375 -0.0625 1.328125 0.25 2.046875 0.25 C 2.28125 0.25 2.484375 0.234375 2.796875 0.171875 L 2.796875 -0.59375 C 2.65625 -0.5625 2.546875 -0.546875 2.359375 -0.546875 C 1.953125 -0.546875 1.84375 -0.65625 1.84375 -1.0625 L 1.84375 -5.015625 L 2.796875 -5.015625 Z M 2.796875 -5.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-3">
<path style="stroke:none;" d="M 4.265625 -5.765625 L 2.671875 -1.28125 L 1.203125 -5.765625 L 0.21875 -5.765625 L 2.171875 0.015625 L 1.8125 0.9375 C 1.65625 1.34375 1.46875 1.5 1.078125 1.5 C 0.921875 1.5 0.796875 1.46875 0.59375 1.4375 L 0.59375 2.25 C 0.78125 2.359375 0.96875 2.390625 1.203125 2.390625 C 1.5 2.390625 1.828125 2.296875 2.0625 2.125 C 2.359375 1.90625 2.515625 1.671875 2.6875 1.203125 L 5.265625 -5.765625 Z M 4.265625 -5.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-4">
<path style="stroke:none;" d="M 2.84375 -5.765625 L 1.875 -5.765625 L 1.875 -6.671875 C 1.875 -7.046875 2.09375 -7.25 2.515625 -7.25 C 2.59375 -7.25 2.625 -7.25 2.84375 -7.234375 L 2.84375 -8 C 2.625 -8.046875 2.515625 -8.046875 2.328125 -8.046875 C 1.46875 -8.046875 0.96875 -7.5625 0.96875 -6.75 L 0.96875 -5.765625 L 0.203125 -5.765625 L 0.203125 -5.015625 L 0.96875 -5.015625 L 0.96875 0 L 1.875 0 L 1.875 -5.015625 L 2.84375 -5.015625 Z M 2.84375 -5.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-5">
<path style="stroke:none;" d="M 5.890625 -0.53125 C 5.78125 -0.515625 5.734375 -0.515625 5.6875 -0.515625 C 5.375 -0.515625 5.1875 -0.6875 5.1875 -0.96875 L 5.1875 -4.359375 C 5.1875 -5.375 4.4375 -5.921875 3.03125 -5.921875 C 2.171875 -5.921875 1.5 -5.6875 1.109375 -5.265625 C 0.84375 -4.96875 0.734375 -4.625 0.71875 -4.0625 L 1.640625 -4.0625 C 1.71875 -4.765625 2.140625 -5.078125 2.984375 -5.078125 C 3.828125 -5.078125 4.28125 -4.78125 4.28125 -4.21875 L 4.28125 -3.984375 C 4.265625 -3.59375 4.0625 -3.4375 3.328125 -3.34375 C 2.03125 -3.171875 1.828125 -3.140625 1.46875 -2.984375 C 0.796875 -2.703125 0.46875 -2.203125 0.46875 -1.453125 C 0.46875 -0.40625 1.1875 0.25 2.359375 0.25 C 3.078125 0.25 3.65625 0 4.3125 -0.59375 C 4.375 0 4.65625 0.25 5.265625 0.25 C 5.453125 0.25 5.578125 0.234375 5.890625 0.15625 Z M 4.28125 -1.8125 C 4.28125 -1.5 4.1875 -1.3125 3.921875 -1.0625 C 3.546875 -0.71875 3.09375 -0.546875 2.546875 -0.546875 C 1.84375 -0.546875 1.421875 -0.890625 1.421875 -1.46875 C 1.421875 -2.078125 1.8125 -2.390625 2.8125 -2.53125 C 3.78125 -2.65625 3.96875 -2.703125 4.28125 -2.84375 Z M 4.28125 -1.8125 "/>
</symbol>
<symbol overflow="visible" id="glyph1-6">
<path style="stroke:none;" d="M 2.984375 -5.921875 C 1.359375 -5.921875 0.390625 -4.78125 0.390625 -2.84375 C 0.390625 -0.890625 1.359375 0.25 3 0.25 C 4.625 0.25 5.609375 -0.90625 5.609375 -2.796875 C 5.609375 -4.796875 4.65625 -5.921875 2.984375 -5.921875 Z M 3 -5.078125 C 4.03125 -5.078125 4.65625 -4.234375 4.65625 -2.8125 C 4.65625 -1.4375 4.015625 -0.59375 3 -0.59375 C 1.984375 -0.59375 1.359375 -1.4375 1.359375 -2.84375 C 1.359375 -4.234375 1.984375 -5.078125 3 -5.078125 Z M 3 -5.078125 "/>
</symbol>
<symbol overflow="visible" id="glyph1-7">
<path style="stroke:none;" d="M 0.765625 -5.765625 L 0.765625 0 L 1.6875 0 L 1.6875 -2.984375 C 1.6875 -4.375 2.265625 -5 3.53125 -4.96875 L 3.53125 -5.890625 C 3.375 -5.921875 3.28125 -5.921875 3.171875 -5.921875 C 2.578125 -5.921875 2.140625 -5.578125 1.609375 -4.71875 L 1.609375 -5.765625 Z M 0.765625 -5.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-8">
<path style="stroke:none;" d="M 2.59375 -8.015625 C 1.5 -6.578125 0.796875 -4.578125 0.796875 -2.84375 C 0.796875 -1.109375 1.5 0.890625 2.59375 2.328125 L 3.203125 2.328125 C 2.234375 0.765625 1.6875 -1.09375 1.6875 -2.84375 C 1.6875 -4.59375 2.234375 -6.453125 3.203125 -8.015625 Z M 2.59375 -8.015625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-9">
<path style="stroke:none;" d="M 1.671875 -8.015625 L 0.75 -8.015625 L 0.75 0 L 1.671875 0 Z M 1.671875 -8.015625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-10">
<path style="stroke:none;" d="M 1.015625 2.328125 C 2.125 0.890625 2.8125 -1.109375 2.8125 -2.84375 C 2.8125 -4.578125 2.125 -6.578125 1.015625 -8.015625 L 0.421875 -8.015625 C 1.390625 -6.453125 1.921875 -4.59375 1.921875 -2.84375 C 1.921875 -1.09375 1.390625 0.765625 0.421875 2.328125 Z M 1.015625 2.328125 "/>
</symbol>
<symbol overflow="visible" id="glyph2-0">
<path style="stroke:none;" d=""/>
</symbol>
<symbol overflow="visible" id="glyph2-1">
<path style="stroke:none;" d="M -8.015625 -0.765625 L 0 -0.765625 L 0 -1.6875 L -3.171875 -1.6875 C -4.359375 -1.6875 -5.125 -2.296875 -5.125 -3.25 C -5.125 -3.546875 -5.03125 -3.84375 -4.875 -4.0625 C -4.671875 -4.328125 -4.40625 -4.4375 -4 -4.4375 L 0 -4.4375 L 0 -5.34375 L -4.359375 -5.34375 C -5.328125 -5.34375 -5.921875 -4.65625 -5.921875 -3.53125 C -5.921875 -2.71875 -5.671875 -2.21875 -4.96875 -1.6875 L -8.015625 -1.6875 Z M -8.015625 -0.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph2-2">
<path style="stroke:none;" d="M 0 -6.09375 L -5.765625 -7.78125 L -5.765625 -6.75 L -1.28125 -5.609375 L -5.765625 -4.484375 L -5.765625 -3.359375 L -1.28125 -2.25 L -5.765625 -1.078125 L -5.765625 -0.0625 L 0 -1.734375 L 0 -2.765625 L -4.515625 -3.890625 L 0 -5.046875 Z M 0 -6.09375 "/>
</symbol>
<symbol overflow="visible" id="glyph2-3">
<path style="stroke:none;" d="M -5.765625 -4.265625 L -1.28125 -2.671875 L -5.765625 -1.203125 L -5.765625 -0.21875 L 0.015625 -2.171875 L 0.9375 -1.8125 C 1.34375 -1.65625 1.5 -1.46875 1.5 -1.078125 C 1.5 -0.921875 1.46875 -0.796875 1.4375 -0.59375 L 2.25 -0.59375 C 2.359375 -0.78125 2.390625 -0.96875 2.390625 -1.203125 C 2.390625 -1.5 2.296875 -1.828125 2.125 -2.0625 C 1.90625 -2.359375 1.671875 -2.515625 1.203125 -2.6875 L -5.765625 -5.265625 Z M -5.765625 -4.265625 "/>
</symbol>
</g>
<clipPath id="clip1">
  <path d="M 35.152344 23.246094 L 229.136719 23.246094 L 229.136719 234.925781 L 35.152344 234.925781 Z M 35.152344 23.246094 "/>
</clipPath>
<clipPath id="clip2">
  <path d="M 35.152344 203 L 229.136719 203 L 229.136719 205 L 35.152344 205 Z M 35.152344 203 "/>
</clipPath>
<clipPath id="clip3">
  <path d="M 35.152344 150 L 229.136719 150 L 229.136719 152 L 35.152344 152 Z M 35.152344 150 "/>
</clipPath>
<clipPath id="clip4">
  <path d="M 35.152344 97 L 229.136719 97 L 229.136719 98 L 35.152344 98 Z M 35.152344 97 "/>
</clipPath>
<clipPath id="clip5">
  <path d="M 35.152344 44 L 229.136719 44 L 229.136719 45 L 35.152344 45 Z M 35.152344 44 "/>
</clipPath>
<clipPath id="clip6">
  <path d="M 67 23.246094 L 68 23.246094 L 68 234.925781 L 67 234.925781 Z M 67 23.246094 "/>
</clipPath>
<clipPath id="clip7">
  <path d="M 101 23.246094 L 102 23.246094 L 102 234.925781 L 101 234.925781 Z M 101 23.246094 "/>
</clipPath>
<clipPath id="clip8">
  <path d="M 135 23.246094 L 136 23.246094 L 136 234.925781 L 135 234.925781 Z M 135 23.246094 "/>
</clipPath>
<clipPath id="clip9">
  <path d="M 169 23.246094 L 170 23.246094 L 170 234.925781 L 169 234.925781 Z M 169 23.246094 "/>
</clipPath>
<clipPath id="clip10">
  <path d="M 203 23.246094 L 204 23.246094 L 204 234.925781 L 203 234.925781 Z M 203 23.246094 "/>
</clipPath>
<clipPath id="clip11">
  <path d="M 35.152344 230 L 229.136719 230 L 229.136719 232 L 35.152344 232 Z M 35.152344 230 "/>
</clipPath>
<clipPath id="clip12">
  <path d="M 35.152344 177 L 229.136719 177 L 229.136719 179 L 35.152344 179 Z M 35.152344 177 "/>
</clipPath>
<clipPath id="clip13">
  <path d="M 35.152344 123 L 229.136719 123 L 229.136719 125 L 35.152344 125 Z M 35.152344 123 "/>
</clipPath>
<clipPath id="clip14">
  <path d="M 35.152344 70 L 229.136719 70 L 229.136719 72 L 35.152344 72 Z M 35.152344 70 "/>
</clipPath>
<clipPath id="clip15">
  <path d="M 50 23.246094 L 52 23.246094 L 52 234.925781 L 50 234.925781 Z M 50 23.246094 "/>
</clipPath>
<clipPath id="clip16">
  <path d="M 84 23.246094 L 86 23.246094 L 86 234.925781 L 84 234.925781 Z M 84 23.246094 "/>
</clipPath>
<clipPath id="clip17">
  <path d="M 118 23.246094 L 120 23.246094 L 120 234.925781 L 118 234.925781 Z M 118 23.246094 "/>
</clipPath>
<clipPath id="clip18">
  <path d="M 151 23.246094 L 154 23.246094 L 154 234.925781 L 151 234.925781 Z M 151 23.246094 "/>
</clipPath>
<clipPath id="clip19">
  <path d="M 185 23.246094 L 187 23.246094 L 187 234.925781 L 185 234.925781 Z M 185 23.246094 "/>
</clipPath>
<clipPath id="clip20">
  <path d="M 219 23.246094 L 221 23.246094 L 221 234.925781 L 219 234.925781 Z M 219 23.246094 "/>
</clipPath>
<clipPath id="clip21">
  <path d="M 35.152344 258.171875 L 229.136719 258.171875 L 229.136719 469.851562 L 35.152344 469.851562 Z M 35.152344 258.171875 "/>
</clipPath>
<clipPath id="clip22">
  <path d="M 35.152344 438 L 229.136719 438 L 229.136719 440 L 35.152344 440 Z M 35.152344 438 "/>
</clipPath>
<clipPath id="clip23">
  <path d="M 35.152344 385 L 229.136719 385 L 229.136719 387 L 35.152344 387 Z M 35.152344 385 "/>
</clipPath>
<clipPath id="clip24">
  <path d="M 35.152344 332 L 229.136719 332 L 229.136719 333 L 35.152344 333 Z M 35.152344 332 "/>
</clipPath>
<clipPath id="clip25">
  <path d="M 35.152344 278 L 229.136719 278 L 229.136719 280 L 35.152344 280 Z M 35.152344 278 "/>
</clipPath>
<clipPath id="clip26">
  <path d="M 67 258.171875 L 68 258.171875 L 68 469.851562 L 67 469.851562 Z M 67 258.171875 "/>
</clipPath>
<clipPath id="clip27">
  <path d="M 101 258.171875 L 102 258.171875 L 102 469.851562 L 101 469.851562 Z M 101 258.171875 "/>
</clipPath>
<clipPath id="clip28">
  <path d="M 135 258.171875 L 136 258.171875 L 136 469.851562 L 135 469.851562 Z M 135 258.171875 "/>
</clipPath>
<clipPath id="clip29">
  <path d="M 169 258.171875 L 170 258.171875 L 170 469.851562 L 169 469.851562 Z M 169 258.171875 "/>
</clipPath>
<clipPath id="clip30">
  <path d="M 203 258.171875 L 204 258.171875 L 204 469.851562 L 203 469.851562 Z M 203 258.171875 "/>
</clipPath>
<clipPath id="clip31">
  <path d="M 35.152344 465 L 229.136719 465 L 229.136719 467 L 35.152344 467 Z M 35.152344 465 "/>
</clipPath>
<clipPath id="clip32">
  <path d="M 35.152344 411 L 229.136719 411 L 229.136719 414 L 35.152344 414 Z M 35.152344 411 "/>
</clipPath>
<clipPath id="clip33">
  <path d="M 35.152344 358 L 229.136719 358 L 229.136719 360 L 35.152344 360 Z M 35.152344 358 "/>
</clipPath>
<clipPath id="clip34">
  <path d="M 35.152344 305 L 229.136719 305 L 229.136719 307 L 35.152344 307 Z M 35.152344 305 "/>
</clipPath>
<clipPath id="clip35">
  <path d="M 50 258.171875 L 52 258.171875 L 52 469.851562 L 50 469.851562 Z M 50 258.171875 "/>
</clipPath>
<clipPath id="clip36">
  <path d="M 84 258.171875 L 86 258.171875 L 86 469.851562 L 84 469.851562 Z M 84 258.171875 "/>
</clipPath>
<clipPath id="clip37">
  <path d="M 118 258.171875 L 120 258.171875 L 120 469.851562 L 118 469.851562 Z M 118 258.171875 "/>
</clipPath>
<clipPath id="clip38">
  <path d="M 151 258.171875 L 154 258.171875 L 154 469.851562 L 151 469.851562 Z M 151 258.171875 "/>
</clipPath>
<clipPath id="clip39">
  <path d="M 185 258.171875 L 187 258.171875 L 187 469.851562 L 185 469.851562 Z M 185 258.171875 "/>
</clipPath>
<clipPath id="clip40">
  <path d="M 219 258.171875 L 221 258.171875 L 221 469.851562 L 219 469.851562 Z M 219 258.171875 "/>
</clipPath>
<clipPath id="clip41">
  <path d="M 234.617188 23.246094 L 428.601562 23.246094 L 428.601562 234.925781 L 234.617188 234.925781 Z M 234.617188 23.246094 "/>
</clipPath>
<clipPath id="clip42">
  <path d="M 234.617188 203 L 428.601562 203 L 428.601562 205 L 234.617188 205 Z M 234.617188 203 "/>
</clipPath>
<clipPath id="clip43">
  <path d="M 234.617188 150 L 428.601562 150 L 428.601562 152 L 234.617188 152 Z M 234.617188 150 "/>
</clipPath>
<clipPath id="clip44">
  <path d="M 234.617188 97 L 428.601562 97 L 428.601562 98 L 234.617188 98 Z M 234.617188 97 "/>
</clipPath>
<clipPath id="clip45">
  <path d="M 234.617188 44 L 428.601562 44 L 428.601562 45 L 234.617188 45 Z M 234.617188 44 "/>
</clipPath>
<clipPath id="clip46">
  <path d="M 266 23.246094 L 268 23.246094 L 268 234.925781 L 266 234.925781 Z M 266 23.246094 "/>
</clipPath>
<clipPath id="clip47">
  <path d="M 300 23.246094 L 302 23.246094 L 302 234.925781 L 300 234.925781 Z M 300 23.246094 "/>
</clipPath>
<clipPath id="clip48">
  <path d="M 334 23.246094 L 336 23.246094 L 336 234.925781 L 334 234.925781 Z M 334 23.246094 "/>
</clipPath>
<clipPath id="clip49">
  <path d="M 368 23.246094 L 370 23.246094 L 370 234.925781 L 368 234.925781 Z M 368 23.246094 "/>
</clipPath>
<clipPath id="clip50">
  <path d="M 402 23.246094 L 404 23.246094 L 404 234.925781 L 402 234.925781 Z M 402 23.246094 "/>
</clipPath>
<clipPath id="clip51">
  <path d="M 234.617188 230 L 428.601562 230 L 428.601562 232 L 234.617188 232 Z M 234.617188 230 "/>
</clipPath>
<clipPath id="clip52">
  <path d="M 234.617188 177 L 428.601562 177 L 428.601562 179 L 234.617188 179 Z M 234.617188 177 "/>
</clipPath>
<clipPath id="clip53">
  <path d="M 234.617188 123 L 428.601562 123 L 428.601562 125 L 234.617188 125 Z M 234.617188 123 "/>
</clipPath>
<clipPath id="clip54">
  <path d="M 234.617188 70 L 428.601562 70 L 428.601562 72 L 234.617188 72 Z M 234.617188 70 "/>
</clipPath>
<clipPath id="clip55">
  <path d="M 249 23.246094 L 251 23.246094 L 251 234.925781 L 249 234.925781 Z M 249 23.246094 "/>
</clipPath>
<clipPath id="clip56">
  <path d="M 283 23.246094 L 285 23.246094 L 285 234.925781 L 283 234.925781 Z M 283 23.246094 "/>
</clipPath>
<clipPath id="clip57">
  <path d="M 317 23.246094 L 319 23.246094 L 319 234.925781 L 317 234.925781 Z M 317 23.246094 "/>
</clipPath>
<clipPath id="clip58">
  <path d="M 351 23.246094 L 353 23.246094 L 353 234.925781 L 351 234.925781 Z M 351 23.246094 "/>
</clipPath>
<clipPath id="clip59">
  <path d="M 385 23.246094 L 387 23.246094 L 387 234.925781 L 385 234.925781 Z M 385 23.246094 "/>
</clipPath>
<clipPath id="clip60">
  <path d="M 419 23.246094 L 421 23.246094 L 421 234.925781 L 419 234.925781 Z M 419 23.246094 "/>
</clipPath>
<clipPath id="clip61">
  <path d="M 234.617188 258.171875 L 428.601562 258.171875 L 428.601562 469.851562 L 234.617188 469.851562 Z M 234.617188 258.171875 "/>
</clipPath>
<clipPath id="clip62">
  <path d="M 234.617188 438 L 428.601562 438 L 428.601562 440 L 234.617188 440 Z M 234.617188 438 "/>
</clipPath>
<clipPath id="clip63">
  <path d="M 234.617188 385 L 428.601562 385 L 428.601562 387 L 234.617188 387 Z M 234.617188 385 "/>
</clipPath>
<clipPath id="clip64">
  <path d="M 234.617188 332 L 428.601562 332 L 428.601562 333 L 234.617188 333 Z M 234.617188 332 "/>
</clipPath>
<clipPath id="clip65">
  <path d="M 234.617188 278 L 428.601562 278 L 428.601562 280 L 234.617188 280 Z M 234.617188 278 "/>
</clipPath>
<clipPath id="clip66">
  <path d="M 266 258.171875 L 268 258.171875 L 268 469.851562 L 266 469.851562 Z M 266 258.171875 "/>
</clipPath>
<clipPath id="clip67">
  <path d="M 300 258.171875 L 302 258.171875 L 302 469.851562 L 300 469.851562 Z M 300 258.171875 "/>
</clipPath>
<clipPath id="clip68">
  <path d="M 334 258.171875 L 336 258.171875 L 336 469.851562 L 334 469.851562 Z M 334 258.171875 "/>
</clipPath>
<clipPath id="clip69">
  <path d="M 368 258.171875 L 370 258.171875 L 370 469.851562 L 368 469.851562 Z M 368 258.171875 "/>
</clipPath>
<clipPath id="clip70">
  <path d="M 402 258.171875 L 404 258.171875 L 404 469.851562 L 402 469.851562 Z M 402 258.171875 "/>
</clipPath>
<clipPath id="clip71">
  <path d="M 234.617188 465 L 428.601562 465 L 428.601562 467 L 234.617188 467 Z M 234.617188 465 "/>
</clipPath>
<clipPath id="clip72">
  <path d="M 234.617188 411 L 428.601562 411 L 428.601562 414 L 234.617188 414 Z M 234.617188 411 "/>
</clipPath>
<clipPath id="clip73">
  <path d="M 234.617188 358 L 428.601562 358 L 428.601562 360 L 234.617188 360 Z M 234.617188 358 "/>
</clipPath>
<clipPath id="clip74">
  <path d="M 234.617188 305 L 428.601562 305 L 428.601562 307 L 234.617188 307 Z M 234.617188 305 "/>
</clipPath>
<clipPath id="clip75">
  <path d="M 249 258.171875 L 251 258.171875 L 251 469.851562 L 249 469.851562 Z M 249 258.171875 "/>
</clipPath>
<clipPath id="clip76">
  <path d="M 283 258.171875 L 285 258.171875 L 285 469.851562 L 283 469.851562 Z M 283 258.171875 "/>
</clipPath>
<clipPath id="clip77">
  <path d="M 317 258.171875 L 319 258.171875 L 319 469.851562 L 317 469.851562 Z M 317 258.171875 "/>
</clipPath>
<clipPath id="clip78">
  <path d="M 351 258.171875 L 353 258.171875 L 353 469.851562 L 351 469.851562 Z M 351 258.171875 "/>
</clipPath>
<clipPath id="clip79">
  <path d="M 385 258.171875 L 387 258.171875 L 387 469.851562 L 385 469.851562 Z M 385 258.171875 "/>
</clipPath>
<clipPath id="clip80">
  <path d="M 419 258.171875 L 421 258.171875 L 421 469.851562 L 419 469.851562 Z M 419 258.171875 "/>
</clipPath>
<clipPath id="clip81">
  <path d="M 35.152344 240.402344 L 229.136719 240.402344 L 229.136719 258.167969 L 35.152344 258.167969 Z M 35.152344 240.402344 "/>
</clipPath>
<clipPath id="clip82">
  <path d="M 234.617188 240.402344 L 428.601562 240.402344 L 428.601562 258.167969 L 234.617188 258.167969 Z M 234.617188 240.402344 "/>
</clipPath>
<clipPath id="clip83">
  <path d="M 35.152344 5.480469 L 229.136719 5.480469 L 229.136719 23.246094 L 35.152344 23.246094 Z M 35.152344 5.480469 "/>
</clipPath>
<clipPath id="clip84">
  <path d="M 234.617188 5.480469 L 428.601562 5.480469 L 428.601562 23.246094 L 234.617188 23.246094 Z M 234.617188 5.480469 "/>
</clipPath>
</defs>
<g id="surface36">
<rect x="0" y="0" width="504" height="504" style="fill:rgb(100%,100%,100%);fill-opacity:1;stroke:none;"/>
<rect x="0" y="0" width="504" height="504" style="fill:rgb(100%,100%,100%);fill-opacity:1;stroke:none;"/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 0 504 L 504 504 L 504 0 L 0 0 Z M 0 504 "/>
<g clip-path="url(#clip1)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(92.156863%,92.156863%,92.156863%);fill-opacity:1;" d="M 35.152344 234.925781 L 229.136719 234.925781 L 229.136719 23.246094 L 35.152344 23.246094 Z M 35.152344 234.925781 "/>
</g>
<g clip-path="url(#clip2)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 204.210938 L 229.136719 204.210938 "/>
</g>
<g clip-path="url(#clip3)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 150.910156 L 229.136719 150.910156 "/>
</g>
<g clip-path="url(#clip4)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 97.613281 L 229.136719 97.613281 "/>
</g>
<g clip-path="url(#clip5)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 44.3125 L 229.136719 44.3125 "/>
</g>
<g clip-path="url(#clip6)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 67.707031 234.925781 L 67.707031 23.246094 "/>
</g>
<g clip-path="url(#clip7)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 101.621094 234.925781 L 101.621094 23.246094 "/>
</g>
<g clip-path="url(#clip8)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 135.535156 234.925781 L 135.535156 23.246094 "/>
</g>
<g clip-path="url(#clip9)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 169.449219 234.925781 L 169.449219 23.246094 "/>
</g>
<g clip-path="url(#clip10)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 203.363281 234.925781 L 203.363281 23.246094 "/>
</g>
<g clip-path="url(#clip11)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 230.859375 L 229.136719 230.859375 "/>
</g>
<g clip-path="url(#clip12)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 177.558594 L 229.136719 177.558594 "/>
</g>
<g clip-path="url(#clip13)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 124.261719 L 229.136719 124.261719 "/>
</g>
<g clip-path="url(#clip14)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 70.960938 L 229.136719 70.960938 "/>
</g>
<g clip-path="url(#clip15)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 50.75 234.925781 L 50.75 23.246094 "/>
</g>
<g clip-path="url(#clip16)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 84.664062 234.925781 L 84.664062 23.246094 "/>
</g>
<g clip-path="url(#clip17)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 118.578125 234.925781 L 118.578125 23.246094 "/>
</g>
<g clip-path="url(#clip18)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 152.492188 234.925781 L 152.492188 23.246094 "/>
</g>
<g clip-path="url(#clip19)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 186.40625 234.925781 L 186.40625 23.246094 "/>
</g>
<g clip-path="url(#clip20)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 220.320312 234.925781 L 220.320312 23.246094 "/>
</g>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 129.589844 C 106.96875 132.199219 103.058594 132.199219 103.058594 129.589844 C 103.058594 126.984375 106.96875 126.984375 106.96875 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 118.933594 C 120.53125 121.539062 116.625 121.539062 116.625 118.933594 C 116.625 116.324219 120.53125 116.324219 120.53125 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 124.261719 C 127.316406 126.867188 123.40625 126.867188 123.40625 124.261719 C 123.40625 121.65625 127.316406 121.65625 127.316406 124.261719 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 145.582031 C 106.96875 148.1875 103.058594 148.1875 103.058594 145.582031 C 103.058594 142.976562 106.96875 142.976562 106.96875 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 150.910156 C 93.402344 153.515625 89.492188 153.515625 89.492188 150.910156 C 89.492188 148.304688 93.402344 148.304688 93.402344 150.910156 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 134.921875 C 120.53125 137.527344 116.625 137.527344 116.625 134.921875 C 116.625 132.316406 120.53125 132.316406 120.53125 134.921875 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 140.25 C 113.75 142.859375 109.839844 142.859375 109.839844 140.25 C 109.839844 137.644531 113.75 137.644531 113.75 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 140.25 C 113.75 142.859375 109.839844 142.859375 109.839844 140.25 C 109.839844 137.644531 113.75 137.644531 113.75 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 134.097656 124.261719 C 134.097656 126.867188 130.1875 126.867188 130.1875 124.261719 C 130.1875 121.65625 134.097656 121.65625 134.097656 124.261719 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 156.242188 C 106.96875 158.847656 103.058594 158.847656 103.058594 156.242188 C 103.058594 153.632812 106.96875 153.632812 106.96875 156.242188 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 174.796875 108.273438 C 174.796875 110.878906 170.886719 110.878906 170.886719 108.273438 C 170.886719 105.664062 174.796875 105.664062 174.796875 108.273438 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 147.664062 113.601562 C 147.664062 116.207031 143.753906 116.207031 143.753906 113.601562 C 143.753906 110.996094 147.664062 110.996094 147.664062 113.601562 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 154.445312 113.601562 C 154.445312 116.207031 150.539062 116.207031 150.539062 113.601562 C 150.539062 110.996094 154.445312 110.996094 154.445312 113.601562 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 140.882812 129.589844 C 140.882812 132.199219 136.972656 132.199219 136.972656 129.589844 C 136.972656 126.984375 140.882812 126.984375 140.882812 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 147.664062 113.601562 C 147.664062 116.207031 143.753906 116.207031 143.753906 113.601562 C 143.753906 110.996094 147.664062 110.996094 147.664062 113.601562 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 161.230469 102.941406 C 161.230469 105.546875 157.320312 105.546875 157.320312 102.941406 C 157.320312 100.335938 161.230469 100.335938 161.230469 102.941406 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 154.445312 92.28125 C 154.445312 94.890625 150.539062 94.890625 150.539062 92.28125 C 150.539062 89.675781 154.445312 89.675781 154.445312 92.28125 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 147.664062 92.28125 C 147.664062 94.890625 143.753906 94.890625 143.753906 92.28125 C 143.753906 89.675781 147.664062 89.675781 147.664062 92.28125 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 145.582031 C 106.96875 148.1875 103.058594 148.1875 103.058594 145.582031 C 103.058594 142.976562 106.96875 142.976562 106.96875 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 140.25 C 106.96875 142.859375 103.058594 142.859375 103.058594 140.25 C 103.058594 137.644531 106.96875 137.644531 106.96875 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 124.261719 C 127.316406 126.867188 123.40625 126.867188 123.40625 124.261719 C 123.40625 121.65625 127.316406 121.65625 127.316406 124.261719 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 118.933594 C 127.316406 121.539062 123.40625 121.539062 123.40625 118.933594 C 123.40625 116.324219 127.316406 116.324219 127.316406 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 129.589844 C 113.75 132.199219 109.839844 132.199219 109.839844 129.589844 C 109.839844 126.984375 113.75 126.984375 113.75 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 134.921875 C 120.53125 137.527344 116.625 137.527344 116.625 134.921875 C 116.625 132.316406 120.53125 132.316406 120.53125 134.921875 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 140.25 C 120.53125 142.859375 116.625 142.859375 116.625 140.25 C 116.625 137.644531 120.53125 137.644531 120.53125 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 140.25 C 113.75 142.859375 109.839844 142.859375 109.839844 140.25 C 109.839844 137.644531 113.75 137.644531 113.75 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 140.882812 118.933594 C 140.882812 121.539062 136.972656 121.539062 136.972656 118.933594 C 136.972656 116.324219 140.882812 116.324219 140.882812 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 140.882812 113.601562 C 140.882812 116.207031 136.972656 116.207031 136.972656 113.601562 C 136.972656 110.996094 140.882812 110.996094 140.882812 113.601562 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 150.910156 C 106.96875 153.515625 103.058594 153.515625 103.058594 150.910156 C 103.058594 148.304688 106.96875 148.304688 106.96875 150.910156 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 156.242188 C 106.96875 158.847656 103.058594 158.847656 103.058594 156.242188 C 103.058594 153.632812 106.96875 153.632812 106.96875 156.242188 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 140.25 C 120.53125 142.859375 116.625 142.859375 116.625 140.25 C 116.625 137.644531 120.53125 137.644531 120.53125 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 150.910156 C 113.75 153.515625 109.839844 153.515625 109.839844 150.910156 C 109.839844 148.304688 113.75 148.304688 113.75 150.910156 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 145.582031 C 120.53125 148.1875 116.625 148.1875 116.625 145.582031 C 116.625 142.976562 120.53125 142.976562 120.53125 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 161.570312 C 106.96875 164.175781 103.058594 164.175781 103.058594 161.570312 C 103.058594 158.964844 106.96875 158.964844 106.96875 161.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 145.582031 C 127.316406 148.1875 123.40625 148.1875 123.40625 145.582031 C 123.40625 142.976562 127.316406 142.976562 127.316406 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 150.910156 C 120.53125 153.515625 116.625 153.515625 116.625 150.910156 C 116.625 148.304688 120.53125 148.304688 120.53125 150.910156 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 140.25 C 120.53125 142.859375 116.625 142.859375 116.625 140.25 C 116.625 137.644531 120.53125 137.644531 120.53125 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 150.910156 C 113.75 153.515625 109.839844 153.515625 109.839844 150.910156 C 109.839844 148.304688 113.75 148.304688 113.75 150.910156 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 140.25 C 120.53125 142.859375 116.625 142.859375 116.625 140.25 C 116.625 137.644531 120.53125 137.644531 120.53125 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 177.558594 C 86.621094 180.167969 82.710938 180.167969 82.710938 177.558594 C 82.710938 174.953125 86.621094 174.953125 86.621094 177.558594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 177.558594 C 93.402344 180.167969 89.492188 180.167969 89.492188 177.558594 C 89.492188 174.953125 93.402344 174.953125 93.402344 177.558594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 140.25 C 127.316406 142.859375 123.40625 142.859375 123.40625 140.25 C 123.40625 137.644531 127.316406 137.644531 127.316406 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 118.933594 C 127.316406 121.539062 123.40625 121.539062 123.40625 118.933594 C 123.40625 116.324219 127.316406 116.324219 127.316406 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 118.933594 C 127.316406 121.539062 123.40625 121.539062 123.40625 118.933594 C 123.40625 116.324219 127.316406 116.324219 127.316406 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 140.25 C 127.316406 142.859375 123.40625 142.859375 123.40625 140.25 C 123.40625 137.644531 127.316406 137.644531 127.316406 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 118.933594 C 127.316406 121.539062 123.40625 121.539062 123.40625 118.933594 C 123.40625 116.324219 127.316406 116.324219 127.316406 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 134.097656 118.933594 C 134.097656 121.539062 130.1875 121.539062 130.1875 118.933594 C 130.1875 116.324219 134.097656 116.324219 134.097656 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 147.664062 124.261719 C 147.664062 126.867188 143.753906 126.867188 143.753906 124.261719 C 143.753906 121.65625 147.664062 121.65625 147.664062 124.261719 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 147.664062 108.273438 C 147.664062 110.878906 143.753906 110.878906 143.753906 108.273438 C 143.753906 105.664062 147.664062 105.664062 147.664062 108.273438 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 161.230469 97.613281 C 161.230469 100.21875 157.320312 100.21875 157.320312 97.613281 C 157.320312 95.003906 161.230469 95.003906 161.230469 97.613281 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 174.796875 86.953125 C 174.796875 89.558594 170.886719 89.558594 170.886719 86.953125 C 170.886719 84.347656 174.796875 84.347656 174.796875 86.953125 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 161.230469 97.613281 C 161.230469 100.21875 157.320312 100.21875 157.320312 97.613281 C 157.320312 95.003906 161.230469 95.003906 161.230469 97.613281 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 177.558594 C 86.621094 180.167969 82.710938 180.167969 82.710938 177.558594 C 82.710938 174.953125 86.621094 174.953125 86.621094 177.558594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 177.558594 C 93.402344 180.167969 89.492188 180.167969 89.492188 177.558594 C 89.492188 174.953125 93.402344 174.953125 93.402344 177.558594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 166.902344 C 100.183594 169.507812 96.273438 169.507812 96.273438 166.902344 C 96.273438 164.292969 100.183594 164.292969 100.183594 166.902344 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 134.097656 129.589844 C 134.097656 132.199219 130.1875 132.199219 130.1875 129.589844 C 130.1875 126.984375 134.097656 126.984375 134.097656 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 208.707031 49.644531 C 208.707031 52.25 204.800781 52.25 204.800781 49.644531 C 204.800781 47.035156 208.707031 47.035156 208.707031 49.644531 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 134.097656 129.589844 C 134.097656 132.199219 130.1875 132.199219 130.1875 129.589844 C 130.1875 126.984375 134.097656 126.984375 134.097656 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 222.273438 49.644531 C 222.273438 52.25 218.363281 52.25 218.363281 49.644531 C 218.363281 47.035156 222.273438 47.035156 222.273438 49.644531 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 181.578125 65.632812 C 181.578125 68.238281 177.667969 68.238281 177.667969 65.632812 C 177.667969 63.027344 181.578125 63.027344 181.578125 65.632812 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 129.589844 C 106.96875 132.199219 103.058594 132.199219 103.058594 129.589844 C 103.058594 126.984375 106.96875 126.984375 106.96875 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 134.921875 C 113.75 137.527344 109.839844 137.527344 109.839844 134.921875 C 109.839844 132.316406 113.75 132.316406 113.75 134.921875 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 84.664062 165.398438 L 86.382812 163.894531 L 88.097656 162.386719 L 91.535156 159.371094 L 93.25 157.859375 L 94.96875 156.347656 L 96.683594 154.832031 L 98.402344 153.3125 L 100.117188 151.792969 L 103.554688 148.738281 L 105.269531 147.207031 L 106.988281 145.671875 L 108.703125 144.128906 L 110.421875 142.582031 L 112.140625 141.027344 L 113.855469 139.46875 L 115.574219 137.902344 L 117.289062 136.328125 L 119.007812 134.742188 L 120.726562 133.148438 L 122.441406 131.546875 L 124.160156 129.933594 L 125.875 128.3125 L 127.59375 126.679688 L 129.308594 125.035156 L 131.027344 123.382812 L 132.746094 121.71875 L 134.460938 120.050781 L 136.179688 118.371094 L 137.894531 116.683594 L 139.613281 114.988281 L 141.332031 113.289062 L 143.046875 111.582031 L 144.765625 109.871094 L 146.480469 108.152344 L 148.199219 106.433594 L 149.917969 104.710938 L 151.632812 102.980469 L 153.351562 101.25 L 155.066406 99.515625 L 156.785156 97.78125 L 158.503906 96.042969 L 160.21875 94.300781 L 161.9375 92.558594 L 163.652344 90.816406 L 165.371094 89.070312 L 167.085938 87.324219 L 168.804688 85.578125 L 170.523438 83.828125 L 172.238281 82.078125 L 173.957031 80.328125 L 175.671875 78.574219 L 177.390625 76.824219 L 179.109375 75.070312 L 180.824219 73.316406 L 182.542969 71.5625 L 184.257812 69.804688 L 185.976562 68.050781 L 187.695312 66.292969 L 189.410156 64.539062 L 191.128906 62.78125 L 192.84375 61.023438 L 196.28125 57.507812 L 197.996094 55.75 L 199.714844 53.992188 L 201.429688 52.230469 L 203.148438 50.472656 L 204.863281 48.710938 L 206.582031 46.953125 L 208.300781 45.191406 L 210.015625 43.433594 L 211.734375 41.671875 L 213.449219 39.914062 L 216.886719 36.390625 L 218.601562 34.628906 L 220.320312 32.867188 L 220.320312 48.824219 L 218.601562 50.3125 L 216.886719 51.800781 L 215.167969 53.289062 L 213.449219 54.78125 L 211.734375 56.269531 L 210.015625 57.761719 L 208.300781 59.25 L 204.863281 62.234375 L 203.148438 63.722656 L 201.429688 65.214844 L 199.714844 66.707031 L 197.996094 68.199219 L 196.28125 69.691406 L 192.84375 72.675781 L 191.128906 74.167969 L 189.410156 75.664062 L 187.695312 77.15625 L 184.257812 80.148438 L 182.542969 81.640625 L 180.824219 83.136719 L 179.109375 84.636719 L 175.671875 87.628906 L 173.957031 89.128906 L 172.238281 90.628906 L 170.523438 92.128906 L 168.804688 93.632812 L 167.085938 95.132812 L 165.371094 96.636719 L 163.652344 98.144531 L 161.9375 99.648438 L 160.21875 101.160156 L 158.503906 102.667969 L 156.785156 104.179688 L 155.066406 105.695312 L 153.351562 107.210938 L 151.632812 108.730469 L 149.917969 110.253906 L 148.199219 111.78125 L 146.480469 113.3125 L 144.765625 114.84375 L 143.046875 116.382812 L 141.332031 117.929688 L 139.613281 119.476562 L 137.894531 121.035156 L 136.179688 122.597656 L 134.460938 124.167969 L 132.746094 125.75 L 131.027344 127.335938 L 129.308594 128.933594 L 127.59375 130.542969 L 125.875 132.160156 L 124.160156 133.789062 L 122.441406 135.425781 L 120.726562 137.074219 L 119.007812 138.730469 L 117.289062 140.398438 L 115.574219 142.074219 L 113.855469 143.757812 L 112.140625 145.449219 L 110.421875 147.144531 L 108.703125 148.847656 L 106.988281 150.558594 L 105.269531 152.273438 L 103.554688 153.988281 L 101.835938 155.710938 L 100.117188 157.4375 L 98.402344 159.167969 L 96.683594 160.902344 L 94.96875 162.636719 L 93.25 164.371094 L 91.535156 166.113281 L 89.816406 167.851562 L 88.097656 169.597656 L 86.382812 171.339844 L 84.664062 173.085938 Z M 84.664062 165.398438 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 84.664062 169.242188 L 86.382812 167.617188 L 88.097656 165.992188 L 91.535156 162.742188 L 93.25 161.117188 L 94.96875 159.492188 L 96.683594 157.867188 L 98.402344 156.242188 L 100.117188 154.613281 L 103.554688 151.363281 L 105.269531 149.738281 L 106.988281 148.113281 L 108.703125 146.488281 L 112.140625 143.238281 L 113.855469 141.613281 L 115.574219 139.988281 L 117.289062 138.363281 L 120.726562 135.113281 L 122.441406 133.488281 L 124.160156 131.859375 L 125.875 130.234375 L 127.59375 128.609375 L 129.308594 126.984375 L 132.746094 123.734375 L 134.460938 122.109375 L 136.179688 120.484375 L 137.894531 118.859375 L 141.332031 115.609375 L 143.046875 113.984375 L 144.765625 112.359375 L 146.480469 110.734375 L 148.199219 109.105469 L 149.917969 107.480469 L 151.632812 105.855469 L 153.351562 104.230469 L 155.066406 102.605469 L 158.503906 99.355469 L 160.21875 97.730469 L 161.9375 96.105469 L 163.652344 94.480469 L 165.371094 92.855469 L 167.085938 91.230469 L 170.523438 87.980469 L 172.238281 86.351562 L 173.957031 84.726562 L 175.671875 83.101562 L 179.109375 79.851562 L 180.824219 78.226562 L 182.542969 76.601562 L 184.257812 74.976562 L 187.695312 71.726562 L 189.410156 70.101562 L 191.128906 68.476562 L 192.84375 66.851562 L 194.5625 65.222656 L 196.28125 63.597656 L 197.996094 61.972656 L 199.714844 60.347656 L 201.429688 58.722656 L 203.148438 57.097656 L 204.863281 55.472656 L 208.300781 52.222656 L 210.015625 50.597656 L 211.734375 48.972656 L 213.449219 47.347656 L 216.886719 44.097656 L 218.601562 42.46875 L 220.320312 40.84375 "/>
<g clip-path="url(#clip21)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(92.156863%,92.156863%,92.156863%);fill-opacity:1;" d="M 35.152344 469.847656 L 229.136719 469.847656 L 229.136719 258.167969 L 35.152344 258.167969 Z M 35.152344 469.847656 "/>
</g>
<g clip-path="url(#clip22)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 439.132812 L 229.136719 439.132812 "/>
</g>
<g clip-path="url(#clip23)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 385.835938 L 229.136719 385.835938 "/>
</g>
<g clip-path="url(#clip24)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 332.535156 L 229.136719 332.535156 "/>
</g>
<g clip-path="url(#clip25)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 279.238281 L 229.136719 279.238281 "/>
</g>
<g clip-path="url(#clip26)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 67.707031 469.847656 L 67.707031 258.171875 "/>
</g>
<g clip-path="url(#clip27)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 101.621094 469.847656 L 101.621094 258.171875 "/>
</g>
<g clip-path="url(#clip28)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 135.535156 469.847656 L 135.535156 258.171875 "/>
</g>
<g clip-path="url(#clip29)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 169.449219 469.847656 L 169.449219 258.171875 "/>
</g>
<g clip-path="url(#clip30)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 203.363281 469.847656 L 203.363281 258.171875 "/>
</g>
<g clip-path="url(#clip31)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 465.785156 L 229.136719 465.785156 "/>
</g>
<g clip-path="url(#clip32)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 412.484375 L 229.136719 412.484375 "/>
</g>
<g clip-path="url(#clip33)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 359.1875 L 229.136719 359.1875 "/>
</g>
<g clip-path="url(#clip34)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 305.886719 L 229.136719 305.886719 "/>
</g>
<g clip-path="url(#clip35)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 50.75 469.847656 L 50.75 258.171875 "/>
</g>
<g clip-path="url(#clip36)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 84.664062 469.847656 L 84.664062 258.171875 "/>
</g>
<g clip-path="url(#clip37)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 118.578125 469.847656 L 118.578125 258.171875 "/>
</g>
<g clip-path="url(#clip38)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 152.492188 469.847656 L 152.492188 258.171875 "/>
</g>
<g clip-path="url(#clip39)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 186.40625 469.847656 L 186.40625 258.171875 "/>
</g>
<g clip-path="url(#clip40)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 220.320312 469.847656 L 220.320312 258.171875 "/>
</g>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 380.503906 C 93.402344 383.113281 89.492188 383.113281 89.492188 380.503906 C 89.492188 377.898438 93.402344 377.898438 93.402344 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 375.175781 C 106.96875 377.78125 103.058594 377.78125 103.058594 375.175781 C 103.058594 372.570312 106.96875 372.570312 106.96875 375.175781 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 385.835938 C 86.621094 388.441406 82.710938 388.441406 82.710938 385.835938 C 82.710938 383.230469 86.621094 383.230469 86.621094 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 385.835938 C 100.183594 388.441406 96.273438 388.441406 96.273438 385.835938 C 96.273438 383.230469 100.183594 383.230469 100.183594 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 385.835938 C 100.183594 388.441406 96.273438 388.441406 96.273438 385.835938 C 96.273438 383.230469 100.183594 383.230469 100.183594 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 385.835938 C 86.621094 388.441406 82.710938 388.441406 82.710938 385.835938 C 82.710938 383.230469 86.621094 383.230469 86.621094 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 391.164062 C 86.621094 393.773438 82.710938 393.773438 82.710938 391.164062 C 82.710938 388.558594 86.621094 388.558594 86.621094 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 385.835938 C 100.183594 388.441406 96.273438 388.441406 96.273438 385.835938 C 96.273438 383.230469 100.183594 383.230469 100.183594 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 364.515625 C 106.96875 367.121094 103.058594 367.121094 103.058594 364.515625 C 103.058594 361.910156 106.96875 361.910156 106.96875 364.515625 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 380.503906 C 100.183594 383.113281 96.273438 383.113281 96.273438 380.503906 C 96.273438 377.898438 100.183594 377.898438 100.183594 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 401.824219 C 93.402344 404.429688 89.492188 404.429688 89.492188 401.824219 C 89.492188 399.21875 93.402344 399.21875 93.402344 401.824219 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 401.824219 C 93.402344 404.429688 89.492188 404.429688 89.492188 401.824219 C 89.492188 399.21875 93.402344 399.21875 93.402344 401.824219 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 59.488281 428.472656 C 59.488281 431.082031 55.578125 431.082031 55.578125 428.472656 C 55.578125 425.867188 59.488281 425.867188 59.488281 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 401.824219 C 86.621094 404.429688 82.710938 404.429688 82.710938 401.824219 C 82.710938 399.21875 86.621094 399.21875 86.621094 401.824219 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 407.15625 C 86.621094 409.761719 82.710938 409.761719 82.710938 407.15625 C 82.710938 404.546875 86.621094 404.546875 86.621094 407.15625 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 396.496094 C 93.402344 399.101562 89.492188 399.101562 89.492188 396.496094 C 89.492188 393.890625 93.402344 393.890625 93.402344 396.496094 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 396.496094 C 93.402344 399.101562 89.492188 399.101562 89.492188 396.496094 C 89.492188 393.890625 93.402344 393.890625 93.402344 396.496094 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 417.816406 C 86.621094 420.421875 82.710938 420.421875 82.710938 417.816406 C 82.710938 415.207031 86.621094 415.207031 86.621094 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 423.144531 C 79.835938 425.75 75.925781 425.75 75.925781 423.144531 C 75.925781 420.539062 79.835938 420.539062 79.835938 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 73.054688 428.472656 C 73.054688 431.082031 69.144531 431.082031 69.144531 428.472656 C 69.144531 425.867188 73.054688 425.867188 73.054688 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 73.054688 428.472656 C 73.054688 431.082031 69.144531 431.082031 69.144531 428.472656 C 69.144531 425.867188 73.054688 425.867188 73.054688 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 417.816406 C 86.621094 420.421875 82.710938 420.421875 82.710938 417.816406 C 82.710938 415.207031 86.621094 415.207031 86.621094 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 73.054688 417.816406 C 73.054688 420.421875 69.144531 420.421875 69.144531 417.816406 C 69.144531 415.207031 73.054688 415.207031 73.054688 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 385.835938 C 106.96875 388.441406 103.058594 388.441406 103.058594 385.835938 C 103.058594 383.230469 106.96875 383.230469 106.96875 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 380.503906 C 100.183594 383.113281 96.273438 383.113281 96.273438 380.503906 C 96.273438 377.898438 100.183594 377.898438 100.183594 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 391.164062 C 93.402344 393.773438 89.492188 393.773438 89.492188 391.164062 C 89.492188 388.558594 93.402344 388.558594 93.402344 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 369.847656 C 113.75 372.453125 109.839844 372.453125 109.839844 369.847656 C 109.839844 367.238281 113.75 367.238281 113.75 369.847656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 391.164062 C 93.402344 393.773438 89.492188 393.773438 89.492188 391.164062 C 89.492188 388.558594 93.402344 388.558594 93.402344 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 401.824219 C 100.183594 404.429688 96.273438 404.429688 96.273438 401.824219 C 96.273438 399.21875 100.183594 399.21875 100.183594 401.824219 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 417.816406 C 86.621094 420.421875 82.710938 420.421875 82.710938 417.816406 C 82.710938 415.207031 86.621094 415.207031 86.621094 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 412.484375 C 86.621094 415.089844 82.710938 415.089844 82.710938 412.484375 C 82.710938 409.878906 86.621094 409.878906 86.621094 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 73.054688 417.816406 C 73.054688 420.421875 69.144531 420.421875 69.144531 417.816406 C 69.144531 415.207031 73.054688 415.207031 73.054688 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 375.175781 C 113.75 377.78125 109.839844 377.78125 109.839844 375.175781 C 109.839844 372.570312 113.75 372.570312 113.75 375.175781 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 380.503906 C 113.75 383.113281 109.839844 383.113281 109.839844 380.503906 C 109.839844 377.898438 113.75 377.898438 113.75 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 385.835938 C 113.75 388.441406 109.839844 388.441406 109.839844 385.835938 C 109.839844 383.230469 113.75 383.230469 113.75 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 385.835938 C 113.75 388.441406 109.839844 388.441406 109.839844 385.835938 C 109.839844 383.230469 113.75 383.230469 113.75 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 428.472656 C 86.621094 431.082031 82.710938 431.082031 82.710938 428.472656 C 82.710938 425.867188 86.621094 425.867188 86.621094 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 412.484375 C 79.835938 415.089844 75.925781 415.089844 75.925781 412.484375 C 75.925781 409.878906 79.835938 409.878906 79.835938 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 380.503906 C 93.402344 383.113281 89.492188 383.113281 89.492188 380.503906 C 89.492188 377.898438 93.402344 377.898438 93.402344 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 375.175781 C 100.183594 377.78125 96.273438 377.78125 96.273438 375.175781 C 96.273438 372.570312 100.183594 372.570312 100.183594 375.175781 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 369.847656 C 106.96875 372.453125 103.058594 372.453125 103.058594 369.847656 C 103.058594 367.238281 106.96875 367.238281 106.96875 369.847656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 417.816406 C 86.621094 420.421875 82.710938 420.421875 82.710938 417.816406 C 82.710938 415.207031 86.621094 415.207031 86.621094 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 428.472656 C 86.621094 431.082031 82.710938 431.082031 82.710938 428.472656 C 82.710938 425.867188 86.621094 425.867188 86.621094 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 412.484375 C 93.402344 415.089844 89.492188 415.089844 89.492188 412.484375 C 89.492188 409.878906 93.402344 409.878906 93.402344 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 369.847656 C 113.75 372.453125 109.839844 372.453125 109.839844 369.847656 C 109.839844 367.238281 113.75 367.238281 113.75 369.847656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 375.175781 C 106.96875 377.78125 103.058594 377.78125 103.058594 375.175781 C 103.058594 372.570312 106.96875 372.570312 106.96875 375.175781 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 428.472656 C 86.621094 431.082031 82.710938 431.082031 82.710938 428.472656 C 82.710938 425.867188 86.621094 425.867188 86.621094 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 417.816406 C 86.621094 420.421875 82.710938 420.421875 82.710938 417.816406 C 82.710938 415.207031 86.621094 415.207031 86.621094 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 423.144531 C 86.621094 425.75 82.710938 425.75 82.710938 423.144531 C 82.710938 420.539062 86.621094 420.539062 86.621094 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 412.484375 C 93.402344 415.089844 89.492188 415.089844 89.492188 412.484375 C 89.492188 409.878906 93.402344 409.878906 93.402344 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 396.496094 C 93.402344 399.101562 89.492188 399.101562 89.492188 396.496094 C 89.492188 393.890625 93.402344 393.890625 93.402344 396.496094 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 380.503906 C 93.402344 383.113281 89.492188 383.113281 89.492188 380.503906 C 89.492188 377.898438 93.402344 377.898438 93.402344 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 380.503906 C 100.183594 383.113281 96.273438 383.113281 96.273438 380.503906 C 96.273438 377.898438 100.183594 377.898438 100.183594 380.503906 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 57.535156 440.144531 L 58.21875 439.300781 L 60.28125 436.769531 L 60.96875 435.921875 L 61.65625 435.078125 L 62.339844 434.234375 L 63.027344 433.386719 L 63.714844 432.542969 L 65.089844 430.847656 L 65.777344 430.003906 L 66.464844 429.15625 L 67.148438 428.308594 L 67.835938 427.457031 L 68.523438 426.609375 L 69.210938 425.757812 L 69.898438 424.910156 L 70.585938 424.058594 L 71.269531 423.203125 L 72.644531 421.5 L 74.019531 419.789062 L 74.707031 418.929688 L 75.390625 418.070312 L 76.765625 416.351562 L 78.140625 414.625 L 78.828125 413.757812 L 79.511719 412.890625 L 80.886719 411.148438 L 81.574219 410.273438 L 82.261719 409.394531 L 82.949219 408.511719 L 83.632812 407.628906 L 85.007812 405.847656 L 86.382812 404.050781 L 87.070312 403.140625 L 87.753906 402.230469 L 88.441406 401.3125 L 89.128906 400.390625 L 89.816406 399.460938 L 90.503906 398.523438 L 91.191406 397.578125 L 91.875 396.628906 L 92.5625 395.667969 L 93.25 394.703125 L 93.9375 393.730469 L 94.625 392.75 L 95.3125 391.761719 L 95.996094 390.765625 L 96.683594 389.765625 L 97.371094 388.757812 L 98.058594 387.742188 L 98.746094 386.722656 L 99.433594 385.699219 L 100.117188 384.667969 L 100.804688 383.632812 L 101.492188 382.59375 L 102.179688 381.550781 L 102.867188 380.503906 L 103.554688 379.453125 L 104.238281 378.402344 L 105.613281 376.285156 L 106.988281 374.160156 L 107.675781 373.09375 L 108.359375 372.027344 L 109.046875 370.960938 L 109.734375 369.886719 L 110.421875 368.816406 L 111.796875 366.667969 L 111.796875 375.09375 L 111.109375 375.949219 L 109.046875 378.527344 L 108.359375 379.390625 L 107.675781 380.253906 L 106.300781 381.988281 L 105.613281 382.859375 L 104.238281 384.609375 L 103.554688 385.484375 L 102.867188 386.367188 L 101.492188 388.140625 L 100.804688 389.03125 L 100.117188 389.929688 L 99.433594 390.832031 L 98.746094 391.738281 L 98.058594 392.648438 L 97.371094 393.566406 L 96.683594 394.488281 L 95.996094 395.421875 L 95.3125 396.355469 L 94.625 397.300781 L 93.9375 398.25 L 93.25 399.210938 L 92.5625 400.175781 L 91.875 401.148438 L 91.191406 402.128906 L 90.503906 403.117188 L 89.816406 404.113281 L 89.128906 405.113281 L 88.441406 406.121094 L 87.753906 407.136719 L 87.070312 408.15625 L 86.382812 409.179688 L 85.695312 410.210938 L 85.007812 411.246094 L 84.320312 412.285156 L 83.632812 413.328125 L 82.949219 414.375 L 82.261719 415.425781 L 80.886719 417.535156 L 80.199219 418.59375 L 79.511719 419.65625 L 78.828125 420.71875 L 77.453125 422.851562 L 75.390625 426.0625 L 74.707031 427.136719 L 73.332031 429.285156 L 71.269531 432.519531 L 70.585938 433.597656 L 67.148438 439.007812 L 66.464844 440.089844 L 65.777344 441.175781 L 65.089844 442.257812 L 62.339844 446.601562 L 61.65625 447.6875 L 60.96875 448.773438 L 60.28125 449.863281 L 59.59375 450.949219 L 58.90625 452.039062 L 58.21875 453.125 L 57.535156 454.214844 Z M 57.535156 440.144531 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 57.535156 447.179688 L 58.21875 446.210938 L 60.28125 443.316406 L 60.96875 442.347656 L 61.65625 441.382812 L 62.339844 440.417969 L 63.027344 439.453125 L 63.714844 438.484375 L 65.777344 435.589844 L 66.464844 434.621094 L 67.148438 433.65625 L 69.210938 430.761719 L 69.898438 429.792969 L 70.585938 428.828125 L 71.269531 427.863281 L 71.957031 426.898438 L 72.644531 425.929688 L 74.707031 423.035156 L 75.390625 422.066406 L 77.453125 419.171875 L 78.140625 418.203125 L 78.828125 417.238281 L 79.511719 416.273438 L 80.199219 415.308594 L 80.886719 414.339844 L 82.949219 411.445312 L 83.632812 410.476562 L 85.695312 407.582031 L 86.382812 406.613281 L 87.070312 405.648438 L 87.753906 404.683594 L 88.441406 403.71875 L 89.128906 402.75 L 91.191406 399.855469 L 91.875 398.886719 L 94.625 395.027344 L 95.3125 394.058594 L 95.996094 393.09375 L 97.371094 391.164062 L 98.058594 390.195312 L 99.433594 388.265625 L 100.117188 387.300781 L 100.804688 386.332031 L 102.867188 383.4375 L 103.554688 382.46875 L 104.238281 381.503906 L 105.613281 379.574219 L 106.300781 378.605469 L 107.675781 376.675781 L 108.359375 375.710938 L 109.046875 374.742188 L 111.109375 371.847656 L 111.796875 370.878906 "/>
<g clip-path="url(#clip41)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(92.156863%,92.156863%,92.156863%);fill-opacity:1;" d="M 234.617188 234.925781 L 428.601562 234.925781 L 428.601562 23.246094 L 234.617188 23.246094 Z M 234.617188 234.925781 "/>
</g>
<g clip-path="url(#clip42)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 204.210938 L 428.601562 204.210938 "/>
</g>
<g clip-path="url(#clip43)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 150.910156 L 428.601562 150.910156 "/>
</g>
<g clip-path="url(#clip44)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 97.613281 L 428.601562 97.613281 "/>
</g>
<g clip-path="url(#clip45)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 44.3125 L 428.601562 44.3125 "/>
</g>
<g clip-path="url(#clip46)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 267.171875 234.925781 L 267.171875 23.246094 "/>
</g>
<g clip-path="url(#clip47)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 301.085938 234.925781 L 301.085938 23.246094 "/>
</g>
<g clip-path="url(#clip48)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 335 234.925781 L 335 23.246094 "/>
</g>
<g clip-path="url(#clip49)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 368.914062 234.925781 L 368.914062 23.246094 "/>
</g>
<g clip-path="url(#clip50)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 402.828125 234.925781 L 402.828125 23.246094 "/>
</g>
<g clip-path="url(#clip51)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 230.859375 L 428.601562 230.859375 "/>
</g>
<g clip-path="url(#clip52)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 177.558594 L 428.601562 177.558594 "/>
</g>
<g clip-path="url(#clip53)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 124.261719 L 428.601562 124.261719 "/>
</g>
<g clip-path="url(#clip54)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 70.960938 L 428.601562 70.960938 "/>
</g>
<g clip-path="url(#clip55)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 250.214844 234.925781 L 250.214844 23.246094 "/>
</g>
<g clip-path="url(#clip56)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 284.128906 234.925781 L 284.128906 23.246094 "/>
</g>
<g clip-path="url(#clip57)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 318.042969 234.925781 L 318.042969 23.246094 "/>
</g>
<g clip-path="url(#clip58)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 351.957031 234.925781 L 351.957031 23.246094 "/>
</g>
<g clip-path="url(#clip59)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 385.871094 234.925781 L 385.871094 23.246094 "/>
</g>
<g clip-path="url(#clip60)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 419.785156 234.925781 L 419.785156 23.246094 "/>
</g>
<path style="fill-rule:nonzero;fill:rgb(48.627451%,68.235294%,0%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 326.78125 129.589844 C 326.78125 132.199219 322.871094 132.199219 322.871094 129.589844 C 322.871094 126.984375 326.78125 126.984375 326.78125 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(48.627451%,68.235294%,0%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 326.78125 129.589844 C 326.78125 132.199219 322.871094 132.199219 322.871094 129.589844 C 322.871094 126.984375 326.78125 126.984375 326.78125 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(48.627451%,68.235294%,0%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 320 134.921875 C 320 137.527344 316.089844 137.527344 316.089844 134.921875 C 316.089844 132.316406 320 132.316406 320 134.921875 "/>
<path style="fill-rule:nonzero;fill:rgb(48.627451%,68.235294%,0%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 320 129.589844 C 320 132.199219 316.089844 132.199219 316.089844 129.589844 C 316.089844 126.984375 320 126.984375 320 129.589844 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 318.042969 124.148438 L 318.128906 124.214844 L 318.214844 124.285156 L 318.300781 124.347656 L 318.386719 124.414062 L 318.472656 124.476562 L 318.558594 124.535156 L 318.644531 124.597656 L 318.730469 124.652344 L 318.816406 124.710938 L 318.902344 124.761719 L 318.988281 124.816406 L 319.074219 124.867188 L 319.246094 124.960938 L 319.417969 125.046875 L 319.503906 125.085938 L 319.675781 125.15625 L 319.847656 125.21875 L 320.105469 125.289062 L 320.277344 125.320312 L 320.363281 125.332031 L 320.535156 125.347656 L 320.707031 125.347656 L 320.878906 125.332031 L 320.964844 125.320312 L 321.050781 125.304688 L 321.136719 125.285156 L 321.308594 125.238281 L 321.390625 125.207031 L 321.5625 125.136719 L 321.734375 125.050781 L 321.90625 124.949219 L 321.992188 124.894531 L 322.078125 124.835938 L 322.164062 124.773438 L 322.25 124.707031 L 322.421875 124.566406 L 322.59375 124.410156 L 322.679688 124.328125 L 322.765625 124.242188 L 322.9375 124.0625 L 323.023438 123.96875 L 323.195312 123.773438 L 323.28125 123.671875 L 323.453125 123.460938 L 323.539062 123.351562 L 323.796875 123.011719 L 323.882812 122.894531 L 324.054688 122.652344 L 324.3125 122.277344 L 324.484375 122.019531 L 324.742188 121.621094 L 324.828125 121.484375 L 324.828125 137.699219 L 324.742188 137.632812 L 324.65625 137.5625 L 324.570312 137.5 L 324.484375 137.433594 L 324.398438 137.371094 L 324.3125 137.3125 L 324.226562 137.25 L 324.140625 137.195312 L 324.054688 137.136719 L 323.96875 137.085938 L 323.882812 137.03125 L 323.796875 136.980469 L 323.625 136.886719 L 323.453125 136.800781 L 323.367188 136.761719 L 323.195312 136.691406 L 323.023438 136.628906 L 322.765625 136.558594 L 322.59375 136.527344 L 322.507812 136.515625 L 322.421875 136.507812 L 322.25 136.5 L 322.164062 136.5 L 321.992188 136.515625 L 321.90625 136.527344 L 321.820312 136.542969 L 321.734375 136.5625 L 321.5625 136.609375 L 321.476562 136.640625 L 321.390625 136.675781 L 321.308594 136.710938 L 321.136719 136.796875 L 320.964844 136.898438 L 320.878906 136.953125 L 320.792969 137.011719 L 320.707031 137.074219 L 320.621094 137.140625 L 320.449219 137.28125 L 320.277344 137.4375 L 320.191406 137.519531 L 320.105469 137.605469 L 319.933594 137.785156 L 319.847656 137.878906 L 319.675781 138.074219 L 319.589844 138.175781 L 319.417969 138.386719 L 319.332031 138.496094 L 319.074219 138.835938 L 318.988281 138.953125 L 318.816406 139.195312 L 318.558594 139.570312 L 318.386719 139.828125 L 318.214844 140.09375 L 318.128906 140.230469 L 318.042969 140.363281 Z M 318.042969 124.148438 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 318.042969 132.257812 L 318.214844 132.1875 L 318.300781 132.15625 L 318.472656 132.085938 L 318.558594 132.054688 L 318.644531 132.019531 L 318.730469 131.988281 L 318.902344 131.917969 L 318.988281 131.886719 L 319.160156 131.816406 L 319.246094 131.785156 L 319.417969 131.714844 L 319.503906 131.683594 L 319.589844 131.648438 L 319.675781 131.617188 L 319.847656 131.546875 L 319.933594 131.515625 L 320.105469 131.445312 L 320.191406 131.414062 L 320.363281 131.34375 L 320.449219 131.3125 L 320.535156 131.277344 L 320.621094 131.246094 L 320.792969 131.175781 L 320.878906 131.144531 L 321.050781 131.074219 L 321.136719 131.042969 L 321.308594 130.972656 L 321.390625 130.941406 L 321.476562 130.90625 L 321.5625 130.875 L 321.734375 130.804688 L 321.820312 130.773438 L 321.992188 130.703125 L 322.078125 130.671875 L 322.25 130.601562 L 322.335938 130.570312 L 322.421875 130.535156 L 322.507812 130.503906 L 322.679688 130.433594 L 322.765625 130.402344 L 322.9375 130.332031 L 323.023438 130.300781 L 323.195312 130.230469 L 323.28125 130.199219 L 323.367188 130.164062 L 323.453125 130.132812 L 323.625 130.0625 L 323.710938 130.03125 L 323.882812 129.960938 L 323.96875 129.929688 L 324.140625 129.859375 L 324.226562 129.828125 L 324.3125 129.792969 L 324.398438 129.761719 L 324.570312 129.691406 L 324.65625 129.660156 L 324.828125 129.589844 "/>
<g clip-path="url(#clip61)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(92.156863%,92.156863%,92.156863%);fill-opacity:1;" d="M 234.617188 469.847656 L 428.601562 469.847656 L 428.601562 258.167969 L 234.617188 258.167969 Z M 234.617188 469.847656 "/>
</g>
<g clip-path="url(#clip62)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 439.132812 L 428.601562 439.132812 "/>
</g>
<g clip-path="url(#clip63)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 385.835938 L 428.601562 385.835938 "/>
</g>
<g clip-path="url(#clip64)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 332.535156 L 428.601562 332.535156 "/>
</g>
<g clip-path="url(#clip65)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 279.238281 L 428.601562 279.238281 "/>
</g>
<g clip-path="url(#clip66)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 267.171875 469.847656 L 267.171875 258.171875 "/>
</g>
<g clip-path="url(#clip67)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 301.085938 469.847656 L 301.085938 258.171875 "/>
</g>
<g clip-path="url(#clip68)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 335 469.847656 L 335 258.171875 "/>
</g>
<g clip-path="url(#clip69)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 368.914062 469.847656 L 368.914062 258.171875 "/>
</g>
<g clip-path="url(#clip70)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 402.828125 469.847656 L 402.828125 258.171875 "/>
</g>
<g clip-path="url(#clip71)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 465.785156 L 428.601562 465.785156 "/>
</g>
<g clip-path="url(#clip72)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 412.484375 L 428.601562 412.484375 "/>
</g>
<g clip-path="url(#clip73)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 359.1875 L 428.601562 359.1875 "/>
</g>
<g clip-path="url(#clip74)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 305.886719 L 428.601562 305.886719 "/>
</g>
<g clip-path="url(#clip75)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 250.214844 469.847656 L 250.214844 258.171875 "/>
</g>
<g clip-path="url(#clip76)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 284.128906 469.847656 L 284.128906 258.171875 "/>
</g>
<g clip-path="url(#clip77)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 318.042969 469.847656 L 318.042969 258.171875 "/>
</g>
<g clip-path="url(#clip78)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 351.957031 469.847656 L 351.957031 258.171875 "/>
</g>
<g clip-path="url(#clip79)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 385.871094 469.847656 L 385.871094 258.171875 "/>
</g>
<g clip-path="url(#clip80)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 419.785156 469.847656 L 419.785156 258.171875 "/>
</g>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 292.867188 396.496094 C 292.867188 399.101562 288.957031 399.101562 288.957031 396.496094 C 288.957031 393.890625 292.867188 393.890625 292.867188 396.496094 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 412.484375 C 279.300781 415.089844 275.394531 415.089844 275.394531 412.484375 C 275.394531 409.878906 279.300781 409.878906 279.300781 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 412.484375 C 279.300781 415.089844 275.394531 415.089844 275.394531 412.484375 C 275.394531 409.878906 279.300781 409.878906 279.300781 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 428.472656 C 265.738281 431.082031 261.828125 431.082031 261.828125 428.472656 C 261.828125 425.867188 265.738281 425.867188 265.738281 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 292.867188 380.503906 C 292.867188 383.113281 288.957031 383.113281 288.957031 380.503906 C 288.957031 377.898438 292.867188 377.898438 292.867188 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 396.496094 C 286.085938 399.101562 282.175781 399.101562 282.175781 396.496094 C 282.175781 393.890625 286.085938 393.890625 286.085938 396.496094 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 292.867188 380.503906 C 292.867188 383.113281 288.957031 383.113281 288.957031 380.503906 C 288.957031 377.898438 292.867188 377.898438 292.867188 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 385.835938 C 286.085938 388.441406 282.175781 388.441406 282.175781 385.835938 C 282.175781 383.230469 286.085938 383.230469 286.085938 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 391.164062 C 286.085938 393.773438 282.175781 393.773438 282.175781 391.164062 C 282.175781 388.558594 286.085938 388.558594 286.085938 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 417.816406 C 279.300781 420.421875 275.394531 420.421875 275.394531 417.816406 C 275.394531 415.207031 279.300781 415.207031 279.300781 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 444.464844 C 258.953125 447.070312 255.042969 447.070312 255.042969 444.464844 C 255.042969 441.859375 258.953125 441.859375 258.953125 444.464844 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 428.472656 C 279.300781 431.082031 275.394531 431.082031 275.394531 428.472656 C 275.394531 425.867188 279.300781 425.867188 279.300781 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 417.816406 C 279.300781 420.421875 275.394531 420.421875 275.394531 417.816406 C 275.394531 415.207031 279.300781 415.207031 279.300781 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 417.816406 C 279.300781 420.421875 275.394531 420.421875 275.394531 417.816406 C 275.394531 415.207031 279.300781 415.207031 279.300781 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 245.390625 455.125 C 245.390625 457.730469 241.480469 457.730469 241.480469 455.125 C 241.480469 452.515625 245.390625 452.515625 245.390625 455.125 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 428.472656 C 258.953125 431.082031 255.042969 431.082031 255.042969 428.472656 C 255.042969 425.867188 258.953125 425.867188 258.953125 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 245.390625 455.125 C 245.390625 457.730469 241.480469 457.730469 241.480469 455.125 C 241.480469 452.515625 245.390625 452.515625 245.390625 455.125 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 433.804688 C 258.953125 436.410156 255.042969 436.410156 255.042969 433.804688 C 255.042969 431.199219 258.953125 431.199219 258.953125 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 423.144531 C 272.519531 425.75 268.609375 425.75 268.609375 423.144531 C 268.609375 420.539062 272.519531 420.539062 272.519531 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 433.804688 C 265.738281 436.410156 261.828125 436.410156 261.828125 433.804688 C 261.828125 431.199219 265.738281 431.199219 265.738281 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 245.390625 455.125 C 245.390625 457.730469 241.480469 457.730469 241.480469 455.125 C 241.480469 452.515625 245.390625 452.515625 245.390625 455.125 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 433.804688 C 265.738281 436.410156 261.828125 436.410156 261.828125 433.804688 C 261.828125 431.199219 265.738281 431.199219 265.738281 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 245.390625 455.125 C 245.390625 457.730469 241.480469 457.730469 241.480469 455.125 C 241.480469 452.515625 245.390625 452.515625 245.390625 455.125 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 433.804688 C 258.953125 436.410156 255.042969 436.410156 255.042969 433.804688 C 255.042969 431.199219 258.953125 431.199219 258.953125 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 428.472656 C 258.953125 431.082031 255.042969 431.082031 255.042969 428.472656 C 255.042969 425.867188 258.953125 425.867188 258.953125 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 428.472656 C 258.953125 431.082031 255.042969 431.082031 255.042969 428.472656 C 255.042969 425.867188 258.953125 425.867188 258.953125 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 423.144531 C 265.738281 425.75 261.828125 425.75 261.828125 423.144531 C 261.828125 420.539062 265.738281 420.539062 265.738281 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 417.816406 C 272.519531 420.421875 268.609375 420.421875 268.609375 417.816406 C 268.609375 415.207031 272.519531 415.207031 272.519531 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 433.804688 C 272.519531 436.410156 268.609375 436.410156 268.609375 433.804688 C 268.609375 431.199219 272.519531 431.199219 272.519531 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 433.804688 C 272.519531 436.410156 268.609375 436.410156 268.609375 433.804688 C 268.609375 431.199219 272.519531 431.199219 272.519531 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 407.15625 C 286.085938 409.761719 282.175781 409.761719 282.175781 407.15625 C 282.175781 404.546875 286.085938 404.546875 286.085938 407.15625 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 401.824219 C 286.085938 404.429688 282.175781 404.429688 282.175781 401.824219 C 282.175781 399.21875 286.085938 399.21875 286.085938 401.824219 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 396.496094 C 286.085938 399.101562 282.175781 399.101562 282.175781 396.496094 C 282.175781 393.890625 286.085938 393.890625 286.085938 396.496094 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 401.824219 C 286.085938 404.429688 282.175781 404.429688 282.175781 401.824219 C 282.175781 399.21875 286.085938 399.21875 286.085938 401.824219 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 412.484375 C 279.300781 415.089844 275.394531 415.089844 275.394531 412.484375 C 275.394531 409.878906 279.300781 409.878906 279.300781 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 428.472656 C 279.300781 431.082031 275.394531 431.082031 275.394531 428.472656 C 275.394531 425.867188 279.300781 425.867188 279.300781 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 245.390625 455.125 C 245.390625 457.730469 241.480469 457.730469 241.480469 455.125 C 241.480469 452.515625 245.390625 452.515625 245.390625 455.125 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 417.816406 C 279.300781 420.421875 275.394531 420.421875 275.394531 417.816406 C 275.394531 415.207031 279.300781 415.207031 279.300781 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 423.144531 C 272.519531 425.75 268.609375 425.75 268.609375 423.144531 C 268.609375 420.539062 272.519531 420.539062 272.519531 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 444.464844 C 258.953125 447.070312 255.042969 447.070312 255.042969 444.464844 C 255.042969 441.859375 258.953125 441.859375 258.953125 444.464844 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 423.144531 C 265.738281 425.75 261.828125 425.75 261.828125 423.144531 C 261.828125 420.539062 265.738281 420.539062 265.738281 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 423.144531 C 265.738281 425.75 261.828125 425.75 261.828125 423.144531 C 261.828125 420.539062 265.738281 420.539062 265.738281 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 428.472656 C 258.953125 431.082031 255.042969 431.082031 255.042969 428.472656 C 255.042969 425.867188 258.953125 425.867188 258.953125 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 433.804688 C 258.953125 436.410156 255.042969 436.410156 255.042969 433.804688 C 255.042969 431.199219 258.953125 431.199219 258.953125 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 423.144531 C 265.738281 425.75 261.828125 425.75 261.828125 423.144531 C 261.828125 420.539062 265.738281 420.539062 265.738281 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 417.816406 C 272.519531 420.421875 268.609375 420.421875 268.609375 417.816406 C 268.609375 415.207031 272.519531 415.207031 272.519531 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 423.144531 C 265.738281 425.75 261.828125 425.75 261.828125 423.144531 C 261.828125 420.539062 265.738281 420.539062 265.738281 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 292.867188 385.835938 C 292.867188 388.441406 288.957031 388.441406 288.957031 385.835938 C 288.957031 383.230469 292.867188 383.230469 292.867188 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 428.472656 C 279.300781 431.082031 275.394531 431.082031 275.394531 428.472656 C 275.394531 425.867188 279.300781 425.867188 279.300781 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 423.144531 C 272.519531 425.75 268.609375 425.75 268.609375 423.144531 C 268.609375 420.539062 272.519531 420.539062 272.519531 423.144531 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 243.433594 452.394531 L 244.035156 451.695312 L 245.238281 450.304688 L 245.839844 449.605469 L 246.4375 448.90625 L 248.84375 446.109375 L 249.445312 445.40625 L 250.046875 444.707031 L 250.644531 444.003906 L 251.246094 443.300781 L 251.847656 442.59375 L 252.449219 441.890625 L 253.652344 440.476562 L 254.253906 439.765625 L 254.851562 439.058594 L 255.453125 438.347656 L 257.257812 436.203125 L 257.859375 435.484375 L 258.460938 434.761719 L 259.058594 434.039062 L 259.660156 433.316406 L 260.261719 432.589844 L 260.863281 431.859375 L 262.066406 430.390625 L 262.667969 429.652344 L 263.265625 428.910156 L 263.867188 428.164062 L 264.46875 427.414062 L 265.070312 426.660156 L 265.671875 425.902344 L 266.273438 425.140625 L 266.875 424.375 L 267.472656 423.601562 L 268.074219 422.828125 L 268.675781 422.050781 L 269.878906 420.480469 L 271.082031 418.894531 L 271.679688 418.09375 L 272.882812 416.484375 L 273.484375 415.675781 L 274.085938 414.863281 L 274.6875 414.046875 L 275.285156 413.226562 L 275.886719 412.40625 L 277.089844 410.757812 L 277.691406 409.929688 L 278.894531 408.265625 L 279.492188 407.433594 L 281.296875 404.925781 L 282.5 403.246094 L 283.101562 402.402344 L 283.699219 401.5625 L 284.902344 399.875 L 285.503906 399.027344 L 286.105469 398.183594 L 287.308594 396.488281 L 287.90625 395.640625 L 288.507812 394.792969 L 289.109375 393.941406 L 289.710938 393.09375 L 290.3125 392.242188 L 290.914062 391.394531 L 290.914062 398.976562 L 290.3125 399.675781 L 289.710938 400.371094 L 287.90625 402.46875 L 287.308594 403.167969 L 286.707031 403.867188 L 286.105469 404.570312 L 285.503906 405.269531 L 283.699219 407.378906 L 283.101562 408.085938 L 281.296875 410.207031 L 279.492188 412.339844 L 278.894531 413.054688 L 278.292969 413.769531 L 276.488281 415.925781 L 275.285156 417.378906 L 274.6875 418.105469 L 273.484375 419.574219 L 272.28125 421.050781 L 271.679688 421.796875 L 271.082031 422.546875 L 270.480469 423.296875 L 269.277344 424.8125 L 268.675781 425.578125 L 268.074219 426.347656 L 267.472656 427.121094 L 266.875 427.898438 L 266.273438 428.679688 L 265.671875 429.464844 L 265.070312 430.253906 L 264.46875 431.046875 L 263.867188 431.84375 L 263.265625 432.644531 L 262.667969 433.453125 L 262.066406 434.257812 L 261.464844 435.070312 L 260.261719 436.703125 L 259.660156 437.523438 L 259.058594 438.347656 L 258.460938 439.171875 L 257.257812 440.828125 L 255.453125 443.324219 L 254.851562 444.160156 L 254.253906 445 L 251.847656 448.359375 L 250.644531 450.046875 L 250.046875 450.894531 L 249.445312 451.738281 L 248.84375 452.585938 L 248.242188 453.429688 L 247.640625 454.277344 L 247.039062 455.128906 L 246.4375 455.976562 L 245.839844 456.824219 L 244.636719 458.527344 L 244.035156 459.375 L 243.433594 460.226562 Z M 243.433594 452.394531 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 243.433594 456.308594 L 245.839844 453.214844 L 246.4375 452.441406 L 250.046875 447.800781 L 250.644531 447.027344 L 251.246094 446.25 L 254.253906 442.382812 L 254.851562 441.609375 L 258.460938 436.96875 L 259.058594 436.195312 L 259.660156 435.417969 L 262.667969 431.550781 L 263.265625 430.777344 L 266.875 426.136719 L 267.472656 425.359375 L 271.082031 420.71875 L 271.679688 419.945312 L 274.6875 416.078125 L 275.285156 415.304688 L 275.886719 414.527344 L 278.894531 410.660156 L 279.492188 409.886719 L 283.101562 405.246094 L 283.699219 404.46875 L 287.308594 399.828125 L 287.90625 399.054688 L 290.914062 395.1875 "/>
<g clip-path="url(#clip81)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(85.098039%,85.098039%,85.098039%);fill-opacity:1;" d="M 35.152344 258.171875 L 229.136719 258.171875 L 229.136719 240.40625 L 35.152344 240.40625 Z M 35.152344 258.171875 "/>
</g>
<g style="fill:rgb(10.196078%,10.196078%,10.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-1" x="129.644531" y="252.204102"/>
</g>
<g clip-path="url(#clip82)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(85.098039%,85.098039%,85.098039%);fill-opacity:1;" d="M 234.617188 258.171875 L 428.601562 258.171875 L 428.601562 240.40625 L 234.617188 240.40625 Z M 234.617188 258.171875 "/>
</g>
<g style="fill:rgb(10.196078%,10.196078%,10.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-2" x="329.109375" y="252.204102"/>
</g>
<g clip-path="url(#clip83)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(85.098039%,85.098039%,85.098039%);fill-opacity:1;" d="M 35.152344 23.246094 L 229.136719 23.246094 L 229.136719 5.480469 L 35.152344 5.480469 Z M 35.152344 23.246094 "/>
</g>
<g style="fill:rgb(10.196078%,10.196078%,10.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-3" x="129.644531" y="17.27832"/>
</g>
<g clip-path="url(#clip84)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(85.098039%,85.098039%,85.098039%);fill-opacity:1;" d="M 234.617188 23.246094 L 428.601562 23.246094 L 428.601562 5.480469 L 234.617188 5.480469 Z M 234.617188 23.246094 "/>
</g>
<g style="fill:rgb(10.196078%,10.196078%,10.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-4" x="329.109375" y="17.27832"/>
</g>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 50.75 472.589844 L 50.75 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 84.664062 472.589844 L 84.664062 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 118.578125 472.589844 L 118.578125 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 152.492188 472.589844 L 152.492188 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 186.40625 472.589844 L 186.40625 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 220.320312 472.589844 L 220.320312 469.847656 "/>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="45.75" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="50.75" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="79.664062" y="481.196289"/>
  <use xlink:href="#glyph0-4" x="84.664062" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-7" x="113.578125" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="118.578125" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-7" x="147.492188" y="481.196289"/>
  <use xlink:href="#glyph0-4" x="152.492188" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-8" x="181.40625" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="186.40625" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-8" x="215.320312" y="481.196289"/>
  <use xlink:href="#glyph0-4" x="220.320312" y="481.196289"/>
</g>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 250.214844 472.589844 L 250.214844 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 284.128906 472.589844 L 284.128906 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 318.042969 472.589844 L 318.042969 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 351.957031 472.589844 L 351.957031 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 385.871094 472.589844 L 385.871094 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 419.785156 472.589844 L 419.785156 469.847656 "/>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="245.214844" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="250.214844" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="279.128906" y="481.196289"/>
  <use xlink:href="#glyph0-4" x="284.128906" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-7" x="313.042969" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="318.042969" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-7" x="346.957031" y="481.196289"/>
  <use xlink:href="#glyph0-4" x="351.957031" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-8" x="380.871094" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="385.871094" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-8" x="414.785156" y="481.196289"/>
  <use xlink:href="#glyph0-4" x="419.785156" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="20.21875" y="233.774414"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="233.774414"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-7" x="20.21875" y="180.473633"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="180.473633"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-8" x="20.21875" y="127.176758"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="127.176758"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-3" x="20.21875" y="73.875977"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="73.875977"/>
</g>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 230.859375 L 35.152344 230.859375 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 177.558594 L 35.152344 177.558594 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 124.261719 L 35.152344 124.261719 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 70.960938 L 35.152344 70.960938 "/>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="20.21875" y="468.700195"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="468.700195"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-7" x="20.21875" y="415.399414"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="415.399414"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-8" x="20.21875" y="362.102539"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="362.102539"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-3" x="20.21875" y="308.801758"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="308.801758"/>
</g>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 465.785156 L 35.152344 465.785156 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 412.484375 L 35.152344 412.484375 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 359.1875 L 35.152344 359.1875 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 305.886719 L 35.152344 305.886719 "/>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph1-1" x="224.875" y="494.538086"/>
  <use xlink:href="#glyph1-2" x="230.875" y="494.538086"/>
  <use xlink:href="#glyph1-3" x="233.875" y="494.538086"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph2-1" x="13.499023" y="256.046875"/>
  <use xlink:href="#glyph2-2" x="13.499023" y="250.046875"/>
  <use xlink:href="#glyph2-3" x="13.499023" y="242.046875"/>
</g>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(100%,100%,100%);fill-opacity:1;" d="M 439.5625 295.328125 L 498.519531 295.328125 L 498.519531 197.769531 L 439.5625 197.769531 Z M 439.5625 295.328125 "/>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph1-4" x="445.042969" y="212.768555"/>
  <use xlink:href="#glyph1-5" x="448.042969" y="212.768555"/>
  <use xlink:href="#glyph1-1" x="454.042969" y="212.768555"/>
  <use xlink:href="#glyph1-2" x="460.042969" y="212.768555"/>
  <use xlink:href="#glyph1-6" x="463.042969" y="212.768555"/>
  <use xlink:href="#glyph1-7" x="469.042969" y="212.768555"/>
  <use xlink:href="#glyph1-8" x="473.042969" y="212.768555"/>
  <use xlink:href="#glyph1-1" x="477.042969" y="212.768555"/>
  <use xlink:href="#glyph1-3" x="482.042969" y="212.768555"/>
  <use xlink:href="#glyph1-9" x="487.042969" y="212.768555"/>
  <use xlink:href="#glyph1-10" x="489.042969" y="212.768555"/>
</g>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(94.901961%,94.901961%,94.901961%);fill-opacity:1;" d="M 445.042969 238.007812 L 462.324219 238.007812 L 462.324219 220.726562 L 445.042969 220.726562 Z M 445.042969 238.007812 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 455.636719 229.367188 C 455.636719 231.972656 451.726562 231.972656 451.726562 229.367188 C 451.726562 226.761719 455.636719 226.761719 455.636719 229.367188 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 445.042969 238.007812 L 462.324219 238.007812 L 462.324219 220.726562 L 445.042969 220.726562 Z M 445.042969 238.007812 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 446.769531 229.367188 L 460.59375 229.367188 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(94.901961%,94.901961%,94.901961%);fill-opacity:1;" d="M 445.042969 255.289062 L 462.324219 255.289062 L 462.324219 238.007812 L 445.042969 238.007812 Z M 445.042969 255.289062 "/>
<path style="fill-rule:nonzero;fill:rgb(48.627451%,68.235294%,0%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 455.636719 246.648438 C 455.636719 249.253906 451.726562 249.253906 451.726562 246.648438 C 451.726562 244.042969 455.636719 244.042969 455.636719 246.648438 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 445.042969 255.289062 L 462.324219 255.289062 L 462.324219 238.007812 L 445.042969 238.007812 Z M 445.042969 255.289062 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 446.769531 246.648438 L 460.59375 246.648438 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(94.901961%,94.901961%,94.901961%);fill-opacity:1;" d="M 445.042969 272.566406 L 462.324219 272.566406 L 462.324219 255.285156 L 445.042969 255.285156 Z M 445.042969 272.566406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 455.636719 263.925781 C 455.636719 266.535156 451.726562 266.535156 451.726562 263.925781 C 451.726562 261.320312 455.636719 261.320312 455.636719 263.925781 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 445.042969 272.566406 L 462.324219 272.566406 L 462.324219 255.285156 L 445.042969 255.285156 Z M 445.042969 272.566406 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 446.769531 263.925781 L 460.59375 263.925781 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(94.901961%,94.901961%,94.901961%);fill-opacity:1;" d="M 445.042969 289.847656 L 462.324219 289.847656 L 462.324219 272.566406 L 445.042969 272.566406 Z M 445.042969 289.847656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 455.636719 281.207031 C 455.636719 283.8125 451.726562 283.8125 451.726562 281.207031 C 451.726562 278.601562 455.636719 278.601562 455.636719 281.207031 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 445.042969 289.847656 L 462.324219 289.847656 L 462.324219 272.566406 L 445.042969 272.566406 Z M 445.042969 289.847656 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 446.769531 281.207031 L 460.59375 281.207031 "/>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph0-3" x="467.800781" y="232.282227"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph0-4" x="467.800781" y="249.563477"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph0-1" x="467.800781" y="266.84082"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph0-2" x="467.800781" y="284.12207"/>
</g>
</g>
</svg>

::: {.sourceClojure}
```clojure
(-> plot
    r/r->clj)
```
:::

::: {.printedClojure}
```clojure
{:data
 {:manufacturer
  ["audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "audi"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "chevrolet"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "dodge"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "ford"
   "honda"
   "honda"
   "honda"
   "honda"
   "honda"
   "honda"
   "honda"
   "honda"
   "honda"
   "hyundai"
   "hyundai"
   "hyundai"
   "hyundai"
   "hyundai"
   "hyundai"
   "hyundai"
   "hyundai"
   "hyundai"
   "hyundai"
   "hyundai"
   "hyundai"
   "hyundai"
   "hyundai"
   "jeep"
   "jeep"
   "jeep"
   "jeep"
   "jeep"
   "jeep"
   "jeep"
   "jeep"
   "land rover"
   "land rover"
   "land rover"
   "land rover"
   "lincoln"
   "lincoln"
   "lincoln"
   "mercury"
   "mercury"
   "mercury"
   "mercury"
   "nissan"
   "nissan"
   "nissan"
   "nissan"
   "nissan"
   "nissan"
   "nissan"
   "nissan"
   "nissan"
   "nissan"
   "nissan"
   "nissan"
   "nissan"
   "pontiac"
   "pontiac"
   "pontiac"
   "pontiac"
   "pontiac"
   "subaru"
   "subaru"
   "subaru"
   "subaru"
   "subaru"
   "subaru"
   "subaru"
   "subaru"
   "subaru"
   "subaru"
   "subaru"
   "subaru"
   "subaru"
   "subaru"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "toyota"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"
   "volkswagen"],
  :model
  ["a4"
   "a4"
   "a4"
   "a4"
   "a4"
   "a4"
   "a4"
   "a4 quattro"
   "a4 quattro"
   "a4 quattro"
   "a4 quattro"
   "a4 quattro"
   "a4 quattro"
   "a4 quattro"
   "a4 quattro"
   "a6 quattro"
   "a6 quattro"
   "a6 quattro"
   "c1500 suburban 2wd"
   "c1500 suburban 2wd"
   "c1500 suburban 2wd"
   "c1500 suburban 2wd"
   "c1500 suburban 2wd"
   "corvette"
   "corvette"
   "corvette"
   "corvette"
   "corvette"
   "k1500 tahoe 4wd"
   "k1500 tahoe 4wd"
   "k1500 tahoe 4wd"
   "k1500 tahoe 4wd"
   "malibu"
   "malibu"
   "malibu"
   "malibu"
   "malibu"
   "caravan 2wd"
   "caravan 2wd"
   "caravan 2wd"
   "caravan 2wd"
   "caravan 2wd"
   "caravan 2wd"
   "caravan 2wd"
   "caravan 2wd"
   "caravan 2wd"
   "caravan 2wd"
   "caravan 2wd"
   "dakota pickup 4wd"
   "dakota pickup 4wd"
   "dakota pickup 4wd"
   "dakota pickup 4wd"
   "dakota pickup 4wd"
   "dakota pickup 4wd"
   "dakota pickup 4wd"
   "dakota pickup 4wd"
   "dakota pickup 4wd"
   "durango 4wd"
   "durango 4wd"
   "durango 4wd"
   "durango 4wd"
   "durango 4wd"
   "durango 4wd"
   "durango 4wd"
   "ram 1500 pickup 4wd"
   "ram 1500 pickup 4wd"
   "ram 1500 pickup 4wd"
   "ram 1500 pickup 4wd"
   "ram 1500 pickup 4wd"
   "ram 1500 pickup 4wd"
   "ram 1500 pickup 4wd"
   "ram 1500 pickup 4wd"
   "ram 1500 pickup 4wd"
   "ram 1500 pickup 4wd"
   "expedition 2wd"
   "expedition 2wd"
   "expedition 2wd"
   "explorer 4wd"
   "explorer 4wd"
   "explorer 4wd"
   "explorer 4wd"
   "explorer 4wd"
   "explorer 4wd"
   "f150 pickup 4wd"
   "f150 pickup 4wd"
   "f150 pickup 4wd"
   "f150 pickup 4wd"
   "f150 pickup 4wd"
   "f150 pickup 4wd"
   "f150 pickup 4wd"
   "mustang"
   "mustang"
   "mustang"
   "mustang"
   "mustang"
   "mustang"
   "mustang"
   "mustang"
   "mustang"
   "civic"
   "civic"
   "civic"
   "civic"
   "civic"
   "civic"
   "civic"
   "civic"
   "civic"
   "sonata"
   "sonata"
   "sonata"
   "sonata"
   "sonata"
   "sonata"
   "sonata"
   "tiburon"
   "tiburon"
   "tiburon"
   "tiburon"
   "tiburon"
   "tiburon"
   "tiburon"
   "grand cherokee 4wd"
   "grand cherokee 4wd"
   "grand cherokee 4wd"
   "grand cherokee 4wd"
   "grand cherokee 4wd"
   "grand cherokee 4wd"
   "grand cherokee 4wd"
   "grand cherokee 4wd"
   "range rover"
   "range rover"
   "range rover"
   "range rover"
   "navigator 2wd"
   "navigator 2wd"
   "navigator 2wd"
   "mountaineer 4wd"
   "mountaineer 4wd"
   "mountaineer 4wd"
   "mountaineer 4wd"
   "altima"
   "altima"
   "altima"
   "altima"
   "altima"
   "altima"
   "maxima"
   "maxima"
   "maxima"
   "pathfinder 4wd"
   "pathfinder 4wd"
   "pathfinder 4wd"
   "pathfinder 4wd"
   "grand prix"
   "grand prix"
   "grand prix"
   "grand prix"
   "grand prix"
   "forester awd"
   "forester awd"
   "forester awd"
   "forester awd"
   "forester awd"
   "forester awd"
   "impreza awd"
   "impreza awd"
   "impreza awd"
   "impreza awd"
   "impreza awd"
   "impreza awd"
   "impreza awd"
   "impreza awd"
   "4runner 4wd"
   "4runner 4wd"
   "4runner 4wd"
   "4runner 4wd"
   "4runner 4wd"
   "4runner 4wd"
   "camry"
   "camry"
   "camry"
   "camry"
   "camry"
   "camry"
   "camry"
   "camry solara"
   "camry solara"
   "camry solara"
   "camry solara"
   "camry solara"
   "camry solara"
   "camry solara"
   "corolla"
   "corolla"
   "corolla"
   "corolla"
   "corolla"
   "land cruiser wagon 4wd"
   "land cruiser wagon 4wd"
   "toyota tacoma 4wd"
   "toyota tacoma 4wd"
   "toyota tacoma 4wd"
   "toyota tacoma 4wd"
   "toyota tacoma 4wd"
   "toyota tacoma 4wd"
   "toyota tacoma 4wd"
   "gti"
   "gti"
   "gti"
   "gti"
   "gti"
   "jetta"
   "jetta"
   "jetta"
   "jetta"
   "jetta"
   "jetta"
   "jetta"
   "jetta"
   "jetta"
   "new beetle"
   "new beetle"
   "new beetle"
   "new beetle"
   "new beetle"
   "new beetle"
   "passat"
   "passat"
   "passat"
   "passat"
   "passat"
   "passat"
   "passat"],
  :displ
  [1.8
   1.8
   2.0
   2.0
   2.8
   2.8
   3.1
   1.8
   1.8
   2.0
   2.0
   2.8
   2.8
   3.1
   3.1
   2.8
   3.1
   4.2
   5.3
   5.3
   5.3
   5.7
   6.0
   5.7
   5.7
   6.2
   6.2
   7.0
   5.3
   5.3
   5.7
   6.5
   2.4
   2.4
   3.1
   3.5
   3.6
   2.4
   3.0
   3.3
   3.3
   3.3
   3.3
   3.3
   3.8
   3.8
   3.8
   4.0
   3.7
   3.7
   3.9
   3.9
   4.7
   4.7
   4.7
   5.2
   5.2
   3.9
   4.7
   4.7
   4.7
   5.2
   5.7
   5.9
   4.7
   4.7
   4.7
   4.7
   4.7
   4.7
   5.2
   5.2
   5.7
   5.9
   4.6
   5.4
   5.4
   4.0
   4.0
   4.0
   4.0
   4.6
   5.0
   4.2
   4.2
   4.6
   4.6
   4.6
   5.4
   5.4
   3.8
   3.8
   4.0
   4.0
   4.6
   4.6
   4.6
   4.6
   5.4
   1.6
   1.6
   1.6
   1.6
   1.6
   1.8
   1.8
   1.8
   2.0
   2.4
   2.4
   2.4
   2.4
   2.5
   2.5
   3.3
   2.0
   2.0
   2.0
   2.0
   2.7
   2.7
   2.7
   3.0
   3.7
   4.0
   4.7
   4.7
   4.7
   5.7
   6.1
   4.0
   4.2
   4.4
   4.6
   5.4
   5.4
   5.4
   4.0
   4.0
   4.6
   5.0
   2.4
   2.4
   2.5
   2.5
   3.5
   3.5
   3.0
   3.0
   3.5
   3.3
   3.3
   4.0
   5.6
   3.1
   3.8
   3.8
   3.8
   5.3
   2.5
   2.5
   2.5
   2.5
   2.5
   2.5
   2.2
   2.2
   2.5
   2.5
   2.5
   2.5
   2.5
   2.5
   2.7
   2.7
   3.4
   3.4
   4.0
   4.7
   2.2
   2.2
   2.4
   2.4
   3.0
   3.0
   3.5
   2.2
   2.2
   2.4
   2.4
   3.0
   3.0
   3.3
   1.8
   1.8
   1.8
   1.8
   1.8
   4.7
   5.7
   2.7
   2.7
   2.7
   3.4
   3.4
   4.0
   4.0
   2.0
   2.0
   2.0
   2.0
   2.8
   1.9
   2.0
   2.0
   2.0
   2.0
   2.5
   2.5
   2.8
   2.8
   1.9
   1.9
   2.0
   2.0
   2.5
   2.5
   1.8
   1.8
   2.0
   2.0
   2.8
   2.8
   3.6],
  :year
  [1999
   1999
   2008
   2008
   1999
   1999
   2008
   1999
   1999
   2008
   2008
   1999
   1999
   2008
   2008
   1999
   2008
   2008
   2008
   2008
   2008
   1999
   2008
   1999
   1999
   2008
   2008
   2008
   2008
   2008
   1999
   1999
   1999
   2008
   1999
   2008
   2008
   1999
   1999
   1999
   1999
   2008
   2008
   2008
   1999
   1999
   2008
   2008
   2008
   2008
   1999
   1999
   2008
   2008
   2008
   1999
   1999
   1999
   2008
   2008
   2008
   1999
   2008
   1999
   2008
   2008
   2008
   2008
   2008
   2008
   1999
   1999
   2008
   1999
   1999
   1999
   2008
   1999
   1999
   1999
   2008
   2008
   1999
   1999
   1999
   1999
   1999
   2008
   1999
   2008
   1999
   1999
   2008
   2008
   1999
   1999
   2008
   2008
   2008
   1999
   1999
   1999
   1999
   1999
   2008
   2008
   2008
   2008
   1999
   1999
   2008
   2008
   1999
   1999
   2008
   1999
   1999
   2008
   2008
   2008
   2008
   2008
   2008
   2008
   1999
   1999
   2008
   2008
   2008
   2008
   1999
   2008
   2008
   1999
   1999
   1999
   2008
   1999
   2008
   2008
   1999
   1999
   1999
   2008
   2008
   2008
   2008
   1999
   1999
   2008
   1999
   1999
   2008
   2008
   1999
   1999
   1999
   2008
   2008
   1999
   1999
   2008
   2008
   2008
   2008
   1999
   1999
   1999
   1999
   2008
   2008
   2008
   2008
   1999
   1999
   1999
   1999
   2008
   2008
   1999
   1999
   2008
   2008
   1999
   1999
   2008
   1999
   1999
   2008
   2008
   1999
   1999
   2008
   1999
   1999
   1999
   2008
   2008
   1999
   2008
   1999
   1999
   2008
   1999
   1999
   2008
   2008
   1999
   1999
   2008
   2008
   1999
   1999
   1999
   1999
   2008
   2008
   2008
   2008
   1999
   1999
   1999
   1999
   1999
   1999
   2008
   2008
   1999
   1999
   2008
   2008
   1999
   1999
   2008],
  :cyl
  [4
   4
   4
   4
   6
   6
   6
   4
   4
   4
   4
   6
   6
   6
   6
   6
   6
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   4
   4
   6
   6
   6
   4
   6
   6
   6
   6
   6
   6
   6
   6
   6
   6
   6
   6
   6
   6
   8
   8
   8
   8
   8
   6
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   6
   6
   6
   6
   8
   8
   6
   6
   8
   8
   8
   8
   8
   6
   6
   6
   6
   8
   8
   8
   8
   8
   4
   4
   4
   4
   4
   4
   4
   4
   4
   4
   4
   4
   4
   6
   6
   6
   4
   4
   4
   4
   6
   6
   6
   6
   6
   6
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   8
   6
   6
   8
   8
   4
   4
   4
   4
   6
   6
   6
   6
   6
   6
   6
   6
   8
   6
   6
   6
   6
   8
   4
   4
   4
   4
   4
   4
   4
   4
   4
   4
   4
   4
   4
   4
   4
   4
   6
   6
   6
   8
   4
   4
   4
   4
   6
   6
   6
   4
   4
   4
   4
   6
   6
   6
   4
   4
   4
   4
   4
   8
   8
   4
   4
   4
   6
   6
   6
   6
   4
   4
   4
   4
   6
   4
   4
   4
   4
   4
   5
   5
   6
   6
   4
   4
   4
   4
   5
   5
   4
   4
   4
   4
   6
   6
   6],
  :trans
  ["auto(l5)"
   "manual(m5)"
   "manual(m6)"
   "auto(av)"
   "auto(l5)"
   "manual(m5)"
   "auto(av)"
   "manual(m5)"
   "auto(l5)"
   "manual(m6)"
   "auto(s6)"
   "auto(l5)"
   "manual(m5)"
   "auto(s6)"
   "manual(m6)"
   "auto(l5)"
   "auto(s6)"
   "auto(s6)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "manual(m6)"
   "auto(l4)"
   "manual(m6)"
   "auto(s6)"
   "manual(m6)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(s6)"
   "auto(l3)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l6)"
   "auto(l6)"
   "manual(m6)"
   "auto(l4)"
   "auto(l4)"
   "manual(m5)"
   "auto(l5)"
   "auto(l5)"
   "auto(l5)"
   "manual(m5)"
   "auto(l4)"
   "auto(l4)"
   "auto(l5)"
   "auto(l5)"
   "auto(l5)"
   "auto(l4)"
   "auto(l5)"
   "auto(l4)"
   "manual(m6)"
   "auto(l5)"
   "auto(l5)"
   "auto(l5)"
   "manual(m6)"
   "manual(m6)"
   "auto(l4)"
   "manual(m5)"
   "auto(l5)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l6)"
   "auto(l5)"
   "manual(m5)"
   "auto(l5)"
   "auto(l5)"
   "auto(l6)"
   "auto(l4)"
   "auto(l4)"
   "manual(m5)"
   "manual(m5)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "manual(m5)"
   "auto(l4)"
   "manual(m5)"
   "auto(l5)"
   "auto(l4)"
   "manual(m5)"
   "manual(m5)"
   "auto(l5)"
   "manual(m6)"
   "manual(m5)"
   "auto(l4)"
   "manual(m5)"
   "manual(m5)"
   "auto(l4)"
   "manual(m5)"
   "auto(l5)"
   "auto(l5)"
   "manual(m6)"
   "auto(l4)"
   "manual(m5)"
   "auto(l4)"
   "manual(m5)"
   "auto(l4)"
   "manual(m5)"
   "auto(l5)"
   "auto(l4)"
   "manual(m5)"
   "manual(m5)"
   "auto(l4)"
   "auto(l4)"
   "manual(m6)"
   "manual(m5)"
   "auto(l5)"
   "auto(l5)"
   "auto(l4)"
   "auto(l4)"
   "auto(l5)"
   "auto(l5)"
   "auto(l5)"
   "auto(l5)"
   "auto(l4)"
   "auto(s6)"
   "auto(s6)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l6)"
   "auto(l5)"
   "auto(l5)"
   "auto(l6)"
   "auto(l4)"
   "manual(m5)"
   "auto(l4)"
   "auto(av)"
   "manual(m6)"
   "manual(m6)"
   "auto(av)"
   "auto(l4)"
   "manual(m5)"
   "auto(av)"
   "auto(l4)"
   "manual(m5)"
   "auto(l5)"
   "auto(s5)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "auto(s4)"
   "manual(m5)"
   "auto(l4)"
   "manual(m5)"
   "manual(m5)"
   "auto(l4)"
   "auto(l4)"
   "auto(l4)"
   "manual(m5)"
   "manual(m5)"
   "auto(l4)"
   "auto(s4)"
   "auto(s4)"
   "manual(m5)"
   "manual(m5)"
   "manual(m5)"
   "auto(l4)"
   "auto(l4)"
   "manual(m5)"
   "auto(l5)"
   "auto(l5)"
   "manual(m5)"
   "auto(l4)"
   "manual(m5)"
   "auto(l5)"
   "auto(l4)"
   "manual(m5)"
   "auto(s6)"
   "auto(l4)"
   "manual(m5)"
   "manual(m5)"
   "auto(s5)"
   "auto(l4)"
   "manual(m5)"
   "auto(s5)"
   "auto(l3)"
   "auto(l4)"
   "manual(m5)"
   "manual(m5)"
   "auto(l4)"
   "auto(l4)"
   "auto(s6)"
   "manual(m5)"
   "auto(l4)"
   "manual(m5)"
   "manual(m5)"
   "auto(l4)"
   "manual(m6)"
   "auto(l5)"
   "manual(m5)"
   "auto(l4)"
   "manual(m6)"
   "auto(s6)"
   "manual(m5)"
   "manual(m5)"
   "manual(m5)"
   "auto(l4)"
   "auto(s6)"
   "manual(m6)"
   "auto(s6)"
   "manual(m5)"
   "auto(l4)"
   "manual(m5)"
   "manual(m5)"
   "auto(l4)"
   "manual(m5)"
   "auto(l4)"
   "manual(m5)"
   "auto(s6)"
   "manual(m5)"
   "auto(l5)"
   "auto(s6)"
   "manual(m6)"
   "auto(l5)"
   "manual(m5)"
   "auto(s6)"],
  :drv
  ["f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "4"
   "4"
   "4"
   "4"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "r"
   "r"
   "r"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "r"
   "r"
   "r"
   "4"
   "4"
   "4"
   "4"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "4"
   "4"
   "4"
   "4"
   "f"
   "f"
   "f"
   "f"
   "f"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "4"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"
   "f"],
  :cty
  [18
   21
   20
   21
   16
   18
   18
   18
   16
   20
   19
   15
   17
   17
   15
   15
   17
   16
   14
   11
   14
   13
   12
   16
   15
   16
   15
   15
   14
   11
   11
   14
   19
   22
   18
   18
   17
   18
   17
   16
   16
   17
   17
   11
   15
   15
   16
   16
   15
   14
   13
   14
   14
   14
   9
   11
   11
   13
   13
   9
   13
   11
   13
   11
   12
   9
   13
   13
   12
   9
   11
   11
   13
   11
   11
   11
   12
   14
   15
   14
   13
   13
   13
   14
   14
   13
   13
   13
   11
   13
   18
   18
   17
   16
   15
   15
   15
   15
   14
   28
   24
   25
   23
   24
   26
   25
   24
   21
   18
   18
   21
   21
   18
   18
   19
   19
   19
   20
   20
   17
   16
   17
   17
   15
   15
   14
   9
   14
   13
   11
   11
   12
   12
   11
   11
   11
   12
   14
   13
   13
   13
   21
   19
   23
   23
   19
   19
   18
   19
   19
   14
   15
   14
   12
   18
   16
   17
   18
   16
   18
   18
   20
   19
   20
   18
   21
   19
   19
   19
   20
   20
   19
   20
   15
   16
   15
   15
   16
   14
   21
   21
   21
   21
   18
   18
   19
   21
   21
   21
   22
   18
   18
   18
   24
   24
   26
   28
   26
   11
   13
   15
   16
   17
   15
   15
   15
   16
   21
   19
   21
   22
   17
   33
   21
   19
   22
   21
   21
   21
   16
   17
   35
   29
   21
   19
   20
   20
   21
   18
   19
   21
   16
   18
   17],
  :hwy
  [29
   29
   31
   30
   26
   26
   27
   26
   25
   28
   27
   25
   25
   25
   25
   24
   25
   23
   20
   15
   20
   17
   17
   26
   23
   26
   25
   24
   19
   14
   15
   17
   27
   30
   26
   29
   26
   24
   24
   22
   22
   24
   24
   17
   22
   21
   23
   23
   19
   18
   17
   17
   19
   19
   12
   17
   15
   17
   17
   12
   17
   16
   18
   15
   16
   12
   17
   17
   16
   12
   15
   16
   17
   15
   17
   17
   18
   17
   19
   17
   19
   19
   17
   17
   17
   16
   16
   17
   15
   17
   26
   25
   26
   24
   21
   22
   23
   22
   20
   33
   32
   32
   29
   32
   34
   36
   36
   29
   26
   27
   30
   31
   26
   26
   28
   26
   29
   28
   27
   24
   24
   24
   22
   19
   20
   17
   12
   19
   18
   14
   15
   18
   18
   15
   17
   16
   18
   17
   19
   19
   17
   29
   27
   31
   32
   27
   26
   26
   25
   25
   17
   17
   20
   18
   26
   26
   27
   28
   25
   25
   24
   27
   25
   26
   23
   26
   26
   26
   26
   25
   27
   25
   27
   20
   20
   19
   17
   20
   17
   29
   27
   31
   31
   26
   26
   28
   27
   29
   31
   31
   26
   26
   27
   30
   33
   35
   37
   35
   15
   18
   20
   20
   22
   17
   19
   18
   20
   29
   26
   29
   29
   24
   44
   29
   26
   29
   29
   29
   29
   23
   24
   44
   41
   29
   26
   28
   29
   29
   29
   28
   29
   26
   26
   26],
  :fl
  ["p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "r"
   "e"
   "r"
   "r"
   "r"
   "p"
   "p"
   "p"
   "p"
   "p"
   "r"
   "e"
   "r"
   "d"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "e"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "e"
   "r"
   "r"
   "r"
   "r"
   "e"
   "r"
   "r"
   "r"
   "r"
   "r"
   "e"
   "r"
   "r"
   "r"
   "e"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "p"
   "r"
   "r"
   "r"
   "p"
   "r"
   "r"
   "r"
   "c"
   "p"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "d"
   "r"
   "r"
   "r"
   "e"
   "r"
   "r"
   "p"
   "p"
   "r"
   "r"
   "p"
   "r"
   "p"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "p"
   "p"
   "r"
   "r"
   "p"
   "r"
   "r"
   "p"
   "p"
   "r"
   "p"
   "r"
   "r"
   "p"
   "r"
   "r"
   "r"
   "p"
   "r"
   "p"
   "r"
   "r"
   "r"
   "r"
   "p"
   "r"
   "p"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "r"
   "p"
   "p"
   "r"
   "d"
   "r"
   "r"
   "p"
   "p"
   "r"
   "r"
   "r"
   "r"
   "d"
   "d"
   "r"
   "r"
   "r"
   "r"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"
   "p"],
  :class
  ["compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "midsize"
   "midsize"
   "midsize"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "2seater"
   "2seater"
   "2seater"
   "2seater"
   "2seater"
   "suv"
   "suv"
   "suv"
   "suv"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "minivan"
   "minivan"
   "minivan"
   "minivan"
   "minivan"
   "minivan"
   "minivan"
   "minivan"
   "minivan"
   "minivan"
   "minivan"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "compact"
   "compact"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "suv"
   "suv"
   "suv"
   "suv"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "compact"
   "compact"
   "compact"
   "compact"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "suv"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "suv"
   "suv"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "pickup"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "compact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "subcompact"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"
   "midsize"]},
 :layers [nil nil],
 :scales nil,
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ [factor cyl]]},
 :theme [],
 :coordinates nil,
 :facet nil,
 :plot_env nil,
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

```
:::

<div style="background-color:grey;height:2px;width:100%;"></div>

<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/main/projects/noj/notebooks/ggplot.clj">projects/noj/notebooks/ggplot.clj</a></small></small></pre></div>