### 2024-02-05 docs/projects/noj/index.qmd
Noj - rerendered

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: index.html}
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="index_files/md-default0.js" type="text/javascript"></script><script src="index_files/md-default1.js" type="text/javascript"></script>

```{=html}
<table><tr><td>This is part of the Scicloj <a href="https://scicloj.github.io/clojure-data-scrapbook/">Clojure Data Scrapbook</a>.</td><a href="https://scicloj.github.io/clojure-data-scrapbook/"><img align="left" alt="SciCloj logo" src="https://scicloj.github.io/sci-cloj-logo-transparent.png" width="40" /></a></tr></table>
```

# Preface
[Noj](https://scicloj.github.io/noj/) (short for Scinojure)
is an opinionated way to use the emerging Clojure data stack.
It collects a few of the relevant libraries together
and addes some layers of composition on top of them.

```{=html}
<div style="background-color:grey;height:2px;width:100%;"></div>
```

```{=html}
<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/main/projects/noj/notebooks/index.clj">projects/noj/notebooks/index.clj</a></small></small></pre></div>
```