### 2024-02-05 docs/projects/noj/ggplot.qmd
Noj - rerendered

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: ggplot.html}
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="ggplot_files/md-default4.js" type="text/javascript"></script><script src="ggplot_files/md-default5.js" type="text/javascript"></script><script src="ggplot_files/portal6.js" type="text/javascript"></script>

```{=html}
<table><tr><td>This is part of the Scicloj <a href="https://scicloj.github.io/clojure-data-scrapbook/">Clojure Data Scrapbook</a>.</td><a href="https://scicloj.github.io/clojure-data-scrapbook/"><img align="left" alt="SciCloj logo" src="https://scicloj.github.io/sci-cloj-logo-transparent.png" width="40" /></a></tr></table>
```

# Exploring ggplot
Here we explore [ggplot2](https://ggplot2.tidyverse.org/) from Clojure. The goal is to eventually implement a similar grammar in Clojure.

## Setup
We will use [ClojisR](https://scicloj.github.io/clojisr/) to call R from Clojure.

::: {.sourceClojure}
```clojure
(ns ggplot
  (:require [clojisr.v1.r :as r :refer [r r$ r->clj]]
            [clojisr.v1.applications.plotting :as plotting]
            [scicloj.kindly.v4.kind :as kind]
            [clojure.walk :as walk]
            [tablecloth.api :as tc]))
```
:::

::: {.sourceClojure}
```clojure
(r/library "ggplot2")
```
:::

::: {.printedClojure}
```clojure
[1] "ggplot2"   "Rserve"    "stats"     "graphics"  "grDevices" "utils"    
[7] "datasets"  "methods"   "base"

```
:::

::: {.sourceClojure}
```clojure
(r/require-r '[base])
```
:::

::: {.printedClojure}
```clojure
nil

```
:::

the base R library

## An example plot

::: {.sourceClojure}
```clojure
(-> "(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method=\"lm\")
         + facet_wrap(~cyl))"
    r
    plotting/plot->svg
    kind/html)
```
:::

```{=html}
<?xml version="1.0" encoding="UTF-8"?>
<svg xmlns="http://www.w3.org/2000/svg" xmlns:xlink="http://www.w3.org/1999/xlink" width="504pt" height="504pt" viewBox="0 0 504 504" version="1.1">
<defs>
<g>
<symbol overflow="visible" id="glyph0-0">
<path style="stroke:none;" d=""/>
</symbol>
<symbol overflow="visible" id="glyph0-1">
<path style="stroke:none;" d="M 4.375 -4.703125 C 4.234375 -5.75 3.5625 -6.359375 2.609375 -6.359375 C 1.921875 -6.359375 1.3125 -6.03125 0.9375 -5.453125 C 0.546875 -4.828125 0.375 -4.0625 0.375 -2.90625 C 0.375 -1.828125 0.53125 -1.15625 0.90625 -0.59375 C 1.234375 -0.078125 1.78125 0.203125 2.46875 0.203125 C 3.65625 0.203125 4.515625 -0.6875 4.515625 -1.921875 C 4.515625 -3.09375 3.71875 -3.921875 2.609375 -3.921875 C 1.984375 -3.921875 1.5 -3.6875 1.171875 -3.234375 C 1.171875 -4.796875 1.671875 -5.671875 2.5625 -5.671875 C 3.109375 -5.671875 3.484375 -5.328125 3.609375 -4.703125 Z M 2.5 -3.234375 C 3.25 -3.234375 3.71875 -2.703125 3.71875 -1.859375 C 3.71875 -1.0625 3.1875 -0.484375 2.484375 -0.484375 C 1.765625 -0.484375 1.21875 -1.09375 1.21875 -1.90625 C 1.21875 -2.6875 1.734375 -3.234375 2.5 -3.234375 Z M 2.5 -3.234375 "/>
</symbol>
<symbol overflow="visible" id="glyph0-2">
<path style="stroke:none;" d="M 3.4375 -3.359375 C 4.09375 -3.75 4.296875 -4.0625 4.296875 -4.65625 C 4.296875 -5.671875 3.53125 -6.359375 2.421875 -6.359375 C 1.3125 -6.359375 0.546875 -5.671875 0.546875 -4.671875 C 0.546875 -4.0625 0.75 -3.75 1.390625 -3.359375 C 0.671875 -2.984375 0.328125 -2.46875 0.328125 -1.765625 C 0.328125 -0.609375 1.171875 0.203125 2.421875 0.203125 C 3.65625 0.203125 4.515625 -0.609375 4.515625 -1.765625 C 4.515625 -2.46875 4.15625 -2.984375 3.4375 -3.359375 Z M 2.421875 -5.671875 C 3.078125 -5.671875 3.5 -5.28125 3.5 -4.640625 C 3.5 -4.046875 3.0625 -3.65625 2.421875 -3.65625 C 1.765625 -3.65625 1.34375 -4.046875 1.34375 -4.65625 C 1.34375 -5.28125 1.765625 -5.671875 2.421875 -5.671875 Z M 2.421875 -3 C 3.1875 -3 3.71875 -2.484375 3.71875 -1.75 C 3.71875 -0.984375 3.203125 -0.484375 2.40625 -0.484375 C 1.640625 -0.484375 1.125 -1 1.125 -1.734375 C 1.125 -2.5 1.640625 -3 2.421875 -3 Z M 2.421875 -3 "/>
</symbol>
<symbol overflow="visible" id="glyph0-3">
<path style="stroke:none;" d="M 2.875 -1.546875 L 2.875 0 L 3.65625 0 L 3.65625 -1.546875 L 4.578125 -1.546875 L 4.578125 -2.234375 L 3.65625 -2.234375 L 3.65625 -6.359375 L 3.078125 -6.359375 L 0.25 -2.359375 L 0.25 -1.546875 Z M 2.875 -2.234375 L 0.921875 -2.234375 L 2.875 -5.046875 Z M 2.875 -2.234375 "/>
</symbol>
<symbol overflow="visible" id="glyph0-4">
<path style="stroke:none;" d="M 4.1875 -6.234375 L 0.96875 -6.234375 L 0.5 -2.84375 L 1.21875 -2.84375 C 1.578125 -3.265625 1.875 -3.421875 2.359375 -3.421875 C 3.203125 -3.421875 3.71875 -2.84375 3.71875 -1.921875 C 3.71875 -1.03125 3.203125 -0.484375 2.359375 -0.484375 C 1.6875 -0.484375 1.265625 -0.828125 1.078125 -1.53125 L 0.3125 -1.53125 C 0.40625 -1.015625 0.5 -0.78125 0.6875 -0.546875 C 1.03125 -0.078125 1.671875 0.203125 2.375 0.203125 C 3.640625 0.203125 4.515625 -0.71875 4.515625 -2.03125 C 4.515625 -3.265625 3.6875 -4.109375 2.5 -4.109375 C 2.0625 -4.109375 1.703125 -4 1.34375 -3.734375 L 1.59375 -5.46875 L 4.1875 -5.46875 Z M 4.1875 -6.234375 "/>
</symbol>
<symbol overflow="visible" id="glyph0-5">
<path style="stroke:none;" d="M 2.28125 -4.53125 L 2.28125 0 L 3.046875 0 L 3.046875 -6.359375 L 2.546875 -6.359375 C 2.265625 -5.390625 2.09375 -5.25 0.890625 -5.09375 L 0.890625 -4.53125 Z M 2.28125 -4.53125 "/>
</symbol>
<symbol overflow="visible" id="glyph0-6">
<path style="stroke:none;" d="M 2.421875 -6.359375 C 1.84375 -6.359375 1.3125 -6.09375 0.984375 -5.671875 C 0.578125 -5.109375 0.375 -4.25 0.375 -3.078125 C 0.375 -0.9375 1.078125 0.203125 2.421875 0.203125 C 3.734375 0.203125 4.453125 -0.9375 4.453125 -3.03125 C 4.453125 -4.25 4.265625 -5.078125 3.859375 -5.671875 C 3.53125 -6.109375 3.015625 -6.359375 2.421875 -6.359375 Z M 2.421875 -5.671875 C 3.25 -5.671875 3.671875 -4.828125 3.671875 -3.09375 C 3.671875 -1.28125 3.265625 -0.4375 2.40625 -0.4375 C 1.578125 -0.4375 1.171875 -1.3125 1.171875 -3.0625 C 1.171875 -4.828125 1.578125 -5.671875 2.421875 -5.671875 Z M 2.421875 -5.671875 "/>
</symbol>
<symbol overflow="visible" id="glyph0-7">
<path style="stroke:none;" d="M 4.453125 -0.765625 L 1.171875 -0.765625 C 1.25 -1.296875 1.53125 -1.625 2.296875 -2.09375 L 3.171875 -2.59375 C 4.046875 -3.0625 4.5 -3.71875 4.5 -4.5 C 4.5 -5.03125 4.28125 -5.53125 3.921875 -5.875 C 3.546875 -6.203125 3.09375 -6.359375 2.5 -6.359375 C 1.703125 -6.359375 1.125 -6.078125 0.78125 -5.53125 C 0.546875 -5.203125 0.453125 -4.796875 0.4375 -4.15625 L 1.21875 -4.15625 C 1.234375 -4.59375 1.296875 -4.84375 1.40625 -5.0625 C 1.59375 -5.4375 2 -5.6875 2.46875 -5.6875 C 3.171875 -5.6875 3.703125 -5.171875 3.703125 -4.484375 C 3.703125 -3.96875 3.40625 -3.53125 2.859375 -3.21875 L 2.046875 -2.75 C 0.75 -2 0.375 -1.40625 0.296875 -0.015625 L 4.453125 -0.015625 Z M 4.453125 -0.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph0-8">
<path style="stroke:none;" d="M 1.9375 -2.921875 L 2.359375 -2.921875 C 3.203125 -2.921875 3.65625 -2.53125 3.65625 -1.765625 C 3.65625 -0.96875 3.171875 -0.484375 2.375 -0.484375 C 1.515625 -0.484375 1.109375 -0.921875 1.0625 -1.84375 L 0.28125 -1.84375 C 0.3125 -1.34375 0.40625 -1 0.546875 -0.71875 C 0.875 -0.109375 1.484375 0.203125 2.34375 0.203125 C 3.625 0.203125 4.453125 -0.578125 4.453125 -1.78125 C 4.453125 -2.59375 4.140625 -3.03125 3.390625 -3.296875 C 3.96875 -3.53125 4.265625 -3.96875 4.265625 -4.609375 C 4.265625 -5.703125 3.546875 -6.359375 2.359375 -6.359375 C 1.109375 -6.359375 0.4375 -5.65625 0.40625 -4.3125 L 1.1875 -4.3125 C 1.203125 -4.703125 1.234375 -4.921875 1.328125 -5.109375 C 1.5 -5.46875 1.890625 -5.6875 2.375 -5.6875 C 3.0625 -5.6875 3.46875 -5.265625 3.46875 -4.578125 C 3.46875 -4.140625 3.3125 -3.859375 2.96875 -3.71875 C 2.765625 -3.625 2.484375 -3.59375 1.9375 -3.578125 Z M 1.9375 -2.921875 "/>
</symbol>
<symbol overflow="visible" id="glyph1-0">
<path style="stroke:none;" d=""/>
</symbol>
<symbol overflow="visible" id="glyph1-1">
<path style="stroke:none;" d="M 5.1875 -3.828125 C 5.140625 -4.390625 5.015625 -4.75 4.796875 -5.078125 C 4.40625 -5.609375 3.703125 -5.921875 2.90625 -5.921875 C 1.34375 -5.921875 0.34375 -4.703125 0.34375 -2.78125 C 0.34375 -0.921875 1.328125 0.25 2.890625 0.25 C 4.265625 0.25 5.140625 -0.578125 5.25 -1.984375 L 4.328125 -1.984375 C 4.171875 -1.0625 3.703125 -0.59375 2.921875 -0.59375 C 1.90625 -0.59375 1.296875 -1.421875 1.296875 -2.78125 C 1.296875 -4.21875 1.890625 -5.078125 2.890625 -5.078125 C 3.65625 -5.078125 4.140625 -4.625 4.25 -3.828125 Z M 5.1875 -3.828125 "/>
</symbol>
<symbol overflow="visible" id="glyph1-2">
<path style="stroke:none;" d="M 2.796875 -5.765625 L 1.84375 -5.765625 L 1.84375 -7.34375 L 0.9375 -7.34375 L 0.9375 -5.765625 L 0.15625 -5.765625 L 0.15625 -5.015625 L 0.9375 -5.015625 L 0.9375 -0.65625 C 0.9375 -0.0625 1.328125 0.25 2.046875 0.25 C 2.28125 0.25 2.484375 0.234375 2.796875 0.171875 L 2.796875 -0.59375 C 2.65625 -0.5625 2.546875 -0.546875 2.359375 -0.546875 C 1.953125 -0.546875 1.84375 -0.65625 1.84375 -1.0625 L 1.84375 -5.015625 L 2.796875 -5.015625 Z M 2.796875 -5.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-3">
<path style="stroke:none;" d="M 4.265625 -5.765625 L 2.671875 -1.28125 L 1.203125 -5.765625 L 0.21875 -5.765625 L 2.171875 0.015625 L 1.8125 0.9375 C 1.65625 1.34375 1.46875 1.5 1.078125 1.5 C 0.921875 1.5 0.796875 1.46875 0.59375 1.4375 L 0.59375 2.25 C 0.78125 2.359375 0.96875 2.390625 1.203125 2.390625 C 1.5 2.390625 1.828125 2.296875 2.0625 2.125 C 2.359375 1.90625 2.515625 1.671875 2.6875 1.203125 L 5.265625 -5.765625 Z M 4.265625 -5.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-4">
<path style="stroke:none;" d="M 2.84375 -5.765625 L 1.875 -5.765625 L 1.875 -6.671875 C 1.875 -7.046875 2.09375 -7.25 2.515625 -7.25 C 2.59375 -7.25 2.625 -7.25 2.84375 -7.234375 L 2.84375 -8 C 2.625 -8.046875 2.515625 -8.046875 2.328125 -8.046875 C 1.46875 -8.046875 0.96875 -7.5625 0.96875 -6.75 L 0.96875 -5.765625 L 0.203125 -5.765625 L 0.203125 -5.015625 L 0.96875 -5.015625 L 0.96875 0 L 1.875 0 L 1.875 -5.015625 L 2.84375 -5.015625 Z M 2.84375 -5.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-5">
<path style="stroke:none;" d="M 5.890625 -0.53125 C 5.78125 -0.515625 5.734375 -0.515625 5.6875 -0.515625 C 5.375 -0.515625 5.1875 -0.6875 5.1875 -0.96875 L 5.1875 -4.359375 C 5.1875 -5.375 4.4375 -5.921875 3.03125 -5.921875 C 2.171875 -5.921875 1.5 -5.6875 1.109375 -5.265625 C 0.84375 -4.96875 0.734375 -4.625 0.71875 -4.0625 L 1.640625 -4.0625 C 1.71875 -4.765625 2.140625 -5.078125 2.984375 -5.078125 C 3.828125 -5.078125 4.28125 -4.78125 4.28125 -4.21875 L 4.28125 -3.984375 C 4.265625 -3.59375 4.0625 -3.4375 3.328125 -3.34375 C 2.03125 -3.171875 1.828125 -3.140625 1.46875 -2.984375 C 0.796875 -2.703125 0.46875 -2.203125 0.46875 -1.453125 C 0.46875 -0.40625 1.1875 0.25 2.359375 0.25 C 3.078125 0.25 3.65625 0 4.3125 -0.59375 C 4.375 0 4.65625 0.25 5.265625 0.25 C 5.453125 0.25 5.578125 0.234375 5.890625 0.15625 Z M 4.28125 -1.8125 C 4.28125 -1.5 4.1875 -1.3125 3.921875 -1.0625 C 3.546875 -0.71875 3.09375 -0.546875 2.546875 -0.546875 C 1.84375 -0.546875 1.421875 -0.890625 1.421875 -1.46875 C 1.421875 -2.078125 1.8125 -2.390625 2.8125 -2.53125 C 3.78125 -2.65625 3.96875 -2.703125 4.28125 -2.84375 Z M 4.28125 -1.8125 "/>
</symbol>
<symbol overflow="visible" id="glyph1-6">
<path style="stroke:none;" d="M 2.984375 -5.921875 C 1.359375 -5.921875 0.390625 -4.78125 0.390625 -2.84375 C 0.390625 -0.890625 1.359375 0.25 3 0.25 C 4.625 0.25 5.609375 -0.90625 5.609375 -2.796875 C 5.609375 -4.796875 4.65625 -5.921875 2.984375 -5.921875 Z M 3 -5.078125 C 4.03125 -5.078125 4.65625 -4.234375 4.65625 -2.8125 C 4.65625 -1.4375 4.015625 -0.59375 3 -0.59375 C 1.984375 -0.59375 1.359375 -1.4375 1.359375 -2.84375 C 1.359375 -4.234375 1.984375 -5.078125 3 -5.078125 Z M 3 -5.078125 "/>
</symbol>
<symbol overflow="visible" id="glyph1-7">
<path style="stroke:none;" d="M 0.765625 -5.765625 L 0.765625 0 L 1.6875 0 L 1.6875 -2.984375 C 1.6875 -4.375 2.265625 -5 3.53125 -4.96875 L 3.53125 -5.890625 C 3.375 -5.921875 3.28125 -5.921875 3.171875 -5.921875 C 2.578125 -5.921875 2.140625 -5.578125 1.609375 -4.71875 L 1.609375 -5.765625 Z M 0.765625 -5.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-8">
<path style="stroke:none;" d="M 2.59375 -8.015625 C 1.5 -6.578125 0.796875 -4.578125 0.796875 -2.84375 C 0.796875 -1.109375 1.5 0.890625 2.59375 2.328125 L 3.203125 2.328125 C 2.234375 0.765625 1.6875 -1.09375 1.6875 -2.84375 C 1.6875 -4.59375 2.234375 -6.453125 3.203125 -8.015625 Z M 2.59375 -8.015625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-9">
<path style="stroke:none;" d="M 1.671875 -8.015625 L 0.75 -8.015625 L 0.75 0 L 1.671875 0 Z M 1.671875 -8.015625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-10">
<path style="stroke:none;" d="M 1.015625 2.328125 C 2.125 0.890625 2.8125 -1.109375 2.8125 -2.84375 C 2.8125 -4.578125 2.125 -6.578125 1.015625 -8.015625 L 0.421875 -8.015625 C 1.390625 -6.453125 1.921875 -4.59375 1.921875 -2.84375 C 1.921875 -1.09375 1.390625 0.765625 0.421875 2.328125 Z M 1.015625 2.328125 "/>
</symbol>
<symbol overflow="visible" id="glyph2-0">
<path style="stroke:none;" d=""/>
</symbol>
<symbol overflow="visible" id="glyph2-1">
<path style="stroke:none;" d="M -8.015625 -0.765625 L 0 -0.765625 L 0 -1.6875 L -3.171875 -1.6875 C -4.359375 -1.6875 -5.125 -2.296875 -5.125 -3.25 C -5.125 -3.546875 -5.03125 -3.84375 -4.875 -4.0625 C -4.671875 -4.328125 -4.40625 -4.4375 -4 -4.4375 L 0 -4.4375 L 0 -5.34375 L -4.359375 -5.34375 C -5.328125 -5.34375 -5.921875 -4.65625 -5.921875 -3.53125 C -5.921875 -2.71875 -5.671875 -2.21875 -4.96875 -1.6875 L -8.015625 -1.6875 Z M -8.015625 -0.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph2-2">
<path style="stroke:none;" d="M 0 -6.09375 L -5.765625 -7.78125 L -5.765625 -6.75 L -1.28125 -5.609375 L -5.765625 -4.484375 L -5.765625 -3.359375 L -1.28125 -2.25 L -5.765625 -1.078125 L -5.765625 -0.0625 L 0 -1.734375 L 0 -2.765625 L -4.515625 -3.890625 L 0 -5.046875 Z M 0 -6.09375 "/>
</symbol>
<symbol overflow="visible" id="glyph2-3">
<path style="stroke:none;" d="M -5.765625 -4.265625 L -1.28125 -2.671875 L -5.765625 -1.203125 L -5.765625 -0.21875 L 0.015625 -2.171875 L 0.9375 -1.8125 C 1.34375 -1.65625 1.5 -1.46875 1.5 -1.078125 C 1.5 -0.921875 1.46875 -0.796875 1.4375 -0.59375 L 2.25 -0.59375 C 2.359375 -0.78125 2.390625 -0.96875 2.390625 -1.203125 C 2.390625 -1.5 2.296875 -1.828125 2.125 -2.0625 C 1.90625 -2.359375 1.671875 -2.515625 1.203125 -2.6875 L -5.765625 -5.265625 Z M -5.765625 -4.265625 "/>
</symbol>
</g>
<clipPath id="clip1">
  <path d="M 35.152344 23.246094 L 229.136719 23.246094 L 229.136719 234.925781 L 35.152344 234.925781 Z M 35.152344 23.246094 "/>
</clipPath>
<clipPath id="clip2">
  <path d="M 35.152344 203 L 229.136719 203 L 229.136719 205 L 35.152344 205 Z M 35.152344 203 "/>
</clipPath>
<clipPath id="clip3">
  <path d="M 35.152344 150 L 229.136719 150 L 229.136719 152 L 35.152344 152 Z M 35.152344 150 "/>
</clipPath>
<clipPath id="clip4">
  <path d="M 35.152344 97 L 229.136719 97 L 229.136719 98 L 35.152344 98 Z M 35.152344 97 "/>
</clipPath>
<clipPath id="clip5">
  <path d="M 35.152344 44 L 229.136719 44 L 229.136719 45 L 35.152344 45 Z M 35.152344 44 "/>
</clipPath>
<clipPath id="clip6">
  <path d="M 67 23.246094 L 68 23.246094 L 68 234.925781 L 67 234.925781 Z M 67 23.246094 "/>
</clipPath>
<clipPath id="clip7">
  <path d="M 101 23.246094 L 102 23.246094 L 102 234.925781 L 101 234.925781 Z M 101 23.246094 "/>
</clipPath>
<clipPath id="clip8">
  <path d="M 135 23.246094 L 136 23.246094 L 136 234.925781 L 135 234.925781 Z M 135 23.246094 "/>
</clipPath>
<clipPath id="clip9">
  <path d="M 169 23.246094 L 170 23.246094 L 170 234.925781 L 169 234.925781 Z M 169 23.246094 "/>
</clipPath>
<clipPath id="clip10">
  <path d="M 203 23.246094 L 204 23.246094 L 204 234.925781 L 203 234.925781 Z M 203 23.246094 "/>
</clipPath>
<clipPath id="clip11">
  <path d="M 35.152344 230 L 229.136719 230 L 229.136719 232 L 35.152344 232 Z M 35.152344 230 "/>
</clipPath>
<clipPath id="clip12">
  <path d="M 35.152344 177 L 229.136719 177 L 229.136719 179 L 35.152344 179 Z M 35.152344 177 "/>
</clipPath>
<clipPath id="clip13">
  <path d="M 35.152344 123 L 229.136719 123 L 229.136719 125 L 35.152344 125 Z M 35.152344 123 "/>
</clipPath>
<clipPath id="clip14">
  <path d="M 35.152344 70 L 229.136719 70 L 229.136719 72 L 35.152344 72 Z M 35.152344 70 "/>
</clipPath>
<clipPath id="clip15">
  <path d="M 50 23.246094 L 52 23.246094 L 52 234.925781 L 50 234.925781 Z M 50 23.246094 "/>
</clipPath>
<clipPath id="clip16">
  <path d="M 84 23.246094 L 86 23.246094 L 86 234.925781 L 84 234.925781 Z M 84 23.246094 "/>
</clipPath>
<clipPath id="clip17">
  <path d="M 118 23.246094 L 120 23.246094 L 120 234.925781 L 118 234.925781 Z M 118 23.246094 "/>
</clipPath>
<clipPath id="clip18">
  <path d="M 151 23.246094 L 154 23.246094 L 154 234.925781 L 151 234.925781 Z M 151 23.246094 "/>
</clipPath>
<clipPath id="clip19">
  <path d="M 185 23.246094 L 187 23.246094 L 187 234.925781 L 185 234.925781 Z M 185 23.246094 "/>
</clipPath>
<clipPath id="clip20">
  <path d="M 219 23.246094 L 221 23.246094 L 221 234.925781 L 219 234.925781 Z M 219 23.246094 "/>
</clipPath>
<clipPath id="clip21">
  <path d="M 35.152344 258.171875 L 229.136719 258.171875 L 229.136719 469.851562 L 35.152344 469.851562 Z M 35.152344 258.171875 "/>
</clipPath>
<clipPath id="clip22">
  <path d="M 35.152344 438 L 229.136719 438 L 229.136719 440 L 35.152344 440 Z M 35.152344 438 "/>
</clipPath>
<clipPath id="clip23">
  <path d="M 35.152344 385 L 229.136719 385 L 229.136719 387 L 35.152344 387 Z M 35.152344 385 "/>
</clipPath>
<clipPath id="clip24">
  <path d="M 35.152344 332 L 229.136719 332 L 229.136719 333 L 35.152344 333 Z M 35.152344 332 "/>
</clipPath>
<clipPath id="clip25">
  <path d="M 35.152344 278 L 229.136719 278 L 229.136719 280 L 35.152344 280 Z M 35.152344 278 "/>
</clipPath>
<clipPath id="clip26">
  <path d="M 67 258.171875 L 68 258.171875 L 68 469.851562 L 67 469.851562 Z M 67 258.171875 "/>
</clipPath>
<clipPath id="clip27">
  <path d="M 101 258.171875 L 102 258.171875 L 102 469.851562 L 101 469.851562 Z M 101 258.171875 "/>
</clipPath>
<clipPath id="clip28">
  <path d="M 135 258.171875 L 136 258.171875 L 136 469.851562 L 135 469.851562 Z M 135 258.171875 "/>
</clipPath>
<clipPath id="clip29">
  <path d="M 169 258.171875 L 170 258.171875 L 170 469.851562 L 169 469.851562 Z M 169 258.171875 "/>
</clipPath>
<clipPath id="clip30">
  <path d="M 203 258.171875 L 204 258.171875 L 204 469.851562 L 203 469.851562 Z M 203 258.171875 "/>
</clipPath>
<clipPath id="clip31">
  <path d="M 35.152344 465 L 229.136719 465 L 229.136719 467 L 35.152344 467 Z M 35.152344 465 "/>
</clipPath>
<clipPath id="clip32">
  <path d="M 35.152344 411 L 229.136719 411 L 229.136719 414 L 35.152344 414 Z M 35.152344 411 "/>
</clipPath>
<clipPath id="clip33">
  <path d="M 35.152344 358 L 229.136719 358 L 229.136719 360 L 35.152344 360 Z M 35.152344 358 "/>
</clipPath>
<clipPath id="clip34">
  <path d="M 35.152344 305 L 229.136719 305 L 229.136719 307 L 35.152344 307 Z M 35.152344 305 "/>
</clipPath>
<clipPath id="clip35">
  <path d="M 50 258.171875 L 52 258.171875 L 52 469.851562 L 50 469.851562 Z M 50 258.171875 "/>
</clipPath>
<clipPath id="clip36">
  <path d="M 84 258.171875 L 86 258.171875 L 86 469.851562 L 84 469.851562 Z M 84 258.171875 "/>
</clipPath>
<clipPath id="clip37">
  <path d="M 118 258.171875 L 120 258.171875 L 120 469.851562 L 118 469.851562 Z M 118 258.171875 "/>
</clipPath>
<clipPath id="clip38">
  <path d="M 151 258.171875 L 154 258.171875 L 154 469.851562 L 151 469.851562 Z M 151 258.171875 "/>
</clipPath>
<clipPath id="clip39">
  <path d="M 185 258.171875 L 187 258.171875 L 187 469.851562 L 185 469.851562 Z M 185 258.171875 "/>
</clipPath>
<clipPath id="clip40">
  <path d="M 219 258.171875 L 221 258.171875 L 221 469.851562 L 219 469.851562 Z M 219 258.171875 "/>
</clipPath>
<clipPath id="clip41">
  <path d="M 234.617188 23.246094 L 428.601562 23.246094 L 428.601562 234.925781 L 234.617188 234.925781 Z M 234.617188 23.246094 "/>
</clipPath>
<clipPath id="clip42">
  <path d="M 234.617188 203 L 428.601562 203 L 428.601562 205 L 234.617188 205 Z M 234.617188 203 "/>
</clipPath>
<clipPath id="clip43">
  <path d="M 234.617188 150 L 428.601562 150 L 428.601562 152 L 234.617188 152 Z M 234.617188 150 "/>
</clipPath>
<clipPath id="clip44">
  <path d="M 234.617188 97 L 428.601562 97 L 428.601562 98 L 234.617188 98 Z M 234.617188 97 "/>
</clipPath>
<clipPath id="clip45">
  <path d="M 234.617188 44 L 428.601562 44 L 428.601562 45 L 234.617188 45 Z M 234.617188 44 "/>
</clipPath>
<clipPath id="clip46">
  <path d="M 266 23.246094 L 268 23.246094 L 268 234.925781 L 266 234.925781 Z M 266 23.246094 "/>
</clipPath>
<clipPath id="clip47">
  <path d="M 300 23.246094 L 302 23.246094 L 302 234.925781 L 300 234.925781 Z M 300 23.246094 "/>
</clipPath>
<clipPath id="clip48">
  <path d="M 334 23.246094 L 336 23.246094 L 336 234.925781 L 334 234.925781 Z M 334 23.246094 "/>
</clipPath>
<clipPath id="clip49">
  <path d="M 368 23.246094 L 370 23.246094 L 370 234.925781 L 368 234.925781 Z M 368 23.246094 "/>
</clipPath>
<clipPath id="clip50">
  <path d="M 402 23.246094 L 404 23.246094 L 404 234.925781 L 402 234.925781 Z M 402 23.246094 "/>
</clipPath>
<clipPath id="clip51">
  <path d="M 234.617188 230 L 428.601562 230 L 428.601562 232 L 234.617188 232 Z M 234.617188 230 "/>
</clipPath>
<clipPath id="clip52">
  <path d="M 234.617188 177 L 428.601562 177 L 428.601562 179 L 234.617188 179 Z M 234.617188 177 "/>
</clipPath>
<clipPath id="clip53">
  <path d="M 234.617188 123 L 428.601562 123 L 428.601562 125 L 234.617188 125 Z M 234.617188 123 "/>
</clipPath>
<clipPath id="clip54">
  <path d="M 234.617188 70 L 428.601562 70 L 428.601562 72 L 234.617188 72 Z M 234.617188 70 "/>
</clipPath>
<clipPath id="clip55">
  <path d="M 249 23.246094 L 251 23.246094 L 251 234.925781 L 249 234.925781 Z M 249 23.246094 "/>
</clipPath>
<clipPath id="clip56">
  <path d="M 283 23.246094 L 285 23.246094 L 285 234.925781 L 283 234.925781 Z M 283 23.246094 "/>
</clipPath>
<clipPath id="clip57">
  <path d="M 317 23.246094 L 319 23.246094 L 319 234.925781 L 317 234.925781 Z M 317 23.246094 "/>
</clipPath>
<clipPath id="clip58">
  <path d="M 351 23.246094 L 353 23.246094 L 353 234.925781 L 351 234.925781 Z M 351 23.246094 "/>
</clipPath>
<clipPath id="clip59">
  <path d="M 385 23.246094 L 387 23.246094 L 387 234.925781 L 385 234.925781 Z M 385 23.246094 "/>
</clipPath>
<clipPath id="clip60">
  <path d="M 419 23.246094 L 421 23.246094 L 421 234.925781 L 419 234.925781 Z M 419 23.246094 "/>
</clipPath>
<clipPath id="clip61">
  <path d="M 234.617188 258.171875 L 428.601562 258.171875 L 428.601562 469.851562 L 234.617188 469.851562 Z M 234.617188 258.171875 "/>
</clipPath>
<clipPath id="clip62">
  <path d="M 234.617188 438 L 428.601562 438 L 428.601562 440 L 234.617188 440 Z M 234.617188 438 "/>
</clipPath>
<clipPath id="clip63">
  <path d="M 234.617188 385 L 428.601562 385 L 428.601562 387 L 234.617188 387 Z M 234.617188 385 "/>
</clipPath>
<clipPath id="clip64">
  <path d="M 234.617188 332 L 428.601562 332 L 428.601562 333 L 234.617188 333 Z M 234.617188 332 "/>
</clipPath>
<clipPath id="clip65">
  <path d="M 234.617188 278 L 428.601562 278 L 428.601562 280 L 234.617188 280 Z M 234.617188 278 "/>
</clipPath>
<clipPath id="clip66">
  <path d="M 266 258.171875 L 268 258.171875 L 268 469.851562 L 266 469.851562 Z M 266 258.171875 "/>
</clipPath>
<clipPath id="clip67">
  <path d="M 300 258.171875 L 302 258.171875 L 302 469.851562 L 300 469.851562 Z M 300 258.171875 "/>
</clipPath>
<clipPath id="clip68">
  <path d="M 334 258.171875 L 336 258.171875 L 336 469.851562 L 334 469.851562 Z M 334 258.171875 "/>
</clipPath>
<clipPath id="clip69">
  <path d="M 368 258.171875 L 370 258.171875 L 370 469.851562 L 368 469.851562 Z M 368 258.171875 "/>
</clipPath>
<clipPath id="clip70">
  <path d="M 402 258.171875 L 404 258.171875 L 404 469.851562 L 402 469.851562 Z M 402 258.171875 "/>
</clipPath>
<clipPath id="clip71">
  <path d="M 234.617188 465 L 428.601562 465 L 428.601562 467 L 234.617188 467 Z M 234.617188 465 "/>
</clipPath>
<clipPath id="clip72">
  <path d="M 234.617188 411 L 428.601562 411 L 428.601562 414 L 234.617188 414 Z M 234.617188 411 "/>
</clipPath>
<clipPath id="clip73">
  <path d="M 234.617188 358 L 428.601562 358 L 428.601562 360 L 234.617188 360 Z M 234.617188 358 "/>
</clipPath>
<clipPath id="clip74">
  <path d="M 234.617188 305 L 428.601562 305 L 428.601562 307 L 234.617188 307 Z M 234.617188 305 "/>
</clipPath>
<clipPath id="clip75">
  <path d="M 249 258.171875 L 251 258.171875 L 251 469.851562 L 249 469.851562 Z M 249 258.171875 "/>
</clipPath>
<clipPath id="clip76">
  <path d="M 283 258.171875 L 285 258.171875 L 285 469.851562 L 283 469.851562 Z M 283 258.171875 "/>
</clipPath>
<clipPath id="clip77">
  <path d="M 317 258.171875 L 319 258.171875 L 319 469.851562 L 317 469.851562 Z M 317 258.171875 "/>
</clipPath>
<clipPath id="clip78">
  <path d="M 351 258.171875 L 353 258.171875 L 353 469.851562 L 351 469.851562 Z M 351 258.171875 "/>
</clipPath>
<clipPath id="clip79">
  <path d="M 385 258.171875 L 387 258.171875 L 387 469.851562 L 385 469.851562 Z M 385 258.171875 "/>
</clipPath>
<clipPath id="clip80">
  <path d="M 419 258.171875 L 421 258.171875 L 421 469.851562 L 419 469.851562 Z M 419 258.171875 "/>
</clipPath>
<clipPath id="clip81">
  <path d="M 35.152344 240.402344 L 229.136719 240.402344 L 229.136719 258.167969 L 35.152344 258.167969 Z M 35.152344 240.402344 "/>
</clipPath>
<clipPath id="clip82">
  <path d="M 234.617188 240.402344 L 428.601562 240.402344 L 428.601562 258.167969 L 234.617188 258.167969 Z M 234.617188 240.402344 "/>
</clipPath>
<clipPath id="clip83">
  <path d="M 35.152344 5.480469 L 229.136719 5.480469 L 229.136719 23.246094 L 35.152344 23.246094 Z M 35.152344 5.480469 "/>
</clipPath>
<clipPath id="clip84">
  <path d="M 234.617188 5.480469 L 428.601562 5.480469 L 428.601562 23.246094 L 234.617188 23.246094 Z M 234.617188 5.480469 "/>
</clipPath>
</defs>
<g id="surface929">
<rect x="0" y="0" width="504" height="504" style="fill:rgb(100%,100%,100%);fill-opacity:1;stroke:none;"/>
<rect x="0" y="0" width="504" height="504" style="fill:rgb(100%,100%,100%);fill-opacity:1;stroke:none;"/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 0 504 L 504 504 L 504 0 L 0 0 Z M 0 504 "/>
<g clip-path="url(#clip1)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(92.156863%,92.156863%,92.156863%);fill-opacity:1;" d="M 35.152344 234.925781 L 229.136719 234.925781 L 229.136719 23.246094 L 35.152344 23.246094 Z M 35.152344 234.925781 "/>
</g>
<g clip-path="url(#clip2)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 204.210938 L 229.136719 204.210938 "/>
</g>
<g clip-path="url(#clip3)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 150.910156 L 229.136719 150.910156 "/>
</g>
<g clip-path="url(#clip4)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 97.613281 L 229.136719 97.613281 "/>
</g>
<g clip-path="url(#clip5)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 44.3125 L 229.136719 44.3125 "/>
</g>
<g clip-path="url(#clip6)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 67.707031 234.925781 L 67.707031 23.246094 "/>
</g>
<g clip-path="url(#clip7)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 101.621094 234.925781 L 101.621094 23.246094 "/>
</g>
<g clip-path="url(#clip8)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 135.535156 234.925781 L 135.535156 23.246094 "/>
</g>
<g clip-path="url(#clip9)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 169.449219 234.925781 L 169.449219 23.246094 "/>
</g>
<g clip-path="url(#clip10)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 203.363281 234.925781 L 203.363281 23.246094 "/>
</g>
<g clip-path="url(#clip11)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 230.859375 L 229.136719 230.859375 "/>
</g>
<g clip-path="url(#clip12)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 177.558594 L 229.136719 177.558594 "/>
</g>
<g clip-path="url(#clip13)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 124.261719 L 229.136719 124.261719 "/>
</g>
<g clip-path="url(#clip14)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 70.960938 L 229.136719 70.960938 "/>
</g>
<g clip-path="url(#clip15)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 50.75 234.925781 L 50.75 23.246094 "/>
</g>
<g clip-path="url(#clip16)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 84.664062 234.925781 L 84.664062 23.246094 "/>
</g>
<g clip-path="url(#clip17)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 118.578125 234.925781 L 118.578125 23.246094 "/>
</g>
<g clip-path="url(#clip18)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 152.492188 234.925781 L 152.492188 23.246094 "/>
</g>
<g clip-path="url(#clip19)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 186.40625 234.925781 L 186.40625 23.246094 "/>
</g>
<g clip-path="url(#clip20)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 220.320312 234.925781 L 220.320312 23.246094 "/>
</g>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 129.589844 C 106.96875 132.199219 103.058594 132.199219 103.058594 129.589844 C 103.058594 126.984375 106.96875 126.984375 106.96875 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 118.933594 C 120.53125 121.539062 116.625 121.539062 116.625 118.933594 C 116.625 116.324219 120.53125 116.324219 120.53125 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 124.261719 C 127.316406 126.867188 123.40625 126.867188 123.40625 124.261719 C 123.40625 121.65625 127.316406 121.65625 127.316406 124.261719 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 145.582031 C 106.96875 148.1875 103.058594 148.1875 103.058594 145.582031 C 103.058594 142.976562 106.96875 142.976562 106.96875 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 150.910156 C 93.402344 153.515625 89.492188 153.515625 89.492188 150.910156 C 89.492188 148.304688 93.402344 148.304688 93.402344 150.910156 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 134.921875 C 120.53125 137.527344 116.625 137.527344 116.625 134.921875 C 116.625 132.316406 120.53125 132.316406 120.53125 134.921875 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 140.25 C 113.75 142.859375 109.839844 142.859375 109.839844 140.25 C 109.839844 137.644531 113.75 137.644531 113.75 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 140.25 C 113.75 142.859375 109.839844 142.859375 109.839844 140.25 C 109.839844 137.644531 113.75 137.644531 113.75 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 134.097656 124.261719 C 134.097656 126.867188 130.1875 126.867188 130.1875 124.261719 C 130.1875 121.65625 134.097656 121.65625 134.097656 124.261719 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 156.242188 C 106.96875 158.847656 103.058594 158.847656 103.058594 156.242188 C 103.058594 153.632812 106.96875 153.632812 106.96875 156.242188 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 174.796875 108.273438 C 174.796875 110.878906 170.886719 110.878906 170.886719 108.273438 C 170.886719 105.664062 174.796875 105.664062 174.796875 108.273438 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 147.664062 113.601562 C 147.664062 116.207031 143.753906 116.207031 143.753906 113.601562 C 143.753906 110.996094 147.664062 110.996094 147.664062 113.601562 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 154.445312 113.601562 C 154.445312 116.207031 150.539062 116.207031 150.539062 113.601562 C 150.539062 110.996094 154.445312 110.996094 154.445312 113.601562 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 140.882812 129.589844 C 140.882812 132.199219 136.972656 132.199219 136.972656 129.589844 C 136.972656 126.984375 140.882812 126.984375 140.882812 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 147.664062 113.601562 C 147.664062 116.207031 143.753906 116.207031 143.753906 113.601562 C 143.753906 110.996094 147.664062 110.996094 147.664062 113.601562 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 161.230469 102.941406 C 161.230469 105.546875 157.320312 105.546875 157.320312 102.941406 C 157.320312 100.335938 161.230469 100.335938 161.230469 102.941406 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 154.445312 92.28125 C 154.445312 94.890625 150.539062 94.890625 150.539062 92.28125 C 150.539062 89.675781 154.445312 89.675781 154.445312 92.28125 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 147.664062 92.28125 C 147.664062 94.890625 143.753906 94.890625 143.753906 92.28125 C 143.753906 89.675781 147.664062 89.675781 147.664062 92.28125 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 145.582031 C 106.96875 148.1875 103.058594 148.1875 103.058594 145.582031 C 103.058594 142.976562 106.96875 142.976562 106.96875 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 140.25 C 106.96875 142.859375 103.058594 142.859375 103.058594 140.25 C 103.058594 137.644531 106.96875 137.644531 106.96875 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 124.261719 C 127.316406 126.867188 123.40625 126.867188 123.40625 124.261719 C 123.40625 121.65625 127.316406 121.65625 127.316406 124.261719 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 118.933594 C 127.316406 121.539062 123.40625 121.539062 123.40625 118.933594 C 123.40625 116.324219 127.316406 116.324219 127.316406 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 129.589844 C 113.75 132.199219 109.839844 132.199219 109.839844 129.589844 C 109.839844 126.984375 113.75 126.984375 113.75 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 134.921875 C 120.53125 137.527344 116.625 137.527344 116.625 134.921875 C 116.625 132.316406 120.53125 132.316406 120.53125 134.921875 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 140.25 C 120.53125 142.859375 116.625 142.859375 116.625 140.25 C 116.625 137.644531 120.53125 137.644531 120.53125 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 140.25 C 113.75 142.859375 109.839844 142.859375 109.839844 140.25 C 109.839844 137.644531 113.75 137.644531 113.75 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 140.882812 118.933594 C 140.882812 121.539062 136.972656 121.539062 136.972656 118.933594 C 136.972656 116.324219 140.882812 116.324219 140.882812 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 140.882812 113.601562 C 140.882812 116.207031 136.972656 116.207031 136.972656 113.601562 C 136.972656 110.996094 140.882812 110.996094 140.882812 113.601562 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 150.910156 C 106.96875 153.515625 103.058594 153.515625 103.058594 150.910156 C 103.058594 148.304688 106.96875 148.304688 106.96875 150.910156 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 156.242188 C 106.96875 158.847656 103.058594 158.847656 103.058594 156.242188 C 103.058594 153.632812 106.96875 153.632812 106.96875 156.242188 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 140.25 C 120.53125 142.859375 116.625 142.859375 116.625 140.25 C 116.625 137.644531 120.53125 137.644531 120.53125 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 150.910156 C 113.75 153.515625 109.839844 153.515625 109.839844 150.910156 C 109.839844 148.304688 113.75 148.304688 113.75 150.910156 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 145.582031 C 120.53125 148.1875 116.625 148.1875 116.625 145.582031 C 116.625 142.976562 120.53125 142.976562 120.53125 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 161.570312 C 106.96875 164.175781 103.058594 164.175781 103.058594 161.570312 C 103.058594 158.964844 106.96875 158.964844 106.96875 161.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 145.582031 C 127.316406 148.1875 123.40625 148.1875 123.40625 145.582031 C 123.40625 142.976562 127.316406 142.976562 127.316406 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 150.910156 C 120.53125 153.515625 116.625 153.515625 116.625 150.910156 C 116.625 148.304688 120.53125 148.304688 120.53125 150.910156 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 140.25 C 120.53125 142.859375 116.625 142.859375 116.625 140.25 C 116.625 137.644531 120.53125 137.644531 120.53125 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 150.910156 C 113.75 153.515625 109.839844 153.515625 109.839844 150.910156 C 109.839844 148.304688 113.75 148.304688 113.75 150.910156 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 120.53125 140.25 C 120.53125 142.859375 116.625 142.859375 116.625 140.25 C 116.625 137.644531 120.53125 137.644531 120.53125 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 177.558594 C 86.621094 180.167969 82.710938 180.167969 82.710938 177.558594 C 82.710938 174.953125 86.621094 174.953125 86.621094 177.558594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 177.558594 C 93.402344 180.167969 89.492188 180.167969 89.492188 177.558594 C 89.492188 174.953125 93.402344 174.953125 93.402344 177.558594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 140.25 C 127.316406 142.859375 123.40625 142.859375 123.40625 140.25 C 123.40625 137.644531 127.316406 137.644531 127.316406 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 118.933594 C 127.316406 121.539062 123.40625 121.539062 123.40625 118.933594 C 123.40625 116.324219 127.316406 116.324219 127.316406 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 118.933594 C 127.316406 121.539062 123.40625 121.539062 123.40625 118.933594 C 123.40625 116.324219 127.316406 116.324219 127.316406 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 140.25 C 127.316406 142.859375 123.40625 142.859375 123.40625 140.25 C 123.40625 137.644531 127.316406 137.644531 127.316406 140.25 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 118.933594 C 127.316406 121.539062 123.40625 121.539062 123.40625 118.933594 C 123.40625 116.324219 127.316406 116.324219 127.316406 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 134.097656 118.933594 C 134.097656 121.539062 130.1875 121.539062 130.1875 118.933594 C 130.1875 116.324219 134.097656 116.324219 134.097656 118.933594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 147.664062 124.261719 C 147.664062 126.867188 143.753906 126.867188 143.753906 124.261719 C 143.753906 121.65625 147.664062 121.65625 147.664062 124.261719 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 147.664062 108.273438 C 147.664062 110.878906 143.753906 110.878906 143.753906 108.273438 C 143.753906 105.664062 147.664062 105.664062 147.664062 108.273438 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 161.230469 97.613281 C 161.230469 100.21875 157.320312 100.21875 157.320312 97.613281 C 157.320312 95.003906 161.230469 95.003906 161.230469 97.613281 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 174.796875 86.953125 C 174.796875 89.558594 170.886719 89.558594 170.886719 86.953125 C 170.886719 84.347656 174.796875 84.347656 174.796875 86.953125 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 161.230469 97.613281 C 161.230469 100.21875 157.320312 100.21875 157.320312 97.613281 C 157.320312 95.003906 161.230469 95.003906 161.230469 97.613281 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 177.558594 C 86.621094 180.167969 82.710938 180.167969 82.710938 177.558594 C 82.710938 174.953125 86.621094 174.953125 86.621094 177.558594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 177.558594 C 93.402344 180.167969 89.492188 180.167969 89.492188 177.558594 C 89.492188 174.953125 93.402344 174.953125 93.402344 177.558594 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 166.902344 C 100.183594 169.507812 96.273438 169.507812 96.273438 166.902344 C 96.273438 164.292969 100.183594 164.292969 100.183594 166.902344 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 134.097656 129.589844 C 134.097656 132.199219 130.1875 132.199219 130.1875 129.589844 C 130.1875 126.984375 134.097656 126.984375 134.097656 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 208.707031 49.644531 C 208.707031 52.25 204.800781 52.25 204.800781 49.644531 C 204.800781 47.035156 208.707031 47.035156 208.707031 49.644531 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 134.097656 129.589844 C 134.097656 132.199219 130.1875 132.199219 130.1875 129.589844 C 130.1875 126.984375 134.097656 126.984375 134.097656 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 222.273438 49.644531 C 222.273438 52.25 218.363281 52.25 218.363281 49.644531 C 218.363281 47.035156 222.273438 47.035156 222.273438 49.644531 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 181.578125 65.632812 C 181.578125 68.238281 177.667969 68.238281 177.667969 65.632812 C 177.667969 63.027344 181.578125 63.027344 181.578125 65.632812 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 145.582031 C 113.75 148.1875 109.839844 148.1875 109.839844 145.582031 C 109.839844 142.976562 113.75 142.976562 113.75 145.582031 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 129.589844 C 106.96875 132.199219 103.058594 132.199219 103.058594 129.589844 C 103.058594 126.984375 106.96875 126.984375 106.96875 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 134.921875 C 113.75 137.527344 109.839844 137.527344 109.839844 134.921875 C 109.839844 132.316406 113.75 132.316406 113.75 134.921875 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 127.316406 129.589844 C 127.316406 132.199219 123.40625 132.199219 123.40625 129.589844 C 123.40625 126.984375 127.316406 126.984375 127.316406 129.589844 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 84.664062 165.398438 L 86.382812 163.894531 L 88.097656 162.386719 L 91.535156 159.371094 L 93.25 157.859375 L 94.96875 156.347656 L 96.683594 154.832031 L 98.402344 153.3125 L 100.117188 151.792969 L 103.554688 148.738281 L 105.269531 147.207031 L 106.988281 145.671875 L 108.703125 144.128906 L 110.421875 142.582031 L 112.140625 141.027344 L 113.855469 139.46875 L 115.574219 137.902344 L 117.289062 136.328125 L 119.007812 134.742188 L 120.726562 133.148438 L 122.441406 131.546875 L 124.160156 129.933594 L 125.875 128.3125 L 127.59375 126.679688 L 129.308594 125.035156 L 131.027344 123.382812 L 132.746094 121.71875 L 134.460938 120.050781 L 136.179688 118.371094 L 137.894531 116.683594 L 139.613281 114.988281 L 141.332031 113.289062 L 143.046875 111.582031 L 144.765625 109.871094 L 146.480469 108.152344 L 148.199219 106.433594 L 149.917969 104.710938 L 151.632812 102.980469 L 153.351562 101.25 L 155.066406 99.515625 L 156.785156 97.78125 L 158.503906 96.042969 L 160.21875 94.300781 L 161.9375 92.558594 L 163.652344 90.816406 L 165.371094 89.070312 L 167.085938 87.324219 L 168.804688 85.578125 L 170.523438 83.828125 L 172.238281 82.078125 L 173.957031 80.328125 L 175.671875 78.574219 L 177.390625 76.824219 L 179.109375 75.070312 L 180.824219 73.316406 L 182.542969 71.5625 L 184.257812 69.804688 L 185.976562 68.050781 L 187.695312 66.292969 L 189.410156 64.539062 L 191.128906 62.78125 L 192.84375 61.023438 L 196.28125 57.507812 L 197.996094 55.75 L 199.714844 53.992188 L 201.429688 52.230469 L 203.148438 50.472656 L 204.863281 48.710938 L 206.582031 46.953125 L 208.300781 45.191406 L 210.015625 43.433594 L 211.734375 41.671875 L 213.449219 39.914062 L 216.886719 36.390625 L 218.601562 34.628906 L 220.320312 32.867188 L 220.320312 48.824219 L 218.601562 50.3125 L 216.886719 51.800781 L 215.167969 53.289062 L 213.449219 54.78125 L 211.734375 56.269531 L 210.015625 57.761719 L 208.300781 59.25 L 204.863281 62.234375 L 203.148438 63.722656 L 201.429688 65.214844 L 199.714844 66.707031 L 197.996094 68.199219 L 196.28125 69.691406 L 192.84375 72.675781 L 191.128906 74.167969 L 189.410156 75.664062 L 187.695312 77.15625 L 184.257812 80.148438 L 182.542969 81.640625 L 180.824219 83.136719 L 179.109375 84.636719 L 175.671875 87.628906 L 173.957031 89.128906 L 172.238281 90.628906 L 170.523438 92.128906 L 168.804688 93.632812 L 167.085938 95.132812 L 165.371094 96.636719 L 163.652344 98.144531 L 161.9375 99.648438 L 160.21875 101.160156 L 158.503906 102.667969 L 156.785156 104.179688 L 155.066406 105.695312 L 153.351562 107.210938 L 151.632812 108.730469 L 149.917969 110.253906 L 148.199219 111.78125 L 146.480469 113.3125 L 144.765625 114.84375 L 143.046875 116.382812 L 141.332031 117.929688 L 139.613281 119.476562 L 137.894531 121.035156 L 136.179688 122.597656 L 134.460938 124.167969 L 132.746094 125.75 L 131.027344 127.335938 L 129.308594 128.933594 L 127.59375 130.542969 L 125.875 132.160156 L 124.160156 133.789062 L 122.441406 135.425781 L 120.726562 137.074219 L 119.007812 138.730469 L 117.289062 140.398438 L 115.574219 142.074219 L 113.855469 143.757812 L 112.140625 145.449219 L 110.421875 147.144531 L 108.703125 148.847656 L 106.988281 150.558594 L 105.269531 152.273438 L 103.554688 153.988281 L 101.835938 155.710938 L 100.117188 157.4375 L 98.402344 159.167969 L 96.683594 160.902344 L 94.96875 162.636719 L 93.25 164.371094 L 91.535156 166.113281 L 89.816406 167.851562 L 88.097656 169.597656 L 86.382812 171.339844 L 84.664062 173.085938 Z M 84.664062 165.398438 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 84.664062 169.242188 L 86.382812 167.617188 L 88.097656 165.992188 L 91.535156 162.742188 L 93.25 161.117188 L 94.96875 159.492188 L 96.683594 157.867188 L 98.402344 156.242188 L 100.117188 154.613281 L 103.554688 151.363281 L 105.269531 149.738281 L 106.988281 148.113281 L 108.703125 146.488281 L 112.140625 143.238281 L 113.855469 141.613281 L 115.574219 139.988281 L 117.289062 138.363281 L 120.726562 135.113281 L 122.441406 133.488281 L 124.160156 131.859375 L 125.875 130.234375 L 127.59375 128.609375 L 129.308594 126.984375 L 132.746094 123.734375 L 134.460938 122.109375 L 136.179688 120.484375 L 137.894531 118.859375 L 141.332031 115.609375 L 143.046875 113.984375 L 144.765625 112.359375 L 146.480469 110.734375 L 148.199219 109.105469 L 149.917969 107.480469 L 151.632812 105.855469 L 153.351562 104.230469 L 155.066406 102.605469 L 158.503906 99.355469 L 160.21875 97.730469 L 161.9375 96.105469 L 163.652344 94.480469 L 165.371094 92.855469 L 167.085938 91.230469 L 170.523438 87.980469 L 172.238281 86.351562 L 173.957031 84.726562 L 175.671875 83.101562 L 179.109375 79.851562 L 180.824219 78.226562 L 182.542969 76.601562 L 184.257812 74.976562 L 187.695312 71.726562 L 189.410156 70.101562 L 191.128906 68.476562 L 192.84375 66.851562 L 194.5625 65.222656 L 196.28125 63.597656 L 197.996094 61.972656 L 199.714844 60.347656 L 201.429688 58.722656 L 203.148438 57.097656 L 204.863281 55.472656 L 208.300781 52.222656 L 210.015625 50.597656 L 211.734375 48.972656 L 213.449219 47.347656 L 216.886719 44.097656 L 218.601562 42.46875 L 220.320312 40.84375 "/>
<g clip-path="url(#clip21)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(92.156863%,92.156863%,92.156863%);fill-opacity:1;" d="M 35.152344 469.847656 L 229.136719 469.847656 L 229.136719 258.167969 L 35.152344 258.167969 Z M 35.152344 469.847656 "/>
</g>
<g clip-path="url(#clip22)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 439.132812 L 229.136719 439.132812 "/>
</g>
<g clip-path="url(#clip23)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 385.835938 L 229.136719 385.835938 "/>
</g>
<g clip-path="url(#clip24)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 332.535156 L 229.136719 332.535156 "/>
</g>
<g clip-path="url(#clip25)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 279.238281 L 229.136719 279.238281 "/>
</g>
<g clip-path="url(#clip26)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 67.707031 469.847656 L 67.707031 258.171875 "/>
</g>
<g clip-path="url(#clip27)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 101.621094 469.847656 L 101.621094 258.171875 "/>
</g>
<g clip-path="url(#clip28)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 135.535156 469.847656 L 135.535156 258.171875 "/>
</g>
<g clip-path="url(#clip29)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 169.449219 469.847656 L 169.449219 258.171875 "/>
</g>
<g clip-path="url(#clip30)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 203.363281 469.847656 L 203.363281 258.171875 "/>
</g>
<g clip-path="url(#clip31)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 465.785156 L 229.136719 465.785156 "/>
</g>
<g clip-path="url(#clip32)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 412.484375 L 229.136719 412.484375 "/>
</g>
<g clip-path="url(#clip33)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 359.1875 L 229.136719 359.1875 "/>
</g>
<g clip-path="url(#clip34)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 305.886719 L 229.136719 305.886719 "/>
</g>
<g clip-path="url(#clip35)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 50.75 469.847656 L 50.75 258.171875 "/>
</g>
<g clip-path="url(#clip36)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 84.664062 469.847656 L 84.664062 258.171875 "/>
</g>
<g clip-path="url(#clip37)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 118.578125 469.847656 L 118.578125 258.171875 "/>
</g>
<g clip-path="url(#clip38)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 152.492188 469.847656 L 152.492188 258.171875 "/>
</g>
<g clip-path="url(#clip39)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 186.40625 469.847656 L 186.40625 258.171875 "/>
</g>
<g clip-path="url(#clip40)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 220.320312 469.847656 L 220.320312 258.171875 "/>
</g>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 380.503906 C 93.402344 383.113281 89.492188 383.113281 89.492188 380.503906 C 89.492188 377.898438 93.402344 377.898438 93.402344 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 375.175781 C 106.96875 377.78125 103.058594 377.78125 103.058594 375.175781 C 103.058594 372.570312 106.96875 372.570312 106.96875 375.175781 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 385.835938 C 86.621094 388.441406 82.710938 388.441406 82.710938 385.835938 C 82.710938 383.230469 86.621094 383.230469 86.621094 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 385.835938 C 100.183594 388.441406 96.273438 388.441406 96.273438 385.835938 C 96.273438 383.230469 100.183594 383.230469 100.183594 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 385.835938 C 100.183594 388.441406 96.273438 388.441406 96.273438 385.835938 C 96.273438 383.230469 100.183594 383.230469 100.183594 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 385.835938 C 86.621094 388.441406 82.710938 388.441406 82.710938 385.835938 C 82.710938 383.230469 86.621094 383.230469 86.621094 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 391.164062 C 86.621094 393.773438 82.710938 393.773438 82.710938 391.164062 C 82.710938 388.558594 86.621094 388.558594 86.621094 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 385.835938 C 100.183594 388.441406 96.273438 388.441406 96.273438 385.835938 C 96.273438 383.230469 100.183594 383.230469 100.183594 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 364.515625 C 106.96875 367.121094 103.058594 367.121094 103.058594 364.515625 C 103.058594 361.910156 106.96875 361.910156 106.96875 364.515625 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 380.503906 C 100.183594 383.113281 96.273438 383.113281 96.273438 380.503906 C 96.273438 377.898438 100.183594 377.898438 100.183594 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 401.824219 C 93.402344 404.429688 89.492188 404.429688 89.492188 401.824219 C 89.492188 399.21875 93.402344 399.21875 93.402344 401.824219 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 401.824219 C 93.402344 404.429688 89.492188 404.429688 89.492188 401.824219 C 89.492188 399.21875 93.402344 399.21875 93.402344 401.824219 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 59.488281 428.472656 C 59.488281 431.082031 55.578125 431.082031 55.578125 428.472656 C 55.578125 425.867188 59.488281 425.867188 59.488281 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 401.824219 C 86.621094 404.429688 82.710938 404.429688 82.710938 401.824219 C 82.710938 399.21875 86.621094 399.21875 86.621094 401.824219 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 407.15625 C 86.621094 409.761719 82.710938 409.761719 82.710938 407.15625 C 82.710938 404.546875 86.621094 404.546875 86.621094 407.15625 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 396.496094 C 93.402344 399.101562 89.492188 399.101562 89.492188 396.496094 C 89.492188 393.890625 93.402344 393.890625 93.402344 396.496094 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 396.496094 C 93.402344 399.101562 89.492188 399.101562 89.492188 396.496094 C 89.492188 393.890625 93.402344 393.890625 93.402344 396.496094 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 417.816406 C 86.621094 420.421875 82.710938 420.421875 82.710938 417.816406 C 82.710938 415.207031 86.621094 415.207031 86.621094 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 423.144531 C 79.835938 425.75 75.925781 425.75 75.925781 423.144531 C 75.925781 420.539062 79.835938 420.539062 79.835938 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 73.054688 428.472656 C 73.054688 431.082031 69.144531 431.082031 69.144531 428.472656 C 69.144531 425.867188 73.054688 425.867188 73.054688 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 73.054688 428.472656 C 73.054688 431.082031 69.144531 431.082031 69.144531 428.472656 C 69.144531 425.867188 73.054688 425.867188 73.054688 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 417.816406 C 86.621094 420.421875 82.710938 420.421875 82.710938 417.816406 C 82.710938 415.207031 86.621094 415.207031 86.621094 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 73.054688 417.816406 C 73.054688 420.421875 69.144531 420.421875 69.144531 417.816406 C 69.144531 415.207031 73.054688 415.207031 73.054688 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 385.835938 C 106.96875 388.441406 103.058594 388.441406 103.058594 385.835938 C 103.058594 383.230469 106.96875 383.230469 106.96875 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 380.503906 C 100.183594 383.113281 96.273438 383.113281 96.273438 380.503906 C 96.273438 377.898438 100.183594 377.898438 100.183594 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 391.164062 C 93.402344 393.773438 89.492188 393.773438 89.492188 391.164062 C 89.492188 388.558594 93.402344 388.558594 93.402344 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 369.847656 C 113.75 372.453125 109.839844 372.453125 109.839844 369.847656 C 109.839844 367.238281 113.75 367.238281 113.75 369.847656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 391.164062 C 93.402344 393.773438 89.492188 393.773438 89.492188 391.164062 C 89.492188 388.558594 93.402344 388.558594 93.402344 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 401.824219 C 100.183594 404.429688 96.273438 404.429688 96.273438 401.824219 C 96.273438 399.21875 100.183594 399.21875 100.183594 401.824219 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 417.816406 C 86.621094 420.421875 82.710938 420.421875 82.710938 417.816406 C 82.710938 415.207031 86.621094 415.207031 86.621094 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 412.484375 C 86.621094 415.089844 82.710938 415.089844 82.710938 412.484375 C 82.710938 409.878906 86.621094 409.878906 86.621094 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 73.054688 417.816406 C 73.054688 420.421875 69.144531 420.421875 69.144531 417.816406 C 69.144531 415.207031 73.054688 415.207031 73.054688 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 375.175781 C 113.75 377.78125 109.839844 377.78125 109.839844 375.175781 C 109.839844 372.570312 113.75 372.570312 113.75 375.175781 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 380.503906 C 113.75 383.113281 109.839844 383.113281 109.839844 380.503906 C 109.839844 377.898438 113.75 377.898438 113.75 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 385.835938 C 113.75 388.441406 109.839844 388.441406 109.839844 385.835938 C 109.839844 383.230469 113.75 383.230469 113.75 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 385.835938 C 113.75 388.441406 109.839844 388.441406 109.839844 385.835938 C 109.839844 383.230469 113.75 383.230469 113.75 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 428.472656 C 79.835938 431.082031 75.925781 431.082031 75.925781 428.472656 C 75.925781 425.867188 79.835938 425.867188 79.835938 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 428.472656 C 86.621094 431.082031 82.710938 431.082031 82.710938 428.472656 C 82.710938 425.867188 86.621094 425.867188 86.621094 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 79.835938 412.484375 C 79.835938 415.089844 75.925781 415.089844 75.925781 412.484375 C 75.925781 409.878906 79.835938 409.878906 79.835938 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 380.503906 C 93.402344 383.113281 89.492188 383.113281 89.492188 380.503906 C 89.492188 377.898438 93.402344 377.898438 93.402344 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 375.175781 C 100.183594 377.78125 96.273438 377.78125 96.273438 375.175781 C 96.273438 372.570312 100.183594 372.570312 100.183594 375.175781 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 369.847656 C 106.96875 372.453125 103.058594 372.453125 103.058594 369.847656 C 103.058594 367.238281 106.96875 367.238281 106.96875 369.847656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 417.816406 C 86.621094 420.421875 82.710938 420.421875 82.710938 417.816406 C 82.710938 415.207031 86.621094 415.207031 86.621094 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 428.472656 C 86.621094 431.082031 82.710938 431.082031 82.710938 428.472656 C 82.710938 425.867188 86.621094 425.867188 86.621094 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 412.484375 C 93.402344 415.089844 89.492188 415.089844 89.492188 412.484375 C 89.492188 409.878906 93.402344 409.878906 93.402344 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 113.75 369.847656 C 113.75 372.453125 109.839844 372.453125 109.839844 369.847656 C 109.839844 367.238281 113.75 367.238281 113.75 369.847656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 375.175781 C 106.96875 377.78125 103.058594 377.78125 103.058594 375.175781 C 103.058594 372.570312 106.96875 372.570312 106.96875 375.175781 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 428.472656 C 86.621094 431.082031 82.710938 431.082031 82.710938 428.472656 C 82.710938 425.867188 86.621094 425.867188 86.621094 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 417.816406 C 86.621094 420.421875 82.710938 420.421875 82.710938 417.816406 C 82.710938 415.207031 86.621094 415.207031 86.621094 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.621094 423.144531 C 86.621094 425.75 82.710938 425.75 82.710938 423.144531 C 82.710938 420.539062 86.621094 420.539062 86.621094 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 412.484375 C 93.402344 415.089844 89.492188 415.089844 89.492188 412.484375 C 89.492188 409.878906 93.402344 409.878906 93.402344 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 396.496094 C 93.402344 399.101562 89.492188 399.101562 89.492188 396.496094 C 89.492188 393.890625 93.402344 393.890625 93.402344 396.496094 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 391.164062 C 100.183594 393.773438 96.273438 393.773438 96.273438 391.164062 C 96.273438 388.558594 100.183594 388.558594 100.183594 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 93.402344 380.503906 C 93.402344 383.113281 89.492188 383.113281 89.492188 380.503906 C 89.492188 377.898438 93.402344 377.898438 93.402344 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 106.96875 380.503906 C 106.96875 383.113281 103.058594 383.113281 103.058594 380.503906 C 103.058594 377.898438 106.96875 377.898438 106.96875 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.183594 380.503906 C 100.183594 383.113281 96.273438 383.113281 96.273438 380.503906 C 96.273438 377.898438 100.183594 377.898438 100.183594 380.503906 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 57.535156 440.144531 L 58.21875 439.300781 L 60.28125 436.769531 L 60.96875 435.921875 L 61.65625 435.078125 L 62.339844 434.234375 L 63.027344 433.386719 L 63.714844 432.542969 L 65.089844 430.847656 L 65.777344 430.003906 L 66.464844 429.15625 L 67.148438 428.308594 L 67.835938 427.457031 L 68.523438 426.609375 L 69.210938 425.757812 L 69.898438 424.910156 L 70.585938 424.058594 L 71.269531 423.203125 L 72.644531 421.5 L 74.019531 419.789062 L 74.707031 418.929688 L 75.390625 418.070312 L 76.765625 416.351562 L 78.140625 414.625 L 78.828125 413.757812 L 79.511719 412.890625 L 80.886719 411.148438 L 81.574219 410.273438 L 82.261719 409.394531 L 82.949219 408.511719 L 83.632812 407.628906 L 85.007812 405.847656 L 86.382812 404.050781 L 87.070312 403.140625 L 87.753906 402.230469 L 88.441406 401.3125 L 89.128906 400.390625 L 89.816406 399.460938 L 90.503906 398.523438 L 91.191406 397.578125 L 91.875 396.628906 L 92.5625 395.667969 L 93.25 394.703125 L 93.9375 393.730469 L 94.625 392.75 L 95.3125 391.761719 L 95.996094 390.765625 L 96.683594 389.765625 L 97.371094 388.757812 L 98.058594 387.742188 L 98.746094 386.722656 L 99.433594 385.699219 L 100.117188 384.667969 L 100.804688 383.632812 L 101.492188 382.59375 L 102.179688 381.550781 L 102.867188 380.503906 L 103.554688 379.453125 L 104.238281 378.402344 L 105.613281 376.285156 L 106.988281 374.160156 L 107.675781 373.09375 L 108.359375 372.027344 L 109.046875 370.960938 L 109.734375 369.886719 L 110.421875 368.816406 L 111.796875 366.667969 L 111.796875 375.09375 L 111.109375 375.949219 L 109.046875 378.527344 L 108.359375 379.390625 L 107.675781 380.253906 L 106.300781 381.988281 L 105.613281 382.859375 L 104.238281 384.609375 L 103.554688 385.484375 L 102.867188 386.367188 L 101.492188 388.140625 L 100.804688 389.03125 L 100.117188 389.929688 L 99.433594 390.832031 L 98.746094 391.738281 L 98.058594 392.648438 L 97.371094 393.566406 L 96.683594 394.488281 L 95.996094 395.421875 L 95.3125 396.355469 L 94.625 397.300781 L 93.9375 398.25 L 93.25 399.210938 L 92.5625 400.175781 L 91.875 401.148438 L 91.191406 402.128906 L 90.503906 403.117188 L 89.816406 404.113281 L 89.128906 405.113281 L 88.441406 406.121094 L 87.753906 407.136719 L 87.070312 408.15625 L 86.382812 409.179688 L 85.695312 410.210938 L 85.007812 411.246094 L 84.320312 412.285156 L 83.632812 413.328125 L 82.949219 414.375 L 82.261719 415.425781 L 80.886719 417.535156 L 80.199219 418.59375 L 79.511719 419.65625 L 78.828125 420.71875 L 77.453125 422.851562 L 75.390625 426.0625 L 74.707031 427.136719 L 73.332031 429.285156 L 71.269531 432.519531 L 70.585938 433.597656 L 67.148438 439.007812 L 66.464844 440.089844 L 65.777344 441.175781 L 65.089844 442.257812 L 62.339844 446.601562 L 61.65625 447.6875 L 60.96875 448.773438 L 60.28125 449.863281 L 59.59375 450.949219 L 58.90625 452.039062 L 58.21875 453.125 L 57.535156 454.214844 Z M 57.535156 440.144531 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 57.535156 447.179688 L 58.21875 446.210938 L 60.28125 443.316406 L 60.96875 442.347656 L 61.65625 441.382812 L 62.339844 440.417969 L 63.027344 439.453125 L 63.714844 438.484375 L 65.777344 435.589844 L 66.464844 434.621094 L 67.148438 433.65625 L 69.210938 430.761719 L 69.898438 429.792969 L 70.585938 428.828125 L 71.269531 427.863281 L 71.957031 426.898438 L 72.644531 425.929688 L 74.707031 423.035156 L 75.390625 422.066406 L 77.453125 419.171875 L 78.140625 418.203125 L 78.828125 417.238281 L 79.511719 416.273438 L 80.199219 415.308594 L 80.886719 414.339844 L 82.949219 411.445312 L 83.632812 410.476562 L 85.695312 407.582031 L 86.382812 406.613281 L 87.070312 405.648438 L 87.753906 404.683594 L 88.441406 403.71875 L 89.128906 402.75 L 91.191406 399.855469 L 91.875 398.886719 L 94.625 395.027344 L 95.3125 394.058594 L 95.996094 393.09375 L 97.371094 391.164062 L 98.058594 390.195312 L 99.433594 388.265625 L 100.117188 387.300781 L 100.804688 386.332031 L 102.867188 383.4375 L 103.554688 382.46875 L 104.238281 381.503906 L 105.613281 379.574219 L 106.300781 378.605469 L 107.675781 376.675781 L 108.359375 375.710938 L 109.046875 374.742188 L 111.109375 371.847656 L 111.796875 370.878906 "/>
<g clip-path="url(#clip41)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(92.156863%,92.156863%,92.156863%);fill-opacity:1;" d="M 234.617188 234.925781 L 428.601562 234.925781 L 428.601562 23.246094 L 234.617188 23.246094 Z M 234.617188 234.925781 "/>
</g>
<g clip-path="url(#clip42)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 204.210938 L 428.601562 204.210938 "/>
</g>
<g clip-path="url(#clip43)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 150.910156 L 428.601562 150.910156 "/>
</g>
<g clip-path="url(#clip44)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 97.613281 L 428.601562 97.613281 "/>
</g>
<g clip-path="url(#clip45)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 44.3125 L 428.601562 44.3125 "/>
</g>
<g clip-path="url(#clip46)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 267.171875 234.925781 L 267.171875 23.246094 "/>
</g>
<g clip-path="url(#clip47)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 301.085938 234.925781 L 301.085938 23.246094 "/>
</g>
<g clip-path="url(#clip48)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 335 234.925781 L 335 23.246094 "/>
</g>
<g clip-path="url(#clip49)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 368.914062 234.925781 L 368.914062 23.246094 "/>
</g>
<g clip-path="url(#clip50)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 402.828125 234.925781 L 402.828125 23.246094 "/>
</g>
<g clip-path="url(#clip51)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 230.859375 L 428.601562 230.859375 "/>
</g>
<g clip-path="url(#clip52)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 177.558594 L 428.601562 177.558594 "/>
</g>
<g clip-path="url(#clip53)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 124.261719 L 428.601562 124.261719 "/>
</g>
<g clip-path="url(#clip54)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 70.960938 L 428.601562 70.960938 "/>
</g>
<g clip-path="url(#clip55)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 250.214844 234.925781 L 250.214844 23.246094 "/>
</g>
<g clip-path="url(#clip56)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 284.128906 234.925781 L 284.128906 23.246094 "/>
</g>
<g clip-path="url(#clip57)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 318.042969 234.925781 L 318.042969 23.246094 "/>
</g>
<g clip-path="url(#clip58)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 351.957031 234.925781 L 351.957031 23.246094 "/>
</g>
<g clip-path="url(#clip59)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 385.871094 234.925781 L 385.871094 23.246094 "/>
</g>
<g clip-path="url(#clip60)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 419.785156 234.925781 L 419.785156 23.246094 "/>
</g>
<path style="fill-rule:nonzero;fill:rgb(48.627451%,68.235294%,0%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 326.78125 129.589844 C 326.78125 132.199219 322.871094 132.199219 322.871094 129.589844 C 322.871094 126.984375 326.78125 126.984375 326.78125 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(48.627451%,68.235294%,0%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 326.78125 129.589844 C 326.78125 132.199219 322.871094 132.199219 322.871094 129.589844 C 322.871094 126.984375 326.78125 126.984375 326.78125 129.589844 "/>
<path style="fill-rule:nonzero;fill:rgb(48.627451%,68.235294%,0%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 320 134.921875 C 320 137.527344 316.089844 137.527344 316.089844 134.921875 C 316.089844 132.316406 320 132.316406 320 134.921875 "/>
<path style="fill-rule:nonzero;fill:rgb(48.627451%,68.235294%,0%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 320 129.589844 C 320 132.199219 316.089844 132.199219 316.089844 129.589844 C 316.089844 126.984375 320 126.984375 320 129.589844 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 318.042969 124.148438 L 318.128906 124.214844 L 318.214844 124.285156 L 318.300781 124.347656 L 318.386719 124.414062 L 318.472656 124.476562 L 318.558594 124.535156 L 318.644531 124.597656 L 318.730469 124.652344 L 318.816406 124.710938 L 318.902344 124.761719 L 318.988281 124.816406 L 319.074219 124.867188 L 319.246094 124.960938 L 319.417969 125.046875 L 319.503906 125.085938 L 319.675781 125.15625 L 319.847656 125.21875 L 320.105469 125.289062 L 320.277344 125.320312 L 320.363281 125.332031 L 320.535156 125.347656 L 320.707031 125.347656 L 320.878906 125.332031 L 320.964844 125.320312 L 321.050781 125.304688 L 321.136719 125.285156 L 321.308594 125.238281 L 321.390625 125.207031 L 321.5625 125.136719 L 321.734375 125.050781 L 321.90625 124.949219 L 321.992188 124.894531 L 322.078125 124.835938 L 322.164062 124.773438 L 322.25 124.707031 L 322.421875 124.566406 L 322.59375 124.410156 L 322.679688 124.328125 L 322.765625 124.242188 L 322.9375 124.0625 L 323.023438 123.96875 L 323.195312 123.773438 L 323.28125 123.671875 L 323.453125 123.460938 L 323.539062 123.351562 L 323.796875 123.011719 L 323.882812 122.894531 L 324.054688 122.652344 L 324.3125 122.277344 L 324.484375 122.019531 L 324.742188 121.621094 L 324.828125 121.484375 L 324.828125 137.699219 L 324.742188 137.632812 L 324.65625 137.5625 L 324.570312 137.5 L 324.484375 137.433594 L 324.398438 137.371094 L 324.3125 137.3125 L 324.226562 137.25 L 324.140625 137.195312 L 324.054688 137.136719 L 323.96875 137.085938 L 323.882812 137.03125 L 323.796875 136.980469 L 323.625 136.886719 L 323.453125 136.800781 L 323.367188 136.761719 L 323.195312 136.691406 L 323.023438 136.628906 L 322.765625 136.558594 L 322.59375 136.527344 L 322.507812 136.515625 L 322.421875 136.507812 L 322.25 136.5 L 322.164062 136.5 L 321.992188 136.515625 L 321.90625 136.527344 L 321.820312 136.542969 L 321.734375 136.5625 L 321.5625 136.609375 L 321.476562 136.640625 L 321.390625 136.675781 L 321.308594 136.710938 L 321.136719 136.796875 L 320.964844 136.898438 L 320.878906 136.953125 L 320.792969 137.011719 L 320.707031 137.074219 L 320.621094 137.140625 L 320.449219 137.28125 L 320.277344 137.4375 L 320.191406 137.519531 L 320.105469 137.605469 L 319.933594 137.785156 L 319.847656 137.878906 L 319.675781 138.074219 L 319.589844 138.175781 L 319.417969 138.386719 L 319.332031 138.496094 L 319.074219 138.835938 L 318.988281 138.953125 L 318.816406 139.195312 L 318.558594 139.570312 L 318.386719 139.828125 L 318.214844 140.09375 L 318.128906 140.230469 L 318.042969 140.363281 Z M 318.042969 124.148438 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 318.042969 132.257812 L 318.214844 132.1875 L 318.300781 132.15625 L 318.472656 132.085938 L 318.558594 132.054688 L 318.644531 132.019531 L 318.730469 131.988281 L 318.902344 131.917969 L 318.988281 131.886719 L 319.160156 131.816406 L 319.246094 131.785156 L 319.417969 131.714844 L 319.503906 131.683594 L 319.589844 131.648438 L 319.675781 131.617188 L 319.847656 131.546875 L 319.933594 131.515625 L 320.105469 131.445312 L 320.191406 131.414062 L 320.363281 131.34375 L 320.449219 131.3125 L 320.535156 131.277344 L 320.621094 131.246094 L 320.792969 131.175781 L 320.878906 131.144531 L 321.050781 131.074219 L 321.136719 131.042969 L 321.308594 130.972656 L 321.390625 130.941406 L 321.476562 130.90625 L 321.5625 130.875 L 321.734375 130.804688 L 321.820312 130.773438 L 321.992188 130.703125 L 322.078125 130.671875 L 322.25 130.601562 L 322.335938 130.570312 L 322.421875 130.535156 L 322.507812 130.503906 L 322.679688 130.433594 L 322.765625 130.402344 L 322.9375 130.332031 L 323.023438 130.300781 L 323.195312 130.230469 L 323.28125 130.199219 L 323.367188 130.164062 L 323.453125 130.132812 L 323.625 130.0625 L 323.710938 130.03125 L 323.882812 129.960938 L 323.96875 129.929688 L 324.140625 129.859375 L 324.226562 129.828125 L 324.3125 129.792969 L 324.398438 129.761719 L 324.570312 129.691406 L 324.65625 129.660156 L 324.828125 129.589844 "/>
<g clip-path="url(#clip61)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(92.156863%,92.156863%,92.156863%);fill-opacity:1;" d="M 234.617188 469.847656 L 428.601562 469.847656 L 428.601562 258.167969 L 234.617188 258.167969 Z M 234.617188 469.847656 "/>
</g>
<g clip-path="url(#clip62)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 439.132812 L 428.601562 439.132812 "/>
</g>
<g clip-path="url(#clip63)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 385.835938 L 428.601562 385.835938 "/>
</g>
<g clip-path="url(#clip64)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 332.535156 L 428.601562 332.535156 "/>
</g>
<g clip-path="url(#clip65)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 279.238281 L 428.601562 279.238281 "/>
</g>
<g clip-path="url(#clip66)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 267.171875 469.847656 L 267.171875 258.171875 "/>
</g>
<g clip-path="url(#clip67)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 301.085938 469.847656 L 301.085938 258.171875 "/>
</g>
<g clip-path="url(#clip68)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 335 469.847656 L 335 258.171875 "/>
</g>
<g clip-path="url(#clip69)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 368.914062 469.847656 L 368.914062 258.171875 "/>
</g>
<g clip-path="url(#clip70)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 402.828125 469.847656 L 402.828125 258.171875 "/>
</g>
<g clip-path="url(#clip71)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 465.785156 L 428.601562 465.785156 "/>
</g>
<g clip-path="url(#clip72)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 412.484375 L 428.601562 412.484375 "/>
</g>
<g clip-path="url(#clip73)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 359.1875 L 428.601562 359.1875 "/>
</g>
<g clip-path="url(#clip74)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 234.617188 305.886719 L 428.601562 305.886719 "/>
</g>
<g clip-path="url(#clip75)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 250.214844 469.847656 L 250.214844 258.171875 "/>
</g>
<g clip-path="url(#clip76)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 284.128906 469.847656 L 284.128906 258.171875 "/>
</g>
<g clip-path="url(#clip77)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 318.042969 469.847656 L 318.042969 258.171875 "/>
</g>
<g clip-path="url(#clip78)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 351.957031 469.847656 L 351.957031 258.171875 "/>
</g>
<g clip-path="url(#clip79)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 385.871094 469.847656 L 385.871094 258.171875 "/>
</g>
<g clip-path="url(#clip80)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 419.785156 469.847656 L 419.785156 258.171875 "/>
</g>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 292.867188 396.496094 C 292.867188 399.101562 288.957031 399.101562 288.957031 396.496094 C 288.957031 393.890625 292.867188 393.890625 292.867188 396.496094 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 412.484375 C 279.300781 415.089844 275.394531 415.089844 275.394531 412.484375 C 275.394531 409.878906 279.300781 409.878906 279.300781 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 412.484375 C 279.300781 415.089844 275.394531 415.089844 275.394531 412.484375 C 275.394531 409.878906 279.300781 409.878906 279.300781 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 428.472656 C 265.738281 431.082031 261.828125 431.082031 261.828125 428.472656 C 261.828125 425.867188 265.738281 425.867188 265.738281 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 292.867188 380.503906 C 292.867188 383.113281 288.957031 383.113281 288.957031 380.503906 C 288.957031 377.898438 292.867188 377.898438 292.867188 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 396.496094 C 286.085938 399.101562 282.175781 399.101562 282.175781 396.496094 C 282.175781 393.890625 286.085938 393.890625 286.085938 396.496094 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 292.867188 380.503906 C 292.867188 383.113281 288.957031 383.113281 288.957031 380.503906 C 288.957031 377.898438 292.867188 377.898438 292.867188 380.503906 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 385.835938 C 286.085938 388.441406 282.175781 388.441406 282.175781 385.835938 C 282.175781 383.230469 286.085938 383.230469 286.085938 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 391.164062 C 286.085938 393.773438 282.175781 393.773438 282.175781 391.164062 C 282.175781 388.558594 286.085938 388.558594 286.085938 391.164062 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 417.816406 C 279.300781 420.421875 275.394531 420.421875 275.394531 417.816406 C 275.394531 415.207031 279.300781 415.207031 279.300781 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 444.464844 C 258.953125 447.070312 255.042969 447.070312 255.042969 444.464844 C 255.042969 441.859375 258.953125 441.859375 258.953125 444.464844 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 428.472656 C 279.300781 431.082031 275.394531 431.082031 275.394531 428.472656 C 275.394531 425.867188 279.300781 425.867188 279.300781 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 417.816406 C 279.300781 420.421875 275.394531 420.421875 275.394531 417.816406 C 275.394531 415.207031 279.300781 415.207031 279.300781 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 417.816406 C 279.300781 420.421875 275.394531 420.421875 275.394531 417.816406 C 275.394531 415.207031 279.300781 415.207031 279.300781 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 245.390625 455.125 C 245.390625 457.730469 241.480469 457.730469 241.480469 455.125 C 241.480469 452.515625 245.390625 452.515625 245.390625 455.125 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 428.472656 C 258.953125 431.082031 255.042969 431.082031 255.042969 428.472656 C 255.042969 425.867188 258.953125 425.867188 258.953125 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 245.390625 455.125 C 245.390625 457.730469 241.480469 457.730469 241.480469 455.125 C 241.480469 452.515625 245.390625 452.515625 245.390625 455.125 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 433.804688 C 258.953125 436.410156 255.042969 436.410156 255.042969 433.804688 C 255.042969 431.199219 258.953125 431.199219 258.953125 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 423.144531 C 272.519531 425.75 268.609375 425.75 268.609375 423.144531 C 268.609375 420.539062 272.519531 420.539062 272.519531 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 433.804688 C 265.738281 436.410156 261.828125 436.410156 261.828125 433.804688 C 261.828125 431.199219 265.738281 431.199219 265.738281 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 245.390625 455.125 C 245.390625 457.730469 241.480469 457.730469 241.480469 455.125 C 241.480469 452.515625 245.390625 452.515625 245.390625 455.125 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 433.804688 C 265.738281 436.410156 261.828125 436.410156 261.828125 433.804688 C 261.828125 431.199219 265.738281 431.199219 265.738281 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 245.390625 455.125 C 245.390625 457.730469 241.480469 457.730469 241.480469 455.125 C 241.480469 452.515625 245.390625 452.515625 245.390625 455.125 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 433.804688 C 258.953125 436.410156 255.042969 436.410156 255.042969 433.804688 C 255.042969 431.199219 258.953125 431.199219 258.953125 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 428.472656 C 258.953125 431.082031 255.042969 431.082031 255.042969 428.472656 C 255.042969 425.867188 258.953125 425.867188 258.953125 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 428.472656 C 258.953125 431.082031 255.042969 431.082031 255.042969 428.472656 C 255.042969 425.867188 258.953125 425.867188 258.953125 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 423.144531 C 265.738281 425.75 261.828125 425.75 261.828125 423.144531 C 261.828125 420.539062 265.738281 420.539062 265.738281 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 417.816406 C 272.519531 420.421875 268.609375 420.421875 268.609375 417.816406 C 268.609375 415.207031 272.519531 415.207031 272.519531 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 433.804688 C 272.519531 436.410156 268.609375 436.410156 268.609375 433.804688 C 268.609375 431.199219 272.519531 431.199219 272.519531 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 433.804688 C 272.519531 436.410156 268.609375 436.410156 268.609375 433.804688 C 268.609375 431.199219 272.519531 431.199219 272.519531 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 407.15625 C 286.085938 409.761719 282.175781 409.761719 282.175781 407.15625 C 282.175781 404.546875 286.085938 404.546875 286.085938 407.15625 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 401.824219 C 286.085938 404.429688 282.175781 404.429688 282.175781 401.824219 C 282.175781 399.21875 286.085938 399.21875 286.085938 401.824219 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 396.496094 C 286.085938 399.101562 282.175781 399.101562 282.175781 396.496094 C 282.175781 393.890625 286.085938 393.890625 286.085938 396.496094 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 286.085938 401.824219 C 286.085938 404.429688 282.175781 404.429688 282.175781 401.824219 C 282.175781 399.21875 286.085938 399.21875 286.085938 401.824219 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 412.484375 C 279.300781 415.089844 275.394531 415.089844 275.394531 412.484375 C 275.394531 409.878906 279.300781 409.878906 279.300781 412.484375 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 428.472656 C 279.300781 431.082031 275.394531 431.082031 275.394531 428.472656 C 275.394531 425.867188 279.300781 425.867188 279.300781 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 245.390625 455.125 C 245.390625 457.730469 241.480469 457.730469 241.480469 455.125 C 241.480469 452.515625 245.390625 452.515625 245.390625 455.125 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 417.816406 C 279.300781 420.421875 275.394531 420.421875 275.394531 417.816406 C 275.394531 415.207031 279.300781 415.207031 279.300781 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 423.144531 C 272.519531 425.75 268.609375 425.75 268.609375 423.144531 C 268.609375 420.539062 272.519531 420.539062 272.519531 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 444.464844 C 258.953125 447.070312 255.042969 447.070312 255.042969 444.464844 C 255.042969 441.859375 258.953125 441.859375 258.953125 444.464844 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 423.144531 C 265.738281 425.75 261.828125 425.75 261.828125 423.144531 C 261.828125 420.539062 265.738281 420.539062 265.738281 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 423.144531 C 265.738281 425.75 261.828125 425.75 261.828125 423.144531 C 261.828125 420.539062 265.738281 420.539062 265.738281 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 428.472656 C 258.953125 431.082031 255.042969 431.082031 255.042969 428.472656 C 255.042969 425.867188 258.953125 425.867188 258.953125 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 433.804688 C 258.953125 436.410156 255.042969 436.410156 255.042969 433.804688 C 255.042969 431.199219 258.953125 431.199219 258.953125 433.804688 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 423.144531 C 265.738281 425.75 261.828125 425.75 261.828125 423.144531 C 261.828125 420.539062 265.738281 420.539062 265.738281 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 417.816406 C 272.519531 420.421875 268.609375 420.421875 268.609375 417.816406 C 268.609375 415.207031 272.519531 415.207031 272.519531 417.816406 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 428.472656 C 272.519531 431.082031 268.609375 431.082031 268.609375 428.472656 C 268.609375 425.867188 272.519531 425.867188 272.519531 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 265.738281 423.144531 C 265.738281 425.75 261.828125 425.75 261.828125 423.144531 C 261.828125 420.539062 265.738281 420.539062 265.738281 423.144531 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 292.867188 385.835938 C 292.867188 388.441406 288.957031 388.441406 288.957031 385.835938 C 288.957031 383.230469 292.867188 383.230469 292.867188 385.835938 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 279.300781 428.472656 C 279.300781 431.082031 275.394531 431.082031 275.394531 428.472656 C 275.394531 425.867188 279.300781 425.867188 279.300781 428.472656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 258.953125 439.132812 C 258.953125 441.742188 255.042969 441.742188 255.042969 439.132812 C 255.042969 436.527344 258.953125 436.527344 258.953125 439.132812 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 272.519531 423.144531 C 272.519531 425.75 268.609375 425.75 268.609375 423.144531 C 268.609375 420.539062 272.519531 420.539062 272.519531 423.144531 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 243.433594 452.394531 L 244.035156 451.695312 L 245.238281 450.304688 L 245.839844 449.605469 L 246.4375 448.90625 L 248.84375 446.109375 L 249.445312 445.40625 L 250.046875 444.707031 L 250.644531 444.003906 L 251.246094 443.300781 L 251.847656 442.59375 L 252.449219 441.890625 L 253.652344 440.476562 L 254.253906 439.765625 L 254.851562 439.058594 L 255.453125 438.347656 L 257.257812 436.203125 L 257.859375 435.484375 L 258.460938 434.761719 L 259.058594 434.039062 L 259.660156 433.316406 L 260.261719 432.589844 L 260.863281 431.859375 L 262.066406 430.390625 L 262.667969 429.652344 L 263.265625 428.910156 L 263.867188 428.164062 L 264.46875 427.414062 L 265.070312 426.660156 L 265.671875 425.902344 L 266.273438 425.140625 L 266.875 424.375 L 267.472656 423.601562 L 268.074219 422.828125 L 268.675781 422.050781 L 269.878906 420.480469 L 271.082031 418.894531 L 271.679688 418.09375 L 272.882812 416.484375 L 273.484375 415.675781 L 274.085938 414.863281 L 274.6875 414.046875 L 275.285156 413.226562 L 275.886719 412.40625 L 277.089844 410.757812 L 277.691406 409.929688 L 278.894531 408.265625 L 279.492188 407.433594 L 281.296875 404.925781 L 282.5 403.246094 L 283.101562 402.402344 L 283.699219 401.5625 L 284.902344 399.875 L 285.503906 399.027344 L 286.105469 398.183594 L 287.308594 396.488281 L 287.90625 395.640625 L 288.507812 394.792969 L 289.109375 393.941406 L 289.710938 393.09375 L 290.3125 392.242188 L 290.914062 391.394531 L 290.914062 398.976562 L 290.3125 399.675781 L 289.710938 400.371094 L 287.90625 402.46875 L 287.308594 403.167969 L 286.707031 403.867188 L 286.105469 404.570312 L 285.503906 405.269531 L 283.699219 407.378906 L 283.101562 408.085938 L 281.296875 410.207031 L 279.492188 412.339844 L 278.894531 413.054688 L 278.292969 413.769531 L 276.488281 415.925781 L 275.285156 417.378906 L 274.6875 418.105469 L 273.484375 419.574219 L 272.28125 421.050781 L 271.679688 421.796875 L 271.082031 422.546875 L 270.480469 423.296875 L 269.277344 424.8125 L 268.675781 425.578125 L 268.074219 426.347656 L 267.472656 427.121094 L 266.875 427.898438 L 266.273438 428.679688 L 265.671875 429.464844 L 265.070312 430.253906 L 264.46875 431.046875 L 263.867188 431.84375 L 263.265625 432.644531 L 262.667969 433.453125 L 262.066406 434.257812 L 261.464844 435.070312 L 260.261719 436.703125 L 259.660156 437.523438 L 259.058594 438.347656 L 258.460938 439.171875 L 257.257812 440.828125 L 255.453125 443.324219 L 254.851562 444.160156 L 254.253906 445 L 251.847656 448.359375 L 250.644531 450.046875 L 250.046875 450.894531 L 249.445312 451.738281 L 248.84375 452.585938 L 248.242188 453.429688 L 247.640625 454.277344 L 247.039062 455.128906 L 246.4375 455.976562 L 245.839844 456.824219 L 244.636719 458.527344 L 244.035156 459.375 L 243.433594 460.226562 Z M 243.433594 452.394531 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 243.433594 456.308594 L 245.839844 453.214844 L 246.4375 452.441406 L 250.046875 447.800781 L 250.644531 447.027344 L 251.246094 446.25 L 254.253906 442.382812 L 254.851562 441.609375 L 258.460938 436.96875 L 259.058594 436.195312 L 259.660156 435.417969 L 262.667969 431.550781 L 263.265625 430.777344 L 266.875 426.136719 L 267.472656 425.359375 L 271.082031 420.71875 L 271.679688 419.945312 L 274.6875 416.078125 L 275.285156 415.304688 L 275.886719 414.527344 L 278.894531 410.660156 L 279.492188 409.886719 L 283.101562 405.246094 L 283.699219 404.46875 L 287.308594 399.828125 L 287.90625 399.054688 L 290.914062 395.1875 "/>
<g clip-path="url(#clip81)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(85.098039%,85.098039%,85.098039%);fill-opacity:1;" d="M 35.152344 258.171875 L 229.136719 258.171875 L 229.136719 240.40625 L 35.152344 240.40625 Z M 35.152344 258.171875 "/>
</g>
<g style="fill:rgb(10.196078%,10.196078%,10.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-1" x="129.644531" y="252.204102"/>
</g>
<g clip-path="url(#clip82)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(85.098039%,85.098039%,85.098039%);fill-opacity:1;" d="M 234.617188 258.171875 L 428.601562 258.171875 L 428.601562 240.40625 L 234.617188 240.40625 Z M 234.617188 258.171875 "/>
</g>
<g style="fill:rgb(10.196078%,10.196078%,10.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-2" x="329.109375" y="252.204102"/>
</g>
<g clip-path="url(#clip83)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(85.098039%,85.098039%,85.098039%);fill-opacity:1;" d="M 35.152344 23.246094 L 229.136719 23.246094 L 229.136719 5.480469 L 35.152344 5.480469 Z M 35.152344 23.246094 "/>
</g>
<g style="fill:rgb(10.196078%,10.196078%,10.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-3" x="129.644531" y="17.27832"/>
</g>
<g clip-path="url(#clip84)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(85.098039%,85.098039%,85.098039%);fill-opacity:1;" d="M 234.617188 23.246094 L 428.601562 23.246094 L 428.601562 5.480469 L 234.617188 5.480469 Z M 234.617188 23.246094 "/>
</g>
<g style="fill:rgb(10.196078%,10.196078%,10.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-4" x="329.109375" y="17.27832"/>
</g>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 50.75 472.589844 L 50.75 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 84.664062 472.589844 L 84.664062 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 118.578125 472.589844 L 118.578125 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 152.492188 472.589844 L 152.492188 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 186.40625 472.589844 L 186.40625 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 220.320312 472.589844 L 220.320312 469.847656 "/>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="45.75" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="50.75" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="79.664062" y="481.196289"/>
  <use xlink:href="#glyph0-4" x="84.664062" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-7" x="113.578125" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="118.578125" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-7" x="147.492188" y="481.196289"/>
  <use xlink:href="#glyph0-4" x="152.492188" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-8" x="181.40625" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="186.40625" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-8" x="215.320312" y="481.196289"/>
  <use xlink:href="#glyph0-4" x="220.320312" y="481.196289"/>
</g>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 250.214844 472.589844 L 250.214844 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 284.128906 472.589844 L 284.128906 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 318.042969 472.589844 L 318.042969 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 351.957031 472.589844 L 351.957031 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 385.871094 472.589844 L 385.871094 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 419.785156 472.589844 L 419.785156 469.847656 "/>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="245.214844" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="250.214844" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="279.128906" y="481.196289"/>
  <use xlink:href="#glyph0-4" x="284.128906" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-7" x="313.042969" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="318.042969" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-7" x="346.957031" y="481.196289"/>
  <use xlink:href="#glyph0-4" x="351.957031" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-8" x="380.871094" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="385.871094" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-8" x="414.785156" y="481.196289"/>
  <use xlink:href="#glyph0-4" x="419.785156" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="20.21875" y="233.774414"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="233.774414"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-7" x="20.21875" y="180.473633"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="180.473633"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-8" x="20.21875" y="127.176758"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="127.176758"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-3" x="20.21875" y="73.875977"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="73.875977"/>
</g>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 230.859375 L 35.152344 230.859375 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 177.558594 L 35.152344 177.558594 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 124.261719 L 35.152344 124.261719 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 70.960938 L 35.152344 70.960938 "/>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="20.21875" y="468.700195"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="468.700195"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-7" x="20.21875" y="415.399414"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="415.399414"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-8" x="20.21875" y="362.102539"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="362.102539"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-3" x="20.21875" y="308.801758"/>
  <use xlink:href="#glyph0-6" x="25.21875" y="308.801758"/>
</g>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 465.785156 L 35.152344 465.785156 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 412.484375 L 35.152344 412.484375 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 359.1875 L 35.152344 359.1875 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 305.886719 L 35.152344 305.886719 "/>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph1-1" x="224.875" y="494.538086"/>
  <use xlink:href="#glyph1-2" x="230.875" y="494.538086"/>
  <use xlink:href="#glyph1-3" x="233.875" y="494.538086"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph2-1" x="13.499023" y="256.046875"/>
  <use xlink:href="#glyph2-2" x="13.499023" y="250.046875"/>
  <use xlink:href="#glyph2-3" x="13.499023" y="242.046875"/>
</g>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(100%,100%,100%);fill-opacity:1;" d="M 439.5625 295.328125 L 498.519531 295.328125 L 498.519531 197.769531 L 439.5625 197.769531 Z M 439.5625 295.328125 "/>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph1-4" x="445.042969" y="212.768555"/>
  <use xlink:href="#glyph1-5" x="448.042969" y="212.768555"/>
  <use xlink:href="#glyph1-1" x="454.042969" y="212.768555"/>
  <use xlink:href="#glyph1-2" x="460.042969" y="212.768555"/>
  <use xlink:href="#glyph1-6" x="463.042969" y="212.768555"/>
  <use xlink:href="#glyph1-7" x="469.042969" y="212.768555"/>
  <use xlink:href="#glyph1-8" x="473.042969" y="212.768555"/>
  <use xlink:href="#glyph1-1" x="477.042969" y="212.768555"/>
  <use xlink:href="#glyph1-3" x="482.042969" y="212.768555"/>
  <use xlink:href="#glyph1-9" x="487.042969" y="212.768555"/>
  <use xlink:href="#glyph1-10" x="489.042969" y="212.768555"/>
</g>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(94.901961%,94.901961%,94.901961%);fill-opacity:1;" d="M 445.042969 238.007812 L 462.324219 238.007812 L 462.324219 220.726562 L 445.042969 220.726562 Z M 445.042969 238.007812 "/>
<path style="fill-rule:nonzero;fill:rgb(97.254902%,46.27451%,42.745098%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 455.636719 229.367188 C 455.636719 231.972656 451.726562 231.972656 451.726562 229.367188 C 451.726562 226.761719 455.636719 226.761719 455.636719 229.367188 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 445.042969 238.007812 L 462.324219 238.007812 L 462.324219 220.726562 L 445.042969 220.726562 Z M 445.042969 238.007812 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(97.254902%,46.27451%,42.745098%);stroke-opacity:1;stroke-miterlimit:10;" d="M 446.769531 229.367188 L 460.59375 229.367188 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(94.901961%,94.901961%,94.901961%);fill-opacity:1;" d="M 445.042969 255.289062 L 462.324219 255.289062 L 462.324219 238.007812 L 445.042969 238.007812 Z M 445.042969 255.289062 "/>
<path style="fill-rule:nonzero;fill:rgb(48.627451%,68.235294%,0%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 455.636719 246.648438 C 455.636719 249.253906 451.726562 249.253906 451.726562 246.648438 C 451.726562 244.042969 455.636719 244.042969 455.636719 246.648438 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 445.042969 255.289062 L 462.324219 255.289062 L 462.324219 238.007812 L 445.042969 238.007812 Z M 445.042969 255.289062 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(48.627451%,68.235294%,0%);stroke-opacity:1;stroke-miterlimit:10;" d="M 446.769531 246.648438 L 460.59375 246.648438 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(94.901961%,94.901961%,94.901961%);fill-opacity:1;" d="M 445.042969 272.566406 L 462.324219 272.566406 L 462.324219 255.285156 L 445.042969 255.285156 Z M 445.042969 272.566406 "/>
<path style="fill-rule:nonzero;fill:rgb(0%,74.901961%,76.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 455.636719 263.925781 C 455.636719 266.535156 451.726562 266.535156 451.726562 263.925781 C 451.726562 261.320312 455.636719 261.320312 455.636719 263.925781 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 445.042969 272.566406 L 462.324219 272.566406 L 462.324219 255.285156 L 445.042969 255.285156 Z M 445.042969 272.566406 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(0%,74.901961%,76.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 446.769531 263.925781 L 460.59375 263.925781 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(94.901961%,94.901961%,94.901961%);fill-opacity:1;" d="M 445.042969 289.847656 L 462.324219 289.847656 L 462.324219 272.566406 L 445.042969 272.566406 Z M 445.042969 289.847656 "/>
<path style="fill-rule:nonzero;fill:rgb(78.039216%,48.627451%,100%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 455.636719 281.207031 C 455.636719 283.8125 451.726562 283.8125 451.726562 281.207031 C 451.726562 278.601562 455.636719 278.601562 455.636719 281.207031 "/>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(60%,60%,60%);fill-opacity:0.4;" d="M 445.042969 289.847656 L 462.324219 289.847656 L 462.324219 272.566406 L 445.042969 272.566406 Z M 445.042969 289.847656 "/>
<path style="fill:none;stroke-width:2.133957;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(78.039216%,48.627451%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 446.769531 281.207031 L 460.59375 281.207031 "/>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph0-3" x="467.800781" y="232.282227"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph0-4" x="467.800781" y="249.563477"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph0-1" x="467.800781" y="266.84082"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph0-2" x="467.800781" y="284.12207"/>
</g>
</g>
</svg>

```

## Representing plots as Clojure data
Inspired by
[cxplot](https://cxplot.com/index.html)'s internal ggplot.as.list finction, let us represent ggplot objects as R data structures.

::: {.sourceClojure}
```clojure
(defn ggolot->clj
  ([r-obj
    options]
   (ggolot->clj r-obj options []))
  ([r-obj
    {:as options
     :keys [avoid]}
    path]
   #_(prn path)
   (let [relevant-names (some->> r-obj
                                 r.base/names
                                 r->clj
                                 (filter (complement avoid)))]
     (cond
       ;;
       ;; a named list or a ggproto object
       (seq relevant-names) (->> relevant-names
                                 (map (fn [nam]
                                        [(keyword nam) (-> r-obj
                                                           (r$ nam)
                                                           (ggolot->clj options
                                                                        (conj path nam)))]))
                                 (into {}))
       ;;
       ;; a ggproto method
       (-> r-obj
           r.base/class
           r->clj
           first
           (= "ggproto_method"))
       :ggproto-method
       ;;
       ;; an unnamed list
       (-> r-obj
           r.base/is-list
           r->clj
           first)
       (-> r-obj
           r.base/length
           r->clj
           first
           range
           (->> (mapv (fn [i]
                        (prn [path (inc i)])
                        (-> r-obj
                            (r/brabra (inc i))
                            (ggolot->clj options
                                         (conj path [i])))))))
       ;;
       (r.base/is-atomic r-obj) (try (r->clj r-obj)
                                     (catch Exception e
                                       (-> r-obj println with-out-str)))
       :else r-obj))))
```
:::

For example:

::: {.sourceClojure}
```clojure
(-> "(ggplot(mpg, aes(cty, hwy))
         + geom_point())"
    r
    (ggolot->clj {:avoid #{"data" "plot_env"}}))
```
:::

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position
   {:compute_layer :ggproto-method, :super :ggproto-method}}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"]}}

```
:::

## Exlploring a few plots
Let us explore and compare a few plots this way:

::: {.sourceClojure}
```clojure
(defn h4 [title]
  (kind/hiccup [:h3 title]))
```
:::

::: {.sourceClojure}
```clojure
(defn ggplot-summary [r-code]
  (let [plot (r r-code)
        clj (-> plot
                (ggolot->clj {:avoid #{"data" "plot_env"}}))]
    (kind/fragment
     [(h4 "R code")
      (kind/md
       (format "\n```{r eval=FALSE}\n%s\n```\n"
               r-code))
      (h4 "plot")
      (-> plot
          plotting/plot->buffered-image)
      (h4 "clj data")
      (kind/pprint clj)
      (h4 "portal view")
      (->> clj
           (walk/postwalk (fn [form]
                            ;; Avoiding symbols that
                            (if (or (and (symbol? form)
                                         (-> form str (= "~")))
                                    (and (keyword? form)
                                         (-> form str (= "."))))
                              (str form)
                              form)))
           kind/portal)])))
```
:::

### A scatterplot

::: {.sourceClojure}
```clojure
(ggplot-summary
 "(ggplot(mpg, aes(cty, hwy))
         + geom_point())")
```
:::

```{=html}
<h3>R code</h3>
```

```{r eval=FALSE}
(ggplot(mpg, aes(cty, hwy))
         + geom_point())
```

```{=html}
<h3>plot</h3>
```

```{=html}
<img src="ggplot_files/0.png" style="width:100%;" />
```

```{=html}
<h3>clj data</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position
   {:compute_layer :ggproto-method, :super :ggproto-method}}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"]}}

```
:::

```{=html}
<h3>portal view</h3>
```

```{=html}
<div><script>portal_api.embed().renderOutputItem(
                {'mime': 'x-application/edn',
                 'text': (() => "^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :layers]] [^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :constructor]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector geom_point]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stat_params]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :na.rm]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :position]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :compute_layer]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :inherit.aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :show.legend]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geom_params]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :na.rm]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :mapping]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :aes_params]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stat]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :compute_layer]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geom]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :non_missing_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"size\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"shape\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"colour\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :draw_key]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :default_aes]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :shape]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 19.0]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :colour]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"black\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :size]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 1.5]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :fill]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :alpha]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stroke]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.5]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :required_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"x\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"y\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :draw_panel]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}}], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :scales]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :scales]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :mapping]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :x]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"~\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector cty]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :y]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"~\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector hwy]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :theme]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :coordinates]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :expand]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :clip]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"on\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :limits]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :x]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :y]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :default]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :facet]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :shrink]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :labels]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :x]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"cty\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :y]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"hwy\"]]]}}")}
                , document.currentScript.parentElement);</script></div>
```

### A scatterplot with colours

::: {.sourceClojure}
```clojure
(ggplot-summary
 "(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point())")
```
:::

```{=html}
<h3>R code</h3>
```

```{r eval=FALSE}
(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point())
```

```{=html}
<h3>plot</h3>
```

```{=html}
<img src="ggplot_files/1.png" style="width:100%;" />
```

```{=html}
<h3>clj data</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position
   {:compute_layer :ggproto-method, :super :ggproto-method}}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ [factor cyl]]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

```
:::

```{=html}
<h3>portal view</h3>
```

```{=html}
<div><script>portal_api.embed().renderOutputItem(
                {'mime': 'x-application/edn',
                 'text': (() => "^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :layers]] [^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :constructor]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector geom_point]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stat_params]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :na.rm]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :position]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :compute_layer]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :inherit.aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :show.legend]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geom_params]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :na.rm]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :mapping]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :aes_params]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stat]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :compute_layer]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geom]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :non_missing_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"size\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"shape\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"colour\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :draw_key]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :default_aes]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :shape]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 19.0]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :colour]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"black\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :size]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 1.5]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :fill]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :alpha]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stroke]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.5]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :required_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"x\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"y\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :draw_panel]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}}], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :scales]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :scales]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :mapping]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :x]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"~\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector cty]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :y]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"~\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector hwy]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :colour]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"~\"]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector factor]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector cyl]]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :theme]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :coordinates]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :expand]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :clip]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"on\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :limits]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :x]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :y]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :default]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :facet]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :shrink]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :labels]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :x]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"cty\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :y]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"hwy\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :colour]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"factor(cyl)\"]]]}}")}
                , document.currentScript.parentElement);</script></div>
```

### A scatterplot with colours and smoothing

::: {.sourceClojure}
```clojure
(ggplot-summary
 "(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method=\"lm\"))")
```
:::

```{=html}
<h3>R code</h3>
```

```{r eval=FALSE}
(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method="lm"))
```

```{=html}
<h3>plot</h3>
```

```{=html}
<img src="ggplot_files/2.png" style="width:100%;" />
```

```{=html}
<h3>clj data</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method}}
  {:aes_params [],
   :stat
   {:extra_params ["na.rm" "orientation"],
    :super :ggproto-method,
    :compute_group :ggproto-method,
    :required_aes ["x" "y"],
    :setup_params :ggproto-method,
    :dropped_aes ["weight"]},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:se [true], :na.rm [false], :orientation [nil]},
   :geom
   {:setup_params :ggproto-method,
    :super :ggproto-method,
    :draw_key :ggproto-method,
    :extra_params ["na.rm" "orientation"],
    :required_aes ["x" "y"],
    :draw_group :ggproto-method,
    :optional_aes ["ymin" "ymax"],
    :rename_size [true],
    :setup_data :ggproto-method,
    :default_aes
    {:colour ["#3366FF"],
     :fill ["grey60"],
     :linewidth [1.0],
     :linetype [1.0],
     :weight [1.0],
     :alpha [0.4]}},
   :stat_params
   {:na.rm [false],
    :method ["lm"],
    :n [80.0],
    :orientation [nil],
    :level [0.95],
    :fullrange [false],
    :se [true],
    :formula nil,
    :method.args [],
    :span [0.75]},
   :constructor {: nil, :method ["lm"]},
   :position
   {:compute_layer :ggproto-method, :super :ggproto-method}}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ [factor cyl]]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

```
:::

```{=html}
<h3>portal view</h3>
```

```{=html}
<div><script>portal_api.embed().renderOutputItem(
                {'mime': 'x-application/edn',
                 'text': (() => "^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :layers]] [^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :constructor]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector geom_point]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stat_params]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :na.rm]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :position]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :compute_layer]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :inherit.aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :show.legend]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geom_params]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :na.rm]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :mapping]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :aes_params]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stat]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :compute_layer]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geom]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :non_missing_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"size\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"shape\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"colour\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :draw_key]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :default_aes]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :shape]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 19.0]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :colour]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"black\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :size]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 1.5]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :fill]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :alpha]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stroke]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.5]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :required_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"x\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"y\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :draw_panel]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}} ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :constructor]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :method]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"lm\"]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stat_params]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :method.args]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :span]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.75]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :formula]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :level]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.95]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :orientation]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :n]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 80.0]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :se]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :method]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"lm\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :fullrange]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :na.rm]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :position]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :compute_layer]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :inherit.aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :show.legend]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geom_params]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :se]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :na.rm]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :orientation]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :mapping]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :aes_params]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stat]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :extra_params]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"na.rm\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"orientation\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :compute_group]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :required_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"x\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"y\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :setup_params]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :dropped_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"weight\"]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geom]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :draw_group]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :draw_key]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :default_aes]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :colour]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"#3366FF\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :fill]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"grey60\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :linewidth]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 1.0]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :linetype]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 1.0]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :weight]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 1.0]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :alpha]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.4]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :rename_size]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :required_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"x\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"y\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :setup_data]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :optional_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"ymin\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"ymax\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :setup_params]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :extra_params]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"na.rm\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"orientation\"]]]}}], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :scales]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :scales]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :mapping]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :x]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"~\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector cty]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :y]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"~\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector hwy]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :colour]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"~\"]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector factor]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector cyl]]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :theme]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :coordinates]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :expand]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :clip]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"on\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :limits]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :x]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :y]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :default]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :facet]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :shrink]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :labels]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :x]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"cty\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :y]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"hwy\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :colour]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"factor(cyl)\"]]]}}")}
                , document.currentScript.parentElement);</script></div>
```

### A scatterplot with colours, smoothing, and facets

::: {.sourceClojure}
```clojure
(ggplot-summary
 "(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method=\"lm\")
         + facet_wrap(~cyl))")
```
:::

```{=html}
<h3>R code</h3>
```

```{r eval=FALSE}
(ggplot(mpg, aes(cty, hwy, color=factor(cyl)))
         + geom_point()
         + stat_smooth(method="lm")
         + facet_wrap(~cyl))
```

```{=html}
<h3>plot</h3>
```

```{=html}
<img src="ggplot_files/3.png" style="width:100%;" />
```

```{=html}
<h3>clj data</h3>
```

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method}}
  {:aes_params [],
   :stat
   {:extra_params ["na.rm" "orientation"],
    :super :ggproto-method,
    :compute_group :ggproto-method,
    :required_aes ["x" "y"],
    :setup_params :ggproto-method,
    :dropped_aes ["weight"]},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:se [true], :na.rm [false], :orientation [nil]},
   :geom
   {:setup_params :ggproto-method,
    :super :ggproto-method,
    :draw_key :ggproto-method,
    :extra_params ["na.rm" "orientation"],
    :required_aes ["x" "y"],
    :draw_group :ggproto-method,
    :optional_aes ["ymin" "ymax"],
    :rename_size [true],
    :setup_data :ggproto-method,
    :default_aes
    {:colour ["#3366FF"],
     :fill ["grey60"],
     :linewidth [1.0],
     :linetype [1.0],
     :weight [1.0],
     :alpha [0.4]}},
   :stat_params
   {:na.rm [false],
    :method ["lm"],
    :n [80.0],
    :orientation [nil],
    :level [0.95],
    :fullrange [false],
    :se [true],
    :formula nil,
    :method.args [],
    :span [0.75]},
   :constructor {: nil, :method ["lm"]},
   :position
   {:compute_layer :ggproto-method, :super :ggproto-method}}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ [factor cyl]]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet
 {:params
  {:dir ["h"],
   :nrow nil,
   :as.table [true],
   :free {:x [false], :y [false]},
   :drop [true],
   :strip.position ["top"],
   :ncol nil,
   :facets {:cyl [~ cyl]},
   :labeller
   "function (labels, multi_line = TRUE) \n{\n    labels &lt;- lapply(labels, as.character)\n    if (multi_line) {\n        labels\n    }\n    else {\n        collapse_labels_lines(labels)\n    }\n}\n&lt;bytecode: 0x55eafa698ac0&gt;\n&lt;environment: namespace:ggplot2&gt;\nattr(,\"class\")\n[1] \"function\" \"labeller\"\n"},
  :shrink [true],
  :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["factor(cyl)"]}}

```
:::

```{=html}
<h3>portal view</h3>
```

```{=html}
<div><script>portal_api.embed().renderOutputItem(
                {'mime': 'x-application/edn',
                 'text': (() => "^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :layers]] [^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :constructor]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector geom_point]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stat_params]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :na.rm]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :position]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :compute_layer]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :inherit.aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :show.legend]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geom_params]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :na.rm]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :mapping]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :aes_params]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stat]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :compute_layer]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geom]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :non_missing_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"size\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"shape\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"colour\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :draw_key]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :default_aes]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :shape]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 19.0]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :colour]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"black\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :size]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 1.5]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :fill]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :alpha]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stroke]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.5]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :required_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"x\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"y\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :draw_panel]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}} ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :constructor]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :method]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"lm\"]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stat_params]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :method.args]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :span]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.75]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :formula]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :level]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.95]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :orientation]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :n]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 80.0]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :se]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :method]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"lm\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :fullrange]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :na.rm]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :position]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :compute_layer]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :inherit.aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :show.legend]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geom_params]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :se]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :na.rm]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :orientation]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :mapping]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :aes_params]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :stat]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :extra_params]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"na.rm\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"orientation\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :compute_group]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :required_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"x\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"y\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :setup_params]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :dropped_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"weight\"]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :geom]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :draw_group]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :draw_key]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :default_aes]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :colour]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"#3366FF\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :fill]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"grey60\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :linewidth]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 1.0]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :linetype]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 1.0]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :weight]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 1.0]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :alpha]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector 0.4]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :rename_size]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :required_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"x\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"y\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :setup_data]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :optional_aes]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"ymin\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"ymax\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :setup_params]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :extra_params]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"na.rm\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"orientation\"]]]}}], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :scales]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :scales]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :mapping]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :x]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"~\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector cty]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :y]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"~\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector hwy]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :colour]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"~\"]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector factor]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector cyl]]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :theme]] [], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :coordinates]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :expand]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :clip]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"on\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :limits]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :x]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :y]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :default]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :facet]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :params]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :free]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :x]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :y]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector false]]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :drop]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :nrow]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :labeller]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"function (labels, multi_line = TRUE) \\n{\\n    labels <- lapply(labels, as.character)\\n    if (multi_line) {\\n        labels\\n    }\\n    else {\\n        collapse_labels_lines(labels)\\n    }\\n}\\n<bytecode: 0x55eafa698ac0>\\n<environment: namespace:ggplot2>\\nattr(,\\\"class\\\")\\n[1] \\\"function\\\" \\\"labeller\\\"\\n\"]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ncol]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector nil]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :as.table]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :dir]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"h\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :strip.position]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"top\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :facets]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :cyl]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"~\"]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector cyl]]]}}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :shrink]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector true]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :super]] ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :ggproto-method]]}, ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :labels]] ^{:source \"{}\", :line 58, :column 40, :end-line 58, :end-column 42} {^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :x]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"cty\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :y]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"hwy\"]]], ^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector :colour]] [^#:portal.viewer{:default :portal.viewer/hiccup} [:div nil [:portal.viewer/inspector \"factor(cyl)\"]]]}}")}
                , document.currentScript.parentElement);</script></div>
```

```{=html}
<div style="background-color:grey;height:2px;width:100%;"></div>
```

```{=html}
<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/main/projects/noj/notebooks/ggplot.clj">projects/noj/notebooks/ggplot.clj</a></small></small></pre></div>
```