### 2024-02-15 docs/projects/math/wolframite/index.qmd
Wolframite - rerendered

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: index.html}
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}
title: Wolfram Langauge interop with Wolframite
author: Daniel Slutsky

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="index_files/md-default0.js" type="text/javascript"></script><script src="index_files/md-default1.js" type="text/javascript"></script>

```{=html}
<table><tr><td>This is part of the Scicloj <a href="https://scicloj.github.io/clojure-data-scrapbook/">Clojure Data Scrapbook</a>.</td><a href="https://scicloj.github.io/clojure-data-scrapbook/"><img align="left" alt="SciCloj logo" src="https://scicloj.github.io/sci-cloj-logo-transparent.png" width="40" /></a></tr></table>
```

This notebook demonstrates basic usage of [Wolframite](https://github.com/scicloj/wolframite/) in a way that would work in visual tools supporting [Kindly](https://scicloj.github.io/kindly-noted/kindly).
It is also appears at the Wolrfamite repo as [dev/kindly-demo.clj](https://github.com/scicloj/wolframite/blob/main/dev/kindly_demo.clj).
Note that to use Wolframite, you need Wolfamite in the dependencies,
and if Wolfraimte fails to find the correct install path automatically,
you may need to have the `WOLFRAM_INSTALL_PATH` environment variable
set up in your system, as explained in
[Wolframite's README](https://github.com/scicloj/wolframite/).

::: {.sourceClojure}
```clojure
(ns index
  (:refer-clojure
   ;; Exclude symbols also used by Wolfram:
   :exclude [Byte Character Integer Number Short String Thread])
  (:require
   [wolframite.core :as wl]
   [wolframite.tools.hiccup :refer [view]]
   [wolframite.base.parse :as parse]
   [wolframite.jlink]
   [scicloj.kindly.v4.kind :as kind]
   [scicloj.kindly.v4.api :as kindly])
  (:import (com.wolfram.jlink MathCanvas KernelLink)
           (java.awt Color Frame)
           (java.awt.event WindowAdapter ActionEvent)))
```
:::

## Init (base example)

::: {.sourceClojure}
```clojure
(wl/eval '(Dot [1 2 3] [4 5 6]))
```
:::

::: {.printedClojure}
```clojure
32

```
:::

## Strings of WL code

::: {.sourceClojure}
```clojure
(wl/eval "{1 , 2, 3} . {4, 5, 6}")
```
:::

::: {.printedClojure}
```clojure
32

```
:::

## Def / intern WL fns, i.e. effectively define WL fns as clojure fns:

::: {.sourceClojure}
```clojure
(def W:Plus (parse/parse-fn 'Plus {:kernel/link @wl/kernel-link-atom}))
```
:::

::: {.sourceClojure}
```clojure
(W:Plus 1 2 3)
```
:::

::: {.printedClojure}
```clojure
6

```
:::

... and call it

::: {.sourceClojure}
```clojure
(def greetings
  (wl/eval
   '(Function [x] (StringJoin "Hello, " x "! This is a Mathematica function's output."))))
```
:::

::: {.sourceClojure}
```clojure
(greetings "Stephen")
```
:::

::: {.printedClojure}
```clojure
"Hello, Stephen! This is a Mathematica function's output."

```
:::

## Bidirectional translation
(Somewhat experimental, especially in the wl->clj direction)

::: {.sourceClojure}
```clojure
(wl/->clj! "GridGraph[{5, 5}]")
```
:::

::: {.printedClojure}
```clojure
(GridGraph [5 5])

```
:::

::: {.sourceClojure}
```clojure
(wl/->wl! '(GridGraph [5 5]) {:output-fn str})
```
:::

::: {.printedClojure}
```clojure
"GridGraph[{5, 5}]"

```
:::

## Graphics

::: {.sourceClojure}
```clojure
(view
 '(GridGraph [5 5]))
```
:::

```{=html}
<div><details open="open"><summary><h5 style="cursor:pointer;display:inline;padding:0.5rem 1rem 0 1rem;">GridGraph[{5, 5}]</h5></summary><div class="wl-results"><hr /><img src="data:image/jpeg;base64,R0lGODlhwAO/A/frAB8mMSIpMyIpNSQrNSYsNikvNyMrOCMsOSQtOyYvPSkvOCYwPiswOC4zOi80OzA0OzE2PTM4PjU5Py83QzM4QDY6QTI5RTY8RDo9Qjw/Qzk+Rz0/RDA8TjI+UT1AQz5BRTtBSTxCSz1DTD5FTjNAUzVCVTlHXDtJXjxKYD1MY0FDRkJER0RFR0VGSEdISUpKS0tLTEtMTExNTUFHUUJJUkNKVERLVUVMVklOVkZOWEJNXkhPWkhQWk5SWEtSXUxUX1FRUVVVVVFUW1JWXVpaWl1dXUVPYEVRYk5WYU9YY0FRakNUbUVWb05bb1FaZlpdYVNcaFVdalZfbEpbdkpcd0tdeExeek1ffF1gZVdgbVlibk5hfVljcFtkclxmc11mdF9odlhmfGFhYWVlZWRlaGpqam1tbWJpc2NqdG5ucGFreWJsemNtfGRvfmZwf21ye290fXJycnV1dXp6en19fVlvkFpwklxzlV50l191mGZxgGhzgml0g2t2hmx3h2JyimZ2jm56inl9gnF9jnN/kH6CiHyDjXWBknaDlHiFlnqHmXuImnyJm36Lnn+Mn2eApnaJp3mMqn6Ts4KCgoWFhYWHi4mJiY6Ojo2SmIuSnZGRkZKSlJaWlpmZmZ2dnYGOoYKQo4ORpISSpoaUp4aUqImXq4+ZqIqZrYybr5CaqYGWtY2bsI6dspCftJ2iqZqir5Ggtp+ms5OiuJSjuZWkupemvZiovqGhoaWlpamqqq6urq6zvLGxsbW1tbq6ur29vZqqwZyqwJuswpysw52txp+wx6Cwx6GyyqSzyqO0zKu3ybO/0r7CyLjB0b3G1cHBwcXFxcnJyc7OzsHL28TN3M/Q0s7V39LS0tDR1NXV1dnZ2d7e3tLZ4tTb5dbd5+Li4uXl5eDk6urq6u7u7uXp8PHx8fLz9vb29vb3+vj5+/r7/P///////////////////////////////////////////////////////////////////////////////////yH/C05FVFNDQVBFMi4wAwEAAAAh+QQECgD/ACwAAAAAwAO/AwAI/wDXCRxIsKDBgwgTKlzIsKHDhxAjSpxIsaLFixgzatzIsaPHjyBDihxJsqTJkyhTqlzJsqXLlzBjypxJs6bNmzhz6tzJs6fPn0CDCh1KtKjRo0iTKl3KtKnTp1CjSp1KtarVq1izat3KtavXr2DDih1LtqzZs2jTql3Ltq3bt3Djyp1Lt67du3jz6t3Lt6/fv4ADCx5MuLDhw4gTK17MuLHjx5AjS55MubLly5gza97MubPnz6BDix5NurTp06hTq17NurXr17Bjy55N+2a6bM9+XTNXG6U2aL6kiet98ls0X9G+ETcpTpovaNqWlzR37dezbOlegrsEREYMGDGATP+KLt1juU5FZMiAoV4OtPIe0+kqAx58jDG8ssPnCE1O/e9FeFLOfhxpMwkQ38UQQxCXgMMSLjK0gIEEEEAQAQYtxKCJfgRa9EwQMHxAYYUVqADDHMN1aNE2YsCgwogQSGCiGNuoaJE4/qmAQYUxZgBDEM/YWFE6msAgIY8SZNCCDLeoZAkMOyYRSCOfDBJFBBLAIAdvQkakiwwZQHADH4yAgkgXF0jgghgOdglRNECsAIEIbCgCiiJriBDBCkBI4yZE4ojRggQXeIEIKIzwcQMEH8igy58PneOfBBFEMQgojQSCBAQYxGAJSp7AIIENqyRj6qmyIBEBDJRA2tAzMmD/AIIjp54qDBgQtFAGl64mtE0QKlAQCDK1IoNMIBF8EESNvSZkzhhygiFMrckg4wgIGMgQZLMJUSIqErOcSmwyq9yQpScmZQNDBTwQQ624UGT5C7cHlRNEBiDI8u6pfEAAAyf0HmTGChGEsq+poUCwghkBG8QJDBD4cXAysoCg7IAND/SLqFCM+y4xPFQQQzYlybFCBbRMnAwxIXwgBocNdwLDA6ConEwS2baZ8TOi8mHzHhJom7FA4MSKhMfvIgMKBC6gO7Q6RXwAgrsT01LBCnKQVLQEYNicTCASxODn0ERgcAPS+5biLy5Dr0PHyVRPTMzVk7SNC8SlqGxsDhgQ/9G2NDFEEIjXagStM0i9QCyK17Tkyok6kEcu+eSUV2755ZhnrvnmnHfu+eegh+65uhJIbDMyFsch+uqst+7667DHDvo5QGAQhdfJQIEBEObI7vvvwAcvfOtxqAAC7nyEjd3wzDfv/POvc9ICBCnbHMoDMPQykifT2+I1MhVAUMIj5Jdv/vnop6/++uy37/778Mcv//z01y//FgFA0AjuSEiQgP0ADKAAB0jAAhqQfngAQAXagLs2SAAAeTigBCdIwQpaUIAJkAAScNcICARgCxcMoQhHSMICkgACFcBd41rgtJBoYnrT8poGIMCBEtrwhjjM4SOokL+aec0JFQiADv+HSMQiDtAOCvSZ15IHADsY8YlQjKL6AiABJ+AuYQGwghS3yMUicgACGsDdMCLQAk2MBBcugIC+bFYMTkWBGHCMoxznSMc62vGOeMyjHvfIxz768Y+ADKQfG/EACCgCdzzAgAgEychGOvKRkIykJAEpCwdUQA24W4MEHFCLSXryk6AMpSgdOQIM8AB3iahQI0bJyla68pWRjMKO0LYvWTCNbSKBBsQS4TVS+CsXQxPHurpmM2JQYAWtGlrUcoC7G3ygCG2jBAsoELeJgaECMEhRw3IBMVJ4LZUweI9IzBEEDPjAa10I2+ECFocPaKCa+0KEv+Y1NOlBoFQqWwXTADb/tI1BABE2G4YGVBCHtn1DBhXowul8sLtzkKQTaVyEytTWgrq1bWMSWIPKagGCDLzMoLHyQTEmhgyGykA5QzuHGKRWC5WpYVXbGtokppe3iS3CATDoREnKQYQVUMCH71oFtoCA0rbFYXqD25ctdoDNmA5NZhHgwkg/xoVV6bRt63hGDNjlvaQFIlcFxeo3gIAvfL4LFBV4JsZIkg0gfCACYKieqWzBhwp0yqnBLIKcoHCKWhEDESHI0lWxmg45uCACN3CEx6x1g1XJAWZPXRcIEBE3ZJwCCg5gQRG0edEYYKACfABGrWihBkYBgWQniUYQWEAhGjjBCYuSwApk4Aus/xIEHGNwwY5EgIQohIxROVWHbQVSjjlAyQEXQAIUfjBDDJxorVhVRydi8AEU+gAKSBABBC7ggjGss22+kMEKKHSD19IgRi0IQjRU8o05REhESVqSGVA7XOJq4jvwxRAMxIBX26YjF0CAwQrSdIEP/AgXkLXtM8QQgxVMKK3s0QR0h5sNMxhpQhL4QIbmUFSVXIMTRThAAHqgifXW1yDfuIUZEBCAEEzCFw49MUHEwQs5UCAAE5iDLsYhY4Kc4xeTCEEADmCGW3S4x9G4RA8MYIAicIK+LvEG+abRY4Skg3zGqDJCVPEIVWj5IMUgn3C/TJBpkM8bM5HyI6hM5oFc+f8RWW7zQLjsZTkLxBjkS7CWzfwINMtEzWyW85vjbGc62/nOeT40n/0cE0AfetCHXgeXJRFpPD9Cz1VedJqn/GgsR1oSXa50ou2s6T9z2s6QPrShD21pTPe41I0+taA9repHUJrVo5YzrGHiaFTTutChxvWlFX3mTa+503D+dLDt3Gpi99nYgW5zqoF9a2bnus27fkmvZ53sWtfZ2sMmdbFNfWxfd9vOoP62nJst7meTO9pknracVw1uV8s421GWtbR/PW9bizrcuh53rMvNbULLOd3/tveJ8d2Sbe/73P1Wd5vZHXB3DxzeX5Z3myedcGczmtf6jje/N77sdV+bzAz/Z4nDRQ7xNiNc2Aqvb8pXsvKMj5zMHIe5x6GNbIOTXOJkpji2BQ5ygj/c52R+eb13/u6eR5reJgf40C1edIxrWeM497fO2/1xbYfc5i1PesknfvIvz1wlNb/6zb8MdbJLHeVE97rRWY50tmt96VznubnrruW2B73se457vucOdr5XWelRj/lwz56StFcZ63avduKZfnGne7vjeW/63pUN9C8LHe5Ul7vVH7/2vo/97283u+Ab/nW1hz3ymK941wc/+h5DXsuId7vibct4lDje9qWvst89D/hMr17lrSf9600ved1TvuqWB3bntfx51Yee9tE/+OmJn/rAX5/1hHe9/+F7nHO8y17vBX/69qlf/Fcfn+bJB/7yD7/+Klff+7MHf+1lfHvh17/H92d834d84ad84ydj5Td5mVd5m3douYd6u4dVvXcSv8d/wUd+/ydjAeh+Awh/BSh/B3hiCeh8Cwh9DSh9sTd1+UeA+3di/ddjD8h9Edg2E2gSFeiCF4iAd6eA56d56Xd5W9eDDPiD6JaBJ7aB9/Z+aBd/Fjh/GDh99td+SdiBS/iBTRiC9TWCEPh8opd9LmeE9YWEC6eEjceEOOiEOgiFACiFY0iFZWiFZ4iFw6WFMsiF2HeCEZeCoLeCHtiC9fWCMhaD7Nd9AsiHVeiHwwWIIriDJCiEJv9IhHkYhCqIfkfHeXpofYb4hojoXzm4iGqogWwoc2Toe2b4h52YhYy4hSXYhXj4hZ94hKG4eKNIgaWYiKc4h2A4XGIoim5IinBoimi4iM2nio7IipD4c5eIf5RId5YoiXu4jIWnfsNYh6t4h8eYda8YhrHIe7Nog7XIicFYX4IYhYTIgZnoi5tIWLdoW8M3iDM4NDVYEjcIjHLIjqlIjcVojZXogLloW7soi71Ii79oi+GIi9moi9sogd0oj9+ojgVpj9Pojnaof16IjclYiNAofs1ofpPog/tIbRdpjhlpgOp3kP6YkDS4kCQxjwRZj1g1jmtYjlN4jgKZjm2jiKj/aJJY9Y/cGJDeOJDg6JJtQ4cSWY0U2Ypip5Ntw5MK6ZMMCZQOKZRD047k+I4ZE48r2ZA3uY5YRZRVOZEsWJF2p5RDw5Qp6ZRZCZVb+ZAv2Y87iZLwqJIjwZJBKY0hOZMjCYIleZdtSJM/aZNDg5PDBZOgKJN9mZdXuJfOiImIGYd2uZjK6JHMyI9kmTFmGZdoOZdaGZhcOZRuuZRweZVyKRJ0GZWPyZHPKJnRCISoyZiqqZGUyZe86JdPCZgZI5gQKZsASZtpaZsNg5td+ZllGZoNg5WaqZacyZZtQ5iwaJiz2Zj0qJitGZlD+JH9FpFfaZRhiZRfxpza6Jy7CZ0t/ymdPNiR1TmZIAmZGPmaJMma5Zma57maRViZDXOZopmZpLmZt9mZU3mPRZmPR3mNY6mbPcmbx+mbAQOcyymclkmcAWOc+Ymc+6mc/UmfAWOfxTmaIVGaaymVGeOVMWmVGYqfG6qfv8mfGeOdCAmeBSqedemejWiej2idJIedIQqWfSiWpmeh9IKhD6qhIMGhyemhDaOiJ8miTWmgEYqg9KKg/WmjhSmiP0qiQWqiCYqiDUOVN6qdOcqduMeg9emg9AKhJSqhJ0qhHwqmFyqm3EKmVWqmV4qmWeqf2Qmg2ymgX8qj3OKjYwqkHyGkE0qkAaOlUYqjh6ij/gelzSmlff9KpX9qpU2KpYOqpj3Kps3ipo8Kp5EqpwFjpG+JpGeppGXKpNzipGmqqN/JqG3qpx4BqGcqqPRCqItqqJqIqDBIqXtqqb2Cqa0KqaUqqbGKq83Cp6vqqL2qqb/KqcGKqiuqqpfKqh3hqnEKq9ziqaAJqpgpqm9Kqs1iqnOqp8Oqq67Cq9Hqq90KrNwCooXKpYfqpf4Hrr1CrM9qrOWKrOeqrNUqrPEqrpBCrhwhrZtKrc2irrPKrrXqrk9IoEnqoqYZm+opkuyplzBKjDJqjDSadcx6pM66q9D6r+baK97aqfrqKvLKsfTqsfYKsug6sCMLKSU7rh27EQCbrALbKwT/m6q0io62moYKG6oM26Eb+Z6uGZ+wmZ7TuZ5E254o+LB4GbGJ6bBHC7FJK7FLG7VNO7VPa7RCS50zip7aB68ky69/4q8y+7GuErLBCrYuK7ZuQrYaMbP3WrOucrPNmrM1ubOe2LPZ+rNDGrQxCp9dK5/Xqbf3qa2Zyq0qi68sq7Z/8rL9GrNva7aQgrb5yrhu4rhjC7kZAbeJK7eQIqs4a7A6i7A6mLGfurEwe7Jlm7Jnu7K9Yq3Dia2Fy7eBSp5/O7SBW7SDy7SH6bSOObH4WLH66LWuSLgjarjHiritq7g2S6dbaqddiqfvarxTirz1qryT67quArsNKrvHS7uv/2q7FAu4Fku8GEu9jWq9KIu9f0K5zcK9Yeq91Qu+0yq+wUu+wyu4NYq+xaq+q8u+buK+zWu5XYK5bau5GMG5y+u5fwK/ayq/6Uu/AXuaW4u0uau0kWi1vYu1vzuf/Duv/hu5rJu9zDu3Ldu4bNslbru5ktu+2vu5zruu0Nuu0nurBCwkBqzCCHwRCkzCDOwmoFu3onu3pCuMH2yyIczCI+zCJQzDN2wjOSwkK5zALRzAL9zAJ3y5KSzFO2wRPczEP9wldKuxdvuXeJuTR5y6SUzFS2zFTYzFT6wiUWwjU8zDVdwlAmzCcdwhc6widezFdywkeQzDpnutqPu4qivCAP+Mx1fsJg5cqRDcvxJMs/b7n8IboBcLe7z7nL4bncBryfiLyebbnVlcwFtMx11cEV/sxmEsJGN8umVcm2dskGmMyGtsx23MyG/syKWMw6fsx6lMEausy61sI0FMxkNsxkWMioUcu4ecuYmsxIssyI3cJY+cq5EMwpMct5Vcp5fMEtyAB3jgDId2DnlgB7YQaZBwB5AQabZwB3gQY3LmDOLMDTCRDr1gMgAAABhgBrkwYSf2DAeyzw8gBp3AWTKWDZdABAMAAANABJcAZTIGDp0gBg+wz+LRX/VVDrlgBhiwzyogB73wzBERDS3yARMSARBGBLUlY9sQBzDAAhiWJDD/AASeQNLSUQ6UECGfJQESoAEsIAOUANA3eQvr8QEVoNIYYGBxcGS21QtEECJJLQFLvV8mhhK6QF0RoAFaoAd60AUhAAE+wk/DJQ1AME0UAAVtsAdqcF7ONQdE3TDfkFvh4wNrwAdr4AMotCbflTHFBSUPcANgoAdtAEQKEwRjM1wPEyYi4AVerQUz1CiPchIbgwEaQAho4wgikCW4hFXa4FYQoAfVZFk84C90MFzlIAZykgVyZSq0kAW5IgZxTS90ICo80Fd+pQemRR5YdTcVIAL7UyvFgAgX0Cn0pDW1ky8HYws34FzXQFiqTQGhQEvJYAxRkCXAhFXeAgHDsi/I/+AHTJNMbcNNEBAF1J0MobAnH9U21+BZNtBV71IxGQAEfe0RlJBGNXUwVrMwWKULEDMIKlMMzR0Es+0q1yADEpAFxjIxWlABMvDcQ2MvGUADU3Uwg+Avkz00A1MB4XIwyEBRnzIS5RArt2MzDhQDvN0wY6ACInDep+II/sILbWMJLRABrb0vK3QJbcMLEEMreiMCHzAGbaMNMSABDGQzsiQDBZ4R/hTc+YQ9TZIxByUBbHA64KMCpz00QaABG+Q1SKABQdA2dKACF1DhE8MGQePUzRIq9+Q1hAQDxw0StzA9N34wMWIEqpDner7nfN7nfv7ngB7ogj7ohF7ohn7oiP+e6Ip+6IDAABAgUV7jAxVgAYte6ZZ+6Zie6Zq+6YkeCQVA5biD5gQQCZxe6qZ+6qie6pduAeaEO4sAAQwACKo+67Re67au6UZAKbgzC0wT5S40PfB0MCBAQ11U7MZOQlQwABBgMD8URMf+7NAuQUgkAUpkM8kTAE4U7dq+7fVDRVBwRR5UBdw+7uTePl90PF4jDLliRiJxC2lU5/uCJTogCapA7/Ze7/h+7/qe7/y+7/7e7wD/7wIf8AQ/8AZf8Ah/8Aqf8Ay/8AIvCX/g6JBuM5JO6Q7f8Bh/8Rqf8Ry/8R7f8SDv8ZIgCZBAAKDuNZpEAJDw8Swf8i7f8jD/8jL/H/MDPwEVcE5e8+oM8Ac0P/M+3/NA//NCT/OSYAQREAG73usj4Qs97jX65AK+HjDbsFVHHuAXsAJz0Db30uU2gwQYEOZDMwdkbuYH40AywCwN4wlpZFYT00Ex0NIhIQ6xkgVeowcRMDJts1ItbjNLAwMyPjQ0TgEttVF7EuIZw+MQ4OMkBeRCPjTZgFB64DVScAFKThKUMD24rd8XoAIMM94wEAGEEOADjtD0AjgSoFAqk04xAOEZIw7ldANkTy0XDgPZreFkDu+ncgq5It4iAQ61EwId/i7CkAMik9i3KQbBsjjenQXYbVszBQEAbuG5YlGeDwFaEPumIgolst5D/yMNsZIDMRTfIrA79e0RiWPZh1QsoEADWRL1Q/PZb+UH8LQKPhABLjAHOE0bgSInXTD4pwIQtbpAWCFG3DqECRUuZNjQ4UOIESVOpFjR4kWMGTOmm9MiApJVyUSKJMYnwgcg2TSuZNnS5UuYMVl6giFhBqiRIpEp0oABhi+ZEnHF+PAgRBc+gcCMgJABhqagUaVOpVq1JbQgKyJciLIn0BoeEHzKKWfV7Fm0aWF+K+JCAgUkbQK1SVJBggsx39Tu5dvX77pycmBcgMBjTaA9WS5EYAEk2l/IkSXD1AQDAwQaYPzw6QICwocYufhCKwJDBYYIEiR8gBGE12TYsWVL3P9WBsYKDKorZHAho1O62cGFxxY3SQZuuxIyrIgx6eBw6NH3pusUo0WG5BhUwDCzTfp38FV5BYHxITtrItD8mtNlZgUAABjG3Cob3v79mL7myIAPocgmvfATcECMpLEkCAIAICAIS6Qh8EEIIfqGEyIegE+GOX6JcEMO1xnnljIwgE+FNHQxRzJr7LCjmQ5b5NCcO+qgxUUaH1QHkjogUadGHvGjpY47TuxxyO+aUdGa2Lx55JFpiHQSunSWNOZJKoNT5RFVqtQSNmOWBG5LMP2aZklvklyyyTDTVCvKR6ZU882zrswSTjqp6vKRL+vUM6YxHykTNiWZ3HPQl9h0k1D/RDWSM1FGMbozz0YjhajPPycLFE1JM23IUE07ZUgSLD0VFaFHR/WUUjMFNVVTTlfV9EpJXNW0VFkjRRXQM2uNtFVdGV20V0ZpBZbQWy3NdVhCeUV2T1DnXHZPYZ+ls1jJLpVWT2WvffNXbeGMttswqY3MWnDVzLbcLWFFV81v16VSXMjIdVfLc+d1sll7t2w33x7h/UtefomsN+AauSV4yH0PbtFfvwBWmMaBH+ZQXYl5TLhiCBnuy2GMN4y4YwINBnnDi0fGT2O+ODZ5wI9Xtg9flyMsOebvUN5LZZrDaznn6CjmWcCZfxbOZrVwFhpKKY/GT2SlpQu6adiITsto/6hl27nqyWDGOrqnt/5LarSo9jqyq8f2i2mzufQy7eDAPktstvkqO+60fKZ7sq7vNstts+DW+6y5/65Ka8H/yrtwqfi2ym/Epwq88aDQhvyswyeHSfGqGLc8psc3d8luz82qPPSVMKdKc9JZ6jz1jCRnXabRX7fI9KlQl/2i1W+niHDdYV+796Bol8p24CXKvfiHQEe+pdiXb0j4qIh3ftOkp/88VOuZ/z17jaAPSnruETo+fN7Dt6j58L2XCfzwx+fedfMjQp979WNin3v3s1c+/onmz75+mNwve/mzXvn4JxH/WQ+ALxGg9Qg4PfgdsCEJnN4CXdLA6T3QefuT4P9DKOg8C7YEg87T4PIi2EGFfHB5IWTJCJdXQuQZEIUT3N4MH8LClbgQeTAsHgdtmMIa/pAhONSIDovHQ+Cd0IYqRB4RM2JE4CGxdzIUIqmCWMWEOBEjUOydFHWnxBkysXhavAgXdefF2/mwimIEHhktYsbboVF2VFzjFbG4DjdWBI6yk+PrwIhCNvYujxTZ4+v6yDo1CjGQuhvkRArJukOm7o8dXOTtGimRR6YukqSjoyLtiMVLRiSTpNtk6BL5w0rKLpQQGWXoSum5SUowla9b5UNa6blXbq6TqPxkFWvpkFtuLpeWi+UBZ8m6XzYkmJYb5uROucReCjGZDFnm5Jr/CbldQhNPdxwimVKFKW6Kr3rhTEgx+XfM1E1zIdWE3DUb98wwRvOH6lQIOxvnTsSZM37oJB09E2JPxOGzcNmM5zbJiRB/IgSghROo4OAJSHnaMKHrWKjgGvo3fZqPn6GbaEX/dlG9ERSiBj1oR491UJDeLaPh26jnTKoqlI7zoA+lZERn+FJwhjOldBNpTUlKTpweVJxtEipCVsq9lm4uqELdadxoKkubonCpMSVqUY+avaRabqrkbCrbegrVn4ZzqzqVKTmfasyodnCs3Oxq2q5qvaxObq13bKvZvorWsHJzrlis69jeOr24Qm6vVeyr1856zrRKcLBCLOzW7orY/7zecbE/bCzW/uq8wDZusjasbNUOu8/EHnCzM+ws1C67vMwibrQoLG3THgvayILSm7iCKVfLGs7PajS0/FttB1urtNMiL7WF660Ef3u01+o2tr6crbFqS9aqCjW4xRuu4Ip7wOMKLbcs3W38rsu/7P4sudxdrjSbW62T2ja6M8WeUKv7t+/GL7w82y5Su2u++JpvvjmbLvDeq7f8tu+23Byvfcs7z/OOK73QPRR7Y1XU/94twPgb8B3727sI023CA6wwFguM1fumL8HxWjBbO1zFC+suw3HbsANPLMT6gvjAEh3xv0pM1xf/8MNwDTH9atywG/M1xzZM8e1WzLYWZ/9wyDOMMY9nfNMfbyzIhF0yCossuyOnLckkrHIHdwzYHv8vyimbMmO7LMEmg/nJUh3zzcpM2TMf8Mqvy7LZtvzCOPPvy5gNswLbXLQ3czbP8Zsz6+o8tjvvcNDmSzOf16zWP08t0KRdNPnae9BDey3RR6z0+y5NzkxvbdNR7LT+HvFg9/a5gpEO26RZW2rrFTp1ocbaqLsI6+ntGbWqBiGr3+Zq3+J6g6cu6jpoXTVbn1HYJvx0OI8NtWTHcdkxbDY3n920aPNx2j2s9h2vrbRsG3LbSSQ2hHm9Ql/3DdjGHfcUu43Fbx8t3JBs9xffXcdHKzbdi1s3duudxnKnOt//ot135voN3n/LTtaki7fQ5q3JhL9O18I9dxMLfrqDyzfiiAw4pis+xovXLuP63bgk7+3JgfM25MMbuYDXS86JU/fjbVx59FpO4Zfj9uS8TLl3a/69m3M459xs9K57jt+fry/oLh76HWPu35kLMun2W7qSm47FhYeu4T97OClLTrqiU/zoIvbTN4u6X5pl3XNb51nXXfn10D0dw1Fn5NQDWHUuXx3FHQc13S1pdwbiHc96h/HOtQkpsQL+goJXNOF1bPiCIl6vihch4znteCJDfqSSlyzlW2h5UmOeyXx3tt9V6fkcgv7Wokeh3FVselqivoiqVzbrO6j2zbE9Z27H/yXcYUl6a8MembJ/Iu2lbXs0a96nnJdt2WmbUxMj/4CuN7Lw00n8LRpf29LnX9hlPnYfO9+50Mcx9wmtfLAyn7niR+9zo99gmKMfr+o3L/sV7P7yw1/nzvI4+MVsfxLDPyEzP0YDPm+zvn7CvjLSPnEjQEvjv77zPz8DQBsTQCpzQE+DwNKTwFWjQCCzQDPDQFNDtf6jPwTzQCkDQTgTwViTP8gyQRpDQTJTQUFjwVxzQdiCQSiTQTejQUqzwWEjwQjUQTbjQUDzwVcDQmbTwODjwF4zQklDwmBTQmpjwgN0QnSDwlaTQnajQm6zQnhDQI5SwDdiQHrzQnITwg0kQv9I08Jf40J/Q0N3A0N8Y0N9c0N1g0OEk0N7o0OUs0OCw0N+00ON40OAU8MmBESVE0SDI0SSM0SFw0HlUkSfY0SMc0SX0z8Ck0TyokSks0SRw0Sc00QLM8AwxEKLA0WWE0WhI0Ws40QD80Syq5T2I78BdMUqoj4sE0OXIkM9MkOIg0Q/gkUZk8Xwo8X7s8ULxMXCQ8QrNMb/Q8YAVMYQZMbH80Oeg8YJlMYKpMYVtMbMw8bDKyreEybf2xzvgzpUBDlVtDlWZDpwHD1xjDxy9EVCAkavE0bW0UU640WlskdHwse300eTc8ZT1MYO5MYP9MYajEcrI0YnQ8gnVMgUZMj/H3RIL4NINZPILKTIGbTIJMTI5JvHzavHdgS6d7Q6kZQzU6xDk/TIHgTJKVxJPdNIR+PIVITJI5TJLqTJ8yPJ5XtJs2OqcySmlvxDoXy+sytKZ7JJo8NJdtTJKOTJOPRJ8+FHQ/NHrQJITBLI3iNIsDvKbEzK8VtKsDQlpxQ7qKS5k1S6lMw7q3zAYtM9milHZmJKyME9y6HLmLFLa8LLdxLLcRQqv2wnwEQcrJw1rZQrrhQlrzTHs/w9oEw/sqxFs4xLTzNIlyTMxmSlx7zLyETHtPy+tZS6tqS6txw8zMyexGS4xRSszrSlz/zL0DRKzURKzjzNu0vNxlvNFpzM//mrzGS8zGJrTa17Tc2KTWCaTcOszaYEzhcUzmkkTqsSTHrMTancQqrcQ9+8QejMQensRuqUrtFUx9KsO90MPN68vO4MwrlETtVSTmVizns6zMLRy8nhS5cpzPp0Tmwqz7lbR7bMzjfczkJsT2a7zbHEzqGkKgRFHvyEHP1cGf4MKPsUHONcO/gkLvmkJvq0UP8MzO+cxPBcyPF0sPcUUNMk0Dw00Ed80OLJ0NzbUOvq0HX6UIa6UIwC0NdTUfRk0UF00UwstnQM0PP8u/RcvPUMPRgltxHtxBKtyBONvyeNxSj9yCnVOQUdzJKy0XrCUYvSUb2J0MaZUJPxJ23wBP8sQAAEGAJOcBBuAgdckIMEMIAZsIRnONLQKYdeoAMLMAALoINeqA++egZLmIEBWIA4wAVwCCdp4IQhYFMs8ARt+AtwmIQYcIEMUA0MaIEYiINKFSJz4AQZMA1OXQEYEAP1qCJdII8VqIAIqABUDQJdwCJoEIPbyI3VgAEg4AQh+SFtiAPr0NUPcAEYmIRG3QtpCIIWkAAImAEkSAIagAAJWAEg0BAbAocycAEMeAAQ8AEnuIEIgAAViAFP+KFyoAMYyAAIqAAecAIeqACxgAE6IFQUookPgAAIuAEnQAINeACfGINknaFfAIJmxYwkQAIRoNYWCAI4RQttAAIVgAD/L5CFnKiFNtgNGXiGGRIHMWgBCEgCUkCGkRiGQQCBCoCBc2WtOYCBCqgBRiBZkSiGRqgBCWiBObAhmqgADSCEYRgJZCiFJHiAFjCIGXoGGcCONqCFnJAFMCBXIAhVsyiHIlCBCnCEnMiJU9CADACCgT0gOQBZP8jakaCFG5AAGFhVCfIEF4gAKSgGsk0GYsgCCFBZFIKGmriBWojbZPADCHCBnO0gcACCDNCAU8haZCBZR4gAFSgCe6UKmoCAROBbkSAFCGgBSuigZ6gJMKDcZKgFEFABMdBTvQEHGcAAH4BbvjUGH8AAGfha/RKDDwABW/BcMEBbjpWgSQBZUvBc/0RwALu1inQAAgzgAZmlXC2ogBiA3fCJgxXQAGHw3GQYhAiAAWzln04A2cP13FP4206QoF+AAQgYBOkVBg1QgTiQINPFAC6Q3mTYAQwIgnOwCryFAERwX1GoW9HgH3GQAQno3PKNAMyVoCL4gBtw32S4gQ8oAgmihBWQgJ+VXi+oABl4jvjJBfEVBfdFBAiQAbWdClyAgQeQhePlW2QQCy1IXBVeYRZuYRd+YRiOYRmeYRquYRu+YRzOYR3GYVDQV0UoYb7NAQyggR0uYiM+YiROYiVe4hwOhnZVAwRWAwl4gGBgYiu+YizOYi024mSYgeJFYEV4AAkAhS0uYzM+Y/80VmItuAwgjltZ+FtcsApNAFlgQGANgAAOWBI93mM+7mM//mNADmRBHmRCLmRDPmRETmRFRuQqCAAIIGP3dYIKCIBFrmRLvmRMzmRN3uREtgMAqAA+QGA+kAAAsANOPmVUTmVVXuVLDgAJgAIE7uEAmAJWrmVbvmVc1mQOgAANQGBguFyoqIpOcAEI2FvpRQYKgIASyGVmbmZnfuZH2AJHbgQERoIKSABozmZt3mZMxgMAkAA9QGA9IGU84GZzPmd0/uMEkAAkQOBGgIAA2IJ0nmd63mYSaFcEroXLXVmq4AXxDQX3pYXL3QR0OIeCPmiDTmiEXmiFbmiGfmiHjmj/iJ5oia5oir5oi85ojN5oje5ojqboanDZUHZfEPiANPhoj05plF5plW5pln5pl45pmC5oczjdKEDgKHBdc0iHmZZpn+5poP5poQ5qok7odEiD2UVgv4UBbChqpx5qqH5qqY5qn55jCGBa6Q2Fun2NqjDdCoBi6R2EDn4M8CKCDDBe6e1eGLgFCaKDFbgAYpBeYpBVOpCgEIaA7fVcHsgAIpCgaJCBCCBf6VUDCmbeqHDeCjBmviUGEVhg0r2bYYYAgPZcJ3DdAOGfX4iBCBhpyjUJ65Wgb4iBC3CCNs4JUADe78WuAhaBuKbcWpDV9DWLa4gBCfCB1o5bLZCAGAAK/wkShyDIABBQbLINhL8NZgkqAxagAA3mW1KIgBUwAxTSBGIOBMqlBRDoWgvmH1+oiSygXGLwAd2+BrTghLxtBbKtBSe43EkgWNEGgcnOCWJYAwhggTF43PjZBiD4AAogBCBGBkKgAJTwjg4qhzFogQdgg9seCVAAAQyIgett67aFAqzOiVbIgepN7bNQB0qwDAiAAkQAhVBQhORNWTP41Q7CBRnIVx8IBBBnhDUAAQEugsvuIGgAghV4Vj1wBFFwBD2gAecGgg8G7SJoVg1YA0YQhU8IBCR4ANCI4xkyhzhwiwrgAkUABVAghCiAgAs41h1RC08YVgh4gDDHgBWQAf9LmN8f+oUg4FZn1dcKYA05yG4UygYxcAHz0Nd2LVYxUIkfEgdh/YDLiAAxxgAXCALe5ixLiAHcwHOxaAEZ4Oe92AZLsPEAIIAMkAE64PMqEodOIAIYqPQIiAEzyF3CygUxkAEGMAAGiAExyIXHnpxfMAPNJoABaI1OkHMbugY6SNoBGAAVAAJLEPC/MIddsIIqeAX7JqxqsIIpYARcx6Jt+AMq+ANhDydxYAQqsIJq6PJwKodXsAIr2AUTn06ifARiKE4erT4fRVIgbUQhHUUiTfddXPfTS9LKW9LVa9LeIVPEMdORqdAcDdF8kvd+pPfYs/fPw/fa03d729Lr7FL/hE89hT8+hr8dGd1LGoUvL/0nMP0oMVUpgs9Kgx++iJ+9id++io9Ehy9JBlXKck/5YazSYrzSmMzSTZT5iKT5nbT5Ulz5oGz5snz5eMf5jdT5qeR5Dwt5xRz56yv54jv5BoR5jiP6mzR67UT6vfN5ygR6yxT6orr4/Mx4ANt4hep4vUG7mOH3wvF3kAH4MBX4+7ROlof4dr/Ed2/FoU/RV09Ap88+qD9DqU8dsJdQsZcwsqcos78btHeZIu3RvR/Dvl/Avw/GwA9Lqn9Kqy9QrBeiwS/Twteww/eos/94ulF7wWH7jnF7j4d7h5L7n6f7BlWvyo87pXdNpuf7ug/F/7uHx7w3t9uH/Nxfxd1XyXjX+uDk+uH0evK8fLXM/Bbd/GvUe+cPUujPPOOPTuQndwftfYF7/F6M/DKc/Hycfc/p/H7/fBYL/cSnm8VfGdP/G9THGNUffdbf0esHz+wXT+U/KPNfe/QHiHUCBxIsaPAgwoQKFzJs6PAhxIgQpz165E0iRofeKk7L6PEjyJAiR5IsaTBdRWMmV7Js6fLlS1WPVMGsafMmzpvGKqbL6fMn0KAMKVoUanDjo45GlzJtyhLlI5VOp1KtClGmJKtat3Jdt/NRz65ixwYletEpUqVk17J9CVVq27hyR8qkOfcuXolfw+bt6/eg2alp/xIu/P+2MGK/kmYmbnx3r+PIbAOj5Sj5stjDmDdrxcr5M1XIoEcbpdx0MOnUPzWrbn2zruvYNUXLrm3SNFPUtneLZM37t8fFdoET18uzOPKHuJfqTu4coe/n0gnCnm59IO3r0pcbba7defTvyT2Ll569fHHuQr2jBx6+/W/h8JOfn79bfVD29mu/3y+7un+81Rega/gBpR+BqvWXYGrkMSjbgA+OZuBPCEr42YIXcgaghqlF2CFmFPpkIYiSZVhiZPKhONqHKzYmYk4kuojYiTMW5qCNl7WY418w4iQjj33VGGReHBKZ2I5H3uXjTUAqKdeQT8alopSFJVnlZBWddZplWP7/FaWXYxkZ5mPHkZkXkzY5eSZXYLLZ2SNZvVkmWHPOlWZNa9pJlZt7OkWln2tdGahTeMKkJ6FL9ZmoUGMyqtWgjwpl6EuISrpaSpd2haOmVUXaaU6UumQpqDYtWmpNjqJq1KerwiRqS6S62tKps7IEqK1CtZrrSrCyJCuvJNUaLF1xEqurmcf65OtKwCr70bDPeqSqtC7tWq1HzJrkLLYRRdstRLiC29K1406kpWBdmktrpuu6RK27IZUb70LalsQtvdC1m29JnPIrb7L/hmQvSfgKPNC3Bw8krsIYzduwQASPZLDCCTcML8QNPQyxxCJRfLDFCvubMUQbN9xxSB8L/xzywRiTrJDJCqMMksr/siwwwy9rHLDODs38Uc383vzvyD0vFPPBP3sUdL5D8+uy0QUhLbDSGTFNr9P55hw1QlP/WzVGV8ebNb1Qcy2Q1/yCLZHY7pIdb9FnG5R2vmtH1Pa6b7u7tdwE0U2v3RDhba7e65p99t/xBv7Q4OMWbm7cfWPHs+QFLa6RupWftK/mBh3OdeLuXt5Q4+A+Pi7flYe+7ugMld7t6eBGrjrlna/T+kKvYxt7t59Hvbq5uCuke7W8Y5u65MCPK3xCxEtrfLW+G608uMwj5Pyz0Es7e/K1d279Qdgrq/2zyPdNfbfgH5W57QKRr6z0PaOPrfoFif9/7PvHcn++95rXT9D9iJU/YsVPZ/Or1v8GEsBgDTBY5pPbAaWVQIEskFcN5NX+INi/yk1wHRXM1QVzVcCXRfBZHfygrUJoqwciboOSOyH72qfCWY2QZCVUFgyT0r6CzNBVGWxhnXY4kByqRYg9XBULQefCvhFRiAjjnBBrmLEbHquJTlzHEVH1QyUG0YlWdGIWSyVFiFGRWF80IhR3mMTfLVFuZ9xhGEG1RTZ2UYhvlGEa2zfGhpUxWHe0XRw7tcbptfFsf+xcIDW1R4X1kVeH1FwiLzVHQtZxh4+sXCQlNUj5FZJrl5RcJh+1yIM1Mlef7FsoGTVJTlayfaeUWyr/EzVKgZXSVq88WywJtUkDdjJqt+RaLgO1Sl620na/jFow/TTLf9VyVsc0WjL3tEsS9tJoz+xZNO20TH4101XX1Fk25zRMahbze+iqjA6vGM43TdOG1ezZN1+2TjZtM1/dXFU8STbPM43TneX03zm5lE4w5tF29aTXPVGVz4ztk0ztnOI7dbZQiDU0TP2E6D85GNDcxBCQBe3cQeOV0FJNtGEV9dJDyRjRl5W0Yh/VXEjdNVJQtRRkL63cRVWa0RdulDkdReRNJZdSPq6UZDVdWVD7FtN1zbRTR7VZUuWWU6LulIk97c5PIRnVsy3VXE3V1FOFtlWuDZWRRc1YWJs2/9aoTdWsVXXjVdeTVUyu1WhdHddXL5VWrNW1Z2Ul5Vk5Ftf8zBWUfdXZXcGVV0ntdWyHfVlbAftWQw72QIVF5WNJ9ldaBvZkla3QZWGZ2YwltluLfVRj3TZaiEWWs5P15GdHFFpcrvZijLmiVzors9jGaLbArK3CNstM3SaNtz/yLTKB2zJj4fa0jEpt3pQrsNJiy7mJgi7hpPsv4XKTuFQzbpOQC03tPu22V7QuobDrOPLmq7XDfa0vwasm8WKTvfTirj29+zX55om+4LQv3MzrRPQGSr2mA7C73Ntd+FqTv4fyrzwRbDgBC5HAfjIw7CRsLvwiVL9qc3ClIKxPDf+PS8H5ZTA8QTwqETOUxLKj8A4tvCcM787F3eKwSD1cNxXHisUUtTG2qFstGduJxsUDcvSYe14dA47Hv/KxSZEsLRzLlMmKc3KzoOzSqOB2HUKWFpHnZOTnSflZJu4wiiWK5W1p2aZcxu2XnxXmN405e2VWFpWZamXRrflebUbqm6945hynmaV9LtifoRpoJ8ZZWXNmU53Hd+dj5dmre2bdoSeWaLEuOoowbt+jzxRp/E2agEoe8KWDl2mPbVqtnVbjp20XajKNWoClDlajjzXrMNWagbfG4KkrnOrlrTplrebrq/UY687t2ku9tuCvc1VpvA67esWm2bEdm2yDBjv/xtVO37WBlm3Vbhuky9Zcs7H0bBBGe4Xnpl2hjRrupY07uuWG6bu7F2+0zttq9c7uvXHabVB/m379Dtu/1xtwoeabf/sWbFHQWUQ4tpuGDdfgwz0bcYFOHI8LV+rAZV1wBB6cbQk/8MelenEg8sWOJb/byTOc8rNNW7Ejl+DLBRfzGs+ca4OucsZ3u3GODhSNPWfryrnYckvmnHE7P/LRjVZz097chE3HXNEpHvWe5ZpY6a7SulNYcR+GnNlVx+HVSfd0Mm9dZ1Ov7tmrmHbXrd3ObYds0um4dFfOPXd1l/TdSfZzPQe9uEP3adY9DhdG552Sezdm34f3d1IHPmNv/x9y3M0Y+eZN3taVZ23Z0Z15P27+ep339edtOxxUF/67h8dq4j2a+uA2npWPN+fr5Rp7oM5+uasXduv3m3vC7l6rvZ9u6OF9e4AO37LFp+vxt1t7Yi5fo80H7fMNG/3y/t7bwf/w9WWbfcxuv73J13f1eRr+3o5ftOUv2/TJmX6rrv+47aft++N1eTCP3pGlD9/pQVv+Jdj5Odz8wVX9hdf9/dYATlj3Edz37VgCztcCJlcDblj8+dMBUtYE9lcFjtcFllgGYtQGwlYHPtgH1lcIvpicLFkENtkJhlgK/tcK3tgI6lQJxlcMsoQ5QEMhkAAJwIEvlAMcSUMnlEAHaP+BLoiDEGkDLhwBBxwBLmiDEImDLmgBB5RAJ0hDDpJMOfgCHJBACRTCM3RhSZxDLgRBDHwABEBABsAAEHQCEWrOL4iBDLRAG2IADMQAJYBD51xDHMQADFAABFCAC8RAHFxD54ADJcAADGAABERAC8iAGPxC55RDJwABDLAhBKgADARBLpzDT4hDGcBABkTAA1xABeQhDIjBNkiOOTSiCrQhBVxAG0pACwABNFTOLciACkhAG14AIUrACsQALlTOMwBBC0jAA0SAKkKABHwADFCCOUjONhTBIz4ABFTAKj6ANJoBE+KEOBRBC0RADSiCLSRDMgxDIyABBKwAELwiLsn/AQxIQAgEgiyoIzKEghZAwAfIgCXKDSc8YgW0wSkggzqeQhtQgB5yQt/8whpCABeIQjGooywEAghIAAzIgTrIzTYEwSwiQSMMQzIggy0owg1AQAsUQTjWRDqkAR7ugTGoI02WJCK8oxjMYdRoAgw8QBQQQ03SJCloQAYAARVyTS/EgATYwCwEpTrOgg1oZC+cjTYAQQZogCg4ZTIQQxRAAAxowtmUQxHMIiIgpFO2gSSaQUfWBC/EQAS0gVbSJCF4ZSdwjTYoJRKYpVaeAjHGAdeUQxB8gAgAQ1wmAzCIwAcEgU72TBysQAScQmEiAxJUAAwcpdF0Qk8iQknGJVrG/wAv2AQRCKZeaiUy/AAGyMBivswkOGZTFmYytIFGSkPU4AIMQIAjuGYyOEIEwMAxGo00yIAE6AFuykIErMAkRE05yAAG5KVrIgNiFkFNXEM9BgJu5qZX+oLRlAMQYIAUVKctqCRYGk0ZiGZ1ImYZ7CQe1kJ1SgEGAEFqZkwv1KYjjKZTIkMgSIAMKOJL5EJPrkJ1EoMbvgE1DCiBFqiBHiiCJqiCLiiDNqiDPiiERqiETiiFRmgsYEAEEEJ1JsMNYIAQVCiIhqiIjiiJlqiJTqgzYMAFgAF9auUXYAAGOMOJziiN1qiN3qiICgEG1MCG3uQHxAKOBqmQDimRlugbfP/AAwAlbq6CV/amS3ACHgrDhmoABHBARVwplmaplm4pl3apl34pmIapmI4pmZapmZ5pmVpBADwAKGyoE0jAAaCpnM4pndapnd4pnprpHQCABOzBhvKBBACAHeQpoRaqoR4qotLpAUiAE2woKDxAAFhBok4qpVaqpd5pB0DABWwoMLyjQ74ElEIAYVanLZLApZ4qqqaqqj6CmkJAm1bnZCLAqs4qrdZqne5pcG7oHgQqHtiqr/4qsHIpAjDqhn4CBATAFQSrsi5rrWaqBnCqStblftYmZOJmp6qAIHSDtm4rt3art34ruIaruI4ruZaruZ4ruqaruq7ruXrDLqhABAz/wobOQAY8AbveK77mq77uK7/2q7pawwdcQBdsaBdgwAdYg78mrMIuLMM2bL4+QQbMwIYOQifugsNeLMZmrMbyqyDM4qi65ik4AAzkAkxkA3DyQXUqglcGpM6YQxBgQKPi5ixEa9TEwQeAwIaCwAf45WXiIS1UpxO0ZzX2DERCwCJUJx9QgAxYpkuIwc1WZGTmAAYoZtRYAh76p2uCAX5mQ9ToQm0qAm4mwgPAgC5ETTYopRosqUpaQtSYg1XmQIvWJDHorBjYxC/U48AWJh/sppP2zDYo5w1ArVaGgkrSAdeYQxF8gAa0plbOggZ8QBGIYtTQwQq4amEWg9TKwDdw/w1tQgDKxiUydAFlsixMzEFtdoGS1iQy6IFKlsHQRs0t1KMPpKNTKkIFJObmcg00xAAGiEC1BuUpiMBp7iLXfENgVsDROmUt+IBGesLZmEMZVC4fCC5NEkMXQIALzAFOlEMZuIAEgIAfnIIwDMMqDAINqCQR+OHZqEMjSkAFrIEo2AIxyIIi+IA/AoFsyg1/QmIWOEItEEMtOEIWQAAGxADJyk00AMEHRAASKIIsEIMtiMIacOM09g04EIELRMAMDMIqDAMwnEIfgEAEuEAZvOdLlAM9fgAwtiE0viE4So46dEIMrMAqsjABuwARcK3k9IIyYoADRIADBGMuYqfkZP8DETwiKrJwBayADHSCGfILKcKABlDAD/9wNMIAHZhwTfiCGMxwAzRABsgAEejCEwuMNAQiDDSAA1wAHMrhIk6CDMQABARxDMjAJORu5YxDJsZABTiAyCJi/mKSLhCBDHyAH6tADFTiUmQDJqDACRgCFwrRN+SCI7PBM7yu7YhDLzTBCTRBL7Sk7ZjDM7SBCaRALuCx7aiDNBjCCaAAJugw0XWc7C2epz2gyL3gle1gj81ghNVgkN0gVZWxvOnyk/HyiPlyktmy2eEynxFzlhlziyHzlAGzWwkzvzkzm0Hzj0mzmRUgxlkzxG1JLOPWSWFJ1wXL10lJ2M1KOVfJ/sn/Wf+Z0v+tjzZHGTfDDzVLFjhrnDgjnizzHi3v0OBZGjNjGjb7WT1vWUC3zzs7Wjzb0jzbTwCy2z3rTz671j4LXT/D3j8b30JzWwuyXkYb3kbrXkdD30ebmzKLXkGr2kEjWkK7WUprTkPr2kM7U0QD0ESLXUWbWkgD30i7XkkT30lr30wL3EorH26ts6u0s5TUtNfdtDfltALtNDuP3aqcM6+k85Mw9ao49ZMMNLW1NLG9tKbFNKAdNcMlNfotNVVTkFU3NVZr0UW/V1AL31A7X1GTn1qD3E97312DX15j3167X1+rHFsboFubNauhtaIdNs3V9YIFtgQOtvgVNv5B/7bPeTPLLbZlsx9mM6BmI11if7Nnp4tjc9poS51knxhlw+Bn219oW+Bqc11ro9lr53JsK+Bsg2BtIxZnK91pSxw5z3WpQDU6SzU+vbUHxfVXG7cc3Tah5XYz7zYF9rYK/jbe/TUEUrdBW7cHYjcNarfgSTfQebdLgzcKincvk7flmTfhoXdZq7cMsvcxuzfocfcty7e1Mbaxpbar4bfqdRlXK4lXowpYKwlyb7VyKxRzo9BV9zSuwTdB8ze4+Te2ATiyCbjIBLfeDTfHFbeE88qC50qBH8mBl0qCH4lWm3iDk9SDOzeCQ3eniLXNkXV/0/eKabi2cbjvEfiL01SM8/84ufk4zlD4WFu4wWG4uBG5vRk50Xi444H4OKsTjSsSkt+4kpMck9ObkwMclPNLidvKiRNJioPKihNJi5N5kDvVkNt3NIe5+en3Mm85znW5v325wsn5fWU51eH4hev4LsP5NvN5gJV2Z1/RmXdKmgeJjf+5nVsdniOcnqOcoRMgogu3or+5iF/6uoz5rJR5kCy6pjQ6jzw63AH6kgt6MRO6PXs65Ph5qkc62k26yVW6zME66sg65qk6l7P6M7u6QnfZmod6m4MVp1v5iIuQlNselftzp3cZqLuKqPMIqV+KqedIsVP7setVshPUstsKqvc6rcudrcMcrvOcrr9Ypn//+Kafu86lO9Stuw22+5S/O7Bns7DLNLE3O/U9O0dHO5zxOv/5+p3nO0Lve1pLO8HDs8FLOsLDtMI/dr/bu7PjO2pPvGr3O52zdLlrHrw7nbyzHb0fT8M79MPXesSftcYHeMUDecqb+8o3dstvOMfD/MeTXshjXc33+Ms3V7cz1rcbXclXy7SvSrXnyLVLSrbbyLgXfM77386r3cjbXdFvz8nbdMyD/Mz/d88XOcNb/L9jPHEr+9V3s9jLH8CbtMALmr+rPdmHuNmHPc6vPVG3PeOlvQbavV7jvad1vFLHfZWD+9njs96TIN8Ttt/D2uHjYOJf9uIrG+C3teBD+9wP/3zjB/Pjg3bk287Ro0rS28jSP0rTz8i2I33Qo9bQa13hW/TkK3blB/zlX9Hnl0roz8joM0rpu8jpg37qP9fqK97NA/3W63zXZ/jXP/nPu2DUy/PU013VA17rE0vtg8rtu0juJ8rur8jTO3zzQ/Tz+130U970T3jmV/Pmy3bnd071d8r1r0j2E8r2o0jv2/7vX1fwz/LwM3/68/b6A8Q6gQMJFjR4EGFChQsZNnT4EGJEiRMpEpT0SFVFjRs5dvT4EeRHY48epQt5EmVKlStVTiPpjSVEbySnxbR5E2fOlOlIGtP5E2hQoQhVYRx6FGlSlSNLKnX6FCpEl49gKp35qP9mVK1bt/J85JNrWLFIi0oaexZtTqYm07Z1G3Jq1aRXs761e3eiV7B4+fZdWDSjX8GDB64lfLhvXKd0ETe2q9dx5LYXA0u2HNbwZc1RFVuluRm0U8ihSR8tWxp10MypWdvsPPdza9krR8+2fRLwbd0gV+/2TfE1Usa/iUusXRx5Q8rJmTPs3Rz6wOBHh0e3vu74dei5tVt/3h359KHVwSfPXr74afTMv6/fLV4oefe+z8/Xvdw+8fb5ZcMPKp+/2eoLsDXuCLxtvwNL8w8oABVEbcAHSVNPwtYSrFAzBn9yEMPNIuzwMgNBJO3CER3TUCcOTYzswxUbw8/FzUqMcTD/FHNSkUbCWsxRMAp5jGzGH/GyESccheRrxyPvElFJwoJssi0ibzISSreSrDItGLH068ktxZLSJiq9HOvKMcNi0sy3ukyTs5cWi43Nt8qMEyof6UxrzTuTAjMmMfUUrac/JzNKULTyLDQoPlnyE9Gh5mw0KDQhherQSXFSdCVGLdXp0U1xstNTpSoNdSVMVdKU1Jg6TXUlSVlVjSS2Xj3K1JRQnRWlVXENSctdgRrV145qRenWYD3S1ViOQE3WJmCZBc5Nz7B6NidkqZ3I1WtRclbbh4Y9qdhuI7JWXId6LfckbtFV6NuQwl2XIXLhTSjbeTVS196C2gXp3XwPktdf/4KWDZgifAne96N+CR4I4IXPXXgigwNG2COFIW6Y4HohdkhifynuyOKFMQ544I0f6jjfjzkKmeCR/dXY5IVQtlfljVgO2OV8H46Z41h5nqhmjW72N2d7S/5ZoZnnDbqiofMtel6YkTZIaXiZpshpe6GGd+epE6p63asnynrerdeV2uvCfE67IbElIhtes9E9mm2CwEbX7YjgXlfucruu2+61AU8ob5ngHPzfQBGnl9DFD7q73MIf2hvdvsWle3DIxZXcIcrLtbxbtNnWvFvOG/JcXNC1/Rtx0rU1nSHUu1X9WswBd/1a2BeSXVvaqRU9bdyp1V0h3q/1/VnWMxfccf+CiE/IeGqRZxZ4r4V/9nmEon92+mRtr/t6ZrM/aHtmuzdW+duZb16g8Q0qP9nzg61+6vCTdb8g+I2V39fvR1+fffgjiP6Cxb9d0Q9p9jOWAAdCQF8ZEFfpAx8Am8dAgThwVxCclf+CR0HHWXAdGMSVBl+FwJ8pMFggFOGsSMgqCf6vKexzXrRgMy0ZMkxxNxSICXmGQl+p8HAybGGqOGg9Dy4OiDbU4RBJ9cIOxlCHSazLDZkYKh7GzIe7kqIOBVJFTxWxfkdE3Ba56MVNXdFkWcQVGZeYQx060YhQvCEbqejGG4IxgWIcHB2FaEcZonFjapwVH9lnRkvBMYxylCH/IZtnyEkBEmKCfBUjHedISOHxhHoEHCUXZ8lGITKPigwgDYUTxEL6kX2QXJgkWcVJxHkSUZjsoSbr5srBwbJQqiQYK1NlS8DhUlCgzKQoK0hK6piykahsniyxSEu2+bJuwPyTLgPGS1JBk23S1JMwZ0nMDxpzPMispDIdR01/WTNU2EybNu/EzDQ6M23q9Bo76cTNZnoTieCMjzg7Sc7FmTNf6PSUPKdGzzi5M5Dw9BpBkWZQNgHUXgLdFEN/5tA02fOd+ByjPv/Dz1f6E3EIjaRCp0ZRnlnUTBCdl0QtZdKYoXRMGE2oRvfI0QZ59JYgHZxK4cXSSbnUZDD1kkhX/0lSpAF1Y0LdkkxHStNN2nRDOP2lTgHH03X5FFJIvRhV60bUXRr1Z1oVGVfZZlV0YbVRYm0ZWdPG1KI6tZZQTZFUo8lWr3m1mmDlmVpxZtepmbVcaEUUX4nmV6S59atwfaZcb0TXbBr2Z4AVl2ALRdinQZZneD2nXmNmWa1hNmaIzati48nYIjl2naA1mWS7RVlBebZsqt2YZgPKWZPBNm6yhRhrteXaP+GWb7p1WON06Fs9Abdyws3YI8zCxXUY907I/ZxySUbcG0KXThYUhyuWoIRMfEOH5fgFE5QQiGwsMRphUEIYoiGr5qkjG4FQAhN+UQ4dbiMTSliCK8RxFP9pxAEGKigAAzIAgzL8onnfmIQMYFCAAkhABkXIhXsBJw5OBEEGDWBAA2AQhE70t5O5KEIMJMAABsBABpTYRvOeUYYYZMDEKohBHKQBlHRYQgYrkAAEeAwBCagABnOw7+ByIYMWYKDHPv4ADMQA3sFFIwguQHKSMdDhaCBuG2OAwQck8IAH8BgDLZBBLhBXDjlsecc9rsAKZGAJCrMkHQCugATU8IlVtAIUbdCABFwwBhCzjRMwQHIUFHEKWYjCDzSAgAqCoA3A+UIGH4CADwhRClmcYhBIeMAHZOALwGkjCC2AAA38IApZoGIRUYBAlTlR4TG4QAIaaAMoZLEKR4D/Yc4wkMObVWIJGEAgCbZARjKITWxieCECuuZ1zHQBAwnQgBXFLjYyArFoIvx5ateQQQY0AIphSxsZjtAABmRwDbaJowiSDsS3pb0KGkgABrpgmzrk0IIHfGEY0ia2LZAAARhY4ibQcLYT2K3vZKjB3/L2mjiAgIEZBMPgxEaGIiDQAoDPcwwquMAqklHwYq9izWNY9sYsIWpFRJzYwaABBoCA7Z81GwJrQHnHoQBvaNikDB/QADBm3nEaZCAII1+YJkR9ip4nQwsVgMGKp9aLXw/i6ISIQAx64bVtxKACWjh6KR7QAk3MMwgfoIExeg4MDaigDDHZhgwk0IajIwMR/xCQwc0bGgQMIOHoyViFA2DgCa/JQeMeNzgyLqCCOXjNE7+O9tGRAHShB0zgEUBE3tsggRgwXSUwL0Xea7HoSpAD9KEX/ehJX3rTnx71qVf96lnfete/Hvaxdz0zWhCBQOQ9GSL4ABlk33vf/x74wRf+8GEfjiNnAfdZCHM4iN985z8f+tH/PRkyMAPBGzwQEmgBM6Tffe9/H/zCr4QKHmCLvJfC37xgSSdEbf68VwACJCDJ/Olff/vfH//51//++d9///8fAANQAAcwAK0gAB7gE3APCSoAAQjQAR8QAiNQAieQAgUQDwBAAvQA9/RAAgAADyoQBENQBEeQBCEQAf8kAAmuT98cAQICYAtKEAZjUAZncAJJ4AEuAPdsoeI6gSU4QdSEAfc0AAI4gAaL0AiPEAkfwQAhABRwzwkkwACSUAqnkAoj0A4AoAL4APf4IAIAwA6qEAzDUAzxLwAqwAlwDxRasArGkA3bkAo5IAI0APeCoeJabSVwQdQ47uiIYdXQYBn+EBADURAHkRAL0RAPERETUREXkREb0REfERIb0RTgb/Ly7gYwAAciURM3kRM70RM/ERQfURkkoALAAPdwTQKUIRRXkRVb0RVfkRNxAANuAPfirgJSARZzURd3kRc/8QwyAALI7uhWwd/IbCWkgcSg7ujSEAbUD2nKIQb/MEDrjg4YKu4SvGYMxA73aOADxsBrLmEFIIDnjk4LyG3IfoYXYOABmvDoBsHfrmwl1CEIMsAGVLDYoIDcXC5m5GAFKsD9UI7apo7ukCbxIIAUji4U/M3vpkbgIOD2eq4WKsDwFi4ao+DtbuADgkAdYiIXfu0hUY4FvS5tpIHtLHLmWqHw0m7hGq4GiGHmiKEGWE4fTUYdykAFKqAVei4KKkAGauwbRc0Req7aYMAYWcIcxiAcCeH6PuECPuDa2IYSXCACwOD6VkEEyM3c0iYXSAwJxlHfgAEJLK8op4YkMWAE9FDfkAEMIMAFKOHciOADKoAd05IQIGAFxuAcbuIb/4IgHJMgFNjtFLQAAj4ACLIybcyhDKSyBhjBJYlNFtrgAqpM4aDS2UBgEGqh2GphEEAA3tyybnQhGiugDWTh24ZhEWwgAlygDMyhbq4BCCRNC07h25AhFJLALoMAHHJiG8RAyiLAAkaABi7Ax1wgCHyybsoBwCQNAmbgBoRQAtjMGaPpEraMAhwgBGwgBHjsA2JAEx7PZHghxyggAi7ABkQgAiCgwOTgHNlGGqIM/i6ABmigOjHABcQA83BiHDiBwVbgAz5gBWAgBixhJucpF4IABvjTP11gxhxtcZ5BDGCgBfpTBcRMDJ7BcbIhDmTABfhTBVYgBoJgwhYHHCwhBq0CrD9Z4AVkgBMG1CbEgRcmIQ7MYBJy4T4XR7wsIQ7KgA464bwKKRo0YQ7KQA40ob1kKBs6gQ7KIA4sob5kaBtygQ7MIA4mgRdY1LmuFEuzVEu3lEu71Eu/FEzDVEzHlEzL1EzPFE3TVE3XlE3b1E3fFE7jVE7nlE7r1E7vFE/zVE/3lE/71E//FFADVVAHlVAL1VAPFVETVVEXlVEb1VEfFVIjVVInlVIr1ToCAgA7" style="margin-top:1rem;" /></div></details></div>
```

::: {.sourceClojure}
```clojure
(view
 '(GridGraph [5 5])
 {:folded? true})
```
:::

```{=html}
<div><details><summary><h5 style="cursor:pointer;display:inline;padding:0.5rem 1rem 0 1rem;">GridGraph[{5, 5}]</h5></summary><div class="wl-results"><hr /><img src="data:image/jpeg;base64,R0lGODlhwAO/A/frAB8mMSIpMyIpNSQrNSYsNikvNyMrOCMsOSQtOyYvPSkvOCYwPiswOC4zOi80OzA0OzE2PTM4PjU5Py83QzM4QDY6QTI5RTY8RDo9Qjw/Qzk+Rz0/RDA8TjI+UT1AQz5BRTtBSTxCSz1DTD5FTjNAUzVCVTlHXDtJXjxKYD1MY0FDRkJER0RFR0VGSEdISUpKS0tLTEtMTExNTUFHUUJJUkNKVERLVUVMVklOVkZOWEJNXkhPWkhQWk5SWEtSXUxUX1FRUVVVVVFUW1JWXVpaWl1dXUVPYEVRYk5WYU9YY0FRakNUbUVWb05bb1FaZlpdYVNcaFVdalZfbEpbdkpcd0tdeExeek1ffF1gZVdgbVlibk5hfVljcFtkclxmc11mdF9odlhmfGFhYWVlZWRlaGpqam1tbWJpc2NqdG5ucGFreWJsemNtfGRvfmZwf21ye290fXJycnV1dXp6en19fVlvkFpwklxzlV50l191mGZxgGhzgml0g2t2hmx3h2JyimZ2jm56inl9gnF9jnN/kH6CiHyDjXWBknaDlHiFlnqHmXuImnyJm36Lnn+Mn2eApnaJp3mMqn6Ts4KCgoWFhYWHi4mJiY6Ojo2SmIuSnZGRkZKSlJaWlpmZmZ2dnYGOoYKQo4ORpISSpoaUp4aUqImXq4+ZqIqZrYybr5CaqYGWtY2bsI6dspCftJ2iqZqir5Ggtp+ms5OiuJSjuZWkupemvZiovqGhoaWlpamqqq6urq6zvLGxsbW1tbq6ur29vZqqwZyqwJuswpysw52txp+wx6Cwx6GyyqSzyqO0zKu3ybO/0r7CyLjB0b3G1cHBwcXFxcnJyc7OzsHL28TN3M/Q0s7V39LS0tDR1NXV1dnZ2d7e3tLZ4tTb5dbd5+Li4uXl5eDk6urq6u7u7uXp8PHx8fLz9vb29vb3+vj5+/r7/P///////////////////////////////////////////////////////////////////////////////////yH/C05FVFNDQVBFMi4wAwEAAAAh+QQECgD/ACwAAAAAwAO/AwAI/wDXCRxIsKDBgwgTKlzIsKHDhxAjSpxIsaLFixgzatzIsaPHjyBDihxJsqTJkyhTqlzJsqXLlzBjypxJs6bNmzhz6tzJs6fPn0CDCh1KtKjRo0iTKl3KtKnTp1CjSp1KtarVq1izat3KtavXr2DDih1LtqzZs2jTql3Ltq3bt3Djyp1Lt67du3jz6t3Lt6/fv4ADCx5MuLDhw4gTK17MuLHjx5AjS55MubLly5gza97MubPnz6BDix5NurTp06hTq17NurXr17Bjy55N+2a6bM9+XTNXG6U2aL6kiet98ls0X9G+ETcpTpovaNqWlzR37dezbOlegrsEREYMGDGATP+KLt1juU5FZMiAoV4OtPIe0+kqAx58jDG8ssPnCE1O/e9FeFLOfhxpMwkQ38UQQxCXgMMSLjK0gIEEEEAQAQYtxKCJfgRa9EwQMHxAYYUVqADDHMN1aNE2YsCgwogQSGCiGNuoaJE4/qmAQYUxZgBDEM/YWFE6msAgIY8SZNCCDLeoZAkMOyYRSCOfDBJFBBLAIAdvQkakiwwZQHADH4yAgkgXF0jgghgOdglRNECsAIEIbCgCiiJriBDBCkBI4yZE4ojRggQXeIEIKIzwcQMEH8igy58PneOfBBFEMQgojQSCBAQYxGAJSp7AIIENqyRj6qmyIBEBDJRA2tAzMmD/AIIjp54qDBgQtFAGl64mtE0QKlAQCDK1IoNMIBF8EESNvSZkzhhygiFMrckg4wgIGMgQZLMJUSIqErOcSmwyq9yQpScmZQNDBTwQQ624UGT5C7cHlRNEBiDI8u6pfEAAAyf0HmTGChGEsq+poUCwghkBG8QJDBD4cXAysoCg7IAND/SLqFCM+y4xPFQQQzYlybFCBbRMnAwxIXwgBocNdwLDA6ConEwS2baZ8TOi8mHzHhJom7FA4MSKhMfvIgMKBC6gO7Q6RXwAgrsT01LBCnKQVLQEYNicTCASxODn0ERgcAPS+5biLy5Dr0PHyVRPTMzVk7SNC8SlqGxsDhgQ/9G2NDFEEIjXagStM0i9QCyK17Tkyok6kEcu+eSUV2755ZhnrvnmnHfu+eegh+65uhJIbDMyFsch+uqst+7667DHDvo5QGAQhdfJQIEBEObI7vvvwAcvfOtxqAAC7nyEjd3wzDfv/POvc9ICBCnbHMoDMPQykifT2+I1MhVAUMIj5Jdv/vnop6/++uy37/778Mcv//z01y//FgFA0AjuSEiQgP0ADKAAB0jAAhqQfngAQAXagLs2SAAAeTigBCdIwQpaUIAJkAAScNcICARgCxcMoQhHSMICkgACFcBd41rgtJBoYnrT8poGIMCBEtrwhjjM4SOokL+aec0JFQiADv+HSMQiDtAOCvSZ15IHADsY8YlQjKL6AiABJ+AuYQGwghS3yMUicgACGsDdMCLQAk2MBBcugIC+bFYMTkWBGHCMoxznSMc62vGOeMyjHvfIxz768Y+ADKQfG/EACCgCdzzAgAgEychGOvKRkIykJAEpCwdUQA24W4MEHFCLSXryk6AMpSgdOQIM8AB3iahQI0bJyla68pWRjMKO0LYvWTCNbSKBBsQS4TVS+CsXQxPHurpmM2JQYAWtGlrUcoC7G3ygCG2jBAsoELeJgaECMEhRw3IBMVJ4LZUweI9IzBEEDPjAa10I2+ECFocPaKCa+0KEv+Y1NOlBoFQqWwXTADb/tI1BABE2G4YGVBCHtn1DBhXowul8sLtzkKQTaVyEytTWgrq1bWMSWIPKagGCDLzMoLHyQTEmhgyGykA5QzuHGKRWC5WpYVXbGtokppe3iS3CATDoREnKQYQVUMCH71oFtoCA0rbFYXqD25ctdoDNmA5NZhHgwkg/xoVV6bRt63hGDNjlvaQFIlcFxeo3gIAvfL4LFBV4JsZIkg0gfCACYKieqWzBhwp0yqnBLIKcoHCKWhEDESHI0lWxmg45uCACN3CEx6x1g1XJAWZPXRcIEBE3ZJwCCg5gQRG0edEYYKACfABGrWihBkYBgWQniUYQWEAhGjjBCYuSwApk4Aus/xIEHGNwwY5EgIQohIxROVWHbQVSjjlAyQEXQAIUfjBDDJxorVhVRydi8AEU+gAKSBABBC7ggjGss22+kMEKKHSD19IgRi0IQjRU8o05REhESVqSGVA7XOJq4jvwxRAMxIBX26YjF0CAwQrSdIEP/AgXkLXtM8QQgxVMKK3s0QR0h5sNMxhpQhL4QIbmUFSVXIMTRThAAHqgifXW1yDfuIUZEBCAEEzCFw49MUHEwQs5UCAAE5iDLsYhY4Kc4xeTCEEADmCGW3S4x9G4RA8MYIAicIK+LvEG+abRY4Skg3zGqDJCVPEIVWj5IMUgn3C/TJBpkM8bM5HyI6hM5oFc+f8RWW7zQLjsZTkLxBjkS7CWzfwINMtEzWyW85vjbGc62/nOeT40n/0cE0AfetCHXgeXJRFpPD9Cz1VedJqn/GgsR1oSXa50ou2s6T9z2s6QPrShD21pTPe41I0+taA9repHUJrVo5YzrGHiaFTTutChxvWlFX3mTa+503D+dLDt3Gpi99nYgW5zqoF9a2bnus27fkmvZ53sWtfZ2sMmdbFNfWxfd9vOoP62nJst7meTO9pknracVw1uV8s421GWtbR/PW9bizrcuh53rMvNbULLOd3/tveJ8d2Sbe/73P1Wd5vZHXB3DxzeX5Z3myedcGczmtf6jje/N77sdV+bzAz/Z4nDRQ7xNiNc2Aqvb8pXsvKMj5zMHIe5x6GNbIOTXOJkpji2BQ5ygj/c52R+eb13/u6eR5reJgf40C1edIxrWeM497fO2/1xbYfc5i1PesknfvIvz1wlNb/6zb8MdbJLHeVE97rRWY50tmt96VznubnrruW2B73se457vucOdr5XWelRj/lwz56StFcZ63avduKZfnGne7vjeW/63pUN9C8LHe5Ul7vVH7/2vo/97283u+Ab/nW1hz3ymK941wc/+h5DXsuId7vibct4lDje9qWvst89D/hMr17lrSf9600ved1TvuqWB3bntfx51Yee9tE/+OmJn/rAX5/1hHe9/+F7nHO8y17vBX/69qlf/Fcfn+bJB/7yD7/+Klff+7MHf+1lfHvh17/H92d834d84ad84ydj5Td5mVd5m3douYd6u4dVvXcSv8d/wUd+/ydjAeh+Awh/BSh/B3hiCeh8Cwh9DSh9sTd1+UeA+3di/ddjD8h9Edg2E2gSFeiCF4iAd6eA56d56Xd5W9eDDPiD6JaBJ7aB9/Z+aBd/Fjh/GDh99td+SdiBS/iBTRiC9TWCEPh8opd9LmeE9YWEC6eEjceEOOiEOgiFACiFY0iFZWiFZ4iFw6WFMsiF2HeCEZeCoLeCHtiC9fWCMhaD7Nd9AsiHVeiHwwWIIriDJCiEJv9IhHkYhCqIfkfHeXpofYb4hojoXzm4iGqogWwoc2Toe2b4h52YhYy4hSXYhXj4hZ94hKG4eKNIgaWYiKc4h2A4XGIoim5IinBoimi4iM2nio7IipD4c5eIf5RId5YoiXu4jIWnfsNYh6t4h8eYda8YhrHIe7Nog7XIicFYX4IYhYTIgZnoi5tIWLdoW8M3iDM4NDVYEjcIjHLIjqlIjcVojZXogLloW7soi71Ii79oi+GIi9moi9sogd0oj9+ojgVpj9Pojnaof16IjclYiNAofs1ofpPog/tIbRdpjhlpgOp3kP6YkDS4kCQxjwRZj1g1jmtYjlN4jgKZjm2jiKj/aJJY9Y/cGJDeOJDg6JJtQ4cSWY0U2Ypip5Ntw5MK6ZMMCZQOKZRD047k+I4ZE48r2ZA3uY5YRZRVOZEsWJF2p5RDw5Qp6ZRZCZVb+ZAv2Y87iZLwqJIjwZJBKY0hOZMjCYIleZdtSJM/aZNDg5PDBZOgKJN9mZdXuJfOiImIGYd2uZjK6JHMyI9kmTFmGZdoOZdaGZhcOZRuuZRweZVyKRJ0GZWPyZHPKJnRCISoyZiqqZGUyZe86JdPCZgZI5gQKZsASZtpaZsNg5td+ZllGZoNg5WaqZacyZZtQ5iwaJiz2Zj0qJitGZlD+JH9FpFfaZRhiZRfxpza6Jy7CZ0t/ymdPNiR1TmZIAmZGPmaJMma5Zma57maRViZDXOZopmZpLmZt9mZU3mPRZmPR3mNY6mbPcmbx+mbAQOcyymclkmcAWOc+Ymc+6mc/UmfAWOfxTmaIVGaaymVGeOVMWmVGYqfG6qfv8mfGeOdCAmeBSqedemejWiej2idJIedIQqWfSiWpmeh9IKhD6qhIMGhyemhDaOiJ8miTWmgEYqg9KKg/WmjhSmiP0qiQWqiCYqiDUOVN6qdOcqduMeg9emg9AKhJSqhJ0qhHwqmFyqm3EKmVWqmV4qmWeqf2Qmg2ymgX8qj3OKjYwqkHyGkE0qkAaOlUYqjh6ij/gelzSmlff9KpX9qpU2KpYOqpj3Kps3ipo8Kp5EqpwFjpG+JpGeppGXKpNzipGmqqN/JqG3qpx4BqGcqqPRCqItqqJqIqDBIqXtqqb2Cqa0KqaUqqbGKq83Cp6vqqL2qqb/KqcGKqiuqqpfKqh3hqnEKq9ziqaAJqpgpqm9Kqs1iqnOqp8Oqq67Cq9Hqq90KrNwCooXKpYfqpf4Hrr1CrM9qrOWKrOeqrNUqrPEqrpBCrhwhrZtKrc2irrPKrrXqrk9IoEnqoqYZm+opkuyplzBKjDJqjDSadcx6pM66q9D6r+baK97aqfrqKvLKsfTqsfYKsug6sCMLKSU7rh27EQCbrALbKwT/m6q0io62moYKG6oM26Eb+Z6uGZ+wmZ7TuZ5E254o+LB4GbGJ6bBHC7FJK7FLG7VNO7VPa7RCS50zip7aB68ky69/4q8y+7GuErLBCrYuK7ZuQrYaMbP3WrOucrPNmrM1ubOe2LPZ+rNDGrQxCp9dK5/Xqbf3qa2Zyq0qi68sq7Z/8rL9GrNva7aQgrb5yrhu4rhjC7kZAbeJK7eQIqs4a7A6i7A6mLGfurEwe7Jlm7Jnu7K9Yq3Dia2Fy7eBSp5/O7SBW7SDy7SH6bSOObH4WLH66LWuSLgjarjHiritq7g2S6dbaqddiqfvarxTirz1qryT67quArsNKrvHS7uv/2q7FAu4Fku8GEu9jWq9KIu9f0K5zcK9Yeq91Qu+0yq+wUu+wyu4NYq+xaq+q8u+buK+zWu5XYK5bau5GMG5y+u5fwK/ayq/6Uu/AXuaW4u0uau0kWi1vYu1vzuf/Duv/hu5rJu9zDu3Ldu4bNslbru5ktu+2vu5zruu0Nuu0nurBCwkBqzCCHwRCkzCDOwmoFu3onu3pCuMH2yyIczCI+zCJQzDN2wjOSwkK5zALRzAL9zAJ3y5KSzFO2wRPczEP9wldKuxdvuXeJuTR5y6SUzFS2zFTYzFT6wiUWwjU8zDVdwlAmzCcdwhc6widezFdywkeQzDpnutqPu4qivCAP+Mx1fsJg5cqRDcvxJMs/b7n8IboBcLe7z7nL4bncBryfiLyebbnVlcwFtMx11cEV/sxmEsJGN8umVcm2dskGmMyGtsx23MyG/syKWMw6fsx6lMEausy61sI0FMxkNsxkWMioUcu4ecuYmsxIssyI3cJY+cq5EMwpMct5Vcp5fMEtyAB3jgDId2DnlgB7YQaZBwB5AQabZwB3gQY3LmDOLMDTCRDr1gMgAAABhgBrkwYSf2DAeyzw8gBp3AWTKWDZdABAMAAANABJcAZTIGDp0gBg+wz+LRX/VVDrlgBhiwzyogB73wzBERDS3yARMSARBGBLUlY9sQBzDAAhiWJDD/AASeQNLSUQ6UECGfJQESoAEsIAOUANA3eQvr8QEVoNIYYGBxcGS21QtEECJJLQFLvV8mhhK6QF0RoAFaoAd60AUhAAE+wk/DJQ1AME0UAAVtsAdqcF7ONQdE3TDfkFvh4wNrwAdr4AMotCbflTHFBSUPcANgoAdtAEQKEwRjM1wPEyYi4AVerQUz1CiPchIbgwEaQAho4wgikCW4hFXa4FYQoAfVZFk84C90MFzlIAZykgVyZSq0kAW5IgZxTS90ICo80Fd+pQemRR5YdTcVIAL7UyvFgAgX0Cn0pDW1ky8HYws34FzXQFiqTQGhQEvJYAxRkCXAhFXeAgHDsi/I/+AHTJNMbcNNEBAF1J0MobAnH9U21+BZNtBV71IxGQAEfe0RlJBGNXUwVrMwWKULEDMIKlMMzR0Es+0q1yADEpAFxjIxWlABMvDcQ2MvGUADU3Uwg+Avkz00A1MB4XIwyEBRnzIS5RArt2MzDhQDvN0wY6ACInDep+II/sILbWMJLRABrb0vK3QJbcMLEEMreiMCHzAGbaMNMSABDGQzsiQDBZ4R/hTc+YQ9TZIxByUBbHA64KMCpz00QaABG+Q1SKABQdA2dKACF1DhE8MGQePUzRIq9+Q1hAQDxw0StzA9N34wMWIEqpDner7nfN7nfv7ngB7ogj7ohF7ohn7oiP+e6Ip+6IDAABAgUV7jAxVgAYte6ZZ+6Zie6Zq+6YkeCQVA5biD5gQQCZxe6qZ+6qie6pduAeaEO4sAAQwACKo+67Re67au6UZAKbgzC0wT5S40PfB0MCBAQ11U7MZOQlQwABBgMD8URMf+7NAuQUgkAUpkM8kTAE4U7dq+7fVDRVBwRR5UBdw+7uTePl90PF4jDLliRiJxC2lU5/uCJTogCapA7/Ze7/h+7/qe7/y+7/7e7wD/7wIf8AQ/8AZf8Ah/8Aqf8Ay/8AIvCX/g6JBuM5JO6Q7f8Bh/8Rqf8Ry/8R7f8SDv8ZIgCZBAAKDuNZpEAJDw8Swf8i7f8jD/8jL/H/MDPwEVcE5e8+oM8Ac0P/M+3/NA//NCT/OSYAQREAG73usj4Qs97jX65AK+HjDbsFVHHuAXsAJz0Db30uU2gwQYEOZDMwdkbuYH40AywCwN4wlpZFYT00Ex0NIhIQ6xkgVeowcRMDJts1ItbjNLAwMyPjQ0TgEttVF7EuIZw+MQ4OMkBeRCPjTZgFB64DVScAFKThKUMD24rd8XoAIMM94wEAGEEOADjtD0AjgSoFAqk04xAOEZIw7ldANkTy0XDgPZreFkDu+ncgq5It4iAQ61EwId/i7CkAMik9i3KQbBsjjenQXYbVszBQEAbuG5YlGeDwFaEPumIgolst5D/yMNsZIDMRTfIrA79e0RiWPZh1QsoEADWRL1Q/PZb+UH8LQKPhABLjAHOE0bgSInXTD4pwIQtbpAWCFG3DqECRUuZNjQ4UOIESVOpFjR4kWMGTOmm9MiApJVyUSKJMYnwgcg2TSuZNnS5UuYMVl6giFhBqiRIpEp0oABhi+ZEnHF+PAgRBc+gcCMgJABhqagUaVOpVq1JbQgKyJciLIn0BoeEHzKKWfV7Fm0aWF+K+JCAgUkbQK1SVJBggsx39Tu5dvX77pycmBcgMBjTaA9WS5EYAEk2l/IkSXD1AQDAwQaYPzw6QICwocYufhCKwJDBYYIEiR8gBGE12TYsWVL3P9WBsYKDKorZHAho1O62cGFxxY3SQZuuxIyrIgx6eBw6NH3pusUo0WG5BhUwDCzTfp38FV5BYHxITtrItD8mtNlZgUAABjG3Cob3v79mL7myIAPocgmvfATcECMpLEkCAIAICAIS6Qh8EEIIfqGEyIegE+GOX6JcEMO1xnnljIwgE+FNHQxRzJr7LCjmQ5b5NCcO+qgxUUaH1QHkjogUadGHvGjpY47TuxxyO+aUdGa2Lx55JFpiHQSunSWNOZJKoNT5RFVqtQSNmOWBG5LMP2aZklvklyyyTDTVCvKR6ZU882zrswSTjqp6vKRL+vUM6YxHykTNiWZ3HPQl9h0k1D/RDWSM1FGMbozz0YjhajPPycLFE1JM23IUE07ZUgSLD0VFaFHR/WUUjMFNVVTTlfV9EpJXNW0VFkjRRXQM2uNtFVdGV20V0ZpBZbQWy3NdVhCeUV2T1DnXHZPYZ+ls1jJLpVWT2WvffNXbeGMttswqY3MWnDVzLbcLWFFV81v16VSXMjIdVfLc+d1sll7t2w33x7h/UtefomsN+AauSV4yH0PbtFfvwBWmMaBH+ZQXYl5TLhiCBnuy2GMN4y4YwINBnnDi0fGT2O+ODZ5wI9Xtg9flyMsOebvUN5LZZrDaznn6CjmWcCZfxbOZrVwFhpKKY/GT2SlpQu6adiITsto/6hl27nqyWDGOrqnt/5LarSo9jqyq8f2i2mzufQy7eDAPktstvkqO+60fKZ7sq7vNstts+DW+6y5/65Ka8H/yrtwqfi2ym/Epwq88aDQhvyswyeHSfGqGLc8psc3d8luz82qPPSVMKdKc9JZ6jz1jCRnXabRX7fI9KlQl/2i1W+niHDdYV+796Bol8p24CXKvfiHQEe+pdiXb0j4qIh3ftOkp/88VOuZ/z17jaAPSnruETo+fN7Dt6j58L2XCfzwx+fedfMjQp979WNin3v3s1c+/onmz75+mNwve/mzXvn4JxH/WQ+ALxGg9Qg4PfgdsCEJnN4CXdLA6T3QefuT4P9DKOg8C7YEg87T4PIi2EGFfHB5IWTJCJdXQuQZEIUT3N4MH8LClbgQeTAsHgdtmMIa/pAhONSIDovHQ+Cd0IYqRB4RM2JE4CGxdzIUIqmCWMWEOBEjUOydFHWnxBkysXhavAgXdefF2/mwimIEHhktYsbboVF2VFzjFbG4DjdWBI6yk+PrwIhCNvYujxTZ4+v6yDo1CjGQuhvkRArJukOm7o8dXOTtGimRR6YukqSjoyLtiMVLRiSTpNtk6BL5w0rKLpQQGWXoSum5SUowla9b5UNa6blXbq6TqPxkFWvpkFtuLpeWi+UBZ8m6XzYkmJYb5uROucReCjGZDFnm5Jr/CbldQhNPdxwimVKFKW6Kr3rhTEgx+XfM1E1zIdWE3DUb98wwRvOH6lQIOxvnTsSZM37oJB09E2JPxOGzcNmM5zbJiRB/IgSghROo4OAJSHnaMKHrWKjgGvo3fZqPn6GbaEX/dlG9ERSiBj1oR491UJDeLaPh26jnTKoqlI7zoA+lZERn+FJwhjOldBNpTUlKTpweVJxtEipCVsq9lm4uqELdadxoKkubonCpMSVqUY+avaRabqrkbCrbegrVn4ZzqzqVKTmfasyodnCs3Oxq2q5qvaxObq13bKvZvorWsHJzrlis69jeOr24Qm6vVeyr1856zrRKcLBCLOzW7orY/7zecbE/bCzW/uq8wDZusjasbNUOu8/EHnCzM+ws1C67vMwibrQoLG3THgvayILSm7iCKVfLGs7PajS0/FttB1urtNMiL7WF660Ef3u01+o2tr6crbFqS9aqCjW4xRuu4Ip7wOMKLbcs3W38rsu/7P4sudxdrjSbW62T2ja6M8WeUKv7t+/GL7w82y5Su2u++JpvvjmbLvDeq7f8tu+23Byvfcs7z/OOK73QPRR7Y1XU/94twPgb8B3727sI023CA6wwFguM1fumL8HxWjBbO1zFC+suw3HbsANPLMT6gvjAEh3xv0pM1xf/8MNwDTH9atywG/M1xzZM8e1WzLYWZ/9wyDOMMY9nfNMfbyzIhF0yCossuyOnLckkrHIHdwzYHv8vyimbMmO7LMEmg/nJUh3zzcpM2TMf8Mqvy7LZtvzCOPPvy5gNswLbXLQ3czbP8Zsz6+o8tjvvcNDmSzOf16zWP08t0KRdNPnae9BDey3RR6z0+y5NzkxvbdNR7LT+HvFg9/a5gpEO26RZW2rrFTp1ocbaqLsI6+ntGbWqBiGr3+Zq3+J6g6cu6jpoXTVbn1HYJvx0OI8NtWTHcdkxbDY3n920aPNx2j2s9h2vrbRsG3LbSSQ2hHm9Ql/3DdjGHfcUu43Fbx8t3JBs9xffXcdHKzbdi1s3duudxnKnOt//ot135voN3n/LTtaki7fQ5q3JhL9O18I9dxMLfrqDyzfiiAw4pis+xovXLuP63bgk7+3JgfM25MMbuYDXS86JU/fjbVx59FpO4Zfj9uS8TLl3a/69m3M459xs9K57jt+fry/oLh76HWPu35kLMun2W7qSm47FhYeu4T97OClLTrqiU/zoIvbTN4u6X5pl3XNb51nXXfn10D0dw1Fn5NQDWHUuXx3FHQc13S1pdwbiHc96h/HOtQkpsQL+goJXNOF1bPiCIl6vihch4znteCJDfqSSlyzlW2h5UmOeyXx3tt9V6fkcgv7Wokeh3FVselqivoiqVzbrO6j2zbE9Z27H/yXcYUl6a8MembJ/Iu2lbXs0a96nnJdt2WmbUxMj/4CuN7Lw00n8LRpf29LnX9hlPnYfO9+50Mcx9wmtfLAyn7niR+9zo99gmKMfr+o3L/sV7P7yw1/nzvI4+MVsfxLDPyEzP0YDPm+zvn7CvjLSPnEjQEvjv77zPz8DQBsTQCpzQE+DwNKTwFWjQCCzQDPDQFNDtf6jPwTzQCkDQTgTwViTP8gyQRpDQTJTQUFjwVxzQdiCQSiTQTejQUqzwWEjwQjUQTbjQUDzwVcDQmbTwODjwF4zQklDwmBTQmpjwgN0QnSDwlaTQnajQm6zQnhDQI5SwDdiQHrzQnITwg0kQv9I08Jf40J/Q0N3A0N8Y0N9c0N1g0OEk0N7o0OUs0OCw0N+00ON40OAU8MmBESVE0SDI0SSM0SFw0HlUkSfY0SMc0SX0z8Ck0TyokSks0SRw0Sc00QLM8AwxEKLA0WWE0WhI0Ws40QD80Syq5T2I78BdMUqoj4sE0OXIkM9MkOIg0Q/gkUZk8Xwo8X7s8ULxMXCQ8QrNMb/Q8YAVMYQZMbH80Oeg8YJlMYKpMYVtMbMw8bDKyreEybf2xzvgzpUBDlVtDlWZDpwHD1xjDxy9EVCAkavE0bW0UU640WlskdHwse300eTc8ZT1MYO5MYP9MYajEcrI0YnQ8gnVMgUZMj/H3RIL4NINZPILKTIGbTIJMTI5JvHzavHdgS6d7Q6kZQzU6xDk/TIHgTJKVxJPdNIR+PIVITJI5TJLqTJ8yPJ5XtJs2OqcySmlvxDoXy+sytKZ7JJo8NJdtTJKOTJOPRJ8+FHQ/NHrQJITBLI3iNIsDvKbEzK8VtKsDQlpxQ7qKS5k1S6lMw7q3zAYtM9milHZmJKyME9y6HLmLFLa8LLdxLLcRQqv2wnwEQcrJw1rZQrrhQlrzTHs/w9oEw/sqxFs4xLTzNIlyTMxmSlx7zLyETHtPy+tZS6tqS6txw8zMyexGS4xRSszrSlz/zL0DRKzURKzjzNu0vNxlvNFpzM//mrzGS8zGJrTa17Tc2KTWCaTcOszaYEzhcUzmkkTqsSTHrMTancQqrcQ9+8QejMQensRuqUrtFUx9KsO90MPN68vO4MwrlETtVSTmVizns6zMLRy8nhS5cpzPp0Tmwqz7lbR7bMzjfczkJsT2a7zbHEzqGkKgRFHvyEHP1cGf4MKPsUHONcO/gkLvmkJvq0UP8MzO+cxPBcyPF0sPcUUNMk0Dw00Ed80OLJ0NzbUOvq0HX6UIa6UIwC0NdTUfRk0UF00UwstnQM0PP8u/RcvPUMPRgltxHtxBKtyBONvyeNxSj9yCnVOQUdzJKy0XrCUYvSUb2J0MaZUJPxJ23wBP8sQAAEGAJOcBBuAgdckIMEMIAZsIRnONLQKYdeoAMLMAALoINeqA++egZLmIEBWIA4wAVwCCdp4IQhYFMs8ARt+AtwmIQYcIEMUA0MaIEYiINKFSJz4AQZMA1OXQEYEAP1qCJdII8VqIAIqABUDQJdwCJoEIPbyI3VgAEg4AQh+SFtiAPr0NUPcAEYmIRG3QtpCIIWkAAImAEkSAIagAAJWAEg0BAbAocycAEMeAAQ8AEnuIEIgAAViAFP+KFyoAMYyAAIqAAecAIeqACxgAE6IFQUookPgAAIuAEnQAINeACfGINknaFfAIJmxYwkQAIRoNYWCAI4RQttAAIVgAD/L5CFnKiFNtgNGXiGGRIHMWgBCEgCUkCGkRiGQQCBCoCBc2WtOYCBCqgBRiBZkSiGRqgBCWiBObAhmqgADSCEYRgJZCiFJHiAFjCIGXoGGcCONqCFnJAFMCBXIAhVsyiHIlCBCnCEnMiJU9CADACCgT0gOQBZP8jakaCFG5AAGFhVCfIEF4gAKSgGsk0GYsgCCFBZFIKGmriBWojbZPADCHCBnO0gcACCDNCAU8haZCBZR4gAFSgCe6UKmoCAROBbkSAFCGgBSuigZ6gJMKDcZKgFEFABMdBTvQEHGcAAH4BbvjUGH8AAGfha/RKDDwABW/BcMEBbjpWgSQBZUvBc/0RwALu1inQAAgzgAZmlXC2ogBiA3fCJgxXQAGHw3GQYhAiAAWzln04A2cP13FP4206QoF+AAQgYBOkVBg1QgTiQINPFAC6Q3mTYAQwIgnOwCryFAERwX1GoW9HgH3GQAQno3PKNAMyVoCL4gBtw32S4gQ8oAgmihBWQgJ+VXi+oABl4jvjJBfEVBfdFBAiQAbWdClyAgQeQhePlW2QQCy1IXBVeYRZuYRd+YRiOYRmeYRquYRu+YRzOYR3GYVDQV0UoYb7NAQyggR0uYiM+YiROYiVe4hwOhnZVAwRWAwl4gGBgYiu+YizOYi024mSYgeJFYEV4AAkAhS0uYzM+Y/80VmItuAwgjltZ+FtcsApNAFlgQGANgAAOWBI93mM+7mM//mNADmRBHmRCLmRDPmRETmRFRuQqCAAIIGP3dYIKCIBFrmRLvmRMzmRN3uREtgMAqAA+QGA+kAAAsANOPmVUTmVVXuVLDgAJgAIE7uEAmAJWrmVbvmVc1mQOgAANQGBguFyoqIpOcAEI2FvpRQYKgIASyGVmbmZnfuZH2AJHbgQERoIKSABozmZt3mZMxgMAkAA9QGA9IGU84GZzPmd0/uMEkAAkQOBGgIAA2IJ0nmd63mYSaFcEroXLXVmq4AXxDQX3pYXL3QR0OIeCPmiDTmiEXmiFbmiGfmiHjmj/iJ5oia5oir5oi85ojN5oje5ojqboanDZUHZfEPiANPhoj05plF5plW5pln5pl45pmC5oczjdKEDgKHBdc0iHmZZpn+5poP5poQ5qok7odEiD2UVgv4UBbChqpx5qqH5qqY5qn55jCGBa6Q2Fun2NqjDdCoBi6R2EDn4M8CKCDDBe6e1eGLgFCaKDFbgAYpBeYpBVOpCgEIaA7fVcHsgAIpCgaJCBCCBf6VUDCmbeqHDeCjBmviUGEVhg0r2bYYYAgPZcJ3DdAOGfX4iBCBhpyjUJ65Wgb4iBC3CCNs4JUADe78WuAhaBuKbcWpDV9DWLa4gBCfCB1o5bLZCAGAAK/wkShyDIABBQbLINhL8NZgkqAxagAA3mW1KIgBUwAxTSBGIOBMqlBRDoWgvmH1+oiSygXGLwAd2+BrTghLxtBbKtBSe43EkgWNEGgcnOCWJYAwhggTF43PjZBiD4AAogBCBGBkKgAJTwjg4qhzFogQdgg9seCVAAAQyIgett67aFAqzOiVbIgepN7bNQB0qwDAiAAkQAhVBQhORNWTP41Q7CBRnIVx8IBBBnhDUAAQEugsvuIGgAghV4Vj1wBFFwBD2gAecGgg8G7SJoVg1YA0YQhU8IBCR4ANCI4xkyhzhwiwrgAkUABVAghCiAgAs41h1RC08YVgh4gDDHgBWQAf9LmN8f+oUg4FZn1dcKYA05yG4UygYxcAHz0Nd2LVYxUIkfEgdh/YDLiAAxxgAXCALe5ixLiAHcwHOxaAEZ4Oe92AZLsPEAIIAMkAE64PMqEodOIAIYqPQIiAEzyF3CygUxkAEGMAAGiAExyIXHnpxfMAPNJoABaI1OkHMbugY6SNoBGAAVAAJLEPC/MIddsIIqeAX7JqxqsIIpYARcx6Jt+AMq+ANhDydxYAQqsIJq6PJwKodXsAIr2AUTn06ifARiKE4erT4fRVIgbUQhHUUiTfddXPfTS9LKW9LVa9LeIVPEMdORqdAcDdF8kvd+pPfYs/fPw/fa03d729Lr7FL/hE89hT8+hr8dGd1LGoUvL/0nMP0oMVUpgs9Kgx++iJ+9id++io9Ehy9JBlXKck/5YazSYrzSmMzSTZT5iKT5nbT5Ulz5oGz5snz5eMf5jdT5qeR5Dwt5xRz56yv54jv5BoR5jiP6mzR67UT6vfN5ygR6yxT6orr4/Mx4ANt4hep4vUG7mOH3wvF3kAH4MBX4+7ROlof4dr/Ed2/FoU/RV09Ap88+qD9DqU8dsJdQsZcwsqcos78btHeZIu3RvR/Dvl/Avw/GwA9Lqn9Kqy9QrBeiwS/Twteww/eos/94ulF7wWH7jnF7j4d7h5L7n6f7BlWvyo87pXdNpuf7ug/F/7uHx7w3t9uH/Nxfxd1XyXjX+uDk+uH0evK8fLXM/Bbd/GvUe+cPUujPPOOPTuQndwftfYF7/F6M/DKc/Hycfc/p/H7/fBYL/cSnm8VfGdP/G9THGNUffdbf0esHz+wXT+U/KPNfe/QHiHUCBxIsaPAgwoQKFzJs6PAhxIgQpz165E0iRofeKk7L6PEjyJAiR5IsaTBdRWMmV7Js6fLlS1WPVMGsafMmzpvGKqbL6fMn0KAMKVoUanDjo45GlzJtyhLlI5VOp1KtClGmJKtat3Jdt/NRz65ixwYletEpUqVk17J9CVVq27hyR8qkOfcuXolfw+bt6/eg2alp/xIu/P+2MGK/kmYmbnx3r+PIbAOj5Sj5stjDmDdrxcr5M1XIoEcbpdx0MOnUPzWrbn2zruvYNUXLrm3SNFPUtneLZM37t8fFdoET18uzOPKHuJfqTu4coe/n0gnCnm59IO3r0pcbba7defTvyT2Ll569fHHuQr2jBx6+/W/h8JOfn79bfVD29mu/3y+7un+81Rega/gBpR+BqvWXYGrkMSjbgA+OZuBPCEr42YIXcgaghqlF2CFmFPpkIYiSZVhiZPKhONqHKzYmYk4kuojYiTMW5qCNl7WY418w4iQjj33VGGReHBKZ2I5H3uXjTUAqKdeQT8alopSFJVnlZBWddZplWP7/FaWXYxkZ5mPHkZkXkzY5eSZXYLLZ2SNZvVkmWHPOlWZNa9pJlZt7OkWln2tdGahTeMKkJ6FL9ZmoUGMyqtWgjwpl6EuISrpaSpd2haOmVUXaaU6UumQpqDYtWmpNjqJq1KerwiRqS6S62tKps7IEqK1CtZrrSrCyJCuvJNUaLF1xEqurmcf65OtKwCr70bDPeqSqtC7tWq1HzJrkLLYRRdstRLiC29K1406kpWBdmktrpuu6RK27IZUb70LalsQtvdC1m29JnPIrb7L/hmQvSfgKPNC3Bw8krsIYzduwQASPZLDCCTcML8QNPQyxxCJRfLDFCvubMUQbN9xxSB8L/xzywRiTrJDJCqMMksr/siwwwy9rHLDODs38Uc383vzvyD0vFPPBP3sUdL5D8+uy0QUhLbDSGTFNr9P55hw1QlP/WzVGV8ebNb1Qcy2Q1/yCLZHY7pIdb9FnG5R2vmtH1Pa6b7u7tdwE0U2v3RDhba7e65p99t/xBv7Q4OMWbm7cfWPHs+QFLa6RupWftK/mBh3OdeLuXt5Q4+A+Pi7flYe+7ugMld7t6eBGrjrlna/T+kKvYxt7t59Hvbq5uCuke7W8Y5u65MCPK3xCxEtrfLW+G608uMwj5Pyz0Es7e/K1d279Qdgrq/2zyPdNfbfgH5W57QKRr6z0PaOPrfoFif9/7PvHcn++95rXT9D9iJU/YsVPZ/Or1v8GEsBgDTBY5pPbAaWVQIEskFcN5NX+INi/yk1wHRXM1QVzVcCXRfBZHfygrUJoqwciboOSOyH72qfCWY2QZCVUFgyT0r6CzNBVGWxhnXY4kByqRYg9XBULQefCvhFRiAjjnBBrmLEbHquJTlzHEVH1QyUG0YlWdGIWSyVFiFGRWF80IhR3mMTfLVFuZ9xhGEG1RTZ2UYhvlGEa2zfGhpUxWHe0XRw7tcbptfFsf+xcIDW1R4X1kVeH1FwiLzVHQtZxh4+sXCQlNUj5FZJrl5RcJh+1yIM1Mlef7FsoGTVJTlayfaeUWyr/EzVKgZXSVq88WywJtUkDdjJqt+RaLgO1Sl620na/jFow/TTLf9VyVsc0WjL3tEsS9tJoz+xZNO20TH4101XX1Fk25zRMahbze+iqjA6vGM43TdOG1ezZN1+2TjZtM1/dXFU8STbPM43TneX03zm5lE4w5tF29aTXPVGVz4ztk0ztnOI7dbZQiDU0TP2E6D85GNDcxBCQBe3cQeOV0FJNtGEV9dJDyRjRl5W0Yh/VXEjdNVJQtRRkL63cRVWa0RdulDkdReRNJZdSPq6UZDVdWVD7FtN1zbRTR7VZUuWWU6LulIk97c5PIRnVsy3VXE3V1FOFtlWuDZWRRc1YWJs2/9aoTdWsVXXjVdeTVUyu1WhdHddXL5VWrNW1Z2Ul5Vk5Ftf8zBWUfdXZXcGVV0ntdWyHfVlbAftWQw72QIVF5WNJ9ldaBvZkla3QZWGZ2YwltluLfVRj3TZaiEWWs5P15GdHFFpcrvZijLmiVzors9jGaLbArK3CNstM3SaNtz/yLTKB2zJj4fa0jEpt3pQrsNJiy7mJgi7hpPsv4XKTuFQzbpOQC03tPu22V7QuobDrOPLmq7XDfa0vwasm8WKTvfTirj29+zX55om+4LQv3MzrRPQGSr2mA7C73Ntd+FqTv4fyrzwRbDgBC5HAfjIw7CRsLvwiVL9qc3ClIKxPDf+PS8H5ZTA8QTwqETOUxLKj8A4tvCcM787F3eKwSD1cNxXHisUUtTG2qFstGduJxsUDcvSYe14dA47Hv/KxSZEsLRzLlMmKc3KzoOzSqOB2HUKWFpHnZOTnSflZJu4wiiWK5W1p2aZcxu2XnxXmN405e2VWFpWZamXRrflebUbqm6945hynmaV9LtifoRpoJ8ZZWXNmU53Hd+dj5dmre2bdoSeWaLEuOoowbt+jzxRp/E2agEoe8KWDl2mPbVqtnVbjp20XajKNWoClDlajjzXrMNWagbfG4KkrnOrlrTplrebrq/UY687t2ku9tuCvc1VpvA67esWm2bEdm2yDBjv/xtVO37WBlm3Vbhuky9Zcs7H0bBBGe4Xnpl2hjRrupY07uuWG6bu7F2+0zttq9c7uvXHabVB/m379Dtu/1xtwoeabf/sWbFHQWUQ4tpuGDdfgwz0bcYFOHI8LV+rAZV1wBB6cbQk/8MelenEg8sWOJb/byTOc8rNNW7Ejl+DLBRfzGs+ca4OucsZ3u3GODhSNPWfryrnYckvmnHE7P/LRjVZz097chE3HXNEpHvWe5ZpY6a7SulNYcR+GnNlVx+HVSfd0Mm9dZ1Ov7tmrmHbXrd3ObYds0um4dFfOPXd1l/TdSfZzPQe9uEP3adY9DhdG552Sezdm34f3d1IHPmNv/x9y3M0Y+eZN3taVZ23Z0Z15P27+ep339edtOxxUF/67h8dq4j2a+uA2npWPN+fr5Rp7oM5+uasXduv3m3vC7l6rvZ9u6OF9e4AO37LFp+vxt1t7Yi5fo80H7fMNG/3y/t7bwf/w9WWbfcxuv73J13f1eRr+3o5ftOUv2/TJmX6rrv+47aft++N1eTCP3pGlD9/pQVv+Jdj5Odz8wVX9hdf9/dYATlj3Edz37VgCztcCJlcDblj8+dMBUtYE9lcFjtcFllgGYtQGwlYHPtgH1lcIvpicLFkENtkJhlgK/tcK3tgI6lQJxlcMsoQ5QEMhkAAJwIEvlAMcSUMnlEAHaP+BLoiDEGkDLhwBBxwBLmiDEImDLmgBB5RAJ0hDDpJMOfgCHJBACRTCM3RhSZxDLgRBDHwABEBABsAAEHQCEWrOL4iBDLRAG2IADMQAJYBD51xDHMQADFAABFCAC8RAHFxD54ADJcAADGAABERAC8iAGPxC55RDJwABDLAhBKgADARBLpzDT4hDGcBABkTAA1xABeQhDIjBNkiOOTSiCrQhBVxAG0pACwABNFTOLciACkhAG14AIUrACsQALlTOMwBBC0jAA0SAKkKABHwADFCCOUjONhTBIz4ABFTAKj6ANJoBE+KEOBRBC0RADSiCLSRDMgxDIyABBKwAELwiLsn/AQxIQAgEgiyoIzKEghZAwAfIgCXKDSc8YgW0wSkggzqeQhtQgB5yQt/8whpCABeIQjGooywEAghIAAzIgTrIzTYEwSwiQSMMQzIggy0owg1AQAsUQTjWRDqkAR7ugTGoI02WJCK8oxjMYdRoAgw8QBQQQ03SJCloQAYAARVyTS/EgATYwCwEpTrOgg1oZC+cjTYAQQZogCg4ZTIQQxRAAAxowtmUQxHMIiIgpFO2gSSaQUfWBC/EQAS0gVbSJCF4ZSdwjTYoJRKYpVaeAjHGAdeUQxB8gAgAQ1wmAzCIwAcEgU72TBysQAScQmEiAxJUAAwcpdF0Qk8iQknGJVrG/wAv2AQRCKZeaiUy/AAGyMBivswkOGZTFmYytIFGSkPU4AIMQIAjuGYyOEIEwMAxGo00yIAE6AFuykIErMAkRE05yAAG5KVrIgNiFkFNXEM9BgJu5qZX+oLRlAMQYIAUVKctqCRYGk0ZiGZ1ImYZ7CQe1kJ1SgEGAEFqZkwv1KYjjKZTIkMgSIAMKOJL5EJPrkJ1EoMbvgE1DCiBFqiBHiiCJqiCLiiDNqiDPiiERqiETiiFRmgsYEAEEEJ1JsMNYIAQVCiIhqiIjiiJlqiJTqgzYMAFgAF9auUXYAAGOMOJziiN1qiN3qiICgEG1MCG3uQHxAKOBqmQDimRlugbfP/AAwAlbq6CV/amS3ACHgrDhmoABHBARVwplmaplm4pl3apl34pmIapmI4pmZapmZ5pmVpBADwAKGyoE0jAAaCpnM4pndapnd4pnprpHQCABOzBhvKBBACAHeQpoRaqoR4qotLpAUiAE2woKDxAAFhBok4qpVaqpd5pB0DABWwoMLyjQ74ElEIAYVanLZLApZ4qqqaqqj6CmkJAm1bnZCLAqs4qrdZqne5pcG7oHgQqHtiqr/4qsHIpAjDqhn4CBATAFQSrsi5rrWaqBnCqStblftYmZOJmp6qAIHSDtm4rt3art34ruIaruI4ruZaruZ4ruqaruq7ruXrDLqhABAz/wobOQAY8AbveK77mq77uK7/2q7pawwdcQBdsaBdgwAdYg78mrMIuLMM2bL4+QQbMwIYOQifugsNeLMZmrMbyqyDM4qi65ik4AAzkAkxkA3DyQXUqglcGpM6YQxBgQKPi5ixEa9TEwQeAwIaCwAf45WXiIS1UpxO0ZzX2DERCwCJUJx9QgAxYpkuIwc1WZGTmAAYoZtRYAh76p2uCAX5mQ9ToQm0qAm4mwgPAgC5ETTYopRosqUpaQtSYg1XmQIvWJDHorBjYxC/U48AWJh/sppP2zDYo5w1ArVaGgkrSAdeYQxF8gAa0plbOggZ8QBGIYtTQwQq4amEWg9TKwDdw/w1tQgDKxiUydAFlsixMzEFtdoGS1iQy6IFKlsHQRs0t1KMPpKNTKkIFJObmcg00xAAGiEC1BuUpiMBp7iLXfENgVsDROmUt+IBGesLZmEMZVC4fCC5NEkMXQIALzAFOlEMZuIAEgIAfnIIwDMMqDAINqCQR+OHZqEMjSkAFrIEo2AIxyIIi+IA/AoFsyg1/QmIWOEItEEMtOEIWQAAGxADJyk00AMEHRAASKIIsEIMtiMIacOM09g04EIELRMAMDMIqDAMwnEIfgEAEuEAZvOdLlAM9fgAwtiE0viE4So46dEIMrMAqsjABuwARcK3k9IIyYoADRIADBGMuYqfkZP8DETwiKrJwBayADHSCGfILKcKABlDAD/9wNMIAHZhwTfiCGMxwAzRABsgAEejCEwuMNAQiDDSAA1wAHMrhIk6CDMQABARxDMjAJORu5YxDJsZABTiAyCJi/mKSLhCBDHyAH6tADFTiUmQDJqDACRgCFwrRN+SCI7PBM7yu7YhDLzTBCTRBL7Sk7ZjDM7SBCaRALuCx7aiDNBjCCaAAJugw0XWc7C2epz2gyL3gle1gj81ghNVgkN0gVZWxvOnyk/HyiPlyktmy2eEynxFzlhlziyHzlAGzWwkzvzkzm0Hzj0mzmRUgxlkzxG1JLOPWSWFJ1wXL10lJ2M1KOVfJ/sn/Wf+Z0v+tjzZHGTfDDzVLFjhrnDgjnizzHi3v0OBZGjNjGjb7WT1vWUC3zzs7Wjzb0jzbTwCy2z3rTz671j4LXT/D3j8b30JzWwuyXkYb3kbrXkdD30ebmzKLXkGr2kEjWkK7WUprTkPr2kM7U0QD0ESLXUWbWkgD30i7XkkT30lr30wL3EorH26ts6u0s5TUtNfdtDfltALtNDuP3aqcM6+k85Mw9ao49ZMMNLW1NLG9tKbFNKAdNcMlNfotNVVTkFU3NVZr0UW/V1AL31A7X1GTn1qD3E97312DX15j3167X1+rHFsboFubNauhtaIdNs3V9YIFtgQOtvgVNv5B/7bPeTPLLbZlsx9mM6BmI11if7Nnp4tjc9poS51knxhlw+Bn219oW+Bqc11ro9lr53JsK+Bsg2BtIxZnK91pSxw5z3WpQDU6SzU+vbUHxfVXG7cc3Tah5XYz7zYF9rYK/jbe/TUEUrdBW7cHYjcNarfgSTfQebdLgzcKincvk7flmTfhoXdZq7cMsvcxuzfocfcty7e1Mbaxpbar4bfqdRlXK4lXowpYKwlyb7VyKxRzo9BV9zSuwTdB8ze4+Te2ATiyCbjIBLfeDTfHFbeE88qC50qBH8mBl0qCH4lWm3iDk9SDOzeCQ3eniLXNkXV/0/eKabi2cbjvEfiL01SM8/84ufk4zlD4WFu4wWG4uBG5vRk50Xi444H4OKsTjSsSkt+4kpMck9ObkwMclPNLidvKiRNJioPKihNJi5N5kDvVkNt3NIe5+en3Mm85znW5v325wsn5fWU51eH4hev4LsP5NvN5gJV2Z1/RmXdKmgeJjf+5nVsdniOcnqOcoRMgogu3or+5iF/6uoz5rJR5kCy6pjQ6jzw63AH6kgt6MRO6PXs65Ph5qkc62k26yVW6zME66sg65qk6l7P6M7u6QnfZmod6m4MVp1v5iIuQlNselftzp3cZqLuKqPMIqV+KqedIsVP7setVshPUstsKqvc6rcudrcMcrvOcrr9Ypn//+Kafu86lO9Stuw22+5S/O7Bns7DLNLE3O/U9O0dHO5zxOv/5+p3nO0Lve1pLO8HDs8FLOsLDtMI/dr/bu7PjO2pPvGr3O52zdLlrHrw7nbyzHb0fT8M79MPXesSftcYHeMUDecqb+8o3dstvOMfD/MeTXshjXc33+Ms3V7cz1rcbXclXy7SvSrXnyLVLSrbbyLgXfM77386r3cjbXdFvz8nbdMyD/Mz/d88XOcNb/L9jPHEr+9V3s9jLH8CbtMALmr+rPdmHuNmHPc6vPVG3PeOlvQbavV7jvad1vFLHfZWD+9njs96TIN8Ttt/D2uHjYOJf9uIrG+C3teBD+9wP/3zjB/Pjg3bk287Ro0rS28jSP0rTz8i2I33Qo9bQa13hW/TkK3blB/zlX9Hnl0roz8joM0rpu8jpg37qP9fqK97NA/3W63zXZ/jXP/nPu2DUy/PU013VA17rE0vtg8rtu0juJ8rur8jTO3zzQ/Tz+130U970T3jmV/Pmy3bnd071d8r1r0j2E8r2o0jv2/7vX1fwz/LwM3/68/b6A8Q6gQMJFjR4EGFChQsZNnT4EGJEiRMpEpT0SFVFjRs5dvT4EeRHY48epQt5EmVKlStVTiPpjSVEbySnxbR5E2fOlOlIGtP5E2hQoQhVYRx6FGlSlSNLKnX6FCpEl49gKp35qP9mVK1bt/J85JNrWLFIi0oaexZtTqYm07Z1G3Jq1aRXs761e3eiV7B4+fZdWDSjX8GDB64lfLhvXKd0ETe2q9dx5LYXA0u2HNbwZc1RFVuluRm0U8ihSR8tWxp10MypWdvsPPdza9krR8+2fRLwbd0gV+/2TfE1Usa/iUusXRx5Q8rJmTPs3Rz6wOBHh0e3vu74dei5tVt/3h359KHVwSfPXr74afTMv6/fLV4oefe+z8/Xvdw+8fb5ZcMPKp+/2eoLsDXuCLxtvwNL8w8oABVEbcAHSVNPwtYSrFAzBn9yEMPNIuzwMgNBJO3CER3TUCcOTYzswxUbw8/FzUqMcTD/FHNSkUbCWsxRMAp5jGzGH/GyESccheRrxyPvElFJwoJssi0ibzISSreSrDItGLH068ktxZLSJiq9HOvKMcNi0sy3ukyTs5cWi43Nt8qMEyof6UxrzTuTAjMmMfUUrac/JzNKULTyLDQoPlnyE9Gh5mw0KDQhherQSXFSdCVGLdXp0U1xstNTpSoNdSVMVdKU1Jg6TXUlSVlVjSS2Xj3K1JRQnRWlVXENSctdgRrV145qRenWYD3S1ViOQE3WJmCZBc5Nz7B6NidkqZ3I1WtRclbbh4Y9qdhuI7JWXId6LfckbtFV6NuQwl2XIXLhTSjbeTVS196C2gXp3XwPktdf/4KWDZgifAne96N+CR4I4IXPXXgigwNG2COFIW6Y4HohdkhifynuyOKFMQ544I0f6jjfjzkKmeCR/dXY5IVQtlfljVgO2OV8H46Z41h5nqhmjW72N2d7S/5ZoZnnDbqiofMtel6YkTZIaXiZpshpe6GGd+epE6p63asnynrerdeV2uvCfE67IbElIhtes9E9mm2CwEbX7YjgXlfucruu2+61AU8ob5ngHPzfQBGnl9DFD7q73MIf2hvdvsWle3DIxZXcIcrLtbxbtNnWvFvOG/JcXNC1/Rtx0rU1nSHUu1X9WswBd/1a2BeSXVvaqRU9bdyp1V0h3q/1/VnWMxfccf+CiE/IeGqRZxZ4r4V/9nmEon92+mRtr/t6ZrM/aHtmuzdW+duZb16g8Q0qP9nzg61+6vCTdb8g+I2V39fvR1+fffgjiP6Cxb9d0Q9p9jOWAAdCQF8ZEFfpAx8Am8dAgThwVxCclf+CR0HHWXAdGMSVBl+FwJ8pMFggFOGsSMgqCf6vKexzXrRgMy0ZMkxxNxSICXmGQl+p8HAybGGqOGg9Dy4OiDbU4RBJ9cIOxlCHSazLDZkYKh7GzIe7kqIOBVJFTxWxfkdE3Ba56MVNXdFkWcQVGZeYQx060YhQvCEbqejGG4IxgWIcHB2FaEcZonFjapwVH9lnRkvBMYxylCH/IZtnyEkBEmKCfBUjHedISOHxhHoEHCUXZ8lGITKPigwgDYUTxEL6kX2QXJgkWcVJxHkSUZjsoSbr5srBwbJQqiQYK1NlS8DhUlCgzKQoK0hK6piykahsniyxSEu2+bJuwPyTLgPGS1JBk23S1JMwZ0nMDxpzPMispDIdR01/WTNU2EybNu/EzDQ6M23q9Bo76cTNZnoTieCMjzg7Sc7FmTNf6PSUPKdGzzi5M5Dw9BpBkWZQNgHUXgLdFEN/5tA02fOd+ByjPv/Dz1f6E3EIjaRCp0ZRnlnUTBCdl0QtZdKYoXRMGE2oRvfI0QZ59JYgHZxK4cXSSbnUZDD1kkhX/0lSpAF1Y0LdkkxHStNN2nRDOP2lTgHH03X5FFJIvRhV60bUXRr1Z1oVGVfZZlV0YbVRYm0ZWdPG1KI6tZZQTZFUo8lWr3m1mmDlmVpxZtepmbVcaEUUX4nmV6S59atwfaZcb0TXbBr2Z4AVl2ALRdinQZZneD2nXmNmWa1hNmaIzati48nYIjl2naA1mWS7RVlBebZsqt2YZgPKWZPBNm6yhRhrteXaP+GWb7p1WON06Fs9Abdyws3YI8zCxXUY907I/ZxySUbcG0KXThYUhyuWoIRMfEOH5fgFE5QQiGwsMRphUEIYoiGr5qkjG4FQAhN+UQ4dbiMTSliCK8RxFP9pxAEGKigAAzIAgzL8onnfmIQMYFCAAkhABkXIhXsBJw5OBEEGDWBAA2AQhE70t5O5KEIMJMAABsBABpTYRvOeUYYYZMDEKohBHKQBlHRYQgYrkAAEeAwBCagABnOw7+ByIYMWYKDHPv4ADMQA3sFFIwguQHKSMdDhaCBuG2OAwQck8IAH8BgDLZBBLhBXDjlsecc9rsAKZGAJCrMkHQCugATU8IlVtAIUbdCABFwwBhCzjRMwQHIUFHEKWYjCDzSAgAqCoA3A+UIGH4CADwhRClmcYhBIeMAHZOALwGkjCC2AAA38IApZoGIRUYBAlTlR4TG4QAIaaAMoZLEKR4D/Yc4wkMObVWIJGEAgCbZARjKITWxieCECuuZ1zHQBAwnQgBXFLjYyArFoIvx5ateQQQY0AIphSxsZjtAABmRwDbaJowiSDsS3pb0KGkgABrpgmzrk0IIHfGEY0ia2LZAAARhY4ibQcLYT2K3vZKjB3/L2mjiAgIEZBMPgxEaGIiDQAoDPcwwquMAqklHwYq9izWNY9sYsIWpFRJzYwaABBoCA7Z81GwJrQHnHoQBvaNikDB/QADBm3nEaZCAII1+YJkR9ip4nQwsVgMGKp9aLXw/i6ISIQAx64bVtxKACWjh6KR7QAk3MMwgfoIExeg4MDaigDDHZhgwk0IajIwMR/xCQwc0bGgQMIOHoyViFA2DgCa/JQeMeNzgyLqCCOXjNE7+O9tGRAHShB0zgEUBE3tsggRgwXSUwL0Xea7HoSpAD9KEX/ehJX3rTnx71qVf96lnfete/Hvaxdz0zWhCBQOQ9GSL4ABlk33vf/x74wRf+8GEfjiNnAfdZCHM4iN985z8f+tH/PRkyMAPBGzwQEmgBM6Tffe9/H/zCr4QKHmCLvJfC37xgSSdEbf68VwACJCDJ/Olff/vfH//51//++d9///8fAANQAAcwAK0gAB7gE3APCSoAAQjQAR8QAiNQAieQAgUQDwBAAvQA9/RAAgAADyoQBENQBEeQBCEQAf8kAAmuT98cAQICYAtKEAZjUAZncAJJ4AEuAPdsoeI6gSU4QdSEAfc0AAI4gAaL0AiPEAkfwQAhABRwzwkkwACSUAqnkAoj0A4AoAL4APf4IAIAwA6qEAzDUAzxLwAqwAlwDxRasArGkA3bkAo5IAI0APeCoeJabSVwQdQ47uiIYdXQYBn+EBADURAHkRAL0RAPERETUREXkREb0REfERIb0RTgb/Ly7gYwAAciURM3kRM70RM/ERQfURkkoALAAPdwTQKUIRRXkRVb0RVfkRNxAANuAPfirgJSARZzURd3kRc/8QwyAALI7uhWwd/IbCWkgcSg7ujSEAbUD2nKIQb/MEDrjg4YKu4SvGYMxA73aOADxsBrLmEFIIDnjk4LyG3IfoYXYOABmvDoBsHfrmwl1CEIMsAGVLDYoIDcXC5m5GAFKsD9UI7apo7ukCbxIIAUji4U/M3vpkbgIOD2eq4WKsDwFi4ao+DtbuADgkAdYiIXfu0hUY4FvS5tpIHtLHLmWqHw0m7hGq4GiGHmiKEGWE4fTUYdykAFKqAVei4KKkAGauwbRc0Req7aYMAYWcIcxiAcCeH6PuECPuDa2IYSXCACwOD6VkEEyM3c0iYXSAwJxlHfgAEJLK8op4YkMWAE9FDfkAEMIMAFKOHciOADKoAd05IQIGAFxuAcbuIb/4IgHJMgFNjtFLQAAj4ACLIybcyhDKSyBhjBJYlNFtrgAqpM4aDS2UBgEGqh2GphEEAA3tyybnQhGiugDWTh24ZhEWwgAlygDMyhbq4BCCRNC07h25AhFJLALoMAHHJiG8RAyiLAAkaABi7Ax1wgCHyybsoBwCQNAmbgBoRQAtjMGaPpEraMAhwgBGwgBHjsA2JAEx7PZHghxyggAi7ABkQgAiCgwOTgHNlGGqIM/i6ABmigOjHABcQA83BiHDiBwVbgAz5gBWAgBixhJucpF4IABvjTP11gxhxtcZ5BDGCgBfpTBcRMDJ7BcbIhDmTABfhTBVYgBoJgwhYHHCwhBq0CrD9Z4AVkgBMG1CbEgRcmIQ7MYBJy4T4XR7wsIQ7KgA464bwKKRo0YQ7KQA40ob1kKBs6gQ7KIA4sob5kaBtygQ7MIA4mgRdY1LmuFEuzVEu3lEu71Eu/FEzDVEzHlEzL1EzPFE3TVE3XlE3b1E3fFE7jVE7nlE7r1E7vFE/zVE/3lE/71E//FFADVVAHlVAL1VAPFVETVVEXlVEb1VEfFVIjVVInlVIr1ToCAgA7" style="margin-top:1rem;" /></div></details></div>
```

::: {.sourceClojure}
```clojure
(view
 '(ChemicalData "Ethanol" "StructureDiagram"))
```
:::

```{=html}
<div><details open="open"><summary><h5 style="cursor:pointer;display:inline;padding:0.5rem 1rem 0 1rem;">ChemicalData["Ethanol", "StructureDiagram"]</h5></summary><div class="wl-results"><hr /><img src="data:image/jpeg;base64,R0lGODlhBgFVAPdkADAwMLq6uiwsLNPT0x4eHo+Pj9/f31VVVYyMjICAgFtbW0pKSi8vL3l5eZmZmba2ticnJ3FxcZiYmNfX18jIyLOzs2dnZ+rq6j8/P3x8fHp6eo2Njfn5+aurqyYmJoODgz4+Ps/Pz0dHR9bW1uTk5FxcXKqqqsbGxkFBQdHR0WVlZYeHh0lJSYKCgu7u7m9vbxMTE6+vr/b29lpaWsXFxVFRUZqamjU1NSgoKLy8vNnZ2U9PTwkJCVJSUiEhIV5eXqenp+Dg4MzMzFdXV3t7e5aWll9fXxUVFf39/Q0NDVZWVtra2qysrIqKir6+vuHh4RAQEPT09BcXF+3t7R0dHbKysjo6OhsbG8LCwisrK9LS0gAAAAwMDKOjoyoqKjs7Ow8PD+Xl5XR0dMvLy/////X19UREREJCQkNDQ6SkpFNTU+np6TExMcnJyYaGhjg4OP7+/qGhoX9/f/j4+Pf393BwcGtrawsLC4iIiOPj4wEBAWxsbLe3twcHBxoaGi0tLRwcHAMDAwgICMPDw7GxsRYWFqampnV1dejo6N3d3a6urkhISDMzM2pqar+/v+vr6wICAjIyMsDAwGJiYufn5/z8/FhYWAoKCvPz8/Dw8F1dXaioqCUlJWBgYKCgoMfHx/v7+2RkZGZmZjk5Ofr6+qWlpZGRkQUFBc7Ozg4ODubm5ikpKVRUVJ2dnWlpae/v7wYGBkxMTCAgIEZGRvHx8VlZWZKSkk5OTq2trSQkJBERESIiIh8fH7m5uWFhYTQ0NJ+fn3h4eLW1tZubm4mJiU1NTX5+fi4uLpycnMrKyqKiotjY2BgYGDc3N97e3uzs7EVFRSMjI+Li4hQUFLS0tNTU1I6Ojri4uJSUlBISEtXV1bu7u21tbfLy8ktLSwQEBLCwsHd3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3d3dyH/C05FVFNDQVBFMi4wAwEAAAAh+QQECgD/ACwAAAAABgFVAAAI/wDJCBxIsKDBgwgTKlzIsKHDhxAjSpxIsaLFixgzatzIsaPHjyBDihxJEmOvKSVTqlzJsmXFKGK2FHLksqbNmzg9wklzZQsaWVvUOMtJtKjRowLbvNliZQyZOU2S9CGCCanVq1g9qjKyBZAhOAQpTdLjpwuSrGjTqlVIxw0XQXKiIBTyZcubNmvz6r0KR5GPLUOCLIQDBNCWH2H2Kl7sUsuZLWywPCwjh8cdN3QYc8zDhIlThUE6C9F89FEjSIWQVZL4xNIWH4rAkr64acuWTgs92Q41OycpU1BOiclkEUukLWe09K5Y+3bu3cttXlu1pdiSjJVaFYLk6lH0iM1xK//UvYX395VLYm3x0qsjrUOnoJgidb5h+Ofl65fMJCa+LfofLXHLFqtco59C940H3YEgVYLMdnY8Q1IvXmwRy3UMFpRgQuSZlyFHgxyHhnIpkWKLLsIR9yEZGyLU4YoaPTHEFruAI9tKz9iBmmoftnjQizBaRJkgXOAxx01aoLFFJINk6KNBQAYpERyGGOaLKkTBAc4ugD1xYHOLSCLmmGRKEsyCUkJEF1OfGTUHHlzwIEcZ9TVn25145plfmg5R4oseXt2IlCq+bEEFEILOZqeejO7Jp0JzEJOEIMbIlVYyo2zxxWjLNdeMJ6CGKqonoaD5qEFwEBINYILlBUcpgOj/MQklvTVnCX4enkpQNdBAJtliURgjSBJNHKmZrbjqStAUewSS2mqkBaHEFrlUkWheyCro6KmgYJNNit9hwcYWs2izWLYcmsrnNllUh+F5lQxzRCDcoKQXui6qG6QkXBb4IX+nJDGEsWrh+6O+H3YTzjd6hAPgijoAlUUOaxkMJcIHIgEMM5A0MhSfuOCwRQ2JpGVxQVF+SMMvyJF46m9QwBJMN1mdTFDKA2Giibm95THDazFc++gjrkDCjCdnWWXzQDgnBUMgL7igmQwtXHLHCpkpa1AKKGxxwwkfuvDCNzAUAcpecJjAyxZGJKY1Qn39VQIJH44gwhZ/OJFXCCDY/4XX2wu1dcclH8iQIRzUeLBFDx9jtYYKZKWRNOAMhfHDFrx0IHR0HGwARh8a0GJV56n0kUFVlEPURjNbYBDChxdYoMc3GkyekzCKs9J46hAh0YUfkIhywYcV6LJFMxTkNAELWxxDE+8VYZJBH2BYw8HhHayNWE2v1PHNNBJAC71FzrCyBQQPfCjDB1Zj9hAih7wRDTOcvNHAGhBVIsE030TwyvgbkQQDtrCAZXwoDwp4TWwWIgRONGoLHkBFQxwxwEVMAIAdqYQD+Oc/lbEMDVk7iCj0pAc9uUIhcKgFBIWBQZC8IgLfc4D49KMxOySEebYJxAL4cD06gMMMkLhTMf8UkgAEXK+FIVneFhggCV0Z4044wJ9BpPEX2+ABiTgRBgS2cAAD8MkZJVySQpAgsi1AAhFYvAkHrAEGQWTAUjBS0hbuEEKE0KIPttlBGnESOz1cIQ2b+w4HgriFAjTkTFv4hu326JIQYGALo0jGh1phmz44BBSBsE0VGIkTOGTvNlg60Cxsw4aHrGoLt+IkTtZ3CS5gTT+5sM0eHuIavKmSKCQoAY2Cdh4p2MYaDwmHbQhwy6Kc4AY+GcB3UmEbZTzEGraBQTGLgoQ4cCxCyymEbRDwkAzYBhDTNEo3GgALXXyiN7HcwiwdwpUlhvMoiQgFnWZjBttE4iEV2sIt3smsT4XYoJIOQcIpNNnPghqEDmG0RUMSYJtTGPShBOnbFrhwxIRggge28QZEN6qDMN5gIe1KpOg2ClFh2kYA9iqIKk65hQyQlKSPsc03biGJsyAhACwgJGBe+tLL5UmndyoBT3lKgyoyahcBGKpSpfECRsjiCAQAwAsuqNSqWvWqWM2qVrfK1a569atgDatYx0rWspr1rGhNq1rXyta2uvWtcI2rXOdK17ra1a0BAQA7" style="margin-top:1rem;" /></div></details></div>
```

## More Working Examples

::: {.sourceClojure}
```clojure
(wl/eval '(GeoNearest (Entity "Ocean") Here))
```
:::

::: {.printedClojure}
```clojure
[:entity.ocean/mediterranean-sea-eastern-basin
 :entity.ocean/mediterranean-sea]

```
:::

 TODO: Make this work with `view` as well.

::: {.sourceClojure}
```clojure
(view '(TextStructure "The cat sat on the mat."))
```
:::

```{=html}
<div><details open="open"><summary><h5 style="cursor:pointer;display:inline;padding:0.5rem 1rem 0 1rem;">TextStructure["The cat sat on the mat."]</h5></summary><div class="wl-results"><hr /><img src="data:image/jpeg;base64,R0lGODlhSAM0AfeJAKqqqtnZ2dzc3MjIyKWlpa2trRAQEI+Pj6mpqcbGxsLCwj4+PicnJx4eHoCAgODg4Kurq15eXtDQ0MXFxaenp0dHR6Ghofj4+EFBQdLS0hEREY2Njerq6ouLiw0NDSwsLNHR0fLy8r29vTo6OggICLy8vPHx8RISEm1tbdfX18zMzDw8PPn5+XV1dU5OTgwMDNXV1RoaGm9vbz09PYyMjLq6unZ2dpGRkT8/P1FRUQEBAUpKSoqKioODg3JyciEhIVJSUgMDAzY2NtjY2Dc3N6KiohgYGCIiItvb2x0dHfT09LKyslxcXPX19ff393BwcGtra3FxcZiYmCUlJRUVFbGxscPDw5aWll9fX5OTk1dXV9PT01ZWVhYWFqamprm5uWFhYR8fH8HBwd3d3a6urujo6EhISMfHx/v7+wAAAM/Pz6OjoyoqKlNTU9bW1rCwsAQEBEtLS3d3dzU1NZqamigoKC8vL6CgoFhYWLOzsxMTE87Ozg4ODubm5ikpKZmZmba2tgYGBkxMTO/v72pqajMzM+Tk5AUFBcvLy/////7+/nt7e25ubmdnZ5ycnMrKyi4uLuPj47e3twcHB3Nzc9ra2pCQkBkZGSsrK5KSkllZWf39/WlpaZ6enjs7O4eHh2VlZSYmJsTExOXl5Q8PD09PTwkJCV1dXaioqJeXl2xsbEZGRiAgIH19faSkpJSUlI6Ojri4uDk5OWZmZi0tLRwcHHx8fFBQUEJCQkRERENDQ8DAwDIyMvz8/GJiYufn54SEhK+vr4mJibW1tZubm01NTYWFhX5+fvb29lpaWklJSYKCgpWVlaysrPr6+lVVVbS0tBQUFNTU1OLi4t/f3+vr683Nzb+/vwICAlRUVJ2dnXp6euHh4b6+vuzs7N7e3iMjI0BAQBsbG2NjY/Dw8AoKCoGBgfPz8+7u7u3t7RcXF4aGhjg4ODAwMAsLC8nJyenp6TExMbu7uzQ0NJ+fn5+fn5+fn5+fn5+fn5+fn5+fn5+fn5+fn5+fn5+fn5+fnyH/C05FVFNDQVBFMi4wAwEAAAAh+QQECgD/ACwAAAAASAM0AQAI/wATCRxIsKDBgwgTKlzIsKHDhxAjSpxIsaLFixgzatzIsaPHjyBDihxJsqTJkyhTqlzJsqXLlzBjypxJs6bNmzhz6tzJs6fPn0CDCh1KtKjRo0iTKl3KtKnTp1CjSp1KtarVq1izat3KtavXr2DDih1LtqzZs2jTql3Ltq3bt3Djyp1Lt67du3jz6t3Lt6/fv4ADCx5MuLDhw4gTK17MuLHjx5AjS57clhsiRE0oa95s0TJmzqB5NkqTBlHo06gPji6dujVHaXliy55NOw8ggqtNu97NOTfv3xShkB5OvDjpSbhJ6x5sCIk3rM2fA/+ZDkmkhr4JR5+uUrjx78eTs/8mDCkNJazlz3PvyShNIezKC6dfj1JJtfv4oZHGgx+/dIHZDXaJeVgNqB59OmniHnzjCUgggixZQRoWCwUo2CEPWoXhgRDelMuCDFkY2IYdriRhGhQqtJo1hA1CGodTuZhhiTXZAWKFpLE4mIww0kjSiSkmtJobiSgiSRtsdOGHL2oshIYXXNRBxRSCSOHET03cwQQtRiQhhBxbKOTEEpx4EoYRkIAiCkKopAILMKqQ9qUcdNbpSkmKxDLLPLVQ4cctr/yy0Jhlnpnmmge1+WacacxZJ513NvWLMeGwYoQdnFyTUDByECOQkUgqyWRaYvxhySfHkJKGEY/WuUxBQxb/eWSSSzap0JNRTlnllTkdI8c3zpiDSS3PTJPIJpbYYYQZYR5EqJloqsmmm3DK2SqkPoIEJI5pUDOKLsZho01C3YDznThi9PRKNN9hM845BqGzjR7ggYKGQeqAZ1wbJEXCDniHJPPNQfLS+529+OpbHL9MfWLKd7kIatBo7STybbjjnqWFwsTpAStp3oJbnLjkmmscujm5k0YzCg4XzyacdAxPQQXXe29B+XJMGsPZdrStiqTJY0Qal/jSCihDpzFJJQfRE0QaQQBBxxJ3+DJJGoGkm1MvXAznhw9ZZMFJF6TFUVAZBptSSg9/8LCDDqRtY1AOONRtYxq11K133ceM/+SEH6R5csMwNXgRxRGk3WIQ2qSpzbbbcKchd0F0201a3nvzzZQvpAVRCg1XyJB0LcZ+HAMMMRBtNNLHMW2WOXuT9kLmdbvwcRpCq3500ks3/XTUU1d9ddY4qexJGrLYwAZp7eHSwg+ktUIQ42k43vbbcc+t992YZ953zz5PyK07cKzDwkDnrEBaFAaJAQfeKhSUQi1piMPrTSiQdkIzBTVhizs6IkgOqMADExSkBCRIwyEMkZBrvGglRRBfQTbRjBFU4yADLOABE7jABj5QKlkgjTgCmIhzJIM0hejFx4JwgvKdTyDpWx9bUMgg8pkPfepLA/sK4j74yY9+9ruJyv/SAARFJAId5SCNDARSCbjNgHIENCBBEKhABiLEgTMCX0h+JiTSuEMEBgEBaSpGkF58oHpEMkg3SCMFnLwDbh6QBkI4cJBfSNEg5iDNBjyYRZNEgTRLiIgdEZLHNOzxih+ECjk0gLUMGGQT4SBNJ273xTCOcYY3AloaKlkQMaaBjAMxIxoPssY0tNEmKiMFHQWCgTQc4QIDQdwPCjLIgxTykAfBYo+06BEuImQ1ZEDIgMpRkCpkDyF1SIMycAIE0txgI4ggjSb4uEuS8IA044CmNKkplSuQBgoIUQNpeHG7YB5kmJh8T4hIY06DoJMgxpRcQpK5TFSmoQIEeUYaGkH/kDms6iHRTMM0EdlHXoYPRdxaTkGIQJqbCQQMpElBQk6ojps4IYnuuF9FyKGCN1wzDbrgpkrSEQbSsIMCsKQIRz1KmpAStJpKKQZpJpCQ5emgD+JRKEEYmgaHooWG62zQQhtKEIimQaIIoagQ70mQjfFzIAyt6EJW+lGX5jKRBv2IL1UTn4PggDQa9acOGkHWspp1eUG4CQxIk4uKSIAHcaBfcTwhUpVQYx7DeUEOiFE6hrw1rsah60ulEgqwJiQCpNHUQEQ0kK+mQaNnAWpCEeLYsKZhrGbNLFqXis+BOHWnaZCqQf4qV+II9qoFzapGtjqxrhqksgQpqc6gdhMs/55iImQoxHA8wA4wtCITgaurSjZRBHZEjjRzYIYSEpLb3fb2t8EdbFROsEmF+IA0EMgpZQ2rFslqUqeN5a5AZKuztNqzswL5LFRDa5Dmkoa3vgVuGk5rEF2qVlsS7KJQCQLbgSQzDTcIm4AHHDZP2eQMpMmGRGZBGnVs4xGbGIgTootamJ7kF8Rog6pIY4RgHITBoX1whAUy4fkK1yniSAMcFEII0kBDu14V708zqV/wCqS/AvlvgAlMYAPXRGXoTYR6BRLVgoDYwRCWMIXri9X7boS1t7NxInCciFamoa9AeQBpPhARZpAGCHdUsokRIoYmu4QFVYjkZb9QEC8TMf/MJF4yD83MFN2mwYoH2QFp4rdY1xaEymeBmzonG+PHEsTKWP4JkJu6T9CKNhFuBrNBSkzfOafWyRaBMoxfK2MHkOYPRGkAaYYAkfJEA7JxHvNBVECap9JkGWwtiKlRnQhKJ4TVjY6KDEhjAYQgQ1Wk8Cljbyzjs7wgDQxgkJSp7Ok0gFooi/ZsronM3oHM2llyJgiuXc0VRMSCHHLRdJ/3G15DD8QN2EgDK5YrlOumoRgQeRgO5pjtgXCANAuwySYCkQZwFETe9Fa1Qe6dhnxHRQGkiYcKDXID0kQgyts1d1qWNwlaQ7zQGkW3utkNlGind9qJKPJAAH4QgleaIAT/N/hWFNEGDoMgLuIGkJ/5W+xxkMYM8DLIBaiRk2psWAYjJggLbJCJglAhDVTIOUESAD2BGwRxQSB1Sp6AghAgZAukWYXRka70gTC93rGEmtShYuUdEuQdKoOD68a97GKbhXNp8HGNIw5Zm6cB5wfZefGYKm1ui1wgR0+6Qb7u9IJAfexZgQBxrOqWmCdi2FMudgjaQZof0OGCAhmDMBpQCCPiBBpPcw8CBiGQXzCDFWkoB5axgO+XF0kBeNABHPhQeIIAgzRH+MLNlCCPLIikBvpbhBo8f6xYoD4NVSgI6wvuekXAXva0P7ntca97gfDe90r5hsGK8QgjwkMYQzxA/2vJTWyJo+VE5ZBCzlmwhx5oFPKAnnzlLz8QzXOe+DTxuJBB/vdELH8Bzfd8s1d7A3F7rlR9iXB9VXEMxPFobeF48Od2HGAGw6ED6nAEBjAc4vAAOpEHBkMafJBEpHEIHVAQo5BipKEBrJBAaeAH79By0idhlEcaJCAOHjCCcMYRzrABjASCDEAE7bBhaeAABnGCw6GCLOiCMCgmM5gGNXiDCpSDRWENSTAckyCE2AAMXEV+kWd+aLEaUGMEH3gbbEd3A0eBpGGBGKiBHHhejOZ31SYQRpiCK0gaSkiAJNaET4iDVBFBwzECMJdfvzRz5UZrigABRHBcpAEJWZAZO/8BD9uQNMPBB74wMAYRCTmQbsNBC8SgQu0RgwMxCOPwPsRxAnjADSPRBFewAqQ4HEFQAVpziZlIHJzoiXgoEKLYivpzikyBDnLQg52zCmcwiFwIaGehCMKQgcRBAqvQDZv2Z25XJIioiGnAiI7ohn3naAWBiZpIGrWYCJ+4ELlYHKaIilPhDI7FB4gCGtUgCRZgAYCAU0HhBm/QCaiQAM6wEH3wBtrwBmsXESFQA0WgDQWwBUFXEudgBQXgCARwDeCmj/zojxQRkANZkAfJFL1wBhRAD5KQDoiBBglAAcRAAQOADCDRju8Yj1Oxj/34jxBBkQRpkFexCV9AARKDaTj/mZM6uZM82ZM++ZNAGZRCOZREWZRGeZRImZRKuZRM2ZRO+ZRQGZVSOZVUWZVWeZVYmZVauZVc2ZVe+ZVgGZZiOZZbYQVmeZZomZZquZZs2ZZu+ZZwGZdyGZdjUBHeMJd4mZdwKQEd4Qx6+ZeAiZYXiREqEJiGeZiImZhomVIbAQKK+ZiQGZlryZcUwQKSeZmD+RCFeZmc2ZmQSUIM4QWiOZqkWZqmeZqomZqquZqs2Zqu2ZqgCRHT8Jq0WZusGQsd4QS2uZu8SZoLtxHQ0JvCOZzEWZykeY0aIQ/GuZzM2ZyniZsUoZvO6ZyZ6RB5MJ3YmZ3NOQwQQQHe+Z3gGZ7i/zme5Fme5nme6Jme6pmesfkQ07Ce8Bmf6AmdG+EE8nmf+Amev6kR0JCf/vmfABqg4ImcGSEPAnqgCJqg48lm0amgDlqdDZEHDjqhFJqgZEiWGJqhGrqhHNqhHvqhIBqiIjqiJFqiJnqiKJqiKrqiLNqiLvqiMBqjMjqjNFqjNnqjOJqjOrqjPNqjPvqjQBqkQjqkRFqkRnqkSJqkSrqkTNqkTvqkUBqlUjqlIboHCXClWGoNApCPQ/EI+MgTyHClZYAQApAA1+EUSnClXVcQA5AAY1oRo5AA5tgSIIClWDoAAfAfPvELCVCXCMGnfqoTVmqnCaClXCoUXnqoORGmbv9KpmZKFHVKqHiqp1CRqC7hCDSQqZqaqZYgCvvpEIrQCxBKEsxAAwR6E96QqdrgUwPxBTSQAE+hCFdAA++AEH2QqaRXEYhAAwrwEkWwqcD6BlKIE2pAA/IwEJvQC8RXrMe6E5gKrJzqqRIRqqM6EqV6qjaRqjSwqgfhqrA6FL8KrZkqrCpBrR/RC/t5rZdKA0vwDu6aAHnwrM3wqQyxq73KEuq6E9pKA1bQra8KFddAA712ELtAA5Giq7zqqzQQDO7qrgoAAZlaBNU6E8w6EGRAAzwnEBXrrOzasPAqr/S6EPbaEvmqE/varwbhrUTxqwzbsA8bsRPbESPrETdAA7//WbIrgalyVBCPsAE0wGcAlbAsga49kao+CwurRBAq+xRlkKkeaRB08LMXMbMt8au2QhCVkKmUyRPmKhAXm7FFIqo8obMG0bNSCxFUqxJEyxNGSwNIm7L/Cq40cLUDkbU0sLUmkbYbUbO/ubYtQbYGYQ00wAyKWq9CqxapmgWSYLD4lwhL+xSdQAM0dTY0sAFrKhF6qxJWexDNQAOSQBRfKxSAWxCCS7hoe7hpkbiL6wqN+7hBsbkG0bmfixKZmxF8SxOjSxBoMKtAKxBDAAid4AhkUAme1wQiQAE0sAYigHi/G7zDi39nAEbwAAGpcEHvIAKbMA0FIAXKCy8cUANr/0AHBYBniSAKIvBClSAC6cACCkAPqeAF78CqidC8wku8BBG9iTC91YsRiesEvFsQrrsJiPAGdLAGX5BoYiACn6oAInAzHCACApAIEkAAqWABJcBxJzEA22oQCkADCFAQvTDAdHAHgPCmAsECIsAiAVAEjpAI9noBYkAPUlAAAxCyIQG7BdENNEABAlEGIvANipAAxHChIUzAJGzCicANIjAKlnkHf0AGiDCYATAM9NAJ0NAsZbQHAEAHjgABbkB8o7C8iWAIIhC1wSACYxrGiCfABGzAiXa9m4AOgOAIfwAAATASuTsQu3u2A0G/zysQxou8ysu8wFu/0Cu91Gu92P+rvdwrAt4LvuJLvuaLvurLvu4Lv/Lrx/Y7EPirv5hnEf37v0obtwLBxgV8wDykwAXBwA4MwRJMwRaMwRmBwwShwzycCNMgAkiwjSLgOmggAqbxAAXwB9qwBEgsEG4ADY6wBvIwCoB8vMkrxucgAo9QEEogArWKrNZQj1JAALtwR1sgArBgrOdbvuU8EE0gBgDwB6jgkLoLzIkgzMRszBWRxwMRsNwpEGiwuDRwAONMA1WgQukArM26z5nqz5ka0AMRuRyQCZk6pr8KAhuQCVFLA3QwBP7sCOMMCzMjEH9gqgIRsClADzQgBaVKA/yjx/yM0AD9mwzt0DRwzBORuIn/kAI0cANh9rghQACcmqkbsI6JkAU0ULilmlJIwKvDQANZUNFSYJIoYQIPXRCY6noCYQKukKk1m6nVDMgeLA1Y7cI08AXP6rPJa3UmQcsDEQw0kAe+K7n8bE5WjdWautVg/Q7a4NOZigpOnQhoEAs+fQCZGgyM6QwWgNWA7bkDkQHGmgggAKykptjNmgg73dOVC9S/GglCTQxZnc0gYc8hTQP5zNcr/c8KPdCbWtCjndAuTQMNHdURPdEVfdEZvdEdnQgf7YgiTdImnakprc+p3dILzdowLdMSQdM2jdOj/K2SzdM0YAk+DdRCTdQ0YNRIrdRMvdezPLcHodZsnQiC/7sLBVGsinUBNEAAEtDPGl25eIYGSd3cR8sipq2px8oByVsQqVoE6FPYzZ3VWYAOAgENwJoZtz0QkZAKlJ0Fa0fe5o3e47wB5AsRnp0Ie1DeAyEGFv0Am4AGbvAKNADeinABZ0AD13ABN2PhdIDhGs7h4C0QkUsPxCABkXAvvwoLiGBEkUDasHQOkcugtg3SiRCwUkAP0gEDmYpnJo7iG97hwe3iMC6/xa3UAlEAKJ3cA9G52jAKm+AEKuCzaxfdbTbdAnHUUpAFEZwIZTCrA6ASyLviAkHfsIDdV40Az4EMouCzftoENKDZqDAEgrKrbmsNyMACASDUHnbW2r2NPv+b5vNLAyQdC9/g34kQ53Ne5zTgp7tqCRugAoE+6KA9EBp8A27gDGjwDaUKRgJRsBZwQYowBOMsj5CdCL1wAZ1bCRcQYa8uEFaO5VrO5QPxq1JQBVfSCyXQ3DFLERE+4QRQ4Ree4Unu4SAu4iQuEEfO7Cq+5C8e44kw4zWeCDee0Dm+4wMx4D9e0kIuEEROA0a+7Cmu5CzO6Nfu5BFB04kg5b3tuKSc61m+5TTQ5UP95dQ95mV+5jSg6BqB1tye6ALx3eEt4gJB3sxwAGegQk2AvNklEFZQ0tOQrNbgs4bw4SE+4vdC32tg3zSA3wLhqnfQ0emACjQAnc5wATXbBCn/Je4sUKpk4N8msLiWkHMOD/EST/HGTgM7axACADoCEQKWcACQLhDc4LaOSLVIr/QE0fSw4IiRmwVh9qsXmghLYNEO5dWoEO4+HrCw8JBee7dHn/RLn8RO3+5YvxHybgI1m0b2/q3TcNPIqQKM7nleThBFHeaZSvdg3d0o0djPJhCiQAPtlAgBQAOdsJ+JP7B4brC/6ed4+wuZepMj8avDkAGenwGIMAw+qw1cOgSZqlgC0fiPTxCRLxB+TkJNuwH+7QRCHahm3s9mfdW2PwzM4EiJcOuJELoCcet3fwN5v/cC8avaQHybINRP+xERXvRXkPZSPxBU//SHG/Vrf/1u/5/1NLD1XU8HX08DYe/RY++2Zh/8aC/Zaj/1bZ8IVz+sMw3liSD3NED3S1v8x08PfN/vfg8QNC4kSoSEBg03BAkiopFH4UOIESUqLEJjWAaMGRAN20BDmzOC1mjsgqiGxjWCFw5+eViGxhWCTW7QSPdQHo0aC2koUMiBxhqI3mgUUdiJxrSHo34+nNlL4R8aTQhOoOFK0cNgNOSlXNny5USwCh3RkDZxGo0DOiVFdEWjkk6eatm6JWj0DMSKSB6K0fowEg0CT6MSvEbjDcTC1uRCbPs2kd2wkSUKzaJQJDNkBL/QSEBQBI0SEJ1loeGNIGmQD5kJJGjw1VWFAgxLpv9dW6GTAzQMFUX4kAxZ0R0zNzkY6SHDVE4fFqAxwfZziRUPTp9uwTTBITQsKSf4u+xDZ8ITMWSGBiLzAYkq0aAXEQINNQQJ0Eg/MUPf7jSoKby/NdHn0MAj7bqKVIBoDRo4gO65sb6L6Ky0xqNhLcbokjAuCSl8qLG6aLjrobz2wo+gvwIjCCqpEinssIcSW2xDCyFbECzKLKMBM8048ww00QY8jYbUFFptoIJoeO0h2VicMTrqmrTORpIeMgmlRFSiAZ6HWKChskQkoIEMiMoYpj6GMPQJqIeEIoogcgaBjaBfaLiDKRq4Q5GgO2gIACKlmOHqyiy3XLLBidygoRP/gmqgoZtIGnW0GfrgUkhRRh1tFNL6jHKMIhrKeEgBDx8yBDDBUiwMQ4J2oUGxRCi19NJIH7NwyYhqJEiRtgDJsbNEmEsIIlRoGOPHIAkasjVSHxpjNlpprYLHRISypFhtaNji1Uiu0C0R4mhw4jhmH3rEomafqwgVEdJNdwAB3kwkOwsgqvbaV7XdjaFmIhqXpX0joirKcWkAQIKaIOpPod/2I+jgXnsDVliCKlJYIQQVLBcsQiUyFNFWF8UWU0kT9fhVkGXdNOJOPw1VoVFNTOROFXeCSFVWXSU5Vk0vTlPQW3PdlSBfIwp22ERQg+jYIl0maFkldT5X3XXbfUik/ygVmvJP8xRyhufC6puozJ6W2nnNhxQxgRtrjJpToaZKJWimcyDapCOQrMyaoK25XDDjiAb4kiAAmmxSjJATCVzw6QiXFUsQU1YIVK8TadntmBGZeVXAEU+8Q8Z1jpZngjiARb9ENuPVKOMgGoaGPYg9mrUiwVQ2XM9rW4+ZqxJoqOzRNafhG25B10mEiAKgAZXaI6sovsiya1qR3jUHnqGcIDIegI55fei+Km5NwJLprpAHHf5GTLh8/1CPaPXWE6nI04cqTh4iviHyW/bDNVcc7Mx9V9yozqEMfgSBnKiS9bLBVO5yrMof4v4HqNrZKnSj24/pOpS6h7DPdaqBnf9BZKcQpiVveZKhWklOwpUbiIZnWXGQRPiXiDMFZSgKUUQGUAG+g7xCTnSy02CUQANYuIsgUqBBNapEgxSCR3gMAo5EmEO8hskjI1PMQB8KxxwpUhEjVpTVdTg1wEQUkGUHhFlhLPcQmgFNK1rcYoe8qDMJEsQKNCCGMyyYCARNTCHPAsEGFUIaInkQIiGcH216QZr95EkvdFIBGzNgguD5CVwsgQgMaNC9QjKJeWHJzhIgMpNGshGSDBlGRCxZgESoClUE2QO5FMKCIUiCGAexBDcWZj7S3dI/eYwIHwV4oARlkiD1Ww4NoIhFR3LxhchkIxeN8kb3OY6AsSLI5E7/lEAzLlCNWdSiM0uTvDgmYo51vCMvIeLLogEJIoBE1geXRruLjTAyJZzaCY+YRK3xbAk0yEBYXhjDh/iEbPukAQKs4AYOpGOHbKuT27ZGg7sJiQbkUwk+8bZE2xAzdAfhYjfgE5kXenSTE3kmXqQZRmpKjozYpMEZFZLGRIg0MiWN4BLRUK1d3DEPNJBARBCElHQWS0sd/NvsmibMiSgqFgrNAncIYoFvgoU4klQIQxAQEVHQQHFI5dRIJ9JJiEAVmhBhiBewaswu7Q4xzZlIOgIXr0QwLBHn0yVBdtpTYAL1fcC0WCY1CkOOEkSm/pSZYD8aFpp+UWWRsyYCTdVS/23G9LBgSSwcbYpTnfLUp0fxYyKGGsiighCe5ZJnWOipkMJQqaIq5BKoRAERNCABeBcKG5piM8NE/CUTYFTo2t7W0GumSFsFeyUNNrCJe7J2UE18CDK8QINgKAQENEBl2SBQBEjSliDTrS4Nr5vdykYTjGKs5kof69JUYS4R3IWIIr7rRnBi9CwbQBCvqIK9h6CjI0r4UYoI8o2DgNadiSAkVyfylyxkNRYRgQYNHgERchQBAjG5EVmBGIKHbGKWvzJwNL0qEbA+pMEPfkiEJyyhDZCjbGPp6VmywIIM04EGKeAWAgCAXIW4xBJ1zc/EGHZfiOiXBvwVL1+RqlHnQv9XutRt73u1u14mWxe78G3ceFPa2DJCFo3qZa+UwRtVz4UzEfOt71QEFuT99vchACbqgAvstMmCRST+UUhbVItE5SbCkr4liGyQ92SfcEwhfFmTCtSqkLPwuW3BJUiwPlQ+jq1WiXqDDt8UkQ43yPgAfXVGKlgnLvZUdSRa83T7CDKu9lBZsY+7snkJo+WXqrfTn1YIqnkz1maJuXTT4ZUSZrIngqCBOaUcpsMSgQYEtXmQo+2wQmQMvgdExCVZGAUNsxKgqVr4kjhORFbpwG0DlxYsIc7xlqp9q2vrBLrcoQozYIxHtCrkDDSgg3IyQYNFEsRQaJLrb4gWV/z4Wk//ChG2K4scv2AK09KY1jSnSw3qVEuoarM2dSJsrepfspqxrlbglllFcYjfOr6UVshmDtLrXxN82GIxNrID3M5lH5W0cS7UobhtPHtKOp9cQgNU8JoIRTzXORJXSAiAmN1EkGMmazKUWW/TlognYiZBgplsmEHcEMi4j8md9HJhYQmwR+8VekzEFg5SBQnAYKewiHZIaGABbhTM7JdM+9rb3kWTWnnjLssyelOp3rKfve5AvHt4y6VrJ2hLR2bewBdggAgE3UDFOcoEIiKxBQswwygCjnmzJzLHlwiRIIBAYg1gIIG2SCFukbRwtQigAjVAytgdFvdXaeBJiJD+BqZH/z0NVK+TmbhCBRL4Da0T8YCONEMNIHjDQYagkM/AQh4wcIMIwNdPgNO5+cOIBH/lSpXGPz7ykz84xRLuVyCC3RJiJ/vc0a52wtsI7nIXPPzZLvIqL9aAfGep32EaeLqzv8IDM8siOYJIvJNTCPBzPMhDIvLbjMq7vMzbPJgzKhGiOYkgh9xoBgGghl3YAArIOTzrOoKwJIuQBhCYDzogEpGYv2LTBhiYBhVgBi/YgDUJgZmQBGqghndgBh1iu3cjjUeIBKeAmUQIllewghSYAE+7AxzTuYsywNoYiybJgiIQhcwoHtKYDikANoIog44YkQDYwoPowocIr73SuP2jnP9s8riHGEPqMEP8CzOMIgic055E2IPcqA7iqpJkO4g/+IX34DwL9LxaOQgoighFEIUwPIg16KtsAxc3iAXquIGToT0MjAhyoyFGpI5HFDUR8KjpOICtU4hIIKLpYAYBeAg0eBbqyAQSyz4bcT5ZVAg9pA4L6MM0NL++KiQqpA4rxEKJgEMu9MJEAMODoDNiLENjRMOTIi+V4r/ziiw7JEPfa0YCvBhdS4Q7fIhb5MOHuIA/pIFAHMQKFC2Za5basw8cSsa/uDOLSoS8ecN7m44i6CtkxI9f8LTpQAVkgAWySYGZmA4ACIFgoYHnSwRXHAwjRAOTm44lIBKu2zlDdAL/HnwEJIioLzyDBMCgRLDId8BIjTTEsABJkSTJ51ACARgACegDp1KIb3iHCUiBYkHJTDKBSkiAPQAq24CHR0iAMchCmywXnNRJntQJlCAHN0iASiAy0ZDBR3gAiQyoDJiARwiAbwkLARCFARgEsFBJlnTJodQZk8xIaeNIjyzLkRzLh1BLtowMsGzJl/wvmaTJt3yOEMgAUUAEI6IVJ/iGAbCGBxCiMkBLrUmBBBiAwZQIJZCGCZCGvmwCEEiAyUODDLACFVhLgkAHxIQBXLtL0AxN0RxN0ixNz2MIKjFN1VxN1mxN13xN2IxN2ZxN2qxN1URN28xN3dxN3uxN3/xN/+AMTuFkS9wcTuM8TuRMTuVcTuZsTucEKXt6TumcTuqsTuu8TuzMTok4B0MgP+38TvAMT/EcT/IsT/M8T/RMT/VcT/ZsT/d8T/iMT/mcT/qsT/u8T/zMT/3cT/7sT//8TwANUAEdUAItUIiYAARNUAVdUAZtUAd9UAiNUAmdUAqtUAu9UAzN0AwtRQPFzy3QUBANUREdURItURMFUXDr0NrwAhZtURd9URiNURmdURqtURu9URzNUR3dUR7t0R5dMBXFT3nwUSItUiM9UiRNUiUl0hQN0shYUiiNUimdUipNUiB10voc0irdUi7tUi9d0ibF0olQBDItUzM9UzRNU/81XVM2bVM3fVM4jVM5nVM6rVM7FdP6tFM93VM+7VM//VNA3VM8HVRCLVRDPVRETVRFXVRGbVRHfVRIjVRJnVRKrVRLvVRMzVRN3VRO7VRP/VRQDVVRHVVSLVVTPVVUTVVVXVVWbVVXfVVYjVVZnVVarVVbvVVczVVd3VVe7VVf/VVgDVZhHVZiLVZjPVZkTVZlXVZmbVZnfVZojVZpndZURYYESAAwQpKOFM3JvFZvHYAAgKZ0SABjpNZJtVZsnQgB2NbQ7FZvTcxwhYhxLVdzHU2h8Ii1vCPQlA7EeQOkMwjcq9dJvVdtyNfF21ff8VeFAFiBNc17pQErkAh9vcv/igiGd7jYd1CA9xgKHGPYhpXUh43YiJjYt6xYjM3YjS2Cjr29jyVNoegIWOjFnwnNdVwPzSqSgG1ZR31ZIJLZXcPDksVAm8Urj9VZ0KQMSagKISJZtlzHRIAUCilao2VUpFVaiGDasXRaqEWWnJ3asaQMBDSQh5jYTUCEN6CDNfiCo0wEMRCBl1QAEcgaDhCBVZQAAkgFCygBp5wIp/UoCuBaRVCBNbgCC5CH1fMsEVCMACgCR1AIZFCAZqADR0AFFXg3hagGSViDVFgDeUA6hfiFEnCFP0AFa6hJr9VOsNUWsS25g02Esj3btF3btn3buA0duu2Su83bvdUkifBb/8AVXMI13FdKXG5kXMeFXMmlXMslCMzVXM71XDgJ3dEt3YatkRRAIs/V1xCYD+04iA14rT9SJw4CLQVYnSyQMd8TSt6NiKxwiCIJBubIAqhIXwpDAGk4iCSahtWggUzYwzW4m/s1ri28gTdSgd7BoU7AsNMFT+vF3rE92O2lJe8F385KBKQxiPLdEvSVAvWNCKdtX2SB3y2ZXw6u3wDO3/3tX0cEYO8d4AI+4INIYIG1FebIF9bVHkjRhlHYBCdQgY7YFKMZX2SRgixYxWPUlsjxYAyMhI6oD4PYgFeYrT5YDRIjDmK4AVQYgl8otiVYvWnwNBggCGQAn0cwj3N4D/8bTgQAuwEJYAFFGAV6qIoFZmCeqeEHxmGP2OEe/uHwrckL9r0i/kIkDot1ZOJYeeIoJogpdjAKu+Is3uJEGIsuJogvpoEwToQxdjAzRmOYRKI2fuM4doUZ5hkTmAkOu6OzuAH/SgRDowfYCGKJAq3ZYwj35VuLmKKN6IiPQJaBU4h3oAFdCZ6q4A7imJaH8BsqATDbgiFm0IZbyZNLFmMdyrc5vk5bKWVjQ2UkWuVWfmXxjWVe5jAJqWUmuYiMyGWPSA2D6GWC+OVg7hZXIGbtKBZk/i+xCZ1mfmZLbq5prlcJuowsvCMA6ZGogmVjIaojua105BTNeRJkSQXroQH/aKAwGvDIXhgE6B0XKKIGGkiFDi43OogIvkjEarbOf76RgF68gRaQgv7mgwathO4zZmNoxHHoIoHoN5RoirZojIYIjSYIjvboiXCJkIaIkfZn0MEVYJ7ZoHkYojFoCyYqpSGwmUYZdIEadnEXg7iqh1iPiRbmrIwINEiHMZgAHYKiCyAiZrgGaljeRLheVMAWUDmxkjbppPaZn1UjcW40iAmq1wGtqX4zJb5qdcnqh9hqiPBqig5r2CJrs463tL4RtnZruJZrGqDraY0j0SGdO1If1aE1qP5jN6tqD5MMg8AkhVDs4DHACxCFO2jEs+6JtvBeVMgA2Jg33xHlurZr/0rb7ApaPM/OINB26agmRHQk5EyEiNNObJ1e7Yho7deejtgOndk2rtq+bd+RY3MNp3Gyo8Uzp4dAJ6hmp9jpPLBwWuVm2a5u7ki8jWq5gSp4h0owgekiaRg6AwTAoWa4inFBAEei5t2WTu6mI+/mFfDeIxrYuvFWtkK05Q+LCKlVD/ausLZ87/ie7/qGCA7Ab/3m74L6b6QmuZsaicz6OfMDKoP+rHN8p4VGmQdP765V7fYmCN2xAPXNcIlAA2kYHb1Yj64N8OwMpxHPqcW7q81CcZdW8fJu8PUNiwiXcQpXiBq/8XgT6x3HNwn/8Y8VMzJbPCDLrzRLp1VmM+Nmcf/kfvHDVu/UnnCqIojmM/FEoAooegT7hYjfeIdEUKjyMOoiuEQgb04uN64yi/MzA/MhU7NOZvDjPu/kTvMYZ3OIePMDjTc6b6G5ogE813ONFIM+D3GJ5TWCEDgvLDhii+SWS7Yyp+oWL+3IeHJIf2CRZV4dChCGcAVu64U8MeLfgAZuGwVY2IDd/XPm1LWHRDl2JnWW4zCXU3QzZ3Q0X1g1Jwgob/Ndi/VEqIZZ1wlbVwhcpwFdl+he//Vgh1bEUzztWUDxc0DWjUDM0zxmV/UzN+1ol/CvnnEJP4ABiIRKuAZYWA16KAOzubc7UIEA2IOK6ITU8AbwIYABGIJdmAn/oBX2Ya9DBDxYdG9AyVt3y2t3ClzyRXdweX/0eo9yaUeLfN/3fmcPgDcBgSd4gz+UhF/4hn/41s3sOuTGBLTFPTyIXAzHcSzHdxdsk3p2rmXukaf2mJmODSiBJqhHqRAAfpwOCvDKyzWK6bAECpZ459zGbtR5XOxDcaQOoE91oW8cosdZo6cwpC8MpWd6p08EqG+SqX+IarB6Wsr6P49LsYyImJxJ03XNalCDCQgABfaGkMwaNKgEFRCFDIBkiOiFUVCDBEiBcdd68tT7uVTjuvz71gz8wS/8ww82xWd8x38IyJd8yrd81V991m9913992I992Z992q9927993M99/93ffd7vfd//feAPfuEffuIvfuM/fuRPfieVgAFofud/fuiPfumffuqvfuu/fuzPfu3ffu7vfu//fvAPf/Eff/Iv/+kXPeWHVWj4UvZvf/d/f/iPf/nv0TBNf1dd//nPf/3ff/7vf/8HCC8CBwrclOggwoQKFzJs6PAhxIgSJ1KsaPEixowaN3Ls6PEjyJAiR07sZfIkypQqV7Js6fIlzJgyZ9KsafMmzpw6d/Ls6bMlyaBChxItavQo0qRKlzJt6vQp1KhSp1KtavUq1qxat3Lt6vUr2LBix5Ita/Ys2rRq17Jt6/Yt3Lhy59Kta/cu3rx69/Lt6/cv4MCCBxMubP/4MOLEihczbuz4MeTIkidTrmz5MubMmjdz7uz5M+jQokeTLm36NOrUqlezbu36NezYsmfTrm37Nu7cunfz7u37N/DgwocTL278OPLkypczb+78OfTo0qdT/wgiAfbsAwJ4E/srwRiH38MrvZBAxcPr2bFv754wXYIA1efTP12EBv78+d+YAKuGhjwIbdKLIgj9F6BSJtDwx0P36acffwghQcMS9Vl4oWf3BfMOh+8oAAF+RRjk1YEIkUEDNQYCuJSCDDqkYYcegkiDiAdNWCGGOepI2X1qLFQJfhJ8pUgvIyZyYooHEWkkUi02SIOPCgFJg5CJ3Lgjllky1iNDzdD/IElZSEbl5ItQdvmljRRquSabg3G5UDc0UHBQGSJ8o0gCxACCUC+IvEHHHYCUkRA3IozCghV3/EEGIkweFMAw9HQCzRYL9bIHAHQ4AoEbBR40ighDJGKICHTQEIwIg4IqqoB+0rHGF9Mo9I4Im6ADiCN/ACCfQt7UQEEq9JARgKeJkNnQmwrFOaeVFCqiwhpXWCDPOQixIII1iQRQhCMIIaNAM5qiogILClUjyRqprCFPfwr9UoIrf6BijTNt2ntvUckmFAwNeRw0BA0TSIIfGQeZ4Ap+N+T3CEKI0PCONvhtgB8qyCCERiwSH4BfMBcg5IwFCW+M5kEZrAjCg6Ka/4xgIiEQgJ8lEouS0H2RZEEDMQo/nJAKsOCXxcQ0lIDQsQzpixC//jYbTAE0ZPEHflJYnEgTNCAgTcIHTcMMfpmMvAYaCGFNwwY303CDewf1DDN+nYSAL9xxf3R0IpFMPMC/NNBDQyzfoHMQwgh0h4woE5PnsCUbqIAMCwHcPAxCA5zthjNofMO1CAjtQoMF1SSiyBA+91Hyir1c4GUlFxi0MkJeajPKJk6oMHElCN0nRRVOJNJLCTRYMiI5Pg9Qby8gTNxu0QvRbTcNeDe7wSvfHNQH1wxTjfMNqAzxy0GOUFhtItOkQgMMByET8yNhnwNiMwh9c7YELCgyyt6uyP99P/4W3TdMBv1ngMgwJqaNeiUCYDS4RkICQINO9CIhouDcQRxGg2whpAxk+5sTbkYeOtHgAG9LBMI2mIhhMCMDpGOZmE6otbM1gWd689R9tFGsTdwsHXlDhUJAJACDLehJ/PMfAAVIwAnRgFcHeQcN9nQ9GriigQepmiUIeBDJITAR71tDQjjADG0o6Q7kSwgyXkEDJOSvjGY02oP0Y4G0AcwSTjzIiaShEGdMzGIOY0bYEtI0vAGJHgsBUZRe5jyGsA6OKEJIIUUgtDnezD33QU9C1kADDhzkAoOYGkKaVjtj9bBMacTPGiVEg1QoRIHQeCJ+IpGQXgyiXQh5BA3/MpcIaowSkwqxIB0WIoZYnrGXZ7wPKkQgTGEOQADFKiAEExKxLUSimc68Ag0MkQiHtU8hsPxCIq65kAnQYBcHgSUNACABGyqkkEc6pAoT0TQ3LAQVNCDPfZKEEElSUiEXgMcQ5IGfTSZPIcAcJjGNmZAJIUBKNDjlEnXHEDSkYwwTEKMsLyAFGjDjGtQoV0JSQANUONOZCqABBHwpUvzRbSEAe0NCFOGzT+JHeg6rwUIUCIBE1IAGCViIyaqgpATEDD9XkMff0nlOeRayEzRQpUKGQYM9HOQ+g4rkJBEyiC9ADT+wyAQN+NlJZJkJIhPSaUKAhNCqZWEhFxDFHYJG/wOIIoQDCJMYKjLgqTOwFD/2Gyle8VVShQAMRwhRmAr8J9gM9MdhkFMIDGhQgERsTgEL2QMNDnsQFgxBEsSAGTeEmsJEFFKS8kRIFWgAgqbS4KnzjGoifqEwKUjCGtRwAjSyykMXcTVKD7lSWA+KSmYoxAkRu0EV3lEJE6BMlm09AwJ62owCwRIBg/UfGfMqXTbtNSF9VUjI0sYQh3lhIQ/MnAT6tZBrBKwh6QBAMs252ULmgUoLkaSsEuFUhdDzIAj40hsTEVut0haNtnUIbhEi1t0qJAGcs2VxG4IGafiMjEDy63QjfK/qIuS6CYmt9RBCjiKEdJpkI0dKvSekaf84DaMH2YSpUkA1BACASRa0hGbRydkVJYKbM00IOiamBNKa9iD1TUQNFeJO/j7pvw0J8EEGfD3eJuQN7k0IN2X5iKsp5ETvSEQ6KJpHhIihCJuUMJgvROG8QTgRFszCKBCiCH4NzcOneiM3mYFRSRo3EXSlgxOxGt2DuIEGWJwxCt+JSBorQWFGREPTJDtfqNbTew9ISJ+LONsiRwTJiVBy1ZiMkC/QwAoJqYYY2+wwVxipF17c4TmhYaRRwGIDOw4zrOkzZmSWORGAOFsNYCABhEkBfA5TmCtUIIETLRUhD5hYM9QAAifTgFWJUCQs5AEDN4ggZiYENEKcPIxI7Nj/nNzcwBdggAhJ3gDEPKYvahX5hwxEwg2h5VoeBsHJ/iqvq7dVU27HSlGDHmAAkajENWDBNXqUQREmwOodVBCAPdynEwT0RswIMIAh7EJhN401xqkzawuntHD6WUM9PSyCOOXnAKNNSCQmmh9moPogaAitfjKRYXNaIz8qo/FB9jAyUJLz3Iw+iDPcmZ8sSIDEC5w3pb2KbwHrdsnj1c8GStAErNKghQIYn34oIG+EVMOo+bHEzDIu9vuZoBIJ2EN8G3bARJDDDQmoxKvnOA0VPOIBHlsIBzIwgUcEQKEOEYAoBrB1hihBAAOQQB/yWxFuIGIA1Lj7NCbATqpUQw0T/wjAB73xjvS5vBIqEEUGuGepUaghASmI+9hTf0aHVVH1rn89blgP+9nTPjayrz3uc3+a2+u+977vzDkMYe7fE7/4xj8+8pOv/OUzv/nOfz70oy/96VO/+p1xRoyyr/3tc7/73v8++MMv/vGTv/zmPz/606/+9bO//e5/P/zjr/5jWt8zTiAI/vOv//3zv//+/z8ABqAADiABFqABHiACJqACLiADNqADPiAEJqCj1B9n3F8EXiAGZqAGbiAHdqAHfiAIhuACTiAFboYznCAKpqAKriALtqALviAMxqAMziAN1qAN3iAO5qAO7iAP9qAP/iAQBqEQ1mAJFqERHiESJv+hEi4hEzahEz4hFEahFE4hFVahFV4hFmahFm4hF3ahF34hGIahGI4hGZahGZ4hGqahGq4hG7ahG74hHMahHM4hHdahHd4hHubhaygBdoAPQwxAAvTYRYxCAmTWRajHerSHQsCHEemhI4aGIkDTlTFEH+DH4GGEwzjWRTjIJ0VImtTaI4biZpCXBTTE5tyVRmQiRsBIh3xIiIyIpYmiLGKGBdFAzyWEqUBSKtKAJuqPvTHdk8XiLA4jZRjVBCyEBW2AH2aEKm7iL7YOyQgjMU7jY0gOFynERxXUKrlKoPTYtWTLtnRLJl6AGNCDFBTAACiefzHEsnzis0TLtPjhN2r/C7d4C7iIC7mYC7qoC7ssxLvEy7xIETUOpFYoSGkphPecnMG8lc7QQIZVzdVkjcN8gfeQDX6swQeViZEdRNKkCdM4TVVJDSpF5NloDdfQgNdc5JaNTdkkjHatje+0TUYSJE1aBQV0UxbRACzYUuAMjseRR9XkjPZwjwTBgjUwjuOcCkQsz92kCfRITyJQj0OiklBuT/d8j9aMT/kkwvk4pPqwj/vAj/zQDxPVpFlWBcr01wMVDEIoEAM5UDJVDRO9kQRVyUH8An6IHlf9UP8EEQ0MUJpIGkIgkRLJZRMhBBQJJBUdxBVl0RZ10Rd5ixjt2VlWZlM4wcZI00EY1eQZ/5IcJQQd0YDFyCVSRdAoKV7THKMnfVIopQkpJVDTkeYqtZI18dIs1VJD4JIu2aZl9uZShFabeYPvCOQydVQzQZM0yaXfeRhK1aZkoVEwAVQxHRNBGZS+0cByJgRDORRbJYJEUZRFmdhBaBRHGedHdZhvpudRAAkzFIiBKY2SrBRLSQ9ZKYTD1NmjbFTS3RtYMZ2+lZU9oZVadWciuFV+bEBczVVdlaV6NihR9MLNpIgXUWYiANZz9U9/ZJp90gA2IRYN9Oc6Kl2ZYdq+JYRvnU1wDVeCZRFyKRdzWc2FZgCFOiiNhkRNxUKWZYHiZddDaGhCOIw2wqUY7CeALV2Sxf9miSKEgVkAgvGmQixYg12akdYolQ5FJDjNA8XCQmCYQmxYh/mo2sHCTG7CZXVmiN7biCKpph2Ek9nlQUTZN1FZQlgZlmmZQnTZl1WpnnKEqcTMo92S06SZkrAZgf0oflSBkTwQHZCg7TwjQ1gaia5pInCap3FdqJkmqfHJqRmSqiEEq7nanobqRliBT9HfQdzaDeTartFArxWq2kUMAaiAGngJDZjpmRZpmuqbpAJJv/1bwA1cwR0cDSTcwjXcw0XcxFWcTYkqs2KEcDppx6mVn4UcmJqmG2RMftxAntZWpU1ppD7dgUod1Vkd1uWH1n2a18FM2DUru4ZE2Z1d2k3/BDw8AnjYElRU3uVl3ublERp4Hujp5SqRnumhXrsWrMEeLMImrMIuLMM2rMM+LMRGrMROLMVWrMVeLMZmrMZubEMggsd+LMiGrMiOLMmWrMmeLMqmrMquLMu2rMu+LMzGrMzOLM3WrM3eLM7mrM7uLM/2rM/+LNAGrdAOLdG+rDSIINImrdIuLdM2rdM+LdRGrdROLdVWrdUy7TBcrdZuLdd2rdd+LdiGrdiOLdmWrf4NwwWkrdquLdu2rdu+LdzGrdzOLd3Wrd3eLd7mrd7uLd/2rd/+LeAGruAOLuEWruEeLuImruIuLuM2ruM+rt6KJ8dOLuVWruVeLuZmruZu/y7ndq7nfi7ohq7oji7plq7pni7qpq7qri7rtq7rvi7sxq7szi7t1q7t3i7ucsYFWEMJNEMwXEMGbFnu5u4emI1+SIGtfsSSDO/lGhA0UMMgVEMloNcGlCZINCPzUi7XDBJC6NM1hgT2Zu/GDgLZCCTQ1ZFIhK/4ZuwDdJCpJsI1DENQHUSf/Ek3aqcIIEIiPEAB/IE2LMFTNYEI3OQahIp1AUInOAIZVMIx0Yqt4IquNOJBhMAuoAIdbIjw7g43Csr6TiP5dtr7KsTBJMzCIMQF0AABhNcBOILPbIA0ZZl+IAgaDEwHrVQVvFHN3EzO4MckHgQ3nKTOYCTRMGQJd//wMNIqPQxAyDFET3LlT1YSRR3AGTRQE9xkSCnCBdDVNVxAHu0SHTzAJqCBG4iRN5EW7ugO7/jOiCADVnXDjnkDwtwYCFmNTxqOEc8iMtCqTw0DDNirW+bXA5Wid+JHh3LQFahdL4aAJRzA/CYCN+hkC8nXX85QkDFW0x3EOfjMjv0xXAryHctiGYjBGqjVAQCC38XRHKHvCdMAPCQECzgNIqsdmCgEwmzSI6FbPUGTIFrBMNgQKoMm+n7yMCJDANTASWoDRhWncSLnINOA8DoDLJtmL9ZUNxhnJHiJ88QTLifCK8PCQyhzRzGzME/jJrzDxHiaSi2o9JzwDcxRNHv/WC+iV10NqSQLYn3BAw3kUkOkc11B5TjjISJcg3YlRO/clYVeaH+wszsDKDxnEoDE6OjU8zYLpxQ8xEE/lyv9sx02jZsmhMkQw0HwqEMoNGi+M/bGyUb600H+HBqQjeLBAxKAmEhrtCzu0hq8b8YEw0FwaUJ4KRS3c0kzNPaizGKlFAQUQbss2mnV02VJMHzp11T2NIfRNB6GwMbkQXYiwcRA5ZkJ6ucQ6iAD9ceYNE4C3fgw1SvpTaPaM2r9Bz1g0vtEkZkFqpqBNVXbocnQwBUsgRUgwi68DC8mBKqqKq+BD0mPNUPXnAVwAzltwaFKAAy0Fyz8qUT/XCJsglERlMPEiUHM6OJg61ph3zUeUkO65scfWMMxKYLH5QfImXBJBnUFBQ3LJGV+SEEjKrWPoVYiOAG/lFylKslqX+QSi7YdcoAETEACSEMkqKPBmB3aZUQZnEECWK8TUMPmIUEGVwQHbIHkZTTROHe8Erd4jzd5l7d5nzd6p7d6rzd7t7d7vzd8x7d8zzd917d93zd+5zeWBAQAOw==" style="margin-top:1rem;" /></div></details></div>
```

## Wolfram Alpha

::: {.sourceClojure}
```clojure
(wl/eval '(WolframAlpha "number of moons of Saturn" "Result"))
```
:::

::: {.printedClojure}
```clojure
(Quantity 145 (IndependentUnit "moons"))

```
:::

::: {.sourceClojure}
```clojure
(view '(WolframAlpha "number of moons of Saturn" "Result"))
```
:::

```{=html}
<div><details open="open"><summary><h5 style="cursor:pointer;display:inline;padding:0.5rem 1rem 0 1rem;">WolframAlpha["number of moons of Saturn", "Result"]</h5></summary><div class="wl-results"><hr /><img src="data:image/jpeg;base64,R0lGODlhxgBTAPeXAPHx8fb185F/a6iaivHv7O/v7/Hv7eLd2KKTgt/a1YVyXMW8scW8ssrBuPf29O7r6OPj4+Hd18rCuYh1X9/a1L+1qYt4Y7Gkluvo5Orn5LOnmaSWhtzX0d3Y0tXOx5iIdol3Yfn5+JKAbeDc1uHh4Z2OfO/s6se+tPv7+p6Offb19KWXh5eHdIFtV7muoqaYiNPMxLerntDIwINwWaudjpaGc5SDb7KmmNTNxYRwWrCklq6hksK4rcnAt4d0XoNvWbqvoqCRgIJuV+Hc187Gvd/f36OUhNnTzI9+adPNxcvDue3q58/Iv6yfkN7e3uXh3Ozs7Id0X4JuWMzEu72zp5GAbOTg26ibi6GSgpuLeZ+Qfsa9s+Db1aWXhuLi4tvb25+Qf/T09IRxW+Xl5ZWFcoBsVfDu68G3rPTz8cjAtubm5unm4vr6+eTk5NvVz9zc3I17ZuLe2ambjKqdjpmJd+rq6u/t6raqne7s6djSy/f39+Xh3aOVhIp3Ytra2tnZ2Y99abCjlfPx7/v7+7qvo7SomuDg4O7u7tbPyIVxW/X19cO6r/r6+vLy8rOnmpeGdPj4+Ih1YNfRyc/Hv/7+/rGll4l2Yf////b29ujo6N3d3ePe2dbQyefn5/z8/Ip4Y6qcjfDw8L60qPPz89DJwc7HvtfRytHKwvr5+PXz8paFcpiHdezp5sS7sMe/tdvWz+fk37uwpMC2q9TOxpyMesvDuo17Z+jk4JWEcfDu7IZzXYt5ZNHKwZOCbrWpnPz8+/39/fv6+sS7sc3FvK6ik6eZiZKBbZOCb/j39pSDcNLLw/39/Ofj3/Py8LmuodrVzpqKePf29Y58Z+bj3uzq5unl4d7Z07Spm6GSgZB/a7arnpmJdu3t7Y58aOvo5enp6d3Y0ejl4Z2NfMi/tdrUzdXPx39rVPTy8K2gkritoIx6Zfn4962gkb2ypoFtVqeZivLw7resn6+ilJB+aubi3ebi3ebi3ebi3ebi3ebi3ebi3ebi3ebi3ebi3ebi3ebi3SH/C05FVFNDQVBFMi4wAwEAAAAh+QQECgD/ACwAAAAAxgBTAAAI/wAvCRxIsKDBgwgTKlzIsKHDhxAjSpxIsaLFixgzatzIsaPHjyBDihxJsqTJkyhTqlzJsqXLlzBjypxJs6bNmzhz6tzJs6fPn0CDCh1KtKjRo0g3AmtUR82Yp1CjSp1KtarVq1izapWaiRukpBQHffPzp6zZs2jTql3Ltq3bt3DdGgoDFqKiN3Hz6t3Lt6/aTpTqMsRE9o+mUIwCC57oSVEns2MWJ/SE908nYJIxjqp8KLPBOmU7ec6oh+wXT6MFUvpiGHXqi1DKdk4dRvZrjKv/eHnNrezX2xfb/Pny+ltZ18ArPv6DeXSmss2TU3zOPDX16NIlXrcOPfv07s7Be/+PuD189fHkxWcuj/4h+/Xq2zN8L5m+fIX2Hf46daXYNImU3FFMMa8kVMmACCY4YIEL5XffQQ4iBIwVUzQRhTkYuiERFRiao0RCunQooogfNhjfgwhFeNASI2YY0RFldFiiQcDECIcROOaYIznznYiiQSoaBE8XRHZRhYsPLXHhHBjOWBAaGKZjUZA/UplQO0g2hIoq5iSzRpMIsYJhK1P6+CNBViKEpTkaOtSEOVLAAguYB22CIRNlnneJKLHg4YksuNgCSi6XUOLKI7vwUQ1CwQxTjDGfHDOHMsswxIwz0MAhDS3tUJPQFrFYIxAbw2CRzS7bxCKIQvAIY2Qf3WT/kQ44iqVoZkNrtsnQOBimcYmdHiL0TJYTsZeIOeW82eEqlBQioiXnGESOOi2aU8UmCa3DTowjunMHcgT1Yk48v87TYiT0IJSGGNVaS0StBaV5UK4NDSGEOegINASdBp2C4X8VGWuOOObQQkgyGGpjThBAmGuOKAVNwi00Z9QiCpfm/PDMQcg8gqEPlfTgyg45YIgNuAKJ2wQO99bgCxU0zIChMShfIgOGusQAAzh5jPNCC+a4s4etekpE70LwUMsCG/ryW1AaGHrDyw1drBBIGna4J96x5lwRmArsmhODQNTEmAJBe0hhjhC8EETJCTH6YIJBcmAIRjMEmeCx2AaJ/2uJO32YQhArn2CYh0HZwOmNkDoskpC8Bh09WQnmRDJ308EeJAuGP7RYxg4qNCSwD2gMFIQ52aAwUOLzEHS6OVMcJAqGxBQ0AoYihGBQMyCY04Kn4c5MaEGuYEguQchg+EGxtzIkOUIKl8EBQftmbtDs5nxSiStT8LABt6qEbqKexxpBECjmaEAQNObsMlAGGJKB0C+FuyO+QEya4wFCJ0RZkLhSwJtBxGSODRQEGCESAwa007yFPM8gysCQ46jnNLeNoxa/KEg1xGWOC/SIfOYw30BokD6CZMEcFhjIFjDEg4RcA0MwGAglLiSGSh3kHBjqxf/MkQhGYagEBlkhD/+dsTitFS0iDyRIONglB4NUz0kOMcOx3JGK8UWnfAQhofoGcsIUCuQGxCpIKTBUgYHAA0NZUEg3zFEGeImrhwhB40G20DkM4SIWFKgZkBqokCQKJBpHmoAbhkDIQkoAQy4g5C0g4ggMJcGKA8HiCEvIRRQO5AoYekJC8oAhXwwkHBjqgkI+gKEqDuSNCZHjQR5QCEuIyAc6SNfj+HilMA6kAe1qVxofQgQMMQCSApGkQLRoQksKhB0YikBCZoHIgbDIHFhQCMLMobtT8jCV5tilhLhACBaIaA7RIBp2kGhLgfQgl9XSZkNIgaEFAPMSwrwEMSvpxUs4A0OnSIgwfDn/kORZKyEINIcCdgjHg6iSISYQhQ9+CK+BQK4gfizUMiZK0YomAEMNmOg4F0IIGL4znvMUSBcHIgkMBSIhCMCQFQiyt6wd5AAYMmDwCmqQgzYkF4BIJoRoqaZyMuSJB0GFAbJFrRaYcpYgFOEwKSlSY15iHQkkwEGYESMkwGtz5rgDQgaAoVIQFJvqbMgZMNS2PR4RIhFdCFANIo9dgOMgKHgH7T54xRBmkamXGOlAsFqCag4kFdP0KkGiMQE2csIguDSHAGxoTZoWxKaXwAMdRIWQupkjDjs9a0KsIInOdhZ9D/OsJI7gkLXmjWsv8IApHaAMG2AIEKX7qF0nucWm/9bzEr+gBYaSkYcMsgEGSMDQOxp6CVPEqAXOuBwGbhCjH2Drq3HMZkEItjAPBIAg57gDhrZB3EvIy7LovGa9KiiQW3izQz+IArfMgYsHiG5rs11qbfPqVIGkQgsdcscE1nsFvxaEFDLDkALCZo4obKxv4jWodAmyhpRiqAzSYEEWBMCtbLg3sxtNCHjR6diEmLYglFAGFoAmImm0QnXvTepd56tXgixDCYkTERmU0d2BeIMGauuQGG4wPAR3eCCQFcg0KkGtEcFBFOtAaoZp4gBr8GIYHshAejRrEQwgggiSwMNDQtABUpyCCygOiTfcQIpSSMIbNXYoT4nyUPS0Gdkob/ZOnH0yZ+nUmSd3Bk6edbLn1/QZJ382z5LBApo/DOJMClFDWdKclEOURRGITggJDPMaSJTlG5E+yCDKoobbOKEsesh0QYTzh0bcBgBlKcKgH1SAshgiORAoiyYgHWlPLMcPmEiOJ4pgFghwAwCNCLawh03sYhv72MhOtrKXzWxih+IblflDKLIDDEX75drYzjZf3kCX8WBiDIXRtrjHne0icGPVtwGGHsLQ7Ha7+93wbvcoMMEIUdv73vjOt773ze9++/vfAA+4wAdO8IIb/OAIT7jCAx4QADs=" style="margin-top:1rem;" /></div></details></div>
```

```{=html}
<div style="background-color:grey;height:2px;width:100%;"></div>
```

```{=html}
<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/main/projects/math/wolframite/notebooks/index.clj">projects/math/wolframite/notebooks/index.clj</a></small></small></pre></div>
```