### 2024-02-24 docs/projects/datavis/ggplot/representing.qmd
ggplot - rerendered

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: representing.html}
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="representing_files/md-default0.js" type="text/javascript"></script><script src="representing_files/md-default1.js" type="text/javascript"></script>

```{=html}
<table><tr><td>This is part of the Scicloj <a href="https://scicloj.github.io/clojure-data-scrapbook/">Clojure Data Scrapbook</a>.</td><a href="https://scicloj.github.io/clojure-data-scrapbook/"><img align="left" alt="SciCloj logo" src="https://scicloj.github.io/sci-cloj-logo-transparent.png" width="40" /></a></tr></table>
```

# Representing ggplot plots as Clojure data

::: {.sourceClojure}
```clojure
(ns representing
  (:require [clojisr.v1.r :as r :refer [r r$ r->clj]]
            [clojisr.v1.applications.plotting :as plotting]
            [scicloj.kindly.v4.kind :as kind]
            [clojure.walk :as walk]
            [tablecloth.api :as tc]
            [editscript.core :as editscript]
            [clojure.pprint :as pp]
            [clojure.string :as str]))
```
:::

::: {.sourceClojure}
```clojure
(r/library "ggplot2")
```
:::

::: {.printedClojure}
```clojure
[1] "ggplot2"   "Rserve"    "stats"     "graphics"  "grDevices" "utils"    
[7] "datasets"  "methods"   "base"

```
:::

::: {.sourceClojure}
```clojure
(r/require-r '[base])
```
:::

::: {.printedClojure}
```clojure
nil

```
:::

## A representation function

A ggplot object is an R list of [ggproto](https://bookdown.dongzhuoer.com/hadley/ggplot2-book/introducing-ggproto) objects. We recursively unwrap this structure and convert it to Clojure.

::: {.sourceClojure}
```clojure
(defn ggplot->clj
  ([r-obj]
   (ggplot->clj r-obj {} []))
  ([r-obj
    {:as options
     :keys [avoid]
     :or {avoid #{"plot_env"}}}
    path]
   #_(prn path)
   (let [relevant-names (some->> r-obj
                                 r.base/names
                                 r->clj
                                 (filter (complement avoid)))]
     (cond
       ;;
       ;; a named list or a ggproto object
       (seq relevant-names) (->> relevant-names
                                 (map (fn [nam]
                                        [(keyword nam) (-> r-obj
                                                           (r$ nam)
                                                           (ggplot->clj options
                                                                        (conj path nam)))]))
                                 (into {}))
       ;;
       ;; a ggproto method
       (-> r-obj
           r.base/class
           r->clj
           first
           (= "ggproto_method"))
       :ggproto-method
       ;;
       ;; an unnamed list
       (-> r-obj
           r.base/is-list
           r->clj
           first)
       (-> r-obj
           r.base/length
           r->clj
           first
           range
           (->> (mapv (fn [i]
                        (-> r-obj
                            (r/brabra (inc i))
                            (ggplot->clj options
                                         (conj path [i])))))))
       ;;
       (r.base/is-atomic r-obj) (try (r->clj r-obj)
                                     (catch Exception e
                                       (-> r-obj println with-out-str)))
       :else r-obj))))
```
:::

In the conversion, we avoid some big parts of the structure, e.g., the `"plot_env"` member. We also do no report the toplevel `"data"` member, which is simply the dataset.

## An example

For example:

::: {.sourceClojure}
```clojure
(delay
  (-> "(ggplot(mpg, aes(cty, hwy))
         + geom_point())"
      r
      ggplot->clj
      (dissoc :data)))
```
:::

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [geom_point],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"]}}

```
:::

```{=html}
<div style="background-color:grey;height:2px;width:100%;"></div>
```

```{=html}
<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/main/projects/datavis/ggplot/notebooks/representing.clj">projects/datavis/ggplot/notebooks/representing.clj</a></small></small></pre></div>
```