### 2024-02-28 docs/projects/datavis/ggplot/study_session_20240225.qmd

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: study_session_20240225.html}
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="study_session_20240225_files/md-default3.js" type="text/javascript"></script><script src="study_session_20240225_files/md-default4.js" type="text/javascript"></script>

```{=html}
<table><tr><td>This is part of the Scicloj <a href="https://scicloj.github.io/clojure-data-scrapbook/">Clojure Data Scrapbook</a>.</td><a href="https://scicloj.github.io/clojure-data-scrapbook/"><img align="left" alt="SciCloj logo" src="https://scicloj.github.io/sci-cloj-logo-transparent.png" width="40" /></a></tr></table>
```

::: {.sourceClojure}
```clojure
(ns study-session-20240225
  (:require [clojisr.v1.r :as r :refer [r]]
            [clojisr.v1.applications.plotting :as plotting]
            [scicloj.kindly.v4.kind :as kind]
            [tablecloth.api :as tc]
            [scicloj.metamorph.ml.toydata :as toydata]
            [scicloj.metamorph.ml.toydata.ggplot :as toydata.ggplot]
            [representing]))
```
:::

# Code examples from study session 2024-02-25

Add ggplot2 to the R runtime:

::: {.sourceClojure}
```clojure
(r/library "ggplot2")
```
:::

::: {.printedClojure}
```clojure
 [1] "ggtrace"   "ggplot2"   "Rserve"    "stats"     "graphics"  "grDevices"
 [7] "utils"     "datasets"  "methods"   "base"

```
:::

The `mpg` dataset:

::: {.sourceClojure}
```clojure
(r "mpg")
```
:::

::: {.printedClojure}
```clojure
# A tibble: 234 × 11
   manufacturer model      displ  year   cyl trans drv     cty   hwy fl    class
   &lt;chr&gt;        &lt;chr&gt;      &lt;dbl&gt; &lt;int&gt; &lt;int&gt; &lt;chr&gt; &lt;chr&gt; &lt;int&gt; &lt;int&gt; &lt;chr&gt; &lt;chr&gt;
 1 audi         a4           1.8  1999     4 auto… f        18    29 p     comp…
 2 audi         a4           1.8  1999     4 manu… f        21    29 p     comp…
 3 audi         a4           2    2008     4 manu… f        20    31 p     comp…
 4 audi         a4           2    2008     4 auto… f        21    30 p     comp…
 5 audi         a4           2.8  1999     6 auto… f        16    26 p     comp…
 6 audi         a4           2.8  1999     6 manu… f        18    26 p     comp…
 7 audi         a4           3.1  2008     6 auto… f        18    27 p     comp…
 8 audi         a4 quattro   1.8  1999     4 manu… 4        18    26 p     comp…
 9 audi         a4 quattro   1.8  1999     4 auto… 4        16    25 p     comp…
10 audi         a4 quattro   2    2008     4 manu… 4        20    28 p     comp…
# ℹ 224 more rows

```
:::

Evaluating R and showing a plot as a bitmap image:

::: {.sourceClojure}
```clojure
(plotting/plot->buffered-image
 (r "
(ggplot(mpg, aes(x = cty, y = hwy, color = year)) +
geom_point())
"))
```
:::

```{=html}
<img src="study_session_20240225_files/0.png" style="width:100%;" />
```

Showing a plot as svg:

::: {.sourceClojure}
```clojure
(plotting/plot->svg
 (r "
(ggplot(mpg, aes(x = cty, y = hwy, color = year)) +
geom_point())
"))
```
:::

::: {.printedClojure}
```clojure
"&lt;?xml version=\"1.0\" encoding=\"UTF-8\"?&gt;\n&lt;svg xmlns=\"http://www.w3.org/2000/svg\" xmlns:xlink=\"http://www.w3.org/1999/xlink\" width=\"504pt\" height=\"504pt\" viewBox=\"0 0 504 504\" version=\"1.1\"&gt;\n&lt;defs&gt;\n&lt;g&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-0\"&gt;\n&lt;path style=\"stroke:none;\" d=\"\"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-1\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 4.453125 -0.765625 L 1.171875 -0.765625 C 1.25 -1.296875 1.53125 -1.625 2.296875 -2.09375 L 3.171875 -2.59375 C 4.046875 -3.0625 4.5 -3.71875 4.5 -4.5 C 4.5 -5.03125 4.28125 -5.53125 3.921875 -5.875 C 3.546875 -6.203125 3.09375 -6.359375 2.5 -6.359375 C 1.703125 -6.359375 1.125 -6.078125 0.78125 -5.53125 C 0.546875 -5.203125 0.453125 -4.796875 0.4375 -4.15625 L 1.21875 -4.15625 C 1.234375 -4.59375 1.296875 -4.84375 1.40625 -5.0625 C 1.59375 -5.4375 2 -5.6875 2.46875 -5.6875 C 3.171875 -5.6875 3.703125 -5.171875 3.703125 -4.484375 C 3.703125 -3.96875 3.40625 -3.53125 2.859375 -3.21875 L 2.046875 -2.75 C 0.75 -2 0.375 -1.40625 0.296875 -0.015625 L 4.453125 -0.015625 Z M 4.453125 -0.765625 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-2\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 2.421875 -6.359375 C 1.84375 -6.359375 1.3125 -6.09375 0.984375 -5.671875 C 0.578125 -5.109375 0.375 -4.25 0.375 -3.078125 C 0.375 -0.9375 1.078125 0.203125 2.421875 0.203125 C 3.734375 0.203125 4.453125 -0.9375 4.453125 -3.03125 C 4.453125 -4.25 4.265625 -5.078125 3.859375 -5.671875 C 3.53125 -6.109375 3.015625 -6.359375 2.421875 -6.359375 Z M 2.421875 -5.671875 C 3.25 -5.671875 3.671875 -4.828125 3.671875 -3.09375 C 3.671875 -1.28125 3.265625 -0.4375 2.40625 -0.4375 C 1.578125 -0.4375 1.171875 -1.3125 1.171875 -3.0625 C 1.171875 -4.828125 1.578125 -5.671875 2.421875 -5.671875 Z M 2.421875 -5.671875 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-3\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 1.9375 -2.921875 L 2.359375 -2.921875 C 3.203125 -2.921875 3.65625 -2.53125 3.65625 -1.765625 C 3.65625 -0.96875 3.171875 -0.484375 2.375 -0.484375 C 1.515625 -0.484375 1.109375 -0.921875 1.0625 -1.84375 L 0.28125 -1.84375 C 0.3125 -1.34375 0.40625 -1 0.546875 -0.71875 C 0.875 -0.109375 1.484375 0.203125 2.34375 0.203125 C 3.625 0.203125 4.453125 -0.578125 4.453125 -1.78125 C 4.453125 -2.59375 4.140625 -3.03125 3.390625 -3.296875 C 3.96875 -3.53125 4.265625 -3.96875 4.265625 -4.609375 C 4.265625 -5.703125 3.546875 -6.359375 2.359375 -6.359375 C 1.109375 -6.359375 0.4375 -5.65625 0.40625 -4.3125 L 1.1875 -4.3125 C 1.203125 -4.703125 1.234375 -4.921875 1.328125 -5.109375 C 1.5 -5.46875 1.890625 -5.6875 2.375 -5.6875 C 3.0625 -5.6875 3.46875 -5.265625 3.46875 -4.578125 C 3.46875 -4.140625 3.3125 -3.859375 2.96875 -3.71875 C 2.765625 -3.625 2.484375 -3.59375 1.9375 -3.578125 Z M 1.9375 -2.921875 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-4\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 2.875 -1.546875 L 2.875 0 L 3.65625 0 L 3.65625 -1.546875 L 4.578125 -1.546875 L 4.578125 -2.234375 L 3.65625 -2.234375 L 3.65625 -6.359375 L 3.078125 -6.359375 L 0.25 -2.359375 L 0.25 -1.546875 Z M 2.875 -2.234375 L 0.921875 -2.234375 L 2.875 -5.046875 Z M 2.875 -2.234375 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-5\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 2.28125 -4.53125 L 2.28125 0 L 3.046875 0 L 3.046875 -6.359375 L 2.546875 -6.359375 C 2.265625 -5.390625 2.09375 -5.25 0.890625 -5.09375 L 0.890625 -4.53125 Z M 2.28125 -4.53125 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-6\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 4.1875 -6.234375 L 0.96875 -6.234375 L 0.5 -2.84375 L 1.21875 -2.84375 C 1.578125 -3.265625 1.875 -3.421875 2.359375 -3.421875 C 3.203125 -3.421875 3.71875 -2.84375 3.71875 -1.921875 C 3.71875 -1.03125 3.203125 -0.484375 2.359375 -0.484375 C 1.6875 -0.484375 1.265625 -0.828125 1.078125 -1.53125 L 0.3125 -1.53125 C 0.40625 -1.015625 0.5 -0.78125 0.6875 -0.546875 C 1.03125 -0.078125 1.671875 0.203125 2.375 0.203125 C 3.640625 0.203125 4.515625 -0.71875 4.515625 -2.03125 C 4.515625 -3.265625 3.6875 -4.109375 2.5 -4.109375 C 2.0625 -4.109375 1.703125 -4 1.34375 -3.734375 L 1.59375 -5.46875 L 4.1875 -5.46875 Z M 4.1875 -6.234375 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-7\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 4.375 -4.703125 C 4.234375 -5.75 3.5625 -6.359375 2.609375 -6.359375 C 1.921875 -6.359375 1.3125 -6.03125 0.9375 -5.453125 C 0.546875 -4.828125 0.375 -4.0625 0.375 -2.90625 C 0.375 -1.828125 0.53125 -1.15625 0.90625 -0.59375 C 1.234375 -0.078125 1.78125 0.203125 2.46875 0.203125 C 3.65625 0.203125 4.515625 -0.6875 4.515625 -1.921875 C 4.515625 -3.09375 3.71875 -3.921875 2.609375 -3.921875 C 1.984375 -3.921875 1.5 -3.6875 1.171875 -3.234375 C 1.171875 -4.796875 1.671875 -5.671875 2.5625 -5.671875 C 3.109375 -5.671875 3.484375 -5.328125 3.609375 -4.703125 Z M 2.5 -3.234375 C 3.25 -3.234375 3.71875 -2.703125 3.71875 -1.859375 C 3.71875 -1.0625 3.1875 -0.484375 2.484375 -0.484375 C 1.765625 -0.484375 1.21875 -1.09375 1.21875 -1.90625 C 1.21875 -2.6875 1.734375 -3.234375 2.5 -3.234375 Z M 2.5 -3.234375 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-8\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 3.4375 -3.359375 C 4.09375 -3.75 4.296875 -4.0625 4.296875 -4.65625 C 4.296875 -5.671875 3.53125 -6.359375 2.421875 -6.359375 C 1.3125 -6.359375 0.546875 -5.671875 0.546875 -4.671875 C 0.546875 -4.0625 0.75 -3.75 1.390625 -3.359375 C 0.671875 -2.984375 0.328125 -2.46875 0.328125 -1.765625 C 0.328125 -0.609375 1.171875 0.203125 2.421875 0.203125 C 3.65625 0.203125 4.515625 -0.609375 4.515625 -1.765625 C 4.515625 -2.46875 4.15625 -2.984375 3.4375 -3.359375 Z M 2.421875 -5.671875 C 3.078125 -5.671875 3.5 -5.28125 3.5 -4.640625 C 3.5 -4.046875 3.0625 -3.65625 2.421875 -3.65625 C 1.765625 -3.65625 1.34375 -4.046875 1.34375 -4.65625 C 1.34375 -5.28125 1.765625 -5.671875 2.421875 -5.671875 Z M 2.421875 -3 C 3.1875 -3 3.71875 -2.484375 3.71875 -1.75 C 3.71875 -0.984375 3.203125 -0.484375 2.40625 -0.484375 C 1.640625 -0.484375 1.125 -1 1.125 -1.734375 C 1.125 -2.5 1.640625 -3 2.421875 -3 Z M 2.421875 -3 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-0\"&gt;\n&lt;path style=\"stroke:none;\" d=\"\"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-1\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 5.1875 -3.828125 C 5.140625 -4.390625 5.015625 -4.75 4.796875 -5.078125 C 4.40625 -5.609375 3.703125 -5.921875 2.90625 -5.921875 C 1.34375 -5.921875 0.34375 -4.703125 0.34375 -2.78125 C 0.34375 -0.921875 1.328125 0.25 2.890625 0.25 C 4.265625 0.25 5.140625 -0.578125 5.25 -1.984375 L 4.328125 -1.984375 C 4.171875 -1.0625 3.703125 -0.59375 2.921875 -0.59375 C 1.90625 -0.59375 1.296875 -1.421875 1.296875 -2.78125 C 1.296875 -4.21875 1.890625 -5.078125 2.890625 -5.078125 C 3.65625 -5.078125 4.140625 -4.625 4.25 -3.828125 Z M 5.1875 -3.828125 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-2\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 2.796875 -5.765625 L 1.84375 -5.765625 L 1.84375 -7.34375 L 0.9375 -7.34375 L 0.9375 -5.765625 L 0.15625 -5.765625 L 0.15625 -5.015625 L 0.9375 -5.015625 L 0.9375 -0.65625 C 0.9375 -0.0625 1.328125 0.25 2.046875 0.25 C 2.28125 0.25 2.484375 0.234375 2.796875 0.171875 L 2.796875 -0.59375 C 2.65625 -0.5625 2.546875 -0.546875 2.359375 -0.546875 C 1.953125 -0.546875 1.84375 -0.65625 1.84375 -1.0625 L 1.84375 -5.015625 L 2.796875 -5.015625 Z M 2.796875 -5.765625 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-3\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 4.265625 -5.765625 L 2.671875 -1.28125 L 1.203125 -5.765625 L 0.21875 -5.765625 L 2.171875 0.015625 L 1.8125 0.9375 C 1.65625 1.34375 1.46875 1.5 1.078125 1.5 C 0.921875 1.5 0.796875 1.46875 0.59375 1.4375 L 0.59375 2.25 C 0.78125 2.359375 0.96875 2.390625 1.203125 2.390625 C 1.5 2.390625 1.828125 2.296875 2.0625 2.125 C 2.359375 1.90625 2.515625 1.671875 2.6875 1.203125 L 5.265625 -5.765625 Z M 4.265625 -5.765625 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-4\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 5.640625 -2.578125 C 5.640625 -3.453125 5.578125 -3.984375 5.40625 -4.40625 C 5.03125 -5.359375 4.15625 -5.921875 3.078125 -5.921875 C 1.46875 -5.921875 0.4375 -4.703125 0.4375 -2.8125 C 0.4375 -0.90625 1.4375 0.25 3.0625 0.25 C 4.375 0.25 5.296875 -0.5 5.515625 -1.75 L 4.59375 -1.75 C 4.34375 -0.984375 3.828125 -0.59375 3.09375 -0.59375 C 2.515625 -0.59375 2.015625 -0.859375 1.703125 -1.34375 C 1.484375 -1.671875 1.40625 -2 1.390625 -2.578125 Z M 1.421875 -3.328125 C 1.5 -4.390625 2.140625 -5.078125 3.0625 -5.078125 C 4 -5.078125 4.65625 -4.359375 4.65625 -3.328125 Z M 1.421875 -3.328125 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-5\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 5.890625 -0.53125 C 5.78125 -0.515625 5.734375 -0.515625 5.6875 -0.515625 C 5.375 -0.515625 5.1875 -0.6875 5.1875 -0.96875 L 5.1875 -4.359375 C 5.1875 -5.375 4.4375 -5.921875 3.03125 -5.921875 C 2.171875 -5.921875 1.5 -5.6875 1.109375 -5.265625 C 0.84375 -4.96875 0.734375 -4.625 0.71875 -4.0625 L 1.640625 -4.0625 C 1.71875 -4.765625 2.140625 -5.078125 2.984375 -5.078125 C 3.828125 -5.078125 4.28125 -4.78125 4.28125 -4.21875 L 4.28125 -3.984375 C 4.265625 -3.59375 4.0625 -3.4375 3.328125 -3.34375 C 2.03125 -3.171875 1.828125 -3.140625 1.46875 -2.984375 C 0.796875 -2.703125 0.46875 -2.203125 0.46875 -1.453125 C 0.46875 -0.40625 1.1875 0.25 2.359375 0.25 C 3.078125 0.25 3.65625 0 4.3125 -0.59375 C 4.375 0 4.65625 0.25 5.265625 0.25 C 5.453125 0.25 5.578125 0.234375 5.890625 0.15625 Z M 4.28125 -1.8125 C 4.28125 -1.5 4.1875 -1.3125 3.921875 -1.0625 C 3.546875 -0.71875 3.09375 -0.546875 2.546875 -0.546875 C 1.84375 -0.546875 1.421875 -0.890625 1.421875 -1.46875 C 1.421875 -2.078125 1.8125 -2.390625 2.8125 -2.53125 C 3.78125 -2.65625 3.96875 -2.703125 4.28125 -2.84375 Z M 4.28125 -1.8125 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-6\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 0.765625 -5.765625 L 0.765625 0 L 1.6875 0 L 1.6875 -2.984375 C 1.6875 -4.375 2.265625 -5 3.53125 -4.96875 L 3.53125 -5.890625 C 3.375 -5.921875 3.28125 -5.921875 3.171875 -5.921875 C 2.578125 -5.921875 2.140625 -5.578125 1.609375 -4.71875 L 1.609375 -5.765625 Z M 0.765625 -5.765625 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph2-0\"&gt;\n&lt;path style=\"stroke:none;\" d=\"\"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph2-1\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M -8.015625 -0.765625 L 0 -0.765625 L 0 -1.6875 L -3.171875 -1.6875 C -4.359375 -1.6875 -5.125 -2.296875 -5.125 -3.25 C -5.125 -3.546875 -5.03125 -3.84375 -4.875 -4.0625 C -4.671875 -4.328125 -4.40625 -4.4375 -4 -4.4375 L 0 -4.4375 L 0 -5.34375 L -4.359375 -5.34375 C -5.328125 -5.34375 -5.921875 -4.65625 -5.921875 -3.53125 C -5.921875 -2.71875 -5.671875 -2.21875 -4.96875 -1.6875 L -8.015625 -1.6875 Z M -8.015625 -0.765625 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph2-2\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 0 -6.09375 L -5.765625 -7.78125 L -5.765625 -6.75 L -1.28125 -5.609375 L -5.765625 -4.484375 L -5.765625 -3.359375 L -1.28125 -2.25 L -5.765625 -1.078125 L -5.765625 -0.0625 L 0 -1.734375 L 0 -2.765625 L -4.515625 -3.890625 L 0 -5.046875 Z M 0 -6.09375 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph2-3\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M -5.765625 -4.265625 L -1.28125 -2.671875 L -5.765625 -1.203125 L -5.765625 -0.21875 L 0.015625 -2.171875 L 0.9375 -1.8125 C 1.34375 -1.65625 1.5 -1.46875 1.5 -1.078125 C 1.5 -0.921875 1.46875 -0.796875 1.4375 -0.59375 L 2.25 -0.59375 C 2.359375 -0.78125 2.390625 -0.96875 2.390625 -1.203125 C 2.390625 -1.5 2.296875 -1.828125 2.125 -2.0625 C 1.90625 -2.359375 1.671875 -2.515625 1.203125 -2.6875 L -5.765625 -5.265625 Z M -5.765625 -4.265625 \"/&gt;\n&lt;/symbol&gt;\n&lt;/g&gt;\n&lt;clipPath id=\"clip1\"&gt;\n  &lt;path d=\"M 35.152344 5.480469 L 433.84375 5.480469 L 433.84375 469.851562 L 35.152344 469.851562 Z M 35.152344 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip2\"&gt;\n  &lt;path d=\"M 35.152344 408 L 433.84375 408 L 433.84375 410 L 35.152344 410 Z M 35.152344 408 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip3\"&gt;\n  &lt;path d=\"M 35.152344 276 L 433.84375 276 L 433.84375 278 L 35.152344 278 Z M 35.152344 276 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip4\"&gt;\n  &lt;path d=\"M 35.152344 145 L 433.84375 145 L 433.84375 146 L 35.152344 146 Z M 35.152344 145 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip5\"&gt;\n  &lt;path d=\"M 35.152344 13 L 433.84375 13 L 433.84375 14 L 35.152344 14 Z M 35.152344 13 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip6\"&gt;\n  &lt;path d=\"M 101 5.480469 L 103 5.480469 L 103 469.851562 L 101 469.851562 Z M 101 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip7\"&gt;\n  &lt;path d=\"M 171 5.480469 L 173 5.480469 L 173 469.851562 L 171 469.851562 Z M 171 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip8\"&gt;\n  &lt;path d=\"M 241 5.480469 L 242 5.480469 L 242 469.851562 L 241 469.851562 Z M 241 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip9\"&gt;\n  &lt;path d=\"M 310 5.480469 L 312 5.480469 L 312 469.851562 L 310 469.851562 Z M 310 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip10\"&gt;\n  &lt;path d=\"M 380 5.480469 L 382 5.480469 L 382 469.851562 L 380 469.851562 Z M 380 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip11\"&gt;\n  &lt;path d=\"M 35.152344 342 L 433.84375 342 L 433.84375 344 L 35.152344 344 Z M 35.152344 342 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip12\"&gt;\n  &lt;path d=\"M 35.152344 210 L 433.84375 210 L 433.84375 212 L 35.152344 212 Z M 35.152344 210 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip13\"&gt;\n  &lt;path d=\"M 35.152344 78 L 433.84375 78 L 433.84375 80 L 35.152344 80 Z M 35.152344 78 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip14\"&gt;\n  &lt;path d=\"M 66 5.480469 L 68 5.480469 L 68 469.851562 L 66 469.851562 Z M 66 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip15\"&gt;\n  &lt;path d=\"M 136 5.480469 L 138 5.480469 L 138 469.851562 L 136 469.851562 Z M 136 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip16\"&gt;\n  &lt;path d=\"M 206 5.480469 L 208 5.480469 L 208 469.851562 L 206 469.851562 Z M 206 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip17\"&gt;\n  &lt;path d=\"M 275 5.480469 L 277 5.480469 L 277 469.851562 L 275 469.851562 Z M 275 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip18\"&gt;\n  &lt;path d=\"M 345 5.480469 L 347 5.480469 L 347 469.851562 L 345 469.851562 Z M 345 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip19\"&gt;\n  &lt;path d=\"M 415 5.480469 L 417 5.480469 L 417 469.851562 L 415 469.851562 Z M 415 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip20\"&gt;\n  &lt;path d=\"M 450.28125 203.203125 L 467.5625 203.203125 L 467.5625 289.601562 L 450.28125 289.601562 Z M 450.28125 203.203125 \"/&gt;\n&lt;/clipPath&gt;\n&lt;image id=\"image249\" width=\"1\" height=\"300\" xlink:href=\"data:image/png;base64,iVBORw0KGgoAAAANSUhEUgAAAAEAAAEsCAIAAACNMk6CAAAABmJLR0QA/wD/AP+gvaeTAAAAjklEQVQ4jaWTQRKAMAgDd3ik//R1HuOlji3SjNYjhZAEKNt+BCgQIRRIAQRSi2l5hdr79/q1fqBQVw+ErvyipqGH5eh7fNFJi1M+9bC8OT/lrTD9fNYwT/0FpvQzwVju0eObndsd/uUa7qX2NffqNE5w7j7dbCudpXbj2ehX0n3f0ot5pX9/1xkPj79W8p41NH/nSHk1sQAAAABJRU5ErkJggg==\"/&gt;\n&lt;pattern id=\"pattern0\" patternUnits=\"userSpaceOnUse\" width=\"1\" height=\"300\"  patternTransform=\"matrix(17.28,0,0,0.288,450.281644,203.20411)\"&gt;\n  &lt;use xlink:href=\"#image249\"/&gt;\n&lt;/pattern&gt;\n&lt;/defs&gt;\n&lt;g id=\"surface245\"&gt;\n&lt;rect x=\"0\" y=\"0\" width=\"504\" height=\"504\" style=\"fill:rgb(100%,100%,100%);fill-opacity:1;stroke:none;\"/&gt;\n&lt;rect x=\"0\" y=\"0\" width=\"504\" height=\"504\" style=\"fill:rgb(100%,100%,100%);fill-opacity:1;stroke:none;\"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 0 504 L 504 504 L 504 0 L 0 0 Z M 0 504 \"/&gt;\n&lt;g clip-path=\"url(#clip1)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\" stroke:none;fill-rule:nonzero;fill:rgb(92.156863%,92.156863%,92.156863%);fill-opacity:1;\" d=\"M 35.152344 469.847656 L 433.84375 469.847656 L 433.84375 5.476562 L 35.152344 5.476562 Z M 35.152344 469.847656 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip2)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 409.164062 L 433.84375 409.164062 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip3)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 277.242188 L 433.84375 277.242188 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip4)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 145.316406 L 433.84375 145.316406 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip5)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 13.394531 L 433.84375 13.394531 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip6)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 102.0625 469.847656 L 102.0625 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip7)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 171.765625 469.847656 L 171.765625 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip8)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 241.46875 469.847656 L 241.46875 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip9)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 311.167969 469.847656 L 311.167969 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip10)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 380.871094 469.847656 L 380.871094 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip11)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 343.203125 L 433.84375 343.203125 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip12)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 211.28125 L 433.84375 211.28125 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip13)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 79.355469 L 433.84375 79.355469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip14)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 67.214844 469.847656 L 67.214844 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip15)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 136.914062 469.847656 L 136.914062 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip16)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 206.617188 469.847656 L 206.617188 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip17)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 276.316406 469.847656 L 276.316406 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip18)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 346.019531 469.847656 L 346.019531 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip19)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 415.722656 469.847656 L 415.722656 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 224.472656 C 180.691406 227.078125 176.78125 227.078125 176.78125 224.472656 C 176.78125 221.867188 180.691406 221.867188 180.691406 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 198.085938 C 208.570312 200.695312 204.660156 200.695312 204.660156 198.085938 C 204.660156 195.480469 208.570312 195.480469 208.570312 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 211.28125 C 222.511719 213.886719 218.601562 213.886719 218.601562 211.28125 C 218.601562 208.671875 222.511719 208.671875 222.511719 211.28125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 264.050781 C 152.808594 266.65625 148.898438 266.65625 148.898438 264.050781 C 148.898438 261.441406 152.808594 261.441406 152.808594 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 250.855469 C 180.691406 253.464844 176.78125 253.464844 176.78125 250.855469 C 176.78125 248.25 180.691406 248.25 180.691406 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 277.242188 C 152.808594 279.847656 148.898438 279.847656 148.898438 277.242188 C 148.898438 274.636719 152.808594 274.636719 152.808594 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 237.664062 C 208.570312 240.269531 204.660156 240.269531 204.660156 237.664062 C 204.660156 235.058594 208.570312 235.058594 208.570312 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 250.855469 C 194.632812 253.464844 190.722656 253.464844 190.722656 250.855469 C 190.722656 248.25 194.632812 248.25 194.632812 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 277.242188 C 138.871094 279.847656 134.960938 279.847656 134.960938 277.242188 C 134.960938 274.636719 138.871094 274.636719 138.871094 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 277.242188 C 166.75 279.847656 162.839844 279.847656 162.839844 277.242188 C 162.839844 274.636719 166.75 274.636719 166.75 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 277.242188 C 166.75 279.847656 162.839844 279.847656 162.839844 277.242188 C 162.839844 274.636719 166.75 274.636719 166.75 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 277.242188 C 138.871094 279.847656 134.960938 279.847656 134.960938 277.242188 C 134.960938 274.636719 138.871094 274.636719 138.871094 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 290.433594 C 138.871094 293.039062 134.960938 293.039062 134.960938 290.433594 C 134.960938 287.828125 138.871094 287.828125 138.871094 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 277.242188 C 166.75 279.847656 162.839844 279.847656 162.839844 277.242188 C 162.839844 274.636719 166.75 274.636719 166.75 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 303.625 C 152.808594 306.230469 148.898438 306.230469 148.898438 303.625 C 148.898438 301.019531 152.808594 301.019531 152.808594 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 343.203125 C 124.929688 345.808594 121.019531 345.808594 121.019531 343.203125 C 121.019531 340.597656 124.929688 340.597656 124.929688 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 343.203125 C 124.929688 345.808594 121.019531 345.808594 121.019531 343.203125 C 121.019531 340.597656 124.929688 340.597656 124.929688 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 382.78125 C 97.046875 385.386719 93.140625 385.386719 93.140625 382.78125 C 93.140625 380.171875 97.046875 380.171875 97.046875 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 264.050781 C 152.808594 266.65625 148.898438 266.65625 148.898438 264.050781 C 148.898438 261.441406 152.808594 261.441406 152.808594 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 303.625 C 138.871094 306.230469 134.960938 306.230469 134.960938 303.625 C 134.960938 301.019531 138.871094 301.019531 138.871094 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 264.050781 C 152.808594 266.65625 148.898438 266.65625 148.898438 264.050781 C 148.898438 261.441406 152.808594 261.441406 152.808594 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 277.242188 C 138.871094 279.847656 134.960938 279.847656 134.960938 277.242188 C 134.960938 274.636719 138.871094 274.636719 138.871094 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 290.433594 C 138.871094 293.039062 134.960938 293.039062 134.960938 290.433594 C 134.960938 287.828125 138.871094 287.828125 138.871094 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 356.394531 C 124.929688 359 121.019531 359 121.019531 356.394531 C 121.019531 353.789062 124.929688 353.789062 124.929688 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 422.355469 C 83.109375 424.964844 79.199219 424.964844 79.199219 422.355469 C 79.199219 419.75 83.109375 419.75 83.109375 422.355469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 250.855469 C 194.632812 253.464844 190.722656 253.464844 190.722656 250.855469 C 190.722656 248.25 194.632812 248.25 194.632812 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 236.453125 211.28125 C 236.453125 213.886719 232.542969 213.886719 232.542969 211.28125 C 232.542969 208.671875 236.453125 208.671875 236.453125 211.28125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 224.472656 C 180.691406 227.078125 176.78125 227.078125 176.78125 224.472656 C 176.78125 221.867188 180.691406 221.867188 180.691406 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 264.050781 C 166.75 266.65625 162.839844 266.65625 162.839844 264.050781 C 162.839844 261.441406 166.75 261.441406 166.75 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 290.433594 C 180.691406 293.039062 176.78125 293.039062 176.78125 290.433594 C 176.78125 287.828125 180.691406 287.828125 180.691406 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 316.816406 C 152.808594 319.425781 148.898438 319.425781 148.898438 316.816406 C 148.898438 314.210938 152.808594 314.210938 152.808594 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 316.816406 C 152.808594 319.425781 148.898438 319.425781 148.898438 316.816406 C 148.898438 314.210938 152.808594 314.210938 152.808594 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 382.78125 C 83.109375 385.386719 79.199219 385.386719 79.199219 382.78125 C 79.199219 380.171875 83.109375 380.171875 83.109375 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 316.816406 C 138.871094 319.425781 134.960938 319.425781 134.960938 316.816406 C 134.960938 314.210938 138.871094 314.210938 138.871094 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 330.011719 C 138.871094 332.617188 134.960938 332.617188 134.960938 330.011719 C 134.960938 327.402344 138.871094 327.402344 138.871094 330.011719 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 303.625 C 152.808594 306.230469 148.898438 306.230469 148.898438 303.625 C 148.898438 301.019531 152.808594 301.019531 152.808594 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 303.625 C 152.808594 306.230469 148.898438 306.230469 148.898438 303.625 C 148.898438 301.019531 152.808594 301.019531 152.808594 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 356.394531 C 138.871094 359 134.960938 359 134.960938 356.394531 C 134.960938 353.789062 138.871094 353.789062 138.871094 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 369.585938 C 124.929688 372.195312 121.019531 372.195312 121.019531 369.585938 C 121.019531 366.980469 124.929688 366.980469 124.929688 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 356.394531 C 124.929688 359 121.019531 359 121.019531 356.394531 C 121.019531 353.789062 124.929688 353.789062 124.929688 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 356.394531 C 124.929688 359 121.019531 359 121.019531 356.394531 C 121.019531 353.789062 124.929688 353.789062 124.929688 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 55.226562 448.742188 C 55.226562 451.347656 51.316406 451.347656 51.316406 448.742188 C 51.316406 446.136719 55.226562 446.136719 55.226562 448.742188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 382.78125 C 83.109375 385.386719 79.199219 385.386719 79.199219 382.78125 C 79.199219 380.171875 83.109375 380.171875 83.109375 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 55.226562 448.742188 C 55.226562 451.347656 51.316406 451.347656 51.316406 448.742188 C 51.316406 446.136719 55.226562 446.136719 55.226562 448.742188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 395.972656 C 83.109375 398.578125 79.199219 398.578125 79.199219 395.972656 C 79.199219 393.367188 83.109375 393.367188 83.109375 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 369.585938 C 110.988281 372.195312 107.078125 372.195312 107.078125 369.585938 C 107.078125 366.980469 110.988281 366.980469 110.988281 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 395.972656 C 97.046875 398.578125 93.140625 398.578125 93.140625 395.972656 C 93.140625 393.367188 97.046875 393.367188 97.046875 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 55.226562 448.742188 C 55.226562 451.347656 51.316406 451.347656 51.316406 448.742188 C 51.316406 446.136719 55.226562 446.136719 55.226562 448.742188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 395.972656 C 97.046875 398.578125 93.140625 398.578125 93.140625 395.972656 C 93.140625 393.367188 97.046875 393.367188 97.046875 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 55.226562 448.742188 C 55.226562 451.347656 51.316406 451.347656 51.316406 448.742188 C 51.316406 446.136719 55.226562 446.136719 55.226562 448.742188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 395.972656 C 83.109375 398.578125 79.199219 398.578125 79.199219 395.972656 C 79.199219 393.367188 83.109375 393.367188 83.109375 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 382.78125 C 83.109375 385.386719 79.199219 385.386719 79.199219 382.78125 C 79.199219 380.171875 83.109375 380.171875 83.109375 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 382.78125 C 83.109375 385.386719 79.199219 385.386719 79.199219 382.78125 C 79.199219 380.171875 83.109375 380.171875 83.109375 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 369.585938 C 97.046875 372.195312 93.140625 372.195312 93.140625 369.585938 C 93.140625 366.980469 97.046875 366.980469 97.046875 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 356.394531 C 138.871094 359 134.960938 359 134.960938 356.394531 C 134.960938 353.789062 138.871094 353.789062 138.871094 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 356.394531 C 110.988281 359 107.078125 359 107.078125 356.394531 C 107.078125 353.789062 110.988281 353.789062 110.988281 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 356.394531 C 110.988281 359 107.078125 359 107.078125 356.394531 C 107.078125 353.789062 110.988281 353.789062 110.988281 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 395.972656 C 110.988281 398.578125 107.078125 398.578125 107.078125 395.972656 C 107.078125 393.367188 110.988281 393.367188 110.988281 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 395.972656 C 110.988281 398.578125 107.078125 398.578125 107.078125 395.972656 C 107.078125 393.367188 110.988281 393.367188 110.988281 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 277.242188 C 180.691406 279.847656 176.78125 279.847656 176.78125 277.242188 C 176.78125 274.636719 180.691406 274.636719 180.691406 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 264.050781 C 166.75 266.65625 162.839844 266.65625 162.839844 264.050781 C 162.839844 261.441406 166.75 261.441406 166.75 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 290.433594 C 152.808594 293.039062 148.898438 293.039062 148.898438 290.433594 C 148.898438 287.828125 152.808594 287.828125 152.808594 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 330.011719 C 138.871094 332.617188 134.960938 332.617188 134.960938 330.011719 C 134.960938 327.402344 138.871094 327.402344 138.871094 330.011719 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 316.816406 C 138.871094 319.425781 134.960938 319.425781 134.960938 316.816406 C 134.960938 314.210938 138.871094 314.210938 138.871094 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 303.625 C 138.871094 306.230469 134.960938 306.230469 134.960938 303.625 C 134.960938 301.019531 138.871094 301.019531 138.871094 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 316.816406 C 138.871094 319.425781 134.960938 319.425781 134.960938 316.816406 C 134.960938 314.210938 138.871094 314.210938 138.871094 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 343.203125 C 124.929688 345.808594 121.019531 345.808594 121.019531 343.203125 C 121.019531 340.597656 124.929688 340.597656 124.929688 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 320.09375 171.703125 C 320.09375 174.308594 316.183594 174.308594 316.183594 171.703125 C 316.183594 169.097656 320.09375 169.097656 320.09375 171.703125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 264.332031 184.894531 C 264.332031 187.5 260.421875 187.5 260.421875 184.894531 C 260.421875 182.289062 264.332031 182.289062 264.332031 184.894531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 278.273438 184.894531 C 278.273438 187.5 274.363281 187.5 274.363281 184.894531 C 274.363281 182.289062 278.273438 182.289062 278.273438 184.894531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 250.390625 224.472656 C 250.390625 227.078125 246.484375 227.078125 246.484375 224.472656 C 246.484375 221.867188 250.390625 221.867188 250.390625 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 264.332031 184.894531 C 264.332031 187.5 260.421875 187.5 260.421875 184.894531 C 260.421875 182.289062 264.332031 182.289062 264.332031 184.894531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 292.214844 158.511719 C 292.214844 161.117188 288.304688 161.117188 288.304688 158.511719 C 288.304688 155.902344 292.214844 155.902344 292.214844 158.511719 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 278.273438 132.125 C 278.273438 134.730469 274.363281 134.730469 274.363281 132.125 C 274.363281 129.519531 278.273438 129.519531 278.273438 132.125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 264.332031 132.125 C 264.332031 134.730469 260.421875 134.730469 260.421875 132.125 C 260.421875 129.519531 264.332031 129.519531 264.332031 132.125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 250.855469 C 180.691406 253.464844 176.78125 253.464844 176.78125 250.855469 C 176.78125 248.25 180.691406 248.25 180.691406 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 211.28125 C 222.511719 213.886719 218.601562 213.886719 218.601562 211.28125 C 218.601562 208.671875 222.511719 208.671875 222.511719 211.28125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 198.085938 C 222.511719 200.695312 218.601562 200.695312 218.601562 198.085938 C 218.601562 195.480469 222.511719 195.480469 222.511719 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 237.664062 C 194.632812 240.269531 190.722656 240.269531 190.722656 237.664062 C 190.722656 235.058594 194.632812 235.058594 194.632812 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 224.472656 C 194.632812 227.078125 190.722656 227.078125 190.722656 224.472656 C 190.722656 221.867188 194.632812 221.867188 194.632812 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 237.664062 C 208.570312 240.269531 204.660156 240.269531 204.660156 237.664062 C 204.660156 235.058594 208.570312 235.058594 208.570312 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 250.855469 C 208.570312 253.464844 204.660156 253.464844 204.660156 250.855469 C 204.660156 248.25 208.570312 248.25 208.570312 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 290.433594 C 152.808594 293.039062 148.898438 293.039062 148.898438 290.433594 C 148.898438 287.828125 152.808594 287.828125 152.808594 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 316.816406 C 166.75 319.425781 162.839844 319.425781 162.839844 316.816406 C 162.839844 314.210938 166.75 314.210938 166.75 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 356.394531 C 138.871094 359 134.960938 359 134.960938 356.394531 C 134.960938 353.789062 138.871094 353.789062 138.871094 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 343.203125 C 138.871094 345.808594 134.960938 345.808594 134.960938 343.203125 C 134.960938 340.597656 138.871094 340.597656 138.871094 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 55.226562 448.742188 C 55.226562 451.347656 51.316406 451.347656 51.316406 448.742188 C 51.316406 446.136719 55.226562 446.136719 55.226562 448.742188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 356.394531 C 124.929688 359 121.019531 359 121.019531 356.394531 C 121.019531 353.789062 124.929688 353.789062 124.929688 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 369.585938 C 110.988281 372.195312 107.078125 372.195312 107.078125 369.585938 C 107.078125 366.980469 110.988281 366.980469 110.988281 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 422.355469 C 83.109375 424.964844 79.199219 424.964844 79.199219 422.355469 C 79.199219 419.75 83.109375 419.75 83.109375 422.355469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 369.585938 C 97.046875 372.195312 93.140625 372.195312 93.140625 369.585938 C 93.140625 366.980469 97.046875 366.980469 97.046875 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 369.585938 C 97.046875 372.195312 93.140625 372.195312 93.140625 369.585938 C 93.140625 366.980469 97.046875 366.980469 97.046875 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 382.78125 C 83.109375 385.386719 79.199219 385.386719 79.199219 382.78125 C 79.199219 380.171875 83.109375 380.171875 83.109375 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 395.972656 C 83.109375 398.578125 79.199219 398.578125 79.199219 395.972656 C 79.199219 393.367188 83.109375 393.367188 83.109375 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 369.585938 C 97.046875 372.195312 93.140625 372.195312 93.140625 369.585938 C 93.140625 366.980469 97.046875 366.980469 97.046875 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 356.394531 C 110.988281 359 107.078125 359 107.078125 356.394531 C 107.078125 353.789062 110.988281 353.789062 110.988281 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 356.394531 C 110.988281 359 107.078125 359 107.078125 356.394531 C 107.078125 353.789062 110.988281 353.789062 110.988281 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 250.855469 C 194.632812 253.464844 190.722656 253.464844 190.722656 250.855469 C 190.722656 248.25 194.632812 248.25 194.632812 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 250.390625 198.085938 C 250.390625 200.695312 246.484375 200.695312 246.484375 198.085938 C 246.484375 195.480469 250.390625 195.480469 250.390625 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 250.390625 184.894531 C 250.390625 187.5 246.484375 187.5 246.484375 184.894531 C 246.484375 182.289062 250.390625 182.289062 250.390625 184.894531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 250.855469 C 194.632812 253.464844 190.722656 253.464844 190.722656 250.855469 C 190.722656 248.25 194.632812 248.25 194.632812 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 277.242188 C 194.632812 279.847656 190.722656 279.847656 190.722656 277.242188 C 190.722656 274.636719 194.632812 274.636719 194.632812 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 277.242188 C 194.632812 279.847656 190.722656 279.847656 190.722656 277.242188 C 190.722656 274.636719 194.632812 274.636719 194.632812 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 382.78125 C 138.871094 385.386719 134.960938 385.386719 134.960938 382.78125 C 134.960938 380.171875 138.871094 380.171875 138.871094 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 343.203125 C 124.929688 345.808594 121.019531 345.808594 121.019531 343.203125 C 121.019531 340.597656 124.929688 340.597656 124.929688 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 369.585938 C 97.046875 372.195312 93.140625 372.195312 93.140625 369.585938 C 93.140625 366.980469 97.046875 366.980469 97.046875 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 264.050781 C 152.808594 266.65625 148.898438 266.65625 148.898438 264.050781 C 148.898438 261.441406 152.808594 261.441406 152.808594 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 250.855469 C 166.75 253.464844 162.839844 253.464844 162.839844 250.855469 C 162.839844 248.25 166.75 248.25 166.75 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 237.664062 C 180.691406 240.269531 176.78125 240.269531 176.78125 237.664062 C 176.78125 235.058594 180.691406 235.058594 180.691406 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 277.242188 C 152.808594 279.847656 148.898438 279.847656 148.898438 277.242188 C 148.898438 274.636719 152.808594 274.636719 152.808594 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 277.242188 C 180.691406 279.847656 176.78125 279.847656 176.78125 277.242188 C 176.78125 274.636719 180.691406 274.636719 180.691406 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 290.433594 C 180.691406 293.039062 176.78125 293.039062 176.78125 290.433594 C 176.78125 287.828125 180.691406 287.828125 180.691406 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 250.855469 C 208.570312 253.464844 204.660156 253.464844 204.660156 250.855469 C 204.660156 248.25 208.570312 248.25 208.570312 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 277.242188 C 194.632812 279.847656 190.722656 279.847656 190.722656 277.242188 C 190.722656 274.636719 194.632812 274.636719 194.632812 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 264.050781 C 208.570312 266.65625 204.660156 266.65625 204.660156 264.050781 C 204.660156 261.441406 208.570312 261.441406 208.570312 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 303.625 C 180.691406 306.230469 176.78125 306.230469 176.78125 303.625 C 176.78125 301.019531 180.691406 301.019531 180.691406 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 264.050781 C 222.511719 266.65625 218.601562 266.65625 218.601562 264.050781 C 218.601562 261.441406 222.511719 261.441406 222.511719 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 277.242188 C 208.570312 279.847656 204.660156 279.847656 204.660156 277.242188 C 204.660156 274.636719 208.570312 274.636719 208.570312 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 250.855469 C 208.570312 253.464844 204.660156 253.464844 204.660156 250.855469 C 204.660156 248.25 208.570312 248.25 208.570312 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 277.242188 C 194.632812 279.847656 190.722656 279.847656 190.722656 277.242188 C 190.722656 274.636719 194.632812 274.636719 194.632812 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 250.855469 C 208.570312 253.464844 204.660156 253.464844 204.660156 250.855469 C 204.660156 248.25 208.570312 248.25 208.570312 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 343.203125 C 138.871094 345.808594 134.960938 345.808594 134.960938 343.203125 C 134.960938 340.597656 138.871094 340.597656 138.871094 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 343.203125 C 152.808594 345.808594 148.898438 345.808594 148.898438 343.203125 C 148.898438 340.597656 152.808594 340.597656 152.808594 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 356.394531 C 138.871094 359 134.960938 359 134.960938 356.394531 C 134.960938 353.789062 138.871094 353.789062 138.871094 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 382.78125 C 138.871094 385.386719 134.960938 385.386719 134.960938 382.78125 C 134.960938 380.171875 138.871094 380.171875 138.871094 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 343.203125 C 152.808594 345.808594 148.898438 345.808594 148.898438 343.203125 C 148.898438 340.597656 152.808594 340.597656 152.808594 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 250.855469 C 222.511719 253.464844 218.601562 253.464844 218.601562 250.855469 C 218.601562 248.25 222.511719 248.25 222.511719 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 198.085938 C 222.511719 200.695312 218.601562 200.695312 218.601562 198.085938 C 218.601562 195.480469 222.511719 195.480469 222.511719 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 198.085938 C 222.511719 200.695312 218.601562 200.695312 218.601562 198.085938 C 218.601562 195.480469 222.511719 195.480469 222.511719 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 237.664062 C 194.632812 240.269531 190.722656 240.269531 190.722656 237.664062 C 190.722656 235.058594 194.632812 235.058594 194.632812 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 250.855469 C 222.511719 253.464844 218.601562 253.464844 218.601562 250.855469 C 218.601562 248.25 222.511719 248.25 222.511719 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 198.085938 C 222.511719 200.695312 218.601562 200.695312 218.601562 198.085938 C 218.601562 195.480469 222.511719 195.480469 222.511719 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 236.453125 198.085938 C 236.453125 200.695312 232.542969 200.695312 232.542969 198.085938 C 232.542969 195.480469 236.453125 195.480469 236.453125 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 250.855469 C 180.691406 253.464844 176.78125 253.464844 176.78125 250.855469 C 176.78125 248.25 180.691406 248.25 180.691406 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 264.332031 211.28125 C 264.332031 213.886719 260.421875 213.886719 260.421875 211.28125 C 260.421875 208.671875 264.332031 208.671875 264.332031 211.28125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 264.332031 171.703125 C 264.332031 174.308594 260.421875 174.308594 260.421875 171.703125 C 260.421875 169.097656 264.332031 169.097656 264.332031 171.703125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 292.214844 145.316406 C 292.214844 147.925781 288.304688 147.925781 288.304688 145.316406 C 288.304688 142.710938 292.214844 142.710938 292.214844 145.316406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 320.09375 118.933594 C 320.09375 121.539062 316.183594 121.539062 316.183594 118.933594 C 316.183594 116.328125 320.09375 116.328125 320.09375 118.933594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 292.214844 145.316406 C 292.214844 147.925781 288.304688 147.925781 288.304688 145.316406 C 288.304688 142.710938 292.214844 142.710938 292.214844 145.316406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 369.585938 C 110.988281 372.195312 107.078125 372.195312 107.078125 369.585938 C 107.078125 366.980469 110.988281 366.980469 110.988281 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 343.203125 C 138.871094 345.808594 134.960938 345.808594 134.960938 343.203125 C 134.960938 340.597656 138.871094 340.597656 138.871094 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 343.203125 C 152.808594 345.808594 148.898438 345.808594 148.898438 343.203125 C 148.898438 340.597656 152.808594 340.597656 152.808594 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 316.816406 C 166.75 319.425781 162.839844 319.425781 162.839844 316.816406 C 162.839844 314.210938 166.75 314.210938 166.75 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 382.78125 C 138.871094 385.386719 134.960938 385.386719 134.960938 382.78125 C 134.960938 380.171875 138.871094 380.171875 138.871094 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 356.394531 C 138.871094 359 134.960938 359 134.960938 356.394531 C 134.960938 353.789062 138.871094 353.789062 138.871094 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 369.585938 C 138.871094 372.195312 134.960938 372.195312 134.960938 369.585938 C 134.960938 366.980469 138.871094 366.980469 138.871094 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 343.203125 C 152.808594 345.808594 148.898438 345.808594 148.898438 343.203125 C 148.898438 340.597656 152.808594 340.597656 152.808594 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 236.453125 224.472656 C 236.453125 227.078125 232.542969 227.078125 232.542969 224.472656 C 232.542969 221.867188 236.453125 221.867188 236.453125 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 389.796875 26.585938 C 389.796875 29.195312 385.886719 29.195312 385.886719 26.585938 C 385.886719 23.980469 389.796875 23.980469 389.796875 26.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 236.453125 224.472656 C 236.453125 227.078125 232.542969 227.078125 232.542969 224.472656 C 232.542969 221.867188 236.453125 221.867188 236.453125 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 303.625 C 152.808594 306.230469 148.898438 306.230469 148.898438 303.625 C 148.898438 301.019531 152.808594 301.019531 152.808594 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 417.675781 26.585938 C 417.675781 29.195312 413.765625 29.195312 413.765625 26.585938 C 413.765625 23.980469 417.675781 23.980469 417.675781 26.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 334.035156 66.164062 C 334.035156 68.769531 330.125 68.769531 330.125 66.164062 C 330.125 63.558594 334.035156 63.558594 334.035156 66.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 237.664062 C 208.570312 240.269531 204.660156 240.269531 204.660156 237.664062 C 204.660156 235.058594 208.570312 235.058594 208.570312 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 224.472656 C 208.570312 227.078125 204.660156 227.078125 204.660156 224.472656 C 204.660156 221.867188 208.570312 221.867188 208.570312 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 224.472656 C 180.691406 227.078125 176.78125 227.078125 176.78125 224.472656 C 176.78125 221.867188 180.691406 221.867188 180.691406 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 237.664062 C 194.632812 240.269531 190.722656 240.269531 190.722656 237.664062 C 190.722656 235.058594 194.632812 235.058594 194.632812 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 264.050781 C 152.808594 266.65625 148.898438 266.65625 148.898438 264.050781 C 148.898438 261.441406 152.808594 261.441406 152.808594 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 264.050781 C 166.75 266.65625 162.839844 266.65625 162.839844 264.050781 C 162.839844 261.441406 166.75 261.441406 166.75 264.050781 \"/&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"20.21875\" y=\"346.118164\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"25.21875\" y=\"346.118164\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-3\" x=\"20.21875\" y=\"214.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"25.21875\" y=\"214.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-4\" x=\"20.21875\" y=\"82.270508\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"25.21875\" y=\"82.270508\"/&gt;\n&lt;/g&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 32.410156 343.203125 L 35.152344 343.203125 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 32.410156 211.28125 L 35.152344 211.28125 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 32.410156 79.355469 L 35.152344 79.355469 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 67.214844 472.589844 L 67.214844 469.847656 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 136.914062 472.589844 L 136.914062 469.847656 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 206.617188 472.589844 L 206.617188 469.847656 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 276.316406 472.589844 L 276.316406 469.847656 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 346.019531 472.589844 L 346.019531 469.847656 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 415.722656 472.589844 L 415.722656 469.847656 \"/&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-5\" x=\"62.214844\" y=\"481.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"67.214844\" y=\"481.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-5\" x=\"131.914062\" y=\"481.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-6\" x=\"136.914062\" y=\"481.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"201.617188\" y=\"481.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"206.617188\" y=\"481.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"271.316406\" y=\"481.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-6\" x=\"276.316406\" y=\"481.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-3\" x=\"341.019531\" y=\"481.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"346.019531\" y=\"481.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-3\" x=\"410.722656\" y=\"481.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-6\" x=\"415.722656\" y=\"481.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph1-1\" x=\"227.496094\" y=\"494.538086\"/&gt;\n  &lt;use xlink:href=\"#glyph1-2\" x=\"233.496094\" y=\"494.538086\"/&gt;\n  &lt;use xlink:href=\"#glyph1-3\" x=\"236.496094\" y=\"494.538086\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph2-1\" x=\"13.499023\" y=\"247.164062\"/&gt;\n  &lt;use xlink:href=\"#glyph2-2\" x=\"13.499023\" y=\"241.164062\"/&gt;\n  &lt;use xlink:href=\"#glyph2-3\" x=\"13.499023\" y=\"233.164062\"/&gt;\n&lt;/g&gt;\n&lt;path style=\" stroke:none;fill-rule:nonzero;fill:rgb(100%,100%,100%);fill-opacity:1;\" d=\"M 444.800781 295.082031 L 498.519531 295.082031 L 498.519531 180.242188 L 444.800781 180.242188 Z M 444.800781 295.082031 \"/&gt;\n&lt;g clip-path=\"url(#clip20)\" clip-rule=\"nonzero\"&gt;\n&lt;rect x=\"0\" y=\"0\" width=\"504\" height=\"504\" style=\"fill:url(#pattern0);stroke:none;\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"473.042969\" y=\"282.805664\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"478.042969\" y=\"282.805664\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"483.042969\" y=\"282.805664\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"488.042969\" y=\"282.805664\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"473.042969\" y=\"263.672852\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"478.042969\" y=\"263.672852\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"483.042969\" y=\"263.672852\"/&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"488.042969\" y=\"263.672852\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"473.042969\" y=\"244.536133\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"478.042969\" y=\"244.536133\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"483.042969\" y=\"244.536133\"/&gt;\n  &lt;use xlink:href=\"#glyph0-4\" x=\"488.042969\" y=\"244.536133\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"473.042969\" y=\"225.399414\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"478.042969\" y=\"225.399414\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"483.042969\" y=\"225.399414\"/&gt;\n  &lt;use xlink:href=\"#glyph0-7\" x=\"488.042969\" y=\"225.399414\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"473.042969\" y=\"206.262695\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"478.042969\" y=\"206.262695\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"483.042969\" y=\"206.262695\"/&gt;\n  &lt;use xlink:href=\"#glyph0-8\" x=\"488.042969\" y=\"206.262695\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph1-3\" x=\"450.28125\" y=\"195.245117\"/&gt;\n  &lt;use xlink:href=\"#glyph1-4\" x=\"455.28125\" y=\"195.245117\"/&gt;\n  &lt;use xlink:href=\"#glyph1-5\" x=\"461.28125\" y=\"195.245117\"/&gt;\n  &lt;use xlink:href=\"#glyph1-6\" x=\"467.28125\" y=\"195.245117\"/&gt;\n&lt;/g&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 450.28125 279.890625 L 453.738281 279.890625 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 450.28125 260.757812 L 453.738281 260.757812 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 450.28125 241.621094 L 453.738281 241.621094 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 450.28125 222.484375 L 453.738281 222.484375 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 450.28125 203.347656 L 453.738281 203.347656 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 464.105469 279.890625 L 467.5625 279.890625 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 464.105469 260.757812 L 467.5625 260.757812 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 464.105469 241.621094 L 467.5625 241.621094 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 464.105469 222.484375 L 467.5625 222.484375 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 464.105469 203.347656 L 467.5625 203.347656 \"/&gt;\n&lt;/g&gt;\n&lt;/svg&gt;\n"

```
:::

Use ggplot2 as a Clojure namespace:

::: {.sourceClojure}
```clojure
(r/require-r '[ggplot2 :as gg])
```
:::

::: {.printedClojure}
```clojure
nil

```
:::

The `mpg` dataset as a Clojure (tech.ml.dataset) dataset:

::: {.sourceClojure}
```clojure
toydata.ggplot/mpg
```
:::

_unnamed [234 11]:

| :manufacturer |     :model | :displ | :year | :cyl |     :trans | :drv | :cty | :hwy | :fl |     :class |
|---------------|------------|-------:|------:|-----:|------------|------|-----:|-----:|-----|------------|
|          audi |         a4 |    1.8 |  1999 |    4 |   auto(l5) |    f |   18 |   29 |   p |    compact |
|          audi |         a4 |    1.8 |  1999 |    4 | manual(m5) |    f |   21 |   29 |   p |    compact |
|          audi |         a4 |    2.0 |  2008 |    4 | manual(m6) |    f |   20 |   31 |   p |    compact |
|          audi |         a4 |    2.0 |  2008 |    4 |   auto(av) |    f |   21 |   30 |   p |    compact |
|          audi |         a4 |    2.8 |  1999 |    6 |   auto(l5) |    f |   16 |   26 |   p |    compact |
|          audi |         a4 |    2.8 |  1999 |    6 | manual(m5) |    f |   18 |   26 |   p |    compact |
|          audi |         a4 |    3.1 |  2008 |    6 |   auto(av) |    f |   18 |   27 |   p |    compact |
|          audi | a4 quattro |    1.8 |  1999 |    4 | manual(m5) |    4 |   18 |   26 |   p |    compact |
|          audi | a4 quattro |    1.8 |  1999 |    4 |   auto(l5) |    4 |   16 |   25 |   p |    compact |
|          audi | a4 quattro |    2.0 |  2008 |    4 | manual(m6) |    4 |   20 |   28 |   p |    compact |
|           ... |        ... |    ... |   ... |  ... |        ... |  ... |  ... |  ... | ... |        ... |
|    volkswagen | new beetle |    2.0 |  1999 |    4 | manual(m5) |    f |   21 |   29 |   r | subcompact |
|    volkswagen | new beetle |    2.0 |  1999 |    4 |   auto(l4) |    f |   19 |   26 |   r | subcompact |
|    volkswagen | new beetle |    2.5 |  2008 |    5 | manual(m5) |    f |   20 |   28 |   r | subcompact |
|    volkswagen | new beetle |    2.5 |  2008 |    5 |   auto(s6) |    f |   20 |   29 |   r | subcompact |
|    volkswagen |     passat |    1.8 |  1999 |    4 | manual(m5) |    f |   21 |   29 |   p |    midsize |
|    volkswagen |     passat |    1.8 |  1999 |    4 |   auto(l5) |    f |   18 |   29 |   p |    midsize |
|    volkswagen |     passat |    2.0 |  2008 |    4 |   auto(s6) |    f |   19 |   28 |   p |    midsize |
|    volkswagen |     passat |    2.0 |  2008 |    4 | manual(m6) |    f |   21 |   29 |   p |    midsize |
|    volkswagen |     passat |    2.8 |  1999 |    6 |   auto(l5) |    f |   16 |   26 |   p |    midsize |
|    volkswagen |     passat |    2.8 |  1999 |    6 | manual(m5) |    f |   18 |   26 |   p |    midsize |
|    volkswagen |     passat |    3.6 |  2008 |    6 |   auto(s6) |    f |   17 |   26 |   p |    midsize |

Evaluating R code generated from the Clojure namespace,
with Clojure data.

::: {.sourceClojure}
```clojure
(-> toydata.ggplot/mpg
    (gg/ggplot (gg/aes :x 'cty
                       :y 'hwy
                       :color 'year))
    (r/r+ (gg/geom_point))
    r
    plotting/plot->buffered-image)
```
:::

```{=html}
<img src="study_session_20240225_files/1.png" style="width:100%;" />
```

Representing a ggplot plot as Clojure data:

::: {.sourceClojure}
```clojure
(-> toydata.ggplot/mpg
    (gg/ggplot (gg/aes :x 'cty
                       :y 'hwy
                       :color 'year))
    (r/r+ (gg/geom_point))
    r
    representing/ggplot->clj
    ;; (dissoc the dataset itself
    ;; to keep the structure small)
    (dissoc :data))
```
:::

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [[$ .MEM xfeea946278794b6e]],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ year]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["year"]}}

```
:::

Using log scale
(keeping axis labels, changing the spacing):

::: {.sourceClojure}
```clojure
(-> toydata.ggplot/mpg
    (gg/ggplot (gg/aes :x 'cty
                       :y 'hwy
                       :color 'year))
    (r/r+ (gg/geom_point)
          (gg/scale_y_log10))
    r
    plotting/plot->buffered-image)
```
:::

```{=html}
<img src="study_session_20240225_files/2.png" style="width:100%;" />
```

```{=html}
<div style="background-color:grey;height:2px;width:100%;"></div>
```

```{=html}
<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/main/projects/datavis/ggplot/notebooks/study_session_20240225.clj">projects/datavis/ggplot/notebooks/study_session_20240225.clj</a></small></small></pre></div>
```