### 2024-03-07 docs/projects/visual-tools/clay-calva-demo-20231216/index.qmd
Clay, Calva, & Noj - rerendered

---
format:
  html: {toc: true, toc-depth: 4, theme: spacelab, output-file: index.html}
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style>
<script src="index_files/md-default1.js" type="text/javascript"></script><script src="index_files/md-default2.js" type="text/javascript"></script><script src="index_files/vega3.js" type="text/javascript"></script><script src="index_files/vega4.js" type="text/javascript"></script><script src="index_files/vega5.js" type="text/javascript"></script>

```{=html}
<table><tr><td>This is part of the Scicloj <a href="https://scicloj.github.io/clojure-data-scrapbook/">Clojure Data Scrapbook</a>.</td><a href="https://scicloj.github.io/clojure-data-scrapbook/"><img align="left" alt="SciCloj logo" src="https://scicloj.github.io/sci-cloj-logo-transparent.png" width="40" /></a></tr></table>
```

# Clay & Noj demo: data visualization

::: {.sourceClojure}
```clojure
(ns index
  (:require [scicloj.metamorph.ml.toydata
             :as toydata]
            [scicloj.noj.v1.vis.hanami
             :as hanami]
            [scicloj.kindly.v4.kind
             :as kind]))
```
:::

::: {.sourceClojure}
```clojure
(kind/video {:youtube-id "X_SsjhmG5Ok"})
```
:::

```{=html}
<iframe allowfullscreen="allowfullscreen" src="https://www.youtube.com/embed/X_SsjhmG5Ok"></iframe>
```

## Arithmetic

::: {.sourceClojure}
```clojure
(+ 1 2)
```
:::

::: {.printedClojure}
```clojure
3

```
:::

## Datasets

::: {.sourceClojure}
```clojure
(def iris
  (toydata/iris-ds))
```
:::

::: {.sourceClojure}
```clojure
iris
```
:::

_unnamed [150 5]:

| :sepal_length | :sepal_width | :petal_length | :petal_width | :species |
|--------------:|-------------:|--------------:|-------------:|---------:|
|           5.1 |          3.5 |           1.4 |          0.2 |        0 |
|           4.9 |          3.0 |           1.4 |          0.2 |        0 |
|           4.7 |          3.2 |           1.3 |          0.2 |        0 |
|           4.6 |          3.1 |           1.5 |          0.2 |        0 |
|           5.0 |          3.6 |           1.4 |          0.2 |        0 |
|           5.4 |          3.9 |           1.7 |          0.4 |        0 |
|           4.6 |          3.4 |           1.4 |          0.3 |        0 |
|           5.0 |          3.4 |           1.5 |          0.2 |        0 |
|           4.4 |          2.9 |           1.4 |          0.2 |        0 |
|           4.9 |          3.1 |           1.5 |          0.1 |        0 |
|           ... |          ... |           ... |          ... |      ... |
|           6.9 |          3.1 |           5.4 |          2.1 |        1 |
|           6.7 |          3.1 |           5.6 |          2.4 |        1 |
|           6.9 |          3.1 |           5.1 |          2.3 |        1 |
|           5.8 |          2.7 |           5.1 |          1.9 |        1 |
|           6.8 |          3.2 |           5.9 |          2.3 |        1 |
|           6.7 |          3.3 |           5.7 |          2.5 |        1 |
|           6.7 |          3.0 |           5.2 |          2.3 |        1 |
|           6.3 |          2.5 |           5.0 |          1.9 |        1 |
|           6.5 |          3.0 |           5.2 |          2.0 |        1 |
|           6.2 |          3.4 |           5.4 |          2.3 |        1 |
|           5.9 |          3.0 |           5.1 |          1.8 |        1 |

## Tables

::: {.sourceClojure}
```clojure
(-> iris
    (kind/table {:element/max-height "500px"}))
```
:::

```{=html}
<div style="max-height:500px;overflow-y:auto;"><table class="table table-hover table-responsive"><thead><tr><th>:sepal_length</th><th>:sepal_width</th><th>:petal_length</th><th>:petal_width</th><th>:species</th></tr></thead><tbody><tr><td><span data-qmd="5.1"></span></td><td><span data-qmd="3.5"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.9"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.7"></span></td><td><span data-qmd="3.2"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.6"></span></td><td><span data-qmd="3.1"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.0"></span></td><td><span data-qmd="3.6"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.4"></span></td><td><span data-qmd="3.9"></span></td><td><span data-qmd="1.7"></span></td><td><span data-qmd="0.4"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.6"></span></td><td><span data-qmd="3.4"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="0.3"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.0"></span></td><td><span data-qmd="3.4"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.4"></span></td><td><span data-qmd="2.9"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.9"></span></td><td><span data-qmd="3.1"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="0.1"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.4"></span></td><td><span data-qmd="3.7"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.8"></span></td><td><span data-qmd="3.4"></span></td><td><span data-qmd="1.6"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.8"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="0.1"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.3"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="1.1"></span></td><td><span data-qmd="0.1"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.8"></span></td><td><span data-qmd="4.0"></span></td><td><span data-qmd="1.2"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.7"></span></td><td><span data-qmd="4.4"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="0.4"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.4"></span></td><td><span data-qmd="3.9"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="0.4"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.1"></span></td><td><span data-qmd="3.5"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="0.3"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.7"></span></td><td><span data-qmd="3.8"></span></td><td><span data-qmd="1.7"></span></td><td><span data-qmd="0.3"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.1"></span></td><td><span data-qmd="3.8"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="0.3"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.4"></span></td><td><span data-qmd="3.4"></span></td><td><span data-qmd="1.7"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.1"></span></td><td><span data-qmd="3.7"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="0.4"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.6"></span></td><td><span data-qmd="3.6"></span></td><td><span data-qmd="1.0"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.1"></span></td><td><span data-qmd="3.3"></span></td><td><span data-qmd="1.7"></span></td><td><span data-qmd="0.5"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.8"></span></td><td><span data-qmd="3.4"></span></td><td><span data-qmd="1.9"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.0"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="1.6"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.0"></span></td><td><span data-qmd="3.4"></span></td><td><span data-qmd="1.6"></span></td><td><span data-qmd="0.4"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.2"></span></td><td><span data-qmd="3.5"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.2"></span></td><td><span data-qmd="3.4"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.7"></span></td><td><span data-qmd="3.2"></span></td><td><span data-qmd="1.6"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.8"></span></td><td><span data-qmd="3.1"></span></td><td><span data-qmd="1.6"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.4"></span></td><td><span data-qmd="3.4"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="0.4"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.2"></span></td><td><span data-qmd="4.1"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="0.1"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.5"></span></td><td><span data-qmd="4.2"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.9"></span></td><td><span data-qmd="3.1"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.0"></span></td><td><span data-qmd="3.2"></span></td><td><span data-qmd="1.2"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.5"></span></td><td><span data-qmd="3.5"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.9"></span></td><td><span data-qmd="3.6"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="0.1"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.4"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.1"></span></td><td><span data-qmd="3.4"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.0"></span></td><td><span data-qmd="3.5"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="0.3"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.5"></span></td><td><span data-qmd="2.3"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="0.3"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.4"></span></td><td><span data-qmd="3.2"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.0"></span></td><td><span data-qmd="3.5"></span></td><td><span data-qmd="1.6"></span></td><td><span data-qmd="0.6"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.1"></span></td><td><span data-qmd="3.8"></span></td><td><span data-qmd="1.9"></span></td><td><span data-qmd="0.4"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.8"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="0.3"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.1"></span></td><td><span data-qmd="3.8"></span></td><td><span data-qmd="1.6"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="4.6"></span></td><td><span data-qmd="3.2"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.3"></span></td><td><span data-qmd="3.7"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="5.0"></span></td><td><span data-qmd="3.3"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="0.2"></span></td><td><span data-qmd="0"></span></td></tr><tr><td><span data-qmd="7.0"></span></td><td><span data-qmd="3.2"></span></td><td><span data-qmd="4.7"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.4"></span></td><td><span data-qmd="3.2"></span></td><td><span data-qmd="4.5"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.9"></span></td><td><span data-qmd="3.1"></span></td><td><span data-qmd="4.9"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.5"></span></td><td><span data-qmd="2.3"></span></td><td><span data-qmd="4.0"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.5"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="4.6"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.7"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="4.5"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.3"></span></td><td><span data-qmd="3.3"></span></td><td><span data-qmd="4.7"></span></td><td><span data-qmd="1.6"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="4.9"></span></td><td><span data-qmd="2.4"></span></td><td><span data-qmd="3.3"></span></td><td><span data-qmd="1.0"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.6"></span></td><td><span data-qmd="2.9"></span></td><td><span data-qmd="4.6"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.2"></span></td><td><span data-qmd="2.7"></span></td><td><span data-qmd="3.9"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.0"></span></td><td><span data-qmd="2.0"></span></td><td><span data-qmd="3.5"></span></td><td><span data-qmd="1.0"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.9"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="4.2"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.0"></span></td><td><span data-qmd="2.2"></span></td><td><span data-qmd="4.0"></span></td><td><span data-qmd="1.0"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.1"></span></td><td><span data-qmd="2.9"></span></td><td><span data-qmd="4.7"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.6"></span></td><td><span data-qmd="2.9"></span></td><td><span data-qmd="3.6"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.7"></span></td><td><span data-qmd="3.1"></span></td><td><span data-qmd="4.4"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.6"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="4.5"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.8"></span></td><td><span data-qmd="2.7"></span></td><td><span data-qmd="4.1"></span></td><td><span data-qmd="1.0"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.2"></span></td><td><span data-qmd="2.2"></span></td><td><span data-qmd="4.5"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.6"></span></td><td><span data-qmd="2.5"></span></td><td><span data-qmd="3.9"></span></td><td><span data-qmd="1.1"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.9"></span></td><td><span data-qmd="3.2"></span></td><td><span data-qmd="4.8"></span></td><td><span data-qmd="1.8"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.1"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="4.0"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.3"></span></td><td><span data-qmd="2.5"></span></td><td><span data-qmd="4.9"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.1"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="4.7"></span></td><td><span data-qmd="1.2"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.4"></span></td><td><span data-qmd="2.9"></span></td><td><span data-qmd="4.3"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.6"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="4.4"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.8"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="4.8"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.7"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="5.0"></span></td><td><span data-qmd="1.7"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.0"></span></td><td><span data-qmd="2.9"></span></td><td><span data-qmd="4.5"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.7"></span></td><td><span data-qmd="2.6"></span></td><td><span data-qmd="3.5"></span></td><td><span data-qmd="1.0"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.5"></span></td><td><span data-qmd="2.4"></span></td><td><span data-qmd="3.8"></span></td><td><span data-qmd="1.1"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.5"></span></td><td><span data-qmd="2.4"></span></td><td><span data-qmd="3.7"></span></td><td><span data-qmd="1.0"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.8"></span></td><td><span data-qmd="2.7"></span></td><td><span data-qmd="3.9"></span></td><td><span data-qmd="1.2"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.0"></span></td><td><span data-qmd="2.7"></span></td><td><span data-qmd="5.1"></span></td><td><span data-qmd="1.6"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.4"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="4.5"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.0"></span></td><td><span data-qmd="3.4"></span></td><td><span data-qmd="4.5"></span></td><td><span data-qmd="1.6"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.7"></span></td><td><span data-qmd="3.1"></span></td><td><span data-qmd="4.7"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.3"></span></td><td><span data-qmd="2.3"></span></td><td><span data-qmd="4.4"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.6"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="4.1"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.5"></span></td><td><span data-qmd="2.5"></span></td><td><span data-qmd="4.0"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.5"></span></td><td><span data-qmd="2.6"></span></td><td><span data-qmd="4.4"></span></td><td><span data-qmd="1.2"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.1"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="4.6"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.8"></span></td><td><span data-qmd="2.6"></span></td><td><span data-qmd="4.0"></span></td><td><span data-qmd="1.2"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.0"></span></td><td><span data-qmd="2.3"></span></td><td><span data-qmd="3.3"></span></td><td><span data-qmd="1.0"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.6"></span></td><td><span data-qmd="2.7"></span></td><td><span data-qmd="4.2"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.7"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="4.2"></span></td><td><span data-qmd="1.2"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.7"></span></td><td><span data-qmd="2.9"></span></td><td><span data-qmd="4.2"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.2"></span></td><td><span data-qmd="2.9"></span></td><td><span data-qmd="4.3"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.1"></span></td><td><span data-qmd="2.5"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="1.1"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="5.7"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="4.1"></span></td><td><span data-qmd="1.3"></span></td><td><span data-qmd="2"></span></td></tr><tr><td><span data-qmd="6.3"></span></td><td><span data-qmd="3.3"></span></td><td><span data-qmd="6.0"></span></td><td><span data-qmd="2.5"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="5.8"></span></td><td><span data-qmd="2.7"></span></td><td><span data-qmd="5.1"></span></td><td><span data-qmd="1.9"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="7.1"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="5.9"></span></td><td><span data-qmd="2.1"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.3"></span></td><td><span data-qmd="2.9"></span></td><td><span data-qmd="5.6"></span></td><td><span data-qmd="1.8"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.5"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="5.8"></span></td><td><span data-qmd="2.2"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="7.6"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="6.6"></span></td><td><span data-qmd="2.1"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="4.9"></span></td><td><span data-qmd="2.5"></span></td><td><span data-qmd="4.5"></span></td><td><span data-qmd="1.7"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="7.3"></span></td><td><span data-qmd="2.9"></span></td><td><span data-qmd="6.3"></span></td><td><span data-qmd="1.8"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.7"></span></td><td><span data-qmd="2.5"></span></td><td><span data-qmd="5.8"></span></td><td><span data-qmd="1.8"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="7.2"></span></td><td><span data-qmd="3.6"></span></td><td><span data-qmd="6.1"></span></td><td><span data-qmd="2.5"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.5"></span></td><td><span data-qmd="3.2"></span></td><td><span data-qmd="5.1"></span></td><td><span data-qmd="2.0"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.4"></span></td><td><span data-qmd="2.7"></span></td><td><span data-qmd="5.3"></span></td><td><span data-qmd="1.9"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.8"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="5.5"></span></td><td><span data-qmd="2.1"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="5.7"></span></td><td><span data-qmd="2.5"></span></td><td><span data-qmd="5.0"></span></td><td><span data-qmd="2.0"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="5.8"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="5.1"></span></td><td><span data-qmd="2.4"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.4"></span></td><td><span data-qmd="3.2"></span></td><td><span data-qmd="5.3"></span></td><td><span data-qmd="2.3"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.5"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="5.5"></span></td><td><span data-qmd="1.8"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="7.7"></span></td><td><span data-qmd="3.8"></span></td><td><span data-qmd="6.7"></span></td><td><span data-qmd="2.2"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="7.7"></span></td><td><span data-qmd="2.6"></span></td><td><span data-qmd="6.9"></span></td><td><span data-qmd="2.3"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.0"></span></td><td><span data-qmd="2.2"></span></td><td><span data-qmd="5.0"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.9"></span></td><td><span data-qmd="3.2"></span></td><td><span data-qmd="5.7"></span></td><td><span data-qmd="2.3"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="5.6"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="4.9"></span></td><td><span data-qmd="2.0"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="7.7"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="6.7"></span></td><td><span data-qmd="2.0"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.3"></span></td><td><span data-qmd="2.7"></span></td><td><span data-qmd="4.9"></span></td><td><span data-qmd="1.8"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.7"></span></td><td><span data-qmd="3.3"></span></td><td><span data-qmd="5.7"></span></td><td><span data-qmd="2.1"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="7.2"></span></td><td><span data-qmd="3.2"></span></td><td><span data-qmd="6.0"></span></td><td><span data-qmd="1.8"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.2"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="4.8"></span></td><td><span data-qmd="1.8"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.1"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="4.9"></span></td><td><span data-qmd="1.8"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.4"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="5.6"></span></td><td><span data-qmd="2.1"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="7.2"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="5.8"></span></td><td><span data-qmd="1.6"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="7.4"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="6.1"></span></td><td><span data-qmd="1.9"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="7.9"></span></td><td><span data-qmd="3.8"></span></td><td><span data-qmd="6.4"></span></td><td><span data-qmd="2.0"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.4"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="5.6"></span></td><td><span data-qmd="2.2"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.3"></span></td><td><span data-qmd="2.8"></span></td><td><span data-qmd="5.1"></span></td><td><span data-qmd="1.5"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.1"></span></td><td><span data-qmd="2.6"></span></td><td><span data-qmd="5.6"></span></td><td><span data-qmd="1.4"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="7.7"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="6.1"></span></td><td><span data-qmd="2.3"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.3"></span></td><td><span data-qmd="3.4"></span></td><td><span data-qmd="5.6"></span></td><td><span data-qmd="2.4"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.4"></span></td><td><span data-qmd="3.1"></span></td><td><span data-qmd="5.5"></span></td><td><span data-qmd="1.8"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.0"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="4.8"></span></td><td><span data-qmd="1.8"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.9"></span></td><td><span data-qmd="3.1"></span></td><td><span data-qmd="5.4"></span></td><td><span data-qmd="2.1"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.7"></span></td><td><span data-qmd="3.1"></span></td><td><span data-qmd="5.6"></span></td><td><span data-qmd="2.4"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.9"></span></td><td><span data-qmd="3.1"></span></td><td><span data-qmd="5.1"></span></td><td><span data-qmd="2.3"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="5.8"></span></td><td><span data-qmd="2.7"></span></td><td><span data-qmd="5.1"></span></td><td><span data-qmd="1.9"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.8"></span></td><td><span data-qmd="3.2"></span></td><td><span data-qmd="5.9"></span></td><td><span data-qmd="2.3"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.7"></span></td><td><span data-qmd="3.3"></span></td><td><span data-qmd="5.7"></span></td><td><span data-qmd="2.5"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.7"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="5.2"></span></td><td><span data-qmd="2.3"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.3"></span></td><td><span data-qmd="2.5"></span></td><td><span data-qmd="5.0"></span></td><td><span data-qmd="1.9"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.5"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="5.2"></span></td><td><span data-qmd="2.0"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="6.2"></span></td><td><span data-qmd="3.4"></span></td><td><span data-qmd="5.4"></span></td><td><span data-qmd="2.3"></span></td><td><span data-qmd="1"></span></td></tr><tr><td><span data-qmd="5.9"></span></td><td><span data-qmd="3.0"></span></td><td><span data-qmd="5.1"></span></td><td><span data-qmd="1.8"></span></td><td><span data-qmd="1"></span></td></tr></tbody></table></div>
```

## Visualization

::: {.sourceClojure}
```clojure
(-> iris
    (hanami/linear-regression-plot
     :petal_length
     :petal_width
     {:line-options {:MCOLOR "brown"
                     :MSIZE 10
                     :OPACITY 0.5}
      :point-options {:MSIZE 100}}))
```
:::

```{=html}
<div><script>vegaEmbed(document.currentScript.parentElement, {"width":400,"background":"floralwhite","title":{"text":"R^2 = 0.927"},"layer":[{"encoding":{"y":{"field":"petal_length","type":"quantitative"},"x":{"field":"petal_width","type":"quantitative"}},"mark":{"type":"circle","size":100,"tooltip":true},"width":400,"background":"floralwhite","title":{"text":"R^2 = 0.927"},"height":300},{"encoding":{"y":{"field":"petal_length-prediction","type":"quantitative","axis":{"title":"petal_length"}},"opacity":{"value":0.5},"x":{"field":"petal_width","type":"quantitative"}},"mark":{"type":"line","size":10,"color":"brown","tooltip":true},"width":400,"background":"floralwhite","title":{"text":"R^2 = 0.927"},"height":300}],"height":300,"data":{"url":"index_files\/0.csv","format":{"type":"csv"}}});</script></div>
```

```{=html}
<div style="background-color:grey;height:2px;width:100%;"></div>
```

```{=html}
<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/main/projects/visual-tools/clay-calva-demo-20231216/notebooks/index.clj">projects/visual-tools/clay-calva-demo-20231216/notebooks/index.clj</a></small></small></pre></div>
```