### 2024-03-19 docs/projects/datavis/ggplot/study_session_20240225.qmd

---
format:
  html: {toc: true, toc-depth: 4, theme: cosmo, output-file: study_session_20240225.html}
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style><style>.clay-limit-image-width .clay-image {max-width: 100%}
</style>
<script src="study_session_20240225_files/md-default2.js" type="text/javascript"></script><script src="study_session_20240225_files/md-default3.js" type="text/javascript"></script>

```{=html}
<table><tr><td>This is part of the Scicloj <a href="https://scicloj.github.io/clojure-data-scrapbook/">Clojure Data Scrapbook</a>.</td><a href="https://scicloj.github.io/clojure-data-scrapbook/"><img align="left" alt="SciCloj logo" src="https://scicloj.github.io/sci-cloj-logo-transparent.png" width="40" /></a></tr></table>
```

::: {.sourceClojure}
```clojure
(ns study-session-20240225
  (:require [clojisr.v1.r :as r :refer [r]]
            [clojisr.v1.applications.plotting :as plotting]
            [scicloj.kindly.v4.kind :as kind]
            [tablecloth.api :as tc]
            [scicloj.metamorph.ml.toydata :as toydata]
            [scicloj.metamorph.ml.toydata.ggplot :as toydata.ggplot]
            [representing]))
```
:::

# Code examples from study session 2024-02-25

Add ggplot2 to the R runtime:

::: {.sourceClojure}
```clojure
(r/library "ggplot2")
```
:::

::: {.printedClojure}
```clojure
[1] "ggplot2"   "Rserve"    "stats"     "graphics"  "grDevices" "utils"    
[7] "datasets"  "methods"   "base"

```
:::

The `mpg` dataset:

::: {.sourceClojure}
```clojure
(r "mpg")
```
:::

::: {.printedClojure}
```clojure
# A tibble: 234 × 11
   manufacturer model      displ  year   cyl trans drv     cty   hwy fl    class
   &lt;chr&gt;        &lt;chr&gt;      &lt;dbl&gt; &lt;int&gt; &lt;int&gt; &lt;chr&gt; &lt;chr&gt; &lt;int&gt; &lt;int&gt; &lt;chr&gt; &lt;chr&gt;
 1 audi         a4           1.8  1999     4 auto… f        18    29 p     comp…
 2 audi         a4           1.8  1999     4 manu… f        21    29 p     comp…
 3 audi         a4           2    2008     4 manu… f        20    31 p     comp…
 4 audi         a4           2    2008     4 auto… f        21    30 p     comp…
 5 audi         a4           2.8  1999     6 auto… f        16    26 p     comp…
 6 audi         a4           2.8  1999     6 manu… f        18    26 p     comp…
 7 audi         a4           3.1  2008     6 auto… f        18    27 p     comp…
 8 audi         a4 quattro   1.8  1999     4 manu… 4        18    26 p     comp…
 9 audi         a4 quattro   1.8  1999     4 auto… 4        16    25 p     comp…
10 audi         a4 quattro   2    2008     4 manu… 4        20    28 p     comp…
# ℹ 224 more rows

```
:::

Evaluating R and showing a plot as a bitmap image:

::: {.sourceClojure}
```clojure
(plotting/plot->buffered-image
 (r "
(ggplot(mpg, aes(x = cty, y = hwy, color = year)) +
geom_point())
"))
```
:::

::: {.clay-image}

```{=html}
<img src="study_session_20240225_files/0.png" />
```

:::

Showing a plot as svg:

::: {.sourceClojure}
```clojure
(plotting/plot->svg
 (r "
(ggplot(mpg, aes(x = cty, y = hwy, color = year)) +
geom_point())
"))
```
:::

::: {.printedClojure}
```clojure
"&lt;?xml version=\"1.0\" encoding=\"UTF-8\"?&gt;\n&lt;svg xmlns=\"http://www.w3.org/2000/svg\" xmlns:xlink=\"http://www.w3.org/1999/xlink\" width=\"504pt\" height=\"504pt\" viewBox=\"0 0 504 504\" version=\"1.1\"&gt;\n&lt;defs&gt;\n&lt;g&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-0\"&gt;\n&lt;path style=\"stroke:none;\" d=\"\"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-1\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 4.453125 -0.765625 L 1.171875 -0.765625 C 1.25 -1.296875 1.53125 -1.625 2.296875 -2.09375 L 3.171875 -2.59375 C 4.046875 -3.0625 4.5 -3.71875 4.5 -4.5 C 4.5 -5.03125 4.28125 -5.53125 3.921875 -5.875 C 3.546875 -6.203125 3.09375 -6.359375 2.5 -6.359375 C 1.703125 -6.359375 1.125 -6.078125 0.78125 -5.53125 C 0.546875 -5.203125 0.453125 -4.796875 0.4375 -4.15625 L 1.21875 -4.15625 C 1.234375 -4.59375 1.296875 -4.84375 1.40625 -5.0625 C 1.59375 -5.4375 2 -5.6875 2.46875 -5.6875 C 3.171875 -5.6875 3.703125 -5.171875 3.703125 -4.484375 C 3.703125 -3.96875 3.40625 -3.53125 2.859375 -3.21875 L 2.046875 -2.75 C 0.75 -2 0.375 -1.40625 0.296875 -0.015625 L 4.453125 -0.015625 Z M 4.453125 -0.765625 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-2\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 2.421875 -6.359375 C 1.84375 -6.359375 1.3125 -6.09375 0.984375 -5.671875 C 0.578125 -5.109375 0.375 -4.25 0.375 -3.078125 C 0.375 -0.9375 1.078125 0.203125 2.421875 0.203125 C 3.734375 0.203125 4.453125 -0.9375 4.453125 -3.03125 C 4.453125 -4.25 4.265625 -5.078125 3.859375 -5.671875 C 3.53125 -6.109375 3.015625 -6.359375 2.421875 -6.359375 Z M 2.421875 -5.671875 C 3.25 -5.671875 3.671875 -4.828125 3.671875 -3.09375 C 3.671875 -1.28125 3.265625 -0.4375 2.40625 -0.4375 C 1.578125 -0.4375 1.171875 -1.3125 1.171875 -3.0625 C 1.171875 -4.828125 1.578125 -5.671875 2.421875 -5.671875 Z M 2.421875 -5.671875 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-3\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 1.9375 -2.921875 L 2.359375 -2.921875 C 3.203125 -2.921875 3.65625 -2.53125 3.65625 -1.765625 C 3.65625 -0.96875 3.171875 -0.484375 2.375 -0.484375 C 1.515625 -0.484375 1.109375 -0.921875 1.0625 -1.84375 L 0.28125 -1.84375 C 0.3125 -1.34375 0.40625 -1 0.546875 -0.71875 C 0.875 -0.109375 1.484375 0.203125 2.34375 0.203125 C 3.625 0.203125 4.453125 -0.578125 4.453125 -1.78125 C 4.453125 -2.59375 4.140625 -3.03125 3.390625 -3.296875 C 3.96875 -3.53125 4.265625 -3.96875 4.265625 -4.609375 C 4.265625 -5.703125 3.546875 -6.359375 2.359375 -6.359375 C 1.109375 -6.359375 0.4375 -5.65625 0.40625 -4.3125 L 1.1875 -4.3125 C 1.203125 -4.703125 1.234375 -4.921875 1.328125 -5.109375 C 1.5 -5.46875 1.890625 -5.6875 2.375 -5.6875 C 3.0625 -5.6875 3.46875 -5.265625 3.46875 -4.578125 C 3.46875 -4.140625 3.3125 -3.859375 2.96875 -3.71875 C 2.765625 -3.625 2.484375 -3.59375 1.9375 -3.578125 Z M 1.9375 -2.921875 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-4\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 2.875 -1.546875 L 2.875 0 L 3.65625 0 L 3.65625 -1.546875 L 4.578125 -1.546875 L 4.578125 -2.234375 L 3.65625 -2.234375 L 3.65625 -6.359375 L 3.078125 -6.359375 L 0.25 -2.359375 L 0.25 -1.546875 Z M 2.875 -2.234375 L 0.921875 -2.234375 L 2.875 -5.046875 Z M 2.875 -2.234375 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-5\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 2.28125 -4.53125 L 2.28125 0 L 3.046875 0 L 3.046875 -6.359375 L 2.546875 -6.359375 C 2.265625 -5.390625 2.09375 -5.25 0.890625 -5.09375 L 0.890625 -4.53125 Z M 2.28125 -4.53125 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-6\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 4.1875 -6.234375 L 0.96875 -6.234375 L 0.5 -2.84375 L 1.21875 -2.84375 C 1.578125 -3.265625 1.875 -3.421875 2.359375 -3.421875 C 3.203125 -3.421875 3.71875 -2.84375 3.71875 -1.921875 C 3.71875 -1.03125 3.203125 -0.484375 2.359375 -0.484375 C 1.6875 -0.484375 1.265625 -0.828125 1.078125 -1.53125 L 0.3125 -1.53125 C 0.40625 -1.015625 0.5 -0.78125 0.6875 -0.546875 C 1.03125 -0.078125 1.671875 0.203125 2.375 0.203125 C 3.640625 0.203125 4.515625 -0.71875 4.515625 -2.03125 C 4.515625 -3.265625 3.6875 -4.109375 2.5 -4.109375 C 2.0625 -4.109375 1.703125 -4 1.34375 -3.734375 L 1.59375 -5.46875 L 4.1875 -5.46875 Z M 4.1875 -6.234375 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-7\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 4.375 -4.703125 C 4.234375 -5.75 3.5625 -6.359375 2.609375 -6.359375 C 1.921875 -6.359375 1.3125 -6.03125 0.9375 -5.453125 C 0.546875 -4.828125 0.375 -4.0625 0.375 -2.90625 C 0.375 -1.828125 0.53125 -1.15625 0.90625 -0.59375 C 1.234375 -0.078125 1.78125 0.203125 2.46875 0.203125 C 3.65625 0.203125 4.515625 -0.6875 4.515625 -1.921875 C 4.515625 -3.09375 3.71875 -3.921875 2.609375 -3.921875 C 1.984375 -3.921875 1.5 -3.6875 1.171875 -3.234375 C 1.171875 -4.796875 1.671875 -5.671875 2.5625 -5.671875 C 3.109375 -5.671875 3.484375 -5.328125 3.609375 -4.703125 Z M 2.5 -3.234375 C 3.25 -3.234375 3.71875 -2.703125 3.71875 -1.859375 C 3.71875 -1.0625 3.1875 -0.484375 2.484375 -0.484375 C 1.765625 -0.484375 1.21875 -1.09375 1.21875 -1.90625 C 1.21875 -2.6875 1.734375 -3.234375 2.5 -3.234375 Z M 2.5 -3.234375 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph0-8\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 3.4375 -3.359375 C 4.09375 -3.75 4.296875 -4.0625 4.296875 -4.65625 C 4.296875 -5.671875 3.53125 -6.359375 2.421875 -6.359375 C 1.3125 -6.359375 0.546875 -5.671875 0.546875 -4.671875 C 0.546875 -4.0625 0.75 -3.75 1.390625 -3.359375 C 0.671875 -2.984375 0.328125 -2.46875 0.328125 -1.765625 C 0.328125 -0.609375 1.171875 0.203125 2.421875 0.203125 C 3.65625 0.203125 4.515625 -0.609375 4.515625 -1.765625 C 4.515625 -2.46875 4.15625 -2.984375 3.4375 -3.359375 Z M 2.421875 -5.671875 C 3.078125 -5.671875 3.5 -5.28125 3.5 -4.640625 C 3.5 -4.046875 3.0625 -3.65625 2.421875 -3.65625 C 1.765625 -3.65625 1.34375 -4.046875 1.34375 -4.65625 C 1.34375 -5.28125 1.765625 -5.671875 2.421875 -5.671875 Z M 2.421875 -3 C 3.1875 -3 3.71875 -2.484375 3.71875 -1.75 C 3.71875 -0.984375 3.203125 -0.484375 2.40625 -0.484375 C 1.640625 -0.484375 1.125 -1 1.125 -1.734375 C 1.125 -2.5 1.640625 -3 2.421875 -3 Z M 2.421875 -3 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-0\"&gt;\n&lt;path style=\"stroke:none;\" d=\"\"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-1\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 5.1875 -3.828125 C 5.140625 -4.390625 5.015625 -4.75 4.796875 -5.078125 C 4.40625 -5.609375 3.703125 -5.921875 2.90625 -5.921875 C 1.34375 -5.921875 0.34375 -4.703125 0.34375 -2.78125 C 0.34375 -0.921875 1.328125 0.25 2.890625 0.25 C 4.265625 0.25 5.140625 -0.578125 5.25 -1.984375 L 4.328125 -1.984375 C 4.171875 -1.0625 3.703125 -0.59375 2.921875 -0.59375 C 1.90625 -0.59375 1.296875 -1.421875 1.296875 -2.78125 C 1.296875 -4.21875 1.890625 -5.078125 2.890625 -5.078125 C 3.65625 -5.078125 4.140625 -4.625 4.25 -3.828125 Z M 5.1875 -3.828125 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-2\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 2.796875 -5.765625 L 1.84375 -5.765625 L 1.84375 -7.34375 L 0.9375 -7.34375 L 0.9375 -5.765625 L 0.15625 -5.765625 L 0.15625 -5.015625 L 0.9375 -5.015625 L 0.9375 -0.65625 C 0.9375 -0.0625 1.328125 0.25 2.046875 0.25 C 2.28125 0.25 2.484375 0.234375 2.796875 0.171875 L 2.796875 -0.59375 C 2.65625 -0.5625 2.546875 -0.546875 2.359375 -0.546875 C 1.953125 -0.546875 1.84375 -0.65625 1.84375 -1.0625 L 1.84375 -5.015625 L 2.796875 -5.015625 Z M 2.796875 -5.765625 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-3\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 4.265625 -5.765625 L 2.671875 -1.28125 L 1.203125 -5.765625 L 0.21875 -5.765625 L 2.171875 0.015625 L 1.8125 0.9375 C 1.65625 1.34375 1.46875 1.5 1.078125 1.5 C 0.921875 1.5 0.796875 1.46875 0.59375 1.4375 L 0.59375 2.25 C 0.78125 2.359375 0.96875 2.390625 1.203125 2.390625 C 1.5 2.390625 1.828125 2.296875 2.0625 2.125 C 2.359375 1.90625 2.515625 1.671875 2.6875 1.203125 L 5.265625 -5.765625 Z M 4.265625 -5.765625 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-4\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 5.640625 -2.578125 C 5.640625 -3.453125 5.578125 -3.984375 5.40625 -4.40625 C 5.03125 -5.359375 4.15625 -5.921875 3.078125 -5.921875 C 1.46875 -5.921875 0.4375 -4.703125 0.4375 -2.8125 C 0.4375 -0.90625 1.4375 0.25 3.0625 0.25 C 4.375 0.25 5.296875 -0.5 5.515625 -1.75 L 4.59375 -1.75 C 4.34375 -0.984375 3.828125 -0.59375 3.09375 -0.59375 C 2.515625 -0.59375 2.015625 -0.859375 1.703125 -1.34375 C 1.484375 -1.671875 1.40625 -2 1.390625 -2.578125 Z M 1.421875 -3.328125 C 1.5 -4.390625 2.140625 -5.078125 3.0625 -5.078125 C 4 -5.078125 4.65625 -4.359375 4.65625 -3.328125 Z M 1.421875 -3.328125 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-5\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 5.890625 -0.53125 C 5.78125 -0.515625 5.734375 -0.515625 5.6875 -0.515625 C 5.375 -0.515625 5.1875 -0.6875 5.1875 -0.96875 L 5.1875 -4.359375 C 5.1875 -5.375 4.4375 -5.921875 3.03125 -5.921875 C 2.171875 -5.921875 1.5 -5.6875 1.109375 -5.265625 C 0.84375 -4.96875 0.734375 -4.625 0.71875 -4.0625 L 1.640625 -4.0625 C 1.71875 -4.765625 2.140625 -5.078125 2.984375 -5.078125 C 3.828125 -5.078125 4.28125 -4.78125 4.28125 -4.21875 L 4.28125 -3.984375 C 4.265625 -3.59375 4.0625 -3.4375 3.328125 -3.34375 C 2.03125 -3.171875 1.828125 -3.140625 1.46875 -2.984375 C 0.796875 -2.703125 0.46875 -2.203125 0.46875 -1.453125 C 0.46875 -0.40625 1.1875 0.25 2.359375 0.25 C 3.078125 0.25 3.65625 0 4.3125 -0.59375 C 4.375 0 4.65625 0.25 5.265625 0.25 C 5.453125 0.25 5.578125 0.234375 5.890625 0.15625 Z M 4.28125 -1.8125 C 4.28125 -1.5 4.1875 -1.3125 3.921875 -1.0625 C 3.546875 -0.71875 3.09375 -0.546875 2.546875 -0.546875 C 1.84375 -0.546875 1.421875 -0.890625 1.421875 -1.46875 C 1.421875 -2.078125 1.8125 -2.390625 2.8125 -2.53125 C 3.78125 -2.65625 3.96875 -2.703125 4.28125 -2.84375 Z M 4.28125 -1.8125 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph1-6\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 0.765625 -5.765625 L 0.765625 0 L 1.6875 0 L 1.6875 -2.984375 C 1.6875 -4.375 2.265625 -5 3.53125 -4.96875 L 3.53125 -5.890625 C 3.375 -5.921875 3.28125 -5.921875 3.171875 -5.921875 C 2.578125 -5.921875 2.140625 -5.578125 1.609375 -4.71875 L 1.609375 -5.765625 Z M 0.765625 -5.765625 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph2-0\"&gt;\n&lt;path style=\"stroke:none;\" d=\"\"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph2-1\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M -8.015625 -0.765625 L 0 -0.765625 L 0 -1.6875 L -3.171875 -1.6875 C -4.359375 -1.6875 -5.125 -2.296875 -5.125 -3.25 C -5.125 -3.546875 -5.03125 -3.84375 -4.875 -4.0625 C -4.671875 -4.328125 -4.40625 -4.4375 -4 -4.4375 L 0 -4.4375 L 0 -5.34375 L -4.359375 -5.34375 C -5.328125 -5.34375 -5.921875 -4.65625 -5.921875 -3.53125 C -5.921875 -2.71875 -5.671875 -2.21875 -4.96875 -1.6875 L -8.015625 -1.6875 Z M -8.015625 -0.765625 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph2-2\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M 0 -6.09375 L -5.765625 -7.78125 L -5.765625 -6.75 L -1.28125 -5.609375 L -5.765625 -4.484375 L -5.765625 -3.359375 L -1.28125 -2.25 L -5.765625 -1.078125 L -5.765625 -0.0625 L 0 -1.734375 L 0 -2.765625 L -4.515625 -3.890625 L 0 -5.046875 Z M 0 -6.09375 \"/&gt;\n&lt;/symbol&gt;\n&lt;symbol overflow=\"visible\" id=\"glyph2-3\"&gt;\n&lt;path style=\"stroke:none;\" d=\"M -5.765625 -4.265625 L -1.28125 -2.671875 L -5.765625 -1.203125 L -5.765625 -0.21875 L 0.015625 -2.171875 L 0.9375 -1.8125 C 1.34375 -1.65625 1.5 -1.46875 1.5 -1.078125 C 1.5 -0.921875 1.46875 -0.796875 1.4375 -0.59375 L 2.25 -0.59375 C 2.359375 -0.78125 2.390625 -0.96875 2.390625 -1.203125 C 2.390625 -1.5 2.296875 -1.828125 2.125 -2.0625 C 1.90625 -2.359375 1.671875 -2.515625 1.203125 -2.6875 L -5.765625 -5.265625 Z M -5.765625 -4.265625 \"/&gt;\n&lt;/symbol&gt;\n&lt;/g&gt;\n&lt;clipPath id=\"clip1\"&gt;\n  &lt;path d=\"M 35.152344 5.480469 L 433.84375 5.480469 L 433.84375 469.851562 L 35.152344 469.851562 Z M 35.152344 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip2\"&gt;\n  &lt;path d=\"M 35.152344 408 L 433.84375 408 L 433.84375 410 L 35.152344 410 Z M 35.152344 408 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip3\"&gt;\n  &lt;path d=\"M 35.152344 276 L 433.84375 276 L 433.84375 278 L 35.152344 278 Z M 35.152344 276 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip4\"&gt;\n  &lt;path d=\"M 35.152344 145 L 433.84375 145 L 433.84375 146 L 35.152344 146 Z M 35.152344 145 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip5\"&gt;\n  &lt;path d=\"M 35.152344 13 L 433.84375 13 L 433.84375 14 L 35.152344 14 Z M 35.152344 13 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip6\"&gt;\n  &lt;path d=\"M 101 5.480469 L 103 5.480469 L 103 469.851562 L 101 469.851562 Z M 101 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip7\"&gt;\n  &lt;path d=\"M 171 5.480469 L 173 5.480469 L 173 469.851562 L 171 469.851562 Z M 171 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip8\"&gt;\n  &lt;path d=\"M 241 5.480469 L 242 5.480469 L 242 469.851562 L 241 469.851562 Z M 241 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip9\"&gt;\n  &lt;path d=\"M 310 5.480469 L 312 5.480469 L 312 469.851562 L 310 469.851562 Z M 310 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip10\"&gt;\n  &lt;path d=\"M 380 5.480469 L 382 5.480469 L 382 469.851562 L 380 469.851562 Z M 380 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip11\"&gt;\n  &lt;path d=\"M 35.152344 342 L 433.84375 342 L 433.84375 344 L 35.152344 344 Z M 35.152344 342 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip12\"&gt;\n  &lt;path d=\"M 35.152344 210 L 433.84375 210 L 433.84375 212 L 35.152344 212 Z M 35.152344 210 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip13\"&gt;\n  &lt;path d=\"M 35.152344 78 L 433.84375 78 L 433.84375 80 L 35.152344 80 Z M 35.152344 78 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip14\"&gt;\n  &lt;path d=\"M 66 5.480469 L 68 5.480469 L 68 469.851562 L 66 469.851562 Z M 66 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip15\"&gt;\n  &lt;path d=\"M 136 5.480469 L 138 5.480469 L 138 469.851562 L 136 469.851562 Z M 136 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip16\"&gt;\n  &lt;path d=\"M 206 5.480469 L 208 5.480469 L 208 469.851562 L 206 469.851562 Z M 206 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip17\"&gt;\n  &lt;path d=\"M 275 5.480469 L 277 5.480469 L 277 469.851562 L 275 469.851562 Z M 275 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip18\"&gt;\n  &lt;path d=\"M 345 5.480469 L 347 5.480469 L 347 469.851562 L 345 469.851562 Z M 345 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip19\"&gt;\n  &lt;path d=\"M 415 5.480469 L 417 5.480469 L 417 469.851562 L 415 469.851562 Z M 415 5.480469 \"/&gt;\n&lt;/clipPath&gt;\n&lt;clipPath id=\"clip20\"&gt;\n  &lt;path d=\"M 450.28125 203.203125 L 467.5625 203.203125 L 467.5625 289.601562 L 450.28125 289.601562 Z M 450.28125 203.203125 \"/&gt;\n&lt;/clipPath&gt;\n&lt;image id=\"image217\" width=\"1\" height=\"300\" xlink:href=\"data:image/png;base64,iVBORw0KGgoAAAANSUhEUgAAAAEAAAEsCAIAAACNMk6CAAAABmJLR0QA/wD/AP+gvaeTAAAAjklEQVQ4jaWTQRKAMAgDd3ik//R1HuOlji3SjNYjhZAEKNt+BCgQIRRIAQRSi2l5hdr79/q1fqBQVw+ErvyipqGH5eh7fNFJi1M+9bC8OT/lrTD9fNYwT/0FpvQzwVju0eObndsd/uUa7qX2NffqNE5w7j7dbCudpXbj2ehX0n3f0ot5pX9/1xkPj79W8p41NH/nSHk1sQAAAABJRU5ErkJggg==\"/&gt;\n&lt;pattern id=\"pattern0\" patternUnits=\"userSpaceOnUse\" width=\"1\" height=\"300\"  patternTransform=\"matrix(17.28,0,0,0.288,450.281644,203.20411)\"&gt;\n  &lt;use xlink:href=\"#image217\"/&gt;\n&lt;/pattern&gt;\n&lt;/defs&gt;\n&lt;g id=\"surface213\"&gt;\n&lt;rect x=\"0\" y=\"0\" width=\"504\" height=\"504\" style=\"fill:rgb(100%,100%,100%);fill-opacity:1;stroke:none;\"/&gt;\n&lt;rect x=\"0\" y=\"0\" width=\"504\" height=\"504\" style=\"fill:rgb(100%,100%,100%);fill-opacity:1;stroke:none;\"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 0 504 L 504 504 L 504 0 L 0 0 Z M 0 504 \"/&gt;\n&lt;g clip-path=\"url(#clip1)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\" stroke:none;fill-rule:nonzero;fill:rgb(92.156863%,92.156863%,92.156863%);fill-opacity:1;\" d=\"M 35.152344 469.847656 L 433.84375 469.847656 L 433.84375 5.476562 L 35.152344 5.476562 Z M 35.152344 469.847656 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip2)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 409.164062 L 433.84375 409.164062 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip3)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 277.242188 L 433.84375 277.242188 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip4)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 145.316406 L 433.84375 145.316406 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip5)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 13.394531 L 433.84375 13.394531 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip6)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 102.0625 469.847656 L 102.0625 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip7)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 171.765625 469.847656 L 171.765625 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip8)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 241.46875 469.847656 L 241.46875 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip9)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 311.167969 469.847656 L 311.167969 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip10)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 380.871094 469.847656 L 380.871094 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip11)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 343.203125 L 433.84375 343.203125 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip12)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 211.28125 L 433.84375 211.28125 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip13)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 35.152344 79.355469 L 433.84375 79.355469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip14)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 67.214844 469.847656 L 67.214844 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip15)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 136.914062 469.847656 L 136.914062 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip16)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 206.617188 469.847656 L 206.617188 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip17)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 276.316406 469.847656 L 276.316406 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip18)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 346.019531 469.847656 L 346.019531 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;g clip-path=\"url(#clip19)\" clip-rule=\"nonzero\"&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 415.722656 469.847656 L 415.722656 5.480469 \"/&gt;\n&lt;/g&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 224.472656 C 180.691406 227.078125 176.78125 227.078125 176.78125 224.472656 C 176.78125 221.867188 180.691406 221.867188 180.691406 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 198.085938 C 208.570312 200.695312 204.660156 200.695312 204.660156 198.085938 C 204.660156 195.480469 208.570312 195.480469 208.570312 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 211.28125 C 222.511719 213.886719 218.601562 213.886719 218.601562 211.28125 C 218.601562 208.671875 222.511719 208.671875 222.511719 211.28125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 264.050781 C 152.808594 266.65625 148.898438 266.65625 148.898438 264.050781 C 148.898438 261.441406 152.808594 261.441406 152.808594 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 250.855469 C 180.691406 253.464844 176.78125 253.464844 176.78125 250.855469 C 176.78125 248.25 180.691406 248.25 180.691406 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 277.242188 C 152.808594 279.847656 148.898438 279.847656 148.898438 277.242188 C 148.898438 274.636719 152.808594 274.636719 152.808594 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 237.664062 C 208.570312 240.269531 204.660156 240.269531 204.660156 237.664062 C 204.660156 235.058594 208.570312 235.058594 208.570312 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 250.855469 C 194.632812 253.464844 190.722656 253.464844 190.722656 250.855469 C 190.722656 248.25 194.632812 248.25 194.632812 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 277.242188 C 138.871094 279.847656 134.960938 279.847656 134.960938 277.242188 C 134.960938 274.636719 138.871094 274.636719 138.871094 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 277.242188 C 166.75 279.847656 162.839844 279.847656 162.839844 277.242188 C 162.839844 274.636719 166.75 274.636719 166.75 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 277.242188 C 166.75 279.847656 162.839844 279.847656 162.839844 277.242188 C 162.839844 274.636719 166.75 274.636719 166.75 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 277.242188 C 138.871094 279.847656 134.960938 279.847656 134.960938 277.242188 C 134.960938 274.636719 138.871094 274.636719 138.871094 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 290.433594 C 138.871094 293.039062 134.960938 293.039062 134.960938 290.433594 C 134.960938 287.828125 138.871094 287.828125 138.871094 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 277.242188 C 166.75 279.847656 162.839844 279.847656 162.839844 277.242188 C 162.839844 274.636719 166.75 274.636719 166.75 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 303.625 C 152.808594 306.230469 148.898438 306.230469 148.898438 303.625 C 148.898438 301.019531 152.808594 301.019531 152.808594 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 343.203125 C 124.929688 345.808594 121.019531 345.808594 121.019531 343.203125 C 121.019531 340.597656 124.929688 340.597656 124.929688 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 343.203125 C 124.929688 345.808594 121.019531 345.808594 121.019531 343.203125 C 121.019531 340.597656 124.929688 340.597656 124.929688 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 382.78125 C 97.046875 385.386719 93.140625 385.386719 93.140625 382.78125 C 93.140625 380.171875 97.046875 380.171875 97.046875 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 264.050781 C 152.808594 266.65625 148.898438 266.65625 148.898438 264.050781 C 148.898438 261.441406 152.808594 261.441406 152.808594 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 303.625 C 138.871094 306.230469 134.960938 306.230469 134.960938 303.625 C 134.960938 301.019531 138.871094 301.019531 138.871094 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 264.050781 C 152.808594 266.65625 148.898438 266.65625 148.898438 264.050781 C 148.898438 261.441406 152.808594 261.441406 152.808594 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 277.242188 C 138.871094 279.847656 134.960938 279.847656 134.960938 277.242188 C 134.960938 274.636719 138.871094 274.636719 138.871094 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 290.433594 C 138.871094 293.039062 134.960938 293.039062 134.960938 290.433594 C 134.960938 287.828125 138.871094 287.828125 138.871094 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 356.394531 C 124.929688 359 121.019531 359 121.019531 356.394531 C 121.019531 353.789062 124.929688 353.789062 124.929688 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 422.355469 C 83.109375 424.964844 79.199219 424.964844 79.199219 422.355469 C 79.199219 419.75 83.109375 419.75 83.109375 422.355469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 250.855469 C 194.632812 253.464844 190.722656 253.464844 190.722656 250.855469 C 190.722656 248.25 194.632812 248.25 194.632812 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 236.453125 211.28125 C 236.453125 213.886719 232.542969 213.886719 232.542969 211.28125 C 232.542969 208.671875 236.453125 208.671875 236.453125 211.28125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 224.472656 C 180.691406 227.078125 176.78125 227.078125 176.78125 224.472656 C 176.78125 221.867188 180.691406 221.867188 180.691406 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 264.050781 C 166.75 266.65625 162.839844 266.65625 162.839844 264.050781 C 162.839844 261.441406 166.75 261.441406 166.75 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 290.433594 C 180.691406 293.039062 176.78125 293.039062 176.78125 290.433594 C 176.78125 287.828125 180.691406 287.828125 180.691406 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 316.816406 C 152.808594 319.425781 148.898438 319.425781 148.898438 316.816406 C 148.898438 314.210938 152.808594 314.210938 152.808594 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 316.816406 C 152.808594 319.425781 148.898438 319.425781 148.898438 316.816406 C 148.898438 314.210938 152.808594 314.210938 152.808594 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 382.78125 C 83.109375 385.386719 79.199219 385.386719 79.199219 382.78125 C 79.199219 380.171875 83.109375 380.171875 83.109375 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 316.816406 C 138.871094 319.425781 134.960938 319.425781 134.960938 316.816406 C 134.960938 314.210938 138.871094 314.210938 138.871094 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 330.011719 C 138.871094 332.617188 134.960938 332.617188 134.960938 330.011719 C 134.960938 327.402344 138.871094 327.402344 138.871094 330.011719 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 303.625 C 152.808594 306.230469 148.898438 306.230469 148.898438 303.625 C 148.898438 301.019531 152.808594 301.019531 152.808594 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 303.625 C 152.808594 306.230469 148.898438 306.230469 148.898438 303.625 C 148.898438 301.019531 152.808594 301.019531 152.808594 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 356.394531 C 138.871094 359 134.960938 359 134.960938 356.394531 C 134.960938 353.789062 138.871094 353.789062 138.871094 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 369.585938 C 124.929688 372.195312 121.019531 372.195312 121.019531 369.585938 C 121.019531 366.980469 124.929688 366.980469 124.929688 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 356.394531 C 124.929688 359 121.019531 359 121.019531 356.394531 C 121.019531 353.789062 124.929688 353.789062 124.929688 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 356.394531 C 124.929688 359 121.019531 359 121.019531 356.394531 C 121.019531 353.789062 124.929688 353.789062 124.929688 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 55.226562 448.742188 C 55.226562 451.347656 51.316406 451.347656 51.316406 448.742188 C 51.316406 446.136719 55.226562 446.136719 55.226562 448.742188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 382.78125 C 83.109375 385.386719 79.199219 385.386719 79.199219 382.78125 C 79.199219 380.171875 83.109375 380.171875 83.109375 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 55.226562 448.742188 C 55.226562 451.347656 51.316406 451.347656 51.316406 448.742188 C 51.316406 446.136719 55.226562 446.136719 55.226562 448.742188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 395.972656 C 83.109375 398.578125 79.199219 398.578125 79.199219 395.972656 C 79.199219 393.367188 83.109375 393.367188 83.109375 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 369.585938 C 110.988281 372.195312 107.078125 372.195312 107.078125 369.585938 C 107.078125 366.980469 110.988281 366.980469 110.988281 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 395.972656 C 97.046875 398.578125 93.140625 398.578125 93.140625 395.972656 C 93.140625 393.367188 97.046875 393.367188 97.046875 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 55.226562 448.742188 C 55.226562 451.347656 51.316406 451.347656 51.316406 448.742188 C 51.316406 446.136719 55.226562 446.136719 55.226562 448.742188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 395.972656 C 97.046875 398.578125 93.140625 398.578125 93.140625 395.972656 C 93.140625 393.367188 97.046875 393.367188 97.046875 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 55.226562 448.742188 C 55.226562 451.347656 51.316406 451.347656 51.316406 448.742188 C 51.316406 446.136719 55.226562 446.136719 55.226562 448.742188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 395.972656 C 83.109375 398.578125 79.199219 398.578125 79.199219 395.972656 C 79.199219 393.367188 83.109375 393.367188 83.109375 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 382.78125 C 83.109375 385.386719 79.199219 385.386719 79.199219 382.78125 C 79.199219 380.171875 83.109375 380.171875 83.109375 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 382.78125 C 83.109375 385.386719 79.199219 385.386719 79.199219 382.78125 C 79.199219 380.171875 83.109375 380.171875 83.109375 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 369.585938 C 97.046875 372.195312 93.140625 372.195312 93.140625 369.585938 C 93.140625 366.980469 97.046875 366.980469 97.046875 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 356.394531 C 138.871094 359 134.960938 359 134.960938 356.394531 C 134.960938 353.789062 138.871094 353.789062 138.871094 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 356.394531 C 110.988281 359 107.078125 359 107.078125 356.394531 C 107.078125 353.789062 110.988281 353.789062 110.988281 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 356.394531 C 110.988281 359 107.078125 359 107.078125 356.394531 C 107.078125 353.789062 110.988281 353.789062 110.988281 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 395.972656 C 110.988281 398.578125 107.078125 398.578125 107.078125 395.972656 C 107.078125 393.367188 110.988281 393.367188 110.988281 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 395.972656 C 110.988281 398.578125 107.078125 398.578125 107.078125 395.972656 C 107.078125 393.367188 110.988281 393.367188 110.988281 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 277.242188 C 180.691406 279.847656 176.78125 279.847656 176.78125 277.242188 C 176.78125 274.636719 180.691406 274.636719 180.691406 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 264.050781 C 166.75 266.65625 162.839844 266.65625 162.839844 264.050781 C 162.839844 261.441406 166.75 261.441406 166.75 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 290.433594 C 152.808594 293.039062 148.898438 293.039062 148.898438 290.433594 C 148.898438 287.828125 152.808594 287.828125 152.808594 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 330.011719 C 138.871094 332.617188 134.960938 332.617188 134.960938 330.011719 C 134.960938 327.402344 138.871094 327.402344 138.871094 330.011719 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 316.816406 C 138.871094 319.425781 134.960938 319.425781 134.960938 316.816406 C 134.960938 314.210938 138.871094 314.210938 138.871094 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 303.625 C 138.871094 306.230469 134.960938 306.230469 134.960938 303.625 C 134.960938 301.019531 138.871094 301.019531 138.871094 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 316.816406 C 138.871094 319.425781 134.960938 319.425781 134.960938 316.816406 C 134.960938 314.210938 138.871094 314.210938 138.871094 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 343.203125 C 124.929688 345.808594 121.019531 345.808594 121.019531 343.203125 C 121.019531 340.597656 124.929688 340.597656 124.929688 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 320.09375 171.703125 C 320.09375 174.308594 316.183594 174.308594 316.183594 171.703125 C 316.183594 169.097656 320.09375 169.097656 320.09375 171.703125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 264.332031 184.894531 C 264.332031 187.5 260.421875 187.5 260.421875 184.894531 C 260.421875 182.289062 264.332031 182.289062 264.332031 184.894531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 278.273438 184.894531 C 278.273438 187.5 274.363281 187.5 274.363281 184.894531 C 274.363281 182.289062 278.273438 182.289062 278.273438 184.894531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 250.390625 224.472656 C 250.390625 227.078125 246.484375 227.078125 246.484375 224.472656 C 246.484375 221.867188 250.390625 221.867188 250.390625 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 264.332031 184.894531 C 264.332031 187.5 260.421875 187.5 260.421875 184.894531 C 260.421875 182.289062 264.332031 182.289062 264.332031 184.894531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 292.214844 158.511719 C 292.214844 161.117188 288.304688 161.117188 288.304688 158.511719 C 288.304688 155.902344 292.214844 155.902344 292.214844 158.511719 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 278.273438 132.125 C 278.273438 134.730469 274.363281 134.730469 274.363281 132.125 C 274.363281 129.519531 278.273438 129.519531 278.273438 132.125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 264.332031 132.125 C 264.332031 134.730469 260.421875 134.730469 260.421875 132.125 C 260.421875 129.519531 264.332031 129.519531 264.332031 132.125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 250.855469 C 180.691406 253.464844 176.78125 253.464844 176.78125 250.855469 C 176.78125 248.25 180.691406 248.25 180.691406 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 211.28125 C 222.511719 213.886719 218.601562 213.886719 218.601562 211.28125 C 218.601562 208.671875 222.511719 208.671875 222.511719 211.28125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 198.085938 C 222.511719 200.695312 218.601562 200.695312 218.601562 198.085938 C 218.601562 195.480469 222.511719 195.480469 222.511719 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 237.664062 C 194.632812 240.269531 190.722656 240.269531 190.722656 237.664062 C 190.722656 235.058594 194.632812 235.058594 194.632812 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 224.472656 C 194.632812 227.078125 190.722656 227.078125 190.722656 224.472656 C 190.722656 221.867188 194.632812 221.867188 194.632812 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 237.664062 C 208.570312 240.269531 204.660156 240.269531 204.660156 237.664062 C 204.660156 235.058594 208.570312 235.058594 208.570312 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 250.855469 C 208.570312 253.464844 204.660156 253.464844 204.660156 250.855469 C 204.660156 248.25 208.570312 248.25 208.570312 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 290.433594 C 152.808594 293.039062 148.898438 293.039062 148.898438 290.433594 C 148.898438 287.828125 152.808594 287.828125 152.808594 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 316.816406 C 166.75 319.425781 162.839844 319.425781 162.839844 316.816406 C 162.839844 314.210938 166.75 314.210938 166.75 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 356.394531 C 138.871094 359 134.960938 359 134.960938 356.394531 C 134.960938 353.789062 138.871094 353.789062 138.871094 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 343.203125 C 138.871094 345.808594 134.960938 345.808594 134.960938 343.203125 C 134.960938 340.597656 138.871094 340.597656 138.871094 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 55.226562 448.742188 C 55.226562 451.347656 51.316406 451.347656 51.316406 448.742188 C 51.316406 446.136719 55.226562 446.136719 55.226562 448.742188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 356.394531 C 124.929688 359 121.019531 359 121.019531 356.394531 C 121.019531 353.789062 124.929688 353.789062 124.929688 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 369.585938 C 110.988281 372.195312 107.078125 372.195312 107.078125 369.585938 C 107.078125 366.980469 110.988281 366.980469 110.988281 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 422.355469 C 83.109375 424.964844 79.199219 424.964844 79.199219 422.355469 C 79.199219 419.75 83.109375 419.75 83.109375 422.355469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 369.585938 C 97.046875 372.195312 93.140625 372.195312 93.140625 369.585938 C 93.140625 366.980469 97.046875 366.980469 97.046875 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 369.585938 C 97.046875 372.195312 93.140625 372.195312 93.140625 369.585938 C 93.140625 366.980469 97.046875 366.980469 97.046875 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 382.78125 C 83.109375 385.386719 79.199219 385.386719 79.199219 382.78125 C 79.199219 380.171875 83.109375 380.171875 83.109375 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 395.972656 C 83.109375 398.578125 79.199219 398.578125 79.199219 395.972656 C 79.199219 393.367188 83.109375 393.367188 83.109375 395.972656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 369.585938 C 97.046875 372.195312 93.140625 372.195312 93.140625 369.585938 C 93.140625 366.980469 97.046875 366.980469 97.046875 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 356.394531 C 110.988281 359 107.078125 359 107.078125 356.394531 C 107.078125 353.789062 110.988281 353.789062 110.988281 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 356.394531 C 110.988281 359 107.078125 359 107.078125 356.394531 C 107.078125 353.789062 110.988281 353.789062 110.988281 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 382.78125 C 110.988281 385.386719 107.078125 385.386719 107.078125 382.78125 C 107.078125 380.171875 110.988281 380.171875 110.988281 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 250.855469 C 194.632812 253.464844 190.722656 253.464844 190.722656 250.855469 C 190.722656 248.25 194.632812 248.25 194.632812 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 250.390625 198.085938 C 250.390625 200.695312 246.484375 200.695312 246.484375 198.085938 C 246.484375 195.480469 250.390625 195.480469 250.390625 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 250.390625 184.894531 C 250.390625 187.5 246.484375 187.5 246.484375 184.894531 C 246.484375 182.289062 250.390625 182.289062 250.390625 184.894531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 250.855469 C 194.632812 253.464844 190.722656 253.464844 190.722656 250.855469 C 190.722656 248.25 194.632812 248.25 194.632812 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 277.242188 C 194.632812 279.847656 190.722656 279.847656 190.722656 277.242188 C 190.722656 274.636719 194.632812 274.636719 194.632812 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 277.242188 C 194.632812 279.847656 190.722656 279.847656 190.722656 277.242188 C 190.722656 274.636719 194.632812 274.636719 194.632812 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 382.78125 C 138.871094 385.386719 134.960938 385.386719 134.960938 382.78125 C 134.960938 380.171875 138.871094 380.171875 138.871094 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 343.203125 C 124.929688 345.808594 121.019531 345.808594 121.019531 343.203125 C 121.019531 340.597656 124.929688 340.597656 124.929688 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 97.046875 369.585938 C 97.046875 372.195312 93.140625 372.195312 93.140625 369.585938 C 93.140625 366.980469 97.046875 366.980469 97.046875 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 264.050781 C 152.808594 266.65625 148.898438 266.65625 148.898438 264.050781 C 148.898438 261.441406 152.808594 261.441406 152.808594 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 250.855469 C 166.75 253.464844 162.839844 253.464844 162.839844 250.855469 C 162.839844 248.25 166.75 248.25 166.75 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 237.664062 C 180.691406 240.269531 176.78125 240.269531 176.78125 237.664062 C 176.78125 235.058594 180.691406 235.058594 180.691406 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 277.242188 C 152.808594 279.847656 148.898438 279.847656 148.898438 277.242188 C 148.898438 274.636719 152.808594 274.636719 152.808594 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 277.242188 C 180.691406 279.847656 176.78125 279.847656 176.78125 277.242188 C 176.78125 274.636719 180.691406 274.636719 180.691406 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 290.433594 C 180.691406 293.039062 176.78125 293.039062 176.78125 290.433594 C 176.78125 287.828125 180.691406 287.828125 180.691406 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 250.855469 C 208.570312 253.464844 204.660156 253.464844 204.660156 250.855469 C 204.660156 248.25 208.570312 248.25 208.570312 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 277.242188 C 194.632812 279.847656 190.722656 279.847656 190.722656 277.242188 C 190.722656 274.636719 194.632812 274.636719 194.632812 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 264.050781 C 208.570312 266.65625 204.660156 266.65625 204.660156 264.050781 C 204.660156 261.441406 208.570312 261.441406 208.570312 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 303.625 C 180.691406 306.230469 176.78125 306.230469 176.78125 303.625 C 176.78125 301.019531 180.691406 301.019531 180.691406 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 264.050781 C 222.511719 266.65625 218.601562 266.65625 218.601562 264.050781 C 218.601562 261.441406 222.511719 261.441406 222.511719 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 277.242188 C 208.570312 279.847656 204.660156 279.847656 204.660156 277.242188 C 204.660156 274.636719 208.570312 274.636719 208.570312 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 250.855469 C 208.570312 253.464844 204.660156 253.464844 204.660156 250.855469 C 204.660156 248.25 208.570312 248.25 208.570312 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 277.242188 C 194.632812 279.847656 190.722656 279.847656 190.722656 277.242188 C 190.722656 274.636719 194.632812 274.636719 194.632812 277.242188 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 250.855469 C 208.570312 253.464844 204.660156 253.464844 204.660156 250.855469 C 204.660156 248.25 208.570312 248.25 208.570312 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 343.203125 C 138.871094 345.808594 134.960938 345.808594 134.960938 343.203125 C 134.960938 340.597656 138.871094 340.597656 138.871094 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 343.203125 C 152.808594 345.808594 148.898438 345.808594 148.898438 343.203125 C 148.898438 340.597656 152.808594 340.597656 152.808594 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 356.394531 C 138.871094 359 134.960938 359 134.960938 356.394531 C 134.960938 353.789062 138.871094 353.789062 138.871094 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 382.78125 C 138.871094 385.386719 134.960938 385.386719 134.960938 382.78125 C 134.960938 380.171875 138.871094 380.171875 138.871094 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 343.203125 C 152.808594 345.808594 148.898438 345.808594 148.898438 343.203125 C 148.898438 340.597656 152.808594 340.597656 152.808594 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 124.929688 382.78125 C 124.929688 385.386719 121.019531 385.386719 121.019531 382.78125 C 121.019531 380.171875 124.929688 380.171875 124.929688 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 250.855469 C 222.511719 253.464844 218.601562 253.464844 218.601562 250.855469 C 218.601562 248.25 222.511719 248.25 222.511719 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 198.085938 C 222.511719 200.695312 218.601562 200.695312 218.601562 198.085938 C 218.601562 195.480469 222.511719 195.480469 222.511719 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 198.085938 C 222.511719 200.695312 218.601562 200.695312 218.601562 198.085938 C 218.601562 195.480469 222.511719 195.480469 222.511719 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 237.664062 C 194.632812 240.269531 190.722656 240.269531 190.722656 237.664062 C 190.722656 235.058594 194.632812 235.058594 194.632812 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 250.855469 C 222.511719 253.464844 218.601562 253.464844 218.601562 250.855469 C 218.601562 248.25 222.511719 248.25 222.511719 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 198.085938 C 222.511719 200.695312 218.601562 200.695312 218.601562 198.085938 C 218.601562 195.480469 222.511719 195.480469 222.511719 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 236.453125 198.085938 C 236.453125 200.695312 232.542969 200.695312 232.542969 198.085938 C 232.542969 195.480469 236.453125 195.480469 236.453125 198.085938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 250.855469 C 180.691406 253.464844 176.78125 253.464844 176.78125 250.855469 C 176.78125 248.25 180.691406 248.25 180.691406 250.855469 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 264.332031 211.28125 C 264.332031 213.886719 260.421875 213.886719 260.421875 211.28125 C 260.421875 208.671875 264.332031 208.671875 264.332031 211.28125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 264.332031 171.703125 C 264.332031 174.308594 260.421875 174.308594 260.421875 171.703125 C 260.421875 169.097656 264.332031 169.097656 264.332031 171.703125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 292.214844 145.316406 C 292.214844 147.925781 288.304688 147.925781 288.304688 145.316406 C 288.304688 142.710938 292.214844 142.710938 292.214844 145.316406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 320.09375 118.933594 C 320.09375 121.539062 316.183594 121.539062 316.183594 118.933594 C 316.183594 116.328125 320.09375 116.328125 320.09375 118.933594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 292.214844 145.316406 C 292.214844 147.925781 288.304688 147.925781 288.304688 145.316406 C 288.304688 142.710938 292.214844 142.710938 292.214844 145.316406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 83.109375 409.164062 C 83.109375 411.769531 79.199219 411.769531 79.199219 409.164062 C 79.199219 406.558594 83.109375 406.558594 83.109375 409.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 110.988281 369.585938 C 110.988281 372.195312 107.078125 372.195312 107.078125 369.585938 C 107.078125 366.980469 110.988281 366.980469 110.988281 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 343.203125 C 138.871094 345.808594 134.960938 345.808594 134.960938 343.203125 C 134.960938 340.597656 138.871094 340.597656 138.871094 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 343.203125 C 152.808594 345.808594 148.898438 345.808594 148.898438 343.203125 C 148.898438 340.597656 152.808594 340.597656 152.808594 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 316.816406 C 166.75 319.425781 162.839844 319.425781 162.839844 316.816406 C 162.839844 314.210938 166.75 314.210938 166.75 316.816406 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 382.78125 C 138.871094 385.386719 134.960938 385.386719 134.960938 382.78125 C 134.960938 380.171875 138.871094 380.171875 138.871094 382.78125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 356.394531 C 138.871094 359 134.960938 359 134.960938 356.394531 C 134.960938 353.789062 138.871094 353.789062 138.871094 356.394531 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 138.871094 369.585938 C 138.871094 372.195312 134.960938 372.195312 134.960938 369.585938 C 134.960938 366.980469 138.871094 366.980469 138.871094 369.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 343.203125 C 152.808594 345.808594 148.898438 345.808594 148.898438 343.203125 C 148.898438 340.597656 152.808594 340.597656 152.808594 343.203125 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 236.453125 224.472656 C 236.453125 227.078125 232.542969 227.078125 232.542969 224.472656 C 232.542969 221.867188 236.453125 221.867188 236.453125 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 389.796875 26.585938 C 389.796875 29.195312 385.886719 29.195312 385.886719 26.585938 C 385.886719 23.980469 389.796875 23.980469 389.796875 26.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 236.453125 224.472656 C 236.453125 227.078125 232.542969 227.078125 232.542969 224.472656 C 232.542969 221.867188 236.453125 221.867188 236.453125 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 303.625 C 152.808594 306.230469 148.898438 306.230469 148.898438 303.625 C 148.898438 301.019531 152.808594 301.019531 152.808594 303.625 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 290.433594 C 166.75 293.039062 162.839844 293.039062 162.839844 290.433594 C 162.839844 287.828125 166.75 287.828125 166.75 290.433594 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 417.675781 26.585938 C 417.675781 29.195312 413.765625 29.195312 413.765625 26.585938 C 413.765625 23.980469 417.675781 23.980469 417.675781 26.585938 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 334.035156 66.164062 C 334.035156 68.769531 330.125 68.769531 330.125 66.164062 C 330.125 63.558594 334.035156 63.558594 334.035156 66.164062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 264.050781 C 194.632812 266.65625 190.722656 266.65625 190.722656 264.050781 C 190.722656 261.441406 194.632812 261.441406 194.632812 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 237.664062 C 208.570312 240.269531 204.660156 240.269531 204.660156 237.664062 C 204.660156 235.058594 208.570312 235.058594 208.570312 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 208.570312 224.472656 C 208.570312 227.078125 204.660156 227.078125 204.660156 224.472656 C 204.660156 221.867188 208.570312 221.867188 208.570312 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 224.472656 C 180.691406 227.078125 176.78125 227.078125 176.78125 224.472656 C 176.78125 221.867188 180.691406 221.867188 180.691406 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 194.632812 237.664062 C 194.632812 240.269531 190.722656 240.269531 190.722656 237.664062 C 190.722656 235.058594 194.632812 235.058594 194.632812 237.664062 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 222.511719 224.472656 C 222.511719 227.078125 218.601562 227.078125 218.601562 224.472656 C 218.601562 221.867188 222.511719 221.867188 222.511719 224.472656 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 152.808594 264.050781 C 152.808594 266.65625 148.898438 266.65625 148.898438 264.050781 C 148.898438 261.441406 152.808594 261.441406 152.808594 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 180.691406 264.050781 C 180.691406 266.65625 176.78125 266.65625 176.78125 264.050781 C 176.78125 261.441406 180.691406 261.441406 180.691406 264.050781 \"/&gt;\n&lt;path style=\"fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 166.75 264.050781 C 166.75 266.65625 162.839844 266.65625 162.839844 264.050781 C 162.839844 261.441406 166.75 261.441406 166.75 264.050781 \"/&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"20.21875\" y=\"346.118164\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"25.21875\" y=\"346.118164\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-3\" x=\"20.21875\" y=\"214.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"25.21875\" y=\"214.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-4\" x=\"20.21875\" y=\"82.270508\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"25.21875\" y=\"82.270508\"/&gt;\n&lt;/g&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 32.410156 343.203125 L 35.152344 343.203125 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 32.410156 211.28125 L 35.152344 211.28125 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 32.410156 79.355469 L 35.152344 79.355469 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 67.214844 472.589844 L 67.214844 469.847656 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 136.914062 472.589844 L 136.914062 469.847656 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 206.617188 472.589844 L 206.617188 469.847656 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 276.316406 472.589844 L 276.316406 469.847656 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 346.019531 472.589844 L 346.019531 469.847656 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 415.722656 472.589844 L 415.722656 469.847656 \"/&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-5\" x=\"62.214844\" y=\"481.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"67.214844\" y=\"481.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-5\" x=\"131.914062\" y=\"481.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-6\" x=\"136.914062\" y=\"481.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"201.617188\" y=\"481.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"206.617188\" y=\"481.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"271.316406\" y=\"481.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-6\" x=\"276.316406\" y=\"481.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-3\" x=\"341.019531\" y=\"481.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"346.019531\" y=\"481.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-3\" x=\"410.722656\" y=\"481.196289\"/&gt;\n  &lt;use xlink:href=\"#glyph0-6\" x=\"415.722656\" y=\"481.196289\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph1-1\" x=\"227.496094\" y=\"494.538086\"/&gt;\n  &lt;use xlink:href=\"#glyph1-2\" x=\"233.496094\" y=\"494.538086\"/&gt;\n  &lt;use xlink:href=\"#glyph1-3\" x=\"236.496094\" y=\"494.538086\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph2-1\" x=\"13.499023\" y=\"247.164062\"/&gt;\n  &lt;use xlink:href=\"#glyph2-2\" x=\"13.499023\" y=\"241.164062\"/&gt;\n  &lt;use xlink:href=\"#glyph2-3\" x=\"13.499023\" y=\"233.164062\"/&gt;\n&lt;/g&gt;\n&lt;path style=\" stroke:none;fill-rule:nonzero;fill:rgb(100%,100%,100%);fill-opacity:1;\" d=\"M 444.800781 295.082031 L 498.519531 295.082031 L 498.519531 180.242188 L 444.800781 180.242188 Z M 444.800781 295.082031 \"/&gt;\n&lt;g clip-path=\"url(#clip20)\" clip-rule=\"nonzero\"&gt;\n&lt;rect x=\"0\" y=\"0\" width=\"504\" height=\"504\" style=\"fill:url(#pattern0);stroke:none;\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"473.042969\" y=\"282.805664\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"478.042969\" y=\"282.805664\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"483.042969\" y=\"282.805664\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"488.042969\" y=\"282.805664\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"473.042969\" y=\"263.672852\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"478.042969\" y=\"263.672852\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"483.042969\" y=\"263.672852\"/&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"488.042969\" y=\"263.672852\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"473.042969\" y=\"244.536133\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"478.042969\" y=\"244.536133\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"483.042969\" y=\"244.536133\"/&gt;\n  &lt;use xlink:href=\"#glyph0-4\" x=\"488.042969\" y=\"244.536133\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"473.042969\" y=\"225.399414\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"478.042969\" y=\"225.399414\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"483.042969\" y=\"225.399414\"/&gt;\n  &lt;use xlink:href=\"#glyph0-7\" x=\"488.042969\" y=\"225.399414\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph0-1\" x=\"473.042969\" y=\"206.262695\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"478.042969\" y=\"206.262695\"/&gt;\n  &lt;use xlink:href=\"#glyph0-2\" x=\"483.042969\" y=\"206.262695\"/&gt;\n  &lt;use xlink:href=\"#glyph0-8\" x=\"488.042969\" y=\"206.262695\"/&gt;\n&lt;/g&gt;\n&lt;g style=\"fill:rgb(0%,0%,0%);fill-opacity:1;\"&gt;\n  &lt;use xlink:href=\"#glyph1-3\" x=\"450.28125\" y=\"195.245117\"/&gt;\n  &lt;use xlink:href=\"#glyph1-4\" x=\"455.28125\" y=\"195.245117\"/&gt;\n  &lt;use xlink:href=\"#glyph1-5\" x=\"461.28125\" y=\"195.245117\"/&gt;\n  &lt;use xlink:href=\"#glyph1-6\" x=\"467.28125\" y=\"195.245117\"/&gt;\n&lt;/g&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 450.28125 279.890625 L 453.738281 279.890625 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 450.28125 260.757812 L 453.738281 260.757812 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 450.28125 241.621094 L 453.738281 241.621094 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 450.28125 222.484375 L 453.738281 222.484375 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 450.28125 203.347656 L 453.738281 203.347656 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 464.105469 279.890625 L 467.5625 279.890625 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 464.105469 260.757812 L 467.5625 260.757812 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 464.105469 241.621094 L 467.5625 241.621094 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 464.105469 222.484375 L 467.5625 222.484375 \"/&gt;\n&lt;path style=\"fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;\" d=\"M 464.105469 203.347656 L 467.5625 203.347656 \"/&gt;\n&lt;/g&gt;\n&lt;/svg&gt;\n"

```
:::

Use ggplot2 as a Clojure namespace:

::: {.sourceClojure}
```clojure
(r/require-r '[ggplot2 :as gg])
```
:::

::: {.printedClojure}
```clojure
nil

```
:::

The `mpg` dataset as a Clojure (tech.ml.dataset) dataset:

::: {.sourceClojure}
```clojure
toydata.ggplot/mpg
```
:::

::: {.clay-dataset}
_unnamed [234 11]:

| :manufacturer |     :model | :displ | :year | :cyl |     :trans | :drv | :cty | :hwy | :fl |     :class |
|---------------|------------|-------:|------:|-----:|------------|------|-----:|-----:|-----|------------|
|          audi |         a4 |    1.8 |  1999 |    4 |   auto(l5) |    f |   18 |   29 |   p |    compact |
|          audi |         a4 |    1.8 |  1999 |    4 | manual(m5) |    f |   21 |   29 |   p |    compact |
|          audi |         a4 |    2.0 |  2008 |    4 | manual(m6) |    f |   20 |   31 |   p |    compact |
|          audi |         a4 |    2.0 |  2008 |    4 |   auto(av) |    f |   21 |   30 |   p |    compact |
|          audi |         a4 |    2.8 |  1999 |    6 |   auto(l5) |    f |   16 |   26 |   p |    compact |
|          audi |         a4 |    2.8 |  1999 |    6 | manual(m5) |    f |   18 |   26 |   p |    compact |
|          audi |         a4 |    3.1 |  2008 |    6 |   auto(av) |    f |   18 |   27 |   p |    compact |
|          audi | a4 quattro |    1.8 |  1999 |    4 | manual(m5) |    4 |   18 |   26 |   p |    compact |
|          audi | a4 quattro |    1.8 |  1999 |    4 |   auto(l5) |    4 |   16 |   25 |   p |    compact |
|          audi | a4 quattro |    2.0 |  2008 |    4 | manual(m6) |    4 |   20 |   28 |   p |    compact |
|           ... |        ... |    ... |   ... |  ... |        ... |  ... |  ... |  ... | ... |        ... |
|    volkswagen | new beetle |    2.0 |  1999 |    4 | manual(m5) |    f |   21 |   29 |   r | subcompact |
|    volkswagen | new beetle |    2.0 |  1999 |    4 |   auto(l4) |    f |   19 |   26 |   r | subcompact |
|    volkswagen | new beetle |    2.5 |  2008 |    5 | manual(m5) |    f |   20 |   28 |   r | subcompact |
|    volkswagen | new beetle |    2.5 |  2008 |    5 |   auto(s6) |    f |   20 |   29 |   r | subcompact |
|    volkswagen |     passat |    1.8 |  1999 |    4 | manual(m5) |    f |   21 |   29 |   p |    midsize |
|    volkswagen |     passat |    1.8 |  1999 |    4 |   auto(l5) |    f |   18 |   29 |   p |    midsize |
|    volkswagen |     passat |    2.0 |  2008 |    4 |   auto(s6) |    f |   19 |   28 |   p |    midsize |
|    volkswagen |     passat |    2.0 |  2008 |    4 | manual(m6) |    f |   21 |   29 |   p |    midsize |
|    volkswagen |     passat |    2.8 |  1999 |    6 |   auto(l5) |    f |   16 |   26 |   p |    midsize |
|    volkswagen |     passat |    2.8 |  1999 |    6 | manual(m5) |    f |   18 |   26 |   p |    midsize |
|    volkswagen |     passat |    3.6 |  2008 |    6 |   auto(s6) |    f |   17 |   26 |   p |    midsize |

:::

Evaluating R code generated from the Clojure namespace,
with Clojure data.

::: {.sourceClojure}
```clojure
(-> toydata.ggplot/mpg
    (gg/ggplot (gg/aes :x 'cty
                       :y 'hwy
                       :color 'year))
    (r/r+ (gg/geom_point))
    r
    plotting/plot->buffered-image)
```
:::

::: {.clay-image}

```{=html}
<img src="study_session_20240225_files/1.png" />
```

:::

Representing a ggplot plot as Clojure data:

::: {.sourceClojure}
```clojure
(-> toydata.ggplot/mpg
    (gg/ggplot (gg/aes :x 'cty
                       :y 'hwy
                       :color 'year))
    (r/r+ (gg/geom_point))
    r
    representing/ggplot->clj
    ;; (dissoc the dataset itself
    ;; to keep the structure small)
    (dissoc :data))
```
:::

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor [[$ .MEM x1a549472598b4a60]],
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy], :colour [~ year]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["year"]}}

```
:::

Using log scale
(keeping axis labels,
changing the spacing):

::: {.sourceClojure}
```clojure
(-> toydata.ggplot/mpg
    (gg/ggplot (gg/aes :x 'cty
                       :y 'hwy
                       :color 'year))
    (r/r+ (gg/geom_point :size 5)
          (gg/scale_y_log10))
    r
    plotting/plot->svg
    kind/html)
```
:::

```{=html}
<?xml version="1.0" encoding="UTF-8"?>
<svg xmlns="http://www.w3.org/2000/svg" xmlns:xlink="http://www.w3.org/1999/xlink" width="504pt" height="504pt" viewBox="0 0 504 504" version="1.1">
<defs>
<g>
<symbol overflow="visible" id="glyph0-0">
<path style="stroke:none;" d=""/>
</symbol>
<symbol overflow="visible" id="glyph0-1">
<path style="stroke:none;" d="M 4.453125 -0.765625 L 1.171875 -0.765625 C 1.25 -1.296875 1.53125 -1.625 2.296875 -2.09375 L 3.171875 -2.59375 C 4.046875 -3.0625 4.5 -3.71875 4.5 -4.5 C 4.5 -5.03125 4.28125 -5.53125 3.921875 -5.875 C 3.546875 -6.203125 3.09375 -6.359375 2.5 -6.359375 C 1.703125 -6.359375 1.125 -6.078125 0.78125 -5.53125 C 0.546875 -5.203125 0.453125 -4.796875 0.4375 -4.15625 L 1.21875 -4.15625 C 1.234375 -4.59375 1.296875 -4.84375 1.40625 -5.0625 C 1.59375 -5.4375 2 -5.6875 2.46875 -5.6875 C 3.171875 -5.6875 3.703125 -5.171875 3.703125 -4.484375 C 3.703125 -3.96875 3.40625 -3.53125 2.859375 -3.21875 L 2.046875 -2.75 C 0.75 -2 0.375 -1.40625 0.296875 -0.015625 L 4.453125 -0.015625 Z M 4.453125 -0.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph0-2">
<path style="stroke:none;" d="M 2.421875 -6.359375 C 1.84375 -6.359375 1.3125 -6.09375 0.984375 -5.671875 C 0.578125 -5.109375 0.375 -4.25 0.375 -3.078125 C 0.375 -0.9375 1.078125 0.203125 2.421875 0.203125 C 3.734375 0.203125 4.453125 -0.9375 4.453125 -3.03125 C 4.453125 -4.25 4.265625 -5.078125 3.859375 -5.671875 C 3.53125 -6.109375 3.015625 -6.359375 2.421875 -6.359375 Z M 2.421875 -5.671875 C 3.25 -5.671875 3.671875 -4.828125 3.671875 -3.09375 C 3.671875 -1.28125 3.265625 -0.4375 2.40625 -0.4375 C 1.578125 -0.4375 1.171875 -1.3125 1.171875 -3.0625 C 1.171875 -4.828125 1.578125 -5.671875 2.421875 -5.671875 Z M 2.421875 -5.671875 "/>
</symbol>
<symbol overflow="visible" id="glyph0-3">
<path style="stroke:none;" d="M 1.9375 -2.921875 L 2.359375 -2.921875 C 3.203125 -2.921875 3.65625 -2.53125 3.65625 -1.765625 C 3.65625 -0.96875 3.171875 -0.484375 2.375 -0.484375 C 1.515625 -0.484375 1.109375 -0.921875 1.0625 -1.84375 L 0.28125 -1.84375 C 0.3125 -1.34375 0.40625 -1 0.546875 -0.71875 C 0.875 -0.109375 1.484375 0.203125 2.34375 0.203125 C 3.625 0.203125 4.453125 -0.578125 4.453125 -1.78125 C 4.453125 -2.59375 4.140625 -3.03125 3.390625 -3.296875 C 3.96875 -3.53125 4.265625 -3.96875 4.265625 -4.609375 C 4.265625 -5.703125 3.546875 -6.359375 2.359375 -6.359375 C 1.109375 -6.359375 0.4375 -5.65625 0.40625 -4.3125 L 1.1875 -4.3125 C 1.203125 -4.703125 1.234375 -4.921875 1.328125 -5.109375 C 1.5 -5.46875 1.890625 -5.6875 2.375 -5.6875 C 3.0625 -5.6875 3.46875 -5.265625 3.46875 -4.578125 C 3.46875 -4.140625 3.3125 -3.859375 2.96875 -3.71875 C 2.765625 -3.625 2.484375 -3.59375 1.9375 -3.578125 Z M 1.9375 -2.921875 "/>
</symbol>
<symbol overflow="visible" id="glyph0-4">
<path style="stroke:none;" d="M 2.875 -1.546875 L 2.875 0 L 3.65625 0 L 3.65625 -1.546875 L 4.578125 -1.546875 L 4.578125 -2.234375 L 3.65625 -2.234375 L 3.65625 -6.359375 L 3.078125 -6.359375 L 0.25 -2.359375 L 0.25 -1.546875 Z M 2.875 -2.234375 L 0.921875 -2.234375 L 2.875 -5.046875 Z M 2.875 -2.234375 "/>
</symbol>
<symbol overflow="visible" id="glyph0-5">
<path style="stroke:none;" d="M 2.28125 -4.53125 L 2.28125 0 L 3.046875 0 L 3.046875 -6.359375 L 2.546875 -6.359375 C 2.265625 -5.390625 2.09375 -5.25 0.890625 -5.09375 L 0.890625 -4.53125 Z M 2.28125 -4.53125 "/>
</symbol>
<symbol overflow="visible" id="glyph0-6">
<path style="stroke:none;" d="M 4.1875 -6.234375 L 0.96875 -6.234375 L 0.5 -2.84375 L 1.21875 -2.84375 C 1.578125 -3.265625 1.875 -3.421875 2.359375 -3.421875 C 3.203125 -3.421875 3.71875 -2.84375 3.71875 -1.921875 C 3.71875 -1.03125 3.203125 -0.484375 2.359375 -0.484375 C 1.6875 -0.484375 1.265625 -0.828125 1.078125 -1.53125 L 0.3125 -1.53125 C 0.40625 -1.015625 0.5 -0.78125 0.6875 -0.546875 C 1.03125 -0.078125 1.671875 0.203125 2.375 0.203125 C 3.640625 0.203125 4.515625 -0.71875 4.515625 -2.03125 C 4.515625 -3.265625 3.6875 -4.109375 2.5 -4.109375 C 2.0625 -4.109375 1.703125 -4 1.34375 -3.734375 L 1.59375 -5.46875 L 4.1875 -5.46875 Z M 4.1875 -6.234375 "/>
</symbol>
<symbol overflow="visible" id="glyph0-7">
<path style="stroke:none;" d="M 4.375 -4.703125 C 4.234375 -5.75 3.5625 -6.359375 2.609375 -6.359375 C 1.921875 -6.359375 1.3125 -6.03125 0.9375 -5.453125 C 0.546875 -4.828125 0.375 -4.0625 0.375 -2.90625 C 0.375 -1.828125 0.53125 -1.15625 0.90625 -0.59375 C 1.234375 -0.078125 1.78125 0.203125 2.46875 0.203125 C 3.65625 0.203125 4.515625 -0.6875 4.515625 -1.921875 C 4.515625 -3.09375 3.71875 -3.921875 2.609375 -3.921875 C 1.984375 -3.921875 1.5 -3.6875 1.171875 -3.234375 C 1.171875 -4.796875 1.671875 -5.671875 2.5625 -5.671875 C 3.109375 -5.671875 3.484375 -5.328125 3.609375 -4.703125 Z M 2.5 -3.234375 C 3.25 -3.234375 3.71875 -2.703125 3.71875 -1.859375 C 3.71875 -1.0625 3.1875 -0.484375 2.484375 -0.484375 C 1.765625 -0.484375 1.21875 -1.09375 1.21875 -1.90625 C 1.21875 -2.6875 1.734375 -3.234375 2.5 -3.234375 Z M 2.5 -3.234375 "/>
</symbol>
<symbol overflow="visible" id="glyph0-8">
<path style="stroke:none;" d="M 3.4375 -3.359375 C 4.09375 -3.75 4.296875 -4.0625 4.296875 -4.65625 C 4.296875 -5.671875 3.53125 -6.359375 2.421875 -6.359375 C 1.3125 -6.359375 0.546875 -5.671875 0.546875 -4.671875 C 0.546875 -4.0625 0.75 -3.75 1.390625 -3.359375 C 0.671875 -2.984375 0.328125 -2.46875 0.328125 -1.765625 C 0.328125 -0.609375 1.171875 0.203125 2.421875 0.203125 C 3.65625 0.203125 4.515625 -0.609375 4.515625 -1.765625 C 4.515625 -2.46875 4.15625 -2.984375 3.4375 -3.359375 Z M 2.421875 -5.671875 C 3.078125 -5.671875 3.5 -5.28125 3.5 -4.640625 C 3.5 -4.046875 3.0625 -3.65625 2.421875 -3.65625 C 1.765625 -3.65625 1.34375 -4.046875 1.34375 -4.65625 C 1.34375 -5.28125 1.765625 -5.671875 2.421875 -5.671875 Z M 2.421875 -3 C 3.1875 -3 3.71875 -2.484375 3.71875 -1.75 C 3.71875 -0.984375 3.203125 -0.484375 2.40625 -0.484375 C 1.640625 -0.484375 1.125 -1 1.125 -1.734375 C 1.125 -2.5 1.640625 -3 2.421875 -3 Z M 2.421875 -3 "/>
</symbol>
<symbol overflow="visible" id="glyph1-0">
<path style="stroke:none;" d=""/>
</symbol>
<symbol overflow="visible" id="glyph1-1">
<path style="stroke:none;" d="M 5.1875 -3.828125 C 5.140625 -4.390625 5.015625 -4.75 4.796875 -5.078125 C 4.40625 -5.609375 3.703125 -5.921875 2.90625 -5.921875 C 1.34375 -5.921875 0.34375 -4.703125 0.34375 -2.78125 C 0.34375 -0.921875 1.328125 0.25 2.890625 0.25 C 4.265625 0.25 5.140625 -0.578125 5.25 -1.984375 L 4.328125 -1.984375 C 4.171875 -1.0625 3.703125 -0.59375 2.921875 -0.59375 C 1.90625 -0.59375 1.296875 -1.421875 1.296875 -2.78125 C 1.296875 -4.21875 1.890625 -5.078125 2.890625 -5.078125 C 3.65625 -5.078125 4.140625 -4.625 4.25 -3.828125 Z M 5.1875 -3.828125 "/>
</symbol>
<symbol overflow="visible" id="glyph1-2">
<path style="stroke:none;" d="M 2.796875 -5.765625 L 1.84375 -5.765625 L 1.84375 -7.34375 L 0.9375 -7.34375 L 0.9375 -5.765625 L 0.15625 -5.765625 L 0.15625 -5.015625 L 0.9375 -5.015625 L 0.9375 -0.65625 C 0.9375 -0.0625 1.328125 0.25 2.046875 0.25 C 2.28125 0.25 2.484375 0.234375 2.796875 0.171875 L 2.796875 -0.59375 C 2.65625 -0.5625 2.546875 -0.546875 2.359375 -0.546875 C 1.953125 -0.546875 1.84375 -0.65625 1.84375 -1.0625 L 1.84375 -5.015625 L 2.796875 -5.015625 Z M 2.796875 -5.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-3">
<path style="stroke:none;" d="M 4.265625 -5.765625 L 2.671875 -1.28125 L 1.203125 -5.765625 L 0.21875 -5.765625 L 2.171875 0.015625 L 1.8125 0.9375 C 1.65625 1.34375 1.46875 1.5 1.078125 1.5 C 0.921875 1.5 0.796875 1.46875 0.59375 1.4375 L 0.59375 2.25 C 0.78125 2.359375 0.96875 2.390625 1.203125 2.390625 C 1.5 2.390625 1.828125 2.296875 2.0625 2.125 C 2.359375 1.90625 2.515625 1.671875 2.6875 1.203125 L 5.265625 -5.765625 Z M 4.265625 -5.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph1-4">
<path style="stroke:none;" d="M 5.640625 -2.578125 C 5.640625 -3.453125 5.578125 -3.984375 5.40625 -4.40625 C 5.03125 -5.359375 4.15625 -5.921875 3.078125 -5.921875 C 1.46875 -5.921875 0.4375 -4.703125 0.4375 -2.8125 C 0.4375 -0.90625 1.4375 0.25 3.0625 0.25 C 4.375 0.25 5.296875 -0.5 5.515625 -1.75 L 4.59375 -1.75 C 4.34375 -0.984375 3.828125 -0.59375 3.09375 -0.59375 C 2.515625 -0.59375 2.015625 -0.859375 1.703125 -1.34375 C 1.484375 -1.671875 1.40625 -2 1.390625 -2.578125 Z M 1.421875 -3.328125 C 1.5 -4.390625 2.140625 -5.078125 3.0625 -5.078125 C 4 -5.078125 4.65625 -4.359375 4.65625 -3.328125 Z M 1.421875 -3.328125 "/>
</symbol>
<symbol overflow="visible" id="glyph1-5">
<path style="stroke:none;" d="M 5.890625 -0.53125 C 5.78125 -0.515625 5.734375 -0.515625 5.6875 -0.515625 C 5.375 -0.515625 5.1875 -0.6875 5.1875 -0.96875 L 5.1875 -4.359375 C 5.1875 -5.375 4.4375 -5.921875 3.03125 -5.921875 C 2.171875 -5.921875 1.5 -5.6875 1.109375 -5.265625 C 0.84375 -4.96875 0.734375 -4.625 0.71875 -4.0625 L 1.640625 -4.0625 C 1.71875 -4.765625 2.140625 -5.078125 2.984375 -5.078125 C 3.828125 -5.078125 4.28125 -4.78125 4.28125 -4.21875 L 4.28125 -3.984375 C 4.265625 -3.59375 4.0625 -3.4375 3.328125 -3.34375 C 2.03125 -3.171875 1.828125 -3.140625 1.46875 -2.984375 C 0.796875 -2.703125 0.46875 -2.203125 0.46875 -1.453125 C 0.46875 -0.40625 1.1875 0.25 2.359375 0.25 C 3.078125 0.25 3.65625 0 4.3125 -0.59375 C 4.375 0 4.65625 0.25 5.265625 0.25 C 5.453125 0.25 5.578125 0.234375 5.890625 0.15625 Z M 4.28125 -1.8125 C 4.28125 -1.5 4.1875 -1.3125 3.921875 -1.0625 C 3.546875 -0.71875 3.09375 -0.546875 2.546875 -0.546875 C 1.84375 -0.546875 1.421875 -0.890625 1.421875 -1.46875 C 1.421875 -2.078125 1.8125 -2.390625 2.8125 -2.53125 C 3.78125 -2.65625 3.96875 -2.703125 4.28125 -2.84375 Z M 4.28125 -1.8125 "/>
</symbol>
<symbol overflow="visible" id="glyph1-6">
<path style="stroke:none;" d="M 0.765625 -5.765625 L 0.765625 0 L 1.6875 0 L 1.6875 -2.984375 C 1.6875 -4.375 2.265625 -5 3.53125 -4.96875 L 3.53125 -5.890625 C 3.375 -5.921875 3.28125 -5.921875 3.171875 -5.921875 C 2.578125 -5.921875 2.140625 -5.578125 1.609375 -4.71875 L 1.609375 -5.765625 Z M 0.765625 -5.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph2-0">
<path style="stroke:none;" d=""/>
</symbol>
<symbol overflow="visible" id="glyph2-1">
<path style="stroke:none;" d="M -8.015625 -0.765625 L 0 -0.765625 L 0 -1.6875 L -3.171875 -1.6875 C -4.359375 -1.6875 -5.125 -2.296875 -5.125 -3.25 C -5.125 -3.546875 -5.03125 -3.84375 -4.875 -4.0625 C -4.671875 -4.328125 -4.40625 -4.4375 -4 -4.4375 L 0 -4.4375 L 0 -5.34375 L -4.359375 -5.34375 C -5.328125 -5.34375 -5.921875 -4.65625 -5.921875 -3.53125 C -5.921875 -2.71875 -5.671875 -2.21875 -4.96875 -1.6875 L -8.015625 -1.6875 Z M -8.015625 -0.765625 "/>
</symbol>
<symbol overflow="visible" id="glyph2-2">
<path style="stroke:none;" d="M 0 -6.09375 L -5.765625 -7.78125 L -5.765625 -6.75 L -1.28125 -5.609375 L -5.765625 -4.484375 L -5.765625 -3.359375 L -1.28125 -2.25 L -5.765625 -1.078125 L -5.765625 -0.0625 L 0 -1.734375 L 0 -2.765625 L -4.515625 -3.890625 L 0 -5.046875 Z M 0 -6.09375 "/>
</symbol>
<symbol overflow="visible" id="glyph2-3">
<path style="stroke:none;" d="M -5.765625 -4.265625 L -1.28125 -2.671875 L -5.765625 -1.203125 L -5.765625 -0.21875 L 0.015625 -2.171875 L 0.9375 -1.8125 C 1.34375 -1.65625 1.5 -1.46875 1.5 -1.078125 C 1.5 -0.921875 1.46875 -0.796875 1.4375 -0.59375 L 2.25 -0.59375 C 2.359375 -0.78125 2.390625 -0.96875 2.390625 -1.203125 C 2.390625 -1.5 2.296875 -1.828125 2.125 -2.0625 C 1.90625 -2.359375 1.671875 -2.515625 1.203125 -2.6875 L -5.765625 -5.265625 Z M -5.765625 -4.265625 "/>
</symbol>
</g>
<clipPath id="clip1">
  <path d="M 35.152344 5.480469 L 433.84375 5.480469 L 433.84375 469.851562 L 35.152344 469.851562 Z M 35.152344 5.480469 "/>
</clipPath>
<clipPath id="clip2">
  <path d="M 35.152344 414 L 433.84375 414 L 433.84375 415 L 35.152344 415 Z M 35.152344 414 "/>
</clipPath>
<clipPath id="clip3">
  <path d="M 35.152344 348 L 433.84375 348 L 433.84375 349 L 35.152344 349 Z M 35.152344 348 "/>
</clipPath>
<clipPath id="clip4">
  <path d="M 35.152344 216 L 433.84375 216 L 433.84375 218 L 35.152344 218 Z M 35.152344 216 "/>
</clipPath>
<clipPath id="clip5">
  <path d="M 35.152344 104 L 433.84375 104 L 433.84375 105 L 35.152344 105 Z M 35.152344 104 "/>
</clipPath>
<clipPath id="clip6">
  <path d="M 101 5.480469 L 103 5.480469 L 103 469.851562 L 101 469.851562 Z M 101 5.480469 "/>
</clipPath>
<clipPath id="clip7">
  <path d="M 171 5.480469 L 173 5.480469 L 173 469.851562 L 171 469.851562 Z M 171 5.480469 "/>
</clipPath>
<clipPath id="clip8">
  <path d="M 241 5.480469 L 242 5.480469 L 242 469.851562 L 241 469.851562 Z M 241 5.480469 "/>
</clipPath>
<clipPath id="clip9">
  <path d="M 310 5.480469 L 312 5.480469 L 312 469.851562 L 310 469.851562 Z M 310 5.480469 "/>
</clipPath>
<clipPath id="clip10">
  <path d="M 380 5.480469 L 382 5.480469 L 382 469.851562 L 380 469.851562 Z M 380 5.480469 "/>
</clipPath>
<clipPath id="clip11">
  <path d="M 35.152344 282 L 433.84375 282 L 433.84375 284 L 35.152344 284 Z M 35.152344 282 "/>
</clipPath>
<clipPath id="clip12">
  <path d="M 35.152344 150 L 433.84375 150 L 433.84375 152 L 35.152344 152 Z M 35.152344 150 "/>
</clipPath>
<clipPath id="clip13">
  <path d="M 35.152344 57 L 433.84375 57 L 433.84375 59 L 35.152344 59 Z M 35.152344 57 "/>
</clipPath>
<clipPath id="clip14">
  <path d="M 66 5.480469 L 68 5.480469 L 68 469.851562 L 66 469.851562 Z M 66 5.480469 "/>
</clipPath>
<clipPath id="clip15">
  <path d="M 136 5.480469 L 138 5.480469 L 138 469.851562 L 136 469.851562 Z M 136 5.480469 "/>
</clipPath>
<clipPath id="clip16">
  <path d="M 206 5.480469 L 208 5.480469 L 208 469.851562 L 206 469.851562 Z M 206 5.480469 "/>
</clipPath>
<clipPath id="clip17">
  <path d="M 275 5.480469 L 277 5.480469 L 277 469.851562 L 275 469.851562 Z M 275 5.480469 "/>
</clipPath>
<clipPath id="clip18">
  <path d="M 345 5.480469 L 347 5.480469 L 347 469.851562 L 345 469.851562 Z M 345 5.480469 "/>
</clipPath>
<clipPath id="clip19">
  <path d="M 415 5.480469 L 417 5.480469 L 417 469.851562 L 415 469.851562 Z M 415 5.480469 "/>
</clipPath>
<clipPath id="clip20">
  <path d="M 450.28125 203.203125 L 467.5625 203.203125 L 467.5625 289.601562 L 450.28125 289.601562 Z M 450.28125 203.203125 "/>
</clipPath>
<image id="image244" width="1" height="300" xlink:href="data:image/png;base64,iVBORw0KGgoAAAANSUhEUgAAAAEAAAEsCAIAAACNMk6CAAAABmJLR0QA/wD/AP+gvaeTAAAAjklEQVQ4jaWTQRKAMAgDd3ik//R1HuOlji3SjNYjhZAEKNt+BCgQIRRIAQRSi2l5hdr79/q1fqBQVw+ErvyipqGH5eh7fNFJi1M+9bC8OT/lrTD9fNYwT/0FpvQzwVju0eObndsd/uUa7qX2NffqNE5w7j7dbCudpXbj2ehX0n3f0ot5pX9/1xkPj79W8p41NH/nSHk1sQAAAABJRU5ErkJggg=="/>
<pattern id="pattern0" patternUnits="userSpaceOnUse" width="1" height="300"  patternTransform="matrix(17.28,0,0,0.288,450.281644,203.20411)">
  <use xlink:href="#image244"/>
</pattern>
</defs>
<g id="surface240">
<rect x="0" y="0" width="504" height="504" style="fill:rgb(100%,100%,100%);fill-opacity:1;stroke:none;"/>
<rect x="0" y="0" width="504" height="504" style="fill:rgb(100%,100%,100%);fill-opacity:1;stroke:none;"/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 0 504 L 504 504 L 504 0 L 0 0 Z M 0 504 "/>
<g clip-path="url(#clip1)" clip-rule="nonzero">
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(92.156863%,92.156863%,92.156863%);fill-opacity:1;" d="M 35.152344 469.847656 L 433.84375 469.847656 L 433.84375 5.476562 L 35.152344 5.476562 Z M 35.152344 469.847656 "/>
</g>
<g clip-path="url(#clip2)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 414.507812 L 433.84375 414.507812 "/>
</g>
<g clip-path="url(#clip3)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 348.636719 L 433.84375 348.636719 "/>
</g>
<g clip-path="url(#clip4)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 216.898438 L 433.84375 216.898438 "/>
</g>
<g clip-path="url(#clip5)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 104.289062 L 433.84375 104.289062 "/>
</g>
<g clip-path="url(#clip6)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 102.0625 469.847656 L 102.0625 5.480469 "/>
</g>
<g clip-path="url(#clip7)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 171.765625 469.847656 L 171.765625 5.480469 "/>
</g>
<g clip-path="url(#clip8)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 241.46875 469.847656 L 241.46875 5.480469 "/>
</g>
<g clip-path="url(#clip9)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 311.167969 469.847656 L 311.167969 5.480469 "/>
</g>
<g clip-path="url(#clip10)" clip-rule="nonzero">
<path style="fill:none;stroke-width:0.533489;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 380.871094 469.847656 L 380.871094 5.480469 "/>
</g>
<g clip-path="url(#clip11)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 282.765625 L 433.84375 282.765625 "/>
</g>
<g clip-path="url(#clip12)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 151.027344 L 433.84375 151.027344 "/>
</g>
<g clip-path="url(#clip13)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 35.152344 57.554688 L 433.84375 57.554688 "/>
</g>
<g clip-path="url(#clip14)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 67.214844 469.847656 L 67.214844 5.480469 "/>
</g>
<g clip-path="url(#clip15)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 136.914062 469.847656 L 136.914062 5.480469 "/>
</g>
<g clip-path="url(#clip16)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 206.617188 469.847656 L 206.617188 5.480469 "/>
</g>
<g clip-path="url(#clip17)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 276.316406 469.847656 L 276.316406 5.480469 "/>
</g>
<g clip-path="url(#clip18)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 346.019531 469.847656 L 346.019531 5.480469 "/>
</g>
<g clip-path="url(#clip19)" clip-rule="nonzero">
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 415.722656 469.847656 L 415.722656 5.480469 "/>
</g>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 162.042969 C 184.425781 165.183594 181.878906 167.730469 178.734375 167.730469 C 175.59375 167.730469 173.046875 165.183594 173.046875 162.042969 C 173.046875 158.898438 175.59375 156.351562 178.734375 156.351562 C 181.878906 156.351562 184.425781 158.898438 184.425781 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 212.304688 140.371094 C 212.304688 143.515625 209.757812 146.0625 206.617188 146.0625 C 203.472656 146.0625 200.925781 143.515625 200.925781 140.371094 C 200.925781 137.230469 203.472656 134.683594 206.617188 134.683594 C 209.757812 134.683594 212.304688 137.230469 212.304688 140.371094 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 151.027344 C 226.246094 154.167969 223.699219 156.714844 220.558594 156.714844 C 217.414062 156.714844 214.867188 154.167969 214.867188 151.027344 C 214.867188 147.882812 217.414062 145.335938 220.558594 145.335938 C 223.699219 145.335938 226.246094 147.882812 226.246094 151.027344 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 197.523438 C 156.542969 200.664062 153.996094 203.210938 150.855469 203.210938 C 147.714844 203.210938 145.164062 200.664062 145.164062 197.523438 C 145.164062 194.378906 147.714844 191.832031 150.855469 191.832031 C 153.996094 191.832031 156.542969 194.378906 156.542969 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 185.257812 C 184.425781 188.402344 181.878906 190.949219 178.734375 190.949219 C 175.59375 190.949219 173.046875 188.402344 173.046875 185.257812 C 173.046875 182.117188 175.59375 179.570312 178.734375 179.570312 C 181.878906 179.570312 184.425781 182.117188 184.425781 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 210.265625 C 156.542969 213.40625 153.996094 215.953125 150.855469 215.953125 C 147.714844 215.953125 145.164062 213.40625 145.164062 210.265625 C 145.164062 207.125 147.714844 204.574219 150.855469 204.574219 C 153.996094 204.574219 156.542969 207.125 156.542969 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 212.304688 173.441406 C 212.304688 176.585938 209.757812 179.132812 206.617188 179.132812 C 203.472656 179.132812 200.925781 176.585938 200.925781 173.441406 C 200.925781 170.300781 203.472656 167.753906 206.617188 167.753906 C 209.757812 167.753906 212.304688 170.300781 212.304688 173.441406 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 185.257812 C 198.367188 188.402344 195.816406 190.949219 192.675781 190.949219 C 189.535156 190.949219 186.988281 188.402344 186.988281 185.257812 C 186.988281 182.117188 189.535156 179.570312 192.675781 179.570312 C 195.816406 179.570312 198.367188 182.117188 198.367188 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 210.265625 C 142.605469 213.40625 140.058594 215.953125 136.914062 215.953125 C 133.773438 215.953125 131.226562 213.40625 131.226562 210.265625 C 131.226562 207.125 133.773438 204.574219 136.914062 204.574219 C 140.058594 204.574219 142.605469 207.125 142.605469 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 210.265625 C 170.484375 213.40625 167.9375 215.953125 164.796875 215.953125 C 161.652344 215.953125 159.105469 213.40625 159.105469 210.265625 C 159.105469 207.125 161.652344 204.574219 164.796875 204.574219 C 167.9375 204.574219 170.484375 207.125 170.484375 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 210.265625 C 170.484375 213.40625 167.9375 215.953125 164.796875 215.953125 C 161.652344 215.953125 159.105469 213.40625 159.105469 210.265625 C 159.105469 207.125 161.652344 204.574219 164.796875 204.574219 C 167.9375 204.574219 170.484375 207.125 170.484375 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 210.265625 C 142.605469 213.40625 140.058594 215.953125 136.914062 215.953125 C 133.773438 215.953125 131.226562 213.40625 131.226562 210.265625 C 131.226562 207.125 133.773438 204.574219 136.914062 204.574219 C 140.058594 204.574219 142.605469 207.125 142.605469 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 223.527344 C 142.605469 226.671875 140.058594 229.21875 136.914062 229.21875 C 133.773438 229.21875 131.226562 226.671875 131.226562 223.527344 C 131.226562 220.386719 133.773438 217.839844 136.914062 217.839844 C 140.058594 217.839844 142.605469 220.386719 142.605469 223.527344 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 210.265625 C 170.484375 213.40625 167.9375 215.953125 164.796875 215.953125 C 161.652344 215.953125 159.105469 213.40625 159.105469 210.265625 C 159.105469 207.125 161.652344 204.574219 164.796875 204.574219 C 167.9375 204.574219 170.484375 207.125 170.484375 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 237.355469 C 156.542969 240.5 153.996094 243.046875 150.855469 243.046875 C 147.714844 243.046875 145.164062 240.5 145.164062 237.355469 C 145.164062 234.214844 147.714844 231.667969 150.855469 231.667969 C 153.996094 231.667969 156.542969 234.214844 156.542969 237.355469 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 282.765625 C 128.664062 285.910156 126.117188 288.457031 122.972656 288.457031 C 119.832031 288.457031 117.285156 285.910156 117.285156 282.765625 C 117.285156 279.625 119.832031 277.078125 122.972656 277.078125 C 126.117188 277.078125 128.664062 279.625 128.664062 282.765625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 376.238281 C 86.84375 379.382812 84.296875 381.929688 81.152344 381.929688 C 78.011719 381.929688 75.464844 379.382812 75.464844 376.238281 C 75.464844 373.097656 78.011719 370.550781 81.152344 370.550781 C 84.296875 370.550781 86.84375 373.097656 86.84375 376.238281 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 282.765625 C 128.664062 285.910156 126.117188 288.457031 122.972656 288.457031 C 119.832031 288.457031 117.285156 285.910156 117.285156 282.765625 C 117.285156 279.625 119.832031 277.078125 122.972656 277.078125 C 126.117188 277.078125 128.664062 279.625 128.664062 282.765625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 335.570312 C 114.722656 338.714844 112.175781 341.261719 109.035156 341.261719 C 105.890625 341.261719 103.34375 338.714844 103.34375 335.570312 C 103.34375 332.429688 105.890625 329.882812 109.035156 329.882812 C 112.175781 329.882812 114.722656 332.429688 114.722656 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.78125 335.570312 C 100.78125 338.714844 98.234375 341.261719 95.09375 341.261719 C 91.953125 341.261719 89.40625 338.714844 89.40625 335.570312 C 89.40625 332.429688 91.953125 329.882812 95.09375 329.882812 C 98.234375 329.882812 100.78125 332.429688 100.78125 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 197.523438 C 156.542969 200.664062 153.996094 203.210938 150.855469 203.210938 C 147.714844 203.210938 145.164062 200.664062 145.164062 197.523438 C 145.164062 194.378906 147.714844 191.832031 150.855469 191.832031 C 153.996094 191.832031 156.542969 194.378906 156.542969 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 237.355469 C 142.605469 240.5 140.058594 243.046875 136.914062 243.046875 C 133.773438 243.046875 131.226562 240.5 131.226562 237.355469 C 131.226562 234.214844 133.773438 231.667969 136.914062 231.667969 C 140.058594 231.667969 142.605469 234.214844 142.605469 237.355469 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 197.523438 C 156.542969 200.664062 153.996094 203.210938 150.855469 203.210938 C 147.714844 203.210938 145.164062 200.664062 145.164062 197.523438 C 145.164062 194.378906 147.714844 191.832031 150.855469 191.832031 C 153.996094 191.832031 156.542969 194.378906 156.542969 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 210.265625 C 142.605469 213.40625 140.058594 215.953125 136.914062 215.953125 C 133.773438 215.953125 131.226562 213.40625 131.226562 210.265625 C 131.226562 207.125 133.773438 204.574219 136.914062 204.574219 C 140.058594 204.574219 142.605469 207.125 142.605469 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 223.527344 C 142.605469 226.671875 140.058594 229.21875 136.914062 229.21875 C 133.773438 229.21875 131.226562 226.671875 131.226562 223.527344 C 131.226562 220.386719 133.773438 217.839844 136.914062 217.839844 C 140.058594 217.839844 142.605469 220.386719 142.605469 223.527344 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 299.433594 C 128.664062 302.574219 126.117188 305.121094 122.972656 305.121094 C 119.832031 305.121094 117.285156 302.574219 117.285156 299.433594 C 117.285156 296.292969 119.832031 293.746094 122.972656 293.746094 C 126.117188 293.746094 128.664062 296.292969 128.664062 299.433594 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 398.65625 C 86.84375 401.796875 84.296875 404.34375 81.152344 404.34375 C 78.011719 404.34375 75.464844 401.796875 75.464844 398.65625 C 75.464844 395.515625 78.011719 392.964844 81.152344 392.964844 C 84.296875 392.964844 86.84375 395.515625 86.84375 398.65625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 376.238281 C 86.84375 379.382812 84.296875 381.929688 81.152344 381.929688 C 78.011719 381.929688 75.464844 379.382812 75.464844 376.238281 C 75.464844 373.097656 78.011719 370.550781 81.152344 370.550781 C 84.296875 370.550781 86.84375 373.097656 86.84375 376.238281 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 335.570312 C 128.664062 338.714844 126.117188 341.261719 122.972656 341.261719 C 119.832031 341.261719 117.285156 338.714844 117.285156 335.570312 C 117.285156 332.429688 119.832031 329.882812 122.972656 329.882812 C 126.117188 329.882812 128.664062 332.429688 128.664062 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 185.257812 C 198.367188 188.402344 195.816406 190.949219 192.675781 190.949219 C 189.535156 190.949219 186.988281 188.402344 186.988281 185.257812 C 186.988281 182.117188 189.535156 179.570312 192.675781 179.570312 C 195.816406 179.570312 198.367188 182.117188 198.367188 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 240.1875 151.027344 C 240.1875 154.167969 237.640625 156.714844 234.496094 156.714844 C 231.355469 156.714844 228.808594 154.167969 228.808594 151.027344 C 228.808594 147.882812 231.355469 145.335938 234.496094 145.335938 C 237.640625 145.335938 240.1875 147.882812 240.1875 151.027344 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 162.042969 C 184.425781 165.183594 181.878906 167.730469 178.734375 167.730469 C 175.59375 167.730469 173.046875 165.183594 173.046875 162.042969 C 173.046875 158.898438 175.59375 156.351562 178.734375 156.351562 C 181.878906 156.351562 184.425781 158.898438 184.425781 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 197.523438 C 170.484375 200.664062 167.9375 203.210938 164.796875 203.210938 C 161.652344 203.210938 159.105469 200.664062 159.105469 197.523438 C 159.105469 194.378906 161.652344 191.832031 164.796875 191.832031 C 167.9375 191.832031 170.484375 194.378906 170.484375 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 223.527344 C 184.425781 226.671875 181.878906 229.21875 178.734375 229.21875 C 175.59375 229.21875 173.046875 226.671875 173.046875 223.527344 C 173.046875 220.386719 175.59375 217.839844 178.734375 217.839844 C 181.878906 217.839844 184.425781 220.386719 184.425781 223.527344 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 223.527344 C 170.484375 226.671875 167.9375 229.21875 164.796875 229.21875 C 161.652344 229.21875 159.105469 226.671875 159.105469 223.527344 C 159.105469 220.386719 161.652344 217.839844 164.796875 217.839844 C 167.9375 217.839844 170.484375 220.386719 170.484375 223.527344 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 251.800781 C 156.542969 254.941406 153.996094 257.488281 150.855469 257.488281 C 147.714844 257.488281 145.164062 254.941406 145.164062 251.800781 C 145.164062 248.65625 147.714844 246.109375 150.855469 246.109375 C 153.996094 246.109375 156.542969 248.65625 156.542969 251.800781 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 251.800781 C 156.542969 254.941406 153.996094 257.488281 150.855469 257.488281 C 147.714844 257.488281 145.164062 254.941406 145.164062 251.800781 C 145.164062 248.65625 147.714844 246.109375 150.855469 246.109375 C 153.996094 246.109375 156.542969 248.65625 156.542969 251.800781 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 223.527344 C 170.484375 226.671875 167.9375 229.21875 164.796875 229.21875 C 161.652344 229.21875 159.105469 226.671875 159.105469 223.527344 C 159.105469 220.386719 161.652344 217.839844 164.796875 217.839844 C 167.9375 217.839844 170.484375 220.386719 170.484375 223.527344 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 223.527344 C 170.484375 226.671875 167.9375 229.21875 164.796875 229.21875 C 161.652344 229.21875 159.105469 226.671875 159.105469 223.527344 C 159.105469 220.386719 161.652344 217.839844 164.796875 217.839844 C 167.9375 217.839844 170.484375 220.386719 170.484375 223.527344 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 335.570312 C 86.84375 338.714844 84.296875 341.261719 81.152344 341.261719 C 78.011719 341.261719 75.464844 338.714844 75.464844 335.570312 C 75.464844 332.429688 78.011719 329.882812 81.152344 329.882812 C 84.296875 329.882812 86.84375 332.429688 86.84375 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 251.800781 C 142.605469 254.941406 140.058594 257.488281 136.914062 257.488281 C 133.773438 257.488281 131.226562 254.941406 131.226562 251.800781 C 131.226562 248.65625 133.773438 246.109375 136.914062 246.109375 C 140.058594 246.109375 142.605469 248.65625 142.605469 251.800781 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 266.914062 C 142.605469 270.058594 140.058594 272.605469 136.914062 272.605469 C 133.773438 272.605469 131.226562 270.058594 131.226562 266.914062 C 131.226562 263.773438 133.773438 261.226562 136.914062 261.226562 C 140.058594 261.226562 142.605469 263.773438 142.605469 266.914062 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 237.355469 C 156.542969 240.5 153.996094 243.046875 150.855469 243.046875 C 147.714844 243.046875 145.164062 240.5 145.164062 237.355469 C 145.164062 234.214844 147.714844 231.667969 150.855469 231.667969 C 153.996094 231.667969 156.542969 234.214844 156.542969 237.355469 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 237.355469 C 156.542969 240.5 153.996094 243.046875 150.855469 243.046875 C 147.714844 243.046875 145.164062 240.5 145.164062 237.355469 C 145.164062 234.214844 147.714844 231.667969 150.855469 231.667969 C 153.996094 231.667969 156.542969 234.214844 156.542969 237.355469 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 299.433594 C 142.605469 302.574219 140.058594 305.121094 136.914062 305.121094 C 133.773438 305.121094 131.226562 302.574219 131.226562 299.433594 C 131.226562 296.292969 133.773438 293.746094 136.914062 293.746094 C 140.058594 293.746094 142.605469 296.292969 142.605469 299.433594 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 317 C 128.664062 320.144531 126.117188 322.691406 122.972656 322.691406 C 119.832031 322.691406 117.285156 320.144531 117.285156 317 C 117.285156 313.859375 119.832031 311.3125 122.972656 311.3125 C 126.117188 311.3125 128.664062 313.859375 128.664062 317 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 335.570312 C 114.722656 338.714844 112.175781 341.261719 109.035156 341.261719 C 105.890625 341.261719 103.34375 338.714844 103.34375 335.570312 C 103.34375 332.429688 105.890625 329.882812 109.035156 329.882812 C 112.175781 329.882812 114.722656 332.429688 114.722656 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 335.570312 C 128.664062 338.714844 126.117188 341.261719 122.972656 341.261719 C 119.832031 341.261719 117.285156 338.714844 117.285156 335.570312 C 117.285156 332.429688 119.832031 329.882812 122.972656 329.882812 C 126.117188 329.882812 128.664062 332.429688 128.664062 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 299.433594 C 128.664062 302.574219 126.117188 305.121094 122.972656 305.121094 C 119.832031 305.121094 117.285156 302.574219 117.285156 299.433594 C 117.285156 296.292969 119.832031 293.746094 122.972656 293.746094 C 126.117188 293.746094 128.664062 296.292969 128.664062 299.433594 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 299.433594 C 128.664062 302.574219 126.117188 305.121094 122.972656 305.121094 C 119.832031 305.121094 117.285156 302.574219 117.285156 299.433594 C 117.285156 296.292969 119.832031 293.746094 122.972656 293.746094 C 126.117188 293.746094 128.664062 296.292969 128.664062 299.433594 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 58.960938 448.742188 C 58.960938 451.882812 56.414062 454.429688 53.273438 454.429688 C 50.132812 454.429688 47.582031 451.882812 47.582031 448.742188 C 47.582031 445.597656 50.132812 443.050781 53.273438 443.050781 C 56.414062 443.050781 58.960938 445.597656 58.960938 448.742188 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 335.570312 C 86.84375 338.714844 84.296875 341.261719 81.152344 341.261719 C 78.011719 341.261719 75.464844 338.714844 75.464844 335.570312 C 75.464844 332.429688 78.011719 329.882812 81.152344 329.882812 C 84.296875 329.882812 86.84375 332.429688 86.84375 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 376.238281 C 86.84375 379.382812 84.296875 381.929688 81.152344 381.929688 C 78.011719 381.929688 75.464844 379.382812 75.464844 376.238281 C 75.464844 373.097656 78.011719 370.550781 81.152344 370.550781 C 84.296875 370.550781 86.84375 373.097656 86.84375 376.238281 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 335.570312 C 114.722656 338.714844 112.175781 341.261719 109.035156 341.261719 C 105.890625 341.261719 103.34375 338.714844 103.34375 335.570312 C 103.34375 332.429688 105.890625 329.882812 109.035156 329.882812 C 112.175781 329.882812 114.722656 332.429688 114.722656 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 335.570312 C 114.722656 338.714844 112.175781 341.261719 109.035156 341.261719 C 105.890625 341.261719 103.34375 338.714844 103.34375 335.570312 C 103.34375 332.429688 105.890625 329.882812 109.035156 329.882812 C 112.175781 329.882812 114.722656 332.429688 114.722656 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 58.960938 448.742188 C 58.960938 451.882812 56.414062 454.429688 53.273438 454.429688 C 50.132812 454.429688 47.582031 451.882812 47.582031 448.742188 C 47.582031 445.597656 50.132812 443.050781 53.273438 443.050781 C 56.414062 443.050781 58.960938 445.597656 58.960938 448.742188 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 335.570312 C 114.722656 338.714844 112.175781 341.261719 109.035156 341.261719 C 105.890625 341.261719 103.34375 338.714844 103.34375 335.570312 C 103.34375 332.429688 105.890625 329.882812 109.035156 329.882812 C 112.175781 329.882812 114.722656 332.429688 114.722656 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 355.269531 C 86.84375 358.410156 84.296875 360.960938 81.152344 360.960938 C 78.011719 360.960938 75.464844 358.410156 75.464844 355.269531 C 75.464844 352.128906 78.011719 349.582031 81.152344 349.582031 C 84.296875 349.582031 86.84375 352.128906 86.84375 355.269531 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 317 C 114.722656 320.144531 112.175781 322.691406 109.035156 322.691406 C 105.890625 322.691406 103.34375 320.144531 103.34375 317 C 103.34375 313.859375 105.890625 311.3125 109.035156 311.3125 C 112.175781 311.3125 114.722656 313.859375 114.722656 317 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 376.238281 C 86.84375 379.382812 84.296875 381.929688 81.152344 381.929688 C 78.011719 381.929688 75.464844 379.382812 75.464844 376.238281 C 75.464844 373.097656 78.011719 370.550781 81.152344 370.550781 C 84.296875 370.550781 86.84375 373.097656 86.84375 376.238281 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.78125 355.269531 C 100.78125 358.410156 98.234375 360.960938 95.09375 360.960938 C 91.953125 360.960938 89.40625 358.410156 89.40625 355.269531 C 89.40625 352.128906 91.953125 349.582031 95.09375 349.582031 C 98.234375 349.582031 100.78125 352.128906 100.78125 355.269531 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 58.960938 448.742188 C 58.960938 451.882812 56.414062 454.429688 53.273438 454.429688 C 50.132812 454.429688 47.582031 451.882812 47.582031 448.742188 C 47.582031 445.597656 50.132812 443.050781 53.273438 443.050781 C 56.414062 443.050781 58.960938 445.597656 58.960938 448.742188 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 335.570312 C 114.722656 338.714844 112.175781 341.261719 109.035156 341.261719 C 105.890625 341.261719 103.34375 338.714844 103.34375 335.570312 C 103.34375 332.429688 105.890625 329.882812 109.035156 329.882812 C 112.175781 329.882812 114.722656 332.429688 114.722656 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 335.570312 C 114.722656 338.714844 112.175781 341.261719 109.035156 341.261719 C 105.890625 341.261719 103.34375 338.714844 103.34375 335.570312 C 103.34375 332.429688 105.890625 329.882812 109.035156 329.882812 C 112.175781 329.882812 114.722656 332.429688 114.722656 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.78125 355.269531 C 100.78125 358.410156 98.234375 360.960938 95.09375 360.960938 C 91.953125 360.960938 89.40625 358.410156 89.40625 355.269531 C 89.40625 352.128906 91.953125 349.582031 95.09375 349.582031 C 98.234375 349.582031 100.78125 352.128906 100.78125 355.269531 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 58.960938 448.742188 C 58.960938 451.882812 56.414062 454.429688 53.273438 454.429688 C 50.132812 454.429688 47.582031 451.882812 47.582031 448.742188 C 47.582031 445.597656 50.132812 443.050781 53.273438 443.050781 C 56.414062 443.050781 58.960938 445.597656 58.960938 448.742188 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 376.238281 C 86.84375 379.382812 84.296875 381.929688 81.152344 381.929688 C 78.011719 381.929688 75.464844 379.382812 75.464844 376.238281 C 75.464844 373.097656 78.011719 370.550781 81.152344 370.550781 C 84.296875 370.550781 86.84375 373.097656 86.84375 376.238281 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 355.269531 C 86.84375 358.410156 84.296875 360.960938 81.152344 360.960938 C 78.011719 360.960938 75.464844 358.410156 75.464844 355.269531 C 75.464844 352.128906 78.011719 349.582031 81.152344 349.582031 C 84.296875 349.582031 86.84375 352.128906 86.84375 355.269531 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 335.570312 C 114.722656 338.714844 112.175781 341.261719 109.035156 341.261719 C 105.890625 341.261719 103.34375 338.714844 103.34375 335.570312 C 103.34375 332.429688 105.890625 329.882812 109.035156 329.882812 C 112.175781 329.882812 114.722656 332.429688 114.722656 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 376.238281 C 86.84375 379.382812 84.296875 381.929688 81.152344 381.929688 C 78.011719 381.929688 75.464844 379.382812 75.464844 376.238281 C 75.464844 373.097656 78.011719 370.550781 81.152344 370.550781 C 84.296875 370.550781 86.84375 373.097656 86.84375 376.238281 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 335.570312 C 86.84375 338.714844 84.296875 341.261719 81.152344 341.261719 C 78.011719 341.261719 75.464844 338.714844 75.464844 335.570312 C 75.464844 332.429688 78.011719 329.882812 81.152344 329.882812 C 84.296875 329.882812 86.84375 332.429688 86.84375 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 335.570312 C 86.84375 338.714844 84.296875 341.261719 81.152344 341.261719 C 78.011719 341.261719 75.464844 338.714844 75.464844 335.570312 C 75.464844 332.429688 78.011719 329.882812 81.152344 329.882812 C 84.296875 329.882812 86.84375 332.429688 86.84375 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.78125 317 C 100.78125 320.144531 98.234375 322.691406 95.09375 322.691406 C 91.953125 322.691406 89.40625 320.144531 89.40625 317 C 89.40625 313.859375 91.953125 311.3125 95.09375 311.3125 C 98.234375 311.3125 100.78125 313.859375 100.78125 317 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 335.570312 C 128.664062 338.714844 126.117188 341.261719 122.972656 341.261719 C 119.832031 341.261719 117.285156 338.714844 117.285156 335.570312 C 117.285156 332.429688 119.832031 329.882812 122.972656 329.882812 C 126.117188 329.882812 128.664062 332.429688 128.664062 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 299.433594 C 142.605469 302.574219 140.058594 305.121094 136.914062 305.121094 C 133.773438 305.121094 131.226562 302.574219 131.226562 299.433594 C 131.226562 296.292969 133.773438 293.746094 136.914062 293.746094 C 140.058594 293.746094 142.605469 296.292969 142.605469 299.433594 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 335.570312 C 128.664062 338.714844 126.117188 341.261719 122.972656 341.261719 C 119.832031 341.261719 117.285156 338.714844 117.285156 335.570312 C 117.285156 332.429688 119.832031 329.882812 122.972656 329.882812 C 126.117188 329.882812 128.664062 332.429688 128.664062 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 299.433594 C 114.722656 302.574219 112.175781 305.121094 109.035156 305.121094 C 105.890625 305.121094 103.34375 302.574219 103.34375 299.433594 C 103.34375 296.292969 105.890625 293.746094 109.035156 293.746094 C 112.175781 293.746094 114.722656 296.292969 114.722656 299.433594 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 299.433594 C 114.722656 302.574219 112.175781 305.121094 109.035156 305.121094 C 105.890625 305.121094 103.34375 302.574219 103.34375 299.433594 C 103.34375 296.292969 105.890625 293.746094 109.035156 293.746094 C 112.175781 293.746094 114.722656 296.292969 114.722656 299.433594 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 335.570312 C 114.722656 338.714844 112.175781 341.261719 109.035156 341.261719 C 105.890625 341.261719 103.34375 338.714844 103.34375 335.570312 C 103.34375 332.429688 105.890625 329.882812 109.035156 329.882812 C 112.175781 329.882812 114.722656 332.429688 114.722656 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 335.570312 C 128.664062 338.714844 126.117188 341.261719 122.972656 341.261719 C 119.832031 341.261719 117.285156 338.714844 117.285156 335.570312 C 117.285156 332.429688 119.832031 329.882812 122.972656 329.882812 C 126.117188 329.882812 128.664062 332.429688 128.664062 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 335.570312 C 128.664062 338.714844 126.117188 341.261719 122.972656 341.261719 C 119.832031 341.261719 117.285156 338.714844 117.285156 335.570312 C 117.285156 332.429688 119.832031 329.882812 122.972656 329.882812 C 126.117188 329.882812 128.664062 332.429688 128.664062 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 355.269531 C 114.722656 358.410156 112.175781 360.960938 109.035156 360.960938 C 105.890625 360.960938 103.34375 358.410156 103.34375 355.269531 C 103.34375 352.128906 105.890625 349.582031 109.035156 349.582031 C 112.175781 349.582031 114.722656 352.128906 114.722656 355.269531 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 355.269531 C 114.722656 358.410156 112.175781 360.960938 109.035156 360.960938 C 105.890625 360.960938 103.34375 358.410156 103.34375 355.269531 C 103.34375 352.128906 105.890625 349.582031 109.035156 349.582031 C 112.175781 349.582031 114.722656 352.128906 114.722656 355.269531 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 335.570312 C 114.722656 338.714844 112.175781 341.261719 109.035156 341.261719 C 105.890625 341.261719 103.34375 338.714844 103.34375 335.570312 C 103.34375 332.429688 105.890625 329.882812 109.035156 329.882812 C 112.175781 329.882812 114.722656 332.429688 114.722656 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 376.238281 C 86.84375 379.382812 84.296875 381.929688 81.152344 381.929688 C 78.011719 381.929688 75.464844 379.382812 75.464844 376.238281 C 75.464844 373.097656 78.011719 370.550781 81.152344 370.550781 C 84.296875 370.550781 86.84375 373.097656 86.84375 376.238281 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 335.570312 C 114.722656 338.714844 112.175781 341.261719 109.035156 341.261719 C 105.890625 341.261719 103.34375 338.714844 103.34375 335.570312 C 103.34375 332.429688 105.890625 329.882812 109.035156 329.882812 C 112.175781 329.882812 114.722656 332.429688 114.722656 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 210.265625 C 184.425781 213.40625 181.878906 215.953125 178.734375 215.953125 C 175.59375 215.953125 173.046875 213.40625 173.046875 210.265625 C 173.046875 207.125 175.59375 204.574219 178.734375 204.574219 C 181.878906 204.574219 184.425781 207.125 184.425781 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 197.523438 C 170.484375 200.664062 167.9375 203.210938 164.796875 203.210938 C 161.652344 203.210938 159.105469 200.664062 159.105469 197.523438 C 159.105469 194.378906 161.652344 191.832031 164.796875 191.832031 C 167.9375 191.832031 170.484375 194.378906 170.484375 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 223.527344 C 156.542969 226.671875 153.996094 229.21875 150.855469 229.21875 C 147.714844 229.21875 145.164062 226.671875 145.164062 223.527344 C 145.164062 220.386719 147.714844 217.839844 150.855469 217.839844 C 153.996094 217.839844 156.542969 220.386719 156.542969 223.527344 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 266.914062 C 142.605469 270.058594 140.058594 272.605469 136.914062 272.605469 C 133.773438 272.605469 131.226562 270.058594 131.226562 266.914062 C 131.226562 263.773438 133.773438 261.226562 136.914062 261.226562 C 140.058594 261.226562 142.605469 263.773438 142.605469 266.914062 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 251.800781 C 142.605469 254.941406 140.058594 257.488281 136.914062 257.488281 C 133.773438 257.488281 131.226562 254.941406 131.226562 251.800781 C 131.226562 248.65625 133.773438 246.109375 136.914062 246.109375 C 140.058594 246.109375 142.605469 248.65625 142.605469 251.800781 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 237.355469 C 142.605469 240.5 140.058594 243.046875 136.914062 243.046875 C 133.773438 243.046875 131.226562 240.5 131.226562 237.355469 C 131.226562 234.214844 133.773438 231.667969 136.914062 231.667969 C 140.058594 231.667969 142.605469 234.214844 142.605469 237.355469 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 251.800781 C 142.605469 254.941406 140.058594 257.488281 136.914062 257.488281 C 133.773438 257.488281 131.226562 254.941406 131.226562 251.800781 C 131.226562 248.65625 133.773438 246.109375 136.914062 246.109375 C 140.058594 246.109375 142.605469 248.65625 142.605469 251.800781 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 282.765625 C 128.664062 285.910156 126.117188 288.457031 122.972656 288.457031 C 119.832031 288.457031 117.285156 285.910156 117.285156 282.765625 C 117.285156 279.625 119.832031 277.078125 122.972656 277.078125 C 126.117188 277.078125 128.664062 279.625 128.664062 282.765625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 323.828125 120.058594 C 323.828125 123.199219 321.28125 125.75 318.140625 125.75 C 314.996094 125.75 312.449219 123.199219 312.449219 120.058594 C 312.449219 116.917969 314.996094 114.371094 318.140625 114.371094 C 321.28125 114.371094 323.828125 116.917969 323.828125 120.058594 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 268.066406 130.058594 C 268.066406 133.199219 265.519531 135.746094 262.378906 135.746094 C 259.234375 135.746094 256.6875 133.199219 256.6875 130.058594 C 256.6875 126.914062 259.234375 124.367188 262.378906 124.367188 C 265.519531 124.367188 268.066406 126.914062 268.066406 130.058594 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 282.007812 130.058594 C 282.007812 133.199219 279.460938 135.746094 276.316406 135.746094 C 273.175781 135.746094 270.628906 133.199219 270.628906 130.058594 C 270.628906 126.914062 273.175781 124.367188 276.316406 124.367188 C 279.460938 124.367188 282.007812 126.914062 282.007812 130.058594 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 254.125 162.042969 C 254.125 165.183594 251.578125 167.730469 248.4375 167.730469 C 245.296875 167.730469 242.75 165.183594 242.75 162.042969 C 242.75 158.898438 245.296875 156.351562 248.4375 156.351562 C 251.578125 156.351562 254.125 158.898438 254.125 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 268.066406 130.058594 C 268.066406 133.199219 265.519531 135.746094 262.378906 135.746094 C 259.234375 135.746094 256.6875 133.199219 256.6875 130.058594 C 256.6875 126.914062 259.234375 124.367188 262.378906 124.367188 C 265.519531 124.367188 268.066406 126.914062 268.066406 130.058594 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 295.949219 110.359375 C 295.949219 113.5 293.398438 116.046875 290.257812 116.046875 C 287.117188 116.046875 284.570312 113.5 284.570312 110.359375 C 284.570312 107.21875 287.117188 104.671875 290.257812 104.671875 C 293.398438 104.671875 295.949219 107.21875 295.949219 110.359375 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 282.007812 91.789062 C 282.007812 94.929688 279.460938 97.476562 276.316406 97.476562 C 273.175781 97.476562 270.628906 94.929688 270.628906 91.789062 C 270.628906 88.644531 273.175781 86.097656 276.316406 86.097656 C 279.460938 86.097656 282.007812 88.644531 282.007812 91.789062 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 268.066406 91.789062 C 268.066406 94.929688 265.519531 97.476562 262.378906 97.476562 C 259.234375 97.476562 256.6875 94.929688 256.6875 91.789062 C 256.6875 88.644531 259.234375 86.097656 262.378906 86.097656 C 265.519531 86.097656 268.066406 88.644531 268.066406 91.789062 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 185.257812 C 184.425781 188.402344 181.878906 190.949219 178.734375 190.949219 C 175.59375 190.949219 173.046875 188.402344 173.046875 185.257812 C 173.046875 182.117188 175.59375 179.570312 178.734375 179.570312 C 181.878906 179.570312 184.425781 182.117188 184.425781 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 151.027344 C 226.246094 154.167969 223.699219 156.714844 220.558594 156.714844 C 217.414062 156.714844 214.867188 154.167969 214.867188 151.027344 C 214.867188 147.882812 217.414062 145.335938 220.558594 145.335938 C 223.699219 145.335938 226.246094 147.882812 226.246094 151.027344 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 140.371094 C 226.246094 143.515625 223.699219 146.0625 220.558594 146.0625 C 217.414062 146.0625 214.867188 143.515625 214.867188 140.371094 C 214.867188 137.230469 217.414062 134.683594 220.558594 134.683594 C 223.699219 134.683594 226.246094 137.230469 226.246094 140.371094 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 173.441406 C 198.367188 176.585938 195.816406 179.132812 192.675781 179.132812 C 189.535156 179.132812 186.988281 176.585938 186.988281 173.441406 C 186.988281 170.300781 189.535156 167.753906 192.675781 167.753906 C 195.816406 167.753906 198.367188 170.300781 198.367188 173.441406 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 197.523438 C 198.367188 200.664062 195.816406 203.210938 192.675781 203.210938 C 189.535156 203.210938 186.988281 200.664062 186.988281 197.523438 C 186.988281 194.378906 189.535156 191.832031 192.675781 191.832031 C 195.816406 191.832031 198.367188 194.378906 198.367188 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 162.042969 C 198.367188 165.183594 195.816406 167.730469 192.675781 167.730469 C 189.535156 167.730469 186.988281 165.183594 186.988281 162.042969 C 186.988281 158.898438 189.535156 156.351562 192.675781 156.351562 C 195.816406 156.351562 198.367188 158.898438 198.367188 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 212.304688 173.441406 C 212.304688 176.585938 209.757812 179.132812 206.617188 179.132812 C 203.472656 179.132812 200.925781 176.585938 200.925781 173.441406 C 200.925781 170.300781 203.472656 167.753906 206.617188 167.753906 C 209.757812 167.753906 212.304688 170.300781 212.304688 173.441406 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 212.304688 185.257812 C 212.304688 188.402344 209.757812 190.949219 206.617188 190.949219 C 203.472656 190.949219 200.925781 188.402344 200.925781 185.257812 C 200.925781 182.117188 203.472656 179.570312 206.617188 179.570312 C 209.757812 179.570312 212.304688 182.117188 212.304688 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 223.527344 C 170.484375 226.671875 167.9375 229.21875 164.796875 229.21875 C 161.652344 229.21875 159.105469 226.671875 159.105469 223.527344 C 159.105469 220.386719 161.652344 217.839844 164.796875 217.839844 C 167.9375 217.839844 170.484375 220.386719 170.484375 223.527344 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 223.527344 C 156.542969 226.671875 153.996094 229.21875 150.855469 229.21875 C 147.714844 229.21875 145.164062 226.671875 145.164062 223.527344 C 145.164062 220.386719 147.714844 217.839844 150.855469 217.839844 C 153.996094 217.839844 156.542969 220.386719 156.542969 223.527344 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 223.527344 C 170.484375 226.671875 167.9375 229.21875 164.796875 229.21875 C 161.652344 229.21875 159.105469 226.671875 159.105469 223.527344 C 159.105469 220.386719 161.652344 217.839844 164.796875 217.839844 C 167.9375 217.839844 170.484375 220.386719 170.484375 223.527344 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 251.800781 C 170.484375 254.941406 167.9375 257.488281 164.796875 257.488281 C 161.652344 257.488281 159.105469 254.941406 159.105469 251.800781 C 159.105469 248.65625 161.652344 246.109375 164.796875 246.109375 C 167.9375 246.109375 170.484375 248.65625 170.484375 251.800781 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 299.433594 C 142.605469 302.574219 140.058594 305.121094 136.914062 305.121094 C 133.773438 305.121094 131.226562 302.574219 131.226562 299.433594 C 131.226562 296.292969 133.773438 293.746094 136.914062 293.746094 C 140.058594 293.746094 142.605469 296.292969 142.605469 299.433594 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 282.765625 C 142.605469 285.910156 140.058594 288.457031 136.914062 288.457031 C 133.773438 288.457031 131.226562 285.910156 131.226562 282.765625 C 131.226562 279.625 133.773438 277.078125 136.914062 277.078125 C 140.058594 277.078125 142.605469 279.625 142.605469 282.765625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 335.570312 C 128.664062 338.714844 126.117188 341.261719 122.972656 341.261719 C 119.832031 341.261719 117.285156 338.714844 117.285156 335.570312 C 117.285156 332.429688 119.832031 329.882812 122.972656 329.882812 C 126.117188 329.882812 128.664062 332.429688 128.664062 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 58.960938 448.742188 C 58.960938 451.882812 56.414062 454.429688 53.273438 454.429688 C 50.132812 454.429688 47.582031 451.882812 47.582031 448.742188 C 47.582031 445.597656 50.132812 443.050781 53.273438 443.050781 C 56.414062 443.050781 58.960938 445.597656 58.960938 448.742188 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 299.433594 C 128.664062 302.574219 126.117188 305.121094 122.972656 305.121094 C 119.832031 305.121094 117.285156 302.574219 117.285156 299.433594 C 117.285156 296.292969 119.832031 293.746094 122.972656 293.746094 C 126.117188 293.746094 128.664062 296.292969 128.664062 299.433594 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 317 C 114.722656 320.144531 112.175781 322.691406 109.035156 322.691406 C 105.890625 322.691406 103.34375 320.144531 103.34375 317 C 103.34375 313.859375 105.890625 311.3125 109.035156 311.3125 C 112.175781 311.3125 114.722656 313.859375 114.722656 317 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 398.65625 C 86.84375 401.796875 84.296875 404.34375 81.152344 404.34375 C 78.011719 404.34375 75.464844 401.796875 75.464844 398.65625 C 75.464844 395.515625 78.011719 392.964844 81.152344 392.964844 C 84.296875 392.964844 86.84375 395.515625 86.84375 398.65625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 376.238281 C 86.84375 379.382812 84.296875 381.929688 81.152344 381.929688 C 78.011719 381.929688 75.464844 379.382812 75.464844 376.238281 C 75.464844 373.097656 78.011719 370.550781 81.152344 370.550781 C 84.296875 370.550781 86.84375 373.097656 86.84375 376.238281 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.78125 317 C 100.78125 320.144531 98.234375 322.691406 95.09375 322.691406 C 91.953125 322.691406 89.40625 320.144531 89.40625 317 C 89.40625 313.859375 91.953125 311.3125 95.09375 311.3125 C 98.234375 311.3125 100.78125 313.859375 100.78125 317 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.78125 317 C 100.78125 320.144531 98.234375 322.691406 95.09375 322.691406 C 91.953125 322.691406 89.40625 320.144531 89.40625 317 C 89.40625 313.859375 91.953125 311.3125 95.09375 311.3125 C 98.234375 311.3125 100.78125 313.859375 100.78125 317 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 376.238281 C 86.84375 379.382812 84.296875 381.929688 81.152344 381.929688 C 78.011719 381.929688 75.464844 379.382812 75.464844 376.238281 C 75.464844 373.097656 78.011719 370.550781 81.152344 370.550781 C 84.296875 370.550781 86.84375 373.097656 86.84375 376.238281 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 335.570312 C 86.84375 338.714844 84.296875 341.261719 81.152344 341.261719 C 78.011719 341.261719 75.464844 338.714844 75.464844 335.570312 C 75.464844 332.429688 78.011719 329.882812 81.152344 329.882812 C 84.296875 329.882812 86.84375 332.429688 86.84375 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 355.269531 C 86.84375 358.410156 84.296875 360.960938 81.152344 360.960938 C 78.011719 360.960938 75.464844 358.410156 75.464844 355.269531 C 75.464844 352.128906 78.011719 349.582031 81.152344 349.582031 C 84.296875 349.582031 86.84375 352.128906 86.84375 355.269531 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.78125 317 C 100.78125 320.144531 98.234375 322.691406 95.09375 322.691406 C 91.953125 322.691406 89.40625 320.144531 89.40625 317 C 89.40625 313.859375 91.953125 311.3125 95.09375 311.3125 C 98.234375 311.3125 100.78125 313.859375 100.78125 317 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 335.570312 C 128.664062 338.714844 126.117188 341.261719 122.972656 341.261719 C 119.832031 341.261719 117.285156 338.714844 117.285156 335.570312 C 117.285156 332.429688 119.832031 329.882812 122.972656 329.882812 C 126.117188 329.882812 128.664062 332.429688 128.664062 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 299.433594 C 114.722656 302.574219 112.175781 305.121094 109.035156 305.121094 C 105.890625 305.121094 103.34375 302.574219 103.34375 299.433594 C 103.34375 296.292969 105.890625 293.746094 109.035156 293.746094 C 112.175781 293.746094 114.722656 296.292969 114.722656 299.433594 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 299.433594 C 114.722656 302.574219 112.175781 305.121094 109.035156 305.121094 C 105.890625 305.121094 103.34375 302.574219 103.34375 299.433594 C 103.34375 296.292969 105.890625 293.746094 109.035156 293.746094 C 112.175781 293.746094 114.722656 296.292969 114.722656 299.433594 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 335.570312 C 114.722656 338.714844 112.175781 341.261719 109.035156 341.261719 C 105.890625 341.261719 103.34375 338.714844 103.34375 335.570312 C 103.34375 332.429688 105.890625 329.882812 109.035156 329.882812 C 112.175781 329.882812 114.722656 332.429688 114.722656 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 185.257812 C 198.367188 188.402344 195.816406 190.949219 192.675781 190.949219 C 189.535156 190.949219 186.988281 188.402344 186.988281 185.257812 C 186.988281 182.117188 189.535156 179.570312 192.675781 179.570312 C 195.816406 179.570312 198.367188 182.117188 198.367188 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 254.125 140.371094 C 254.125 143.515625 251.578125 146.0625 248.4375 146.0625 C 245.296875 146.0625 242.75 143.515625 242.75 140.371094 C 242.75 137.230469 245.296875 134.683594 248.4375 134.683594 C 251.578125 134.683594 254.125 137.230469 254.125 140.371094 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 254.125 130.058594 C 254.125 133.199219 251.578125 135.746094 248.4375 135.746094 C 245.296875 135.746094 242.75 133.199219 242.75 130.058594 C 242.75 126.914062 245.296875 124.367188 248.4375 124.367188 C 251.578125 124.367188 254.125 126.914062 254.125 130.058594 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 185.257812 C 198.367188 188.402344 195.816406 190.949219 192.675781 190.949219 C 189.535156 190.949219 186.988281 188.402344 186.988281 185.257812 C 186.988281 182.117188 189.535156 179.570312 192.675781 179.570312 C 195.816406 179.570312 198.367188 182.117188 198.367188 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 197.523438 C 198.367188 200.664062 195.816406 203.210938 192.675781 203.210938 C 189.535156 203.210938 186.988281 200.664062 186.988281 197.523438 C 186.988281 194.378906 189.535156 191.832031 192.675781 191.832031 C 195.816406 191.832031 198.367188 194.378906 198.367188 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 210.265625 C 198.367188 213.40625 195.816406 215.953125 192.675781 215.953125 C 189.535156 215.953125 186.988281 213.40625 186.988281 210.265625 C 186.988281 207.125 189.535156 204.574219 192.675781 204.574219 C 195.816406 204.574219 198.367188 207.125 198.367188 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 210.265625 C 198.367188 213.40625 195.816406 215.953125 192.675781 215.953125 C 189.535156 215.953125 186.988281 213.40625 186.988281 210.265625 C 186.988281 207.125 189.535156 204.574219 192.675781 204.574219 C 195.816406 204.574219 198.367188 207.125 198.367188 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 335.570312 C 128.664062 338.714844 126.117188 341.261719 122.972656 341.261719 C 119.832031 341.261719 117.285156 338.714844 117.285156 335.570312 C 117.285156 332.429688 119.832031 329.882812 122.972656 329.882812 C 126.117188 329.882812 128.664062 332.429688 128.664062 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 335.570312 C 142.605469 338.714844 140.058594 341.261719 136.914062 341.261719 C 133.773438 341.261719 131.226562 338.714844 131.226562 335.570312 C 131.226562 332.429688 133.773438 329.882812 136.914062 329.882812 C 140.058594 329.882812 142.605469 332.429688 142.605469 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 282.765625 C 128.664062 285.910156 126.117188 288.457031 122.972656 288.457031 C 119.832031 288.457031 117.285156 285.910156 117.285156 282.765625 C 117.285156 279.625 119.832031 277.078125 122.972656 277.078125 C 126.117188 277.078125 128.664062 279.625 128.664062 282.765625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 100.78125 317 C 100.78125 320.144531 98.234375 322.691406 95.09375 322.691406 C 91.953125 322.691406 89.40625 320.144531 89.40625 317 C 89.40625 313.859375 91.953125 311.3125 95.09375 311.3125 C 98.234375 311.3125 100.78125 313.859375 100.78125 317 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 197.523438 C 156.542969 200.664062 153.996094 203.210938 150.855469 203.210938 C 147.714844 203.210938 145.164062 200.664062 145.164062 197.523438 C 145.164062 194.378906 147.714844 191.832031 150.855469 191.832031 C 153.996094 191.832031 156.542969 194.378906 156.542969 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 185.257812 C 170.484375 188.402344 167.9375 190.949219 164.796875 190.949219 C 161.652344 190.949219 159.105469 188.402344 159.105469 185.257812 C 159.105469 182.117188 161.652344 179.570312 164.796875 179.570312 C 167.9375 179.570312 170.484375 182.117188 170.484375 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 173.441406 C 184.425781 176.585938 181.878906 179.132812 178.734375 179.132812 C 175.59375 179.132812 173.046875 176.585938 173.046875 173.441406 C 173.046875 170.300781 175.59375 167.753906 178.734375 167.753906 C 181.878906 167.753906 184.425781 170.300781 184.425781 173.441406 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 210.265625 C 156.542969 213.40625 153.996094 215.953125 150.855469 215.953125 C 147.714844 215.953125 145.164062 213.40625 145.164062 210.265625 C 145.164062 207.125 147.714844 204.574219 150.855469 204.574219 C 153.996094 204.574219 156.542969 207.125 156.542969 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 210.265625 C 184.425781 213.40625 181.878906 215.953125 178.734375 215.953125 C 175.59375 215.953125 173.046875 213.40625 173.046875 210.265625 C 173.046875 207.125 175.59375 204.574219 178.734375 204.574219 C 181.878906 204.574219 184.425781 207.125 184.425781 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 223.527344 C 184.425781 226.671875 181.878906 229.21875 178.734375 229.21875 C 175.59375 229.21875 173.046875 226.671875 173.046875 223.527344 C 173.046875 220.386719 175.59375 217.839844 178.734375 217.839844 C 181.878906 217.839844 184.425781 220.386719 184.425781 223.527344 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 212.304688 185.257812 C 212.304688 188.402344 209.757812 190.949219 206.617188 190.949219 C 203.472656 190.949219 200.925781 188.402344 200.925781 185.257812 C 200.925781 182.117188 203.472656 179.570312 206.617188 179.570312 C 209.757812 179.570312 212.304688 182.117188 212.304688 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 210.265625 C 198.367188 213.40625 195.816406 215.953125 192.675781 215.953125 C 189.535156 215.953125 186.988281 213.40625 186.988281 210.265625 C 186.988281 207.125 189.535156 204.574219 192.675781 204.574219 C 195.816406 204.574219 198.367188 207.125 198.367188 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 212.304688 197.523438 C 212.304688 200.664062 209.757812 203.210938 206.617188 203.210938 C 203.472656 203.210938 200.925781 200.664062 200.925781 197.523438 C 200.925781 194.378906 203.472656 191.832031 206.617188 191.832031 C 209.757812 191.832031 212.304688 194.378906 212.304688 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 237.355469 C 184.425781 240.5 181.878906 243.046875 178.734375 243.046875 C 175.59375 243.046875 173.046875 240.5 173.046875 237.355469 C 173.046875 234.214844 175.59375 231.667969 178.734375 231.667969 C 181.878906 231.667969 184.425781 234.214844 184.425781 237.355469 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 197.523438 C 226.246094 200.664062 223.699219 203.210938 220.558594 203.210938 C 217.414062 203.210938 214.867188 200.664062 214.867188 197.523438 C 214.867188 194.378906 217.414062 191.832031 220.558594 191.832031 C 223.699219 191.832031 226.246094 194.378906 226.246094 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 197.523438 C 198.367188 200.664062 195.816406 203.210938 192.675781 203.210938 C 189.535156 203.210938 186.988281 200.664062 186.988281 197.523438 C 186.988281 194.378906 189.535156 191.832031 192.675781 191.832031 C 195.816406 191.832031 198.367188 194.378906 198.367188 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 197.523438 C 198.367188 200.664062 195.816406 203.210938 192.675781 203.210938 C 189.535156 203.210938 186.988281 200.664062 186.988281 197.523438 C 186.988281 194.378906 189.535156 191.832031 192.675781 191.832031 C 195.816406 191.832031 198.367188 194.378906 198.367188 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 197.523438 C 198.367188 200.664062 195.816406 203.210938 192.675781 203.210938 C 189.535156 203.210938 186.988281 200.664062 186.988281 197.523438 C 186.988281 194.378906 189.535156 191.832031 192.675781 191.832031 C 195.816406 191.832031 198.367188 194.378906 198.367188 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 212.304688 210.265625 C 212.304688 213.40625 209.757812 215.953125 206.617188 215.953125 C 203.472656 215.953125 200.925781 213.40625 200.925781 210.265625 C 200.925781 207.125 203.472656 204.574219 206.617188 204.574219 C 209.757812 204.574219 212.304688 207.125 212.304688 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 212.304688 185.257812 C 212.304688 188.402344 209.757812 190.949219 206.617188 190.949219 C 203.472656 190.949219 200.925781 188.402344 200.925781 185.257812 C 200.925781 182.117188 203.472656 179.570312 206.617188 179.570312 C 209.757812 179.570312 212.304688 182.117188 212.304688 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 210.265625 C 198.367188 213.40625 195.816406 215.953125 192.675781 215.953125 C 189.535156 215.953125 186.988281 213.40625 186.988281 210.265625 C 186.988281 207.125 189.535156 204.574219 192.675781 204.574219 C 195.816406 204.574219 198.367188 207.125 198.367188 210.265625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 212.304688 185.257812 C 212.304688 188.402344 209.757812 190.949219 206.617188 190.949219 C 203.472656 190.949219 200.925781 188.402344 200.925781 185.257812 C 200.925781 182.117188 203.472656 179.570312 206.617188 179.570312 C 209.757812 179.570312 212.304688 182.117188 212.304688 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 282.765625 C 142.605469 285.910156 140.058594 288.457031 136.914062 288.457031 C 133.773438 288.457031 131.226562 285.910156 131.226562 282.765625 C 131.226562 279.625 133.773438 277.078125 136.914062 277.078125 C 140.058594 277.078125 142.605469 279.625 142.605469 282.765625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 282.765625 C 156.542969 285.910156 153.996094 288.457031 150.855469 288.457031 C 147.714844 288.457031 145.164062 285.910156 145.164062 282.765625 C 145.164062 279.625 147.714844 277.078125 150.855469 277.078125 C 153.996094 277.078125 156.542969 279.625 156.542969 282.765625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 299.433594 C 142.605469 302.574219 140.058594 305.121094 136.914062 305.121094 C 133.773438 305.121094 131.226562 302.574219 131.226562 299.433594 C 131.226562 296.292969 133.773438 293.746094 136.914062 293.746094 C 140.058594 293.746094 142.605469 296.292969 142.605469 299.433594 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 335.570312 C 142.605469 338.714844 140.058594 341.261719 136.914062 341.261719 C 133.773438 341.261719 131.226562 338.714844 131.226562 335.570312 C 131.226562 332.429688 133.773438 329.882812 136.914062 329.882812 C 140.058594 329.882812 142.605469 332.429688 142.605469 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 282.765625 C 156.542969 285.910156 153.996094 288.457031 150.855469 288.457031 C 147.714844 288.457031 145.164062 285.910156 145.164062 282.765625 C 145.164062 279.625 147.714844 277.078125 150.855469 277.078125 C 153.996094 277.078125 156.542969 279.625 156.542969 282.765625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 128.664062 335.570312 C 128.664062 338.714844 126.117188 341.261719 122.972656 341.261719 C 119.832031 341.261719 117.285156 338.714844 117.285156 335.570312 C 117.285156 332.429688 119.832031 329.882812 122.972656 329.882812 C 126.117188 329.882812 128.664062 332.429688 128.664062 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 185.257812 C 226.246094 188.402344 223.699219 190.949219 220.558594 190.949219 C 217.414062 190.949219 214.867188 188.402344 214.867188 185.257812 C 214.867188 182.117188 217.414062 179.570312 220.558594 179.570312 C 223.699219 179.570312 226.246094 182.117188 226.246094 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 140.371094 C 226.246094 143.515625 223.699219 146.0625 220.558594 146.0625 C 217.414062 146.0625 214.867188 143.515625 214.867188 140.371094 C 214.867188 137.230469 217.414062 134.683594 220.558594 134.683594 C 223.699219 134.683594 226.246094 137.230469 226.246094 140.371094 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 140.371094 C 226.246094 143.515625 223.699219 146.0625 220.558594 146.0625 C 217.414062 146.0625 214.867188 143.515625 214.867188 140.371094 C 214.867188 137.230469 217.414062 134.683594 220.558594 134.683594 C 223.699219 134.683594 226.246094 137.230469 226.246094 140.371094 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 173.441406 C 198.367188 176.585938 195.816406 179.132812 192.675781 179.132812 C 189.535156 179.132812 186.988281 176.585938 186.988281 173.441406 C 186.988281 170.300781 189.535156 167.753906 192.675781 167.753906 C 195.816406 167.753906 198.367188 170.300781 198.367188 173.441406 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 185.257812 C 226.246094 188.402344 223.699219 190.949219 220.558594 190.949219 C 217.414062 190.949219 214.867188 188.402344 214.867188 185.257812 C 214.867188 182.117188 217.414062 179.570312 220.558594 179.570312 C 223.699219 179.570312 226.246094 182.117188 226.246094 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 140.371094 C 226.246094 143.515625 223.699219 146.0625 220.558594 146.0625 C 217.414062 146.0625 214.867188 143.515625 214.867188 140.371094 C 214.867188 137.230469 217.414062 134.683594 220.558594 134.683594 C 223.699219 134.683594 226.246094 137.230469 226.246094 140.371094 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 240.1875 140.371094 C 240.1875 143.515625 237.640625 146.0625 234.496094 146.0625 C 231.355469 146.0625 228.808594 143.515625 228.808594 140.371094 C 228.808594 137.230469 231.355469 134.683594 234.496094 134.683594 C 237.640625 134.683594 240.1875 137.230469 240.1875 140.371094 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 185.257812 C 184.425781 188.402344 181.878906 190.949219 178.734375 190.949219 C 175.59375 190.949219 173.046875 188.402344 173.046875 185.257812 C 173.046875 182.117188 175.59375 179.570312 178.734375 179.570312 C 181.878906 179.570312 184.425781 182.117188 184.425781 185.257812 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 268.066406 151.027344 C 268.066406 154.167969 265.519531 156.714844 262.378906 156.714844 C 259.234375 156.714844 256.6875 154.167969 256.6875 151.027344 C 256.6875 147.882812 259.234375 145.335938 262.378906 145.335938 C 265.519531 145.335938 268.066406 147.882812 268.066406 151.027344 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 268.066406 120.058594 C 268.066406 123.199219 265.519531 125.75 262.378906 125.75 C 259.234375 125.75 256.6875 123.199219 256.6875 120.058594 C 256.6875 116.917969 259.234375 114.371094 262.378906 114.371094 C 265.519531 114.371094 268.066406 116.917969 268.066406 120.058594 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 295.949219 100.941406 C 295.949219 104.082031 293.398438 106.628906 290.257812 106.628906 C 287.117188 106.628906 284.570312 104.082031 284.570312 100.941406 C 284.570312 97.796875 287.117188 95.25 290.257812 95.25 C 293.398438 95.25 295.949219 97.796875 295.949219 100.941406 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 323.828125 82.886719 C 323.828125 86.027344 321.28125 88.574219 318.140625 88.574219 C 314.996094 88.574219 312.449219 86.027344 312.449219 82.886719 C 312.449219 79.742188 314.996094 77.195312 318.140625 77.195312 C 321.28125 77.195312 323.828125 79.742188 323.828125 82.886719 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 295.949219 100.941406 C 295.949219 104.082031 293.398438 106.628906 290.257812 106.628906 C 287.117188 106.628906 284.570312 104.082031 284.570312 100.941406 C 284.570312 97.796875 287.117188 95.25 290.257812 95.25 C 293.398438 95.25 295.949219 97.796875 295.949219 100.941406 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 86.84375 376.238281 C 86.84375 379.382812 84.296875 381.929688 81.152344 381.929688 C 78.011719 381.929688 75.464844 379.382812 75.464844 376.238281 C 75.464844 373.097656 78.011719 370.550781 81.152344 370.550781 C 84.296875 370.550781 86.84375 373.097656 86.84375 376.238281 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 114.722656 317 C 114.722656 320.144531 112.175781 322.691406 109.035156 322.691406 C 105.890625 322.691406 103.34375 320.144531 103.34375 317 C 103.34375 313.859375 105.890625 311.3125 109.035156 311.3125 C 112.175781 311.3125 114.722656 313.859375 114.722656 317 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 282.765625 C 142.605469 285.910156 140.058594 288.457031 136.914062 288.457031 C 133.773438 288.457031 131.226562 285.910156 131.226562 282.765625 C 131.226562 279.625 133.773438 277.078125 136.914062 277.078125 C 140.058594 277.078125 142.605469 279.625 142.605469 282.765625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 282.765625 C 156.542969 285.910156 153.996094 288.457031 150.855469 288.457031 C 147.714844 288.457031 145.164062 285.910156 145.164062 282.765625 C 145.164062 279.625 147.714844 277.078125 150.855469 277.078125 C 153.996094 277.078125 156.542969 279.625 156.542969 282.765625 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 251.800781 C 170.484375 254.941406 167.9375 257.488281 164.796875 257.488281 C 161.652344 257.488281 159.105469 254.941406 159.105469 251.800781 C 159.105469 248.65625 161.652344 246.109375 164.796875 246.109375 C 167.9375 246.109375 170.484375 248.65625 170.484375 251.800781 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 335.570312 C 142.605469 338.714844 140.058594 341.261719 136.914062 341.261719 C 133.773438 341.261719 131.226562 338.714844 131.226562 335.570312 C 131.226562 332.429688 133.773438 329.882812 136.914062 329.882812 C 140.058594 329.882812 142.605469 332.429688 142.605469 335.570312 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 299.433594 C 142.605469 302.574219 140.058594 305.121094 136.914062 305.121094 C 133.773438 305.121094 131.226562 302.574219 131.226562 299.433594 C 131.226562 296.292969 133.773438 293.746094 136.914062 293.746094 C 140.058594 293.746094 142.605469 296.292969 142.605469 299.433594 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 142.605469 317 C 142.605469 320.144531 140.058594 322.691406 136.914062 322.691406 C 133.773438 322.691406 131.226562 320.144531 131.226562 317 C 131.226562 313.859375 133.773438 311.3125 136.914062 311.3125 C 140.058594 311.3125 142.605469 313.859375 142.605469 317 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 282.765625 C 156.542969 285.910156 153.996094 288.457031 150.855469 288.457031 C 147.714844 288.457031 145.164062 285.910156 145.164062 282.765625 C 145.164062 279.625 147.714844 277.078125 150.855469 277.078125 C 153.996094 277.078125 156.542969 279.625 156.542969 282.765625 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 197.523438 C 198.367188 200.664062 195.816406 203.210938 192.675781 203.210938 C 189.535156 203.210938 186.988281 200.664062 186.988281 197.523438 C 186.988281 194.378906 189.535156 191.832031 192.675781 191.832031 C 195.816406 191.832031 198.367188 194.378906 198.367188 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 240.1875 162.042969 C 240.1875 165.183594 237.640625 167.730469 234.496094 167.730469 C 231.355469 167.730469 228.808594 165.183594 228.808594 162.042969 C 228.808594 158.898438 231.355469 156.351562 234.496094 156.351562 C 237.640625 156.351562 240.1875 158.898438 240.1875 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 223.527344 C 170.484375 226.671875 167.9375 229.21875 164.796875 229.21875 C 161.652344 229.21875 159.105469 226.671875 159.105469 223.527344 C 159.105469 220.386719 161.652344 217.839844 164.796875 217.839844 C 167.9375 217.839844 170.484375 220.386719 170.484375 223.527344 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 393.53125 26.585938 C 393.53125 29.730469 390.980469 32.277344 387.839844 32.277344 C 384.699219 32.277344 382.152344 29.730469 382.152344 26.585938 C 382.152344 23.445312 384.699219 20.898438 387.839844 20.898438 C 390.980469 20.898438 393.53125 23.445312 393.53125 26.585938 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 197.523438 C 198.367188 200.664062 195.816406 203.210938 192.675781 203.210938 C 189.535156 203.210938 186.988281 200.664062 186.988281 197.523438 C 186.988281 194.378906 189.535156 191.832031 192.675781 191.832031 C 195.816406 191.832031 198.367188 194.378906 198.367188 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 240.1875 162.042969 C 240.1875 165.183594 237.640625 167.730469 234.496094 167.730469 C 231.355469 167.730469 228.808594 165.183594 228.808594 162.042969 C 228.808594 158.898438 231.355469 156.351562 234.496094 156.351562 C 237.640625 156.351562 240.1875 158.898438 240.1875 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 237.355469 C 156.542969 240.5 153.996094 243.046875 150.855469 243.046875 C 147.714844 243.046875 145.164062 240.5 145.164062 237.355469 C 145.164062 234.214844 147.714844 231.667969 150.855469 231.667969 C 153.996094 231.667969 156.542969 234.214844 156.542969 237.355469 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 223.527344 C 170.484375 226.671875 167.9375 229.21875 164.796875 229.21875 C 161.652344 229.21875 159.105469 226.671875 159.105469 223.527344 C 159.105469 220.386719 161.652344 217.839844 164.796875 217.839844 C 167.9375 217.839844 170.484375 220.386719 170.484375 223.527344 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 421.410156 26.585938 C 421.410156 29.730469 418.863281 32.277344 415.722656 32.277344 C 412.578125 32.277344 410.03125 29.730469 410.03125 26.585938 C 410.03125 23.445312 412.578125 20.898438 415.722656 20.898438 C 418.863281 20.898438 421.410156 23.445312 421.410156 26.585938 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 337.769531 49.53125 C 337.769531 52.675781 335.222656 55.222656 332.078125 55.222656 C 328.9375 55.222656 326.390625 52.675781 326.390625 49.53125 C 326.390625 46.390625 328.9375 43.84375 332.078125 43.84375 C 335.222656 43.84375 337.769531 46.390625 337.769531 49.53125 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 197.523438 C 198.367188 200.664062 195.816406 203.210938 192.675781 203.210938 C 189.535156 203.210938 186.988281 200.664062 186.988281 197.523438 C 186.988281 194.378906 189.535156 191.832031 192.675781 191.832031 C 195.816406 191.832031 198.367188 194.378906 198.367188 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 212.304688 173.441406 C 212.304688 176.585938 209.757812 179.132812 206.617188 179.132812 C 203.472656 179.132812 200.925781 176.585938 200.925781 173.441406 C 200.925781 170.300781 203.472656 167.753906 206.617188 167.753906 C 209.757812 167.753906 212.304688 170.300781 212.304688 173.441406 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 212.304688 162.042969 C 212.304688 165.183594 209.757812 167.730469 206.617188 167.730469 C 203.472656 167.730469 200.925781 165.183594 200.925781 162.042969 C 200.925781 158.898438 203.472656 156.351562 206.617188 156.351562 C 209.757812 156.351562 212.304688 158.898438 212.304688 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 162.042969 C 184.425781 165.183594 181.878906 167.730469 178.734375 167.730469 C 175.59375 167.730469 173.046875 165.183594 173.046875 162.042969 C 173.046875 158.898438 175.59375 156.351562 178.734375 156.351562 C 181.878906 156.351562 184.425781 158.898438 184.425781 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 198.367188 173.441406 C 198.367188 176.585938 195.816406 179.132812 192.675781 179.132812 C 189.535156 179.132812 186.988281 176.585938 186.988281 173.441406 C 186.988281 170.300781 189.535156 167.753906 192.675781 167.753906 C 195.816406 167.753906 198.367188 170.300781 198.367188 173.441406 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 226.246094 162.042969 C 226.246094 165.183594 223.699219 167.730469 220.558594 167.730469 C 217.414062 167.730469 214.867188 165.183594 214.867188 162.042969 C 214.867188 158.898438 217.414062 156.351562 220.558594 156.351562 C 223.699219 156.351562 226.246094 158.898438 226.246094 162.042969 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 156.542969 197.523438 C 156.542969 200.664062 153.996094 203.210938 150.855469 203.210938 C 147.714844 203.210938 145.164062 200.664062 145.164062 197.523438 C 145.164062 194.378906 147.714844 191.832031 150.855469 191.832031 C 153.996094 191.832031 156.542969 194.378906 156.542969 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(7.45098%,16.862745%,26.27451%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(7.45098%,16.862745%,26.27451%);stroke-opacity:1;stroke-miterlimit:10;" d="M 184.425781 197.523438 C 184.425781 200.664062 181.878906 203.210938 178.734375 203.210938 C 175.59375 203.210938 173.046875 200.664062 173.046875 197.523438 C 173.046875 194.378906 175.59375 191.832031 178.734375 191.832031 C 181.878906 191.832031 184.425781 194.378906 184.425781 197.523438 "/>
<path style="fill-rule:nonzero;fill:rgb(33.72549%,69.411765%,96.862745%);fill-opacity:1;stroke-width:0.708661;stroke-linecap:round;stroke-linejoin:round;stroke:rgb(33.72549%,69.411765%,96.862745%);stroke-opacity:1;stroke-miterlimit:10;" d="M 170.484375 197.523438 C 170.484375 200.664062 167.9375 203.210938 164.796875 203.210938 C 161.652344 203.210938 159.105469 200.664062 159.105469 197.523438 C 159.105469 194.378906 161.652344 191.832031 164.796875 191.832031 C 167.9375 191.832031 170.484375 194.378906 170.484375 197.523438 "/>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-1" x="20.21875" y="285.680664"/>
  <use xlink:href="#glyph0-2" x="25.21875" y="285.680664"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-3" x="20.21875" y="153.942383"/>
  <use xlink:href="#glyph0-2" x="25.21875" y="153.942383"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-4" x="20.21875" y="60.469727"/>
  <use xlink:href="#glyph0-2" x="25.21875" y="60.469727"/>
</g>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 282.765625 L 35.152344 282.765625 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 151.027344 L 35.152344 151.027344 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 32.410156 57.554688 L 35.152344 57.554688 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 67.214844 472.589844 L 67.214844 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 136.914062 472.589844 L 136.914062 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 206.617188 472.589844 L 206.617188 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 276.316406 472.589844 L 276.316406 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 346.019531 472.589844 L 346.019531 469.847656 "/>
<path style="fill:none;stroke-width:1.066978;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(20%,20%,20%);stroke-opacity:1;stroke-miterlimit:10;" d="M 415.722656 472.589844 L 415.722656 469.847656 "/>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="62.214844" y="481.196289"/>
  <use xlink:href="#glyph0-2" x="67.214844" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-5" x="131.914062" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="136.914062" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-1" x="201.617188" y="481.196289"/>
  <use xlink:href="#glyph0-2" x="206.617188" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-1" x="271.316406" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="276.316406" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-3" x="341.019531" y="481.196289"/>
  <use xlink:href="#glyph0-2" x="346.019531" y="481.196289"/>
</g>
<g style="fill:rgb(30.196078%,30.196078%,30.196078%);fill-opacity:1;">
  <use xlink:href="#glyph0-3" x="410.722656" y="481.196289"/>
  <use xlink:href="#glyph0-6" x="415.722656" y="481.196289"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph1-1" x="227.496094" y="494.538086"/>
  <use xlink:href="#glyph1-2" x="233.496094" y="494.538086"/>
  <use xlink:href="#glyph1-3" x="236.496094" y="494.538086"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph2-1" x="13.499023" y="247.164062"/>
  <use xlink:href="#glyph2-2" x="13.499023" y="241.164062"/>
  <use xlink:href="#glyph2-3" x="13.499023" y="233.164062"/>
</g>
<path style=" stroke:none;fill-rule:nonzero;fill:rgb(100%,100%,100%);fill-opacity:1;" d="M 444.800781 295.082031 L 498.519531 295.082031 L 498.519531 180.242188 L 444.800781 180.242188 Z M 444.800781 295.082031 "/>
<g clip-path="url(#clip20)" clip-rule="nonzero">
<rect x="0" y="0" width="504" height="504" style="fill:url(#pattern0);stroke:none;"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph0-1" x="473.042969" y="282.805664"/>
  <use xlink:href="#glyph0-2" x="478.042969" y="282.805664"/>
  <use xlink:href="#glyph0-2" x="483.042969" y="282.805664"/>
  <use xlink:href="#glyph0-2" x="488.042969" y="282.805664"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph0-1" x="473.042969" y="263.672852"/>
  <use xlink:href="#glyph0-2" x="478.042969" y="263.672852"/>
  <use xlink:href="#glyph0-2" x="483.042969" y="263.672852"/>
  <use xlink:href="#glyph0-1" x="488.042969" y="263.672852"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph0-1" x="473.042969" y="244.536133"/>
  <use xlink:href="#glyph0-2" x="478.042969" y="244.536133"/>
  <use xlink:href="#glyph0-2" x="483.042969" y="244.536133"/>
  <use xlink:href="#glyph0-4" x="488.042969" y="244.536133"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph0-1" x="473.042969" y="225.399414"/>
  <use xlink:href="#glyph0-2" x="478.042969" y="225.399414"/>
  <use xlink:href="#glyph0-2" x="483.042969" y="225.399414"/>
  <use xlink:href="#glyph0-7" x="488.042969" y="225.399414"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph0-1" x="473.042969" y="206.262695"/>
  <use xlink:href="#glyph0-2" x="478.042969" y="206.262695"/>
  <use xlink:href="#glyph0-2" x="483.042969" y="206.262695"/>
  <use xlink:href="#glyph0-8" x="488.042969" y="206.262695"/>
</g>
<g style="fill:rgb(0%,0%,0%);fill-opacity:1;">
  <use xlink:href="#glyph1-3" x="450.28125" y="195.245117"/>
  <use xlink:href="#glyph1-4" x="455.28125" y="195.245117"/>
  <use xlink:href="#glyph1-5" x="461.28125" y="195.245117"/>
  <use xlink:href="#glyph1-6" x="467.28125" y="195.245117"/>
</g>
<path style="fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 450.28125 279.890625 L 453.738281 279.890625 "/>
<path style="fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 450.28125 260.757812 L 453.738281 260.757812 "/>
<path style="fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 450.28125 241.621094 L 453.738281 241.621094 "/>
<path style="fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 450.28125 222.484375 L 453.738281 222.484375 "/>
<path style="fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 450.28125 203.347656 L 453.738281 203.347656 "/>
<path style="fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 464.105469 279.890625 L 467.5625 279.890625 "/>
<path style="fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 464.105469 260.757812 L 467.5625 260.757812 "/>
<path style="fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 464.105469 241.621094 L 467.5625 241.621094 "/>
<path style="fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 464.105469 222.484375 L 467.5625 222.484375 "/>
<path style="fill:none;stroke-width:0.375;stroke-linecap:butt;stroke-linejoin:round;stroke:rgb(100%,100%,100%);stroke-opacity:1;stroke-miterlimit:10;" d="M 464.105469 203.347656 L 467.5625 203.347656 "/>
</g>
</svg>

```

```{=html}
<div style="background-color:grey;height:2px;width:100%;"></div>
```

```{=html}
<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/main/projects/datavis/ggplot/notebooks/study_session_20240225.clj">projects/datavis/ggplot/notebooks/study_session_20240225.clj</a></small></small></pre></div>
```