### 2024-04-02 docs/projects/datavis/ggplot/draft.qmd

---
format:
  html: {toc: true, toc-depth: 4, theme: cosmo, output-file: draft.html}
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style><style>.clay-limit-image-width .clay-image {max-width: 100%}
</style>
<script src="draft_files/md-default2.js" type="text/javascript"></script><script src="draft_files/md-default3.js" type="text/javascript"></script>

::: {.sourceClojure}
```clojure
(ns draft
  (:require [clojisr.v1.r :as r :refer [r r$ r->clj]]
            [clojisr.v1.applications.plotting :as plotting]
            [representing]))
```
:::

# Draft explorations

::: {.sourceClojure}
```clojure
(delay
  (-> "(ggplot(mpg, aes(cty, hwy)))"
      r
      representing/ggplot->clj
      (dissoc :data)))
```
:::

::: {.printedClojure}
```clojure
{:layers [],
 :scales {:super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"]}}

```
:::

::: {.sourceClojure}
```clojure
(delay
  (-> "(ggplot(mpg, aes(cty, hwy)))"
      r
      representing/ggplot->clj
      (dissoc :data)))
```
:::

::: {.printedClojure}
```clojure
{:layers [],
 :scales {:super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"]}}

```
:::

::: {.sourceClojure}
```clojure
(delay
  (-> "(ggplot(mpg, mapping=aes(x=cty, y=hwy)))"
      r
      representing/ggplot->clj
      (dissoc :data)))
```
:::

::: {.printedClojure}
```clojure
{:layers [],
 :scales {:super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"]}}

```
:::

::: {.sourceClojure}
```clojure
(delay
  (-> "(ggplot(mpg, mapping=aes(x=cty, y=hwy))
      + geom_point(mapping=aes(color=cyl), size=5)
      + ylim(c(20,30)))"
      r
      representing/ggplot->clj
      (dissoc :data)))
```
:::

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params {:size [5.0]},
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping {:colour [~ cyl]},
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor {: nil, :mapping {: nil, :color cyl}, :size [5.0]},
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales
 {:scales
  [{:labels [],
    :scale_name ["position_c"],
    :breaks [],
    :call
    {:labels labels,
     :scale_name ["position_c"],
     :breaks breaks,
     :name name,
     :palette identity,
     :super ScaleContinuousPosition,
     :guide guide,
     :minor_breaks minor_breaks,
     :n.breaks n.breaks,
     : nil,
     :aesthetics [$ ggplot_global y_aes],
     :na.value na.value,
     :trans trans,
     :oob oob,
     :position position,
     :limits limits,
     :expand expand},
    :rescaler :ggproto-method,
    :name [],
    :palette :ggproto-method,
    :super :ggproto-method,
    :guide [],
    :minor_breaks [],
    :n.breaks nil,
    :aesthetics
    ["y"
     "ymin"
     "ymax"
     "yend"
     "yintercept"
     "ymin_final"
     "ymax_final"
     "lower"
     "middle"
     "upper"
     "y0"],
    :na.value [nil],
    :trans
    {:name ["identity"],
     :transform
     "function (x) \nx\n&lt;bytecode: 0x643ee0636140&gt;\n&lt;environment: namespace:base&gt;\n",
     :inverse
     "function (x) \nx\n&lt;bytecode: 0x643ee0636140&gt;\n&lt;environment: namespace:base&gt;\n",
     :breaks
     "function (x, n = n_default) \n{\n    x &lt;- x[is.finite(x)]\n    if (length(x) == 0) {\n        return(numeric())\n    }\n    rng &lt;- range(x)\n    labeling::extended(rng[1], rng[2], n, ...)\n}\n&lt;bytecode: 0x643ee45a5010&gt;\n&lt;environment: 0x643ef149d1b0&gt;\n",
     :minor_breaks
     "function (b, limits, n) \n{\n    b &lt;- b[!is.na(b)]\n    if (length(b) &lt; 2) {\n        return()\n    }\n    bd &lt;- diff(b)[1]\n    if (!reverse) {\n        if (min(limits) &lt; min(b)) \n            b &lt;- c(b[1] - bd, b)\n        if (max(limits) &gt; max(b)) \n            b &lt;- c(b, b[length(b)] + bd)\n    }\n    else {\n        if (max(limits) &gt; max(b)) \n            b &lt;- c(b[1] - bd, b)\n        if (min(limits) &lt; min(b)) \n            b &lt;- c(b, b[length(b)] + bd)\n    }\n    seq_between &lt;- function(a, b) {\n        seq(a, b, length.out = n + 1)[-(n + 1)]\n    }\n    breaks &lt;- unlist(Map(seq_between, b[-length(b)], b[-1]))\n    breaks &lt;- c(breaks, b[length(b)])\n    breaks\n}\n&lt;bytecode: 0x643ee45aa910&gt;\n&lt;environment: 0x643ef149d338&gt;\n",
     :format
     "function (x) \n{\n    if (!is.null(names(x))) {\n        return(names(x))\n    }\n    ret &lt;- format(x, ..., trim = TRUE, justify = \"left\")\n    ret[is.na(x)] &lt;- NA\n    ret\n}\n&lt;bytecode: 0x643ee45adf38&gt;\n&lt;environment: 0x643ef149d4c0&gt;\n",
     :domain [##-Inf ##Inf]},
    :oob :ggproto-method,
    :position ["left"],
    :limits [20.0 30.0],
    :expand [],
    :range {:super :ggproto-method}}],
  :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["cyl"]}}

```
:::

::: {.sourceClojure}
```clojure
(delay
  (-> "(ggplot(mpg, mapping=aes(x=cty, y=hwy))
      + geom_point(mapping=aes(color=factor(cyl)), size=20))"
      r
      plotting/plot->buffered-image))
```
:::

::: {.clay-image}

```{=html}
<img src="draft_files/0.png" />
```

:::

::: {.sourceClojure}
```clojure
(delay
  (-> "(ggplot(mpg, mapping=aes(x=cty, y=hwy))
      + geom_point(mapping=aes(color=cyl), size=5)
      + ylim(c(20,30))
      + scale_x_log10())"
      r
      representing/ggplot->clj
      (dissoc :data)))
```
:::

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params {:size [5.0]},
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping {:colour [~ cyl]},
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor {: nil, :mapping {: nil, :color cyl}, :size [5.0]},
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales
 {:scales
  [{:range {:super :ggproto-method}, :super :ggproto-method}
   {:labels [],
    :scale_name ["position_c"],
    :breaks [],
    :call
    {:labels labels,
     :scale_name ["position_c"],
     :breaks breaks,
     :name name,
     :palette identity,
     :super ScaleContinuousPosition,
     :guide guide,
     :minor_breaks minor_breaks,
     :n.breaks n.breaks,
     : nil,
     :aesthetics [$ ggplot_global x_aes],
     :na.value na.value,
     :trans trans,
     :oob oob,
     :position position,
     :limits limits,
     :expand expand},
    :rescaler :ggproto-method,
    :name [],
    :palette :ggproto-method,
    :super :ggproto-method,
    :guide [],
    :minor_breaks [],
    :n.breaks nil,
    :aesthetics
    ["x"
     "xmin"
     "xmax"
     "xend"
     "xintercept"
     "xmin_final"
     "xmax_final"
     "xlower"
     "xmiddle"
     "xupper"
     "x0"],
    :na.value [nil],
    :trans
    {:name ["log-10"],
     :transform
     "function (x) \nlog(x, base)\n&lt;bytecode: 0x643ee9d7fc60&gt;\n&lt;environment: 0x643ee9f9be70&gt;\n",
     :inverse
     "function (x) \nbase^x\n&lt;bytecode: 0x643ee9d7fdb0&gt;\n&lt;environment: 0x643ee9f9be70&gt;\n",
     :breaks
     "function (x, n = n_default) \n{\n    raw_rng &lt;- suppressWarnings(range(x, na.rm = TRUE))\n    if (any(!is.finite(raw_rng))) {\n        return(numeric())\n    }\n    rng &lt;- log(raw_rng, base = base)\n    min &lt;- floor(rng[1])\n    max &lt;- ceiling(rng[2])\n    if (max == min) {\n        return(base^min)\n    }\n    by &lt;- floor((max - min)/n) + 1\n    breaks &lt;- base^seq(min, max, by = by)\n    relevant_breaks &lt;- base^rng[1] &lt;= breaks & breaks &lt;= base^rng[2]\n    if (sum(relevant_breaks) &gt;= (n - 2)) {\n        return(breaks)\n    }\n    while (by &gt; 1) {\n        by &lt;- by - 1\n        breaks &lt;- base^seq(min, max, by = by)\n        relevant_breaks &lt;- base^rng[1] &lt;= breaks & breaks &lt;= \n            base^rng[2]\n        if (sum(relevant_breaks) &gt;= (n - 2)) {\n            return(breaks)\n        }\n    }\n    log_sub_breaks(rng, n = n, base = base)\n}\n&lt;bytecode: 0x643ee9cc7d78&gt;\n&lt;environment: 0x643ee9f90eb8&gt;\n",
     :minor_breaks
     "function (b, limits, n) \n{\n    b &lt;- b[!is.na(b)]\n    if (length(b) &lt; 2) {\n        return()\n    }\n    bd &lt;- diff(b)[1]\n    if (!reverse) {\n        if (min(limits) &lt; min(b)) \n            b &lt;- c(b[1] - bd, b)\n        if (max(limits) &gt; max(b)) \n            b &lt;- c(b, b[length(b)] + bd)\n    }\n    else {\n        if (max(limits) &gt; max(b)) \n            b &lt;- c(b[1] - bd, b)\n        if (min(limits) &lt; min(b)) \n            b &lt;- c(b, b[length(b)] + bd)\n    }\n    seq_between &lt;- function(a, b) {\n        seq(a, b, length.out = n + 1)[-(n + 1)]\n    }\n    breaks &lt;- unlist(Map(seq_between, b[-length(b)], b[-1]))\n    breaks &lt;- c(breaks, b[length(b)])\n    breaks\n}\n&lt;bytecode: 0x643ee45aa910&gt;\n&lt;environment: 0x643ee9f91350&gt;\n",
     :format
     "function (x) \n{\n    if (!is.null(names(x))) {\n        return(names(x))\n    }\n    ret &lt;- format(x, ..., trim = TRUE, justify = \"left\")\n    ret[is.na(x)] &lt;- NA\n    ret\n}\n&lt;bytecode: 0x643ee45adf38&gt;\n&lt;environment: 0x643ee9f914d8&gt;\n",
     :domain [1.0E-100 ##Inf]},
    :oob :ggproto-method,
    :position ["bottom"],
    :limits nil,
    :expand [],
    :range {:super :ggproto-method}}],
  :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["cyl"]}}

```
:::

::: {.sourceClojure}
```clojure
(delay
  (-> "(ggplot(mpg, mapping=aes(x=cty, y=hwy))
      + geom_point(mapping=aes(color=cyl), size=5)
      + ylim(c(20,30))
      + scale_x_log10()
      + theme_linedraw())"
      r
      representing/ggplot->clj
      (dissoc :data)))
```
:::

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params {:size [5.0]},
   :stat {:compute_layer :ggproto-method, :super :ggproto-method},
   :show.legend [nil],
   :mapping {:colour [~ cyl]},
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false]},
   :geom
   {:non_missing_aes ["size" "shape" "colour"],
    :draw_key :ggproto-method,
    :default_aes
    {:shape [19.0],
     :colour ["black"],
     :size [1.5],
     :fill [nil],
     :alpha [nil],
     :stroke [0.5]},
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method},
   :stat_params {:na.rm [false]},
   :constructor {: nil, :mapping {: nil, :color cyl}, :size [5.0]},
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales
 {:scales
  [{:range {:super :ggproto-method}, :super :ggproto-method}
   {:range {:super :ggproto-method}, :super :ggproto-method}],
  :super :ggproto-method},
 :mapping {:x [~ cty], :y [~ hwy]},
 :theme
 {:axis.ticks
  {:colour ["black"],
   :linewidth [0.5],
   :linetype nil,
   :lineend nil,
   :arrow [false],
   :inherit.blank [true]},
  :legend.justification ["center"],
  :panel.grid.minor
  {:colour nil,
   :linewidth [0.05],
   :linetype nil,
   :lineend nil,
   :arrow [false],
   :inherit.blank [true]},
  :strip.text.x.bottom nil,
  :legend.key.width nil,
  :panel.spacing.x nil,
  :axis.text.y
  {:hjust [1.0],
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size nil,
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust nil,
   :margin [0.0 2.2 0.0 0.0]},
  :panel.ontop [false],
  :axis.ticks.x.top nil,
  :panel.border
  {:fill [nil],
   :colour ["black"],
   :linewidth [1.0],
   :linetype nil,
   :inherit.blank [true]},
  :legend.key.height nil,
  :plot.caption
  {:hjust [1.0],
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size [0.8],
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust [1.0],
   :margin [5.5 0.0 0.0 0.0]},
  :panel.grid.major.y nil,
  :panel.grid
  {:colour ["black"],
   :linewidth nil,
   :linetype nil,
   :lineend nil,
   :arrow [false],
   :inherit.blank [true]},
  :axis.ticks.x.bottom nil,
  :panel.grid.major.x nil,
  :plot.tag
  {:hjust [0.5],
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size [1.2],
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust [0.5],
   :margin nil},
  :axis.text
  {:hjust nil,
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size [0.8],
   :face nil,
   :lineheight nil,
   :colour ["black"],
   :vjust nil,
   :margin nil},
  :axis.line.x nil,
  :panel.spacing.y nil,
  :legend.spacing.x nil,
  :axis.text.x
  {:hjust nil,
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size nil,
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust [1.0],
   :margin [2.2 0.0 0.0 0.0]},
  :axis.line.y.left nil,
  :plot.caption.position ["panel"],
  :plot.subtitle
  {:hjust [0.0],
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size nil,
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust [1.0],
   :margin [0.0 0.0 5.5 0.0]},
  :axis.ticks.y.right nil,
  :legend.box nil,
  :strip.text.y
  {:hjust nil,
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle [-90.0],
   :size nil,
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust nil,
   :margin nil},
  :panel.spacing [5.5],
  :legend.background
  {:fill nil,
   :colour [nil],
   :linewidth nil,
   :linetype nil,
   :inherit.blank [true]},
  :strip.switch.pad.grid [2.75],
  :axis.ticks.length.y.left nil,
  :legend.direction nil,
  :strip.text
  {:hjust nil,
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size [0.8],
   :face nil,
   :lineheight nil,
   :colour ["white"],
   :vjust nil,
   :margin [4.4 4.4 4.4 4.4]},
  :axis.line.x.bottom nil,
  :axis.ticks.length.x.bottom nil,
  :axis.text.y.left nil,
  :axis.ticks.length.x nil,
  :plot.margin [5.5 5.5 5.5 5.5],
  :axis.ticks.length [2.75],
  :plot.title
  {:hjust [0.0],
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size [1.2],
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust [1.0],
   :margin [0.0 0.0 5.5 0.0]},
  :axis.ticks.y nil,
  :legend.text.align nil,
  :legend.text
  {:hjust nil,
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size [0.8],
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust nil,
   :margin nil},
  :legend.key
  {:fill ["white"],
   :colour [nil],
   :linewidth nil,
   :linetype nil,
   :inherit.blank [true]},
  :axis.line.y nil,
  :axis.title.y.left nil,
  :axis.ticks.length.x.top nil,
  :legend.box.margin [0.0 0.0 0.0 0.0],
  :axis.title.y.right
  {:hjust nil,
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle [-90.0],
   :size nil,
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust [0.0],
   :margin [0.0 0.0 0.0 2.75]},
  :title nil,
  :strip.clip ["inherit"],
  :strip.background.x nil,
  :strip.text.x nil,
  :strip.placement ["inside"],
  :axis.ticks.x nil,
  :axis.line.y.right nil,
  :axis.title.x.bottom nil,
  :legend.key.size [1.2],
  :axis.text.x.top
  {:hjust nil,
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size nil,
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust [0.0],
   :margin [0.0 0.0 2.2 0.0]},
  :panel.background
  {:fill ["white"],
   :colour [nil],
   :linewidth nil,
   :linetype nil,
   :inherit.blank [true]},
  :plot.tag.position ["topleft"],
  :legend.margin [5.5 5.5 5.5 5.5],
  :plot.title.position ["panel"],
  :line
  {:colour ["black"],
   :linewidth [0.5],
   :linetype [1.0],
   :lineend ["butt"],
   :arrow [false],
   :inherit.blank [true]},
  :strip.text.y.left
  {:hjust nil,
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle [90.0],
   :size nil,
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust nil,
   :margin nil},
  :strip.background
  {:fill ["black"],
   :colour nil,
   :linewidth nil,
   :linetype nil,
   :inherit.blank [true]},
  :axis.title.y
  {:hjust nil,
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle [90.0],
   :size nil,
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust [1.0],
   :margin [0.0 2.75 0.0 0.0]},
  :legend.box.background [],
  :panel.grid.major
  {:colour nil,
   :linewidth [0.1],
   :linetype nil,
   :lineend nil,
   :arrow [false],
   :inherit.blank [true]},
  :axis.title nil,
  :strip.text.y.right nil,
  :legend.title.align nil,
  :plot.background
  {:fill nil,
   :colour ["white"],
   :linewidth nil,
   :linetype nil,
   :inherit.blank [true]},
  :aspect.ratio nil,
  :legend.spacing [11.0],
  :axis.title.x
  {:hjust nil,
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size nil,
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust [1.0],
   :margin [2.75 0.0 0.0 0.0]},
  :axis.line.x.top nil,
  :legend.position ["right"],
  :axis.ticks.length.y nil,
  :axis.ticks.y.left nil,
  :axis.ticks.length.y.right nil,
  :legend.box.just nil,
  :legend.title
  {:hjust [0.0],
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size nil,
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust nil,
   :margin nil},
  :axis.text.y.right
  {:hjust [0.0],
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size nil,
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust nil,
   :margin [0.0 0.0 0.0 2.2]},
  :axis.text.x.bottom nil,
  :strip.background.y nil,
  :axis.line [],
  :rect
  {:fill ["white"],
   :colour ["black"],
   :linewidth [0.5],
   :linetype [1.0],
   :inherit.blank [true]},
  :legend.box.spacing [11.0],
  :panel.grid.minor.y nil,
  :panel.grid.minor.x nil,
  :strip.switch.pad.wrap [2.75],
  :axis.title.x.top
  {:hjust nil,
   :family nil,
   :inherit.blank [true],
   :debug nil,
   :angle nil,
   :size nil,
   :face nil,
   :lineheight nil,
   :colour nil,
   :vjust [0.0],
   :margin [0.0 0.0 2.75 0.0]},
  :strip.text.x.top nil,
  :legend.spacing.y nil,
  :text
  {:hjust [0.5],
   :family [""],
   :inherit.blank [true],
   :debug [false],
   :angle [0.0],
   :size [11.0],
   :face ["plain"],
   :lineheight [0.9],
   :colour ["black"],
   :vjust [0.5],
   :margin [0.0 0.0 0.0 0.0]}},
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["cty"], :y ["hwy"], :colour ["cyl"]}}

```
:::

::: {.sourceClojure}
```clojure
(delay
  (-> "(ggplot(data.frame(x = rnorm(1000, 2, 2)), aes(x)) +
      geom_histogram(aes(y=..density..)) +  # scale histogram y
      geom_density(col = 'red', size=5))"
      r
      plotting/plot->buffered-image))
```
:::

::: {.clay-image}

```{=html}
<img src="draft_files/1.png" />
```

:::

::: {.sourceClojure}
```clojure
(delay
  (-> "(ggplot(data.frame(x = rnorm(1000, 2, 2)), aes(x)) +
      geom_histogram(aes(y=..density..)) +  # scale histogram y
      geom_density(col = 'red'))"
      r
      representing/ggplot->clj
      (dissoc :data)))
```
:::

::: {.printedClojure}
```clojure
{:layers
 [{:aes_params [],
   :stat
   {:default_aes
    {:x [~ [after_stat count]],
     :y [~ [after_stat count]],
     :weight [1.0]},
    :extra_params ["na.rm" "orientation"],
    :super :ggproto-method,
    :compute_group :ggproto-method,
    :required_aes ["x|y"],
    :setup_params :ggproto-method,
    :dropped_aes ["weight"]},
   :show.legend [nil],
   :mapping {:y [~ ..density..]},
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params {:na.rm [false], :orientation [nil]},
   :geom
   {:setup_data :ggproto-method,
    :rename_size [true],
    :non_missing_aes ["xmin" "xmax" "ymin" "ymax"],
    :extra_params ["just" "na.rm" "orientation"],
    :super :ggproto-method,
    :required_aes ["x" "y"],
    :draw_panel :ggproto-method,
    :setup_params :ggproto-method},
   :stat_params
   {:binwidth nil,
    :bins nil,
    :na.rm [false],
    :orientation [nil],
    :pad [false]},
   :constructor [geom_histogram {0 aes, :y ..density..}],
   :position
   {:setup_data :ggproto-method,
    :type nil,
    :compute_panel :ggproto-method,
    :vjust [1.0],
    :reverse [false],
    :super :ggproto-method,
    :fill [false],
    :setup_params :ggproto-method},
   :data []}
  {:aes_params {:colour ["red"]},
   :stat
   {:default_aes
    {:x [~ [after_stat density]],
     :y [~ [after_stat density]],
     :fill [nil],
     :weight nil},
    :extra_params ["na.rm" "orientation"],
    :super :ggproto-method,
    :compute_group :ggproto-method,
    :required_aes ["x|y"],
    :setup_params :ggproto-method,
    :dropped_aes ["weight"]},
   :show.legend [nil],
   :mapping nil,
   :super :ggproto-method,
   :inherit.aes [true],
   :geom_params
   {:na.rm [false], :orientation [nil], :outline.type ["upper"]},
   :geom
   {:default_aes
    {:fill [nil],
     :weight [1.0],
     :colour ["black"],
     :alpha [nil],
     :linewidth [0.5],
     :linetype [1.0]},
    :super :ggproto-method},
   :stat_params {:na.rm [false], :orientation [nil]},
   :constructor {: nil, :col ["red"]},
   :position {:compute_layer :ggproto-method, :super :ggproto-method},
   :data []}],
 :scales {:scales [], :super :ggproto-method},
 :mapping {:x [~ x]},
 :theme [],
 :coordinates
 {:expand [true],
  :clip ["on"],
  :limits {:x nil, :y nil},
  :super :ggproto-method,
  :default [true]},
 :facet {:shrink [true], :super :ggproto-method},
 :labels {:x ["x"], :y ["density"], :fill ["fill"], :weight ["weight"]}}

```
:::

# Exploring ggtrace

[ggtrace](https://yjunechoe.github.io/ggtrace/) - programmatically explore, debug, and manipulate ggplot internals

::: {.sourceClojure}
```clojure
(r/require-r '[ggtrace :as ggtrace])
```
:::

::: {.printedClojure}
```clojure
nil

```
:::

::: {.sourceClojure}
```clojure
(let [bar-plot (r "
ggplot(mpg, aes(class, fill = class)) +
geom_bar() +
theme(legend.position = 'none')")]
  (ggtrace/layer_after_stat bar-plot))
```
:::

::: {.printedClojure}
```clojure
# A tibble: 7 × 8
  count  prop x          width flipped_aes fill       PANEL group
  &lt;dbl&gt; &lt;dbl&gt; &lt;mppd_dsc&gt; &lt;dbl&gt; &lt;lgl&gt;       &lt;chr&gt;      &lt;fct&gt; &lt;int&gt;
1     5     1 1            0.9 FALSE       2seater    1         1
2    47     1 2            0.9 FALSE       compact    1         2
3    41     1 3            0.9 FALSE       midsize    1         3
4    11     1 4            0.9 FALSE       minivan    1         4
5    33     1 5            0.9 FALSE       pickup     1         5
6    35     1 6            0.9 FALSE       subcompact 1         6
7    62     1 7            0.9 FALSE       suv        1         7

```
:::

```{=html}
<div style="background-color:grey;height:2px;width:100%;"></div>
```

```{=html}
<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/main/projects/datavis/ggplot/notebooks/draft.clj">projects/datavis/ggplot/notebooks/draft.clj</a></small></small></pre></div>
```