### 2024-04-15 docs/projects/datavis/ggplot/ggplotly_cont.qmd

---
format:
  html: {toc: true, toc-depth: 4, theme: cosmo, output-file: ggplotly_cont.html}
fontsize: 0.9em
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style><style>.clay-limit-image-width .clay-image {max-width: 100%}
</style>
<script src="ggplotly_cont_files/md-default0.js" type="text/javascript"></script><script src="ggplotly_cont_files/md-default1.js" type="text/javascript"></script><script src="htmlwidgets-ggplotly/gh-repos/scicloj/ggplotly-deps/htmlwidgets-1.6.2/htmlwidgets.js" type="text/javascript"></script><script src="htmlwidgets-ggplotly/gh-repos/scicloj/ggplotly-deps/plotly-binding-4.10.4.9000/plotly.js" type="text/javascript"></script><script src="htmlwidgets-ggplotly/gh-repos/scicloj/ggplotly-deps/typedarray-0.1/typedarray.min.js" type="text/javascript"></script><script src="htmlwidgets-ggplotly/gh-repos/scicloj/ggplotly-deps/jquery-3.5.1/jquery.min.js" type="text/javascript"></script><script src="htmlwidgets-ggplotly/gh-repos/scicloj/ggplotly-deps/crosstalk-1.2.1/js/crosstalk.min.js" type="text/javascript"></script><script src="htmlwidgets-ggplotly/gh-repos/scicloj/ggplotly-deps/plotly-main-2.11.1/plotly-latest.min.js" type="text/javascript"></script><link href="htmlwidgets-ggplotly/gh-repos/scicloj/ggplotly-deps/crosstalk-1.2.1/css/crosstalk.min.css" rel="stylesheet" type="text/css" /><link href="htmlwidgets-ggplotly/gh-repos/scicloj/ggplotly-deps/plotly-htmlwidgets-css-2.11.1/plotly-htmlwidgets.css" rel="stylesheet" type="text/css" />

::: {.sourceClojure}
```clojure
(ns ggplotly-cont
  (:require [tablecloth.api :as tc]
            [scicloj.metamorph.ml.toydata.ggplot :as toydata.ggplot]
            [tech.v3.datatype.functional :as fun]
            [scicloj.kindly.v4.kind :as kind]
            [clojure.string :as str]
            [clojure.math :as math]
            [clojure2d.color :as color]))
```
:::

# Further exploring ggplotly's client side - DRAFT

The following template was generated from the JSON file
in the HTML generated by
`ggplotly(ggplot(mpg, aes(x=hwy, y=displ, color=factor(cyl))) + geom_point() + geom_smooth(method="lm"))`
and then gradually generalizing it as a set of Clojure functions.
This is a work-in-progress draft. Some parts are still hard-coded.

::: {.sourceClojure}
```clojure
(defn layer [{:keys [type x y color text name legendgroup showlegend]}]
  (merge {:hoveron "points",
          :legendgroup legendgroup,
          :showlegend showlegend
          :frame nil,
          :hoverinfo "text",
          :name name,
          :mode (case type
                  :point "markers"
                  :line "lines")
          :type "scatter",

          :xaxis "x",
          :yaxis "y",
          :x (vec x)
          :y (vec y)
          :text (vec text)}
         (case type
           :point {:marker {:autocolorscale false,
                            :color color,
                            :opacity 1,
                            :size 5.66929133858268,
                            :symbol "circle",
                            :line {:width 1.88976377952756, :color color}}}
           :line {:line {:width 3.77952755905512,
                         :color color,
                         :dash "solid"}})))
```
:::

::: {.sourceClojure}
```clojure
(color/palette )
```
:::

::: {.printedClojure}
```clojure
[[223.0 46.0 40.0 255.0]
 [254.0 128.0 26.0 255.0]
 [233.0 191.0 53.0 255.0]
 [129.0 187.0 66.0 255.0]
 [50.0 199.0 169.0 255.0]
 [74.0 155.0 220.0 255.0]]

```
:::

::: {.sourceClojure}
```clojure
(def colors
  ["rgba(248,118,109,1)"
   "rgba(124,174,0,1)"
   "rgba(0,191,196,1)"
   "rgba(199,124,255,1)"])
```
:::

::: {.sourceClojure}
```clojure
(def config
  {:doubleClick "reset",
   :modeBarButtonsToAdd ["hoverclosest" "hovercompare"],
   :showSendToCloud false},)
```
:::

::: {.sourceClojure}
```clojure
(def highlight
  {:on "plotly_click",
   :persistent false,
   :dynamic false,
   :selectize false,
   :opacityDim 0.2,
   :selected {:opacity 1},
   :debounce 0},)
```
:::

::: {.sourceClojure}
```clojure
(defn texts [ds column-names]
  (-> ds
      (tc/select-columns column-names)
      tc/rows
      (->> (map (fn [row]
                  (str/join "<br />"
                            (map (partial format "%s: %s")
                                 column-names row)))))))
```
:::

::: {.sourceClojure}
```clojure
(defn ->tickvals [l r]
  (let [jump (-> (- r l)
                 (/ 6)
                 math/floor
                 int
                 (max 1))]
    (-> l
        math/ceil
        (range r jump))))
```
:::

::: {.sourceClojure}
```clojure
(defn axis [{:keys [minval maxval anchor title]}]
  (let [tickvals (->tickvals minval maxval)
        ticktext (mapv str tickvals)
        range-len (- maxval minval)
        range-expansion (* 0.1 range-len)
        expanded-range [(- minval range-expansion)
                        (+ maxval range-expansion)]]
    {:linewidth 0,
     :ticklen 3.65296803652968,
     :tickcolor "rgba(51,51,51,1)",
     :tickmode "array",
     :gridcolor "rgba(255,255,255,1)",
     :automargin true,
     :type "linear",
     :tickvals tickvals,
     :zeroline false,
     :title
     {:text title,
      :font {:color "rgba(0,0,0,1)", :family "", :size 14.6118721461187}},
     :tickfont {:color "rgba(77,77,77,1)", :family "", :size 11.689497716895},
     :autorange false,
     :showticklabels true,
     :showline false,
     :showgrid true,
     :ticktext ticktext,
     :ticks "outside",
     :gridwidth 0.66417600664176,
     :anchor "y",
     :domain [0 1],
     :hoverformat ".2f",
     :tickangle 0,
     :tickwidth 0.66417600664176,
     :categoryarray ticktext,
     :categoryorder "array",
     :range expanded-range},))
```
:::

::: {.sourceClojure}
```clojure
(defn layout [{:keys [xaxis yaxis]}]
  {:plot_bgcolor "rgba(235,235,235,1)",
   :paper_bgcolor "rgba(255,255,255,1)",
   :legend {:bgcolor "rgba(255,255,255,1)",
            :bordercolor "transparent",
            :borderwidth 1.88976377952756,
            :font {:color "rgba(0,0,0,1)", :family "", :size 11.689497716895},
            :title {:text "factor(cyl)",
                    :font {:color "rgba(0,0,0,1)", :family "", :size 14.6118721461187}}},
   :font {:color "rgba(0,0,0,1)", :family "", :size 14.6118721461187},
   :showlegend true,
   :barmode "relative",
   :hovermode "closest",
   :margin
   {:t 25.7412480974125,
    :r 7.30593607305936,
    :b 39.6955859969559,
    :l 31.4155251141553},
   :shapes [{:yref "paper",
             :fillcolor nil,
             :xref "paper",
             :y1 1,
             :type "rect",
             :line {:color nil, :width 0, :linetype []},
             :y0 0,
             :x1 1,
             :x0 0}]
   :xaxis xaxis
   :yaxis yaxis})
```
:::

::: {.sourceClojure}
```clojure
(defn factor [column]
  (vary-meta column
             assoc :gg/factor? true))
```
:::

::: {.sourceClojure}
```clojure
(defn factor? [column]
  (:gg/factor? column))
```
:::

::: {.sourceClojure}
```clojure
(let [data toydata.ggplot/mpg
      point-layers (-> data
                       (tc/add-column "factor(cyl)" #(factor (:cyl %)))
                       (tc/group-by :cyl {:result-type :as-map})
                       (->> (sort-by key)
                            (map-indexed
                             (fn [i [group-name group-data]]
                               (let [base {:x (:hwy group-data),
                                           :y (:displ group-data)
                                           :color (colors i)
                                           :name group-name
                                           :legendgroup group-name}
                                     predictions (map
                                                  (fun/linear-regressor (:hwy group-data)
                                                                        (:displ group-data))
                                                  (:hwy group-data))]
                                 [(-> base
                                      (assoc :type :point
                                             :showlegend true
                                             :y (:displ group-data)
                                             :text (-> group-data
                                                       (texts [:hwy :displ "factor(cyl)"])))
                                      layer)
                                  (-> base
                                      (assoc :type :line
                                             :showlegend false
                                             :y predictions
                                             :text (-> group-data
                                                       ;; (tc/add-column :displ predictions)
                                                       (texts [:hwy :displ "factor(cyl)"])))
                                      layer)])))
                            (apply concat)))
      xmin (-> data :hwy fun/reduce-min)
      xmax (-> data :hwy fun/reduce-max)
      ymin (-> data :displ fun/reduce-min)
      ymax (-> data :displ fun/reduce-max)
      xaxis (axis {:minval xmin
                   :maxval xmax
                   :anchor "x"
                   :title :hwy})
      yaxis (axis {:minval ymin
                   :maxval ymax
                   :anchor "x"
                   :title :displ})]
  (kind/htmlwidgets-ggplotly
   {:x
    {:config config
     :highlight highlight
     :base_url "https://plot.ly",
     :layout (layout {:xaxis xaxis
                      :yaxis yaxis})
     :data point-layers},
    :evals [],
    :jsHooks []}))
```
:::

```{=html}
<div class="plotly html-widget html-fill-item-overflow-hidden html-fill-item" id="htmlwidget-14" style="width:100%;height:400px;"><script data-for="htmlwidget-14" type="application/htmlwidget-sizing">{"viewer":{"width":"100%","height":400,"padding":"0","fille":true},"browser":{"width":"100%","height":400,"padding":"0","fille":true}}</script><script data-for="htmlwidget-14" type="application/json">{"x":{"config":{"doubleClick":"reset","modeBarButtonsToAdd":["hoverclosest","hovercompare"],"showSendToCloud":false},"highlight":{"on":"plotly_click","persistent":false,"dynamic":false,"selectize":false,"opacityDim":0.2,"selected":{"opacity":1},"debounce":0},"base_url":"https:\/\/plot.ly","layout":{"plot_bgcolor":"rgba(235,235,235,1)","paper_bgcolor":"rgba(255,255,255,1)","legend":{"bgcolor":"rgba(255,255,255,1)","bordercolor":"transparent","borderwidth":1.88976377952756,"font":{"color":"rgba(0,0,0,1)","family":"","size":11.689497716895},"title":{"text":"factor(cyl)","font":{"color":"rgba(0,0,0,1)","family":"","size":14.6118721461187}}},"xaxis":{"linewidth":0,"ticklen":3.65296803652968,"tickcolor":"rgba(51,51,51,1)","tickmode":"array","gridcolor":"rgba(255,255,255,1)","automargin":true,"type":"linear","tickvals":[12.0,17.0,22.0,27.0,32.0,37.0,42.0],"zeroline":false,"title":{"text":"hwy","font":{"color":"rgba(0,0,0,1)","family":"","size":14.6118721461187}},"tickfont":{"color":"rgba(77,77,77,1)","family":"","size":11.689497716895},"autorange":false,"showticklabels":true,"showline":false,"showgrid":true,"ticktext":["12.0","17.0","22.0","27.0","32.0","37.0","42.0"],"ticks":"outside","gridwidth":0.66417600664176,"anchor":"y","domain":[0,1],"hoverformat":".2f","tickangle":0,"tickwidth":0.66417600664176,"categoryarray":["12.0","17.0","22.0","27.0","32.0","37.0","42.0"],"categoryorder":"array","range":[8.8,47.2]},"font":{"color":"rgba(0,0,0,1)","family":"","size":14.6118721461187},"showlegend":true,"barmode":"relative","yaxis":{"linewidth":0,"ticklen":3.65296803652968,"tickcolor":"rgba(51,51,51,1)","tickmode":"array","gridcolor":"rgba(255,255,255,1)","automargin":true,"type":"linear","tickvals":[2.0,3.0,4.0,5.0,6.0],"zeroline":false,"title":{"text":"displ","font":{"color":"rgba(0,0,0,1)","family":"","size":14.6118721461187}},"tickfont":{"color":"rgba(77,77,77,1)","family":"","size":11.689497716895},"autorange":false,"showticklabels":true,"showline":false,"showgrid":true,"ticktext":["2.0","3.0","4.0","5.0","6.0"],"ticks":"outside","gridwidth":0.66417600664176,"anchor":"y","domain":[0,1],"hoverformat":".2f","tickangle":0,"tickwidth":0.66417600664176,"categoryarray":["2.0","3.0","4.0","5.0","6.0"],"categoryorder":"array","range":[1.06,7.54]},"hovermode":"closest","margin":{"t":25.7412480974125,"r":7.30593607305936,"b":39.6955859969559,"l":31.4155251141553},"shapes":[{"yref":"paper","fillcolor":null,"xref":"paper","y1":1,"type":"rect","line":{"color":null,"width":0,"linetype":[]},"y0":0,"x1":1,"x0":0}]},"data":[{"y":[1.8,1.8,2.0,2.0,1.8,1.8,2.0,2.0,2.4,2.4,2.4,1.6,1.6,1.6,1.6,1.6,1.8,1.8,1.8,2.0,2.4,2.4,2.4,2.4,2.0,2.0,2.0,2.0,2.4,2.4,2.5,2.5,2.5,2.5,2.5,2.5,2.5,2.5,2.2,2.2,2.5,2.5,2.5,2.5,2.5,2.5,2.7,2.7,2.2,2.2,2.4,2.4,2.2,2.2,2.4,2.4,1.8,1.8,1.8,1.8,1.8,2.7,2.7,2.7,2.0,2.0,2.0,2.0,1.9,2.0,2.0,2.0,2.0,1.9,1.9,2.0,2.0,1.8,1.8,2.0,2.0],"hoveron":"points","legendgroup":4,"frame":null,"hoverinfo":"text","name":4,"marker":{"autocolorscale":false,"color":"rgba(248,118,109,1)","opacity":1,"size":5.66929133858268,"symbol":"circle","line":{"width":1.88976377952756,"color":"rgba(248,118,109,1)"}},"mode":"markers","type":"scatter","xaxis":"x","showlegend":true,"yaxis":"y","x":[29,29,31,30,26,25,28,27,27,30,24,33,32,32,29,32,34,36,36,29,26,27,30,31,26,29,28,27,29,27,31,32,25,24,27,25,26,23,26,26,26,26,25,27,25,27,20,20,29,27,31,31,27,29,31,31,30,33,35,37,35,20,20,22,29,26,29,29,44,29,26,29,29,44,41,29,26,29,29,28,29],"text":[":hwy: 29<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 30<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 25<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 28<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 30<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 24<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 33<br \/>:displ: 1.6<br \/>factor(cyl): 4",":hwy: 32<br \/>:displ: 1.6<br \/>factor(cyl): 4",":hwy: 32<br \/>:displ: 1.6<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 1.6<br \/>factor(cyl): 4",":hwy: 32<br \/>:displ: 1.6<br \/>factor(cyl): 4",":hwy: 34<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 36<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 36<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 30<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 28<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 32<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 25<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 24<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 25<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 23<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.2<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.2<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 25<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 25<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 20<br \/>:displ: 2.7<br \/>factor(cyl): 4",":hwy: 20<br \/>:displ: 2.7<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.2<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.2<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.2<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.2<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 30<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 33<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 35<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 37<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 35<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 20<br \/>:displ: 2.7<br \/>factor(cyl): 4",":hwy: 20<br \/>:displ: 2.7<br \/>factor(cyl): 4",":hwy: 22<br \/>:displ: 2.7<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 44<br \/>:displ: 1.9<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 44<br \/>:displ: 1.9<br \/>factor(cyl): 4",":hwy: 41<br \/>:displ: 1.9<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 28<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4"]},{"y":[2.1377007978924745,2.1377007978924745,2.0569213765537713,2.0973110872231233,2.2588699299005293,2.299259640569881,2.178090508561826,2.2184802192311777,2.2184802192311777,2.0973110872231233,2.3396493512392325,1.9761419552150683,2.01653166588442,2.01653166588442,2.1377007978924745,2.01653166588442,1.9357522445457167,1.8549728232070137,1.8549728232070137,2.1377007978924745,2.2588699299005293,2.2184802192311777,2.0973110872231233,2.0569213765537713,2.2588699299005293,2.1377007978924745,2.178090508561826,2.2184802192311777,2.1377007978924745,2.2184802192311777,2.0569213765537713,2.01653166588442,2.299259640569881,2.3396493512392325,2.2184802192311777,2.299259640569881,2.2588699299005293,2.380039061908584,2.2588699299005293,2.2588699299005293,2.2588699299005293,2.2588699299005293,2.299259640569881,2.2184802192311777,2.299259640569881,2.2184802192311777,2.5012081939166384,2.5012081939166384,2.1377007978924745,2.2184802192311777,2.0569213765537713,2.0569213765537713,2.2184802192311777,2.1377007978924745,2.0569213765537713,2.0569213765537713,2.0973110872231233,1.9761419552150683,1.895362533876365,1.8145831125376621,1.895362533876365,2.5012081939166384,2.5012081939166384,2.4204287725779356,2.1377007978924745,2.2588699299005293,2.1377007978924745,2.1377007978924745,1.5318551378522012,2.1377007978924745,2.2588699299005293,2.1377007978924745,2.1377007978924745,1.5318551378522012,1.6530242698602557,2.1377007978924745,2.2588699299005293,2.1377007978924745,2.1377007978924745,2.178090508561826,2.1377007978924745],"hoveron":"points","legendgroup":4,"frame":null,"hoverinfo":"text","name":4,"mode":"lines","type":"scatter","xaxis":"x","showlegend":false,"line":{"width":3.77952755905512,"color":"rgba(248,118,109,1)","dash":"solid"},"yaxis":"y","x":[29,29,31,30,26,25,28,27,27,30,24,33,32,32,29,32,34,36,36,29,26,27,30,31,26,29,28,27,29,27,31,32,25,24,27,25,26,23,26,26,26,26,25,27,25,27,20,20,29,27,31,31,27,29,31,31,30,33,35,37,35,20,20,22,29,26,29,29,44,29,26,29,29,44,41,29,26,29,29,28,29],"text":[":hwy: 29<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 30<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 25<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 28<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 30<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 24<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 33<br \/>:displ: 1.6<br \/>factor(cyl): 4",":hwy: 32<br \/>:displ: 1.6<br \/>factor(cyl): 4",":hwy: 32<br \/>:displ: 1.6<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 1.6<br \/>factor(cyl): 4",":hwy: 32<br \/>:displ: 1.6<br \/>factor(cyl): 4",":hwy: 34<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 36<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 36<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 30<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 28<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 32<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 25<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 24<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 25<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 23<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.2<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.2<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 25<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 25<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.5<br \/>factor(cyl): 4",":hwy: 20<br \/>:displ: 2.7<br \/>factor(cyl): 4",":hwy: 20<br \/>:displ: 2.7<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.2<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.2<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 27<br \/>:displ: 2.2<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.2<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 31<br \/>:displ: 2.4<br \/>factor(cyl): 4",":hwy: 30<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 33<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 35<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 37<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 35<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 20<br \/>:displ: 2.7<br \/>factor(cyl): 4",":hwy: 20<br \/>:displ: 2.7<br \/>factor(cyl): 4",":hwy: 22<br \/>:displ: 2.7<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 44<br \/>:displ: 1.9<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 44<br \/>:displ: 1.9<br \/>factor(cyl): 4",":hwy: 41<br \/>:displ: 1.9<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 26<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 1.8<br \/>factor(cyl): 4",":hwy: 28<br \/>:displ: 2.0<br \/>factor(cyl): 4",":hwy: 29<br \/>:displ: 2.0<br \/>factor(cyl): 4"]},{"y":[2.5,2.5,2.5,2.5],"hoveron":"points","legendgroup":5,"frame":null,"hoverinfo":"text","name":5,"marker":{"autocolorscale":false,"color":"rgba(124,174,0,1)","opacity":1,"size":5.66929133858268,"symbol":"circle","line":{"width":1.88976377952756,"color":"rgba(124,174,0,1)"}},"mode":"markers","type":"scatter","xaxis":"x","showlegend":true,"yaxis":"y","x":[29,29,28,29],"text":[":hwy: 29<br \/>:displ: 2.5<br \/>factor(cyl): 5",":hwy: 29<br \/>:displ: 2.5<br \/>factor(cyl): 5",":hwy: 28<br \/>:displ: 2.5<br \/>factor(cyl): 5",":hwy: 29<br \/>:displ: 2.5<br \/>factor(cyl): 5"]},{"y":[2.5,2.5,2.5,2.5],"hoveron":"points","legendgroup":5,"frame":null,"hoverinfo":"text","name":5,"mode":"lines","type":"scatter","xaxis":"x","showlegend":false,"line":{"width":3.77952755905512,"color":"rgba(124,174,0,1)","dash":"solid"},"yaxis":"y","x":[29,29,28,29],"text":[":hwy: 29<br \/>:displ: 2.5<br \/>factor(cyl): 5",":hwy: 29<br \/>:displ: 2.5<br \/>factor(cyl): 5",":hwy: 28<br \/>:displ: 2.5<br \/>factor(cyl): 5",":hwy: 29<br \/>:displ: 2.5<br \/>factor(cyl): 5"]},{"y":[2.8,2.8,3.1,2.8,2.8,3.1,3.1,2.8,3.1,3.1,3.5,3.6,3.0,3.3,3.3,3.3,3.3,3.3,3.8,3.8,3.8,4.0,3.7,3.7,3.9,3.9,3.9,4.0,4.0,4.0,4.0,4.2,4.2,3.8,3.8,4.0,4.0,2.5,2.5,3.3,2.7,2.7,2.7,3.0,3.7,4.0,4.0,4.0,3.5,3.5,3.0,3.0,3.5,3.3,3.3,4.0,3.1,3.8,3.8,3.8,3.4,3.4,4.0,3.0,3.0,3.5,3.0,3.0,3.3,3.4,3.4,4.0,4.0,2.8,2.8,2.8,2.8,2.8,3.6],"hoveron":"points","legendgroup":6,"frame":null,"hoverinfo":"text","name":6,"marker":{"autocolorscale":false,"color":"rgba(0,191,196,1)","opacity":1,"size":5.66929133858268,"symbol":"circle","line":{"width":1.88976377952756,"color":"rgba(0,191,196,1)"}},"mode":"markers","type":"scatter","xaxis":"x","showlegend":true,"yaxis":"y","x":[26,26,27,25,25,25,25,24,25,26,29,26,24,22,22,24,24,17,22,21,23,23,19,18,17,17,17,17,19,17,19,17,17,26,25,26,24,26,26,28,24,24,24,22,19,20,17,19,27,26,26,25,25,17,17,20,26,26,27,28,19,17,20,26,26,28,26,26,27,17,19,18,20,24,23,24,26,26,26],"text":[":hwy: 26<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 27<br \/>:displ: 3.1<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 3.1<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 3.1<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 3.1<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.1<br \/>factor(cyl): 6",":hwy: 29<br \/>:displ: 3.5<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.6<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 22<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 22<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 22<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 21<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 23<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 23<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 3.7<br \/>factor(cyl): 6",":hwy: 18<br \/>:displ: 3.7<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.9<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.9<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.9<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 4.2<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 4.2<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 2.5<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 2.5<br \/>factor(cyl): 6",":hwy: 28<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 2.7<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 2.7<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 2.7<br \/>factor(cyl): 6",":hwy: 22<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 3.7<br \/>factor(cyl): 6",":hwy: 20<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 27<br \/>:displ: 3.5<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.5<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 3.5<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 20<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.1<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 27<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 28<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 3.4<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.4<br \/>factor(cyl): 6",":hwy: 20<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 28<br \/>:displ: 3.5<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 27<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.4<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 3.4<br \/>factor(cyl): 6",":hwy: 18<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 20<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 23<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.6<br \/>factor(cyl): 6"]},{"y":[3.202020262359323,3.202020262359323,3.1369190700341694,3.2671214546844767,3.2671214546844767,3.2671214546844767,3.2671214546844767,3.33222264700963,3.2671214546844767,3.202020262359323,3.0067166853838625,3.202020262359323,3.33222264700963,3.4624250316599374,3.4624250316599374,3.33222264700963,3.33222264700963,3.787930993285705,3.4624250316599374,3.527526223985091,3.397323839334784,3.397323839334784,3.6577286086353977,3.7228298009605516,3.787930993285705,3.787930993285705,3.787930993285705,3.787930993285705,3.6577286086353977,3.787930993285705,3.6577286086353977,3.787930993285705,3.787930993285705,3.202020262359323,3.2671214546844767,3.202020262359323,3.33222264700963,3.202020262359323,3.202020262359323,3.071817877709016,3.33222264700963,3.33222264700963,3.33222264700963,3.4624250316599374,3.6577286086353977,3.5926274163102443,3.787930993285705,3.6577286086353977,3.1369190700341694,3.202020262359323,3.202020262359323,3.2671214546844767,3.2671214546844767,3.787930993285705,3.787930993285705,3.5926274163102443,3.202020262359323,3.202020262359323,3.1369190700341694,3.071817877709016,3.6577286086353977,3.787930993285705,3.5926274163102443,3.202020262359323,3.202020262359323,3.071817877709016,3.202020262359323,3.202020262359323,3.1369190700341694,3.787930993285705,3.6577286086353977,3.7228298009605516,3.5926274163102443,3.33222264700963,3.397323839334784,3.33222264700963,3.202020262359323,3.202020262359323,3.202020262359323],"hoveron":"points","legendgroup":6,"frame":null,"hoverinfo":"text","name":6,"mode":"lines","type":"scatter","xaxis":"x","showlegend":false,"line":{"width":3.77952755905512,"color":"rgba(0,191,196,1)","dash":"solid"},"yaxis":"y","x":[26,26,27,25,25,25,25,24,25,26,29,26,24,22,22,24,24,17,22,21,23,23,19,18,17,17,17,17,19,17,19,17,17,26,25,26,24,26,26,28,24,24,24,22,19,20,17,19,27,26,26,25,25,17,17,20,26,26,27,28,19,17,20,26,26,28,26,26,27,17,19,18,20,24,23,24,26,26,26],"text":[":hwy: 26<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 27<br \/>:displ: 3.1<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 3.1<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 3.1<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 3.1<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.1<br \/>factor(cyl): 6",":hwy: 29<br \/>:displ: 3.5<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.6<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 22<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 22<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 22<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 21<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 23<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 23<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 3.7<br \/>factor(cyl): 6",":hwy: 18<br \/>:displ: 3.7<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.9<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.9<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.9<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 4.2<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 4.2<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 2.5<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 2.5<br \/>factor(cyl): 6",":hwy: 28<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 2.7<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 2.7<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 2.7<br \/>factor(cyl): 6",":hwy: 22<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 3.7<br \/>factor(cyl): 6",":hwy: 20<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 27<br \/>:displ: 3.5<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.5<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 25<br \/>:displ: 3.5<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 20<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.1<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 27<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 28<br \/>:displ: 3.8<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 3.4<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.4<br \/>factor(cyl): 6",":hwy: 20<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 28<br \/>:displ: 3.5<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.0<br \/>factor(cyl): 6",":hwy: 27<br \/>:displ: 3.3<br \/>factor(cyl): 6",":hwy: 17<br \/>:displ: 3.4<br \/>factor(cyl): 6",":hwy: 19<br \/>:displ: 3.4<br \/>factor(cyl): 6",":hwy: 18<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 20<br \/>:displ: 4.0<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 23<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 24<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 2.8<br \/>factor(cyl): 6",":hwy: 26<br \/>:displ: 3.6<br \/>factor(cyl): 6"]},{"y":[4.2,5.3,5.3,5.3,5.7,6.0,5.7,5.7,6.2,6.2,7.0,5.3,5.3,5.7,6.5,4.7,4.7,4.7,5.2,5.2,4.7,4.7,4.7,5.2,5.7,5.9,4.7,4.7,4.7,4.7,4.7,4.7,5.2,5.2,5.7,5.9,4.6,5.4,5.4,4.6,5.0,4.6,4.6,4.6,5.4,5.4,4.6,4.6,4.6,4.6,5.4,4.7,4.7,4.7,5.7,6.1,4.0,4.2,4.4,4.6,5.4,5.4,5.4,4.6,5.0,5.6,5.3,4.7,4.7,5.7],"hoveron":"points","legendgroup":8,"frame":null,"hoverinfo":"text","name":8,"marker":{"autocolorscale":false,"color":"rgba(199,124,255,1)","opacity":1,"size":5.66929133858268,"symbol":"circle","line":{"width":1.88976377952756,"color":"rgba(199,124,255,1)"}},"mode":"markers","type":"scatter","xaxis":"x","showlegend":true,"yaxis":"y","x":[23,20,15,20,17,17,26,23,26,25,24,19,14,15,17,19,19,12,17,15,17,12,17,16,18,15,16,12,17,17,16,12,15,16,17,15,17,17,18,19,17,16,16,17,15,17,21,22,23,22,20,17,12,19,18,14,15,18,18,15,17,16,18,19,17,18,25,17,15,18],"text":[":hwy: 23<br \/>:displ: 4.2<br \/>factor(cyl): 8",":hwy: 20<br \/>:displ: 5.3<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.3<br \/>factor(cyl): 8",":hwy: 20<br \/>:displ: 5.3<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 6.0<br \/>factor(cyl): 8",":hwy: 26<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 23<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 26<br \/>:displ: 6.2<br \/>factor(cyl): 8",":hwy: 25<br \/>:displ: 6.2<br \/>factor(cyl): 8",":hwy: 24<br \/>:displ: 7.0<br \/>factor(cyl): 8",":hwy: 19<br \/>:displ: 5.3<br \/>factor(cyl): 8",":hwy: 14<br \/>:displ: 5.3<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 6.5<br \/>factor(cyl): 8",":hwy: 19<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 19<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 12<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.2<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.2<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 12<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 5.2<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.9<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 12<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 12<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.2<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 5.2<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.9<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 19<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.0<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 21<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 22<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 23<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 22<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 20<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 12<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 19<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 14<br \/>:displ: 6.1<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 4.0<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 4.2<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 4.4<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 19<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.0<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 5.6<br \/>factor(cyl): 8",":hwy: 25<br \/>:displ: 5.3<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 5.7<br \/>factor(cyl): 8"]},{"y":[5.360007003346041,5.2331413897751125,5.021698700490232,5.2331413897751125,5.106275776204184,5.106275776204184,5.486872616916969,5.360007003346041,5.486872616916969,5.444584079059993,5.402295541203017,5.1908528519181365,4.979410162633256,5.021698700490232,5.106275776204184,5.1908528519181365,5.1908528519181365,4.894833086919303,5.106275776204184,5.021698700490232,5.106275776204184,4.894833086919303,5.106275776204184,5.063987238347208,5.14856431406116,5.021698700490232,5.063987238347208,4.894833086919303,5.106275776204184,5.106275776204184,5.063987238347208,4.894833086919303,5.021698700490232,5.063987238347208,5.106275776204184,5.021698700490232,5.106275776204184,5.106275776204184,5.14856431406116,5.1908528519181365,5.106275776204184,5.063987238347208,5.063987238347208,5.106275776204184,5.021698700490232,5.106275776204184,5.275429927632088,5.317718465489064,5.360007003346041,5.317718465489064,5.2331413897751125,5.106275776204184,4.894833086919303,5.1908528519181365,5.14856431406116,4.979410162633256,5.021698700490232,5.14856431406116,5.14856431406116,5.021698700490232,5.106275776204184,5.063987238347208,5.14856431406116,5.1908528519181365,5.106275776204184,5.14856431406116,5.444584079059993,5.106275776204184,5.021698700490232,5.14856431406116],"hoveron":"points","legendgroup":8,"frame":null,"hoverinfo":"text","name":8,"mode":"lines","type":"scatter","xaxis":"x","showlegend":false,"line":{"width":3.77952755905512,"color":"rgba(199,124,255,1)","dash":"solid"},"yaxis":"y","x":[23,20,15,20,17,17,26,23,26,25,24,19,14,15,17,19,19,12,17,15,17,12,17,16,18,15,16,12,17,17,16,12,15,16,17,15,17,17,18,19,17,16,16,17,15,17,21,22,23,22,20,17,12,19,18,14,15,18,18,15,17,16,18,19,17,18,25,17,15,18],"text":[":hwy: 23<br \/>:displ: 4.2<br \/>factor(cyl): 8",":hwy: 20<br \/>:displ: 5.3<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.3<br \/>factor(cyl): 8",":hwy: 20<br \/>:displ: 5.3<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 6.0<br \/>factor(cyl): 8",":hwy: 26<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 23<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 26<br \/>:displ: 6.2<br \/>factor(cyl): 8",":hwy: 25<br \/>:displ: 6.2<br \/>factor(cyl): 8",":hwy: 24<br \/>:displ: 7.0<br \/>factor(cyl): 8",":hwy: 19<br \/>:displ: 5.3<br \/>factor(cyl): 8",":hwy: 14<br \/>:displ: 5.3<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 6.5<br \/>factor(cyl): 8",":hwy: 19<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 19<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 12<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.2<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.2<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 12<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 5.2<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.9<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 12<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 12<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.2<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 5.2<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.9<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 19<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.0<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 21<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 22<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 23<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 22<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 20<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 12<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 19<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 5.7<br \/>factor(cyl): 8",":hwy: 14<br \/>:displ: 6.1<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 4.0<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 4.2<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 4.4<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 16<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 5.4<br \/>factor(cyl): 8",":hwy: 19<br \/>:displ: 4.6<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 5.0<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 5.6<br \/>factor(cyl): 8",":hwy: 25<br \/>:displ: 5.3<br \/>factor(cyl): 8",":hwy: 17<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 15<br \/>:displ: 4.7<br \/>factor(cyl): 8",":hwy: 18<br \/>:displ: 5.7<br \/>factor(cyl): 8"]}]},"evals":[],"jsHooks":[]}</script></div>
```

```{=html}
<div style="background-color:grey;height:2px;width:100%;"></div>
```

```{=html}
<div><pre><small><small>source: <a href="https://github.com/scicloj/clojure-data-scrapbook/blob/main/projects/datavis/ggplot/notebooks/ggplotly_cont.clj">projects/datavis/ggplot/notebooks/ggplotly_cont.clj</a></small></small></pre></div>
```