### 2024-04-16 docs/projects/data-processing/many_json_lines.qmd

---
format:
  html: {toc: true, toc-depth: 4, theme: cosmo, output-file: many_json_lines.html}
fontsize: 0.9em
code-block-background: true
include-in-header: {text: '<link rel = "icon" href = "data:," />'}

---
<style></style><style>.printedClojure .sourceCode {
  background-color: transparent;
  border-style: none;
}
</style><style>.clay-limit-image-width .clay-image {max-width: 100%}
</style>
<script src="many_json_lines_files/md-default0.js" type="text/javascript"></script><script src="many_json_lines_files/md-default1.js" type="text/javascript"></script>

```{=html}
<table><tr><td>This is part of the Scicloj <a href="https://scicloj.github.io/clojure-data-scrapbook/">Clojure Data Scrapbook</a>.</td><a href="https://scicloj.github.io/clojure-data-scrapbook/"><img align="left" alt="SciCloj logo" src="https://scicloj.github.io/sci-cloj-logo-transparent.png" width="40" /></a></tr></table>
```

# Processing many JSON lines

::: {.sourceClojure}
```clojure
(ns many-json-lines
  (:require [charred.api :as charred]
            [scicloj.kindly.v4.kind :as kind]
            [clojure.java.io :as io]))
```
:::

Let us create a file with many JSON lines.

::: {.sourceClojure}
```clojure
(def path "/tmp/lines.jsonl")
```
:::

::: {.sourceClojure}
```clojure
(with-open [writer ((charred/json-writer-fn {}) path)]
  (dotimes [i 3]
    (.writeObject writer
                  {:i i
                   :x (rand)})
    (.write (.w writer)
            "\n")))
```
:::

::: {.printedClojure}
```clojure
nil

```
:::

Let us look at a few of those lines we wrote.

::: {.sourceClojure}
```clojure
(with-open [reader (io/reader path)]
  (->> reader
       line-seq
       (take 4)
       vec))
```
:::

::: {.printedClojure}
```clojure
["{\"i\":0,\"x\":0.8833069538053755}"
 "{\"i\":1,\"x\":0.016778130034834637}"
 "{\"i\":2,\"x\":0.751953143790835}"]

```
:::

Now we will demonstrate processing the data.
We will read all lines, extract the relevant part of each line,
and sum these parts up.

::: {.sourceClojure}
```clojure
(defn extract-relevant-data [json-data]
  (get json-data "x"))
```
:::

::: {.sourceClojure}
```clojure
(with-open [json-fn (-> path
                        io/reader
                        (charred/read-json-supplier {:eof-error? false}))]
  (doall
   (map extract-relevant-data json-fn)))
```
:::

::: {.printedClojure}
```clojure
(0.8833069538053755 0.016778130034834637 0.751953143790835 nil)

```
:::

```{=html}
<div style="background-color:grey;height:2px;width:100%;"></div>
```

```{=html}
<div><pre><small><small>source: projects/data-processing/notebooks/many_json_lines.clj</small></small></pre></div>
```